### 2015-09-17 article/article.tex
remove ignore tex

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{article.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Chaos in a long-term experiment with a plankton community}

    \usepackage{authblk}
                        \author[1]{Owen Petchey}
                    \author[1]{Marco Plebani}
                    \author[1]{Frank Pennekamp}
                            \affil[1]{Institute of Evolutionary Biology and Environmental Studies, University
of Zurich, Zurich, Switzerland}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:owen.petchey@ieu.uzh.ch}{owen.petchey@ieu.uzh.ch}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Benincà, E., Huisman, J., Heerkloss, R., Jöhnk, K.D., Branco, P., Van
Nes, E.H., Scheffer, M. \& Ellner, S.P. (2008) Chaos in a long-term
experiment with a plankton community. Nature, 451, 822--825 DOI:
10.1038/nature06512
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

The original paper describes analyses of fluctuations in the abundance
of organisms in a plankton community derived from the Baltic Sea, housed
in a laboratory environment. The length of the time series (samples
every few days for 2,300 days) allowed for analyses revealing that the
observed dynamics exhibited characteristics consistent with chaos
produced by species interactions. The article concludes that stability
is not required for persistence of complex food webs, and that long-term
prediction of abundances may be fundamentally impossible. The
demonstration of chaotic dynamics and limited forecast horizons (sensu
\textcite{Petchey2015}) are important in the field of ecology, since the
ability to predict dynamics is an open question with considerable
applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.

\section{Methods}\label{methods}

This reproduction started with the raw data (source given below) and
used information from the original paper, the supplementary information
\href{http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf}{the
Supplement to the Nature paper}, and communications with Elisa Benincà
and Stephen Ellner. The latter provided code used to produce results in
the original paper, and its use in this reproduction is indicated below.

\subsection{The data}\label{the-data}

The data are available as an Excel file supplement to
\href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
Ecology Letters publication} \textcite{Beninca2009}. The Excel file
contains several datasheets. Two are particularly important, as they are
the source of the raw data (one contains original species abundances,
the one with the nutrient concentrations). We saved these two datasheets
as comma separated value text files. In the code associated with this
reproduction, these data files are read from the associated github
repository.

Another datasheet in the Ecology Letters supplement contains transformed
variables, though this was not used in this reproduction, as we
reproduced the described transformations and applied them to the raw
data.

We also received some data direct from Steve Ellner, see below for
details.

\subsection{Reproduction environment}\label{reproduction-environment}

The R language and environment for statistical computing and graphics
was used to make the reproduction. Additional R packages required are:

\begin{verbatim}
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(RCurl)
library(pracma)
library(oce)
library(tseriesChaos)
library(reshape2)
library(mgcv)
library(repmis)
\end{verbatim}

\section{Results}\label{results}

\subsection{Population dynamics}\label{population-dynamics}

The raw data show populations dynamics at least very similar to those in
figure 1b-g of the original publication (figure \ref{fig:dynamics}).

\begin{figure}[htbp]
\centering
\includegraphics{figures/unnamed-chunk-21-1}
\caption{\label{fig:dynamics}Observed population dynamics.}
\end{figure}

\subsection{Data transformation}\label{data-transformation}

Did our transformation give data matching the data we received from
Ellner, and that in the Ecology Letters supplement?

\subsection{Spectral analyses}\label{spectral-analyses}

\subsection{Correlations among species
abundances}\label{correlations-among-species-abundances}

\subsection{Lyapunov exponents by direct
method}\label{lyapunov-exponents-by-direct-method}

\subsection{Lyapunov exponents by indirect
method}\label{lyapunov-exponents-by-indirect-method}

\subsection{Predictability decay}\label{predictability-decay}

Only done for non-linear models (the original article also does this for
linear models).

\section{Conclusion}\label{conclusion}

Conclusion, at the very minimum, should indicate very clearly if you
were able to replicate original results. If it was not possible but you
found the reason why (error in the original results), you should exlain
it.

\begin{longtable}[c]{@{}llllll@{}}
\caption{\label{tbl:table}Table caption }\tabularnewline
\toprule
Heading 1 & & & Heading 2 & &\tabularnewline
\midrule
\endfirsthead
\toprule
Heading 1 & & & Heading 2 & &\tabularnewline
\midrule
\endhead
cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
cell6 row 1\tabularnewline
cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
cell6 row 2\tabularnewline
cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
cell6 row 3\tabularnewline
\bottomrule
\end{longtable}

A reference to table \ref{tbl:table}. A reference to figure
\ref{fig:logo}. A reference to equation \ref{eq:1}. A reference to
citation \textcite{markdown}.

\begin{figure}[htbp]
\centering
\includegraphics{rescience-logo.pdf}
\caption{\label{fig:logo}Figure caption}
\end{figure}

\begin{equation} A = \sqrt{\frac{B}{C}} \label{eq:1}\end{equation}

\section{Acknowledgements}\label{acknowledgements}

This reproduction was made as part of the Reproducible Research in
Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES)
Course, lead by Owen Petchey at the University of Zurich. More
information about the course
\href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
on github.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2015-09-24 article/article.tex
added table 1 data

@@
 
+\subsection{Scope of the reproduction}\label{scope-of-the-reproduction}
+
+An attempt was made to reproduce the majority of the results in the
+original article. Instances where we did not attempt to reproduce a
+result are detailed below.
+
 \subsection{The data}\label{the-data}
@@
 Ecology Letters publication} \textcite{Beninca2009}. The Excel file
-contains several datasheets. Two are particularly important, as they are
-the source of the raw data (one contains original species abundances,
-the one with the nutrient concentrations). We saved these two datasheets
-as comma separated value text files. In the code associated with this
-reproduction, these data files are read from the associated github
-repository.
+contains several datasheets. Two were particularly important, as they
+are the source of the raw data (one contains original species
+abundances, the one with the nutrient concentrations). We saved these
+two datasheets as comma separated value (csv) text files. In the code
+associated with this reproduction, these data files are read from the
+associated github repository.
 
 Another datasheet in the Ecology Letters supplement contains transformed
-variables, though this was not used in this reproduction, as we
-reproduced the described transformations and applied them to the raw
-data.
-
-We also received some data direct from Steve Ellner, see below for
-details.
+variables (we also saved this as csv file, in order to use it in this
+reproduction). We also received a dataset direct from Steve Ellner, see
+below for details.
 
@@
 The R language and environment for statistical computing and graphics
-was used to make the reproduction. Additional R packages required are:
+was used to make the reproduction. Additional R packages required are
+specified in the code associated with this reproduction.
 
-\begin{verbatim}
-library(tidyr)
-library(dplyr)
-library(lubridate)
-library(stringr)
-library(ggplot2)
-library(RCurl)
-library(pracma)
-library(oce)
-library(tseriesChaos)
-library(reshape2)
-library(mgcv)
-library(repmis)
-\end{verbatim}
+The code for this reproduction resides in an R markdown document, as
+well as a source file containing some required functions. Some of the
+code takes several minutes to run, so an intermediate data file is
+provided with results from this code.
 
@@
 
-Did our transformation give data matching the data we received from
-Ellner, and that in the Ecology Letters supplement?
+The following transformation steps were used:
 
-\subsection{Spectral analyses}\label{spectral-analyses}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Time series shortened to remove long sequences of zeros.
+\item
+  Interpolation to create equally spaced observations in time series.
+\item
+  Fourth root transformation.
+\item
+  Detrending of five of the time series.
+\item
+  Rescaling to zero mean and unit standard deviation.
+\end{enumerate}
+
+The method for selecting the zeros to remove was unclear. In order to
+reproduce this step, we removed the same data asin the original study,
+by matching to the transformed data with zeros removed in the Ecology
+Letters supplement Excel file mentioned above. All remaining
+transformation steps were performed independently of this dataset. Our
+reproduction of the transformed data closely matched the published
+transformed data. \emph{This could result from us performing the
+interpolation before the zero removal. Should correct this.}
+
+The data received directly from Stephen Ellner was interpolated, but
+without zeros removed. Our interpolated data, without zeros removed,
+matched closely this data.
 
@@
 abundances}\label{correlations-among-species-abundances}
+
+Correlations among species abundances presented in Table 1 of the
+original article closely matched our reproduced correlations, calculated
+from the transformed data with zeros removed.
+
+\begin{longtable}[c]{@{}lcccccccccc@{}}
+\caption{\label{tbl:table}Table caption }\tabularnewline
+\toprule
+& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
+Cala\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
+Cala\tabularnewline
+\midrule
+\endhead
+Bact & & 0.03 & -0.24 *** & 0.05 & 0.18 *** & 0.03 & -0.17 *** & 0.3 ***
+& -0.17 * & 0.22 ***\tabularnewline
+Harp & & & 0.21 *** & -0.11 * & 0.03 & -0.09 & -0.06 & -0.05 & 0.14 &
+0.01\tabularnewline
+Ostr & & & & -0.16 ** & -0.06 & -0.06 & 0 & -0.05 & 0.18 * &
+0.04\tabularnewline
+N & & & & & 0.08 * & 0 & -0.02 & -0.04 & 0.03 & 0.04\tabularnewline
+P & & & & & & -0.04 & 0.03 & 0.1 * & -0.02 & -0.08\tabularnewline
+Pico & & & & & & & -0.17 *** & -0.03 & -0.21 * & 0.28 ***\tabularnewline
+Nano & & & & & & & & -0.19 *** & 0.13 & -0.14 *\tabularnewline
+Roti & & & & & & & & & -0.02 & -0.1 *\tabularnewline
+Prot & & & & & & & & & & NA NA\tabularnewline
+Cala & & & & & & & & & &\tabularnewline
+\bottomrule
+\end{longtable}
+
+\subsection{Spectral analyses}\label{spectral-analyses}
 

### 2015-09-24 article/article.tex
original LE data added

@@
 The method for selecting the zeros to remove was unclear. In order to
-reproduce this step, we removed the same data asin the original study,
+reproduce this step, we removed the same data as in the original study,
 by matching to the transformed data with zeros removed in the Ecology
@@
 reproduction of the transformed data closely matched the published
-transformed data. \emph{This could result from us performing the
-interpolation before the zero removal. Should correct this.}
+transformed data.
 
@@
 original article closely matched our reproduced correlations, calculated
-from the transformed data with zeros removed.
+from the transformed data with zeros removed (figure
+\ref{fig:corr_comp}). Deviations between the original and reproduced
+correlations are relatively small (less than 0.072 units) and
+infrequent. \textbf{These may have resulted from us removing zeros after
+interpolation. Maybe we should check this.}
 
-\begin{longtable}[c]{@{}lcccccccccc@{}}
-\caption{\label{tbl:table}Table caption }\tabularnewline
-\toprule
-& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
-Cala\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
-Cala\tabularnewline
-\midrule
-\endhead
-Bact & & 0.03 & -0.24 *** & 0.05 & 0.18 *** & 0.03 & -0.17 *** & 0.3 ***
-& -0.17 * & 0.22 ***\tabularnewline
-Harp & & & 0.21 *** & -0.11 * & 0.03 & -0.09 & -0.06 & -0.05 & 0.14 &
-0.01\tabularnewline
-Ostr & & & & -0.16 ** & -0.06 & -0.06 & 0 & -0.05 & 0.18 * &
-0.04\tabularnewline
-N & & & & & 0.08 * & 0 & -0.02 & -0.04 & 0.03 & 0.04\tabularnewline
-P & & & & & & -0.04 & 0.03 & 0.1 * & -0.02 & -0.08\tabularnewline
-Pico & & & & & & & -0.17 *** & -0.03 & -0.21 * & 0.28 ***\tabularnewline
-Nano & & & & & & & & -0.19 *** & 0.13 & -0.14 *\tabularnewline
-Roti & & & & & & & & & -0.02 & -0.1 *\tabularnewline
-Prot & & & & & & & & & & NA NA\tabularnewline
-Cala & & & & & & & & & &\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/correlation_comparison.pdf}
+\caption{\label{fig:corr_comp}Comparison of calculated correlations
+among species abundances in the original article and this reproduction.}
+\end{figure}
 

### 2015-09-25 article/article.tex
updates

@@
 
-The raw data show populations dynamics at least very similar to those in
-figure 1b-g of the original publication (figure \ref{fig:dynamics}).
+The reproduced populations dynamics were at least very similar to those
+in figure 1b-g of the original publication (figure \ref{fig:dynamics}).
+Note that we plot fourth root transformed values, rather than raw
+abundances with a y-axis break, as in the original article.
 
@@
 \centering
-\includegraphics{figures/unnamed-chunk-21-1}
+\includegraphics{figures/dynamics.pdf}
 \caption{\label{fig:dynamics}Observed population dynamics.}
@@
 correlations are relatively small (less than 0.072 units) and
-infrequent. \textbf{These may have resulted from us removing zeros after
-interpolation. Maybe we should check this.}
+infrequent.
 
@@
 
+Highlighted in the text of the original paper were: negative
+correlations of picophytoplankton with protozoa, and of
+nanophytoplankton both with rotifers and calanoid copepods, positive
+correlation of picophytoplankton with calanoid copepods, negative
+correlation between bacteria and ostracods, and positive correlation
+between bacteria and phosphorus. All of these correlations were
+reproduced.
+
 \subsection{Spectral analyses}\label{spectral-analyses}
+
+Spectral analyses in the original paper were presented graphically in
+figures S3 (raw spectrograms) and S4 (Welch periodograms). These graphs
+were, apparently, visually inspected to reveal: ``fluctuations covered a
+range of different periodicities'', and ``picophytoplankton, rotifers
+and calanoid copepods seemed to fluctuate predominantly with a
+periodicity of about 30 days.'' It is unclear how these conclusions were
+derived from figures S3 and S4 of the original article. Our reproduced
+spectra (not shown here, but code provided) were not quantitatively
+identical to the spectra in the original article.
 
@@
 
+Reproduced divergence rates (figure \ref{fig:divergence}) and comparison
+of the original and reproduced Lyapunov exponents (figure
+\ref{fig:LE_comparison}). The original article states: ``the distance
+between initially nearby trajectories increased over time, and reached a
+plateau after about 20--30 days''. The reproduced results appear not
+inconsistent with this statement, except for one group of species
+(Harpacticoids). The original article also stated that the analyses
+``yielded significantly positive Lyapunov exponents of strikingly
+similar value for all species (Fig. 3; mean exponent = 0.057 per day,
+s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very similar
+mean value, but had about four times greater standard deviation (mean =
+0.055 and s.d. = 0.019).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/unnamed-chunk-50-1.pdf}
+\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
+exponents (figure 3 in the original article).}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/LE_comparison.pdf}
+\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents in
+the original article and this reproduction.}
+\end{figure}
+
 \subsection{Lyapunov exponents by indirect
@@
 
+The original paper reported global Lyapunov exponent calculated via two
+modelling approaches (neural network and generalised additive models
+{[}GAMs{]}). Only the GAM approach was reproduced, with the assistance
+of code donated by Stephen Ellner. The original article obtained a
+global Lyapunov exponent of λ=0.08 day-1. The reproduced value was 0.04.
+We did not reproduce the bootstrapping used to give confidence intervals
+around this estimate.
+
 \subsection{Predictability decay}\label{predictability-decay}
 
-Only done for non-linear models (the original article also does this for
-linear models).
+The article stated: ``For short-term forecasts of only a few days, most
+species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
+However, the predictability of the species was much reduced when
+prediction times were extended to 15--30days.'' The reproduced
+predictabilities, which were calculated from the GAMs, were consistent
+with these qualitative statements, though were quantitatively different
+(\ref{fig:prediction_distance}). We did not reproduce the predictability
+estimates for linear models.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/prediction_distance.pdf}
+\caption{\label{fig:prediction_distance}Predictability (correlation
+between predicted and observed abundances) and prediction distance
+(days) (figure 2 in the original article).}
+\end{figure}
 
@@
 
-Conclusion, at the very minimum, should indicate very clearly if you
-were able to replicate original results. If it was not possible but you
-found the reason why (error in the original results), you should exlain
-it.
-
-\begin{longtable}[c]{@{}llllll@{}}
-\caption{\label{tbl:table}Table caption }\tabularnewline
-\toprule
-Heading 1 & & & Heading 2 & &\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Heading 1 & & & Heading 2 & &\tabularnewline
-\midrule
-\endhead
-cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
-cell6 row 1\tabularnewline
-cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
-cell6 row 2\tabularnewline
-cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
-cell6 row 3\tabularnewline
-\bottomrule
-\end{longtable}
-
-A reference to table \ref{tbl:table}. A reference to figure
-\ref{fig:logo}. A reference to equation \ref{eq:1}. A reference to
-citation \textcite{markdown}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{rescience-logo.pdf}
-\caption{\label{fig:logo}Figure caption}
-\end{figure}
-
-\begin{equation} A = \sqrt{\frac{B}{C}} \label{eq:1}\end{equation}
+Although we were not able to make a quantitatively accurate reproduction
+of all results of the original article, the qualitative results were
+largely identical. Quantitative differences may have resulted from
+difference in algorithms used. For example,the original used the
+\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
+calculate Lyapunov exponents. As this was available from CRAN
+\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
+mid 2014} and since it is a bit less well integrated with R, we instead
+use the tseriesChaos package \textcite{tseriesChaos}, which in any case
+was largely inspired by the TISEAN project. There may also have been
+some difference in data used for specific analyses, e.g., data with
+zeros removed or not, as it was not always possible to be sure the
+reproduction used exactly the same data.
 

### 2015-09-25 code/indirect_method_functions.R

@@ -1,0 +1,189 @@
+### Assume the data comes in as a matrix, with the time in the rows and species in the columns
+
+Get_GLE_Beninca <- function(X1,gval=1.4) {
+  
+  #X1=X
+  
+  require(mgcv)
+  
+  ## prepare the lagging
+  end.t <- dim(X1)[1]
+  
+  ## change X to be a data.frame
+  X1 <- as.data.frame(X1)
+  
+  ## make columns of X in alphabetical order
+  X1 <- X1[,order(colnames(X1))]
+  
+  ## make a matrix of unlagged and lagged variables
+  Z <- cbind(X1[-1,], X1[-end.t,])
+  colnames(Z) <- c(paste("y_", names(X1), sep=""), names(X1))
+  
+  #1 Cyclopoids=f(Cyclopoids,Protozoa,Rotifers)
+  fit.Cyclopoids <- gam(y_Cyclopoids ~ s(Cyclopoids) + s(Protozoa) + s(Rotifers) +
+                          s(Cyclopoids,Protozoa) + s(Cyclopoids,Rotifers),
+                        gamma=gval,data=Z); 
+  
+  #2 Protozoa = f(Cyclopoids,Protozoa,Picophytoplankton,Nanophytoplankton,Bacteria) 
+  fit.Protozoa <- gam(y_Protozoa ~ s(Cyclopoids) + s(Cyclopoids,Protozoa) + s(Protozoa) +
+                        s(Picophytoplankton) + s(Picophytoplankton,Protozoa) +
+                        s(Nanophytoplankton) + s(Nanophytoplankton,Protozoa) +
+                        s(Bacteria)+s(Bacteria,Protozoa),
+                      gamma=gval,data=Z); 
+  
+  #3 Rotifers = f(Cyclopoids,Rotifers,Picophytoplankton,Nanophytoplankton,Bacteria) 
+  fit.Rotifers <- gam(y_Rotifers ~ s(Cyclopoids) + s(Cyclopoids,Rotifers) +
+                        s(Rotifers) + s(Picophytoplankton) +
+                        s(Picophytoplankton,Rotifers) +
+                        s(Nanophytoplankton) + s(Nanophytoplankton,Rotifers) +
+                        s(Bacteria) + s(Bacteria,Rotifers),
+                      gamma=gval,data=Z); 
+  
[…]

### 2015-09-25 code/report.Rmd
changed repo used for data, to ReScience

---
title: "Reproduce anaylses of Beninca et al (2008)"
author: "Owen Petchey et al"
date: "29 Jan 2015"
output:
  html_document:
    keep_md: yes
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction

This is a reproduction of the anaylses presented in the paper *Chaos in a long-term experiment with a plankton community*, by Elisa Benincà and others ([the paper on the Nature website](http://www.nature.com/nature/journal/v451/n7180/abs/nature06512.html)). Details of the methods are in [the Supplement to the Nature paper](http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf).

This reproduction was made as part of the Reproducible Research in Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES) Course, lead by Owen Petchey at the University of Zurich. More information about the course [here](https://github.com/opetchey/RREEBES/blob/master/README.md) on github.

The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.

# Authors

Lead author was Owen Petchey. Frank Pennekamp and Marco Plebani made sizeable contributions. All contributors are detailed on github in the commit history etc.

# Known issues

The known issues / problems with this reproduction are with the prefix "Beninca" [here](https://github.com/opetchey/RREEBES/issues). Please add and or solve issues there.

# The data

The data are available as an Excel file supplement to [an Ecology Letters publication](http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract). The Excel file contains several datasheets. Two are particularly important, as they are the source of the raw data (one contains original species abundances, the one with the nutrient concentrations). Another datasheet in the ELE supplement contains transformed variables. We also got some data direct from Steve Ellner, see below for details.

In the code below, the data and any other files are read from github, which means there must be a connection to github.

# R session info

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
rm(list=ls())
devtools::session_info()
```

# Repository to read data from

The datasets for this reproduction are read from this github repository:
```{r}
repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/beninca_dev"
```

# First get the raw data into R and tidy it.

All required libraries:

```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(RCurl)
library(pracma)
library(oce)
library(tseriesChaos)
library(reshape2)
library(mgcv)
library(repmis)

spp.abund <- read.csv(text=getURL(paste0(repo,"/data/species_abundances_original.csv")), skip=7, header=T)

spp.abund <- select(spp.abund, -X, -X.1)
spp.abund <- spp.abund[-804:-920,]
str(spp.abund)
```

The Protozoa variable contains some numbers with comman as the decimal separator. This creates a question about what dataset was used for the original analyses, as it could not have been this one.

```{r}
spp.abund$Protozoa <- as.numeric(str_replace(spp.abund$Protozoa, ",", "."))
```

Format the dates as dates

```{r}
spp.abund$Date <- dmy(spp.abund$Date)
```

Ooops... R assumes the experiment was done in the 21st century. Shouldn't matter too much.

Check dates match the Day.number (should give true):
```{r}
sum(spp.abund$Day.number == 1+as.numeric((spp.abund$Date - spp.abund$Date[1]) / 24 / 60 / 60)) == length(spp.abund$Date)
```

Check for duplicate dates:
```{r}
spp.abund$Date[duplicated(spp.abund$Date)]
which(duplicated(spp.abund$Date))
```

Original dataset contains a duplicated date: 28/10/1996 (row 709 and 710 in excel sheet). Lets change the date in row 709 to 26/10/1996, which will put it half way between the two surrounding dates:

```{r}
which(spp.abund$Date==ymd("2096-10-28 UTC"))
spp.abund$Date[701] <- ymd("2096-10-26 UTC")
```

Check dates match the Day.number (should give true):
```{r}
sum(spp.abund$Day.number == 1+as.numeric((spp.abund$Date - spp.abund$Date[1]) / 24 / 60 / 60)) == length(spp.abund$Date)
```

Fix the Day.number problem:
```{r}
spp.abund$Day.number <- 1+as.numeric((spp.abund$Date - spp.abund$Date[1]) / 24 / 60 / 60)
```

Data is in wide format, so change it to long:
```{r}
spp.abund <- gather(spp.abund, "variable", "value", 3:12)
str(spp.abund)
```

Bring in the nutrient data:
```{r}
nuts <- read.csv(text=getURL(paste0(repo,"/data/nutrients_original.csv")), skip=7, header=T)
#nuts <- read.csv("~/Dropbox (Dept of Geography)/RREEBES/Beninca_etal_2008_Nature/data/nutrients_original.csv", skip=7, header=T)

nuts <- select(nuts, -X, -X.1)
nuts <- nuts[-349:-8163,]
nuts$Date <- dmy(nuts$Date)
nuts <- select(nuts, -NO2, -NO3, -NH4)
#nuts$Date[duplicated(nuts$Date)]
#which(duplicated(nuts$Date))
nuts <- gather(nuts, "variable", "value", 3:4)
str(nuts)
```

Now put the two datasets together
```{r}
all.data <- rbind(spp.abund, nuts)
```

Now select only the date range used in the Nature paper. From the supplment *The analysis in Benincà et al. (Nature 2008) covered all data from 16/06/1991 until 20/10/1997*. (Remembering dates in the R dataframes are 2090s.)
```{r}
all.data <- filter(all.data, Date>=dmy("15/06/2091") & Date<=dmy("21/10/2097"))

#all.data[all.data$Date==dmy("16/06/2091"),]
#all.data[all.data$Date==dmy("20/10/2097"),]

```

# Reproducing figure 1b through 1g

(No attempt to reproduce Figure 1a, as its a food web diagram.)

First quick go:
```{r}
ggplot(all.data, aes(x=Day.number, y=value)) + geom_line() +
  facet_wrap(~variable, scales="free_y")
```

Now we add a column that gives the variable types, same as in figure 1b through 1g.
First make a lookup table giving species type:
```{r}
tt <- data.frame(variable=unique(all.data$variable),
                 Type=c("Cyclopoids", "Herbivore", "Herbivore", "Herbivore",
                        "Phytoplankton",  "Phytoplankton", "Phytoplankton",
                        "Detritivore", "Detritivore", "Bacteria", "Nutrient", "Nutrient"))
#tt
```

And add the Type variable to the new dataset:
```{r}
all.data <- merge(all.data, tt)
```

First lets set the colours as in the original:
```{r}
species.colour.mapping <- c("Cyclopoids"="pink",
                            "Calanoid.copepods"="red",
                            "Rotifers"="blue",
                            "Protozoa"="green",
                            "Nanophytoplankton"="red",
                            "Picophytoplankton"="black",
                            "Filamentous.diatoms"="green",
                            "Ostracods"="lightblue",
                            "Harpacticoids"="purple",
                            "Bacteria"="black",
                            "Total.dissolved.inorganic.nitrogen"="red",
                            "Soluble.reactive.phosphorus"="black")                            
```

Next change the order of the levels in the Type variable, so plots appear in the same order as in the original figure:
```{r}
all.data$Type <- factor(all.data$Type, levels=c("Cyclopoids", "Herbivore", "Phytoplankton", "Nutrient",
                                    "Bacteria", "Detritivore"))
```

<!--
Here are some versions of the graphs that are not included in the report, hence the html comment:

Now a version that doesn't try to recreate the "gap" in the y axes of the original figures:
```{r, warning=FALSE, fig.width=10}
g1 <- qplot(as.numeric(Day.number), value, col=variable, data=all.data) +
  facet_wrap(~Type, ncol=2, scales="free_y") +
  geom_point() + geom_line() +
  scale_colour_manual(values = species.colour.mapping)
g1
```
Looks reasonably good.

Now a version that approximates the "gap", by removing data above it:
```{r, warning=FALSE, fig.width=10}
an2 <- filter(all.data, Type=="Cyclopoids" & value<0.6 |
                Type=="Herbivore" & value<13 |
                Type=="Phytoplankton" & value<1400 |
                Type=="Nutrient" & value<50 |
                Type=="Bacteria" & value<10 |
                Type=="Detritivore" & value<0.7) 
g1 <- qplot(as.numeric(Day.number), value, col=variable, data=an2) +
  facet_wrap(~Type, ncol=2, scales="free_y") +
  geom_point() + geom_line() +
  scale_colour_manual(values = species.colour.mapping)
g1
```
Difficult it look like the data go off the top of the graph in ggplot.

Try logarithmic y-axes:
```{r, warning=FALSE, fig.width=10}
g1 <- qplot(as.numeric(Day.number), log10(value+0.00001), col=variable, data=all.data) +
  facet_wrap(~Type, ncol=2, scales="free_y") +
  geom_point() + geom_line() +
  scale_colour_manual(values = species.colour.mapping)
g1
```

End of html comment
-->

The graph with abundances fourth root transformed, as this is the transformation used in the ms.

```{r, warning=FALSE, fig.width=10}
g1 <- ggplot(aes(x=as.numeric(Day.number), y=value^0.25, col=variable), data=all.data) +
  xlab("Time (days)") + ylab("Abundance value ^ 0.25") +
  facet_wrap(~Type, ncol=2, scales="free_y") +
  geom_line() +
  scale_colour_manual(values = species.colour.mapping)
g1
```

# Data transformation

Now we need to work with transformed data. Details of the transformation, copied from the Supplmentary information are in indented quote style in the following sections... looks like this:

> 3. Transformation of the time series. We transformed the original time series, shown in Fig. 1b-g of the main text, to obtain stationary time series with equidistant data and homogeneous units of measurement. The transformation steps are illustrated for the bacteria (Fig. S1).

Aside: The ELE supplement contains the raw data and the transformed data, in separate data sheets. I (Owen) also got the interpolated data from Stephen Ellner directly.

## Interpolation

> First, the time series were interpolated using cubic hermite interpolation, to obtain data with equidistant time intervals of 3.35 days (Fig. S1a).

Make a sequence of times at which to interpolate.

```{r}
#aggregate(Day.number ~ variable, all.data, min)
#aggregate(Day.number ~ variable, all.data, max)
#xout <- seq(343.35, 2657.2, by=3.35)
xout <- seq(343.35, 2658, by=3.35)
#range(xout)
```

```{r}
all.data1 <- na.omit(all.data)

#group_by(all.data1, variable) %>% summarise(out=min(Day.number))

mt <- plyr::dlply(all.data1,
                  "variable",
                  function(xx) pracma::interp1(x=xx$Day.number,
                                               y=xx$value,
                                               xi=xout,
                                               method="cubic"))
## Aside: the duplicated date that was previously fixed was only discovered by a warning message
## given by the pracma::interp1 function!!!

mt <- as.data.frame(mt)
mt <- cbind(Day.number=xout, mt)
mt <- gather(mt, variable, value, 2:13)
#ggplot(mt, aes(x=Day.number, y=value)) + facet_wrap(~variable, scales="free") + geom_line()
```

Check this against the data direct from Steve:

```{r}

from.steve <- read.csv(text=getURL(paste0(repo,"/data/direct_from_Steve/interp_short_allsystem_newnames.csv")), header=T)

from.steve <- gather(from.steve, Species, Abundance, 2:13)

names(from.steve) <- c("Day.number", "variable", "value")

g1 <- ggplot(mt, aes(x=as.numeric(Day.number), y=value)) +
  facet_wrap(~variable, ncol=2, scales="free_y") +
  geom_line(size=0.5, col="black") +
  scale_colour_manual(values = species.colour.mapping)
g2 <- geom_line(data=from.steve, aes(x=Day.number, y=value), col="red")
g1 + g2

```

Looks very good.

## Fourth root transform

> Next, because the original time series showed many sharp spikes, the time series were rescaled using a fourth-root power transformation (Fig. S1b). The sharp spikes bias "direct method" estimates of the Lyapunov exponent, because nearby pairs of reconstructed state vectors mostly occurred in the troughs between spikes. The average rate of subsequent trajectory divergence from these pairs is therefore an estimate of the local Lyapunov exponent in the troughs, which may be very different from the global Lyapunov exponent. By making spikes and troughs more nearly symmetric, the power transformation resulted in a much more even spread of nearby state vector pairs across the full range of the data for all functional groups in the food web. The transformation is also useful for fitting nonlinear models of the deterministic skeleton (used for nonlinear predictability and indirect method estimates of the Lyapunov exponent), which was done by least squares and therefore is most efficient when error variances are stabilized. Fourth-root transformation is intermediate between the square-root transformation that would approximately stabilize the measurement error variance in count data from random subsamples, and the log transformation that is usually recommended for stabilizing process noise variance due to stochastic variation in birth and death rates.

```{r}
mt$fr.value <- mt$value^0.25
```

## Detrend

> The time series were then detrended using a Gaussian kernel with a bandwidth of 300 days (red line in Fig. S1b), to obtain stationary time series. Most species did not show long-term trends, except for the bacteria, detritivores (ostracods and harpacticoid copepods), dissolved inorganic nitrogen and soluble reactive phosphorus. One possible explanation for these trends in the microbial loop could be the slow accumulation of refractory organic material in the mesocosm, but we have not measured this component.

```{r}
ww.td <- filter(mt, variable=="Total.dissolved.inorganic.nitrogen" |
                  variable=="Soluble.reactive.phosphorus" |
                  variable=="Bacteria" |
                  variable=="Ostracods" |
                  variable=="Harpacticoids")
## and to not detrend
ww.ntd <- filter(mt, variable!="Total.dissolved.inorganic.nitrogen" &
                  variable!="Soluble.reactive.phosphorus" &
                  variable!="Bacteria" &
                  variable!="Ostracods" &
                  variable!="Harpacticoids")
## detrend:
ww1 <- group_by(ww.td, variable) %>%
  mutate(trend=ksmooth(Day.number,fr.value,bandwidth=300,kernel="normal")$y)
ww1$dt.value <- ww1$fr.value-ww1$trend
#ww1 <- select(ww1, trend)

## don't detrend
ww2 <- ww.ntd
ww2$trend <- 0

ww2$dt.value <- ww2$fr.value

## rejoin
detrended <- rbind(ww1, ww2)

```

## Zero removal

> The time series of cyclopoid copepods, protozoa, filamentous diatoms, harpacticoid copepods and ostracods contained long sequences of zero values. This does not imply that these species were absent from the food web during these periods, but that their concentrations were below the detection limit. Time series dominated by many zeros can bias the statistical analysis. Therefore, these time series were shortened to remove long sequences of zero values, before the data transformation. The transformed data of all species in the food web are shown in Figure S2.

Since the method for this was not given, we do it by getting the trimmed day range from the fully transformed data in the ELE supplement.

```{r, echo=FALSE}

tr <- read.csv(text=getURL(paste0(repo,"/data/transformed_data_Nature2008.csv")),
               skip=7, na.string="")

#head(tr)

tr <- tr[,-14:-24] ## remove bad columns
tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
tr <- gather(tr, key="Species", value="Abundance", 2:13)
levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods"
levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
names(tr)[2] <- "variable"
names(tr)[1] <- "Day.number"
tr$Day.number <- as.numeric(as.character(tr$Day.number))
## get day range
min.max.day <- group_by(tr, variable) %>%
  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))

```

Now we create a trimmed version of our data:

```{r}
trimmed_detrended <- full_join(detrended, min.max.day)
trimmed_detrended <- filter(trimmed_detrended, Day.number>min.day & Day.number<max.day)
```

## Rescale

> Finally, the time series were linearly rescaled to have zero mean and a standard deviation of 1 (Fig. S1c).

(Note that this standardisation is not done in the code sent by Stephen. Probably shouldn't make a difference in the GAMs?)

```{r}
## standardise
scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
  mutate(stand.y=as.numeric(scale(dt.value)))
summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))

## or don't standardise
final <- detrended
final$y <- final$dt.value
summarise(final, mean=mean(y), sd=sd(y))
```

```{r}
glimpse(final)
```

## Figure S1 (visualising the transformation)

Choose a species to plot:
```{r}
soi <- "Bacteria"
```

Raw and interpolated data:

```{r}
g1 <- ggplot(filter(all.data, variable==soi), aes(x=Day.number, y=value)) +
  facet_wrap(~variable, ncol=2, scales="free_y") +
  geom_point(size=1, col="black") + geom_line(size=0.1) +
  scale_colour_manual(values = species.colour.mapping) + ggtitle("Raw and interpolated data")
g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=value), size=0.25, col="blue")
g1 + g2
```

Fourth root transformed with trend:

```{r}
g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=fr.value)) +
  facet_wrap(~variable, ncol=2, scales="free_y") +
  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
  scale_colour_manual(values = species.colour.mapping) + ggtitle("Quarter power trans. and trend")
g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=trend), size=0.25, col="blue")
g1 + g2
```

Detrended and normalised:

```{r}
g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=y)) +
  facet_wrap(~variable, ncol=2, scales="free_y") +
  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
  scale_colour_manual(values = species.colour.mapping) + ggtitle("Detrended and normalised")
g1
```

## Compare the data made above to published data

Now take a look at the transformed data provided in the ELE Supplement, in the data sheet *transformed_data_Nature2008*.

First note that the data in the ELE supplement include a data point at day 2658.1, whereas the data above finish at 2654.85. This is because the real data end at 2658, and the interpolation method above doesn't want to create data outside the range of the original data.

The graph isn't produced here, as there is a mismatch in days abundances were interpolated to. Should fix this...

```{r, echo=FALSE, eval=T}

## Days are not matching at present, so the following doesn't work with the merge

two.datasets <- full_join(tr, scaled_trimmed_final)
#head(two.datasets)
#match(tr$Day.number, final$Day.number)

#ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point()

ggplot(filter(two.datasets), aes(x=as.numeric(Day.number), y=stand.y)) +
  geom_path() +
  facet_wrap(~variable) +
  geom_line(aes(y=Abundance), col="red")

#ggplot(two.datasets, aes(x=as.numeric(Day.number), y=stand.y)) + geom_line() +
#  facet_wrap(~variable) +
#  geom_line(aes(y=Abundance), col="red")

ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point() +
  facet_wrap(~variable)

```

Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros, whereas we don't do any zero removal (this is the same as in Ellner's code.)

# Spectral analyses

```{r}

# Raw spectrum
spectra <- final %>% group_by(variable) %>% do(spectra = spectrum(ts(data=.$y, end=2650.15, deltat=3.35), log='no', method="pgram", detrend=F, plot=F))
spec <- spectra %>% do(data.frame(spec = .$spec[[2]], freq = .$spec[[1]], group = .[[1]]))

ggplot(spec, aes(y=spec, x=1/freq, group=group)) + geom_line() + facet_wrap(~group, scales="free_y") + scale_x_log10() + 
coord_cartesian(ylim=c(0,40), xlim=c(10,240))

freq.est <- spec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
freq.est$freq <- 1/freq.est$freq
#freq.est

# Welch's periodogram

wspectra <- final %>% group_by(variable) %>% do(spectra = pwelch(ts(data=.$y, end=2650.15, deltat=3.35), window=5, method="pgram", plot=F))
wspec <- wspectra %>% do(data.frame(spec = .$spec[[2]], freq = .$spec[[1]], group = .[[1]]))

ggplot(wspec, aes(y=spec, x=1/freq, group=group)) + geom_line() + facet_wrap(~group, scales="free_y") + scale_x_log10() +
coord_cartesian(ylim=c(0.01,100), xlim=c(10,240)) + scale_y_log10()

freq.est <- wspec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
freq.est$freq <- 1/freq.est$freq
#frequency(final$y)
ts <- as.ts(final$y, frequency = 0.3)
#time(ts)

```

# Reproducing Table 1 using ELE supplement data.

Create dataset with zeros removed for this table (note that this is probably not how Beninca et al did this):
```{r}

#final_nozeros <- final[final$value!=0,]
final_nozeros <- scaled_trimmed_final

#
#final_nozeros <- final_nozeros[final_nozeros$variable!="Cyclopoids" & final_nozeros$variable!="Filamentous.diatoms",c("Day.number","variable","stand.y")]

final_wide <- spread(as.data.frame(final_nozeros[,c("Day.number","variable","stand.y")]), variable, stand.y)
final_wide <- select(final_wide, - Day.number)
# we also removed data on Cyclopoids, not used in the table
```

Calculate correlation coefficients:
```{r}
cor.coefs <- cor(final_wide, use="pairwise.complete.obs")
```

Only keep the upper triangle of the cor.pvals matrix:
```{r}
#for(i in 1:10){
#  for(j in 1:10){
#  cor.coefs[i,j] <- ifelse(i<j, cor.coefs[i,j], NA)
#}}
```

Get p-vals too:
```{r}
cor.pvals <- matrix(NA, length(final_wide[1,]), length(final_wide[1,]))
for(i in 1:length(final_wide[1,]))
  for(j in 1:length(final_wide[1,])) {
    if(sum(!is.na(final_wide[,i]) & !is.na(final_wide[,j]))>10)
    cor.pvals[i,j] <- cor.test(final_wide[,i], final_wide[,j])$p.value
}
    #plot(final_wide[,2], final_wide[,3])
```

Only keep the upper triangle of the cor.pvals matrix:
```{r}
#for(i in 1:10){
#  for(j in 1:10){
#  cor.pvals[i,j] <- ifelse(i<j, cor.pvals[i,j], NA)
#}}
```

Add significance "stars" to cor.coefs from cor.pvals
```{r}
cor.stars <- cor.pvals
cor.stars <- ifelse(cor.pvals<0.0001, "***",
                    ifelse(cor.pvals<0.001, "**",
                           ifelse(cor.pvals<0.05, "*", "")))
```

```{r}
cor.cp <- cor.coefs
for(i in 1:length(final_wide[1,])){
  for(j in 1:length(final_wide[1,])){
  cor.cp[i,j] <- paste(round(cor.coefs[i,j],2), cor.stars[i,j])
}}
```

```{r}

oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)

cor.coefs <- cor.coefs[oo,oo]
cor.cp <- cor.cp[oo,oo]

for(i in 1:length(cor.cp[1,]))
  for(j in 1:i)
  cor.cp[i,j] <- ""

#sn <- c("Bacteria", "Harps", "Ostr", "N", "P", "Picophyt", "Nanophyt", "Rotifers", "Protozoa", "Calanoids")
sn <- substr(rownames(cor.cp), 1, 4)
sn[4:5] <- c("N", "P")
dimnames(cor.cp) <- list(sn, sn)

```

Make it a table:
```{r}
library(knitr)
table1b <- kable(cor.cp, format="markdown", col.names = colnames(cor.cp), align="c",
                caption="Table 1.'Correlations between the species in the food web. Table entries show the product–moment correlation coefficients, after transformation of the data to stationary time series (see Methods). Significance tests were corrected for multiple hypothesis testing by calculation of adjusted P values using the false discovery rate.' Significant correlations are indicated as follows: *: P<0.05; **: P<0.01; ***: P<0.001. 'The correlation between calanoid copepods and protozoa could not be calculated, because their time series did not overlap. Filamentous diatoms and cyclopoid copepods were not included in the correlation analysis, because their time series contained too many zeros.' (Beninca et al. 2008)")

table1b

```

Compare correlations in table 1 of the original article with those calculated here:
```{r}

original.cors <- read.csv(text=getURL(paste0(repo,"/data/table1_original_article.csv")), skip=0, header=T, row.names = 1)

original.cors <- as.matrix(original.cors)

qplot(x=as.vector(original.cors), y=as.vector(cor.coefs), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
      xlab="Correlations in original article", ylab="Correlations calculated in this reproduction") + geom_abline(intercept=0,slope=1)

max(as.vector(original.cors) - as.vector(cor.coefs), na.rm=T)
```

# Predictability (Figure 2)

This will be done after getting global Lyapunov exponents by the indirect method.

# Lyapunov exponents by direct method (Figure 3)

Estimate the Lyapunov exponents of the time series, via time-delayed embedding. The Nature report used the [Tisean software](http://www.mpipks-dresden.mpg.de/~tisean/), which was available from CRAN [until mid 2014](http://cran.r-project.org/web/packages/RTisean/index.html). Based on this, and being a bit less well integrated with R, we'll instead use the [tseriesChaos](http://cran.r-project.org/web/packages/tseriesChaos/index.html) package, which was *largely inspired by the TISEAN project*.

Unclear if this was performed on untransformed or transformed data. First try with the transformed data.
Time delay (1), embedding dimension (6), and Theiler window (50) were used in the Nature report. Other parameters are chosen rather randomly, though don't seem to matter too much.

```{r}
time.delay <- 1
embedding.dimension <- 6
Theiler.window <- 50
```

Note that a time step is 3.35 days in the transformed data. So to get a graph with 80 days on the x-axis (as in Figure 3 in the Nature report), we need 80/3.35 = 24 time steps for the calculation of Lyapunov exponents.
```{r}
time.steps <- 24
```

Remove the species that were not analysed in the Nature report, due to too many zeros in the time series:
```{r}
led <- filter(tr, variable!="Filamentous.diatoms",
                variable!="Protozoa",
                variable!="Cyclopoids")
```

Get the data for the graphs:
```{r, message=FALSE, error=TRUE}
all.species <- unique(as.character(led$variable))
diverg <- matrix(NA, time.steps, length(all.species))
colnames(diverg) <- all.species
for(i in 1:length(all.species)) {
  print(all.species[i])
  tr.fs <- filter(final, variable==all.species[i])$y
  diverg[,i] <- as.numeric(try(lyap_k(tr.fs,
                                      m=embedding.dimension,
                                      d=time.delay,
                                      k=10, # number of considered neighbours 20
                                      ref=40, # number of points to take into account 100
                                      t=Theiler.window,
                                      s=time.steps,
                                      eps=10 # radius where to find nearest neighbours 10
                                      )))
}
## a bit of a fudge with the translation to days
diverg <- as.data.frame(cbind(days=1:time.steps, diverg))
diverg <- gather(diverg, Species, Difference, 2:10)
diverg$days <- diverg$days*3.35
#str(diverg)
```

Next calculate the Lyapunov exponents, noting that 6 or 7 points were used in the regressions in the Nature report
```{r}
diverg$Difference[is.na(diverg$Difference)] <- 0
diverg$Difference[is.infinite(diverg$Difference)] <- 0
diverg.short <- filter(diverg, days<24) ## 24 is about 6 steps, after initial gap
LEs <- group_by(diverg.short, Species) %>%
  summarise(le=coef(lm(Difference[1:6] ~ days[1:6]))[2])
#pval=summary(lm(Difference[1:6] ~ days[1:6]))$coefficients[2,4])
```

Then plot the graphs with LE:
```{r}
LEs <- mutate(LEs, days=20, Difference=-0.5)
g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species) +
  geom_text(data=LEs, aes(label=round(le,3)), group=NULL)
g1
```

Not exactly the same at Figure 3 in the Nature report. Qualitatively the same, except for where the time-delayed embedding failed.

Compare graphically with original LEs
```{r}
original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original_direct_LEs.csv")), skip=0, header=T)
LEs

both_LEs <- full_join(original.LEs, LEs)

ggplot(both_LEs, aes(x=LE, y=le)) + geom_point() + xlim(0.02, 0.09) + ylim(0.02, 0.09) + geom_abline(intercept=0, slope=1) +
  xlab("Lyapunov exponent from original article") + ylab("Reproduced Lyapunov exponent")

```

# Lyapunov exponents by indirect method

The functions used in the code below are based on code received from Stephen Ellner. The modifications have been tested, and produce the same results as Ellner's original code.

The function needs a matrix, X, with species abundances in wide format. Be careful to work on the unstandardised data (or standardised, if you wish) (comment out appropriate lines here.

```{r}
## use next line to work on unstandardised data
final.to.melt <- final[, c("variable", "dt.value", "Day.number")]
## use next line to work on standardised
#final.to.melt <- final[, c("variable", "y", "Day.number")]

names(final.to.melt)[1] <- "Species"
melted <- melt(final.to.melt, id=c("Species", "Day.number"))
X <- acast(melted, formula= Day.number ~ Species)
#str(X)
X <- as.data.frame(X)
```

Restrict range of data appropriately:
```{r}
## Select the time period to use 
start.longest=334; start.longer=808; start.shorter=1035; 
e=as.numeric(row.names(X)) > start.longer; X=X[e,];
e=as.numeric(row.names(X)) < 2654; X=X[e,]; 
```

Load and run the functions, or read in from data file (default option, for which you will need to change the path to the data file.)

```{r}
# read script lines from website
script <- getURL(paste0(repo,"/report/functions/indirect_method_functions.R"), ssl.verifypeer = FALSE)

# parase lines and evealuate in the global environement
eval(parse(text = script))

## don't run this as it takes a while
#LE <- Get_GLE_Beninca(X)
#save(LE, file="~/Desktop/GLE_estimate.Rdata")

## load the already saved data from github (this can take some time depending on the internet connection)
source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))

LE[[1]]
```

This is quite far from the number using code and data from Steve (0.08415112). But recall that the functions have been carefully checked. Probably it is the data going into this function. A final step would be to save the data here, and take it into Steve's function. (This point is an issue on github.)

# Predictability (Figure 2)

```{r, echo=FALSE}
## put the models together into a list
all.gams <- LE[[2]]
#names(all.gams)
```

Need to predict until about 40 days ahead = 40 / 3.35 time steps = 12 time steps.

Do this from each time in the time series as initial abundances.

From each time in the series, predict 12 steps ahead. Put results in an array:
time x species, start.location
time will be 12
species will be 12
start location will be length of time series - 12

```{r}
x <- X

Z <- LE[[3]]
Z1 <- Z[,13:24]

all.species <- names(all.gams)

time.to.predict <- 12

nn <- length(Z1[,1])-time.to.predict

preds <- array(NA, c(12, 12, nn))

dimnames(preds) <- list(dist=1:12,
                        species=all.species,
                        start.time=1:nn)
for(i in 1:length(all.species))
  preds[1,i,] <- predict(all.gams[[i]])[1:nn]

#for(i in 1:length(all.species))
#  print(length( predict(all.gams[[i]])))

```

<!--
Look at the correlation between observed and predicted abundances:

```{r}

layout(matrix(1:12, 4, 3))
for(soi in all.species){
  xxx <- Z1[1:nn,soi]
  yyy <- preds[1, soi, ]
  plot(xxx, yyy,
       xlab="Observed abundance",
       ylab="Predicted abundance",
       main=paste(soi, "rsq =", round(cor(xxx,yyy)^2,2)))
}

```
-->

Now for the predictions at t+2 from predictions at t+1

```{r}
pred.time <- 2
for(pred.time in 2:12) {
  for(i in 1:length(all.species))
    preds[pred.time,i,pred.time:nn] <- predict(all.gams[[i]],
                                                newdata=as.data.frame(t(preds[pred.time-1,,])))[pred.time:nn]
}
```

Get the correlations and do some house keeping:

```{r}
cors <- matrix(NA, 12, 12)
dimnames(cors) = list(dist=1:12,
                        species=names(x))
for(i in 1:12) {
  for(j in 1:12) {
    cors[j,i] <- cor(Z1[1:nn,i], preds[j, i, ], use="complete.obs")^2
}}

cors <- data.frame(dist=as.numeric(I(rownames(cors))), cors)
cors.long <- gather(as.data.frame(cors), key=dist)

names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
```

And plot our version of figure 2:

```{r}

ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
  geom_point() +
  facet_wrap(~Variable, scales="free_y" )
```

First pass working. Need to double check everything. Quite different patterns from in figure 2 of the nature paper. There is relatively little information in the paper or supplement about how figure 2 data was produced, so difficult to pin down the difference without asking authors.

### 2015-10-20 article/article.tex
update

@@
             bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
+            pdfauthor=
+{
+      Owen Petchey,
+      Marco Plebani,
+      Frank Pennekamp,
+  },
+            pdftitle=
+{
+[Re] Chaos in a long-term experiment with a plankton community
+},
             colorlinks=true,
@@
 observed dynamics exhibited characteristics consistent with chaos
-produced by species interactions. The article concludes that stability
-is not required for persistence of complex food webs, and that long-term
-prediction of abundances may be fundamentally impossible. The
+produced by non-linear species interactions. The article concludes that
+stability is not required for persistence of complex food webs, and that
+long-term prediction of abundances may be fundamentally impossible. The
 demonstration of chaotic dynamics and limited forecast horizons (sensu
@@
 
+The original species abundance data contained errors (e.g., a few
+numerica values had a comma in place of a period as the decimal
+separator) that suggested that this was not the exact version of the
+dataset used in the original article, or that this was the exact
+dataset, but with errors corrected.
+
 \subsection{Reproduction environment}\label{reproduction-environment}
@@
 \centering
-\includegraphics{figures/dynamics.pdf}
+\includegraphics{figures/obs_pop_dyn.pdf}
 \caption{\label{fig:dynamics}Observed population dynamics.}
@@
 \centering
-\includegraphics{figures/unnamed-chunk-50-1.pdf}
+\includegraphics{figures/div_rate.pdf}
 \caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
@@
 \includegraphics{figures/LE_comparison.pdf}
-\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents in
-the original article and this reproduction.}
+\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents,
+estimated by direct method, in the original article and this
+reproduction.}
 \end{figure}
@@
 of all results of the original article, the qualitative results were
-largely identical. Quantitative differences may have resulted from
+largely identical. For example, all Lyapunov exponents estimated by
+direct method are positive, as in the original article, consistent with
+chaotic dynamics. Quantitative differences may have resulted from
 difference in algorithms used. For example,the original used the
@@
 use the tseriesChaos package \textcite{tseriesChaos}, which in any case
-was largely inspired by the TISEAN project. There may also have been
-some difference in data used for specific analyses, e.g., data with
-zeros removed or not, as it was not always possible to be sure the
-reproduction used exactly the same data.
+was largely inspired by the TISEAN project. In addition, there may have
+been some difference in algorithm parameters, as not all parameters
+required by the functions we used were reported in the original ms.
+There may also have been some difference in data used for specific
+analyses, e.g., data with zeros removed or not, as it was not always
+possible to be totally sure the reproduction used exactly the same data
+as the original article.
+
+In conclusion, this reproduction supports the general scientific
+conclusions of the original article, but also shows how difficult can be
+an accurate quantitative reproduction, even in the presence of the
+extensive methodological details provided alongside the original
+article.
 

### 2015-10-20 code/report.Rmd
update

@@
 g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species) +
-  geom_text(data=LEs, aes(label=round(le,3)), group=NULL)
+  geom_text(data=LEs, aes(x=60, y=-2.5, label=round(le,3)), group=NULL)
 g1
@@
 ```{r}
-
+levels(cors.long$Variable)
 ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
   geom_point() +
-  facet_wrap(~Variable, scales="free_y" )
+  facet_wrap(~Variable, scales="free_y", ncol=3 )
+
+cors.long$Variable <- factor(cors.long$Variable, sort(levels(cors.long$Variable)))
+ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
+  geom_point() +
+  facet_wrap(~Variable, ncol=3 )
+
+cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
+                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
+                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
+                             "Protozoa"))
+ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
+  geom_point(col="red", size=3) +
+  facet_wrap(~Variable, ncol=3 )
+
 ```

### 2015-10-20 article/article.tex
update

@@
 demonstration of chaotic dynamics and limited forecast horizons (sensu
-\textcite{Petchey2015}) are important in the field of ecology, since the
+\textcite{Petchey2015}) is important in the field of ecology, since the
 ability to predict dynamics is an open question with considerable
@@
 \href{http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf}{the
-Supplement to the Nature paper}, and communications with Elisa Benincà
-and Stephen Ellner. The latter provided code used to produce results in
-the original paper, and its use in this reproduction is indicated below.
+Supplement to the Nature paper}, and communications with Elisa Benincà,
+who provided for comparison digitised data from the original article,
+and Stephen Ellner, who provided code and data used to produce results
+in the original paper. Use of this data and code in this reproduction is
+indicated below.
 
@@
 between bacteria and phosphorus. All of these correlations were
-reproduced.
+qualitatively reproduced.
 
@@
 
-Reproduced divergence rates (figure \ref{fig:divergence}) and comparison
-of the original and reproduced Lyapunov exponents (figure
-\ref{fig:LE_comparison}). The original article states: ``the distance
-between initially nearby trajectories increased over time, and reached a
-plateau after about 20--30 days''. The reproduced results appear not
-inconsistent with this statement, except for one group of species
-(Harpacticoids). The original article also stated that the analyses
-``yielded significantly positive Lyapunov exponents of strikingly
-similar value for all species (Fig. 3; mean exponent = 0.057 per day,
-s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very similar
-mean value, but had about four times greater standard deviation (mean =
-0.055 and s.d. = 0.019).
+Reproduced divergence rates (figure \ref{fig:divergence}) and Lyapunov
+exponents (figure \ref{fig:LE_comparison}) were somewhat different from
+those in the original article.. The original article states: ``the
+distance between initially nearby trajectories increased over time, and
+reached a plateau after about 20--30 days''. The reproduced results
+appear not inconsistent with this statement, except for one group of
+species (Harpacticoids). The original article also stated that the
+analyses ``yielded significantly positive Lyapunov exponents of
+strikingly similar value for all species (Fig. 3; mean exponent = 0.057
+per day, s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very
+similar mean value, but had about four times greater standard deviation
+(mean = 0.055 and s.d. = 0.019).
 
@@
 between predicted and observed abundances) and prediction distance
-(days) (figure 2 in the original article).}
+(days) (figure 2 in the original article). Reproducted data in red, and
+data from original publication in black.}
 \end{figure}

### 2015-10-20 code/report.Rmd
update

@@
 
+```{r}
+original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
+opw <- gather(original_preds, Species, r_squared, 2:25)
+opw <- separate(opw, Species, c("Species", "Model"))
+
+name.mapping <- data.frame(Species=c("cyclo", "roti", "cala", "pico", "nano",
+                                   "dia", "P", "N", "bact", "ostra", "harpa",
+                                   "proto"),
+                           Variable=c("Cyclopoids", "Rotifers", "Calanoid.copepods",
+                                  "Picophytoplankton", "Nanophytoplankton",
+                                  "Filamentous.diatoms", "Soluble.reactive.phosphorus",
+                                  "Total.dissolved.inorganic.nitrogen", "Bacteria",
+                                  "Ostracods", "Harpacticoids", "Protozoa"))
+opw <- full_join(opw, name.mapping)
+
+ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) +
+  xlab("Prediction time (days)") +
+  geom_point(col="red") +
+  facet_wrap(~Variable, ncol=3 ) +
+  geom_point(data=filter(opw, Model=="nonlinear"),
+             aes(x=prediction_time, y=r_squared))
+
+```
+
 First pass working.

### 2015-10-20 code/report.Rmd
update

@@
 
-This is a reproduction of the anaylses presented in the paper *Chaos in a long-term experiment with a plankton community*, by Elisa Benincà and others ([the paper on the Nature website](http://www.nature.com/nature/journal/v451/n7180/abs/nature06512.html)).
+This document contains various reproductions of the anaylses presented in the paper *Chaos in a long-term experiment with a plankton community*, by Elisa Benincà and others ([the paper on the Nature website](http://www.nature.com/nature/journal/v451/n7180/abs/nature06512.html)).
 Details of the methods are in [the Supplement to the Nature paper](http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf).
@@
 
-The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.
+*Update* The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.
 
@@
 
-The datasets for this reproduction are read from this github repository:
+The datasets are read from this github repository:
 ```{r}
@@
 
-# First get the raw data into R and tidy it.
+# First get the raw data into R, and clean and tidy it.
 
@@
 
-First quick go:
-```{r}
-ggplot(all.data, aes(x=Day.number, y=value)) + geom_line() +
-  facet_wrap(~variable, scales="free_y")
-```
-
-Now we add a column that gives the variable types, same as in figure 1b through 1g.
-First make a lookup table giving species type:
+First add a column that gives the variable types, same as in figure 1b through 1g. Make a table giving species type:
 ```{r}
@@
 
+Perform the interpolation:
 ```{r}
 all.data1 <- na.omit(all.data)
-
-#group_by(all.data1, variable) %>% summarise(out=min(Day.number))
-
 mt <- plyr::dlply(all.data1,
@@
 ## Aside: the duplicated date that was previously fixed was only discovered by a warning message
-## given by the pracma::interp1 function!!!
+## given by the pracma::interp1 function!
 
@@
 mt <- gather(mt, variable, value, 2:13)
-#ggplot(mt, aes(x=Day.number, y=value)) + facet_wrap(~variable, scales="free") + geom_line()
 ```
@@
 
-from.steve <- read.csv(text=getURL(paste0(repo,"/data/direct_from_Steve/interp_short_allsystem_newnames.csv")), header=T)
+from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
 
@@
 
-Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros, whereas we don't do any zero removal (this is the same as in Ellner's code.)
+Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros.
 
@@
 
-Get p-vals too:
+Get p-vals:
 ```{r}
@@
 # read script lines from website
-script <- getURL(paste0(repo,"/report/functions/indirect_method_functions.R"), ssl.verifypeer = FALSE)
+script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)
 
@@
 Get the correlations and do some house keeping:
-
 ```{r}
@@
 And plot our version of figure 2:
-
 ```{r}
-levels(cors.long$Variable)
-ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
-  geom_point() +
-  facet_wrap(~Variable, scales="free_y", ncol=3 )
-
-cors.long$Variable <- factor(cors.long$Variable, sort(levels(cors.long$Variable)))
-ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
-  geom_point() +
-  facet_wrap(~Variable, ncol=3 )
 
@@
                              "Protozoa"))
-ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
-  geom_point(col="red", size=3) +
-  facet_wrap(~Variable, ncol=3 )
+# ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
+#   geom_point(col="red", size=3) +
+#   facet_wrap(~Variable, ncol=3 )
 
-```
-
-```{r}
+## read in the data of the original article that came direct from Elisa
 original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
@@
 
-First pass working.
-Need to double check everything.
 Quite different patterns from in figure 2 of the nature paper.

### 2015-11-02 code/irlgraph/R/impute_na.R
copy irlgraph code

@@ -1,0 +1,38 @@
+#'@name impute_na
+#'@title Impute NA cells according to nearest neighbor
+#'@export
+#'@importFrom spatstat as.ppp nncross
+#'@importFrom raster xyFromCell cellFromXY
+#'@param csurf RasterLayer
+#'@param result RasterLayer
+#'@param warn logical print warnings?
+#'@param graph igraph object
+
+impute_na <- function(result, csurf, graph, warn = TRUE){
+  
+  cells <- graph$cells
+  infcells <- cells[which(result[cells]==Inf)]
+  
+  nonnullcells <- (1:length(csurf))[!(1:length(csurf) %in% graph$cells)]
+  nonnullcells <- nonnullcells[!(nonnullcells %in% graph$nullcells)]
+  nonnullcells <- c(nonnullcells, infcells)
+  
+  cells <- cells[!cells %in% nonnullcells]
+  
+  cellcoords <- raster::xyFromCell(csurf, cells)
+  nonnullcellcoords <- raster::xyFromCell(csurf, nonnullcells)
+  
+  fcells <- spatstat::as.ppp(cellcoords,c(0,dim(csurf)[1],0,dim(csurf)[2]))
+  icells <- spatstat::as.ppp(nonnullcellcoords,c(0,dim(csurf)[1],0,dim(csurf)[2]))
+  ncross <- spatstat::nncross(icells,fcells, what = "which")
+  
+  if(length(ncross)<2){
+    if(warn==TRUE){
+      warning("No cells found needing to impute. Check that scoords exist in graph and re-run.")
+    }
+    result
+  }else{
+    result[nonnullcells] <- result[raster::cellFromXY(csurf, cellcoords[ncross,])]
+    result
+  }
+}

### 2015-11-02 code/irlgraph/R/irl_graph.R

@@ -1,0 +1,157 @@
+#'@name irl_graph
+#'@title Generate an irregular landscape graph
+#'@param dm matrix cost raster
+#'@param poicoords matrix
+#'@param grainprop numeric
+#'@param cutoff numeric
+#'@param irregular logical
+#'@import sp
+#'@importFrom Matrix Matrix
+#'@importFrom igraph graph.adjacency E
+#'@importFrom raster raster extract boundaries cellFromCol cellFromRow cellFromXY xyFromCell
+#'@importFrom geometry delaunayn
+#'@export
+#'@examples
+#'
+#'dm<-matrix(c( 1,  1,  1,  1,  1,
+#'  NA, NA, NA, NA, 1, 
+#'  1,  1,  1,  1,  1,  
+#'  1,  1,  1,  1,  1,  
+#'  1,  1,  1,  1,  1
+#'  ), ncol = 5, byrow = TRUE)
+#'
+#'sp::plot(raster::raster(dm))
+#'raster::text(raster::raster(dm), which(raster::raster(dm)[] == 1))    
+#'
+#'graph <- irl_graph(dm)
+#'plot(graph$graph)
+#'geometry::trimesh(graph$tri$tri, graph$allcoords)
+
+irl_graph <- function(dm, poicoords = NA, grainprop = 0.25, cutoff = 0, irregular = TRUE){
+  csurf <- raster::raster(nrows=dim(dm)[1], ncols=dim(dm)[2], resolution=1, xmn=0, xmx = dim(dm)[1], ymn = 0, ymx = dim(dm)[2])
+  csurf[] <- dm
+  
+  get_cells <- function(dm, grainprop, poicoords, cutoff = 0){
+    
+    csurf <- raster::raster(nrows=dim(dm)[1], ncols=dim(dm)[2], resolution=1, xmn=0, xmx = dim(dm)[1], ymn = 0, ymx = dim(dm)[2])
+    csurf[] <- dm
+    
+    xmax <- dim(csurf)[1]
+    ymax <- dim(csurf)[2]
+    
+    allcells <- raster::rasterToPoints(is.na(csurf))
+    nullcells <- which(is.na(raster::extract(csurf,allcells[,1:2])))
+    limitcells <- c(raster::cellFromRow(csurf,c(1,ymax)),raster::cellFromCol(csurf,c(1,xmax)))
+    limitcells <- limitcells[!(limitcells %in% nullcells)]
[…]

### 2016-01-06 article/viejo_girard_khamassi.tex
Version ready for submission

--- 2015-09-25 article/article.tex
@@
             bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
+            pdfauthor=
+{
+      Guillaume Viejo,
+      Benoît Girard,
+      Mehdi Khamassi,
+  },
+            pdftitle=
+{
+[Re] Speed/accuracy trade-off between the habitual and the goal-directed
+processes
+},
             colorlinks=true,
@@
 ]{biblatex}
-\bibliography{article.bib}
+\bibliography{viejo_girard_khamassi.bib}
 
@@
 % --- Document ----------------------------------------------------------------
-\title{[Re] Chaos in a long-term experiment with a plankton community}
+\title{[Re] Speed/accuracy trade-off between the habitual and the goal-directed
+processes}
 
     \usepackage{authblk}
-                        \author[1]{Owen Petchey}
-                    \author[1]{Marco Plebani}
-                    \author[1]{Frank Pennekamp}
-                            \affil[1]{Institute of Evolutionary Biology and Environmental Studies, University
-of Zurich, Zurich, Switzerland}
+                        \author[1]{Guillaume Viejo}
+                    \author[1]{Benoît Girard}
+                    \author[1]{Mehdi Khamassi}
+                            \affil[1]{Sorbonne Universités, UPMC Univ Paris 06, CNRS, Institute of Intelligent
+Systems and Robotics (ISIR), F-75005 Paris, France}
 
 \date{\vspace{-5mm}
-      \sffamily \small \href{mailto:owen.petchey@ieu.uzh.ch}{owen.petchey@ieu.uzh.ch}}
+      \sffamily \small \href{mailto:guillaume.viejo@isir.upmc.fr}{guillaume.viejo@isir.upmc.fr}}
 
@@
 
-  {\bfseries Received}  Sep, 1, 2015\\
+  {\bfseries Received}  Nov, 1, 2015\\
   {\bfseries Accepted}  Sep, 1, 2015\\
@@
 \begin{itemize}
-    \item[→] Benincà, E., Huisman, J., Heerkloss, R., Jöhnk, K.D., Branco, P., Van
-Nes, E.H., Scheffer, M. \& Ellner, S.P. (2008) Chaos in a long-term
-experiment with a plankton community. Nature, 451, 822--825 DOI:
-10.1038/nature06512
+    \item[→] Speed/accuracy trade-off between the habitual and the goal-directed
+processes, M. Keramati, A. Dezfouli, P. Payam, PLoS computational
+biology, 7, 2011
   \end{itemize}\par
@@
 
-The original paper describes analyses of fluctuations in the abundance
-of organisms in a plankton community derived from the Baltic Sea, housed
-in a laboratory environment. The length of the time series (samples
-every few days for 2,300 days) allowed for analyses revealing that the
-observed dynamics exhibited characteristics consistent with chaos
-produced by species interactions. The article concludes that stability
-is not required for persistence of complex food webs, and that long-term
-prediction of abundances may be fundamentally impossible. The
-demonstration of chaotic dynamics and limited forecast horizons (sensu
-\textcite{Petchey2015}) are important in the field of ecology, since the
-ability to predict dynamics is an open question with considerable
-applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.
+This study is a reference implementation of \textcite{keramati} that
+proposed an arbitration mechanism between a goal-directed strategy and a
+habitual strategy, used to model the behavior of rats in instrumental
+conditionning tasks. The habitual strategy is the Kalman Q-Learning from
+\textcite{geist}. We replicate the results of the first task i.e.~the
+devaluation experiment with two states and two actions. The
+implementation is in python with numpy, scipy and matplotlib library.
+The authors couldn't provide the original implementation and we are not
+aware of others implementations elsewhere.
 
@@
 
-This reproduction started with the raw data (source given below) and
-used information from the original paper, the supplementary information
-\href{http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf}{the
-Supplement to the Nature paper}, and communications with Elisa Benincà
-and Stephen Ellner. The latter provided code used to produce results in
-the original paper, and its use in this reproduction is indicated below.
+We used the description of the model from the original article except
+for the implementation of the Kalman Q-Learning which we took from
+\textcite{geist}. We used the same parameters as the original article
+except for the update rate of the transition function \(\phi\), the
+initialization of the covariance matrice and an uncentered transform
+parameter \(\kappa\) that were not mentionned in the original article.
+The largest uncertainty about the model concerned the devaluation
+procedure. Besides setting the reward \(r\) to null, the authors stated
+that ``For modeling the devaluation of the outcome in the first two
+simulations, \(R(S_1, EM)\) is set to -1.'' As this notation
+(\(R(S_1, EM)\)) is not defined in the rest of the article, we assumed
+that it is \(\hat{R}(S_1, EM)\) updated by equation (14) in the original
+article.
 
-\subsection{Scope of the reproduction}\label{scope-of-the-reproduction}
+The parameters are as follows :
 
-An attempt was made to reproduce the majority of the results in the
-original article. Instances where we did not attempt to reproduce a
-result are detailed below.
+\begin{longtable}[c]{@{}lll@{}}
+\toprule
+Name & Description & Value\tabularnewline
+\midrule
+\endhead
+\(\sigma\) & Updating rate of the average reward & 0.02\tabularnewline
+\(\eta\) & Variance of evolution noise & 0.0001\tabularnewline
+\(P_n\) & Variance of observation noise & 0.05\tabularnewline
+\(\beta\) & Rate of exploration & 1.0\tabularnewline
+\(\rho\) & Update rate of the reward function & 0.1\tabularnewline
+\(\gamma\) & Discount factor & 0.95\tabularnewline
+\(\tau\) & Time step of graph exploration & 0.08\tabularnewline
+depth & Depth of search in graph exploration & 3\tabularnewline
+\(\phi\) & Update rate of the transition function & 0.5\tabularnewline
+init cov & Initialisaton of covariance matrice & 1.0\tabularnewline
+\(\kappa\) & Unscentered transform parameters & 0.1\tabularnewline
+\bottomrule
+\end{longtable}
 
-\subsection{The data}\label{the-data}
+We describe the algorithm of our implementation in details. The process
+of action selection and reward update are separated for clarity.
 
-The data are available as an Excel file supplement to
-\href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
-Ecology Letters publication} \textcite{Beninca2009}. The Excel file
-contains several datasheets. Two were particularly important, as they
-are the source of the raw data (one contains original species
-abundances, the one with the nutrient concentrations). We saved these
-two datasheets as comma separated value (csv) text files. In the code
-associated with this reproduction, these data files are read from the
-associated github repository.
+\begin{quote}
+\begin{quote}
+Initialization
+\end{quote}
+\end{quote}
 
-Another datasheet in the Ecology Letters supplement contains transformed
-variables (we also saved this as csv file, in order to use it in this
-reproduction). We also received a dataset direct from Steve Ellner, see
-below for details.
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\[Q(s, a)^{Goal-Directed} = \{ 0, \ldots \}\]
+\end{quote}
+\end{quote}
+\end{quote}
 
-\subsection{Reproduction environment}\label{reproduction-environment}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\[Q(s, a)^{Habitual} = \{ 0, \dots \}\]
+\end{quote}
+\end{quote}
+\end{quote}
 
-The R language and environment for statistical computing and graphics
-was used to make the reproduction. Additional R packages required are
-specified in the code associated with this reproduction.
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\[
+\Sigma = \left(
+    \begin{array}{*4{c}}
+    cov \times \eta &  0 &  \ldots & 0 \\
+     0 &  cov \times \eta &  \ldots & \vdots \\
+    \vdots &  \ldots &  \ddots & 0 \\
+     0 &  \ldots &  0 & cov \times \eta \\     
+  \end{array}\right)
+\]
+\end{quote}
+\end{quote}
+\end{quote}
 
-The code for this reproduction resides in an R markdown document, as
-well as a source file containing some required functions. Some of the
-code takes several minutes to run, so an intermediate data file is
-provided with results from this code.
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(R(S1, EM) = 1\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\bar{R} = 0\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\hat{R}(s,a) = \{0,\ldots\}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+Main Loop
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{FOR}\ i = 1:T\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(s_t = S_0\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{IF}\ i = T_{devaluation}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(R(S1, EM) = 0\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\hat{R}(S1, EM) = -1\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{WHILE}\ s_t \neq S1 \bigwedge a_t \neq EM\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(a_t = \textbf{Selection}(s_t)\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(r_t = R(s_t,a_t)\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(s_{t+1} = transition(s_t, a_t)\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{Update}(s_t,a_t, s_{t+1}, r_t)\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+Selection
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\{a_1,\ldots,ai,\ldots\} \leftarrow sort(Q(s_t,a_i))\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(VPI(s_t, a_1) = (Q(s_t,a_2)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_1)^{H}<Q(s_t,a_2)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_2)^H - Q(s_t,a_1)^H)^2}{2\sigma(s_t,a_t)^2}}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(VPI(s_t, a_i) = (Q(s_t,a_i)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_i)^{H}>Q(s_t,a_1)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_1)^H - Q(s_t,a_i)^H)^2}{2\sigma(s_t,a_t)^2}}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{FOR}\ i \in \{a_1, a_2,\ldots, a_i, \ldots\}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{IF}\ VPI(s_t, a_i) \geq \tau \bar{R}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(Q(s_t,a_i) = \hat{R}(s_t,a_i) + \gamma \sum\limits_{s'}p_{T}(\{s,a\}\rightarrow s') \max\limits_{b \in A} Q(s',b)^{Goal-directed}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{ELSE}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(Q(s_t,a_i) = Q(s_t,a_i)^{Habitual}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(a_t \leftarrow \textit{SoftMax}(Q(s_t,a), \beta)\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+Update
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\bar{R} = (1-\sigma) \bar{R} + \sigma r_t\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\hat{R}(s_t,a_t) =(1 - \rho) \hat{R} + \rho r_t\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(p_{T}(s_t, a_t, s_{t+1}) = (1 - \phi) p_{T}(s_t, a_t, s_{t+1}) + \phi\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+Specific to Kalman Q-Learning
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\Theta = \{ \theta_j, 0 \geq j \geq 2|S.A|\}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\check{W} = \{ w_j, 0 \geq j \geq 2|S.A| \}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\check{R} = \{ \check{r}_j = \theta_j(s_t,a_t) - \gamma \max\limits_{b \in A} \theta_j(s_{t+1},b),\ 0 \geq j \geq 2|S.A|\}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(r_{predicted} = \sum\limits_{j=0}^{2|S.A|} w_j \check{r}_j\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(P_{\theta_j \check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\theta_j - Q^{Habitual}_t)(\check{r}_j - r_{predicted})\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(P_{\check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\check{r}_j - r_{predicted})^2 + P_n\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(K_t = P_{\theta_j \check{r}_j} P_{\check{r}_j}^{-1}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\delta_t = r_t - r_{predicted}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(Q_{t+1}^{Habitual} = Q_{t}^{H} + K_t \delta_t\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(P_{t+1}^{H} = P_{t}^{H} - K_t P_{\Sigma_t} K_t^T\)
+\end{quote}
+\end{quote}
+\end{quote}
 
@@
 
-\subsection{Population dynamics}\label{population-dynamics}
+We only reproduced the results of Figure 3 A, B, G, H in a qualitative
+manner. Results are presented in Figure \ref{fig:figure1}. We can
+observe the strategy shift (from goal-directed to habitual) after
+extensive training around 50 time steps. In the original article, the
+strategy shift occurs after 100 time steps.
 
-The reproduced populations dynamics were at least very similar to those
-in figure 1b-g of the original publication (figure \ref{fig:dynamics}).
-Note that we plot fourth root transformed values, rather than raw
-abundances with a y-axis break, as in the original article.
+However we can observe a difference between the probabilities of action
+for the goal-directed model. In our implementation,
+\[p(s_0, pl) \simeq 0.7\] and \[p(s_0, em) \simeq 0.3\] before
+devaluation. In the original article, \[p(s_0, pl) \simeq 0.6\] and
+\[p(s_0, em) \simeq 0.4\] Nevertheless, the probabilities of action from
+the Kalman Q-Learning after strategy shifting are equivalent.
 
@@
 \centering
-\includegraphics{figures/dynamics.pdf}
-\caption{\label{fig:dynamics}Observed population dynamics.}
-\end{figure}
-
-\subsection{Data transformation}\label{data-transformation}
-
-The following transformation steps were used:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Time series shortened to remove long sequences of zeros.
-\item
-  Interpolation to create equally spaced observations in time series.
-\item
-  Fourth root transformation.
-\item
-  Detrending of five of the time series.
-\item
-  Rescaling to zero mean and unit standard deviation.
-\end{enumerate}
-
-The method for selecting the zeros to remove was unclear. In order to
-reproduce this step, we removed the same data as in the original study,
-by matching to the transformed data with zeros removed in the Ecology
-Letters supplement Excel file mentioned above. All remaining
-transformation steps were performed independently of this dataset. Our
-reproduction of the transformed data closely matched the published
-transformed data.
-
-The data received directly from Stephen Ellner was interpolated, but
-without zeros removed. Our interpolated data, without zeros removed,
-matched closely this data.
-
-\subsection{Correlations among species
-abundances}\label{correlations-among-species-abundances}
-
-Correlations among species abundances presented in Table 1 of the
-original article closely matched our reproduced correlations, calculated
-from the transformed data with zeros removed (figure
-\ref{fig:corr_comp}). Deviations between the original and reproduced
-correlations are relatively small (less than 0.072 units) and
-infrequent.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/correlation_comparison.pdf}
-\caption{\label{fig:corr_comp}Comparison of calculated correlations
-among species abundances in the original article and this reproduction.}
-\end{figure}
-
-Highlighted in the text of the original paper were: negative
-correlations of picophytoplankton with protozoa, and of
-nanophytoplankton both with rotifers and calanoid copepods, positive
-correlation of picophytoplankton with calanoid copepods, negative
-correlation between bacteria and ostracods, and positive correlation
-between bacteria and phosphorus. All of these correlations were
-reproduced.
-
-\subsection{Spectral analyses}\label{spectral-analyses}
-
-Spectral analyses in the original paper were presented graphically in
-figures S3 (raw spectrograms) and S4 (Welch periodograms). These graphs
-were, apparently, visually inspected to reveal: ``fluctuations covered a
-range of different periodicities'', and ``picophytoplankton, rotifers
-and calanoid copepods seemed to fluctuate predominantly with a
-periodicity of about 30 days.'' It is unclear how these conclusions were
-derived from figures S3 and S4 of the original article. Our reproduced
-spectra (not shown here, but code provided) were not quantitatively
-identical to the spectra in the original article.
-
-\subsection{Lyapunov exponents by direct
-method}\label{lyapunov-exponents-by-direct-method}
-
-Reproduced divergence rates (figure \ref{fig:divergence}) and comparison
-of the original and reproduced Lyapunov exponents (figure
-\ref{fig:LE_comparison}). The original article states: ``the distance
-between initially nearby trajectories increased over time, and reached a
-plateau after about 20--30 days''. The reproduced results appear not
-inconsistent with this statement, except for one group of species
-(Harpacticoids). The original article also stated that the analyses
-``yielded significantly positive Lyapunov exponents of strikingly
-similar value for all species (Fig. 3; mean exponent = 0.057 per day,
-s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very similar
-mean value, but had about four times greater standard deviation (mean =
-0.055 and s.d. = 0.019).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/unnamed-chunk-50-1.pdf}
-\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
-exponents (figure 3 in the original article).}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/LE_comparison.pdf}
-\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents in
-the original article and this reproduction.}
-\end{figure}
-
-\subsection{Lyapunov exponents by indirect
-method}\label{lyapunov-exponents-by-indirect-method}
-
-The original paper reported global Lyapunov exponent calculated via two
-modelling approaches (neural network and generalised additive models
-{[}GAMs{]}). Only the GAM approach was reproduced, with the assistance
-of code donated by Stephen Ellner. The original article obtained a
-global Lyapunov exponent of λ=0.08 day-1. The reproduced value was 0.04.
-We did not reproduce the bootstrapping used to give confidence intervals
-around this estimate.
-
-\subsection{Predictability decay}\label{predictability-decay}
-
-The article stated: ``For short-term forecasts of only a few days, most
-species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
-However, the predictability of the species was much reduced when
-prediction times were extended to 15--30days.'' The reproduced
-predictabilities, which were calculated from the GAMs, were consistent
-with these qualitative statements, though were quantitatively different
-(\ref{fig:prediction_distance}). We did not reproduce the predictability
-estimates for linear models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/prediction_distance.pdf}
-\caption{\label{fig:prediction_distance}Predictability (correlation
-between predicted and observed abundances) and prediction distance
-(days) (figure 2 in the original article).}
+\includegraphics{../code/fig.pdf}
+\caption{\label{fig:figure1}A. Value of Precise Information (full lines)
+for action press-lever and enter magazine in state \(S_0\) and reward
+rate (dashed line) in moderate training. Vertical line represents the
+timing of devaluation. B. In extensive training. C. Probability of
+actions in state \(S_0\) in moderate training. D. In extensive
+training.}
 \end{figure}
@@
 
-Although we were not able to make a quantitatively accurate reproduction
-of all results of the original article, the qualitative results were
-largely identical. Quantitative differences may have resulted from
-difference in algorithms used. For example,the original used the
-\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
-calculate Lyapunov exponents. As this was available from CRAN
-\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
-mid 2014} and since it is a bit less well integrated with R, we instead
-use the tseriesChaos package \textcite{tseriesChaos}, which in any case
-was largely inspired by the TISEAN project. There may also have been
-some difference in data used for specific analyses, e.g., data with
-zeros removed or not, as it was not always possible to be sure the
-reproduction used exactly the same data.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-This reproduction was made as part of the Reproducible Research in
-Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES)
-Course, lead by Owen Petchey at the University of Zurich. More
-information about the course
-\href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
-on github.
+We were able to qualitatively reproduce the first simulations of the
+article. Despite the small differences in the exact timing of the
+strategy shifting and in the probabilities of action, the behavior of
+our implementation is similar to the original article. Thus, we confirm
+the correctness of the model presented in the original article.
 

### 2016-01-06 code/fonctions.py

@@ -1,0 +1,173 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# -------------------------------------------------------------------------
+# Copyright (c) 2015, Guillaume Viejo
+# Distributed under the (new) BSD License.
+#
+# Contributors : Guillaume Viejo (guillaume.viejo@isir.upmc.fr)
+# -------------------------------------------------------------------------
+# References:
+#
+# Speed/accuracy trade-off between the habitual and the goal-directed processes.
+# M. Keramati, A. Dezfouli, P. Piray. Plos Comp Bio, 7(5), 2011
+# -------------------------------------------------------------------------
+
+import numpy as np
+from copy import deepcopy
+from scipy.stats import norm
+
+# -----------------------------------
+# FONCTIONS
+# -----------------------------------
+def createQValuesDict(states, actions):
+    #return a big dict
+    values = {0:np.zeros(len(states)*len(actions))}
+    tmp = 0
+    for s in states:
+        values[s] = []
+        n = 0
+        for a in actions:
+            values[(s,a)] = tmp
+            values[s].append(tmp)
+            values[(s,n)] = a
+            tmp = tmp + 1
+            n = n + 1
+    return values
+
+def transitionRules(state, action):
+    if state == 's0' and action == 'pl':
+        return 's1'
+    else:
+        return 's0'
+
+
+
+def createRewardRateDict():
+    return dict({'rate':0,
+                 'reward':0,
+                 'tau':0,
+                 'step':0,
+                 'tau_list':[0],
+                 'r_list':[0],
+                 ('s0', 'pl'):8,
+                 ('s0', 'em'):12,
+                 ('s1', 'pl'):12,
+                 ('s1', 'em'):0,
+                 'r*tau':[]})
+
+def createCovarianceDict(n, init_cov, eta):
+    #cov = np.eye(n)*init_cov+(1-np.eye(n,n))*((1-init_cov)/(n-1))    
+    cov = np.eye(n)
+    return dict({'cov':cov,
+                 'noise':np.eye(n,n)*init_cov*eta})
+
+
+def getBestAction(state, values, ind = 0):
+    return values[(state, np.where(values[ind][values[state]] == np.max(values[ind][values[state]]))[0][0])]
+
+def getBestActionSoftMax(state, values, beta, ind = 0):
[…]

### 2016-01-20 article/viejo_girard_khamassi.tex
First Submission for Viejo-Khamassi-Girard-2016

@@
 \toprule
-Name & Description & Value\tabularnewline
+\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
+Name
+\strut\end{minipage} &
+\begin{minipage}[b]{0.45\columnwidth}\raggedright\strut
+Description
+\strut\end{minipage} &
+\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
+Value
+\strut\end{minipage}\tabularnewline
 \midrule
 \endhead
-\(\sigma\) & Updating rate of the average reward & 0.02\tabularnewline
-\(\eta\) & Variance of evolution noise & 0.0001\tabularnewline
-\(P_n\) & Variance of observation noise & 0.05\tabularnewline
-\(\beta\) & Rate of exploration & 1.0\tabularnewline
-\(\rho\) & Update rate of the reward function & 0.1\tabularnewline
-\(\gamma\) & Discount factor & 0.95\tabularnewline
-\(\tau\) & Time step of graph exploration & 0.08\tabularnewline
-depth & Depth of search in graph exploration & 3\tabularnewline
-\(\phi\) & Update rate of the transition function & 0.5\tabularnewline
-init cov & Initialisaton of covariance matrice & 1.0\tabularnewline
-\(\kappa\) & Unscentered transform parameters & 0.1\tabularnewline
+\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
+\(\sigma\) \(\eta\) \(P_n\) \(\beta\) \(\rho\) \(\gamma\) \(\tau\) depth
+\(\phi\) init cov \(\kappa\)
+\strut\end{minipage} &
+\begin{minipage}[t]{0.45\columnwidth}\raggedright\strut
+Updating rate of the average reward Variance of evolution noise Variance
+of observation noise Rate of exploration Update rate of the reward
+function Discount factor Time step of graph exploration Depth of search
+in graph exploration Update rate of the transition function
+Initialisaton of covariance matrice Unscentered transform parameters
+\strut\end{minipage} &
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+0.02 0.0001 0.05 1.0 0.1 0.95 0.08 3 0.5 1.0 0.1
+\strut\end{minipage}\tabularnewline
 \bottomrule

### 2016-02-09 article/viejo_girard_khamassi.tex
Update after first round of reviewing

--- 2016-01-06 article/viejo_girard_khamassi.tex
@@
 \(\phi\) & Update rate of the transition function & 0.5\tabularnewline
-init cov & Initialisaton of covariance matrice & 1.0\tabularnewline
+init cov & Initialisation of covariance matrice & 1.0\tabularnewline
 \(\kappa\) & Unscentered transform parameters & 0.1\tabularnewline
@@
 \begin{quote}
+\textbf{Initialization}
+\end{quote}
+
 \begin{quote}
-Initialization
+\begin{quote}
+\(Q(s, a)^{Goal-Directed} = \{ 0, \ldots \}\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(Q(s, a)^{Habitual} = \{ 0, \dots \}\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\#\) Covariance matrix
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\Sigma = \left(  \begin{array}{*4{c}}  cov \times \eta & 0 & \ldots & 0 \\  0 & cov \times \eta & \ldots & \vdots \\  \vdots & \ldots & \ddots & 0 \\  0 & \ldots & 0 & cov \times \eta \\  \end{array}\right)\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(R(S1, EM) = 1\) \(\#\) Reward value
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\bar{R} = 0\) \(\#\) Reward rate
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\hat{R}(s,a) = \{0,\ldots\}\) \(\#\) Reward function
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{Main Loop}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\textbf{FOR}\ i = 1:T\)
 \end{quote}
@@
 \begin{quote}
-\[Q(s, a)^{Goal-Directed} = \{ 0, \ldots \}\]
+\(s_t = S_0\) \(\#\) Initial state
 \end{quote}
@@
 \begin{quote}
-\[Q(s, a)^{Habitual} = \{ 0, \dots \}\]
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\[
-\Sigma = \left(
-    \begin{array}{*4{c}}
-    cov \times \eta &  0 &  \ldots & 0 \\
-     0 &  cov \times \eta &  \ldots & \vdots \\
-    \vdots &  \ldots &  \ddots & 0 \\
-     0 &  \ldots &  0 & cov \times \eta \\     
-  \end{array}\right)
-\]
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(R(S1, EM) = 1\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\bar{R} = 0\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\hat{R}(s,a) = \{0,\ldots\}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-Main Loop
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\textbf{FOR}\ i = 1:T\)
+\(\textbf{IF}\ i = T_{devaluation}\) \(\#\) Moderate / Extensive
+training
 \end{quote}
@@
 \begin{quote}
-\(s_t = S_0\)
+\(R(S1, EM) = 0\)
 \end{quote}
@@
 \begin{quote}
-\(\textbf{IF}\ i = T_{devaluation}\)
+\(\hat{R}(S1, EM) = -1\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{WHILE}\ s_t \neq S1 \bigwedge a_t \neq EM\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(a_t = \textbf{Selection}(s_t)\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(R(S1, EM) = 0\)
-\end{quote}
+\(r_t = R(s_t,a_t)\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(\hat{R}(S1, EM) = -1\)
-\end{quote}
+\(s_{t+1} = transition(s_t, a_t)\)
 \end{quote}
@@
 \begin{quote}
-\(\textbf{WHILE}\ s_t \neq S1 \bigwedge a_t \neq EM\)
+\(\textbf{Update}(s_t,a_t, s_{t+1}, r_t)\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{Selection}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\#\) Sort the Q-values in descending order
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\{a_1,\ldots,ai,\ldots\} \leftarrow sort(Q(s_t,a_i))\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\#\) VPI : Value of Precise Information
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(VPI(s_t, a_1) = (Q(s_t,a_2)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_1)^{H}<Q(s_t,a_2)^{H})+ \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_2)^H - Q(s_t,a_1)^H)^2}{2\sigma(s_t,a_t)^2}}\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(VPI(s_t, a_i) = (Q(s_t,a_i)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_i)^{H}>Q(s_t,a_1)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_1)^H - Q(s_t,a_i)^H)^2}{2\sigma(s_t,a_t)^2}}\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\(\textbf{FOR}\ i \in \{a_1, a_2,\ldots, a_i, \ldots\}\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\textbf{IF}\ VPI(s_t, a_i) \geq \tau \bar{R}\)
+\end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\#\) Q-Value from Goal-directed system is evaluated
 \end{quote}
@@
 \begin{quote}
+\(Q(s_t,a_i) = \hat{R}(s_t,a_i) + \gamma \sum\limits_{s'}p_{T}(\{s,a\}\rightarrow s') \max\limits_{b \in A} Q(s',b)^{Goal-directed}\)
+\end{quote}
+\end{quote}
+\end{quote}
+\end{quote}
+
 \begin{quote}
-\(a_t = \textbf{Selection}(s_t)\)
+\begin{quote}
+\begin{quote}
+\(\textbf{ELSE}\)
 \end{quote}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\begin{quote}
+\(\#\) Q-Value from Habitual system is retrieved
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(r_t = R(s_t,a_t)\)
-\end{quote}
+\(Q(s_t,a_i) = Q(s_t,a_i)^{Habitual}\)
 \end{quote}
@@
 \begin{quote}
+\(a_t \leftarrow \textit{SoftMax}(Q(s_t,a), \beta)\)
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{Update}
+\end{quote}
+
 \begin{quote}
 \begin{quote}
-\begin{quote}
-\(s_{t+1} = transition(s_t, a_t)\)
-\end{quote}
-\end{quote}
-\end{quote}
+\(\bar{R} = (1-\sigma) \bar{R} + \sigma r_t\) \(\#\) Reward Rate
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\textbf{Update}(s_t,a_t, s_{t+1}, r_t)\)
-\end{quote}
-\end{quote}
-\end{quote}
+\(\hat{R}(s_t,a_t) =(1 - \rho) \hat{R} + \rho r_t\) \(\#\) Reward
+function
 \end{quote}
@@
 \begin{quote}
-Selection
+\(p_{T}(s_t, a_t, s_{t+1}) = (1 - \phi) p_{T}(s_t, a_t, s_{t+1}) + \phi\)
+\(\#\) Probability of transition
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(\{a_1,\ldots,ai,\ldots\} \leftarrow sort(Q(s_t,a_i))\)
-\end{quote}
+Specific to Kalman Q-Learning
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(VPI(s_t, a_1) = (Q(s_t,a_2)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_1)^{H}<Q(s_t,a_2)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_2)^H - Q(s_t,a_1)^H)^2}{2\sigma(s_t,a_t)^2}}\)
-\end{quote}
+\(\#\) Sigma-points sampling
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(VPI(s_t, a_i) = (Q(s_t,a_i)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_i)^{H}>Q(s_t,a_1)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_1)^H - Q(s_t,a_i)^H)^2}{2\sigma(s_t,a_t)^2}}\)
-\end{quote}
+\(\Theta = \{ \theta_j, 0 \geq j \geq 2|S.A|\}\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(\textbf{FOR}\ i \in \{a_1, a_2,\ldots, a_i, \ldots\}\)
-\end{quote}
+\(\check{W} = \{ w_j, 0 \geq j \geq 2|S.A| \}\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\begin{quote}
-\(\textbf{IF}\ VPI(s_t, a_i) \geq \tau \bar{R}\)
-\end{quote}
-\end{quote}
+\(\check{R} = \{ \check{r}_j = \theta_j(s_t,a_t) - \gamma \max\limits_{b \in A} \theta_j(s_{t+1},b),\ 0 \geq j \geq 2|S.A|\}\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(Q(s_t,a_i) = \hat{R}(s_t,a_i) + \gamma \sum\limits_{s'}p_{T}(\{s,a\}\rightarrow s') \max\limits_{b \in A} Q(s',b)^{Goal-directed}\)
-\end{quote}
-\end{quote}
-\end{quote}
+\(r_{predicted} = \sum\limits_{j=0}^{2|S.A|} w_j \check{r}_j\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\begin{quote}
-\(\textbf{ELSE}\)
-\end{quote}
-\end{quote}
+\(\#\) Covariance computation
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(Q(s_t,a_i) = Q(s_t,a_i)^{Habitual}\)
-\end{quote}
-\end{quote}
-\end{quote}
+\(P_{\theta_j \check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\theta_j - Q^{Habitual}_t)(\check{r}_j - r_{predicted})\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(a_t \leftarrow \textit{SoftMax}(Q(s_t,a), \beta)\)
-\end{quote}
+\(P_{\check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\check{r}_j - r_{predicted})^2 + P_n\)
 \end{quote}
@@
 \begin{quote}
-Update
+\(K_t = P_{\theta_j \check{r}_j} P_{\check{r}_j}^{-1}\) \(\#\) Kalman
+gain
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(\bar{R} = (1-\sigma) \bar{R} + \sigma r_t\)
-\end{quote}
+\(\delta_t = r_t - r_{predicted}\) \(\#\) Reward-prediction error
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(\hat{R}(s_t,a_t) =(1 - \rho) \hat{R} + \rho r_t\)
-\end{quote}
+\(Q_{t+1}^{Habitual} = Q_{t}^{H} + K_t \delta_t\)
 \end{quote}
@@
 \begin{quote}
-\begin{quote}
-\(p_{T}(s_t, a_t, s_{t+1}) = (1 - \phi) p_{T}(s_t, a_t, s_{t+1}) + \phi\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-Specific to Kalman Q-Learning
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\Theta = \{ \theta_j, 0 \geq j \geq 2|S.A|\}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\check{W} = \{ w_j, 0 \geq j \geq 2|S.A| \}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\check{R} = \{ \check{r}_j = \theta_j(s_t,a_t) - \gamma \max\limits_{b \in A} \theta_j(s_{t+1},b),\ 0 \geq j \geq 2|S.A|\}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(r_{predicted} = \sum\limits_{j=0}^{2|S.A|} w_j \check{r}_j\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(P_{\theta_j \check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\theta_j - Q^{Habitual}_t)(\check{r}_j - r_{predicted})\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(P_{\check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\check{r}_j - r_{predicted})^2 + P_n\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(K_t = P_{\theta_j \check{r}_j} P_{\check{r}_j}^{-1}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(\delta_t = r_t - r_{predicted}\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\(Q_{t+1}^{Habitual} = Q_{t}^{H} + K_t \delta_t\)
-\end{quote}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-\begin{quote}
 \(P_{t+1}^{H} = P_{t}^{H} - K_t P_{\Sigma_t} K_t^T\)
-\end{quote}
 \end{quote}

### 2016-02-09 code/fonctions.py

@@ -13,14 +13,20 @@
 # -------------------------------------------------------------------------
 
 import numpy as np
-from copy import deepcopy
 from scipy.stats import norm
 
 # -----------------------------------
 # FONCTIONS
 # -----------------------------------
 def createQValuesDict(states, actions):
-    #return a big dict
+    """
+    Returns a dictionary containing the Q-values of all state-action pairs.
+    The keys are either :
+    *states return action index
+    *(state, action index) return action name
+    *(state, action) return action index
+    * 0 return values
+    """
     values = {0:np.zeros(len(states)*len(actions))}
     tmp = 0
     for s in states:
@@ -35,71 +41,59 @@ def createQValuesDict(states, actions):
     return values
 
 def transitionRules(state, action):
+    """
+    Set before runs
+    Return next state
+    """
     if state == 's0' and action == 'pl':
         return 's1'
     else:
         return 's0'
 
-
-
-def createRewardRateDict():
-    return dict({'rate':0,
-                 'reward':0,
-                 'tau':0,
-                 'step':0,
-                 'tau_list':[0],
-                 'r_list':[0],
-                 ('s0', 'pl'):8,
-                 ('s0', 'em'):12,
-                 ('s1', 'pl'):12,
-                 ('s1', 'em'):0,
-                 'r*tau':[]})
-
 def createCovarianceDict(n, init_cov, eta):
-    #cov = np.eye(n)*init_cov+(1-np.eye(n,n))*((1-init_cov)/(n-1))    
+    """
+    Initialise covariance matrix 
+    """
     cov = np.eye(n)
     return dict({'cov':cov,
                  'noise':np.eye(n,n)*init_cov*eta})
 
-
-def getBestAction(state, values, ind = 0):
-    return values[(state, np.where(values[ind][values[state]] == np.max(values[ind][values[state]]))[0][0])]
-
 def getBestActionSoftMax(state, values, beta, ind = 0):
+    """
+    Compute soft-max function and return action
+    """
     tmp = np.exp(values[ind][values[state]]*float(beta))
     tmp = tmp/float(np.sum(tmp))
     tmp = [np.sum(tmp[0:i]) for i in range(len(tmp))]
     return values[(state, np.sum(np.array(tmp) < np.random.rand())-1)]
 
-def SoftMax(values, beta):
[…]

### 2016-02-09 code/models.py

@@ -34,76 +34,27 @@ class KalmanQLearning():
         self.action = None
         self.reward = None
 
-    def getAllParameters(self):        
-        return dict({#'gamma':[0.1,self.gamma,0.9],
-                     #'beta':[1,self.beta,5],
-                     'eta':[1.0e-6,self.eta,0.001],
-                     'var_obs':[0.01,self.var_obs,0.07]})
-                     #'init_cov':[5,self.init_cov,15]})
-
-    def setAllParameters(self, dict_p):
-        for i in dict_p.iterkeys():
-            self.setParameter(i,dict_p[i][1])
-
-    def getParameter(self, name):
-        if name == 'gamma':
-            return self.gamma
-        elif name == 'beta':
-            return self.beta
-        elif name == 'eta':
-            return self.eta
-        else:
-            print "Parameters not found"
-            sys.exit(0)     
-
-    def setParameter(self, name, value):
-        if name == 'gamma':
-            self.gamma = value
-        elif name == 'beta':
-            self.beta = value
-        elif name == 'eta':
-            self.eta = value
-        elif name == 'var_obs':
-            self.var_obs = value
-        elif name == 'init_cov':
-            self.init_cov = value
-        else:
-            print "Parameters not found"
-            sys.exit(0)
-
     def initialize(self):
+        """
+        Restart the class Kalman-Qlearning by initializing new Q-values and covariance matrix        
+        """
         self.values = createQValuesDict(self.states, self.actions)
         self.covariance = createCovarianceDict(len(self.states)*len(self.actions), self.init_cov, self.eta)
         self.state = None
         self.action = None
         self.reward = None
 
-    def chooseAction(self, state):
-        self.state = state        
-        self.covariance['noise'] = self.covariance['cov']*self.eta
-        self.covariance['cov'][:,:] = self.covariance['cov'][:,:] + self.covariance['noise']
-        self.action = getBestActionSoftMax(state, self.values, self.beta, 0)
-        return action
-
-    def updateValue(self, reward):
-        r = (reward==1)*1        
-        s = self.state
-        a = self.action
-        sigma_points, weights = computeSigmaPoints(self.values[0], self.covariance['cov'], self.kappa)
-        rewards_predicted = (sigma_points[:,self.values[(s,a)]]-self.gamma*np.max(sigma_points[:,self.values[s]], 1)).reshape(len(sigma_points), 1)
-        reward_predicted = np.dot(rewards_predicted.flatten(), weights.flatten())
[…]

### 2016-02-09 code/run.py

@@ -11,13 +11,18 @@
 # Speed/accuracy trade-off between the habitual and the goal-directed processes.
 # M. Keramati, A. Dezfouli, P. Piray. Plos Comp Bio, 7(5), 2011
 # -------------------------------------------------------------------------
-
+from __future__ import print_function
+try:
+	import xrange as range # Overrides range in Python2, does nothing in Python 3
+except Exception:
+	pass
 import numpy as np
 from fonctions import *
 from models import *
 from matplotlib import *
 from pylab import *
 import sys
+
 
 # -------------------------------------------------------------------------
 # PARAMETERS + INITIALIZATION
@@ -67,7 +72,7 @@ for i in xrange(nb_blocs):
 		selection.initialize()
 		state = 's0'
 		rewards[0][rewards[('s1','em')]] = 1.0
-		print exp, nb_trials, deval_time
+		print(exp, nb_trials, deval_time)
 		for j in xrange(nb_trials):
 			#Setting Reward
 			if j == deval_time:
@@ -94,14 +99,6 @@ meandata = {i:{j:np.mean(data[i][j], 0) for j in ['vpi', 'r', 'p']} for i in ['m
 # -----------------------------------
 # Plot
 # -----------------------------------
-params = {'backend':'pdf',
-		  'axes.labelsize':10,
-		  'text.fontsize':10,
-		  'legend.fontsize':10,
-		  'xtick.labelsize':8,
-		  'ytick.labelsize':8,
-		  'text.usetex':False}
-
 dashes = ['--', '-.', '-']
 colors = ['black', 'grey']
 

### 2016-02-11 code/report.Rmd
fixing zero removal issue

@@
 library(repmis)
+library(magrittr)
 
@@
 
-Format the dates as dates
+Format the dates as dates.
 
@@
 spp.abund$Date <- dmy(spp.abund$Date)
+# y123 <- year(spp.abund$Date) %% 100
+# year(spp.abund$Date) <- 1900 + y123
+# rm(y123)
 ```
-
-Ooops... R assumes the experiment was done in the 21st century. Shouldn't matter too much.
 
@@
 ```{r}
-which(spp.abund$Date==ymd("2096-10-28 UTC"))
-spp.abund$Date[701] <- ymd("2096-10-26 UTC")
+which(spp.abund$Date==ymd("1996-10-28 UTC"))
+spp.abund$Date[701] <- ymd("1996-10-26 UTC")
 ```
 
-Check dates match the Day.number (should give true):
+Check dates match the Day.number (should give false, as we changed a date):
 ```{r}
@@
 
-Fix the Day.number problem:
+Recalculate the Day.number variable from the dates:
 ```{r}
@@
 
-Now select only the date range used in the Nature paper.
-From the supplment *The analysis in Benincà et al.
-(Nature 2008) covered all data from 16/06/1991 until 20/10/1997*.
-(Remembering dates in the R dataframes are 2090s.)
+Now select only the date range used in the Nature paper. From the supplment *The analysis in Benincà et al. (Nature 2008) covered all data from 16/06/1991 until 20/10/1997*. 
 ```{r}
-all.data <- filter(all.data, Date>=dmy("15/06/2091") & Date<=dmy("21/10/2097"))
+all.data <- filter(all.data, Date>=dmy("15/06/1991") & Date<=dmy("21/10/1997"))
 
@@
 
-Aside: The ELE supplement contains the raw data and the transformed data, in separate data sheets. I (Owen) also got the interpolated data from Stephen Ellner directly.
+The ELE supplement contains the raw data and the transformed data, in separate data sheets. I (Owen) also got the interpolated data from Stephen Ellner directly.
 
-## Interpolation
+## Removal of long sequences of zeros
+
+> The time series of cyclopoid copepods, protozoa, filamentous diatoms, harpacticoid copepods and ostracods contained long sequences of zero values.
+This does not imply that these species were absent from the food web during these periods, but that their concentrations were below the detection limit.
+Time series dominated by many zeros can bias the statistical analysis.
+Therefore, these time series were shortened to remove long sequences of zero values, before the data transformation.
+
+The exact method for identifying the "long sequences of zeros" was not given.
+However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
+
+```{r}
+tr <- read.csv(text=getURL(paste0(repo,"/data/transformed_data_Nature2008.csv")),
+               skip=7, na.string="")
+
+#head(tr)
+
+tr <- tr[,-14:-24] ## remove bad columns
+tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
+tr <- gather(tr, key="Species", value="Abundance", 2:13)
+levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods"
+levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
+names(tr)[2] <- "variable"
+names(tr)[1] <- "Day.number"
+tr$Day.number <- as.numeric(as.character(tr$Day.number))
+## get day range
+min.max.day <- group_by(tr, variable) %>%
+  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
+```
+
+Now we create a trimmed version of our data:
+
+```{r}
+trimmed_raw <- full_join(all.data, min.max.day)
+trimmed_raw <- filter(trimmed_raw, Day.number>min.day & Day.number<max.day)
+```
+
+## Interpolation (zeros trimmed)
 
@@
 ```{r}
-all.data1 <- na.omit(all.data)
-mt <- plyr::dlply(all.data1,
-                  "variable",
-                  function(xx) pracma::interp1(x=xx$Day.number,
-                                               y=xx$value,
-                                               xi=xout,
-                                               method="cubic"))
-## Aside: the duplicated date that was previously fixed was only discovered by a warning message
-## given by the pracma::interp1 function!
 
-mt <- as.data.frame(mt)
-mt <- cbind(Day.number=xout, mt)
-mt <- gather(mt, variable, value, 2:13)
+# make a function to do the interpolation
+# next two lines make variables for testing
+#x <- trimmed_raw %>% filter(variable=="Cyclopoids") %$% Day.number
+#y <- trimmed_raw %>% filter(variable=="Cyclopoids") %$% value
+Interpolate <- function(x, y) {
+  xout.inner <- xout[xout>=min(x) & xout<=max(x)]
+  res <- pracma::interp1(x=as.numeric(x),
+                  y=as.numeric(y),
+                  xi=xout.inner,
+                  method="cubic")
+  res.df <- data.frame(Day.number=xout.inner,
+                       value=res)
+}
+
+# run the function and tidy the data
+# (somewhat ashamed of this code...)
+
+# remove NAs before interpolation, so they don't cause problems
+trimmed_raw <- na.omit(trimmed_raw)
+res <- group_by(trimmed_raw, variable) %>%
+  do(le=Interpolate(.$Day.number, .$value))
+for(p in 1:nrow(res)){
+    temp <- data.frame(variable=res[p,1],
+                       res[p,2][[1]])
+    if(p==1)
+      trimmed_interp <- temp
+    if(p>1)
+      trimmed_interp <- rbind(trimmed_interp, temp)
+  }
+
+#mt <- as.data.frame(mt)
+#mt <- cbind(Day.number=xout, mt)
+#mt <- gather(mt, variable, value, 2:13)
 ```
 
-Check this against the data direct from Steve:
-
-```{r}
-
-from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
-
-from.steve <- gather(from.steve, Species, Abundance, 2:13)
-
-names(from.steve) <- c("Day.number", "variable", "value")
-
-g1 <- ggplot(mt, aes(x=as.numeric(Day.number), y=value)) +
-  facet_wrap(~variable, ncol=2, scales="free_y") +
-  geom_line(size=0.5, col="black") +
-  scale_colour_manual(values = species.colour.mapping)
-g2 <- geom_line(data=from.steve, aes(x=Day.number, y=value), col="red")
-g1 + g2
-
-```
-
-Looks very good.
-
-## Fourth root transform
+## Fourth root transform  (zeros trimmed)
 
@@
 ```{r}
-mt$fr.value <- mt$value^0.25
+trimmed_interp$fr.value <- trimmed_interp$value^0.25
 ```
 
-## Detrend
+## Detrend  (zeros trimmed)
 
@@
 ```{r}
-ww.td <- filter(mt, variable=="Total.dissolved.inorganic.nitrogen" |
+ww.td <- filter(trimmed_interp, variable=="Total.dissolved.inorganic.nitrogen" |
                   variable=="Soluble.reactive.phosphorus" |
@@
 ## and to not detrend
-ww.ntd <- filter(mt, variable!="Total.dissolved.inorganic.nitrogen" &
+ww.ntd <- filter(trimmed_interp, variable!="Total.dissolved.inorganic.nitrogen" &
                   variable!="Soluble.reactive.phosphorus" &
@@
 
-## Zero removal
-
-> The time series of cyclopoid copepods, protozoa, filamentous diatoms, harpacticoid copepods and ostracods contained long sequences of zero values.
-This does not imply that these species were absent from the food web during these periods, but that their concentrations were below the detection limit.
-Time series dominated by many zeros can bias the statistical analysis.
-Therefore, these time series were shortened to remove long sequences of zero values, before the data transformation.
-The transformed data of all species in the food web are shown in Figure S2.
-
-Since the method for this was not given, we do it by getting the trimmed day range from the fully transformed data in the ELE supplement.
-
-```{r, echo=FALSE}
-
-tr <- read.csv(text=getURL(paste0(repo,"/data/transformed_data_Nature2008.csv")),
-               skip=7, na.string="")
-
-#head(tr)
-
-tr <- tr[,-14:-24] ## remove bad columns
-tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
-tr <- gather(tr, key="Species", value="Abundance", 2:13)
-levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods"
-levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
-names(tr)[2] <- "variable"
-names(tr)[1] <- "Day.number"
-tr$Day.number <- as.numeric(as.character(tr$Day.number))
-## get day range
-min.max.day <- group_by(tr, variable) %>%
-  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
-
-```
-
-Now we create a trimmed version of our data:
-
-```{r}
-trimmed_detrended <- full_join(detrended, min.max.day)
-trimmed_detrended <- filter(trimmed_detrended, Day.number>min.day & Day.number<max.day)
-```
-
-## Rescale
+## Rescale (zeros trimmed)
 
@@
 ## standardise
-scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
+scaled_trimmed_final <- group_by(detrended, variable) %>%
   mutate(stand.y=as.numeric(scale(dt.value)))
@@
 ## or don't standardise
-final <- detrended
-final$y <- final$dt.value
-summarise(final, mean=mean(y), sd=sd(y))
+#final <- detrended
+#final$y <- final$dt.value
+#summarise(final, mean=mean(y), sd=sd(y))
 ```
@@
 ```{r}
-glimpse(final)
+#glimpse(scaled_trimmed_final)
 ```
 
-## Figure S1 (visualising the transformation)
-
-Choose a species to plot:
-```{r}
-soi <- "Bacteria"
-```
-
-Raw and interpolated data:
-
-```{r}
-g1 <- ggplot(filter(all.data, variable==soi), aes(x=Day.number, y=value)) +
-  facet_wrap(~variable, ncol=2, scales="free_y") +
-  geom_point(size=1, col="black") + geom_line(size=0.1) +
-  scale_colour_manual(values = species.colour.mapping) + ggtitle("Raw and interpolated data")
-g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=value), size=0.25, col="blue")
-g1 + g2
-```
-
-Fourth root transformed with trend:
-
-```{r}
-g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=fr.value)) +
-  facet_wrap(~variable, ncol=2, scales="free_y") +
-  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
-  scale_colour_manual(values = species.colour.mapping) + ggtitle("Quarter power trans. and trend")
-g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=trend), size=0.25, col="blue")
-g1 + g2
-```
-
-Detrended and normalised:
-
-```{r}
-g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=y)) +
-  facet_wrap(~variable, ncol=2, scales="free_y") +
-  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
-  scale_colour_manual(values = species.colour.mapping) + ggtitle("Detrended and normalised")
-g1
-```
-
-## Compare the data made above to published data
-
-Now take a look at the transformed data provided in the ELE Supplement, in the data sheet *transformed_data_Nature2008*.
-
-First note that the data in the ELE supplement include a data point at day 2658.1, whereas the data above finish at 2654.85.
-This is because the real data end at 2658, and the interpolation method above doesn't want to create data outside the range of the original data.
-
-The graph isn't produced here, as there is a mismatch in days abundances were interpolated to. Should fix this...
+## Compare the zeros removed transformed
 
@@
 
+tr$Day.number <- as.character(tr$Day.number)
+scaled_trimmed_final$Day.number <- as.character(scaled_trimmed_final$Day.number)
+
 ## Days are not matching at present, so the following doesn't work with the merge
 
-two.datasets <- full_join(tr, scaled_trimmed_final)
+two.datasets <- inner_join(tr, scaled_trimmed_final)
+two.datasets$Day.number <- as.numeric(two.datasets$Day.number)
+
 #head(two.datasets)
@@
 
-ggplot(filter(two.datasets), aes(x=as.numeric(Day.number), y=stand.y)) +
-  geom_path() +
-  facet_wrap(~variable) +
-  geom_line(aes(y=Abundance), col="red")
+# ggplot(filter(two.datasets), aes(x=as.numeric(Day.number), y=stand.y)) +
+#   geom_path() +
+#   facet_wrap(~variable) +
+#   geom_line(aes(y=Abundance), col="red")
+
+# ggplot(filter(two.datasets, variable=="Soluble.reactive.phosphorus"), aes(x=as.numeric(Day.number), y=stand.y)) +
+#   geom_path() +
+#   facet_wrap(~variable) +
+#   geom_line(aes(y=Abundance), col="red")
 
@@
 
+#unique(two.datasets$variable)
+
+two.datasets$variable <- plyr::revalue(two.datasets$variable,
+                                 c(Soluble.reactive.phosphorus="P",
+                                   Total.dissolved.inorganic.nitrogen="N"))
+
 ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point() +
-  facet_wrap(~variable)
+  facet_wrap(~variable) +
+  geom_abline(slope=1, intercept = 0, colour="red") +
+  xlab("Values from published data") +
+  ylab("Reproduced values")
 
@@
 
-Looks OK, but suggests / shows that that data in the ELE supplement were standardised after removal of the zeros.
+Looks pretty good, unclear why N and P are less well reproduced than others.
+
+```{r}
+final <- scaled_trimmed_final
+final$y <- final$stand.y
+```
 
@@
 ggplot(spec, aes(y=spec, x=1/freq, group=group)) + geom_line() + facet_wrap(~group, scales="free_y") + scale_x_log10() + 
-coord_cartesian(ylim=c(0,40), xlim=c(10,240))
+coord_cartesian(ylim=c(0,50), xlim=c(10,240))
 
@@
 #final_nozeros <- final[final$value!=0,]
-final_nozeros <- scaled_trimmed_final
+final_nozeros <- final
 
@@
 qplot(x=as.vector(original.cors), y=as.vector(cor.coefs), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
-      xlab="Correlations in original article", ylab="Correlations calculated in this reproduction") + geom_abline(intercept=0,slope=1)
+      xlab="Correlations in original article", ylab="Correlations calculated in this reproduction") + geom_abline(intercept=0,slope=1, colour="red")
 
@@
 LEs <- mutate(LEs, days=20, Difference=-0.5)
-g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species) +
-  geom_text(data=LEs, aes(x=60, y=-2.5, label=round(le,3)), group=NULL)
+
+diverg$Species_short <- plyr::revalue(diverg$Species,
+                                 c(Soluble.reactive.phosphorus="P",
+                                   Total.dissolved.inorganic.nitrogen="N"))
+LEs$Species_short <- plyr::revalue(LEs$Species,
+                                 c(Soluble.reactive.phosphorus="P",
+                                   Total.dissolved.inorganic.nitrogen="N"))
+
+g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species_short) +
+  geom_text(data=LEs, aes(x=60, y=0, label=round(le,3)))
 g1
@@
 
+Get the mean and standard deviation of the original and reproduced Lyapunov exponents
+
+```{r}
+mean(original.LEs$LE)
+sd(original.LEs$LE)
+
+mean(LEs$le)
+sd(LEs$le)
+
+```
+
 # Lyapunov exponents by indirect method
+
+It seems that this must have been done on data without zeros removed...
+
+## Interpolation (zeros retained)
+
+> First, the time series were interpolated using cubic hermite interpolation, to obtain data with equidistant time intervals of 3.35 days (Fig. S1a).
+
+Make a sequence of times at which to interpolate.
+
+```{r}
+#aggregate(Day.number ~ variable, all.data, min)
+#aggregate(Day.number ~ variable, all.data, max)
+#xout <- seq(343.35, 2657.2, by=3.35)
+xout <- seq(343.35, 2658, by=3.35)
+#range(xout)
+```
+
+Perform the interpolation:
+```{r}
+all.data <- na.omit(all.data)
+mt <- plyr::dlply(all.data,
+                  "variable",
+                  function(xx) pracma::interp1(x=xx$Day.number,
+                                               y=xx$value,
+                                               xi=xout,
+                                               method="cubic"))
+## Aside: the duplicated date that was previously fixed was only discovered by a warning message
+## given by the pracma::interp1 function!
+
+mt <- as.data.frame(mt)
+mt <- cbind(Day.number=xout, mt)
+mt <- gather(mt, variable, value, 2:13)
+```
+
+Check this against the data direct from Steve:
+
+```{r}
+
+from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
+
+from.steve <- gather(from.steve, Species, Abundance, 2:13)
+
+names(from.steve) <- c("Day.number", "variable", "value")
+from.steve$value_steve <- from.steve$value 
+from.steve <- select(from.steve, -value)
+
+## correcting an offset in the two datasets Day numbers
+from.steve$Day.number <- from.steve$Day.number + 1.65 + 3.35
+
+mt$Day.number <- as.character(mt$Day.number)
+from.steve$Day.number <- as.character(from.steve$Day.number)
+ff <- inner_join(mt, from.steve)
+mt$Day.number <- as.numeric(mt$Day.number)
+from.steve$Day.number <- as.numeric(from.steve$Day.number)
+ff$Day.number <- as.numeric(ff$Day.number)
+
+g1 <- ggplot(mt, aes(x=as.numeric(Day.number), y=sqrt(value))) +
+  facet_wrap(~variable, ncol=2, scales="free_y") +
+  geom_line(size=0.5, col="black") +
+  scale_colour_manual(values = species.colour.mapping)
+g2 <- geom_line(data=from.steve, aes(x=Day.number, y=sqrt(value_steve)), col="red")
+g1 + g2
+
+mt$Day.number <- as.character(mt$Day.number)
+from.steve$Day.number <- as.character(from.steve$Day.number)
+ff <- inner_join(mt, from.steve)
+mt$Day.number <- as.numeric(mt$Day.number)
+from.steve$Day.number <- as.numeric(from.steve$Day.number)
+ff$Day.number <- as.numeric(ff$Day.number)
+
+str(ff)
+
+ff$difference <- ff$value - ff$value_steve
+qplot(log10(ff$difference))
+
+qplot(data=ff, x=log10(value_steve), y=log10(value)) +
+  xlab("Values from Ellner (log 10 transformed)") +
+  ylab("Reproduced values (log 10 transformed)") +
+  geom_abline(intercept=0, slope=1, colour="red")
+
+```
+
+Looks very good.
+
+## Fourth root transform (zeros retained)
+
+> Next, because the original time series showed many sharp spikes, the time series were rescaled using a fourth-root power transformation (Fig.
+S1b).
+The sharp spikes bias "direct method" estimates of the Lyapunov exponent, because nearby pairs of reconstructed state vectors mostly occurred in the troughs between spikes.
+The average rate of subsequent trajectory divergence from these pairs is therefore an estimate of the local Lyapunov exponent in the troughs, which may be very different from the global Lyapunov exponent.
+By making spikes and troughs more nearly symmetric, the power transformation resulted in a much more even spread of nearby state vector pairs across the full range of the data for all functional groups in the food web.
+The transformation is also useful for fitting nonlinear models of the deterministic skeleton (used for nonlinear predictability and indirect method estimates of the Lyapunov exponent), which was done by least squares and therefore is most efficient when error variances are stabilized.
+Fourth-root transformation is intermediate between the square-root transformation that would approximately stabilize the measurement error variance in count data from random subsamples, and the log transformation that is usually recommended for stabilizing process noise variance due to stochastic variation in birth and death rates.
+
+```{r}
+mt$fr.value <- mt$value^0.25
+```
+
+## Detrend (zeros retained)
+
+> The time series were then detrended using a Gaussian kernel with a bandwidth of 300 days (red line in Fig.
+S1b), to obtain stationary time series.
+Most species did not show long-term trends, except for the bacteria, detritivores (ostracods and harpacticoid copepods), dissolved inorganic nitrogen and soluble reactive phosphorus.
+One possible explanation for these trends in the microbial loop could be the slow accumulation of refractory organic material in the mesocosm, but we have not measured this component.
+
+```{r}
+ww.td <- filter(mt, variable=="Total.dissolved.inorganic.nitrogen" |
+                  variable=="Soluble.reactive.phosphorus" |
+                  variable=="Bacteria" |
+                  variable=="Ostracods" |
+                  variable=="Harpacticoids")
+## and to not detrend
+ww.ntd <- filter(mt, variable!="Total.dissolved.inorganic.nitrogen" &
+                  variable!="Soluble.reactive.phosphorus" &
+                  variable!="Bacteria" &
+                  variable!="Ostracods" &
+                  variable!="Harpacticoids")
+## detrend:
+ww1 <- group_by(ww.td, variable) %>%
+  mutate(trend=ksmooth(Day.number,fr.value,bandwidth=300,kernel="normal")$y)
+ww1$dt.value <- ww1$fr.value-ww1$trend
+#ww1 <- select(ww1, trend)
+
+## don't detrend
+ww2 <- ww.ntd
+ww2$trend <- 0
+
+ww2$dt.value <- ww2$fr.value
+
+## rejoin
+detrended <- rbind(ww1, ww2)
+
+```
+
+## Rescale (zeros retained)
+
+> Finally, the time series were linearly rescaled to have zero mean and a standard deviation of 1 (Fig. S1c).
+
+(Note that this standardisation is not done in the code sent by Stephen. Probably shouldn't make a difference in the GAMs?)
+
+```{r}
+## standardise
+scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
+  mutate(stand.y=as.numeric(scale(dt.value)))
+summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
+
+## or don't standardise
+final <- detrended
+final$y <- final$dt.value
+summarise(final, mean=mean(y), sd=sd(y))
+```
+
+```{r}
+glimpse(final)
+```
+
+## Figure S1 (visualising the transformation)
+
+Choose a species to plot:
+```{r}
+soi <- "Bacteria"
+```
+
+Raw and interpolated data:
+
+```{r}
+g1 <- ggplot(filter(all.data, variable==soi), aes(x=Day.number, y=value)) +
+  facet_wrap(~variable, ncol=2, scales="free_y") +
+  geom_point(size=1, col="black") + geom_line(size=0.1) +
+  scale_colour_manual(values = species.colour.mapping) + ggtitle("Raw and interpolated data")
+g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=value), size=0.25, col="blue")
+g1 + g2
+```
+
+Fourth root transformed with trend:
+
+```{r}
+g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=fr.value)) +
+  facet_wrap(~variable, ncol=2, scales="free_y") +
+  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
+  scale_colour_manual(values = species.colour.mapping) + ggtitle("Quarter power trans. and trend")
+g2 <- geom_line(data=filter(final, variable==soi), aes(x=Day.number, y=trend), size=0.25, col="blue")
+g1 + g2
+```
+
+Detrended and normalised:
+
+```{r}
+g1 <- ggplot(filter(final, variable==soi), aes(x=Day.number, y=y)) +
+  facet_wrap(~variable, ncol=2, scales="free_y") +
+  geom_point(size=0.5, col="black") + geom_line(size=0.1) +
+  scale_colour_manual(values = species.colour.mapping) + ggtitle("Detrended and normalised")
+g1
+```
 
@@
 
-Load and run the functions, or read in from data file (default option, for which you will need to change the path to the data file.)
+Load and run the functions, or read in from data file.
 

### 2016-02-12 code/report.Rmd
markdown report development

@@
 
-*Update* The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.
+The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.
 
@@
 All required libraries:
-
 ```{r, message=FALSE}
@@
 library(magrittr)
-
+```
+
+Read in the data and remove some blank rows and columns.
+```{r}
 spp.abund <- read.csv(text=getURL(paste0(repo,"/data/species_abundances_original.csv")), skip=7, header=T)
@@
 The Protozoa variable contains some numbers with comman as the decimal separator. This creates a question about what dataset was used for the original analyses, as it could not have been this one.
-
 ```{r}
@@
 Format the dates as dates.
-
 ```{r}
@@
 Original dataset contains a duplicated date: 28/10/1996 (row 709 and 710 in excel sheet). Lets change the date in row 709 to 26/10/1996, which will put it half way between the two surrounding dates:
-
 ```{r}
@@
 
-Now select only the date range used in the Nature paper. From the supplment *The analysis in Benincà et al. (Nature 2008) covered all data from 16/06/1991 until 20/10/1997*. 
+Now select only the date range used in the Nature paper. From the supplement *The analysis in Benincà et al. (Nature 2008) covered all data from 16/06/1991 until 20/10/1997*. 
 ```{r}
 all.data <- filter(all.data, Date>=dmy("15/06/1991") & Date<=dmy("21/10/1997"))
-
-#all.data[all.data$Date==dmy("16/06/2091"),]
-#all.data[all.data$Date==dmy("20/10/2097"),]
-
 ```
@@
                         "Detritivore", "Detritivore", "Bacteria", "Nutrient", "Nutrient"))
-#tt
 ```
@@
 
-The graph with abundances fourth root transformed, as this is the transformation used in the ms.
-
+The graph with abundances fourth root transformed, as this is the transformation used in analyses in the original publication.
 ```{r, warning=FALSE, fig.width=10}
@@
 
-The ELE supplement contains the raw data and the transformed data, in separate data sheets. I (Owen) also got the interpolated data from Stephen Ellner directly.
+The ELE supplement contains the raw data and the transformed data, in separate data sheets.
 
@@
 However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
-
 ```{r}
@@
                skip=7, na.string="")
-
-#head(tr)
-
 tr <- tr[,-14:-24] ## remove bad columns
 tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
-tr <- gather(tr, key="Species", value="Abundance", 2:13)
-levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods"
+tr <- gather(tr, key="Species", value="Abundance", 2:13) ## make long format
+levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
 levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
-names(tr)[2] <- "variable"
+names(tr)[2] <- "variable" # rename a couple of variables...
 names(tr)[1] <- "Day.number"
-tr$Day.number <- as.numeric(as.character(tr$Day.number))
-## get day range
+tr$Day.number <- as.numeric(as.character(tr$Day.number)) # make Day.number numeric
+## get day range, in order to use this to constrain later data
 min.max.day <- group_by(tr, variable) %>%
@@
 
-Now we create a trimmed version of our data:
-
+Now we create a trimmed version of the raw data:
 ```{r}
@@
 Make a sequence of times at which to interpolate. 
-
 ```{r}
@@
 ```{r}
-
 # make a function to do the interpolation
-# next two lines make variables for testing
-#x <- trimmed_raw %>% filter(variable=="Cyclopoids") %$% Day.number
-#y <- trimmed_raw %>% filter(variable=="Cyclopoids") %$% value
 Interpolate <- function(x, y) {
@@
 # (somewhat ashamed of this code...)
-
 # remove NAs before interpolation, so they don't cause problems
@@
   do(le=Interpolate(.$Day.number, .$value))
-for(p in 1:nrow(res)){
-    temp <- data.frame(variable=res[p,1],
-                       res[p,2][[1]])
-    if(p==1)
-      trimmed_interp <- temp
-    if(p>1)
-      trimmed_interp <- rbind(trimmed_interp, temp)
-  }
-
-#mt <- as.data.frame(mt)
-#mt <- cbind(Day.number=xout, mt)
-#mt <- gather(mt, variable, value, 2:13)
+
+trimmed_interp <- res %>% do(data.frame(Day.number = .$le[[1]], value = .$le[[2]], variable = .[[1]]))
+
+# for(p in 1:nrow(res)){
+#     temp <- data.frame(variable=res[p,1],
+#                        res[p,2][[1]])
+#     if(p==1)
+#       trimmed_interp <- temp
+#     if(p>1)
+#       trimmed_interp <- rbind(trimmed_interp, temp)
+#   }
 ```
@@
 ```{r}
+## data to detrend
 ww.td <- filter(trimmed_interp, variable=="Total.dissolved.inorganic.nitrogen" |
@@
 ww1$dt.value <- ww1$fr.value-ww1$trend
-#ww1 <- select(ww1, trend)
 
@@
 ww2$trend <- 0
-
 ww2$dt.value <- ww2$fr.value
@@
 detrended <- rbind(ww1, ww2)
-
 ```
@@
 > Finally, the time series were linearly rescaled to have zero mean and a standard deviation of 1 (Fig. S1c).
-
-(Note that this standardisation is not done in the code sent by Stephen. Probably shouldn't make a difference in the GAMs?)
 
@@
   mutate(stand.y=as.numeric(scale(dt.value)))
-summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
-
-## or don't standardise
-#final <- detrended
-#final$y <- final$dt.value
-#summarise(final, mean=mean(y), sd=sd(y))
-```
-
-```{r}
-#glimpse(scaled_trimmed_final)
+#summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
 ```
@@
 
+## change the Day variable to character, to help with the merge, and back after
 tr$Day.number <- as.character(tr$Day.number)
 scaled_trimmed_final$Day.number <- as.character(scaled_trimmed_final$Day.number)
-
-## Days are not matching at present, so the following doesn't work with the merge
-
 two.datasets <- inner_join(tr, scaled_trimmed_final)
 two.datasets$Day.number <- as.numeric(two.datasets$Day.number)
-
-#head(two.datasets)
-#match(tr$Day.number, final$Day.number)
-
-#ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point()
 
@@
 
-#ggplot(two.datasets, aes(x=as.numeric(Day.number), y=stand.y)) + geom_line() +
+# ggplot(two.datasets, aes(x=as.numeric(Day.number), y=stand.y)) + geom_line() +
 #   facet_wrap(~variable) +
@@
 
-#unique(two.datasets$variable)
-
+## shorten the names of the nutrients
 two.datasets$variable <- plyr::revalue(two.datasets$variable,
@@
 
+## graph
 ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point() +
@@
 ```{r}
-
-# Raw spectrum
-spectra <- final %>% group_by(variable) %>% do(spectra = spectrum(ts(data=.$y, end=2650.15, deltat=3.35), log='no', method="pgram", detrend=F, plot=F))
-spec <- spectra %>% do(data.frame(spec = .$spec[[2]], freq = .$spec[[1]], group = .[[1]]))
-
-ggplot(spec, aes(y=spec, x=1/freq, group=group)) + geom_line() + facet_wrap(~group, scales="free_y") + scale_x_log10() + 
+# Calculate and plot raw spectra
+spectra <- final %>% group_by(variable) %>%
+  do(spectra = spectrum(ts(data=.$y, end=2650.15, deltat=3.35), log='no', method="pgram", detrend=F, plot=F))
+spec <- spectra %>%
+  do(data.frame(spec = .$spec[[2]], freq = .$spec[[1]], group = .[[1]]))
+ggplot(spec, aes(y=spec, x=1/freq, group=group)) +
+  geom_line() +
+  facet_wrap(~group, scales="free_y") +
+  scale_x_log10() + 
   coord_cartesian(ylim=c(0,50), xlim=c(10,240))
 
-freq.est <- spec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
-freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
-freq.est$freq <- 1/freq.est$freq
-#freq.est
-
-# Welch's periodogram
-
-wspectra <- final %>% group_by(variable) %>% do(spectra = pwelch(ts(data=.$y, end=2650.15, deltat=3.35), window=5, method="pgram", plot=F))
+# freq.est <- spec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
+# freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
+# freq.est$freq <- 1/freq.est$freq
+
+# Calculate and plot Welch's periodogram
+wspectra <- final %>% group_by(variable) %>%
+  do(spectra = pwelch(ts(data=.$y, end=2650.15, deltat=3.35), window=5, method="pgram", plot=F))
 wspec <- wspectra %>% do(data.frame(spec = .$spec[[2]], freq = .$spec[[1]], group = .[[1]]))
-
-ggplot(wspec, aes(y=spec, x=1/freq, group=group)) + geom_line() + facet_wrap(~group, scales="free_y") + scale_x_log10() +
-coord_cartesian(ylim=c(0.01,100), xlim=c(10,240)) + scale_y_log10()
-
-freq.est <- wspec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
-freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
-freq.est$freq <- 1/freq.est$freq
-#frequency(final$y)
-ts <- as.ts(final$y, frequency = 0.3)
-#time(ts)
+ggplot(wspec, aes(y=spec, x=1/freq, group=group)) +
+  geom_line() +
+  facet_wrap(~group, scales="free_y") +
+  scale_x_log10() +
+  coord_cartesian(ylim=c(0.01,100), xlim=c(10,240)) +
+  scale_y_log10()
+
+# freq.est <- wspec %>%
+#   group_by(group) %>%
+#   mutate(max_spec = max(spec), freq = freq)
+# freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
+# freq.est$freq <- 1/freq.est$freq
+# #frequency(final$y)
+# ts <- as.ts(final$y, frequency = 0.3)
+# #time(ts)
 
@@
 ```{r}
-
-#final_nozeros <- final[final$value!=0,]
+# some data formatting
 final_nozeros <- final
-
-#
-#final_nozeros <- final_nozeros[final_nozeros$variable!="Cyclopoids" & final_nozeros$variable!="Filamentous.diatoms",c("Day.number","variable","stand.y")]
-
 final_wide <- spread(as.data.frame(final_nozeros[,c("Day.number","variable","stand.y")]), variable, stand.y)
 final_wide <- select(final_wide, - Day.number)
-# we also removed data on Cyclopoids, not used in the table
+
 ```
@@
 
-Add significance "stars" to cor.coefs from cor.pvals
+Add significance "stars" to cor.coefs from cor.pvals:
 ```{r}
@@
 
+Make the table entries, by pasting the values and stars:
 ```{r}
@@
 
-```{r}
-
+Cosmetic work on the table:
+```{r}
 oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)
-
 cor.coefs <- cor.coefs[oo,oo]
 cor.cp <- cor.cp[oo,oo]
-
 for(i in 1:length(cor.cp[1,]))
@@
   cor.cp[i,j] <- ""
-
 #sn <- c("Bacteria", "Harps", "Ostr", "N", "P", "Picophyt", "Nanophyt", "Rotifers", "Protozoa", "Calanoids")
@@
 dimnames(cor.cp) <- list(sn, sn)
-
 ```
@@
 2008)")
-
 table1b
-
 ```
@@
 ```{r}
-
+# read in the correlations given in the original publication
 original.cors <- read.csv(text=getURL(paste0(repo,"/data/table1_original_article.csv")), skip=0, header=T, row.names = 1)
-
 original.cors <- as.matrix(original.cors)
@@
       xlab="Correlations in original article", ylab="Correlations calculated in this reproduction") + geom_abline(intercept=0,slope=1, colour="red")
-
-max(as.vector(original.cors) - as.vector(cor.coefs), na.rm=T)
 ```
@@
 
-Unclear if this was performed on untransformed or transformed data. First try with the transformed data.
-Time delay (1), embedding dimension (6), and Theiler window (50) were used in the Nature report. Other parameters are chosen rather randomly, though don't seem to matter too much.
+Time delay (1), embedding dimension (6), and Theiler window (50) were used in the Nature report. Other parameters are chosen rather randomly!
 
@@
 
-Get the data for the graphs:
+Do the time-delayed embedding, and save the data for the graphs:
 ```{r, message=FALSE, error=TRUE}
@@
 LEs <- mutate(LEs, days=20, Difference=-0.5)
-
 diverg$Species_short <- plyr::revalue(diverg$Species,
@@
                                    Total.dissolved.inorganic.nitrogen="N"))
-
 g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species_short) +
@@
 ```{r}
+## read in LE values given in original publication
 original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original_direct_LEs.csv")), skip=0, header=T)
-LEs
-
-both_LEs <- full_join(original.LEs, LEs)
-
+both_LEs <- full_join(original.LEs, LEs) # merge the reproduced and original values
 ggplot(both_LEs, aes(x=LE, y=le)) + geom_point() + xlim(0.02, 0.09) + ylim(0.02, 0.09) + geom_abline(intercept=0, slope=1) +
   xlab("Lyapunov exponent from original article") + ylab("Reproduced Lyapunov exponent")
-
 ```
@@
 Get the mean and standard deviation of the original and reproduced Lyapunov exponents
-
 ```{r}
@@
 sd(original.LEs$LE)
-
 mean(LEs$le)
 sd(LEs$le)
-
 ```
@@
 
-It seems that this must have been done on data without zeros removed...
+It seems that this must have been done on data without zeros removed, because removal of the zeros would leave too many missing values in the GAMs.
+So we do the transformation steps again, but without removal of zeros.
 
@@
 ## given by the pracma::interp1 function!
-
 mt <- as.data.frame(mt)
@@
 Check this against the data direct from Steve:
-
-```{r}
-
+```{r}
 from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
-
 from.steve <- gather(from.steve, Species, Abundance, 2:13)
-
 names(from.steve) <- c("Day.number", "variable", "value")
@@
 
+## joint the data, again first making days character, then back to numeric after
 mt$Day.number <- as.character(mt$Day.number)
@@
 
+## and plot time series
 g1 <- ggplot(mt, aes(x=as.numeric(Day.number), y=sqrt(value))) +
@@
 
+## same but to allow x-y scatter
 mt$Day.number <- as.character(mt$Day.number)
@@
 ff$Day.number <- as.numeric(ff$Day.number)
-
-str(ff)
-
-ff$difference <- ff$value - ff$value_steve
-qplot(log10(ff$difference))
-
+#str(ff)
+#ff$difference <- ff$value - ff$value_steve
+#qplot(log10(ff$difference))
 qplot(data=ff, x=log10(value_steve), y=log10(value)) +
@@
 
-Looks very good.
+Looks good.
 
@@
 ## standardise
-scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
-  mutate(stand.y=as.numeric(scale(dt.value)))
-summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
+#scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
+#  mutate(stand.y=as.numeric(scale(dt.value)))
+#summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
 
@@
 final$y <- final$dt.value
-summarise(final, mean=mean(y), sd=sd(y))
+#summarise(final, mean=mean(y), sd=sd(y))
 ```
@@
 
-The function needs a matrix, X, with species abundances in wide format. Be careful to work on the unstandardised data (or standardised, if you wish) (comment out appropriate lines here.
+The function needs a matrix, X, with species abundances in wide format. Be careful to work on the unstandardised data (or standardised, if you wish) (comment out appropriate lines here).
 
@@
 #final.to.melt <- final[, c("variable", "y", "Day.number")]
-
 names(final.to.melt)[1] <- "Species"
@@
 ## don't run this as it takes a while
-#LE <- Get_GLE_Beninca(X)
-#save(LE, file="~/Desktop/GLE_estimate.Rdata")
+LE <- Get_GLE_Beninca(X)
+save(LE, file="~/Desktop/GLE_estimate.Rdata")
 
 ## load the already saved data from github (this can take some time depending on the internet connection)
-source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))
+#source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))
 
@@
 
-Need to predict until about 40 days ahead = 40 / 3.35 time steps = 12 time steps.
-
-Do this from each time in the time series as initial abundances.
-
-From each time in the series, predict 12 steps ahead. Put results in an array:
-time x species, start.location
-time will be 12
-species will be 12
-start location will be length of time series - 12
-
-```{r}
+Need to predict until about 40 days ahead = 40 / 3.35 time steps = 12 time steps. Do this from each time in the time series as initial abundances.
+
+From each time in the series, predict 12 steps ahead. Put results in an array: time x species x start.location.
+time will be 12, species will be 12, start location will be length of time series - 12
+
+```{r}
+## some preliminaries
 x <- X 
-
 Z <- LE[[3]]
 Z1 <- Z[,13:24]
-
 all.species <- names(all.gams)
-
 time.to.predict <- 12
-
 nn <- length(Z1[,1])-time.to.predict
-
+## create the array of predictions
 preds <- array(NA, c(12, 12, nn))
-
 dimnames(preds) <- list(dist=1:12,
@@
                         start.time=1:nn)
+## fill in the array with the one step ahead predictions
 for(i in 1:length(all.species))
@@
 
-#for(i in 1:length(all.species))
-#  print(length( predict(all.gams[[i]])))
-
-```
-
-<!--
-Look at the correlation between observed and predicted abundances:
-
-```{r}
-
-layout(matrix(1:12, 4, 3))
-for(soi in all.species){
-  xxx <- Z1[1:nn,soi]
-  yyy <- preds[1, soi, ]
-  plot(xxx, yyy,
-       xlab="Observed abundance",
-       ylab="Predicted abundance",
-       main=paste(soi, "rsq =", round(cor(xxx,yyy)^2,2)))
-}
-
-```
--->
-
-Now for the predictions at t+2 from predictions at t+1
-
-```{r}
+# Now for the predictions at t+2 from predictions at t+1
 pred.time <- 2
@@
 }}
-
 cors <- data.frame(dist=as.numeric(I(rownames(cors))), cors)
 cors.long <- gather(as.data.frame(cors), key=dist)
-
 names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
@@
 ```{r}
-
 cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
@@
                              "Protozoa"))
-# ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) + xlab("Prediction time (days)") +
-#   geom_point(col="red", size=3) +
-#   facet_wrap(~Variable, ncol=3 )
 

### 2016-02-12 article/article.tex
new article build

@@
 contains several datasheets. Two were particularly important, as they
-are the source of the raw data (one contains original species
-abundances, the one with the nutrient concentrations). We saved these
-two datasheets as comma separated value (csv) text files. In the code
-associated with this reproduction, these data files are read from the
-associated github repository.
+were the source of the raw data (one contains original species
+abundances, the other the nutrient concentrations). We saved these two
+datasheets, without any alteration, as comma separated value (csv) text
+files. In the code associated with this reproduction, these data files
+are read from the associated github repository.
 
-Another datasheet in the Ecology Letters supplement contains transformed
-variables (we also saved this as csv file, in order to use it in this
-reproduction). We also received a dataset direct from Steve Ellner, see
-below for details.
+Another datasheet (we also saved this as csv file, in order to use it in
+this reproduction). We also received a dataset direct from Steve Ellner,
+see below for details.
 
-The original species abundance data contained errors (e.g., a few
-numerica values had a comma in place of a period as the decimal
-separator) that suggested that this was not the exact version of the
-dataset used in the original article, or that this was the exact
-dataset, but with errors corrected.
+The original species abundance data contained errors (i.e., a few
+numeric values had a comma in place of a period as the decimal
+separator, and a duplicated date).
+
+The original abundance dataset contains a duplicated date: 28/10/1996
+(row 709 and 710 in excel sheet). We changed this (in R) to 26/10/1996,
+making the observation occur half way between the two surrounding dates.
+The original Protozoa variable in the original dataset contains some
+numbers with a comma (instead of period) as the decimal separator; these
+were changed to periods.
 
@@
 
-The code for this reproduction resides in an R markdown document, as
-well as a source file containing some required functions. Some of the
-code takes several minutes to run, so an intermediate data file is
-provided with results from this code.
+The code for making the analyses and figures presented in this
+reproduction resides in an R markdown document, as well as a source file
+containing some required functions. Some of the code takes several
+minutes to run, so an intermediate data file is provided with results
+from this code.
 
@@
 
-The reproduced populations dynamics were at least very similar to those
-in figure 1b-g of the original publication (figure \ref{fig:dynamics}).
-Note that we plot fourth root transformed values, rather than raw
-abundances with a y-axis break, as in the original article.
+The reproduced populations dynamics were very similar to those in figure
+1b-g of the original publication (figure \ref{fig:dynamics}). As most
+analyses were performed on fourth root transformed values, we plot
+these, rather than raw abundances with a y-axis break as in the original
+article.
 
@@
 
-\subsection{Data transformation}\label{data-transformation}
+\subsection{Data transformations}\label{data-transformations}
 
-The following transformation steps were used:
+The original publication states that long sequences of zeros were
+removed from time series prior to transformation and further analysis.
+We removed zeros by matching the removals to those in the transformed
+version of the data in the published Excel file mentioned above.
+Transformation steps were:
 
@@
 \tightlist
-\item
-  Time series shortened to remove long sequences of zeros.
 \item
@@
 
-The method for selecting the zeros to remove was unclear. In order to
-reproduce this step, we removed the same data as in the original study,
-by matching to the transformed data with zeros removed in the Ecology
-Letters supplement Excel file mentioned above. All remaining
-transformation steps were performed independently of this dataset. Our
-reproduction of the transformed data closely matched the published
-transformed data.
+The reproduced transformed data closely matched the published
+transformed data (figure \ref{fig:trans_comp1}).
 
-The data received directly from Stephen Ellner was interpolated, but
-without zeros removed. Our interpolated data, without zeros removed,
-matched closely this data.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/zero_rem_trans_comparison.pdf}
+\caption{\label{fig:trans_comp1}Comparison of published transformed data
+with reproduced transformed data. The red line is the 1:1 line.}
+\end{figure}
 
@@
 derived from figures S3 and S4 of the original article. Our reproduced
-spectra (not shown here, but code provided) were not quantitatively
-identical to the spectra in the original article.
+spectra (not shown here, but code provided) were a reasonably close
+match to the spectra in the original article.
 
@@
 per day, s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very
-similar mean value, but had about four times greater standard deviation
-(mean = 0.055 and s.d. = 0.019).
+similar mean value, but had much larger standard deviation (mean = 0.057
+and s.d. = 0.014).
 
@@
 \includegraphics{figures/LE_comparison.pdf}
-\caption{\label{fig:LE_comparison}Comparison of Lyapunov exponents,
-estimated by direct method, in the original article and this
-reproduction.}
+\caption{\label{fig:LE_comparison}Comparison of original and reproduced
+directly estimated Lyapunov exponents.}
 \end{figure}
@@
 method}\label{lyapunov-exponents-by-indirect-method}
+
+The data received directly from Stephen Ellner (file
+`interp\_short\_allsystem\_newnames.csv' in the reproduction repository)
+was interpolated, but without zeros removed. The reproduced interpolated
+data, without zeros removed, matched closely this data (figure
+\ref{fig:trans_comp2}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures/repro_ellner_comp.pdf}
+\caption{\label{fig:trans_comp2}Comparison of interpolated data provided
+by Ellner with reproduced interpolated data (no removal of zeros). The
+red line is the 1:1 line.}
+\end{figure}
 
@@
 
-Although we were not able to make a quantitatively accurate reproduction
-of all results of the original article, the qualitative results were
-largely identical. For example, all Lyapunov exponents estimated by
-direct method are positive, as in the original article, consistent with
+The reproduced results were qualitatively identical to those in the
+original article. For example, all Lyapunov exponents estimated by
+direct method were positive, as in the original article, consistent with
 chaotic dynamics. Quantitative differences may have resulted from
@@
 required by the functions we used were reported in the original ms.
-There may also have been some difference in data used for specific
-analyses, e.g., data with zeros removed or not, as it was not always
-possible to be totally sure the reproduction used exactly the same data
-as the original article.
+
+Considerable quantitative difference in predictability decay
+(\ref{fig:prediction_distance}) remain unexplained.
 
@@
 \href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
-on github.
+on github. Many thanks to Elisa Benincà and Steve Ellner for assistance,
+and to Steve for providing unpublished data and code.
 

### 2016-02-12 code/report.Rmd
update report

@@
 qplot(x=as.vector(original.cors), y=as.vector(cor.coefs), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
-      xlab="Correlations in original article", ylab="Correlations calculated in this reproduction") + geom_abline(intercept=0,slope=1, colour="red")
+      xlab="Correlations in original article", ylab="Correlations calculated\nin this reproduction") + geom_abline(intercept=0,slope=1, colour="red")
 ```
@@
 both_LEs <- full_join(original.LEs, LEs) # merge the reproduced and original values
-ggplot(both_LEs, aes(x=LE, y=le)) + geom_point() + xlim(0.02, 0.09) + ylim(0.02, 0.09) + geom_abline(intercept=0, slope=1) +
-  xlab("Lyapunov exponent from original article") + ylab("Reproduced Lyapunov exponent")
+ggplot(both_LEs, aes(x=LE, y=le)) +
+  geom_point() +
+  xlim(0.02, 0.09) +
+  ylim(0.02, 0.09) +
+  geom_abline(intercept=0, slope=1, colour="red") +
+  xlab("Lyapunov exponent from original article") + 
+  ylab("Reproduced Lyapunov exponent")
 ```
@@
   scale_colour_manual(values = species.colour.mapping)
-g2 <- geom_line(data=from.steve, aes(x=Day.number, y=sqrt(value_steve)), col="red")
+g2 <- geom_line(data=from.steve, aes(x=Day.number, y=sqrt(value_steve)), colour="red")
 g1 + g2
@@
 qplot(data=ff, x=log10(value_steve), y=log10(value)) +
-  xlab("Values from Ellner (log 10 transformed)") +
-  ylab("Reproduced values (log 10 transformed)") +
+  xlab("Values from Ellner\n(log 10 transformed)") +
+  ylab("Reproduced values\n(log 10 transformed)") +
   geom_abline(intercept=0, slope=1, colour="red")

### 2016-02-12 code/report.Rmd
update rmarkdown report
small fixes

@@
 ```{r}
-diverg$Difference[is.na(diverg$Difference)] <- 0
-diverg$Difference[is.infinite(diverg$Difference)] <- 0
+#diverg$Difference[is.na(diverg$Difference)] <- 0
+diverg$Difference[is.infinite(diverg$Difference)] <- NA
 diverg.short <- filter(diverg, days<24) ## 24 is about 6 steps, after initial gap
@@
 
-## don't run this as it takes a while
+## don't run this unless you have some time (it takes a while)
 LE <- Get_GLE_Beninca(X)

### 2016-02-12 code/report.Rmd
comment LE indirect method

@@
 ## don't run this unless you have some time (it takes a while)
-LE <- Get_GLE_Beninca(X)
-save(LE, file="~/Desktop/GLE_estimate.Rdata")
+#LE <- Get_GLE_Beninca(X)
+#save(LE, file="~/Desktop/GLE_estimate.Rdata")
 
 ## load the already saved data from github (this can take some time depending on the internet connection)
-#source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))
+source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))
 

### 2016-02-12 article/article.tex
latest build of article pdf

@@
 
+\section{Draft article}\label{draft-article}
+
 \section{Introduction}\label{introduction}
@@
 per day, s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very
-similar mean value, but had much larger standard deviation (mean = 0.057
+similar mean value, but had much larger standard deviation (mean = 0.060
 and s.d. = 0.014).
@@
 Considerable quantitative difference in predictability decay
-(\ref{fig:prediction_distance}) remain unexplained.
+(\ref{fig:prediction_distance}) remain unexplained, though could result
+from original analyses using neural networks, and the reproduction using
+GAMs.
 

### 2016-02-12 article/article.tex
latest build of pdf

@@
 files. In the code associated with this reproduction, these data files
-are read from the associated github repository.
-
-Another datasheet (we also saved this as csv file, in order to use it in
-this reproduction). We also received a dataset direct from Steve Ellner,
-see below for details.
-
-The original species abundance data contained errors (i.e., a few
-numeric values had a comma in place of a period as the decimal
-separator, and a duplicated date).
+are read from the associated github repository. Another datasheet
+contained transformed data (we also saved this as csv file, in order to
+use it in this reproduction). We also received a dataset direct from
+Steve Ellner, see below for details.
 
@@
 was used to make the reproduction. Additional R packages required are
-specified in the code associated with this reproduction.
-
-The code for making the analyses and figures presented in this
-reproduction resides in an R markdown document, as well as a source file
-containing some required functions. Some of the code takes several
-minutes to run, so an intermediate data file is provided with results
-from this code.
+specified in the code associated with this reproduction. The code for
+making the analyses and figures presented in this reproduction resides
+in an R markdown document, as well as a source file containing some
+required functions. Some of the code takes several minutes to run, so an
+intermediate data file is provided with results from this code.
 
@@
 correlation between bacteria and ostracods, and positive correlation
-between bacteria and phosphorus. All of these correlations were
+between bacteria and phosphorus. All of these correlations were at least
 qualitatively reproduced.
@@
 Spectral analyses in the original paper were presented graphically in
-figures S3 (raw spectrograms) and S4 (Welch periodograms). These graphs
-were, apparently, visually inspected to reveal: ``fluctuations covered a
-range of different periodicities'', and ``picophytoplankton, rotifers
-and calanoid copepods seemed to fluctuate predominantly with a
-periodicity of about 30 days.'' It is unclear how these conclusions were
-derived from figures S3 and S4 of the original article. Our reproduced
-spectra (not shown here, but code provided) were a reasonably close
-match to the spectra in the original article.
+figures S3 (raw spectrograms) and S4 (Welch periodograms). The original
+article states: ``fluctuations covered a range of different
+periodicities'', and ``picophytoplankton, rotifers and calanoid copepods
+seemed to fluctuate predominantly with a periodicity of about 30 days.''
+Our reproduced spectra (not shown here, but code provided) were a
+reasonably close match to the spectra in the original article.
 
@@
 
-Reproduced divergence rates (figure \ref{fig:divergence}) and Lyapunov
-exponents (figure \ref{fig:LE_comparison}) were somewhat different from
-those in the original article.. The original article states: ``the
-distance between initially nearby trajectories increased over time, and
-reached a plateau after about 20--30 days''. The reproduced results
-appear not inconsistent with this statement, except for one group of
-species (Harpacticoids). The original article also stated that the
-analyses ``yielded significantly positive Lyapunov exponents of
-strikingly similar value for all species (Fig. 3; mean exponent = 0.057
-per day, s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very
-similar mean value, but had much larger standard deviation (mean = 0.060
-and s.d. = 0.014).
+Reproduced divergence rates (figure \ref{fig:divergence}) and therefore
+Lyapunov exponents (figure \ref{fig:LE_comparison}) were somewhat
+different from those in the original article. The original article
+states: ``the distance between initially nearby trajectories increased
+over time, and reached a plateau after about 20--30 days''. The
+reproduced results are consistent with this statement. The original
+article also stated that the analyses ``yielded significantly positive
+Lyapunov exponents of strikingly similar value for all species (Fig. 3;
+mean exponent = 0.057 per day, s.d. = 0.005 per day, n = 9)''.
+Reproduced exponents had very similar mean value, but had larger
+standard deviation (mean = 0.060 and s.d. = 0.014).
 
@@
 
-The original paper reported global Lyapunov exponent calculated via two
-modelling approaches (neural network and generalised additive models
+The original paper reported a global Lyapunov exponent calculated via
+two modelling approaches (neural network and generalised additive models
 {[}GAMs{]}). Only the GAM approach was reproduced, with the assistance
@@
 The reproduced results were qualitatively identical to those in the
-original article. For example, all Lyapunov exponents estimated by
-direct method were positive, as in the original article, consistent with
-chaotic dynamics. Quantitative differences may have resulted from
-difference in algorithms used. For example,the original used the
+original article, and therefore support the conclusion of chaotic
+dynamics produced by non-linear species interactions. For example, all
+Lyapunov exponents estimated by direct method were positive, as in the
+original article, consistent with chaotic dynamics. Quantitative
+differences may have resulted from difference in algorithms used. For
+example,the original used the
 \href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to

### 2016-02-17 article/article.tex
small edits

@@
 mid 2014} and since it is a bit less well integrated with R, we instead
-use the tseriesChaos package \textcite{tseriesChaos}, which in any case
-was largely inspired by the TISEAN project. In addition, there may have
-been some difference in algorithm parameters, as not all parameters
+use the tseriesChaos package (\textcite{tseriesChaos}), which in any
+case was largely inspired by the TISEAN project. In addition, there may
+have been some difference in algorithm parameters, as not all parameters
 required by the functions we used were reported in the original ms.

### 2016-02-17 article/article.tex
small edits

--- 2016-02-12 article/article.tex
@@
                     \author[1]{Frank Pennekamp}
-                            \affil[1]{Institute of Evolutionary Biology and Environmental Studies, University
+                            \affil[1]{Department of Evolutionary Biology and Environmental Studies, University
 of Zurich, Zurich, Switzerland}
@@
     \bfseries\normalsize
-      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
+      \href{https://github.com/opetchey/ReScience-submission/tree/petchey-plebani-pennekamp-2016/article}{\faGithubAlt~Article repository}
   \end{repobox}
@@
     \bfseries\normalsize
-      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
+      \href{https://github.com/opetchey/ReScience-submission/tree/petchey-plebani-pennekamp-2016/code}{\faGithubAlt~Code repository}
+  \end{repobox}
+      \vspace{-1mm}
+  \begin{repobox}
+    \bfseries\normalsize
+      \href{https://github.com/opetchey/ReScience-submission/tree/petchey-plebani-pennekamp-2016/data}{\faGithubAlt~Data repository}
   \end{repobox}
@@
 ability to predict dynamics is an open question with considerable
-applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.
+applied importance (\textcite{Petchey2015} \textcite{Mouquet2015}.)
 
@@
 This reproduction started with the raw data (source given below) and
-used information from the original paper, the supplementary information
+used information from the original paper,
 \href{http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf}{the
-Supplement to the Nature paper}, and communications with Elisa Benincà,
-who provided for comparison digitised data from the original article,
-and Stephen Ellner, who provided code and data used to produce results
-in the original paper. Use of this data and code in this reproduction is
-indicated below.
+supplement to the original paper}, and communications with Elisa
+Benincà, who provided for comparison digitised data from the original
+article, and Stephen Ellner, who provided code and data used to produce
+results in the original paper. Use of this data and code in this
+reproduction is indicated below.
 
@@
 \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
-Ecology Letters publication} \textcite{Beninca2009}. The Excel file
+Ecology Letters publication} (\textcite{Beninca2009}). The Excel file
 contains several datasheets. Two were particularly important, as they
@@
 making the observation occur half way between the two surrounding dates.
-The original Protozoa variable in the original dataset contains some
-numbers with a comma (instead of period) as the decimal separator; these
-were changed to periods.
+The Protozoa variable in the original dataset contains some numbers with
+a comma (instead of period) as the decimal separator; these were changed
+to periods.
 
@@
 The R language and environment for statistical computing and graphics
-was used to make the reproduction. Additional R packages required are
-specified in the code associated with this reproduction. The code for
-making the analyses and figures presented in this reproduction resides
-in an R markdown document, as well as a source file containing some
-required functions. Some of the code takes several minutes to run, so an
-intermediate data file is provided with results from this code.
+(\textcite{R}) was used to make the reproduction. Additional R packages
+required are specified in the code associated with this reproduction.
+The code for making the analyses and figures presented in this
+reproduction resides in an R markdown document, as well as a source file
+containing some required functions. Some of the code takes several
+minutes to run, so an intermediate data file is provided with results
+from this code.
 
@@
 
-The reproduced populations dynamics were very similar to those in figure
-1b-g of the original publication (figure \ref{fig:dynamics}). As most
-analyses were performed on fourth root transformed values, we plot
+The reproduced populations dynamics (figure \ref{fig:dynamics}) were
+very similar to those in figure 1b-g of the original publication. As
+most analyses were performed on fourth root transformed values, we plot
 these, rather than raw abundances with a y-axis break as in the original
@@
 version of the data in the published Excel file mentioned above.
-Transformation steps were:
+Subsequent transformation steps were:
 
@@
 \ref{fig:corr_comp}). Deviations between the original and reproduced
-correlations are relatively small (less than 0.072 units) and
-infrequent.
+correlations are relatively small and infrequent.
 
@@
 
-Reproduced divergence rates (figure \ref{fig:divergence}) and therefore
-Lyapunov exponents (figure \ref{fig:LE_comparison}) were somewhat
-different from those in the original article. The original article
-states: ``the distance between initially nearby trajectories increased
-over time, and reached a plateau after about 20--30 days''. The
-reproduced results are consistent with this statement. The original
-article also stated that the analyses ``yielded significantly positive
-Lyapunov exponents of strikingly similar value for all species (Fig. 3;
-mean exponent = 0.057 per day, s.d. = 0.005 per day, n = 9)''.
-Reproduced exponents had very similar mean value, but had larger
-standard deviation (mean = 0.060 and s.d. = 0.014).
+The original article states: ``the distance between initially nearby
+trajectories increased over time, and reached a plateau after about
+20--30 days''. The reproduced results (figure \ref{fig:divergence}) are
+consistent with this statement. The original article also stated that
+the analyses ``yielded significantly positive Lyapunov exponents of
+strikingly similar value for all species (Fig. 3; mean exponent = 0.057
+per day, s.d. = 0.005 per day, n = 9)''. Reproduced exponents had very
+similar mean value, but had larger standard deviation (mean = 0.060 and
+s.d. = 0.014), resulting from quantitative differences in reproduced
+divergence rates (figure \ref{fig:divergence}) and therefore Lyapunov
+exponents (figure \ref{fig:LE_comparison}).
 
@@
 \caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
-exponents (figure 3 in the original article).}
+exponents (equivalent to figure 3 in the original article).}
 \end{figure}
@@
 of code donated by Stephen Ellner. The original article obtained a
-global Lyapunov exponent of λ=0.08 day-1. The reproduced value was 0.04.
+global Lyapunov exponent of 0.08 per day. The reproduced value was 0.04.
 We did not reproduce the bootstrapping used to give confidence intervals
@@
 original article, and therefore support the conclusion of chaotic
-dynamics produced by non-linear species interactions. For example, all
-Lyapunov exponents estimated by direct method were positive, as in the
-original article, consistent with chaotic dynamics. Quantitative
-differences may have resulted from difference in algorithms used. For
-example,the original used the
+dynamics. For example, all Lyapunov exponents estimated by direct method
+were positive, as in the original article. Quantitative differences may
+have resulted from difference in algorithms used. For example, the
+original used the
 \href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
@@
 mid 2014} and since it is a bit less well integrated with R, we instead
-use the tseriesChaos package \textcite{tseriesChaos}, which in any case
-was largely inspired by the TISEAN project. In addition, there may have
-been some difference in algorithm parameters, as not all parameters
-required by the functions we used were reported in the original ms.
+used the tseriesChaos package (\textcite{tseriesChaos}), which in any
+case was largely inspired by the TISEAN project. In addition, there may
+have been some difference in algorithm parameters, as not all parameters
+required by the function we used were reported in the original ms.
 
-Considerable quantitative difference in predictability decay
-(\ref{fig:prediction_distance}) remain unexplained, though could result
+Considerable quantitative difference in predictability decay (figure
+\ref{fig:prediction_distance}) remain unexplained, though could result
 from original analyses using neural networks, and the reproduction using
@@
 \href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
-on github. Many thanks to Elisa Benincà and Steve Ellner for assistance,
-and to Steve for providing unpublished data and code.
+on github. Many thanks to Elisa Benincà and Stephen Ellner for
+assistance, and to Stephen Ellner for providing unpublished data and
+code.
 

### 2016-02-19 code/report.Rmd
checking global LE

@@
 ```{r}
-spp.abund <- read.csv(text=getURL(paste0(repo,"/data/species_abundances_original.csv")), skip=7, header=T)
+spp.abund <- read.csv(text=getURL(paste0(repo,"/data/original/species_abundances_original.csv")), skip=7, header=T)
 
@@
 ```{r}
-nuts <- read.csv(text=getURL(paste0(repo,"/data/nutrients_original.csv")), skip=7, header=T)
-#nuts <- read.csv("~/Dropbox (Dept of Geography)/RREEBES/Beninca_etal_2008_Nature/data/nutrients_original.csv", skip=7, header=T)
+nuts <- read.csv(text=getURL(paste0(repo,"/data/original/nutrients_original.csv")), skip=7, header=T)
 
@@
 ```{r}
-tr <- read.csv(text=getURL(paste0(repo,"/data/transformed_data_Nature2008.csv")),
+tr <- read.csv(text=getURL(paste0(repo,"/data/original/transformed_data_Nature2008.csv")),
                skip=7, na.string="")
@@
 # read in the correlations given in the original publication
-original.cors <- read.csv(text=getURL(paste0(repo,"/data/table1_original_article.csv")), skip=0, header=T, row.names = 1)
+original.cors <- read.csv(text=getURL(paste0(repo,"/data/original/table1_original_article.csv")), skip=0, header=T, row.names = 1)
 original.cors <- as.matrix(original.cors)
@@
 ## read in LE values given in original publication
-original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original_direct_LEs.csv")), skip=0, header=T)
+original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original/original_direct_LEs.csv")), skip=0, header=T)
 both_LEs <- full_join(original.LEs, LEs) # merge the reproduced and original values
@@
 ```{r}
-from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
+from.steve <- read.csv(text=getURL(paste0(repo,"/data/original/interp_short_allsystem_newnames.csv")), header=T)
 from.steve <- gather(from.steve, Species, Abundance, 2:13)
@@
 
-## joint the data, again first making days character, then back to numeric after
+## join the data, again first making days character, then back to numeric after
 mt$Day.number <- as.character(mt$Day.number)
@@
 Looks good.
+
+```{r, eval=F, echo=F}
+## code to export the interpolated data, with which to test the GAM functions
+mt_wide <- spread(mt, key=variable, value=value)
+write.csv(mt_wide, "~/Desktop/Repro_data_for_gam_test.csv")
+```
 
@@
 
+```{r, eval=F, echo=F}
+## code to export the interpolated and transformeddata, with which to test the GAM functions
+write.csv(X, "~/Desktop/Repro_data_for_gam_test_transformed.csv")
+```
+
 Load and run the functions, or read in from data file.
@@
 ## load the already saved data from github (this can take some time depending on the internet connection)
-source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))
+source_data(paste0(repo, "/data/reproduction/GLE_estimate.Rdata?raw=True"))
 
@@
 ## read in the data of the original article that came direct from Elisa
-original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
+original_preds <- read.csv(text=getURL(paste0(repo,"/data/original/original_rsquared.csv")), skip=0, header=T)
 opw <- gather(original_preds, Species, r_squared, 2:25)

### 2016-02-19 code/GAM_test/GAM_Owen_method.R

@@ -1,0 +1,42 @@
+rm(list=ls())
+
+
+##  Computing global LE with Owen's version of Steve's code
+library(RCurl)
+repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/beninca_dev"
+script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)
+eval(parse(text = script))
+
+
+# follow Elisa's detrending 
+# green flag = nano 
+# melosira = net 
+
+### Before running this code, you must do at the R command line 
+#	rm(list=ls(all=TRUE)); source("c:/lenns/R/lennsBIC.R"); 
+
+require(mgcv); 
+
+######################################################################
+#  Load the data into a properly structured data frame 
+#######################################################################
+
+## load dataset that allows conversion of names
+name_table <- read.csv("../../data/reproduction/repro_steve_name_table.csv")
+
+## choose which dataset to use
+dataset_to_use <- "reproduced" # 0.03748704
+#dataset_to_use <- "from_steve" # 0.08415112
+
+if(dataset_to_use=="from_steve") {
+  X=read.csv("../../data/original/Steve_data_for_gam_test_transformed.csv", row.names=1)
+  names(X) <- name_table$repro_names[match(names(X), name_table$steve_posttrans_names)]
+  }
+
+if(dataset_to_use =="reproduced") {
+  X=read.csv("../../data/reproduction/Repro_data_for_gam_test_transformed.csv", row.names=1)
+}
+
+
+LE <- Get_GLE_Beninca(X)
+LE[[1]]

### 2016-02-19 code/GAM_test/GAM_Steve_method.R

@@ -1,0 +1,173 @@
+# follow Elisa's detrending 
+# green flag = nano 
+# melosira = net 
+
+### Before running this code, you must do at the R command line 
+#	rm(list=ls(all=TRUE)); source("c:/lenns/R/lennsBIC.R"); 
+
+require(mgcv); 
+
+######################################################################
+#  Load the data into a properly structured data frame 
+#######################################################################
+
+## load dataset that allows conversion of names
+name_table <- read.csv("../../data/reproduction/repro_steve_name_table.csv")
+
+## choose which dataset to use
+#dataset_to_use <- "reproduced" # 0.03668501
+dataset_to_use <- "from_steve" # 0.08415112
+
+if(dataset_to_use=="from_steve") {
+  X=read.csv("../../data/original/interp_short_allsystem_original.csv")
+}
+
+if(dataset_to_use =="reproduced") {
+  X=read.csv("../../data/reproduction/Repro_data_for_gam_test.csv", row.names=1)
+  names(X) <- name_table$steve_names[match(names(X), name_table$repro_names)]
+}
+
+
+## Select the time period to use 
+start.longest=334; start.longer=808; start.shorter=1035; 
+e=X$Day.Number > start.longer; X=X[e,];
+e=X$Day.Number < 2654; X=X[e,]; 
+
+
+
+
+### From here on is code pretty much identical to what Steve sent
+
+## Function to detrend series 
+Detrend=function(z) {
+  fit=ksmooth(X$Day.Number,z,bandwidth=300,kernel="normal"); 
+  return(z-fit$y); 
+}
+
+# create the data series 
+Cyclopoids=X$Cyclopoids^0.25; 
+Protozoa=X$Protozoans^0.25; 
+Rotifers=X$Brachionus^0.25; 
+Calanoids=X$Eurytemora^0.25;   
+Pico=X$Picophytopl^0.25;
+Nano=X$Green.Flagellates^0.25; 
[…]

### 2016-02-19 code/report.Rmd
fix

@@
 ```{r}
-repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/beninca_dev"
+repo <- "https://raw.githubusercontent.com/opetchey/ReScience-submission/petchey-plebani-pennekamp-2016"
 ```

### 2016-02-19 code/report.Rmd
fixes

@@
 ```{r}
-from.steve <- read.csv(text=getURL(paste0(repo,"/data/original/interp_short_allsystem_newnames.csv")), header=T)
+from.steve <- read.csv(text=getURL(paste0(repo,"/data/original/interp_short_allsystem_original.csv")), header=T)
+name_table <- read.csv(text=getURL(paste0(repo,"/data/reproduction/repro_steve_name_table.csv")), header=T)
+names(from.steve) <- name_table$repro_names[match(names(from.steve), name_table$steve_names)]
+
 from.steve <- gather(from.steve, Species, Abundance, 2:13)
@@
 
+## To test the reproduction uncomment the next line of code
 ## don't run this unless you have some time (it takes a while)

### 2016-02-23 code/report.Rmd
new method for calc. r-squared

@@
 
+The process is to transform the data, to fit GAMs, and then to calculated the global Lyapunov exponent.
+
 It seems that this must have been done on data without zeros removed, because removal of the zeros would leave too many missing values in the GAMs.
@@
 
-Load and run the functions, or read in from data file.
-
-```{r}
+Load and run the functions, or read in from data file. Chunk not run by default, as take a minute or so. At last run, gave GLE of 0.03730891
+
+```{r, eval=F}
 # read script lines from website
 script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)
-
-# parase lines and evealuate in the global environement
 eval(parse(text = script))
 
-## To test the reproduction uncomment the next line of code
-## don't run this unless you have some time (it takes a while)
-#LE <- Get_GLE_Beninca(X)
-#save(LE, file="~/Desktop/GLE_estimate.Rdata")
-
-## load the already saved data from github (this can take some time depending on the internet connection)
-source_data(paste0(repo, "/data/reproduction/GLE_estimate.Rdata?raw=True"))
-
-LE[[1]]
+## Fit the GAMs
+gams <- Fit_GAMs(X, gval=1.4, T_lag=1)
+
+## calculate and print the GLE
+Get_LE_from_fit2(gams, X1, epsval=0.01)
+
 ```
@@
 This is quite far from the number using code and data from Steve (0.08415112).
-But recall that the functions have been carefully checked.
-Probably it is the data going into this function.
-A final step would be to save the data here, and take it into Steve's function.
-(This point is an issue on github.)
+This is caused by small differences in the data (and not by the functions), as we've found in using the code in the folder "GAM_test".
 
@@
 
-```{r, echo=FALSE}
-## put the models together into a list
-all.gams <- LE[[2]]
-#names(all.gams)
-```
+The procedure is to fit GAMs (neural nets in the original publication) with response variable a function of explanatory variables at lag T, then to record r-squared of observed and fit data.
+Chunk not run by default, due to time required.
+
+```{r, echo=FALSE, eval=FALSE}
+
+T_lags <- 12
+num_spp <- length(gams)
+rsq_vals <- matrix(NA, T_lags, num_spp)
+dimnames(rsq_vals)[[2]] <- names(gams)
+
+for(i in 1:T_lags) {
+  gams <- Fit_GAMs(X, gval=1.4, T_lag=i)
+  for(j in 1:num_spp)
+    rsq_vals[i,j] <- cor(gams[[j]]$fitted.values, gams[[j]]$model[1])^2
+}
+
+save(rsq_vals, file="~/Desktop/rsq_vals.Rdata")
+
+```
+
+<!--
+Until the close HTML comment are some code Owen was previously using to calculate the predictability decline.
 
@@
 
+End of html comment
+-->
+
 And plot our version of figure 2:
 ```{r}
+## Load the r-squared values that were previously saved (above)
+## load the already saved data from github (this can take some time depending on the internet connection)
+source_data(paste0(repo, "/data/reproduction/rsq_vals.Rdata?raw=True"))
+
+cors <- data.frame(dist=1:12, rsq_vals)
+cors.long <- gather(cors, key=dist)
+names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
+
 cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",

### 2016-02-23 code/GAM_test/GAM_Owen_method.R

@@ -37,6 +37,10 @@ if(dataset_to_use =="reproduced") {
   X=read.csv("../../data/reproduction/Repro_data_for_gam_test_transformed.csv", row.names=1)
 }
 
+## Fit the GAMs
+gams <- Fit_GAMs(X, gval=1.4, T_lag=1)
 
-LE <- Get_GLE_Beninca(X)
-LE[[1]]
+## calculate and print the GLE
+Get_LE_from_fit2(gams, X, epsval=0.01) 
+
+

### 2016-02-23 code/indirect_method_functions.R

@@ -1,6 +1,7 @@
 ### Assume the data comes in as a matrix, with the time in the rows and species in the columns
 
-Get_GLE_Beninca <- function(X1,gval=1.4) {
+
+Fit_GAMs <- function(X1, gval=1.4, T_lag) {
   
   #X1=X
   
@@ -16,7 +17,7 @@ Get_GLE_Beninca <- function(X1,gval=1.4) {
   X1 <- X1[,order(colnames(X1))]
   
   ## make a matrix of unlagged and lagged variables
-  Z <- cbind(X1[-1,], X1[-end.t,])
+  Z <- cbind(X1[-1:-T_lag,], X1[-end.t:(-end.t+T_lag-1),])
   colnames(Z) <- c(paste("y_", names(X1), sep=""), names(X1))
   
   #1 Cyclopoids=f(Cyclopoids,Protozoa,Rotifers)
@@ -142,14 +143,12 @@ Get_GLE_Beninca <- function(X1,gval=1.4) {
                    Soluble.reactive.phosphorus=fit.Soluble.reactive.phosphorus,                  
                    Total.dissolved.inorganic.nitrogen=fit.Total.dissolved.inorganic.nitrogen)
   all.gams
-
-  ##save(all.gams, file="~/Desktop/all.gams.Rdata")
-  ##save(Z, file="~/Desktop/Z.Rdata")
   
-  LE <- Get_LE_from_fit2(all.gams, Z[,13:24])
-  rez <- list(LE, all.gams, Z)
-  rez
 }
+
+
+
+
 
 
 
@@ -169,9 +168,6 @@ gamgrad_v2=function(gamfit, eps, zv2) {
 }
 
 
-
-
-
 Get_LE_from_fit2 <- function(fits, X1, epsval=0.01) {
   
   nt=length(X1[,1]-1)
@@ -187,3 +183,161 @@ Get_LE_from_fit2 <- function(fits, X1, epsval=0.01) {
   LE.ours=LE/(3.35*nt)
   LE.ours
 }
+
+
+
+
+## Below is defunct
+# 
+# 
+# Get_GLE_Beninca <- function(X1,gval=1.4) {
+#   
+#   #X1=X
+#   
+#   require(mgcv)
+#   
+#   ## prepare the lagging
+#   end.t <- dim(X1)[1]
+#   
+#   ## change X to be a data.frame
+#   X1 <- as.data.frame(X1)
+#   
[…]

### 2016-02-23 code/report.Rmd
report updated with latest r-squared

@@
 
-<!--
-Until the close HTML comment are some code Owen was previously using to calculate the predictability decline.
-
-Need to predict until about 40 days ahead = 40 / 3.35 time steps = 12 time steps. Do this from each time in the time series as initial abundances.
-
-From each time in the series, predict 12 steps ahead. Put results in an array: time x species x start.location.
-time will be 12, species will be 12, start location will be length of time series - 12
-
-```{r}
-## some preliminaries
-x <- X 
-Z <- LE[[3]]
-Z1 <- Z[,13:24]
-all.species <- names(all.gams)
-time.to.predict <- 12
-nn <- length(Z1[,1])-time.to.predict
-## create the array of predictions
-preds <- array(NA, c(12, 12, nn))
-dimnames(preds) <- list(dist=1:12,
-                        species=all.species,
-                        start.time=1:nn)
-## fill in the array with the one step ahead predictions
-for(i in 1:length(all.species))
-  preds[1,i,] <- predict(all.gams[[i]])[1:nn]
-
-# Now for the predictions at t+2 from predictions at t+1
-pred.time <- 2
-for(pred.time in 2:12) {
-  for(i in 1:length(all.species))
-    preds[pred.time,i,pred.time:nn] <- predict(all.gams[[i]],
-                                                newdata=as.data.frame(t(preds[pred.time-1,,])))[pred.time:nn]
-}
-```
-
-Get the correlations and do some house keeping:
-```{r}
-cors <- matrix(NA, 12, 12)
-dimnames(cors) = list(dist=1:12,
-                        species=names(x))
-for(i in 1:12) {
-  for(j in 1:12) {
-    cors[j,i] <- cor(Z1[1:nn,i], preds[j, i, ], use="complete.obs")^2
-}}
-cors <- data.frame(dist=as.numeric(I(rownames(cors))), cors)
-cors.long <- gather(as.data.frame(cors), key=dist)
-names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
-```
-
-End of html comment
--->
-
 And plot our version of figure 2:
@@
 
-ggplot(cors.long, aes(x=Prediction_distance*3.75, y=Correlation)) +
+ggplot(filter(cors.long, Prediction_distance<13),
+       aes(x=Prediction_distance*3.35, y=Correlation)) +
   xlab("Prediction time (days)") +
@@
   geom_point(data=filter(opw, Model=="nonlinear"),
-             aes(x=prediction_time, y=r_squared))
+             aes(x=prediction_time, y=r_squared)) +
+  ylim(0,1)
 

### 2016-02-23 article/article.tex
latest build

@@
 predictabilities, which were calculated from the GAMs, were consistent
-with these qualitative statements, though were quantitatively different
-(\ref{fig:prediction_distance}). We did not reproduce the predictability
-estimates for linear models.
+with these qualitative statements, and were most often quantitatively
+similar (\ref{fig:prediction_distance}). We did not reproduce the
+predictability estimates for linear models.
 
@@
 
-Considerable quantitative difference in predictability decay (figure
+Quantitative differences in predictability decay (figure
 \ref{fig:prediction_distance}) remain unexplained, though could result
 from original analyses using neural networks, and the reproduction using
-GAMs.
+GAMs. The difference in estimate of the global Lyapunov exponent (0.08
+in the original article and 0.04 here) appears to be due to difference
+in data used (illustrated in figure \ref{fig:trans_comp2}).
 

### 2016-02-23 code/report.Rmd
report update

@@
 
+## Some preliminaries, including arranging storage
 T_lags <- 12
@@
 
+## fit gams for all the required lags
 for(i in 1:T_lags) {
   gams <- Fit_GAMs(X, gval=1.4, T_lag=i)
+  ## and get the r-squared for each
   for(j in 1:num_spp)
@@
 
-Quite different patterns from in figure 2 of the nature paper.
-There is relatively little information in the paper or supplement about how figure 2 data was produced, so difficult to pin down the difference without asking authors.
+Some species very similar to figure 2 of the nature paper, some more different.
+Difference must be due to data in figure 2 coming from neural net models, whereas here it comes from the GAMs.
+Qualitative result the same as in the original paper, however.

### 2016-03-11 code/report.Rmd
add install.packages command for convenience

@@
 ```{r, message=FALSE}
+#install.packages(c("tidyr","dplyr", "lubridate", "stringr",
+#                   "ggplot2", "RCurl", "pracma", "oce",
+#                   "tseriesChaos", "reshape2", "mgcv", "repmis",
+#                   "magrittr", "knitr"))
 library(tidyr)
@@
 library(magrittr)
+library(knitr)
 ```
@@
 ```{r}
-library(knitr)
 table1b <- kable(cor.cp, format="markdown", col.names = colnames(cor.cp), align="c",

### 2016-03-23 code/report.Rmd
remove some possibly interfering code

@@
 
-Only keep the upper triangle of the cor.pvals matrix:
-```{r}
-#for(i in 1:10){
-#  for(j in 1:10){
-#  cor.coefs[i,j] <- ifelse(i<j, cor.coefs[i,j], NA)
-#}}
-```
-
 Get p-vals:
@@
     #plot(final_wide[,2], final_wide[,3])
-```
-
-Only keep the upper triangle of the cor.pvals matrix:
-```{r}
-#for(i in 1:10){
-#  for(j in 1:10){
-#  cor.pvals[i,j] <- ifelse(i<j, cor.pvals[i,j], NA)
-#}}
 ```
@@
 ```{r}
+dim(cor.coefs) ## should be 12,12
 oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)

### 2016-03-23 code/report.Rmd
protecting against a problem
that could be caused by running lines out of order / repeating lines

@@
 oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)
-cor.coefs <- cor.coefs[oo,oo]
-cor.cp <- cor.cp[oo,oo]
-for(i in 1:length(cor.cp[1,]))
+cor.coefs1 <- cor.coefs[oo,oo]
+cor.cp1 <- cor.cp[oo,oo]
+for(i in 1:length(cor.cp1[1,]))
   for(j in 1:i)
-  cor.cp[i,j] <- ""
+  cor.cp1[i,j] <- ""
 #sn <- c("Bacteria", "Harps", "Ostr", "N", "P", "Picophyt", "Nanophyt", "Rotifers", "Protozoa", "Calanoids")
-sn <- substr(rownames(cor.cp), 1, 4)
+sn <- substr(rownames(cor.cp1), 1, 4)
 sn[4:5] <- c("N", "P")
-dimnames(cor.cp) <- list(sn, sn)
+dimnames(cor.cp1) <- list(sn, sn)
 ```
@@
 ```{r}
-table1b <- kable(cor.cp, format="markdown", col.names = colnames(cor.cp), align="c",
+table1b <- kable(cor.cp1, format="markdown", col.names = colnames(cor.cp1), align="c",
                 caption="Table 1.'Correlations between the species in the food web.
@@
 
-qplot(x=as.vector(original.cors), y=as.vector(cor.coefs), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
+qplot(x=as.vector(original.cors), y=as.vector(cor.coefs1), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
       xlab="Correlations in original article", ylab="Correlations calculated\nin this reproduction") + geom_abline(intercept=0,slope=1, colour="red")

### 2016-03-23 code/report.Rmd
Improved specification of required R packages

@@
 
+# R package requirements
+
+The following packages must be installed.
+
+```{r}
+#install.packages(c("tidyr","dplyr", "lubridate", "stringr",
+#                   "ggplot2", "RCurl", "pracma", "oce",
+#                   "tseriesChaos", "reshape2", "mgcv", "repmis",
+#                   "magrittr", "knitr", "devtools"))
+```
+
 # R session info
@@
 
-All required libraries:
+Load required packages:
 ```{r, message=FALSE}
-#install.packages(c("tidyr","dplyr", "lubridate", "stringr",
-#                   "ggplot2", "RCurl", "pracma", "oce",
-#                   "tseriesChaos", "reshape2", "mgcv", "repmis",
-#                   "magrittr", "knitr"))
 library(tidyr)

### 2016-03-26 code/report.Rmd
added setup info and possible fix

@@
 #                   "magrittr", "knitr", "devtools"))
+```
+
+You may also need to create a dependency on libcurl via apt-get:
+
+```
+sudo apt-get -y build-dep libcurl4-gnutls-dev
+sudo apt-get -y install libcurl4-gnutls-dev
 ```
@@
 tr <- gather(tr, key="Species", value="Abundance", 2:13) ## make long format
+tr$Species <- as.factor(tr$Species)
 levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)

### 2016-03-29 code/report.Rmd
various revisions

@@
 ---
-title: "Reproduce anaylses of Beninca et al (2008)"
-author: "Owen Petchey et al"
-date: "29 Jan 2015"
+title: "Reproducing anaylses of Beninca et al (2008)"
+author: "Owen Petchey, Marco Plebani, Frank Pennekamp"
+date: "28 Mar 2015"
 output:
@@
 This reproduction was made as part of the Reproducible Research in Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES) Course, lead by Owen Petchey at the University of Zurich.
-More information about the course [here](https://github.com/opetchey/RREEBES/blob/master/README.md) on github.
+More information about the [RREEBES course on github](https://github.com/opetchey/RREEBES/blob/master/README.md) .
 
-The code and data for the reproduction are [here](https://github.com/opetchey/RREEBES/tree/Beninca_development) on github.
+[The code and data for the reproduction are on github.](https://github.com/opetchey/RREEBES/tree/Beninca_development).
 
@@
 In the code below, the data and any other files are read from github, which means there must be a connection to github.
+
+# Testing
+
+This code in this document has been tested on:
+
+```
+platform       x86_64-apple-darwin14.5.0   
+arch           x86_64                      
+os             darwin14.5.0                
+system         x86_64, darwin14.5.0        
+status                                     
+major          3                           
+minor          2.3                         
+year           2015                        
+month          12                          
+day            10                          
+svn rev        69752                       
+language       R                           
+version.string R version 3.2.3 (2015-12-10)
+nickname       Wooden Christmas-Tree   
+```
+
+and
+
+```
+Win 7 64-bit with R 3.1.3.
+```
 
@@
 
-The Protozoa variable contains some numbers with comman as the decimal separator. This creates a question about what dataset was used for the original analyses, as it could not have been this one.
+The Protozoa variable contains some numbers with comma as the decimal separator. This creates a question about what dataset was used for the original analyses, as it could not have been this one.
 ```{r}
@@
 spp.abund$Date <- dmy(spp.abund$Date)
-# y123 <- year(spp.abund$Date) %% 100
-# year(spp.abund$Date) <- 1900 + y123
-# rm(y123)
 ```
 
-Check dates match the Day.number (should give true):
+Check dates match the Day.number (should give TRUE):
 ```{r}
-sum(spp.abund$Day.number == 1+as.numeric((spp.abund$Date - spp.abund$Date[1]) / 24 / 60 / 60)) == length(spp.abund$Date)
+stopifnot(sum(spp.abund$Day.number == 1+as.numeric((spp.abund$Date - spp.abund$Date[1]) / 24 / 60 / 60)) == length(spp.abund$Date))
 ```
@@
 ```{r}
-spp.abund <- gather(spp.abund, "variable", "value", 3:12)
+spp.abund <- gather(spp.abund, key="variable", value="value",
+                    3:12)
+spp.abund$variable <- as.factor(spp.abund$variable)
 str(spp.abund)
@@
 nuts <- select(nuts, -NO2, -NO3, -NH4)
-#nuts$Date[duplicated(nuts$Date)]
-#which(duplicated(nuts$Date))
 nuts <- gather(nuts, "variable", "value", 3:4)
+nuts$variable <- as.factor(nuts$variable)
 str(nuts)
@@
 <!--
-Here are some versions of the graphs that are not included in the report, hence the html comment:
+Here are some versions of the graphs that were interesting to produce originally, and may be of some interest, but are not included by default for brevity:
 
@@
 ```{r}
-#aggregate(Day.number ~ variable, all.data, min)
-#aggregate(Day.number ~ variable, all.data, max)
-#xout <- seq(343.35, 2657.2, by=3.35)
 xout <- seq(343.35, 2658, by=3.35)
-#range(xout)
 ```
@@
 # run the function and tidy the data
-# (somewhat ashamed of this code...)
 # remove NAs before interpolation, so they don't cause problems
@@
 
-# for(p in 1:nrow(res)){
-#     temp <- data.frame(variable=res[p,1],
-#                        res[p,2][[1]])
-#     if(p==1)
-#       trimmed_interp <- temp
-#     if(p>1)
-#       trimmed_interp <- rbind(trimmed_interp, temp)
-#   }
 ```
@@
   mutate(stand.y=as.numeric(scale(dt.value)))
-#summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
 ```
@@
 
+## some graphs that are of use for very detailed views of the two datasets. Not included in report by default for brevity.
 # ggplot(filter(two.datasets), aes(x=as.numeric(Day.number), y=stand.y)) +
@@
                                    Total.dissolved.inorganic.nitrogen="N"))
+
+# ensure alphabetical order of panels in figure
+two.datasets$variable <- factor(two.datasets$variable,
+                                 levels = sort(levels(as.factor(two.datasets$variable))))
 
@@
 
-# freq.est <- spec %>% group_by(group) %>% mutate(max_spec = max(spec), freq = freq)
-# freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
-# freq.est$freq <- 1/freq.est$freq
-
 # Calculate and plot Welch's periodogram
@@
   scale_y_log10()
-
-# freq.est <- wspec %>%
-#   group_by(group) %>%
-#   mutate(max_spec = max(spec), freq = freq)
-# freq.est <- subset(freq.est, max_spec==spec, select=c(freq,group))
-# freq.est$freq <- 1/freq.est$freq
-# #frequency(final$y)
-# ts <- as.ts(final$y, frequency = 0.3)
-# #time(ts)
 
@@
 }
-    #plot(final_wide[,2], final_wide[,3])
 ```
@@
 ```{r}
-dim(cor.coefs) ## should be 12,12
 oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)
@@
   cor.cp1[i,j] <- ""
-#sn <- c("Bacteria", "Harps", "Ostr", "N", "P", "Picophyt", "Nanophyt", "Rotifers", "Protozoa", "Calanoids")
 sn <- substr(rownames(cor.cp1), 1, 4)
@@
 
-Time delay (1), embedding dimension (6), and Theiler window (50) were used in the Nature report. Other parameters are chosen rather randomly!
+Time delay (1), embedding dimension (6), and Theiler window (50) were used in the Nature report. Other parameters are chosen rather arbitrarily.
 
@@
 diverg$days <- diverg$days*3.35
-#str(diverg)
 ```
@@
 ```{r}
-#diverg$Difference[is.na(diverg$Difference)] <- 0
 diverg$Difference[is.infinite(diverg$Difference)] <- NA
@@
   summarise(le=coef(lm(Difference[1:6] ~ days[1:6]))[2])
-#pval=summary(lm(Difference[1:6] ~ days[1:6]))$coefficients[2,4])
 ```
@@
                                    Total.dissolved.inorganic.nitrogen="N"))
+
+## ensure order of panels in figure same as in original
+diverg$Species_short <- factor(diverg$Species_short,
+                               levels=c("Calanoid.copepods", "Rotifers",
+                                        "Nanophytoplankton", "Picophytoplankton",
+                                        "Ostracods", "Harpacticoids", "Bacteria",
+                                        "N","P"))
+
 g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species_short) +
@@
 ff$Day.number <- as.numeric(ff$Day.number)
-#str(ff)
-#ff$difference <- ff$value - ff$value_steve
-#qplot(log10(ff$difference))
 qplot(data=ff, x=log10(value_steve), y=log10(value)) +
@@
 ```{r}
-## standardise
+## standardise (commented out by default; uncomment, and comment next chunck in order to work on standardised data)
 #scaled_trimmed_final <- group_by(trimmed_detrended, variable) %>%
@@
 #summarise(scaled_trimmed_final, mean=mean(stand.y), sd=sd(stand.y))
+```
 
+```{r}
 ## or don't standardise
@@
 final$y <- final$dt.value
-#summarise(final, mean=mean(y), sd=sd(y))
 ```
@@
 ## use next line to work on standardised
-#final.to.melt <- final[, c("variable", "y", "Day.number")]
 names(final.to.melt)[1] <- "Species"
@@
 X <- acast(melted, formula= Day.number ~ Species)
-#str(X)
 X <- as.data.frame(X)
@@
 names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
+cors.long$Variable <- as.factor(cors.long$Variable)
 
@@
 opw <- separate(opw, Species, c("Species", "Model"))
-
+opw$Species <- as.factor(opw$Species)
+opw$Model <- as.factor(opw$Model)
 name.mapping <- data.frame(Species=c("cyclo", "roti", "cala", "pico", "nano",

### 2016-03-29 code/report.Rmd
revision to attempt to fix behaviour of gather()

@@
 Win 7 64-bit with R 3.1.3.
+```
+
+# R options
+
+```{r}
+options(stringsAsFactors=FALSE)
 ```
@@
 spp.abund <- gather(spp.abund, key="variable", value="value",
-                    3:12)
-spp.abund$variable <- as.factor(spp.abund$variable)
+                    3:12, convert=FALSE)
+#spp.abund$variable <- as.factor(spp.abund$variable)
 str(spp.abund)
@@
 nuts <- select(nuts, -NO2, -NO3, -NH4)
-nuts <- gather(nuts, "variable", "value", 3:4)
-nuts$variable <- as.factor(nuts$variable)
+nuts <- gather(nuts, "variable", "value", 3:4, convert=F)
+#nuts$variable <- as.factor(nuts$variable)
 str(nuts)
@@
 tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
-tr <- gather(tr, key="Species", value="Abundance", 2:13) ## make long format
-tr$Species <- as.factor(tr$Species)
+tr <- gather(tr, key="Species", value="Abundance", 2:13, convert=F) ## make long format
+#tr$Species <- as.factor(tr$Species)
 levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
@@
 diverg <- as.data.frame(cbind(days=1:time.steps, diverg))
-diverg <- gather(diverg, Species, Difference, 2:10)
+diverg <- gather(diverg, Species, Difference, 2:10, convert=F)
 diverg$days <- diverg$days*3.35
@@
 mt <- cbind(Day.number=xout, mt)
-mt <- gather(mt, variable, value, 2:13)
+mt <- gather(mt, variable, value, 2:13, convert=F)
 ```
@@
 
-from.steve <- gather(from.steve, Species, Abundance, 2:13)
+from.steve <- gather(from.steve, Species, Abundance, 2:13, convert=F)
 names(from.steve) <- c("Day.number", "variable", "value")
@@
 cors <- data.frame(dist=1:12, rsq_vals)
-cors.long <- gather(cors, key=dist)
+cors.long <- gather(cors, key=dist, convert=F)
 names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")
-cors.long$Variable <- as.factor(cors.long$Variable)
+#cors.long$Variable <- as.factor(cors.long$Variable)
 
@@
 original_preds <- read.csv(text=getURL(paste0(repo,"/data/original/original_rsquared.csv")), skip=0, header=T)
-opw <- gather(original_preds, Species, r_squared, 2:25)
+opw <- gather(original_preds, Species, r_squared, 2:25, convert=F)
 opw <- separate(opw, Species, c("Species", "Model"))
-opw$Species <- as.factor(opw$Species)
-opw$Model <- as.factor(opw$Model)
+#opw$Species <- as.factor(opw$Species)
+#opw$Model <- as.factor(opw$Model)
 name.mapping <- data.frame(Species=c("cyclo", "roti", "cala", "pico", "nano",

### 2016-04-04 code/report.Rmd
figure 2, 3, 4 fixed (hopefully)

@@
 ```{r, warning=FALSE, fig.width=10}
+
+##########################
+## Figure 1 in the article
 g1 <- ggplot(aes(x=as.numeric(Day.number), y=value^0.25, col=variable), data=all.data) +
@@
 
-## graph
+##########################
+## Figure 2 in the article
 ggplot(two.datasets, aes(x=Abundance, y=stand.y)) + geom_point() +
@@
 ```{r}
-oo <- c(10, 9, 8, 11, 12, 6, 5, 3, 4, 2)
+oo <- c("Bacteria", "Harpacticoids", "Ostracods",
+        "Total.dissolved.inorganic.nitrogen", "Soluble.reactive.phosphorus",
+        "Picophytoplankton", "Nanophytoplankton", "Rotifers",
+        "Protozoa", "Calanoid.copepods")
 cor.coefs1 <- cor.coefs[oo,oo]
@@
 
+##########################
+## Figure 3 in the article
 qplot(x=as.vector(original.cors), y=as.vector(cor.coefs1), ylim = c(-0.4,0.4), xlim=c(-0.4, 0.4),
@@
 diverg <- gather(diverg, Species, Difference, 2:10, convert=F)
+diverg$Species <- as.factor(diverg$Species)
 diverg$days <- diverg$days*3.35
@@
 ## ensure order of panels in figure same as in original
+## This does affect panel order on Owen's computer
 diverg$Species_short <- factor(diverg$Species_short,
@@
 
+## This puts the panels in alphabetical order if uncommented
+#diverg$Species_short <- factor(diverg$Species_short,
+#                               levels=sort(c("Calanoid.copepods", "Rotifers",
+#                                        "Nanophytoplankton", "Picophytoplankton",
+#                                        "Ostracods", "Harpacticoids", "Bacteria",
+#                                        "N","P")))
+
+str(LEs)
+levels(LEs$Species_short)
+
+##########################
+## Figure 4 in the article
 g1 <- ggplot(diverg, aes(x=days, y=Difference)) + geom_point() + facet_wrap(~Species_short) +

### 2016-05-06 article/detorakis-2016.tex
Submitting

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Georgios Detorakis,
  },
            pdftitle=
{
[Re] Multiple dynamical modes of thalamic relay neurons: rhythmic bursting
        and intermittent phase-locking
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\usepackage{float,lscape}
\usepackage{xcolor,colortbl}
\usepackage{amssymb,amsmath,array,amsthm}
\usepackage{marvosym}
\usepackage{array}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\definecolor{Gray}{gray}{0.75}
\definecolor{LightGray}{gray}{0.95}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{detorakis-2016.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Multiple dynamical modes of thalamic relay neurons: 
        rhythmic bursting and intermittent phase-locking}

    \usepackage{authblk}
                        \author[1]{Georgios Detorakis}
                        \affil[1]{ Department of Cognitive Sciences, UC Irvine, CA, USA}

\date{\vspace{-5mm}
    \sffamily \small \href{mailto:gdetorak@uci.edu}{gdetorak@uci.edu}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{Multiple dynamical modes of thalamic relay neurons: rhythmic
        bursting and intermittent phase-locking}, Wang, X-J, Neuroscience,
        59(1), pg. 21--31, 1994.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

This work introduces a reference implementation of a neuron model 
for thalamocortical relay neurons, \cite{wang:1994} proposed by X-J Wang.
The model is conductance-based and takes advantage of an interplay between 
a T-type calcium current and a non-specific cation sag current and thus it
is able to generate spindle and delta rhythms. Another feature of this model
is the presence of an intermittent phase-locking phenomenon
where action potentials of sodium take place in a non-periodic manner, despite
the fact that they are phase-locked to the periodic input current. Finally,
the model is capable of generating tonic spike patterns. The reference 
implementation in this work, does not aim to reproduce all the results of 
the original article, but some of them in order to test if the model can be
easily reproduced and at what degree the original results are reproducible.
The source code of reference implementation is written in Python (Numpy,
Scipy and Matplotlib).

\section{Methods}\label{methods}

The model is described in this section following the paradigm of 
Nordlie et al, \cite{nordlie:2009}. Therefore, a brief description of
the model, equations, parameters and inputs are given in the form of 
tables.

Table~\ref{Table:1} provides a description of the model and Table~\ref{Table:2}
gives a glimpse of the input used during simulations. Table~\ref{Table:3}
introduces the equations of the model. The neuron model is conductance-based
consisting in four differential equations describing the dynamics of membrane
potential and the kinetics of a T-Type calcium current, a Sag current channel
and a Potassium channel. The rest currents are described by algebraic
equations. Finally, all simulations parameters are given in
Table~\ref{Table:5} and the simulations times can be found in
Table~\ref{Table:2}.

The reference implementation has been done in a Python class (Python $3.5.1$)
along with Numpy (version $1.10.4$), Scipy (version $0.17.0$) and Matplotlib
($1.5.1$). The numerical integration has been done by the \emph{ode} method 
of Scipy \emph{integrate} package taking advantage of the BDF multistep method
\cite{ascher:1998}. This is a diversion from the original implementation, since
in \cite{wang:1994}, the author has numerically integrated the system of 
the four ODEs by using a fifth-order adaptive size Runge-Kutta method. This
could cause some of the numerical deviations we noticed during the
implementation of the model. The values for the input current must be 
slightly elevated in order to achieve exactly the same results as
in the original implementation.

All the simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
generation i$7$ processor, $8$GB of physical memory and running Arch Linux
as operating system. The total execution time of simulations is $38.76$
seconds and the total consumed memory is $66$MB. Table~\ref{Table:3}, shows
all the parameters we used in order to produce our results (see Results). 
No any other implementation of this model found by the authors in order 
to make any further comparison with.  
%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ll}
        \thickhline
        \multicolumn{2}{c}{Model Summary} \\\thickhline
        \rowcolor{Gray}
        Populations  & No population -- one neuron model \\\rowcolor{LightGray}
        Topology     & -- \\ \rowcolor{Gray}
        Connectivity & -- \\ \rowcolor{LightGray}
        Neuron Model & Hodgkin-Huxley conductance-based \\\rowcolor{Gray}
        Channel Models & \\ \rowcolor{LightGray}
        Synapse Model & -- \\ \rowcolor{Gray}
        Plasticity & -- \\ \rowcolor{LightGray}
        Input & Constant current or periodic rectangular pulses \\\rowcolor{Gray}
        Measurements & Membrane potential, channels activation, phase plane \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Summary of the model}} 
    \label{Table:1}
\end{table}
 %%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ccc}
        \thickhline
        \multicolumn{3}{c}{Simulation Time} \\ \thickhline
        Figure & Simulation Time ($ms$) & Integration Step ($ms$) \\ \rowcolor{LightGray}
        $1$ & $6000$ & $0.05$ \\ \rowcolor{Gray}
        $2$ & $2000$ & $0.05$ \\ \rowcolor{LightGray} 
        $3$ & $6000$ & $0.05$  \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Time.}}
    \label{Table:2}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{llcccc}
        \thickhline
        \multicolumn{6}{c}{Input} \\\thickhline
        Figure  & Type & Form & \parbox[t]{1.5cm}{Frequency
            ($\frac{1}{P_0}$,$Hz$)} & 
            \parbox[t]{1.5cm}{Duration ($p$, $ms$)} &
            \parbox[t]{1.5cm}{Amplitude ($\mu A / cm^2$)} \\
        \thickhline \rowcolor{LightGray}
        Figure~$1$ & Periodic & 
            \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} &
            \parbox[t]{1.1cm}{$5$, $8$, $10$} &
            \parbox[t]{1.1cm}{$10$, $40$, $75$, $120$} &
            \parbox[c]{0.3cm}{$-1.0$}
            \\\rowcolor{Gray}
        Figure~$2$ & Constant  & 
        \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
        -- & -- & \parbox[c]{0.3cm}{$+3.0$ \\ $0.0$ $-0.5$ $-0.55$ $-0.6$ $-0.8$
        $-1.3$ $-2.1$} 
        \\\rowcolor{LightGray}
        Figure~$3$ & Constant  & 
            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
            -- & -- & \parbox[c]{0.3cm}{$-0.95$} \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the applied current
        $I_{\text{app}}$ }}
    \label{Table:3}
\end{table}
%%

\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{1.5cm}ll}
        \thickhline
        \multicolumn{2}{c}{Neuron Model} \\\thickhline
        \rowcolor{Gray}
        Name  &  Thalamocortical relay neuron  \\ \rowcolor{LightGray}
        Type  &  Conductance-based neuron  \\ \rowcolor{Gray}
        Membrane Potential & $
            \begin{aligned}
                C_m \frac{dV(t)}{dt} &= -I_T - I_h - I_{Na}
                  - I_K - I_{Na(P)} - I_L + I_{\text{app}}
              \end{aligned}$ \\ \rowcolor{LightGray}
        T-Type Calcium Current ($I_T$) & $
            \begin{aligned}
                I_T &= g_T \cdot s^3_{\infty}(V)\cdot h \cdot (V-V_{Ca}) \\
                s_{\infty}(V) &= \frac{1}{1 + \exp(-\frac{V+65}{7.8})} \\
                \frac{dh(t)}{dt} &= \phi_h \frac{h_{\infty}(V) - h}{\tau_h(V)} \\
                h_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V-\theta_h}{k_h})} \\
                \tau_h(V) &= h_{\infty}\exp(\frac{V+162.3}{17.8}) + 20
            \end{aligned} $
            \\ \rowcolor{Gray}
        Sag Current ($I_h$) & $
            \begin{aligned}
                I_h &= g_h \cdot H^2 \cdot (V-V_h) \\
                H_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V+69}{7.1})} \\
                \frac{dH(t)}{dt} &= \phi_H \frac{H_{\infty}(V) - H}{\tau_H(V)} 
            \end{aligned} $
            \\ \rowcolor{LightGray} 
        Hodgkin-Huxley Currents ($I_K$) and ($I_{Na}$) & $
            \begin{aligned}
                I_{K} &= g_K \cdot n^4 \cdot (V - V_K) \\
                \frac{dn(t)}{dt} &= \phi_n \frac{n_{\infty}(V) - n(t)}{\tau_n(V)} \\
                n_{\infty}(\sigma_K, V) &= \frac{\alpha_n(\sigma_K, V)}
                        {\alpha_n(\sigma_K, V) + \beta_n(\sigma_K, V)} \\
                \tau_n(\sigma_K, V) &= \frac{1}{\alpha_n(\sigma_K, V)
                        + \beta_n(\sigma_K, V)} \\
                \alpha_n(\sigma_K, V) &= \frac{-0.01 (V + 45.7 - \sigma_K)}
                {\exp(-0.1(V + 45.7 - \sigma_K)) - 1} \\ 
                \beta_n(\sigma_K, V) &= 0.125 \exp(-\frac{V + 55.7 -
                \sigma_K}{80}) \\
                I_{Na} &= g_{Na} \cdot m^3_{\infty}(\sigma_{Na}, V)
                    \cdot (0.85 - n) \cdot (V - V_{Na}) \\
                    m_{\infty}(V) &= \frac{\alpha_m(\sigma_{Na}, V)}
                        {\alpha_m(\sigma_{Na}, V) + \beta_m(\sigma_{Na}, V)} \\
                \alpha_m(\sigma_{Na}, V) &= -0.1 \frac{V + 29.7 - \sigma_{Na}}
                    {\exp(-0.1(V + 54.7 - \sigma{Na})) - 1} \\
                \beta_m(\sigma_{Na}, V) &= 4\exp(-\frac{V+54.7-\sigma_{Na}}{18})
            \end{aligned} $ \\ \rowcolor{Gray} 
        Persistent Sodium Currents ($I_{Na(P)}$) & $
                \begin{aligned}
                    I_{Na(P)} &= g_{Na(P)} \cdot m^3_{\infty}(\sigma_{Na(P)},
                    V) \cdot (V - V_{Na}) \\
                \end{aligned} $
            \\ \rowcolor{LightGray}
        Leak Current ($I_L$) &  $
            \begin{aligned}
                I_L &= g_L \cdot (V - V_L) \\
            \end{aligned} $ \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the neuron model}} 
    \label{Table:4}
\end{table}

\begin{landscape}
\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{1.5cm}lllll}
        \thickhline
        \multicolumn{5}{c}{Model Parameters} \\\thickhline \rowcolor{Gray}
        Currents Type  & Common &  Figure $1$ & Figure $2$ & Figure $3$ \\
        \thickhline \rowcolor{LightGray}
        Membrane Potential & $
            \begin{aligned}
                C_m &= 1 \mu F/cm^2 \\
                V_0 &= -74mV
            \end{aligned} $ & & & \\\rowcolor{Gray}
        T-Type Calcium Current & $
            \begin{aligned}
                \phi_h &= 2 \\
                V_{Ca} &= 120mV
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 1 mS/cm^2 \\
                \theta_h &= -81 mV \\
                k_h &= 6.25 mV^{-1}
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 0.3 mS/cm^2\\
                \theta_h &= -79 mV \\
                k_h &= 5 mV^{-1}
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 0.3 mS/cm^2\\
                \theta_h &= -75 mV \\
                k_h &= 5 mV^{-1}
            \end{aligned} $ \\\rowcolor{LightGray}
        Sag Current & $
            \begin{aligned}
                \phi_H &= 1 \\
                g_h &= 0.04 mS/cm^2 \\
                V_h &= -40 mV
            \end{aligned} $
        & & & \\\rowcolor{Gray}
        Hodgkin-Huxley Currents & $
            \begin{aligned}[l]
                g_K &= 30 mS/cm^2 \\
                V_K &= -80 mV \\
                \phi_n &= 28.5\\
                \sigma_K &= 10 \\
            \end{aligned} 
            \begin{aligned}[r]
                g_{Ca} &= 42 mS/cm^2 \\
                V_{Ca} &= 55 mV \\
            \end{aligned} $
            & $\sigma_{Na} = 3$ & $\sigma_{Na} = 6$ & $\sigma_{Na} = 6$
            \\\rowcolor{LightGray}
        Persistent Sodium Currents & $
            \begin{aligned}
                V_{Na(P)} &= 55 mV \\
                \sigma_{Na(P)} &= -5
            \end{aligned} $
            & $g_{Na(P)} = 9 mS/cm^2$ & $g_{Na(P)} = 9 mS/cm^2$ &
                $g_{Na(P)} = 9mS/cm^2$ 
            \\\rowcolor{Gray}
        Leak Current & & $
            \begin{aligned}
                g_L &= 0.1 mS/cm^2\\
                V_L &= -72 mV
            \end{aligned} $ & $
            \begin{aligned}
                g_L &= 0.12 mS/cm^2\\
                V_L &= -70 mV
            \end{aligned} $ & $
            \begin{aligned}
                g_L &= 0.08 mS/cm^2\\
                V_L &= -70 mV
            \end{aligned} $ \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Parameters}} 
    \label{Table:5}
\end{table}
\end{landscape}

\section{Results}\label{results}

We simulated the model described in Table~\ref{Table:4} using the parameters
given in Table~\ref{Table:5} and the corresponding input (see
Table~\ref{Table:3} for three different cases. First, we verified 
that the model is able to generate different types of burst responses such
as in figure $1$ of the original article. We did not reproduce the entire 
diagram, instead we picked up five different cases from the original article
and we tested our implementation on these cases. Thus we applied a periodic 
current pulse of $-1 \mu A/cm^2$ amplitude at three different frequencies
(in the original article frequency is referred as $\frac{1}{P_0}$) $5Hz$, $8Hz$
and $10Hz$. The pulse is on for $p ms$ and the ratio $\frac{p}{P_0}$ is
defined, accordingly.
We chose the frequency and the on-duration of the pulse according to figure $1$
of the original article and such that we have a different type of burst response
for each ratio. The results are shown in Figure~\ref{Fig:1}. It is apparent that
the reference implementation can achieve the same behavior as the original
implementation. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{figs/Figure1.pdf}
    \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
        We stimulate the model with a periodic pulse with frequency
        $\frac{1}{P_0}$ and we record the response of the model. All the
        parameters for this simulation are given in Table~\ref{Table:3}. 
        {\bfseries \sffamily Blue} indicates a frequency $\frac{1}{P_0} = 10Hz$
        and a pulse duration $p = 10 ms$. A resting state activity is 
        generated.
        {\bfseries \sffamily Black} color shows the case where $\frac{1}{P_0} =
        10 Hz$ and $p = 40 ms$. In this case we get a single spike
        periodically. 
        {\bfseries \sffamily Red} color demonstrates a burst with two spikes,
        when the frequency of the periodic pulse is taken to be $\frac{1}{P_0}
        = 5 Hz$ and $p = 40 ms$. 
        {\bfseries \sffamily Cyan} curve shows burst responses consist in 
        three spikes. In this case $\frac{1}{P_0} = 8 Hz$ and $p = 75ms$. 
        {\bfseries \sffamily Magenta} curve indicates a burst periodical 
        response of four spikes per period. }
    \label{Fig:1}
\end{figure}
%%
In a second case, we used a steady current varying only its amplitude and 
keeping the rest parameters fixed. Hence, we used $8$ different current
amplitudes ($3, 0, -0.5, -0.55, -0.6, -0.8, -1.3$ and $-2.1 \mu A/cm^2$) and
Figure~\ref{Fig:2} illustrates the results of our simulations, which
correspond to figure $3$ of \cite{wang:1994}. Again original results
are reproducible and the model is capable of generating repetitive firing,
resting state, subthreshold oscillation, $10 Hz$ bursting, $3Hz$ bursting
and hyperpolarized steady activity. However, the external current amplitudes
used in this case are slightly different from the ones used in the original
article. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.8\textwidth]{figs/Figure2.pdf}
    \caption{{ \bfseries \sffamily Dynamic behavior of neuron.} Here we
    simulate the model with parameters given in Table~\ref{Table:3}. The
    only parameter that changes from panel to panel in this figure is the
    external current $I_{\text{app}}$. The jet colormap indicates the different
    current values. We observe that we get quite similar results as in the original
    article \cite{wang:1994}. For isntance, we obtain, repetitive spiking when
    we apply a constant current at $3 \mu A / cm^2$, or a periodic bursting 
    response, when we apply an external current of $-1.3 \mu A /cm^2$. These
    results correspond to Figure~$3$ of \cite{wang:1994}.}
    \label{Fig:2}
\end{figure}
%%
The last case we simulated is the one that corresponds to figure $6$ of 
\cite{wang:1994}. In this test-case the neuron receives a constant
external current at $-0.95 \frac{\mu A}{cm^2}$ and as Figure~\ref{Fig:3}
points out ``spiral chaos'' is generated as in the original article.
In addition, our results are quantitatively and qualitatively comparable
with the ones provided by \cite{wang:1994} pinpointing that the original
model is fully reproducible apart from the point that we had to properly
tune the amplitude of the external current. The parameters for this simulation
are given in Table~\ref{Table:4}. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{figs/Figure3.pdf}
    \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
    we show that the model is capable of generating spiral chaos as
    it has been shown in Figure~$6$ \cite{wang:1994}. Again all the 
    parameters for this simulation are given in Table~\ref{Table:3}.
    The external current in this simulation is constant and we have 
    traced the phase portrait of the membrane potential and of the
    Sag current ($V$ vs $h$). The top panel illustrates the phase 
    portrait, the middle panel shows the membrane potential ($V$)
    and the bottom panel depicts the $h$ current over time. Our 
    results are almost identical to the original ones rendering 
    the model fully reproducible. }
    \label{Fig:3}
\end{figure}
%%

\section{Conclusion}\label{conclusion}

A conductance-based model for relay thalamocortical neurons proposed 
by \cite{wang:1994} was implemented in Python. After a minor modification
of the external current amplitude we were able to reproduce the results 
presented in \cite{wang:1994}. The original model is easy to implement
since all the equations, parameters and simulation details are given 
in \cite{wang:1994}. All replicated figures are pretty much similar to
the original ones rendering the new results quantitatively and qualitatively 
comparable to the original ones.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2016-05-06 code/neuron_model.py

@@ -1,0 +1,298 @@
+# -*- coding: utf-8 -*-
+# 
+# neuron_model:
+# This is a reference implementation of: "Multiple dynamical modes
+# of thalamic relay neurons: rhythmic bursting and intermittent
+# phase-locking", Wang, X-J, Neuroscience, 59(1), pg.21-31, 1994.
+#
+# Copyright (C) 2016 Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program.  If not, see <http://www.gnu.org/licenses/>.
+#
+import numpy as np
+from scipy.integrate import ode
+
+
+def loadParameters(fname):
+    """ Load all the necessary paremeters from a file.
+
+        |  :param fname: File name containing all parameters
+        |  :return: A dictionaty with all the parameters
+    """
+    import os
+    if os.path.isfile(fname) is False:
+        print('File {} does not exist!'.format(fname))
+    else:
+        import configparser
+
+        config = configparser.RawConfigParser()
+        config.read(fname)
+
+        p = {
+            'V0': config.getfloat('Membrane', 'V0'),
+            'phi_h': config.getfloat('Ttype', 'phi_h'),
+            'gT': config.getfloat('Ttype', 'gT'),
+            'VCa': config.getfloat('Ttype', 'VCa'),
+            'theta_h': config.getfloat('Ttype', 'theta_h'),
+            'k_h': config.getfloat('Ttype', 'k_h'),
+            'phi_H': config.getfloat('Sag', 'phi_H'),
+            'gH': config.getfloat('Sag', 'gH'),
+            'VH': config.getfloat('Sag', 'VH'),
+            'phi_K': config.getfloat('Potasium', 'phi_K'),
[…]

### 2016-05-06 code/plot_figures.py

@@ -1,0 +1,182 @@
+# -*- coding: utf-8 -*-
+# plot_figures: Plots figures for the reference implementation of
+# [1]
+# [1]: "Multiple dynamical modes of thalamic relay neurons: rhythmic
+# bursting and intermittent phase-locking", Wang, X-J, Neuroscience,
+# 59(1), pg.21-31, 1994.
+#
+# Copyright (C) 2016 Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program.  If not, see <http://www.gnu.org/licenses/>.
+#
+import numpy as np
+import matplotlib.cm as cmx
+import matplotlib.pylab as plt
+import matplotlib.colors as colors
+from matplotlib import collections  as mc
+
+
+def plot_figure1():
+    """ Plots Figure 1. See text for more details about Figure 1.
+    """
+    base = "../data/"
+
+    N, M = 10000, 20000         # Time window
+    v, t = [], []               # Voltage and time lists
+    # Load the data
+    for i in range(5):
+        v.append(np.load(base+"Fig1_V"+str(i)+".npy")[N:M])
+        t.append(np.load(base+"Fig1_T"+str(i)+".npy")[N:M])
+
+    # Plot the data
+    color = ['b', 'k', 'r', 'c', 'm']
+    fig = plt.figure(figsize=(10, 10))
+    fig.subplots_adjust(wspace=0.5, hspace=0.5)
+    for i in range(1, 6):
+        ax = fig.add_subplot(5, 1, i)
+        ax.plot(t[i-1], v[i-1][:, 0], color=color[i-1], lw=2.0)
+        ax.set_ylim([-100, 25])
+        if i == 3:
[…]

### 2016-05-06 code/run_all.py

@@ -1,0 +1,27 @@
+# -*- coding: utf-8 -*-
+# run_all - Runs all the simulations and generates all the figures
+# for the reference implementation of [1].
+# [1] "Multiple dynamical modes of thalamic relay neurons: rhythmic
+# bursting and intermittent phase-locking", Wang, X-J, Neuroscience,
+# 59(1), pg.21-31, 1994.
+#
+# Copyright (C) 2016 Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program.  If not, see <http://www.gnu.org/licenses/>.
+#
+import subprocess
+
+if __name__ == '__main__':
+    subprocess.run(["python", "neuron_model.py"])
+    subprocess.run(["python", "plot_figures.py"])

### 2016-06-14 article/article.tex
Revised version

--- 2016-02-17 article/article.tex
@@
 {
-      Owen Petchey,
-      Marco Plebani,
-      Frank Pennekamp,
+      Georgios Detorakis,
   },
@@
 {
-[Re] Chaos in a long-term experiment with a plankton community
+[Re] Multiple dynamical modes of thalamic relay neurons: rhythmic bursting
+        and intermittent phase-locking
 },
@@
 \usepackage{listings}
+\usepackage{float,lscape}
+\usepackage{xcolor,colortbl}
+\usepackage{amssymb,amsmath,array,amsthm}
+\usepackage{marvosym}
+\usepackage{array}
+
+\makeatletter
+\newcommand{\thickhline}{%
+    \noalign {\ifnum 0=`}\fi \hrule height 1pt
+    \futurelet \reserved@a \@xhline
+}
+\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
+\makeatother
+
+\definecolor{Gray}{gray}{0.75}
+\definecolor{LightGray}{gray}{0.95}
 \lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}
@@
 ]{biblatex}
-\bibliography{article.bib}
+\bibliography{detorakis-2016.bib}
 
@@
 % --- Document ----------------------------------------------------------------
-\title{[Re] Chaos in a long-term experiment with a plankton community}
+\title{[Re] Multiple dynamical modes of thalamic relay neurons: 
+        rhythmic bursting and intermittent phase-locking}
 
     \usepackage{authblk}
-                        \author[1]{Owen Petchey}
-                    \author[1]{Marco Plebani}
-                    \author[1]{Frank Pennekamp}
-                            \affil[1]{Institute of Evolutionary Biology and Environmental Studies, University
-of Zurich, Zurich, Switzerland}
+                        \author[1]{Georgios Detorakis}
+                        \affil[1]{ Department of Cognitive Sciences, UC Irvine, CA, USA}
 
 \date{\vspace{-5mm}
-      \sffamily \small \href{mailto:owen.petchey@ieu.uzh.ch}{owen.petchey@ieu.uzh.ch}}
+    \sffamily \small \href{mailto:gdetorak@uci.edu}{gdetorak@uci.edu}}
 
@@
 \begin{itemize}
-    \item[→] Benincà, E., Huisman, J., Heerkloss, R., Jöhnk, K.D., Branco, P., Van
-Nes, E.H., Scheffer, M. \& Ellner, S.P. (2008) Chaos in a long-term
-experiment with a plankton community. Nature, 451, 822--825 DOI:
-10.1038/nature06512
+    \item[→] \emph{Multiple dynamical modes of thalamic relay neurons: rhythmic
+        bursting and intermittent phase-locking}, Wang, X-J, Neuroscience,
+        59(1), pg. 21--31, 1994.
   \end{itemize}\par
@@
 
-\section{Draft article}\label{draft-article}
-
 \section{Introduction}\label{introduction}
 
-The original paper describes analyses of fluctuations in the abundance
-of organisms in a plankton community derived from the Baltic Sea, housed
-in a laboratory environment. The length of the time series (samples
-every few days for 2,300 days) allowed for analyses revealing that the
-observed dynamics exhibited characteristics consistent with chaos
-produced by non-linear species interactions. The article concludes that
-stability is not required for persistence of complex food webs, and that
-long-term prediction of abundances may be fundamentally impossible. The
-demonstration of chaotic dynamics and limited forecast horizons (sensu
-\textcite{Petchey2015}) is important in the field of ecology, since the
-ability to predict dynamics is an open question with considerable
-applied importance \textcite{Petchey2015} \textcite{Mouquet2015}.
+This work introduces a reference implementation of a neuron model 
+for thalamocortical relay neurons, \cite{wang:1994} proposed by X-J Wang.
+The model is conductance-based and takes advantage of an interplay between 
+a T-type calcium current and a non-specific cation sag current and thus it
+is able to generate spindle and delta rhythms. Another feature of this model
+is the presence of an intermittent phase-locking phenomenon
+where action potentials of sodium take place in a non-periodic manner, despite
+the fact that they are phase-locked to the periodic input current. Finally,
+the model is capable of generating tonic spike patterns. The reference 
+implementation in this work, does not aim to reproduce all the results of 
+the original article, but some of them in order to test if the model can be
+easily reproduced and at what degree the original results are reproducible.
+The source code of reference implementation is written in Python (Numpy,
+Scipy, Matplotlib and Scikit-image).
 
@@
 
-This reproduction started with the raw data (source given below) and
-used information from the original paper, the supplementary information
-\href{http://www.nature.com/nature/journal/v451/n7180/extref/nature06512-s1.pdf}{the
-Supplement to the Nature paper}, and communications with Elisa Benincà,
-who provided for comparison digitised data from the original article,
-and Stephen Ellner, who provided code and data used to produce results
-in the original paper. Use of this data and code in this reproduction is
-indicated below.
+The model is described in this section following the paradigm of 
+Nordlie et al, \cite{nordlie:2009}. Therefore, a brief description of
+the model, equations, parameters and inputs are given in the form of 
+tables.
 
-\subsection{Scope of the reproduction}\label{scope-of-the-reproduction}
+Table~\ref{Table:1} provides a description of the model and Table~\ref{Table:2}
+gives a glimpse of the input used during simulations. Table~\ref{Table:3}
+introduces the equations of the model. The neuron model is conductance-based
+consisting in four differential equations describing the dynamics of membrane
+potential and the kinetics of a T-Type calcium current, a Sag current channel
+and a Potassium channel. The rest currents are described by algebraic
+equations. Finally, all simulations parameters are given in
+Table~\ref{Table:5} and the simulations times can be found in
+Table~\ref{Table:2}.
 
-An attempt was made to reproduce the majority of the results in the
-original article. Instances where we did not attempt to reproduce a
-result are detailed below.
+The reference implementation has been done in a Python class (Python $3.5.1$)
+along with Numpy (version $1.10.4$), Scipy (version $0.17.0$) and Matplotlib
+($1.5.1$). The numerical integration has been done by the \emph{ode} method 
+of Scipy \emph{integrate} package taking advantage of the \emph{dopri5} 
+method, which is quite close to the one used by the author of the original 
+article (the author has numerically integrated the system of the four ODEs
+by using a fifth-order adaptive size Runge-Kutta method). In addition, we 
+provide in our implementation the choice of using \emph{BDF} or \emph{Adams}
+multistep methods \cite{ascher:1998} (the choice of \emph{BDF} or \emph{Adams}
+does not alter the numerical results but it's much faster than \emph{dopri5}).
 
-\subsection{The data}\label{the-data}
+All the simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
+generation i$7$ processor, $8$GB of physical memory and running Arch Linux
+as operating system. The total execution time of simulations is $38.76$
+seconds and the total consumed memory is $66$MB (if one choose to run 
+the code for the parameters diagrams, the total execution time elevates 
+to $3$ to $4$ days depending on the machine). Table~\ref{Table:3}, shows
+all the parameters we used in order to produce our results (see Results). 
+No any other implementation of this model found by the authors in order 
+to make any further comparison with.  
+%%
+\begin{table}[!htbp]
+    \centering
+    \begin{tabular}{ll}
+        \thickhline
+        \multicolumn{2}{c}{Model Summary} \\\thickhline
+        \rowcolor{Gray}
+        Populations  & No population -- one neuron model \\\rowcolor{LightGray}
+        Topology     & -- \\ \rowcolor{Gray}
+        Connectivity & -- \\ \rowcolor{LightGray}
+        Neuron Model & Hodgkin-Huxley conductance-based \\\rowcolor{Gray}
+        Channel Models & \\ \rowcolor{LightGray}
+        Synapse Model & -- \\ \rowcolor{Gray}
+        Plasticity & -- \\ \rowcolor{LightGray}
+        Input & Constant current or periodic rectangular pulses \\\rowcolor{Gray}
+        Measurements & Membrane potential, channels activation, phase plane \\
+        \thickhline
+    \end{tabular}
+    \caption{{\bfseries \sffamily Summary of the model}} 
+    \label{Table:1}
+\end{table}
+ %%
 
-The data are available as an Excel file supplement to
-\href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
-Ecology Letters publication} \textcite{Beninca2009}. The Excel file
-contains several datasheets. Two were particularly important, as they
-were the source of the raw data (one contains original species
-abundances, the other the nutrient concentrations). We saved these two
-datasheets, without any alteration, as comma separated value (csv) text
-files. In the code associated with this reproduction, these data files
-are read from the associated github repository. Another datasheet
-contained transformed data (we also saved this as csv file, in order to
-use it in this reproduction). We also received a dataset direct from
-Steve Ellner, see below for details.
+%%
+\begin{table}[!htbp]
+    \centering
+    \begin{tabular}{ccc}
+        \thickhline
+        \multicolumn{3}{c}{Simulation Time} \\ \thickhline
+        Figure & Simulation Time ($s$) & Integration Step ($ms$) \\ \rowcolor{LightGray}
+        $1$ & $2$, $5\times period$ & $0.05$ \\ \rowcolor{Gray}
+        $2$ & $2$ & $0.05$ \\ \rowcolor{LightGray} 
+        $3$ & $8$ & $0.05$  \\ \rowcolor{Gray} 
+        $4$ & $2.5$ & $0.05$  \\ \rowcolor{LightGray}  
+        $5$ & $2$ & $0.05$  \\ \thickhline
+    \end{tabular}
+    \caption{{\bfseries \sffamily Simulations Time.}}
+    \label{Table:2}
+\end{table}
+%%
 
-The original abundance dataset contains a duplicated date: 28/10/1996
-(row 709 and 710 in excel sheet). We changed this (in R) to 26/10/1996,
-making the observation occur half way between the two surrounding dates.
-The original Protozoa variable in the original dataset contains some
-numbers with a comma (instead of period) as the decimal separator; these
-were changed to periods.
+%%
+\begin{table}[!htbp]
+    \centering
+    \begin{tabular}{llcccc}
+        \thickhline
+        \multicolumn{6}{c}{Input} \\\thickhline
+        Figure  & Type & Form & \parbox[t]{1.5cm}{Frequency
+            ($\frac{1}{P_0}$,$Hz$)} & 
+            \parbox[t]{1.5cm}{Duration ($p$, $ms$)} &
+            \parbox[t]{1.5cm}{Amplitude ($\mu A / cm^2$)} \\
+        \thickhline \rowcolor{LightGray}
+        Figure~$1$ & Periodic & 
+            \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} &
+            \parbox[t]{1.1cm}{$5$, $10$} &
+            \parbox[t]{1.1cm}{$10$, $40$} &
+            \parbox[c]{0.3cm}{$-1.0$}
+            \\\rowcolor{Gray}
+        Figure~$2$ & Constant  & 
+        \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+        -- & -- & \parbox[c]{0.3cm}{$+3.0$ \\ $0.0$ $-0.5$ $-0.55$ $-0.6$ $-0.8$
+        $-1.3$ $-2.1$} 
+        \\\rowcolor{LightGray}
+        Figure~$3$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$-0.95$}
+        \\\rowcolor{Gray}
+        Figure~$4$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$[-2,0]$}
+        \\\rowcolor{LightGray}
+        Figure~$5$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$[-2,0]$} \\
+        \thickhline
+    \end{tabular}
+    \caption{{\bfseries \sffamily Description of the applied current
+        $I_{\text{app}}$ }}
+    \label{Table:3}
+\end{table}
+%%
 
-\subsection{Reproduction environment}\label{reproduction-environment}
+\begin{table}[!htbp]
+    \centering
+    \begin{tabular}{p{1.5cm}ll}
+        \thickhline
+        \multicolumn{2}{c}{Neuron Model} \\\thickhline
+        \rowcolor{Gray}
+        Name  &  Thalamocortical relay neuron  \\ \rowcolor{LightGray}
+        Type  &  Conductance-based neuron  \\ \rowcolor{Gray}
+        Membrane Potential & $
+            \begin{aligned}
+                C_m \frac{dV(t)}{dt} &= -I_T - I_h - I_{Na}
+                  - I_K - I_{Na(P)} - I_L + I_{\text{app}}
+              \end{aligned}$ \\ \rowcolor{LightGray}
+        T-Type Calcium Current ($I_T$) & $
+            \begin{aligned}
+                I_T &= g_T \cdot s^3_{\infty}(V)\cdot h \cdot (V-V_{Ca}) \\
+                s_{\infty}(V) &= \frac{1}{1 + \exp(-\frac{V+65}{7.8})} \\
+                \frac{dh(t)}{dt} &= \phi_h \frac{h_{\infty}(V) - h}{\tau_h(V)} \\
+                h_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V-\theta_h}{k_h})} \\
+                \tau_h(V) &= h_{\infty}\exp(\frac{V+162.3}{17.8}) + 20
+            \end{aligned} $
+            \\ \rowcolor{Gray}
+        Sag Current ($I_h$) & $
+            \begin{aligned}
+                I_h &= g_h \cdot H^2 \cdot (V-V_h) \\
+                H_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V+69}{7.1})} \\
+                \frac{dH(t)}{dt} &= \phi_H \frac{H_{\infty}(V) - H}{\tau_H(V)} 
+            \end{aligned} $
+            \\ \rowcolor{LightGray} 
+        Hodgkin-Huxley Currents ($I_K$) and ($I_{Na}$) & $
+            \begin{aligned}
+                I_{K} &= g_K \cdot n^4 \cdot (V - V_K) \\
+                \frac{dn(t)}{dt} &= \phi_n \frac{n_{\infty}(V) - n(t)}{\tau_n(V)} \\
+                n_{\infty}(\sigma_K, V) &= \frac{\alpha_n(\sigma_K, V)}
+                        {\alpha_n(\sigma_K, V) + \beta_n(\sigma_K, V)} \\
+                \tau_n(\sigma_K, V) &= \frac{1}{\alpha_n(\sigma_K, V)
+                        + \beta_n(\sigma_K, V)} \\
+                \alpha_n(\sigma_K, V) &= \frac{-0.01 (V + 45.7 - \sigma_K)}
+                {\exp(-0.1(V + 45.7 - \sigma_K)) - 1} \\ 
+                \beta_n(\sigma_K, V) &= 0.125 \exp(-\frac{V + 55.7 -
+                \sigma_K}{80}) \\
+                I_{Na} &= g_{Na} \cdot m^3_{\infty}(\sigma_{Na}, V)
+                    \cdot (0.85 - n) \cdot (V - V_{Na}) \\
+                    m_{\infty}(V) &= \frac{\alpha_m(\sigma_{Na}, V)}
+                        {\alpha_m(\sigma_{Na}, V) + \beta_m(\sigma_{Na}, V)} \\
+                \alpha_m(\sigma_{Na}, V) &= -0.1 \frac{V + 29.7 - \sigma_{Na}}
+                    {\exp(-0.1(V + 54.7 - \sigma{Na})) - 1} \\
+                \beta_m(\sigma_{Na}, V) &= 4\exp(-\frac{V+54.7-\sigma_{Na}}{18})
+            \end{aligned} $ \\ \rowcolor{Gray} 
+        Persistent Sodium Currents ($I_{Na(P)}$) & $
+                \begin{aligned}
+                    I_{Na(P)} &= g_{Na(P)} \cdot m^3_{\infty}(\sigma_{Na(P)},
+                    V) \cdot (V - V_{Na}) \\
+                \end{aligned} $
+            \\ \rowcolor{LightGray}
+        Leak Current ($I_L$) &  $
+            \begin{aligned}
+                I_L &= g_L \cdot (V - V_L) \\
+            \end{aligned} $ \\ \thickhline
+    \end{tabular}
+    \caption{{\bfseries \sffamily Description of the neuron model}} 
+    \label{Table:4}
+\end{table}
 
-The R language and environment for statistical computing and graphics
-was used to make the reproduction. Additional R packages required are
-specified in the code associated with this reproduction. The code for
-making the analyses and figures presented in this reproduction resides
-in an R markdown document, as well as a source file containing some
-required functions. Some of the code takes several minutes to run, so an
-intermediate data file is provided with results from this code.
+\begin{landscape}
+\begin{table}[!htbp]
+    \centering
+    \begin{tabular}{p{1.5cm}lllllll}
+        \thickhline
+        \multicolumn{5}{c}{Model Parameters} \\\thickhline \rowcolor{Gray}
+        Currents Type  & Common &  Figure $1$ & Figure $2$ & Figure $3$ &
+        Figure $4$  & Figure $5$ \\
+        \thickhline \rowcolor{LightGray}
+        Membrane Potential & $
+            \begin{aligned}
+                C_m &= 1 \mu F/cm^2 \\
+                V_0 &= -74mV
+            \end{aligned} $ & & & \\\rowcolor{Gray}
+        T-Type Calcium Current & $
+            \begin{aligned}
+                \phi_h &= 2 \\
+                V_{Ca} &= 120mV
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_T &= 1 mS/cm^2 \\
+                \theta_h &= -81 mV \\
+                k_h &= 6.25 mV^{-1}
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_T &= 0.3 mS/cm^2\\
+                \theta_h &= -79 mV \\
+                k_h &= 5 mV^{-1}
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_T &= 0.3 mS/cm^2\\
+                \theta_h &= -75 mV \\
+                k_h &= 5 mV^{-1}
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_T &= 0.3/0.25/0.2 mS/cm^2\\
+                \theta_h &= -81 mV \\
+                k_h &= 6.25 mV^{-1}
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_T &= 1.0/0.7 mS/cm^2\\
+                \theta_h &= -79 mV \\
+                k_h &= 5 mV^{-1}
+            \end{aligned} $ 
+            \\\rowcolor{LightGray}
+        Sag Current & $
+            \begin{aligned}
+                \phi_H &= 1 \\
+                g_h &= 0.04 mS/cm^2 \\
+                V_h &= -40 mV
+            \end{aligned} $
+        & & & \\\rowcolor{Gray}
+        Hodgkin-Huxley Currents & $
+            \begin{aligned}[l]
+                g_K &= 30 mS/cm^2 \\
+                V_K &= -80 mV \\
+                \phi_n &= 28.5\\
+                \sigma_K &= 10 \\
+            \end{aligned} 
+            \begin{aligned}[r]
+                g_{Ca} &= 42 mS/cm^2 \\
+                V_{Ca} &= 55 mV \\
+            \end{aligned} $
+            & $\sigma_{Na} = 3mV$ & $\sigma_{Na} = 6mV$ & $\sigma_{Na} = 6mV$ &
+            $\sigma_{Na} = 3mV$ & $\sigma_{Na} = 6mV$
+            \\\rowcolor{LightGray}
+        Persistent Sodium Currents & $
+            \begin{aligned}
+                V_{Na(P)} &= 55 mV \\
+                \sigma_{Na(P)} &= -5 \\
+                g_{Na(P)} = 9 mS / cm^2 
+            \end{aligned} $
+            & & & & &  
+            \\\rowcolor{Gray}
+        Leak Current & & $
+            \begin{aligned}
+                g_L &= 0.1 mS/cm^2\\
+                V_L &= -72 mV
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_L &= 0.12 mS/cm^2\\
+                V_L &= -70 mV
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_L &= 0.08 mS/cm^2\\
+                V_L &= -70 mV
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_L &= 0.08 mS/cm^2\\
+                V_L &= -70 mV 
+            \end{aligned} $ & $
+            \begin{aligned}
+                g_L &= 0.12/0.04 mS/cm^2\\
+                V_L &= -70 mV 
+            \end{aligned} $ \\
+        \thickhline
+    \end{tabular}
+    \caption{{\bfseries \sffamily Simulations Parameters}} 
+    \label{Table:5}
+\end{table}
+\end{landscape}
 
@@
 
-\subsection{Population dynamics}\label{population-dynamics}
+We simulated the model described in Table~\ref{Table:4} using the parameters
+given in Table~\ref{Table:5} and the corresponding input (see
+Table~\ref{Table:3} for three different cases). First, we verified 
+that the model is able to generate different types of burst responses such
+as in figure $1$ of the original article. Thus we applied a periodic 
+current pulse of $-1 \mu A/cm^2$ amplitude at several different frequencies
+(in the original article frequency is referred as $\frac{1}{P_0}$) ranging
+from $0.1Hz$ to $15Hz$ with a resolution (discretization) of $300$ points
+(parameters diagram in Figure~\ref{Fig:1} left panel). In addition, we
+picked up three specific frequencies ($\frac{1}{P_0}=10, 10, 5$ and pulse
+duration $p=10, 40, 40ms$ ($\frac{p}{P_0} = 0.1, 0.4, 0.2$). 
+The results are shown in Figure~\ref{Fig:1}, right panels. It is apparent that
+the reference implementation can achieve the same behavior as the original
+implementation. The parameters correspond to the proper behavior of the model.
+%%
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figs/Figure1.pdf}
+    \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
+        We stimulate the model with a periodic pulse with frequency
+        $\frac{1}{P_0}$ and we record the response of the model. All the
+        parameters for this simulation are given in Table~\ref{Table:3}. 
+        {\bfseries \sffamily Left Panel:} shows a range of frequencies
+        $\frac{1}{P_0}$ from $0.1Hz$ to $15Hz$ versus the ratio
+        $\frac{p}{P_0}$. 
+        {\bfseries \sffamily Right Panels:} 
+        {\bfseries \sffamily Red} color demonstrates a case where only
+        subthreshold spikes are generated, when the frequency of the periodic
+        pulse is $\frac{1}{P_0} = 10 Hz$ and $\frac{p}{P_0} = 0.1$. 
+        {\bfseries \sffamily Blue} indicates bursts of one spike, when the
+        frequency is $\frac{1}{P_0} = 10Hz$ and $\frac{p}{P_0} = 0.4$.
+        {\bfseries \sffamily Black} shows a case of two spikes bursts. Here
+        the frequency is $\frac{1}{P_0} = 5 Hz$ and $\frac{p}{P_0} = 0.2$.}
+    \label{Fig:1}
+\end{figure}
+%%
 
-The reproduced populations dynamics were very similar to those in figure
-1b-g of the original publication (figure \ref{fig:dynamics}). As most
-analyses were performed on fourth root transformed values, we plot
-these, rather than raw abundances with a y-axis break as in the original
-article.
-
-\begin{figure}[htbp]
+In a second case, we used a steady current varying only its amplitude and 
+keeping the rest parameters fixed. Hence, we used $8$ different current
+amplitudes ($3, 0, -0.5, -0.55, -0.6, -0.8, -1.3$ and $-2.1 \mu A/cm^2$) and
+Figure~\ref{Fig:2} illustrates the results of our simulations, which
+correspond to figure $3$ of \cite{wang:1994}. Again original results
+are reproducible and the model is capable of generating repetitive firing,
+resting state, subthreshold oscillation, $10 Hz$ bursting, $3Hz$ bursting
+and hyperpolarized steady activity. 
+%%
+\begin{figure}[!htbp]
     \centering
-\includegraphics{figures/obs_pop_dyn.pdf}
-\caption{\label{fig:dynamics}Observed population dynamics.}
+    \includegraphics[width=.8\textwidth]{figs/Figure2.pdf}
+    \caption{{ \bfseries \sffamily Dynamic behavior of neuron.} Here we
+    simulate the model with parameters given in Table~\ref{Table:3}. The
+    only parameter that changes from panel to panel in this figure is the
+    external current $I_{\text{app}}$. The jet colormap indicates the different
+    current values. We observe that we get quite similar results as in the original
+    article \cite{wang:1994}. For instance, we obtain, repetitive spiking when
+    we apply a constant current at $3 \mu A / cm^2$, or a periodic bursting 
+    response, when we apply an external current of $-1.3 \mu A /cm^2$. These
+    results correspond to Figure~$3$ of \cite{wang:1994}.}
+    \label{Fig:2}
 \end{figure}
 
-\subsection{Data transformations}\label{data-transformations}
+%%
+Another test case we simulated is the one that corresponds to figure $6$ of 
+\cite{wang:1994}. In this test-case the neuron receives a constant
+external current at $-0.95 \frac{\mu A}{cm^2}$ and as Figure~\ref{Fig:3}
+points out ``spiral chaos'' is generated as in the original article.
+The parameters for this simulation are given in Table~\ref{Table:4}. 
+%%
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=.7\textwidth]{figs/Figure3.pdf}
+    \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
+    we show that the model is capable of generating spiral chaos as
+    it has been shown in Figure~$6$ \cite{wang:1994}. Again all the 
+    parameters for this simulation are given in Table~\ref{Table:3}.
+    The external current in this simulation is constant and we have 
+    traced the phase portrait of the membrane potential and of the
+    Sag current ($V$ vs $h$). The top panel illustrates the phase 
+    portrait, the middle panel shows the membrane potential ($V$)
+    and the bottom panel depicts the $h$ current over time. Our 
+    results are identical to the original ones rendering 
+    the model fully reproducible. }
+    \label{Fig:3}
+\end{figure}
+%%
 
-The original publication states that long sequences of zeros were
-removed from time series prior to transformation and further analysis.
-We removed zeros by matching the removals to those in the transformed
-version of the data in the published Excel file mentioned above.
-Transformation steps were:
+A fourth test case has to do with the amplitude of the injected external 
+current and its relation to the frequency of the generated spike bursts. 
+Therefore, we captured the data\footnote{The data are available in the
+accompanying github repository of the present article.} values for the 
+external current from Figure~$7$ of the original article (these are the
+values represented by dots and circles in Figure $7$ and on page $27$ of
+the \cite{wang:1994}) using the software
+PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
+Therefore, we run a simulation on the external current of the original 
+article using our implementation and the results are shown in Figure 
+\ref{Fig:4} upper row. Our results are pretty much the same as the ones
+in Figure~$7$ of the original article. In addition, we run one more simulation
+using a much higher resolution (using $200$ values of the injected current
+in the range $0$ to $-2.0 \mu A/cm^2$. It is clear that the more detailed 
+simulation is quantitatively similar to the simpler one but it reveals 
+a more oscillatory behavior in the shape of the parameters curve. If we
+superpose the two panels, we can see that the upper row is an interpolation 
+of some of the data points of the bottom row.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Interpolation to create equally spaced observations in time series.
-\item
-  Fourth root transformation.
-\item
-  Detrending of five of the time series.
-\item
-  Rescaling to zero mean and unit standard deviation.
-\end{enumerate}
+%%
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=.7\textwidth]{figs/Figure4.pdf}
+    \caption{{\bfseries \sffamily Frequency and period versus steady input current.}
+    {\bfseries \sffamily Top Row:} Indicates the external current (ranging from 
+    $0$ to $-2 \mu A/cm^2$ versus the spike bursts frequency in $Hz$. The
+    numerical values for the external current for this simulation have been
+    taken from the original article (see text for more details). 
+    {\bfseries \sffamily Bottom Row:} The same simulation as in the upper row,
+    but with a finer discretization of the external current values range. In
+    these simulations we used $200$ values for the $I_{\text{app}}$.}
+    \label{Fig:4}
+\end{figure}
+%%
 
-The reproduced transformed data closely matched the published
-transformed data (figure \ref{fig:trans_comp1}).
+%%
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=.7\textwidth]{figs/Figure5.pdf}
+    \caption{{\bfseries \sffamily Symbolic patterns.}}
+    \label{Fig:5}
+\end{figure}
+%%
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/zero_rem_trans_comparison.pdf}
-\caption{\label{fig:trans_comp1}Comparison of published transformed data
-with reproduced transformed data. The red line is the 1:1 line.}
-\end{figure}
-
-\subsection{Correlations among species
-abundances}\label{correlations-among-species-abundances}
-
-Correlations among species abundances presented in Table 1 of the
-original article closely matched our reproduced correlations, calculated
-from the transformed data with zeros removed (figure
-\ref{fig:corr_comp}). Deviations between the original and reproduced
-correlations are relatively small (less than 0.072 units) and
-infrequent.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/correlation_comparison.pdf}
-\caption{\label{fig:corr_comp}Comparison of calculated correlations
-among species abundances in the original article and this reproduction.}
-\end{figure}
-
-Highlighted in the text of the original paper were: negative
-correlations of picophytoplankton with protozoa, and of
-nanophytoplankton both with rotifers and calanoid copepods, positive
-correlation of picophytoplankton with calanoid copepods, negative
-correlation between bacteria and ostracods, and positive correlation
-between bacteria and phosphorus. All of these correlations were at least
-qualitatively reproduced.
-
-\subsection{Spectral analyses}\label{spectral-analyses}
-
-Spectral analyses in the original paper were presented graphically in
-figures S3 (raw spectrograms) and S4 (Welch periodograms). The original
-article states: ``fluctuations covered a range of different
-periodicities'', and ``picophytoplankton, rotifers and calanoid copepods
-seemed to fluctuate predominantly with a periodicity of about 30 days.''
-Our reproduced spectra (not shown here, but code provided) were a
-reasonably close match to the spectra in the original article.
-
-\subsection{Lyapunov exponents by direct
-method}\label{lyapunov-exponents-by-direct-method}
-
-Reproduced divergence rates (figure \ref{fig:divergence}) and therefore
-Lyapunov exponents (figure \ref{fig:LE_comparison}) were somewhat
-different from those in the original article. The original article
-states: ``the distance between initially nearby trajectories increased
-over time, and reached a plateau after about 20--30 days''. The
-reproduced results are consistent with this statement. The original
-article also stated that the analyses ``yielded significantly positive
-Lyapunov exponents of strikingly similar value for all species (Fig. 3;
-mean exponent = 0.057 per day, s.d. = 0.005 per day, n = 9)''.
-Reproduced exponents had very similar mean value, but had larger
-standard deviation (mean = 0.060 and s.d. = 0.014).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/div_rate.pdf}
-\caption{\label{fig:divergence}Reproduced divergence rates and Lyapunov
-exponents (figure 3 in the original article).}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/LE_comparison.pdf}
-\caption{\label{fig:LE_comparison}Comparison of original and reproduced
-directly estimated Lyapunov exponents.}
-\end{figure}
-
-\subsection{Lyapunov exponents by indirect
-method}\label{lyapunov-exponents-by-indirect-method}
-
-The data received directly from Stephen Ellner (file
-`interp\_short\_allsystem\_newnames.csv' in the reproduction repository)
-was interpolated, but without zeros removed. The reproduced interpolated
-data, without zeros removed, matched closely this data (figure
-\ref{fig:trans_comp2}).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/repro_ellner_comp.pdf}
-\caption{\label{fig:trans_comp2}Comparison of interpolated data provided
-by Ellner with reproduced interpolated data (no removal of zeros). The
-red line is the 1:1 line.}
-\end{figure}
-
-The original paper reported a global Lyapunov exponent calculated via
-two modelling approaches (neural network and generalised additive models
-{[}GAMs{]}). Only the GAM approach was reproduced, with the assistance
-of code donated by Stephen Ellner. The original article obtained a
-global Lyapunov exponent of λ=0.08 day-1. The reproduced value was 0.04.
-We did not reproduce the bootstrapping used to give confidence intervals
-around this estimate.
-
-\subsection{Predictability decay}\label{predictability-decay}
-
-The article stated: ``For short-term forecasts of only a few days, most
-species had a high predictability of R2 = 0.70 -- 0.90 (Fig. 2).
-However, the predictability of the species was much reduced when
-prediction times were extended to 15--30days.'' The reproduced
-predictabilities, which were calculated from the GAMs, were consistent
-with these qualitative statements, though were quantitatively different
-(\ref{fig:prediction_distance}). We did not reproduce the predictability
-estimates for linear models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figures/prediction_distance.pdf}
-\caption{\label{fig:prediction_distance}Predictability (correlation
-between predicted and observed abundances) and prediction distance
-(days) (figure 2 in the original article). Reproducted data in red, and
-data from original publication in black.}
-\end{figure}
+Finally, we tried to simulate some of the symbolic patterns as they are 
+presented in Table~$1$ of \cite{wang:1994}. Therefore, we run our
+implementation using the same parameters as in \cite{wang:1994} (see 
+Table~\ref{Table:5}). The results are shown in Figure~\ref{Fig:5}, where
+black dots represent zeros ($0$) and red line segments represent aces ($1$).
+Different patterns emerge for different values of the external current. 
+For instance, the shaded area in Figure~\ref{Fig:5} indicates that the
+current is between $0.0$ and $-0.75$ and thus we have only zeros in the
+formed patterns.
 
@@
 
-The reproduced results were qualitatively identical to those in the
-original article, and therefore support the conclusion of chaotic
-dynamics produced by non-linear species interactions. For example, all
-Lyapunov exponents estimated by direct method were positive, as in the
-original article, consistent with chaotic dynamics. Quantitative
-differences may have resulted from difference in algorithms used. For
-example,the original used the
-\href{http://www.mpipks-dresden.mpg.de/~tisean/}{Tisean software} to
-calculate Lyapunov exponents. As this was available from CRAN
-\href{http://cran.r-project.org/web/packages/RTisean/index.html}{until
-mid 2014} and since it is a bit less well integrated with R, we instead
-use the tseriesChaos package (\textcite{tseriesChaos}), which in any
-case was largely inspired by the TISEAN project. In addition, there may
-have been some difference in algorithm parameters, as not all parameters
-required by the functions we used were reported in the original ms.
-
-Considerable quantitative difference in predictability decay
-(\ref{fig:prediction_distance}) remain unexplained, though could result
-from original analyses using neural networks, and the reproduction using
-GAMs.
-
-In conclusion, this reproduction supports the general scientific
-conclusions of the original article, but also shows how difficult can be
-an accurate quantitative reproduction, even in the presence of the
-extensive methodological details provided alongside the original
-article.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-This reproduction was made as part of the Reproducible Research in
-Ecology, Evolution, Behaviour, and Environmental Studies (RREEBES)
-Course, lead by Owen Petchey at the University of Zurich. More
-information about the course
-\href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
-on github. Many thanks to Elisa Benincà and Steve Ellner for assistance,
-and to Steve for providing unpublished data and code.
+A conductance-based model for relay thalamocortical neurons proposed 
+by \cite{wang:1994} was implemented in Python. The model tested 
+thoroughly in several examples taken from the original article and 
+it succeed to reproduce all of them. The original model is easy to implement
+since all the equations, parameters (except the initial time step of the
+integration method) and simulation details are given in \cite{wang:1994}.
+The choice of a different integration scheme does not affect the results,
+but it increases the temporal performance of the model. 
+All replicated figures are pretty much similar to the original ones rendering
+the new results quantitatively and qualitatively comparable to the original
+ones.
 

### 2016-06-14 code/neuron_model.py

@@ -22,6 +22,7 @@
 #
 import numpy as np
 from scipy.integrate import ode
+from skimage.feature import peak_local_max
 
 
 def loadParameters(fname):
@@ -67,8 +68,13 @@ def loadParameters(fname):
 
 
 class WangModel():
-    def __init__(self, fname, frequency, duration, amplitude,
-                 chunks):
+    def __init__(self, fname,
+                 dt=0.05,   
+                 frequency=1,
+                 duration=1000,
+                 inp_type='const',
+                 amplitude=1,
+                 chunks=1):
         """ Constructor method
 
         |  :param fname: File name containing all parameters
@@ -80,10 +86,13 @@ class WangModel():
         """
         self.p = loadParameters(fname)      # Load the parameters
 
-        self.freq = frequency
+        self.dt = dt                        # Integration time step
+        self.freq = frequency               # Pulse current frequency
         self.P0 = (1.0/self.freq)*1000      # Period in ms
         self.P = duration                   # Duration in ms
-        self.Amp = amplitude                # Amplitude in μA/cm²
+        self.amplitude = amplitude          # Amplitude in μA/cm²
+        self.chunks = chunks                # Number of perios (time bins)
+        self.inp_type = inp_type            # Input type (constant or pulse)
 
     def s_inf(self, v):
         """ T-type calcium channel gating kinetics """
@@ -126,7 +135,7 @@ class WangModel():
     def alpha_m(self, v, sigma):
         """ Sodium (Na) channel gating kinetics """
         return (-0.1*(v + 29.7 - sigma)/(np.exp(-0.1 *
-                (v+29.7-sigma)) - 1.))
+                (v + 29.7 - sigma)) - 1.))
 
     def beta_m(self, v, sigma):
         """ Sodium (Na) channel gating kinetics """
@@ -204,10 +213,20 @@ class WangModel():
         """
         return self.p['gL'] * (v - self.p['VL'])
 
-    def I_ext(self, period, duration, amplitude, time, numChunks):
+    def I_ext(self, period, duration, time):
+        """
+        External current (in μA/cm²)
+
+        |  :param period: Period of current on pulses
+        |  :param duration: Duration of current on pulses
+        |  :param amplitude: Pulse amplitude
+        |  :param time: Total simulation time (ms)
[…]

### 2016-06-14 code/plot_figures.py

@@ -24,39 +24,108 @@ import numpy as np
 import matplotlib.cm as cmx
 import matplotlib.pylab as plt
 import matplotlib.colors as colors
+import matplotlib.patches as patches
 from matplotlib import collections  as mc
+from skimage.feature import peak_local_max
 
 
-def plot_figure1():
+def detect_spikes(data, freq=0, dt=0.05, dist=40, th=0):
+    win = int(((1.0 / freq) * 1000) / dt)
+    tmp = data[4*win:5*win, 0]
+    spikes = peak_local_max(tmp, min_distance=dist)
+    spikes = spikes[tmp[spikes] > th]
+    # Check if the model behaves properly
+    if spikes.shape[0] > 4:
+        raise ValueError('Bad number of spikes!')
+    else:
+        return spikes.shape[0]
+
+
+def plot_figure1(diagram=False):
     """ Plots Figure 1. See text for more details about Figure 1.
     """
     base = "../data/"
-
-    N, M = 10000, 20000         # Time window
     v, t = [], []               # Voltage and time lists
     # Load the data
-    for i in range(5):
-        v.append(np.load(base+"Fig1_V"+str(i)+".npy")[N:M])
-        t.append(np.load(base+"Fig1_T"+str(i)+".npy")[N:M])
+    for i in range(3):
+        v.append(np.load(base+"Fig1B_V"+str(i)+".npy")[:, 0])
+        t.append(np.load(base+"Fig1B_T"+str(i)+".npy"))
+
+
+    if diagram is True:
+        fig = plt.figure(figsize=(19.5, 7.5))
+        fig.subplots_adjust(wspace=0.5, hspace=.5)
+    else:
+        fig = plt.figure(figsize=(10, 10))
+
+    if diagram is True:
+        # Plot parameters diagram
+        ax = plt.subplot2grid((3,7), (0, 0), colspan=4, rowspan=3)
+        samples = 150
+        freq = np.linspace(0.1, 15, samples)
+        mat_ = np.empty((samples, samples))
+        idx = 0
+        for i in range(samples):
+            for j in range(samples):
+                data = np.load(base+'Fig1A_V'+str(idx)+'.npy')[0:, :]
+                mat_[i, j] = detect_spikes(data, freq[i])
+                idx += 1
[…]

### 2016-06-14 code/run_all.py

@@ -23,5 +23,5 @@
 import subprocess
 
 if __name__ == '__main__':
-    subprocess.run(["python", "neuron_model.py"])
+    subprocess.run(["python", "simulations.py"])
     subprocess.run(["python", "plot_figures.py"])

### 2016-06-14 code/simulations.py

@@ -1,0 +1,208 @@
+# -*- coding: utf-8 -*-
+# plot_figures: Plots figures for the reference implementation of
+# [1]
+# [1]: "Multiple dynamical modes of thalamic relay neurons: rhythmic
+# bursting and intermittent phase-locking", Wang, X-J, Neuroscience,
+# 59(1), pg.21-31, 1994.
+#
+# Copyright (C) 2016 Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program.  If not, see <http://www.gnu.org/licenses/>.
+#
+from neuron_model import *
+
+
+########################################################################
+# In the following code, if you want to use BDF or Adams methods for
+# temporal integrations, please use: 
+#            simulation.run(sim_time,
+#                           fname_for_mem_pot,
+#                           fname_for_time,
+#                           'vode', 'Adams' [or 'BDF'] ) 
+########################################################################
+def simulate_figure1(diagram=False):
+    print("Running simulation #1")
+
+    if diagram is True:
+        samples = 150
+        freq = np.linspace(0.1, 15, samples)
+        Iext = -1.0                                   
+
+        name = "Fig1A"
+        idx = 0
+        for i, f in enumerate(freq):
+            # period = np.round(1./f, 3) * 1000
+            period = 1./f * 1000
+            dur = np.linspace(0, period, samples)
+            for j, d in enumerate(dur):
+                print('Frequency: {}, Duration: {}'.format(f, d))
+                nm = 'Fig1A_V'+str(idx)
+                print('Fname: {}'.format(nm))
+                simulation = WangModel(base+"params_figure1.cfg",
+                                       frequency=f,
+                                       duration=d,
+                                       inp_type='pulse',
+                                       amplitude=Iext,
+                                       chunks=100)
[…]

### 2016-07-03 article/article.tex

--- 2016-02-23 article/article.tex
@@
 However, the predictability of the species was much reduced when
-prediction times were extended to 15--30days.'' The reproduced
+prediction times were extended to 15--30 days.'' The reproduced
 predictabilities, which were calculated from the GAMs, were consistent
@@
 \href{https://github.com/opetchey/RREEBES/blob/master/README.md}{here}
-on github. Many thanks to Elisa Benincà and Stephen Ellner for
-assistance, and to Stephen Ellner for providing unpublished data and
+on github. Many thanks to Elisa Benincà, Stephen Ellner and Jef Huisman
+for assistance, and to Stephen Ellner for providing unpublished data and
 code.

### 2016-07-03 code/report.Rmd

@@
 tr <- gather(tr, key="Species", value="Abundance", 2:13, convert=F) ## make long format
-#tr$Species <- as.factor(tr$Species)
+tr$Species <- as.factor(tr$Species)
 levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)

### 2016-07-18 article/detorakis-2016.tex
Corrections after second review round

@@
 \usepackage{array}
+\usepackage[export]{adjustbox}
+
+\newcommand{\Rm}[1]{\mathrm{#1}}
 
@@
 
-This work introduces a reference implementation of a neuron model 
-for thalamocortical relay neurons, \cite{wang:1994} proposed by X-J Wang.
+This work introduces a reference implementation of a neuron model for
+thalamocortical relay neurons, proposed by X-J Wang, \cite{wang:1994}.
 The model is conductance-based and takes advantage of an interplay between
-a T-type calcium current and a non-specific cation sag current and thus it
+a T-type calcium current and a non-specific cation sag current and thus, it
 is able to generate spindle and delta rhythms. Another feature of this model
-is the presence of an intermittent phase-locking phenomenon
-where action potentials of sodium take place in a non-periodic manner, despite
-the fact that they are phase-locked to the periodic input current. Finally,
-the model is capable of generating tonic spike patterns. The reference 
-implementation in this work, does not aim to reproduce all the results of 
-the original article, but some of them in order to test if the model can be
-easily reproduced and at what degree the original results are reproducible.
-The source code of reference implementation is written in Python (Numpy,
-Scipy and Matplotlib).
+is the presence of an intermittent phase-locking phenomenon where action 
+potentials of sodium take place in a non-periodic manner, despite the fact
+that they are phase-locked to the periodic input current. Finally, the model
+is capable of generating tonic spike patterns. The source code of
+reference implementation is written in Python (Numpy, Scipy, Matplotlib,
+and Scikit-image).
 
@@
 
-The model is described in this section following the paradigm of 
-Nordlie et al, \cite{nordlie:2009}. Therefore, a brief description of
-the model, equations, parameters and inputs are given in the form of 
-tables.
+In this section, a detailed description of the model is given following 
+the paradigm of Nordlie et al, \cite{nordlie:2009}. Therefore, a brief
+description of the model, equations, parameters, and inputs are given in
+the form of tables.
 
-Table~\ref{Table:1} provides a description of the model and Table~\ref{Table:2}
-gives a glimpse of the input used during simulations. Table~\ref{Table:3}
-introduces the equations of the model. The neuron model is conductance-based
-consisting in four differential equations describing the dynamics of membrane
-potential and the kinetics of a T-Type calcium current, a Sag current channel
-and a Potassium channel. The rest currents are described by algebraic
-equations. Finally, all simulations parameters are given in
-Table~\ref{Table:5} and the simulations times can be found in
-Table~\ref{Table:2}.
-
-The reference implementation has been done in a Python class (Python $3.5.1$)
-along with Numpy (version $1.10.4$), Scipy (version $0.17.0$) and Matplotlib
-($1.5.1$). The numerical integration has been done by the \emph{ode} method 
-of Scipy \emph{integrate} package taking advantage of the BDF multistep method
-\cite{ascher:1998}. This is a diversion from the original implementation, since
-in \cite{wang:1994}, the author has numerically integrated the system of 
-the four ODEs by using a fifth-order adaptive size Runge-Kutta method. This
-could cause some of the numerical deviations we noticed during the
-implementation of the model. The values for the input current must be 
-slightly elevated in order to achieve exactly the same results as
-in the original implementation.
-
-All the simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
-generation i$7$ processor, $8$GB of physical memory and running Arch Linux
-as operating system. The total execution time of simulations is $38.76$
-seconds and the total consumed memory is $66$MB. Table~\ref{Table:3}, shows
-all the parameters we used in order to produce our results (see Results). 
-No any other implementation of this model found by the authors in order 
-to make any further comparison with.  
+Table~\ref{Table:1} provides a description of the model, Table~\ref{Table:2}
+provides information about simulations duration and temporal integration time
+steps, Table~\ref{Table:3} gives a glimpse of the input signals used in this work.
+Table~\ref{Table:4} introduces the equations of the model and finally 
+Table~\ref{Table:5} summarizes all the parameters for each figure (simulation).
+The neuron model is conductance-based consisting of four differential equations
+describing the dynamics of membrane potential and the kinetics of a T-Type
+calcium current, a Sag current channel and a Potassium channel. The rest
+currents are described by algebraic equations.
 %%
@@
  %%  
-
+The reference implementation has been done in a Python class (Python $3.5.1$)
+along with Numpy (version $1.10.4$), Scipy (version $0.17.0$), Matplotlib
+($1.5.1$) and Scikit-image (version $0.12.3$). The numerical integration
+has been done using the \emph{ode} method of Scipy \emph{integrate} package.
+Three different methods have been tested in this work (\emph{dopri5}, 
+\emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{dopri5} is the closest
+numerical method to the one used by the author in the original article (the
+author has numerically integrated the system of the four ODEs by using a
+fifth-order adaptive size Runge-Kutta method). \emph{BDF} and \emph{Adams}
+provide similar numerical results as the first one, but they are
+faster\footnote{The user has the option to choose one of the three methods
+    during class instantiating. We tested all three methods comparing, spike
+    times, amplitudes and the coefficient of variation using as threshold value
+    $0\, \Rm{mV}$. Spike events were overlapping, and $CV \simeq 1.5$ for all
+    three methods. The membrane potential amplitude was sometimes different
+    for the same spike events, however that difference was not too high as
+    Figure~\ref{Fig:1} shows (up to $4\, \Rm{mV}$ at maximum). Thus we decided
+    to use the \emph{Adams} method.}.
 %%
@@
         \multicolumn{3}{c}{Simulation Time} \\ \thickhline
-        Figure & Simulation Time ($ms$) & Integration Step ($ms$) \\ \rowcolor{LightGray}
-        $1$ & $6000$ & $0.05$ \\ \rowcolor{Gray}
-        $2$ & $2000$ & $0.05$ \\ \rowcolor{LightGray} 
-        $3$ & $6000$ & $0.05$  \\ \thickhline
+        Figure & Simulation Time ($\Rm{s}$) & Integration Step ($\Rm{ms}$) \\ 
+        \rowcolor{LightGray}
+        $1$ & $6$ & $0.05$ \\ \rowcolor{Gray}
+        $2$ & $15\times period$, $2$ & $0.05$ \\ \rowcolor{LightGray}
+        $3$ & $6$ & $0.05$ \\ \rowcolor{Gray} 
+        $4$ & $1.5$, $1.5$, $1$ & $0.05$ \\ \rowcolor{LightGray} 
+        $5$ & $5$ & $0.05$  \\ \rowcolor{Gray} 
+        $6$ & $2.5$ & $0.05$  \\ \rowcolor{LightGray} 
+        $7$ & $40$, $20$ & $0.05$  \\ \thickhline
     \end{tabular}
-    \caption{{\bfseries \sffamily Simulations Time.}}
+    \caption{{\bfseries \sffamily Simulations Time}}
     \label{Table:2}
@@
 
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.6\textwidth]{figs/Figure1.pdf}
+    \caption{{\bfseries \sffamily Spike waveform.} Three integration methods:
+    Runge-Kutta $5$ (\emph{dopri5}), \emph{Adams}, and \emph{BDF} where tested
+    in order to identify potential numerical differences in our simulations
+    (see text for more details). The amplitude for the three methods have no
+    significant differences for the randomly picked up spike waveform shown in 
+    this figure (in general we detected differences up to $5\, \Rm{mV}$, data
+    not shown here). Continuous line -- \emph{dopri5}, dashed line -- \emph{Adams},
+    and dashed-dotted line -- \emph{BDF}.}
+    \label{Fig:1}
+\end{figure}
+
+All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
+generation i$7$ processor, $8\, \Rm{GB}$ of physical memory and running Arch
+Linux.
+The total execution time of all simulations was $526$ minutes and the peak
+consumed memory was $465\, \Rm{MB}$\footnote{Python memory
+    profiler used (\url{https://pypi.python.org/pypi/memory_profiler}).}\@. 
+    All the parameters used during simulations are given in Table~\ref{Table:5}. 
 %%
@@
         Figure  & Type & Form & \parbox[t]{1.5cm}{Frequency
-            ($\frac{1}{P_0}$,$Hz$)} & 
-            \parbox[t]{1.5cm}{Duration ($p$, $ms$)} &
-            \parbox[t]{1.5cm}{Amplitude ($\mu A / cm^2$)} \\
+            ($\frac{1}{P_0}$, $\Rm{Hz}$)} & 
+            \parbox[t]{1.5cm}{Duration ($p$, $\Rm{ms}$)} &
+            \parbox[t]{1.5cm}{Amplitude ($\mu \Rm{A} / \Rm{cm}^2$)} \\
         \thickhline \rowcolor{LightGray}
-        Figure~$1$ & Periodic & 
+        Figure~$1$ & Constant & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$-0.6$}
+            \\\rowcolor{Gray}
+        Figure~$2$ & Periodic & 
             \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} &
-            \parbox[t]{1.1cm}{$5$, $8$, $10$} &
-            \parbox[t]{1.1cm}{$10$, $40$, $75$, $120$} &
+            \parbox[t]{1.1cm}{$5$, $10$} &
+            \parbox[t]{1.1cm}{$10$, $40$} &
             \parbox[c]{0.3cm}{$-1.0$}
             \\\rowcolor{Gray}
-        Figure~$2$ & Constant  & 
-        \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
-        -- & -- & \parbox[c]{0.3cm}{$+3.0$ \\ $0.0$ $-0.5$ $-0.55$ $-0.6$ $-0.8$
-        $-1.3$ $-2.1$} 
-        \\\rowcolor{LightGray}
         Figure~$3$ & Constant  & 
         \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
-            -- & -- & \parbox[c]{0.3cm}{$-0.95$} \\
+        -- & -- & \parbox[c]{0.3cm}{$+3.0$ \\ $3$ $0.0$ $-0.45$ $-0.455$
+            $-0.47$ $-0.55$ $-0.6$ $-0.8$ $-1.3$ $-1.4$ $-2.1$} 
+        \\\rowcolor{LightGray}
+        Figure~$4$ & Pulse/Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & \parbox[c]{0.3cm}{100} & \parbox[c]{0.3cm}{$-1.25$ $0.25$ $-0.47$}
+        \\\rowcolor{Gray}
+        Figure~$5$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$-0.95$}
+        \\\rowcolor{LightGray}
+        Figure~$6$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$[-2,0]$}
+        \\\rowcolor{Gray}
+        Figure~$7$ & Constant  & 
+            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
+            -- & -- & \parbox[c]{0.3cm}{$[-2,0]$} \\
         \thickhline
@@
 
+%%
 \begin{table}[!htbp]
@@
             \begin{aligned}
-                C_m \frac{dV(t)}{dt} &= -I_T - I_h - I_{Na}
-                  - I_K - I_{Na(P)} - I_L + I_{\text{app}}
+                C_{\Rm{m}} \frac{dV(t)}{dt} &= -I_{\Rm{T}} - I_{\Rm{h}} - 
+                I_{\Rm{Na}} - I_{\Rm{K}} - I_{\Rm{Na(P)}} - I_{\Rm{L}} +
+                I_{\text{app}}
               \end{aligned}$ \\ \rowcolor{LightGray}
@@
             \begin{aligned}
-                I_T &= g_T \cdot s^3_{\infty}(V)\cdot h \cdot (V-V_{Ca}) \\
+                I_{\Rm{T}} &= g_{\Rm{T}} \cdot s^3_{\infty}(V)\cdot h \cdot
+                (V-V_{\Rm{Ca}}) \\
                 s_{\infty}(V) &= \frac{1}{1 + \exp(-\frac{V+65}{7.8})} \\
@@
             \begin{aligned}
-                I_{K} &= g_K \cdot n^4 \cdot (V - V_K) \\
+                I_{\Rm{K}} &= g_K \cdot n^4 \cdot (V - V_{\Rm{K}}) \\
                 \frac{dn(t)}{dt} &= \phi_n \frac{n_{\infty}(V) - n(t)}{\tau_n(V)} \\
-                n_{\infty}(\sigma_K, V) &= \frac{\alpha_n(\sigma_K, V)}
-                        {\alpha_n(\sigma_K, V) + \beta_n(\sigma_K, V)} \\
-                \tau_n(\sigma_K, V) &= \frac{1}{\alpha_n(\sigma_K, V)
-                        + \beta_n(\sigma_K, V)} \\
-                \alpha_n(\sigma_K, V) &= \frac{-0.01 (V + 45.7 - \sigma_K)}
-                {\exp(-0.1(V + 45.7 - \sigma_K)) - 1} \\ 
-                \beta_n(\sigma_K, V) &= 0.125 \exp(-\frac{V + 55.7 -
-                \sigma_K}{80}) \\
-                I_{Na} &= g_{Na} \cdot m^3_{\infty}(\sigma_{Na}, V)
-                    \cdot (0.85 - n) \cdot (V - V_{Na}) \\
-                    m_{\infty}(V) &= \frac{\alpha_m(\sigma_{Na}, V)}
-                        {\alpha_m(\sigma_{Na}, V) + \beta_m(\sigma_{Na}, V)} \\
-                \alpha_m(\sigma_{Na}, V) &= -0.1 \frac{V + 29.7 - \sigma_{Na}}
-                    {\exp(-0.1(V + 54.7 - \sigma{Na})) - 1} \\
-                \beta_m(\sigma_{Na}, V) &= 4\exp(-\frac{V+54.7-\sigma_{Na}}{18})
+                n_{\infty}(\sigma_{\Rm{K}}, V) &=
+                \frac{\alpha_n(\sigma_{\Rm{K}}, V)}
+                {\alpha_n(\sigma_{\Rm{K}}, V) + \beta_n(\sigma_{\Rm{K}}, V)} \\
+                \tau_n(\sigma_{\Rm{K}}, V) &= \frac{1}{\alpha_n(\sigma_{\Rm{K}}, V)
+                + \beta_n(\sigma_{\Rm{K}}, V)} \\
+                \alpha_n(\sigma_{\Rm{K}}, V) &= 
+                \frac{-0.01 (V + 45.7 - \sigma_{\Rm{K}})}{\exp(-0.1(V + 45.7 - \sigma_{\Rm{K}})) - 1} \\ 
+                \beta_n(\sigma_{\Rm{K}}, V) &= 0.125 \exp(-\frac{V + 55.7 -
+                    \sigma_{\Rm{K}}}{80}) \\
+                I_{\Rm{Na}} &= g_{\Rm{Na}} \cdot m^3_{\infty}(\sigma_{\Rm{Na}}, V)
+                \cdot (0.85 - n) \cdot (V - V_{\Rm{Na}}) \\
+                m_{\infty}(V) &= \frac{\alpha_m(\sigma_{\Rm{Na}}, V)}
+                {\alpha_m(\sigma_{\Rm{Na}}, V) + \beta_m(\sigma_{\Rm{Na}}, V)} \\
+                \alpha_m(\sigma_{\Rm{Na}}, V) &= -0.1 \frac{V + 29.7 -
+                    \sigma_{\Rm{Na}}}
+                    {\exp(-0.1(V + 54.7 - \sigma{\Rm{Na}})) - 1} \\
+                    \beta_m(\sigma_{\Rm{Na}}, V) &=
+                    4\exp(-\frac{V+54.7-\sigma_{\Rm{Na}}}{18})
             \end{aligned} $ \\ \rowcolor{Gray} 
@@
                 \begin{aligned}
-                    I_{Na(P)} &= g_{Na(P)} \cdot m^3_{\infty}(\sigma_{Na(P)},
-                    V) \cdot (V - V_{Na}) \\
+                    I_{\Rm{Na(P)}} &= g_{\Rm{Na(P)}} \cdot
+                    m^3_{\infty}(\sigma_{\Rm{Na(P)}},
+                    V) \cdot (V - V_{\Rm{Na}}) \\
                 \end{aligned} $
@@
 \end{table}
+%%
 
+%%
 \begin{landscape}
@@
     \centering
-    \begin{tabular}{p{1.5cm}lllll}
+    \begin{tabular}{cccccccc}
         \thickhline
-        \multicolumn{5}{c}{Model Parameters} \\\thickhline \rowcolor{Gray}
-        Currents Type  & Common &  Figure $1$ & Figure $2$ & Figure $3$ \\
-        \thickhline \rowcolor{LightGray}
-        Membrane Potential & $
-            \begin{aligned}
-                C_m &= 1 \mu F/cm^2 \\
-                V_0 &= -74mV
-            \end{aligned} $ & & & \\\rowcolor{Gray}
-        T-Type Calcium Current & $
-            \begin{aligned}
-                \phi_h &= 2 \\
-                V_{Ca} &= 120mV
-            \end{aligned} $ & $
-            \begin{aligned}
-                g_T &= 1 mS/cm^2 \\
-                \theta_h &= -81 mV \\
-                k_h &= 6.25 mV^{-1}
-            \end{aligned} $ & $
-            \begin{aligned}
-                g_T &= 0.3 mS/cm^2\\
-                \theta_h &= -79 mV \\
-                k_h &= 5 mV^{-1}
-            \end{aligned} $ & $
-            \begin{aligned}
-                g_T &= 0.3 mS/cm^2\\
-                \theta_h &= -75 mV \\
-                k_h &= 5 mV^{-1}
-            \end{aligned} $ \\\rowcolor{LightGray}
-        Sag Current & $
-            \begin{aligned}
-                \phi_H &= 1 \\
-                g_h &= 0.04 mS/cm^2 \\
-                V_h &= -40 mV
-            \end{aligned} $
-        & & & \\\rowcolor{Gray}
-        Hodgkin-Huxley Currents & $
-            \begin{aligned}[l]
-                g_K &= 30 mS/cm^2 \\
-                V_K &= -80 mV \\
-                \phi_n &= 28.5\\
-                \sigma_K &= 10 \\
-            \end{aligned} 
-            \begin{aligned}[r]
-                g_{Ca} &= 42 mS/cm^2 \\
-                V_{Ca} &= 55 mV \\
-            \end{aligned} $
-            & $\sigma_{Na} = 3$ & $\sigma_{Na} = 6$ & $\sigma_{Na} = 6$
-            \\\rowcolor{LightGray}
-        Persistent Sodium Currents & $
-            \begin{aligned}
-                V_{Na(P)} &= 55 mV \\
-                \sigma_{Na(P)} &= -5
-            \end{aligned} $
-            & $g_{Na(P)} = 9 mS/cm^2$ & $g_{Na(P)} = 9 mS/cm^2$ &
-                $g_{Na(P)} = 9mS/cm^2$ 
-            \\\rowcolor{Gray}
-        Leak Current & & $
-            \begin{aligned}
-                g_L &= 0.1 mS/cm^2\\
-                V_L &= -72 mV
-            \end{aligned} $ & $
-            \begin{aligned}
-                g_L &= 0.12 mS/cm^2\\
-                V_L &= -70 mV
-            \end{aligned} $ & $
-            \begin{aligned}
-                g_L &= 0.08 mS/cm^2\\
-                V_L &= -70 mV
-            \end{aligned} $ \\
+        \multicolumn{8}{c}{Model Parameters} \\\thickhline
+        Figure & $V_0\, (\Rm{mV})$ & $g_T\, (\Rm{mS/cm^2})$ & $\theta_h\,
+        (\Rm{mV})$ & $k_h\, (\Rm{mV^{-1}})$ & $\sigma_{Na}\, (\Rm{mV})$ &
+        $g_L\, (\Rm{mS/cm^2})$ & $V_L\, (\Rm{mV})$ \\\rowcolor{LightGray}
         \thickhline
+        $1$ & $-74$ & $1$ & $-79$ & $-5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
+        $2$ & $-74$ & $1$ & $-81$ & $6.25$ & $3$ & $0.1$ & $-72$ \\\rowcolor{LightGray}
+        $3$ & $-74$ & $0.3$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
+        $4$ & $-72$/$-64$ & $1.0$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{LightGray}
+        $5$ & $-74$ & $0.3$ & $-75$ & $5$ & $6$ & $0.08$ & $-70$ \\\rowcolor{Gray}
+        $6$ & $-74$ & $1$/$0.7$ & $-79$ & $5$ & $6$ & $0.1$/$0.04$ & $-70$ \\\rowcolor{LightGray}
+        $7$ & $-72$ & $0.3$/$0.25$ & $-81$ & $6.25$ & $3$ & $0.1$ & $-72$ \\
+        \thickhline
+        \multicolumn{8}{l}{Common Parameters} \\\rowcolor{Gray}
+        \multicolumn{8}{c}{$C_m = 1\, \Rm{\mu F/cm^2}$, $\phi_h = 2$,
+        $V_{\Rm{Ca}} = 120\, \Rm{mV}$, $\phi_H = 1$, $g_h = 0.04\,
+        \Rm{mS/cm^2}$, $V_h = -40\, \Rm{mV}$, $g_{\Rm{K}} = 30\, \Rm{mS/cm^2}$,
+        $V_{\Rm{K}} = -80\, \Rm{mV}$, $g_{\Rm{Ca}} = 42\, \Rm{mS/cm^2}$} \\\rowcolor{Gray}
+        \multicolumn{8}{c}{$V_{\Rm{Ca}} = 55\, \Rm{mV}$, $\phi_{n} = 28.5$,
+        $\sigma_{\Rm{K}} = 10\, \Rm{mV}$, $V_{\Rm{Na(P)}} = 55\, \Rm{mV}$,
+        $\sigma_{\Rm{Na(P)}} = -5\, \Rm{mV}$, $g_{\Rm{Na(P)}} = 9 \,
+        \Rm{mS/cm^2}$}
+        \\\thickhline
     \end{tabular}
-    \caption{{\bfseries \sffamily Simulations Parameters}} 
+    \caption{\bfseries \sffamily Simulation Parameters}
     \label{Table:5}
@@
 \end{landscape}
+%%
 
@@
 given in Table~\ref{Table:5} and the corresponding input (see 
-Table~\ref{Table:3} for three different cases. First, we verified 
-that the model is able to generate different types of burst responses such
-as in figure $1$ of the original article. We did not reproduce the entire 
-diagram, instead we picked up five different cases from the original article
-and we tested our implementation on these cases. Thus we applied a periodic 
-current pulse of $-1 \mu A/cm^2$ amplitude at three different frequencies
-(in the original article frequency is referred as $\frac{1}{P_0}$) $5Hz$, $8Hz$
-and $10Hz$. The pulse is on for $p ms$ and the ratio $\frac{p}{P_0}$ is
-defined, accordingly.
-We chose the frequency and the on-duration of the pulse according to figure $1$
-of the original article and such that we have a different type of burst response
-for each ratio. The results are shown in Figure~\ref{Fig:1}. It is apparent that
-the reference implementation can achieve the same behavior as the original
-implementation. 
+Table~\ref{Table:3}). First, we examined what is the response of the reference 
+implementation to rhythmic hyperpolarization. In \cite{wang:1994} this is 
+illustrated in Fig~$1$\footnote{From now and then all the 
+    figures of the original article will be referred as Fig} of the original
+article. 
+Thus, we applied a periodic current pulse of $-1\,\Rm{\mu A/cm^2}$ amplitude at
+several different frequencies ($\frac{1}{P_0}$ is the frequency in $\Rm{Hz}$
+and $P_0$ is the corresponding period in $\Rm{ms}$) ranging from 
+$0.1\,\Rm{Hz}$ to $15\,\Rm{Hz}$ with a resolution (discretization) of $30$ 
+points. The same number of samples used for discretizing the duration for each 
+frequency. The results are shown in Figure~\ref{Fig:2} left panel. Right panel
+shows three different simulations of the reference implementation at specific
+frequencies ($10,\, 5,\, 0.5\, \Rm{Hz}$) and rations $p/P_0$ ($0.1, 0.6, 0.6$),
+respectively (crossed area in Figure~\ref{Fig:2} reflects the transient, which 
+is omitted in the rest of the figures in this work). It is apparent that right
+and left panels are in an agreement indicating that the reference model
+produces reliable results. 
+The author in \cite{wang:1994} shows in Fig~$1$ a shaded are where the number 
+of spikes is $0.5$. This number appears if one computes the average of 
+suprathreshold and subthreshold spikes in one period.
+
 %%
@@
     \centering
-    \includegraphics[width=.7\textwidth]{figs/Figure1.pdf}
+    \includegraphics[width=1.1\textwidth]{figs/Figure2.pdf}
     \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
-        We stimulate the model with a periodic pulse with frequency
-        $\frac{1}{P_0}$ and we record the response of the model. All the
-        parameters for this simulation are given in Table~\ref{Table:3}. 
-        {\bfseries \sffamily Blue} indicates a frequency $\frac{1}{P_0} = 10Hz$
-        and a pulse duration $p = 10 ms$. A resting state activity is 
-        generated.
-        {\bfseries \sffamily Black} color shows the case where $\frac{1}{P_0} =
-        10 Hz$ and $p = 40 ms$. In this case we get a single spike
-        periodically. 
-        {\bfseries \sffamily Red} color demonstrates a burst with two spikes,
-        when the frequency of the periodic pulse is taken to be $\frac{1}{P_0}
-        = 5 Hz$ and $p = 40 ms$. 
-        {\bfseries \sffamily Cyan} curve shows burst responses consist in 
-        three spikes. In this case $\frac{1}{P_0} = 8 Hz$ and $p = 75ms$. 
-        {\bfseries \sffamily Magenta} curve indicates a burst periodical 
-        response of four spikes per period. }
-    \label{Fig:1}
+        We stimulate the model with a periodic pulse with varying frequency
+        $\frac{1}{P_0}$ in $[0.1, 15]Hz$ and ON pulse duration $p$. 
+        {\bfseries \sffamily Left Panel:} shows a range of frequencies
+        $\frac{1}{P_0}$ from $0.1Hz$ to $15Hz$ versus the ratio
+        $\frac{p}{P_0}$ in range $[0, 1]$. Numbers indicate the number of 
+        spikes per period. 
+        {\bfseries \sffamily Right Panels:} In all these plots the 
+        crossed area corresponds to the transient, and thus it has to be
+        omitted. 
+        {\bfseries \sffamily Red} color demonstrates a case where only
+        subthreshold spikes are generated, when the frequency of the periodic
+        pulse is $\frac{1}{P_0} = 10 Hz$ and $\frac{p}{P_0} = 0.1$. 
+        {\bfseries \sffamily Blue} indicates bursts of four spike, when the
+        frequency is $\frac{1}{P_0} = 5Hz$ and $\frac{p}{P_0} = 0.6$.
+        {\bfseries \sffamily Black} shows a case of two spikes bursts. Here
+        the frequency is $\frac{1}{P_0} = 0.5 Hz$ and $\frac{p}{P_0} = 0.6$.}
+    \label{Fig:2}
 \end{figure}
 %%
-In a second case, we used a steady current varying only its amplitude and 
-keeping the rest parameters fixed. Hence, we used $8$ different current
-amplitudes ($3, 0, -0.5, -0.55, -0.6, -0.8, -1.3$ and $-2.1 \mu A/cm^2$) and
-Figure~\ref{Fig:2} illustrates the results of our simulations, which
-correspond to figure $3$ of \cite{wang:1994}. Again original results
-are reproducible and the model is capable of generating repetitive firing,
-resting state, subthreshold oscillation, $10 Hz$ bursting, $3Hz$ bursting
-and hyperpolarized steady activity. However, the external current amplitudes
-used in this case are slightly different from the ones used in the original
-article. 
+
+Then we tested the transition from subthreshold to bursting oscillation via 
+chaos. Therefore, we used a steady current varying only its amplitude keeping
+all the other parameters fixed. We used eight different values (exactly the
+same as in \cite{wang:1994}),
+$$I_{\text{app}} =
+3.0, 0.0, -0.45, -0.455, -0.47, -0.55, -0.6, -0.8, -1.3, -1.4, -2.0\, 
+\Rm{\mu A /cm^2}.$$ Results are depicted in Figure~\ref{Fig:3}. This figure
+corresponds to Fig~$3$ of \cite{wang:1994}. In this case, there was no
+difference between the reference implementation and the original one. 
 %%
@@
     \centering
-    \includegraphics[width=.8\textwidth]{figs/Figure2.pdf}
-    \caption{{ \bfseries \sffamily Dynamic behavior of neuron.} Here we
-    simulate the model with parameters given in Table~\ref{Table:3}. The
-    only parameter that changes from panel to panel in this figure is the
-    external current $I_{\text{app}}$. The jet colormap indicates the different
-    current values. We observe that we get quite similar results as in the original
-    article \cite{wang:1994}. For isntance, we obtain, repetitive spiking when
-    we apply a constant current at $3 \mu A / cm^2$, or a periodic bursting 
-    response, when we apply an external current of $-1.3 \mu A /cm^2$. These
-    results correspond to Figure~$3$ of \cite{wang:1994}.}
-    \label{Fig:2}
+    \includegraphics[width=1.\textwidth]{figs/Figure3.pdf}
+    \caption{{\bfseries \sffamily Dynamic behavior of neuron.} Here we
+    simulate the model with parameters given in Table~\ref{Table:5}. The
+    only parameter that varies from panel to panel in this figure is the
+    external current $I_{\text{ext}}$. These results correspond to Fig~$3$ 
+    of \cite{wang:1994}.}
+    \label{Fig:3}
 \end{figure}
 %%
-The last case we simulated is the one that corresponds to figure $6$ of 
-\cite{wang:1994}. In this test-case the neuron receives a constant
-external current at $-0.95 \frac{\mu A}{cm^2}$ and as Figure~\ref{Fig:3}
-points out ``spiral chaos'' is generated as in the original article.
-In addition, our results are quantitatively and qualitatively comparable
-with the ones provided by \cite{wang:1994} pinpointing that the original
-model is fully reproducible apart from the point that we had to properly
-tune the amplitude of the external current. The parameters for this simulation
-are given in Table~\ref{Table:4}. 
+
+The next simulation we performed reproduces results related to hysteresis as
+in Fig~$4$ of \cite{wang:1994}. First, we run a simulation where the input 
+current $I_{ext}$ varied from $-0.433\, \Rm{\mu A / cm^2}$ to $-0.55 \mu A /cm^2$.
+At every iteration, we detected if there is subthreshold and/or suprathreshold 
+activity in the membrane voltage trace. Thus we created the hysteresis diagram
+given in Figure~\ref{Fig:4}, left panel. 
 %%
@@
     \centering
-    \includegraphics[width=.7\textwidth]{figs/Figure3.pdf}
+    \includegraphics[width=1.1\textwidth]{figs/Figure4.pdf}
+    \caption{{\bfseries \sffamily Hysteresis near the transition from the
+    subthreshold to bursting oscillation.} {\bfseries \sffamily Left panel:}
+    Hysteresis diagram indicates a coexistence of bursting and subthreshold 
+    activity. {\bfseries \sffamily Upper right panel:} Bistability example
+    shows how the same system can produce either subthreshold activity (black
+    curve) or burst activity (gray curve). {\bfseries \sffamily Middle right 
+    panel:} shows the activation of $H$ for the same simulation as in the upper 
+    panel. {\bfseries  \sffamily Lower right panel:} Input to the model for
+    this simulation is a brief current pulse at $0.25\, \Rm{\mu A /cm^2}$
+    (black curve) and $-1.2\, \Rm{\mu A/cm^2}$ (gray curve) followed by a
+    steady current at $-0.47\, \Rm{\mu A/cm^2}$ for both simulations. }
+    \label{Fig:4}
+\end{figure}
+%%
+In addition, we tested the bistability of the current model using the same 
+protocol as in \cite{wang:1994}. According to that protocol, we first apply 
+a brief pulse (for $100\, \Rm{ms}$) of $0.25\, \Rm{\mu A /cm^2}$ and
+$-1.2\, \Rm{\mu A/cm^2}$ followed by a steady current at $-0.47\, 
+\Rm{\mu A /cm^2}$. This causes the state of the model to switch from a purely 
+subthreshold activity pattern to a mixed sub- and suprathreshold activity
+pattern. The results are given in Figure~\ref{Fig:4}, right panel.
+
+Another interesting behavior of the model is the development of a ``spiral''
+chaos (see Fig~$6$ in \cite{wang:1994}). The reference implementation is 
+capable of generating similar behavior as Figure~\ref{Fig:5} shows. In order
+to get these results we applied a constant external current with an amplitude
+of $-0.95\, \Rm{\mu A / cm^2}$. 
+%%
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=.7\textwidth]{figs/Figure5.pdf}
     \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
-    we show that the model is capable of generating spiral chaos as
-    it has been shown in Figure~$6$ \cite{wang:1994}. Again all the 
-    parameters for this simulation are given in Table~\ref{Table:3}.
-    The external current in this simulation is constant and we have 
-    traced the phase portrait of the membrane potential and of the
-    Sag current ($V$ vs $h$). The top panel illustrates the phase 
-    portrait, the middle panel shows the membrane potential ($V$)
-    and the bottom panel depicts the $h$ current over time. Our 
-    results are almost identical to the original ones rendering 
-    the model fully reproducible. }
-    \label{Fig:3}
+    we show that the model is capable of generating ``spiral chaos''
+    as it has been shown in Fig~$6$ in \cite{wang:1994}.
+    The external current in this simulation is constant ($-0.95\Rm{\mu A/cm^2}$). 
+    {\bfseries \sffamily Top panel} shows the phase portrait of the 
+    membrane potential and the Sag current ($V$ vs $h$).
+    {\bfseries \sffamily Middle panel} illustrates the membrane potential ($V$)
+    and, the {\bfseries \sffamily bottom panel} shows the $h$ current over time.}
+    \label{Fig:5}
+\end{figure}
+%%
+
+The next simulation investigates how the bursting frequency is affected by 
+the injected external current (see Fig~$7$ in \cite{wang:1994}, two bursting
+modes). To address this issue we captured data\footnote{Data are available in
+    the accompanying github repository of the present article.} values for
+the external current from Fig~$7$ of the original article (dots in Fig $7$,
+pg. $27$ in \cite{wang:1994}) using a software called 
+PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
+The results from our simulations are illustrated in Figure~\ref{Fig:6}. Black
+curve represents the frequency in $\Rm{Hz}$ and the blue one the period in 
+$\Rm{sec}$. Provided results here are not as smooth as the original ones in
+\cite{wang:1994}. Unfortunately there are no many details in the original 
+article about these diagrams so we cannot reproduce them in full details. 
+%%
+\begin{figure}[!htbp]
+    \includegraphics[width=1.1\textwidth, left]{figs/Figure6.pdf}
+    \caption{{\bfseries \sffamily Frequency and period versus steady input 
+    current.} This figure illustrates how the external current (ranging
+    from $0$ to $-2\, \Rm{\mu A/cm^2}$ affects the bursts frequency (bursting
+    mode) of the model. Black curve indicates the frequency in $\Rm{Hz}$
+    and the blue curve the period in $\Rm{sec}$. The external
+    current varies exactly as in Fig~$7$ in \cite{wang:1994} (see text for
+    more details). }
+    \label{Fig:6}
+\end{figure}
+%%
+
+The last simulation is related to Fig~$2$ of \cite{wang:1994} and to an
+``intermittent'' phase-locking phenomenon. A periodic pulse with frequency 
+$10\,\Rm{Hz}$ and ratio $\frac{p}{P_0} = 0.8$ is applied to the model as
+external input current. The response of the model is registered and sub- and
+supra-threshold spikes are counted. We applied six different current amplitudes
+and two different values for the T-type calcium conductance. 
+$I_{app} = -1.4, -1.5, -1.6\, \Rm{\mu A /cm^2}$ and $g_T = 0.3\, \Rm{mS/cm^2}$
+$I_{app} = -1.2, -1.5, -1.8\, \Rm{\mu A/cm^2}$ and  $g_T = 0.25\,
+\Rm{mS/cm^2}$. Figure~\ref{Fig:7} shows $2$ seconds of membrane potential 
+along with symbolic patterns of $0$ (dark circles) and $1$ (green line
+segments). Results given in Figure~\ref{Fig:7} are similar to the ones of
+Fig~$2$A and Fig~$2$B in \cite{wang:1994}.
+%%
+\begin{figure}[!htbp]
+    \includegraphics[width=1.1\textwidth]{figs/Figure7.pdf}
+    \caption{{\bfseries \sffamily Symbolic patterns.} Six different simulations
+    are illustrated here ($I_{app} = -1.4, -1.5 -1.6 \, \Rm{\mu A/cm^2}$
+    with $g_T = 0.3\, \Rm{mS / cm^2}$) and ($I_{app} = -1.2, -1.5, -1.8\,
+    \Rm{\mu A / cm^2}$ with $g_T = 0.25\, \Rm{mS / cm^2}$). Symbolic patterns
+    are illustrated above membrane potential traces as green line segments -- 
+    suprathreshold spikes and dark circles -- subthreshold spikes. Red curve 
+    indicates the input periodic pulses.}
+    \label{Fig:7}
 \end{figure}
@@
 A conductance-based model for relay thalamocortical neurons proposed 
-by \cite{wang:1994} was implemented in Python. After a minor modification
-of the external current amplitude we were able to reproduce the results 
-presented in \cite{wang:1994}. The original model is easy to implement
-since all the equations, parameters and simulation details are given 
-in \cite{wang:1994}. All replicated figures are pretty much similar to
-the original ones rendering the new results quantitatively and qualitatively 
-comparable to the original ones.
+by \cite{wang:1994} was implemented in Python. The model tested 
+thoroughly in several examples taken from the original article. 
+In general, the original model was easy to be implemented since all the
+equations and the most of the parameters (except the initial time step of the
+integration method) are given. The reference implementation results are similar
+to the original ones. Unfortunately, we could not find any other implementation
+of the original model described in \cite{wang:1994} to compare with.
 

### 2016-07-18 code/neuron_model.py

@@ -50,10 +50,10 @@ def loadParameters(fname):
             'phi_H': config.getfloat('Sag', 'phi_H'),
             'gH': config.getfloat('Sag', 'gH'),
             'VH': config.getfloat('Sag', 'VH'),
-            'phi_K': config.getfloat('Potasium', 'phi_K'),
-            'sigma_K': config.getfloat('Potasium', 'sigma_K'),
-            'gK': config.getfloat('Potasium', 'gK'),
-            'VK': config.getfloat('Potasium', 'VK'),
+            'phi_K': config.getfloat('Potassium', 'phi_K'),
+            'sigma_K': config.getfloat('Potassium', 'sigma_K'),
+            'gK': config.getfloat('Potassium', 'gK'),
+            'VK': config.getfloat('Potassium', 'VK'),
             'gNa': config.getfloat('Sodium', 'gNa'),
             'VNa': config.getfloat('Sodium', 'VNa'),
             'sigma_Na': config.getfloat('Sodium', 'sigma_Na'),
@@ -74,14 +74,27 @@ class WangModel():
                  duration=1000,
                  inp_type='const',
                  amplitude=1,
-                 chunks=1):
+                 pulse_ampl=0,
+                 chunks=1,
+                 store=True,
+                 store_stim=False):
         """ Constructor method
 
         |  :param fname: File name containing all parameters
+        |  :param dt: Time discretization step
         |  :param frequency: Stimulation frequency (in Hz)
         |  :param duration: Duration (ON) of stimulation (in ms)
+        |  :param inp_type: Input signal type ['const' - constant current
+                                               'periodic' - periodic pulses]
+                                               'pulse' - one pulse at the 
+                                               beggining of simulation
+                                               following by a constant current.
         |  :param amplitude: Amplitude of injected current (in μA/cm²)
+        |  :param pulse_ampl: Amplitude of initial pulse (input must be set to
+                                                          'pulse')
         |  :param chunks: Number of repeated periods of pulses
+        |  :param store: Store results of simulation on npy files
+        |  :param store_stim: Store the stimulus signal
         |  :return:
         """
         self.p = loadParameters(fname)      # Load the parameters
@@ -91,8 +104,11 @@ class WangModel():
         self.P0 = (1.0/self.freq)*1000      # Period in ms
         self.P = duration                   # Duration in ms
         self.amplitude = amplitude          # Amplitude in μA/cm²
[…]

### 2016-07-18 code/plot_figures.py

@@ -29,9 +29,48 @@ from matplotlib import collections  as mc
 from skimage.feature import peak_local_max
 
 
+def plot_figure1():
+    dt = 0.05
+    base = '../data/'
+    dopri = np.load(base+'test_dopri_V.npy')
+    adams = np.load(base+'test_adams_V.npy')
+    bdf = np.load(base+'test_bdf_V.npy')
+
+    def spk(x):
+        tmp = np.zeros((x.shape[0], ))
+        spikes = peak_local_max(x, min_distance=10)
+        spikes = spikes[x[spikes] > 0]
+        tmp[spikes] = 1
+        return tmp.astype('f')
+
+    M, N = 7010, 7045
+    fig = plt.figure(figsize=(10, 10))
+    ax = fig.add_subplot(111)
+    ax.plot(dopri[M:N, 0], dopri[M:N, 1], 'k', lw=2, alpha=0.3, label='dopri5')
+    ax.plot(adams[M:N, 0], adams[M:N, 1], 'k--', lw=2, alpha=0.5, label='Adams')
+    ax.plot(bdf[M:N, 0], bdf[M:N, 1], 'k-.', lw=2, alpha=0.7, label='BDF')
+    ax.set_xlabel('Time (ms)', fontsize=16, weight='bold')
+    ax.set_ylabel('Membrane Potential (mV)', fontsize=16, weight='bold')
+    ax.get_xaxis().set_tick_params(which='both', direction='out')
+    ax.get_yaxis().set_tick_params(which='both', direction='out')
+    ticks = ax.get_xticks()
+    ticks = [i*dt for i in ticks]
+    ax.set_xticklabels(ticks, fontsize=13, weight='bold')
+    ticks = ax.get_yticks()
+    ax.set_yticklabels(ticks, fontsize=13, weight='bold')
+    handles, labels = ax.get_legend_handles_labels()
+    ax.legend(handles, labels, loc='upper right')
+    ax.spines['right'].set_visible(False)
+    ax.spines['top'].set_visible(False)
+    ax.yaxis.set_ticks_position('left')
+    ax.xaxis.set_ticks_position('bottom')
+    plt.savefig('../article/figs/Figure1.pdf', axis='tight')
+
 def detect_spikes(data, freq=0, dt=0.05, dist=40, th=0):
+    """ Detect spikes in a period of the input signal
+    as local maxima """
     win = int(((1.0 / freq) * 1000) / dt)
[…]

### 2016-07-18 code/simulations.py

@@ -31,32 +31,71 @@ from neuron_model import *
 #                           fname_for_time,
 #                           'vode', 'Adams' [or 'BDF'] ) 
 ########################################################################
-def simulate_figure1(diagram=False):
-    print("Running simulation #1")
+def simulate_figure1():
+    print("Running test for comparing integration methods")
+    base = './params/'
+    Iext = -0.6
+    name = ['test_dopri', 'test_adams', 'test_bdf']
+
+    simulation = WangModel(base+"params_figure1.cfg",
+                           inp_type='const',
+                           amplitude=Iext)
+    simulation.run(6000, name[0]+"_V", 'dopri5')
+    simulation.run(6000, name[1]+"_V", 'vode', 'Adams')
+    simulation.run(6000, name[2]+"_V", 'vode', 'BDF')
+
+    base = '../data/'
+    def spk(x):
+        tmp = np.zeros((x.shape[0], ))
+        spikes = peak_local_max(x, min_distance=10)
+        spikes = spikes[x[spikes] > 0]
+        tmp[spikes] = 1
+        return tmp
+
+    dopri = np.load(base+'test_dopri_V.npy')
+    adams = np.load(base+'test_adams_V.npy')
+    bdf = np.load(base+'test_bdf_V.npy')
+    spks1 = spk(dopri[:, 1]).astype('i')
+    spks2 = spk(adams[:, 1]).astype('i')
+    spks3 = spk(bdf[:, 1]).astype('i')
+    isi1 = np.diff(dopri[spks1==1, 0])
+    isi2 = np.diff(adams[spks2==1, 0])
+    isi3 = np.diff(bdf[spks3==1, 0])
+    CV1 = np.mean(isi1) / np.std(isi1)
+    CV2 = np.mean(isi2) / np.std(isi2)
+    CV3 = np.mean(isi3) / np.std(isi3)
+    print('sum(Spikes dopri5 - spikes Adams):  ',(dopri[spks1,0]-adams[spks2,0]).sum())
+    print('sum(Spikes dopri5 - spikes BDF):  ',(dopri[spks1,0]-bdf[spks3,0]).sum())
+    print('CV dopri5:', CV1)
+    print('CV Adams:', CV2)
+    print('CV BDF:', CV3)
+
+
+def simulate_figure2(diagram=False):
+    print("Running simulation #2")
 
     if diagram is True:
-        samples = 150
+        samples = 30
         freq = np.linspace(0.1, 15, samples)
[…]

### 2016-08-18 article/detorakis-2016.tex
Review #3 comments addressed

@@
 In this section, a detailed description of the model is given following 
-the paradigm of Nordlie et al, \cite{nordlie:2009}. Therefore, a brief
-description of the model, equations, parameters, and inputs are given in
-the form of tables.
+the paradigm of Nordlie et al, \cite{nordlie:2009}. To this end description of
+the model, equations, parameters, and inputs are given in the form of tables.
 
@@
 provides information about simulations duration and temporal integration time
-steps, Table~\ref{Table:3} gives a glimpse of the input signals used in this work.
-Table~\ref{Table:4} introduces the equations of the model and finally 
+steps, Table~\ref{Table:3} gives a glimpse of the input signals used in this
+work. Table~\ref{Table:4} introduces the equations of the model and finally 
 Table~\ref{Table:5} summarizes all the parameters for each figure (simulation).
@@
 Three different methods have been tested in this work (\emph{dopri5}, 
-\emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{dopri5} is the closest
-numerical method to the one used by the author in the original article (the
-author has numerically integrated the system of the four ODEs by using a
-fifth-order adaptive size Runge-Kutta method). \emph{BDF} and \emph{Adams}
-provide similar numerical results as the first one, but they are
-faster\footnote{The user has the option to choose one of the three methods
-    during class instantiating. We tested all three methods comparing, spike
-    times, amplitudes and the coefficient of variation using as threshold value
-    $0\, \Rm{mV}$. Spike events were overlapping, and $CV \simeq 1.5$ for all
-    three methods. The membrane potential amplitude was sometimes different
-    for the same spike events, however that difference was not too high as
-    Figure~\ref{Fig:1} shows (up to $4\, \Rm{mV}$ at maximum). Thus we decided
-    to use the \emph{Adams} method.}.
+\emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{Dopri5} is the closest
+numerical method to the one used by \cite{wang:1994} (fifth-order adaptive 
+size Runge-Kutta method). \emph{BDF} and \emph{Adams} provide similar numerical
+results as the first one, but they are faster. In the current implementation
+of \cite{wang:1994} the user has the option to choose one of the three methods
+at the stage of class instantiating. All three aforementioned methods have
+been tested by comparing, spike times, amplitudes and the coefficient
+of variation using as threshold value $0\, \Rm{mV}$. Spike events were
+overlapping, and $CV \simeq 3.5$ for all three methods. A difference found 
+in the amplitude of spikes generated by the three methods. Figure~\ref{Fig:1}
+shows two histograms of error distribution (error is defined to be the
+absolute difference of spike amplitude between two methods). Top histogram 
+depicts a distribution of differences in amplitude between \emph{dopri5} 
+and \emph{Adams} methods (gray color). The bottom histogram illustrates the 
+distribution of amplitude differences between \emph{dopri5} and \emph{BDF}
+methods (red color). In the rest of the article all the results have been 
+produced by using the \emph{Adams} method.
 %%
@@
         $1$ & $6$ & $0.05$ \\ \rowcolor{Gray}
-        $2$ & $15\times period$, $2$ & $0.05$ \\ \rowcolor{LightGray}
+        $2$ & $15\times period$, $5$ & $0.05$ \\ \rowcolor{LightGray}
         $3$ & $6$ & $0.05$ \\ \rowcolor{Gray} 
-        $4$ & $1.5$, $1.5$, $1$ & $0.05$ \\ \rowcolor{LightGray} 
+        $4$ & $1.5$, $1$, $1$ & $0.05$ \\ \rowcolor{LightGray} 
         $5$ & $5$ & $0.05$  \\ \rowcolor{Gray} 
-        $6$ & $2.5$ & $0.05$  \\ \rowcolor{LightGray} 
+        $6$ & $5$ & $0.05$  \\ \rowcolor{LightGray} 
         $7$ & $40$, $20$ & $0.05$  \\ \thickhline
@@
     \includegraphics[width=0.6\textwidth]{figs/Figure1.pdf}
-    \caption{{\bfseries \sffamily Spike waveform.} Three integration methods:
-    Runge-Kutta $5$ (\emph{dopri5}), \emph{Adams}, and \emph{BDF} where tested
-    in order to identify potential numerical differences in our simulations
-    (see text for more details). The amplitude for the three methods have no
-    significant differences for the randomly picked up spike waveform shown in 
-    this figure (in general we detected differences up to $5\, \Rm{mV}$, data
-    not shown here). Continuous line -- \emph{dopri5}, dashed line -- \emph{Adams},
-    and dashed-dotted line -- \emph{BDF}.}
+    \caption{{\bfseries \sffamily Amplitude difference distribution} 
+    {\bfseries \sffamily Top histogram} depicts the distribution of differences
+    in spikes amplitudes between \emph{dopri5} and \emph{Adams} integration
+        methods of Scipy \emph{integrate} package (black color). 
+    {\bfseries \sffamily Bottom histogram} illustrates the distribution of
+    differences in spikes amplitudes between \emph{dopri5} and \emph{BDF}
+    methods (red color). In both cases the number of bins is $30$. 
+    }
     \label{Fig:1}
@@
 The total execution time of all simulations was $526$ minutes and the peak
-consumed memory was $465\, \Rm{MB}$\footnote{Python memory
-    profiler used (\url{https://pypi.python.org/pypi/memory_profiler}).}\@. 
-    All the parameters used during simulations are given in Table~\ref{Table:5}. 
+consumed memory was $465\, \Rm{MB}$\footnote{Python memory profiler used 
+    (\url{https://pypi.python.org/pypi/memory_profiler}).}\@. 
+All the parameters used in this work are given in Table~\ref{Table:5}. 
 %%
@@
 implementation to rhythmic hyperpolarization. In \cite{wang:1994} this is 
-illustrated in Fig~$1$\footnote{From now and then all the 
-    figures of the original article will be referred as Fig} of the original
-article. 
+illustrated in Fig~$1$\footnote{From now and then original article's figures
+    will be referred as Fig}. 
 Thus, we applied a periodic current pulse of $-1\,\Rm{\mu A/cm^2}$ amplitude at
@@
 and $P_0$ is the corresponding period in $\Rm{ms}$) ranging from 
-$0.1\,\Rm{Hz}$ to $15\,\Rm{Hz}$ with a resolution (discretization) of $30$ 
-points. The same number of samples used for discretizing the duration for each 
-frequency. The results are shown in Figure~\ref{Fig:2} left panel. Right panel
-shows three different simulations of the reference implementation at specific
-frequencies ($10,\, 5,\, 0.5\, \Rm{Hz}$) and rations $p/P_0$ ($0.1, 0.6, 0.6$),
-respectively (crossed area in Figure~\ref{Fig:2} reflects the transient, which 
-is omitted in the rest of the figures in this work). It is apparent that right
-and left panels are in an agreement indicating that the reference model
-produces reliable results. 
-The author in \cite{wang:1994} shows in Fig~$1$ a shaded are where the number 
-of spikes is $0.5$. This number appears if one computes the average of 
-suprathreshold and subthreshold spikes in one period.
+$0.1\,\Rm{Hz}$ to $15\,\Rm{Hz}$ with a resolution (discretization) of $100$ 
+points. The same number of samples used for discretizing the duration of the
+pulse for each frequency.
+
+Upper panel shows two different simulations of the reference implementation at
+specific frequencies ($5,\, 0.5\, \Rm{Hz}$) and ratios $p/P_0$ ($0.6, 0.6$),
+respectively (following the upper panels of Fig~$1$ in \cite{wang:1994}).
+These panels are comparable to the ones presented in \cite{wang:1994} and the 
+timescales are exactly the same ($400\, \Rm{ms}$ for the left sub-panel and
+$4\, \Rm{s}$ for the second one). However, numbers regarding pulse ON duration
+($p$) in Fig~$1$ of \cite{wang:1994} are not correct. For instance, take the
+frequency to be $\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$ then
+the pulse ON duration must be $p = 120\, \Rm{ms}$ and not $160\, \Rm{ms}$.
+
+Bottom panel illustrates the total results of the $100 \times 100$ simulations
+we ran.  In order to create this figure, we counted the number of spikes
+generated per pulse, both supra-threshold and sub-threshold, and then we 
+computed the mean (this procedure is not described in a solid way in
+\cite{wang:1994}, however it is described in \cite{mccormick:1990}).  
+This figure is different from \cite{wang:1994}, where the shaded area that 
+contains an average number of spikes below $1.0$ (but not zero) is larger
+than in \cite{wang:1994}. Unfortunately, information regarding Fig~$1$
+is limited in \cite{wang:1994} in order to be able to follow a specific recipe
+and try to reproduce exactly the same figure.
 
@@
     \centering
-    \includegraphics[width=1.1\textwidth]{figs/Figure2.pdf}
+    \includegraphics[width=0.6\textwidth]{figs/Figure2.pdf}
     \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
         We stimulate the model with a periodic pulse with varying frequency
-        $\frac{1}{P_0}$ in $[0.1, 15]Hz$ and ON pulse duration $p$. 
-        {\bfseries \sffamily Left Panel:} shows a range of frequencies
-        $\frac{1}{P_0}$ from $0.1Hz$ to $15Hz$ versus the ratio
-        $\frac{p}{P_0}$ in range $[0, 1]$. Numbers indicate the number of 
-        spikes per period. 
-        {\bfseries \sffamily Right Panels:} In all these plots the 
-        crossed area corresponds to the transient, and thus it has to be
-        omitted. 
-        {\bfseries \sffamily Red} color demonstrates a case where only
-        subthreshold spikes are generated, when the frequency of the periodic
-        pulse is $\frac{1}{P_0} = 10 Hz$ and $\frac{p}{P_0} = 0.1$. 
-        {\bfseries \sffamily Blue} indicates bursts of four spike, when the
-        frequency is $\frac{1}{P_0} = 5Hz$ and $\frac{p}{P_0} = 0.6$.
-        {\bfseries \sffamily Black} shows a case of two spikes bursts. Here
-        the frequency is $\frac{1}{P_0} = 0.5 Hz$ and $\frac{p}{P_0} = 0.6$.}
+        $\frac{1}{P_0}$ in $[0.1, 15]\, \Rm{Hz}$ and ON pulse duration $p$. 
+        {\bfseries \sffamily Upper panels} Show the results of two specific
+        simulations out of the $100 \times 100$ we ran in total. The blue 
+        curve illustrates a case where bursts of four spike take place when
+        frequency is $\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$.
+        Black 
+        curve depicts a case of two spikes bursts. In this case the frequency
+        is $\frac{1}{P_0} = 0.5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$.
+        {\bfseries \sffamily Bottom panel} shows a range of frequencies
+        $\frac{1}{P_0}$ from $0.1\, \Rm{Hz}$ to $15\, \Rm{Hz}$ versus the ratio
+        $\frac{p}{P_0}$ in range $[0, 1]$. Numbers indicate the average number
+        of spikes per pulse. }
     \label{Fig:2}
@@
 
-Then we tested the transition from subthreshold to bursting oscillation via 
+Then we tested the transition from subthreshold to bursting oscillations via 
 chaos. Therefore, we used a steady current varying only its amplitude keeping
@@
 activity in the membrane voltage trace. Thus we created the hysteresis diagram
-given in Figure~\ref{Fig:4}, left panel. 
+given in Figure~\ref{Fig:4}, top panel. 
 %%
@@
     \centering
-    \includegraphics[width=1.1\textwidth]{figs/Figure4.pdf}
+    \includegraphics[width=0.6\textwidth]{figs/Figure4.pdf}
     \caption{{\bfseries \sffamily Hysteresis near the transition from the
-    subthreshold to bursting oscillation.} {\bfseries \sffamily Left panel:}
+    subthreshold to bursting oscillation.} {\bfseries \sffamily First panel:}
     Hysteresis diagram indicates a coexistence of bursting and subthreshold 
-    activity. {\bfseries \sffamily Upper right panel:} Bistability example
+    activity. {\bfseries \sffamily Second panel:} Bistability example
     shows how the same system can produce either subthreshold activity (black
-    curve) or burst activity (gray curve). {\bfseries \sffamily Middle right 
+    curve) or burst activity (gray curve). {\bfseries \sffamily Third 
     panel:} shows the activation of $H$ for the same simulation as in the upper
-    panel. {\bfseries  \sffamily Lower right panel:} Input to the model for
+    panel. {\bfseries  \sffamily Fourth panel:} Input to the model for
     this simulation is a brief current pulse at $0.25\, \Rm{\mu A /cm^2}$
@@
     \centering
-    \includegraphics[width=.7\textwidth]{figs/Figure5.pdf}
+    \includegraphics[width=.6\textwidth]{figs/Figure5.pdf}
     \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
@@
 curve represents the frequency in $\Rm{Hz}$ and the blue one the period in 
-$\Rm{sec}$. Provided results here are not as smooth as the original ones in
-\cite{wang:1994}. Unfortunately there are no many details in the original 
-article about these diagrams so we cannot reproduce them in full details. 
+$\Rm{sec}$. Black dots indicate frequency and open circles the corresponding
+period (reference implementation), respectively. Cyan crosses show the original
+data taken from \cite{wang:1994} and magenta pentagons the corresponding 
+period. The original points and the reference implementation share a similar 
+shape (and smoothness). However, there is a slight divergence for a few of 
+the points. 
 %%
 \begin{figure}[!htbp]
-    \includegraphics[width=1.1\textwidth, left]{figs/Figure6.pdf}
+    \includegraphics[width=1.0\textwidth, left]{figs/Figure6.pdf}
     \caption{{\bfseries \sffamily Frequency and period versus steady input 
@@
     from $0$ to $-2\, \Rm{\mu A/cm^2}$ affects the bursts frequency (bursting
-    mode) of the model. Black curve indicates the frequency in $\Rm{Hz}$
-    and the blue curve the period in $\Rm{sec}$. The external
-    current varies exactly as in Fig~$7$ in \cite{wang:1994} (see text for
-    more details). }
+    mode) of the model. Black curve indicates frequency (black dots) in
+    $\Rm{Hz}$ and blue curve period (white circle discs) in $\Rm{sec}$. 
+    Cyan crosses (frequency) and magenta pentagons (period) indicate original
+    data taken from \cite{wang:1994}.}
     \label{Fig:6}
@@
 integration method) are given. The reference implementation results are similar
-to the original ones. Unfortunately, we could not find any other implementation
-of the original model described in \cite{wang:1994} to compare with.
+to the original ones except maybe of Figure~\ref{2} (Fig~$1$ of
+\cite{wang:1994}), where we found some divergence from the original results.
+Furthermore, not any other implementation of \cite{wang:1994} found in order
+to be compared with the current reference implementation.
 

### 2016-08-18 code/neuron_model.py

@@ -87,7 +87,7 @@ class WangModel():
         |  :param inp_type: Input signal type ['const' - constant current
                                                'periodic' - periodic pulses]
                                                'pulse' - one pulse at the 
-                                               beggining of simulation
+                                               beginning of simulation
                                                following by a constant current.
         |  :param amplitude: Amplitude of injected current (in μA/cm²)
         |  :param pulse_ampl: Amplitude of initial pulse (input must be set to
@@ -328,7 +328,7 @@ class WangModel():
             i += 1
 
         # Store the neuron states and time in the same file if the 
-        # flag IsStore is true.
+        # flag store is true.
         v = np.array(v)
         t = np.array(t)
         v = np.hstack([t.reshape(t.shape[0], 1), v])
@@ -337,3 +337,4 @@ class WangModel():
             np.save("../data/"+vname, v)
         if self.IsStoreStim is True:
             np.save("../data/"+vname+"stim", Iapp)
+        return v

### 2016-08-18 code/plot_figures.py

@@ -29,7 +29,7 @@ from matplotlib import collections  as mc
 from skimage.feature import peak_local_max
 
 
-def plot_figure1():
+def plot_figure1_old():
     dt = 0.05
     base = '../data/'
     dopri = np.load(base+'test_dopri_V.npy')
@@ -43,56 +43,93 @@ def plot_figure1():
         tmp[spikes] = 1
         return tmp.astype('f')
 
-    M, N = 7010, 7045
+    M, N = 4200, 4205
     fig = plt.figure(figsize=(10, 10))
+    for i in range(2):
+        if i == 0:
+            M, N = 4270, 4310
+        else:
+            M, N = 6830, 6850
+        ax = fig.add_subplot(2, 1, i+1)
+        ax.plot(dopri[M:N, 0], dopri[M:N, 1], 'k', lw=2, alpha=1.0,
+                label='dopri5')
+        ax.plot(adams[M:N, 0], adams[M:N, 1], 'k--', lw=2, alpha=1.0,
+                label='Adams', zorder=0)
+        ax.plot(bdf[M:N, 0], bdf[M:N, 1], 'k-.', lw=2, alpha=1.0,
+                label='BDF', zorder=1)
+        if i == 1:
+            m = max(bdf[M:N, 1].max(), dopri[M:N, 1].max(), adams[M:N, 1].max())
+            mm = dopri[M:N, 1].max()
+            n = max(bdf[M:N, 1].argmax(), dopri[M:N, 1].argmax(),
+                    adams[M:N, 1].argmax())
+            print(m, n, mm)
+        ax.set_xlabel('Time (ms)', fontsize=16, weight='bold')
+        ax.set_ylabel('Membrane Potential (mV)', fontsize=16, weight='bold')
+        ax.get_xaxis().set_tick_params(which='both', direction='out')
+        ax.get_yaxis().set_tick_params(which='both', direction='out')
+        ticks = ax.get_xticks()
+        ticks = [i*dt for i in ticks]
+        ax.set_xticklabels(ticks, fontsize=13, weight='bold')
+        ticks = ax.get_yticks()
+        ax.set_yticklabels(ticks, fontsize=13, weight='bold')
+        handles, labels = ax.get_legend_handles_labels()
+        ax.legend(handles, labels, loc='upper right')
+        ax.spines['right'].set_visible(False)
+        ax.spines['top'].set_visible(False)
+        ax.yaxis.set_ticks_position('left')
[…]

### 2016-08-18 code/simulations.py

@@ -34,10 +34,10 @@ from neuron_model import *
 def simulate_figure1():
     print("Running test for comparing integration methods")
     base = './params/'
-    Iext = -0.6
+    Iext = -0.55
     name = ['test_dopri', 'test_adams', 'test_bdf']
 
-    simulation = WangModel(base+"params_figure1.cfg",
+    simulation = WangModel(base+"params_figure3.cfg",
                            inp_type='const',
                            amplitude=Iext)
     simulation.run(6000, name[0]+"_V", 'dopri5')
@@ -75,16 +75,15 @@ def simulate_figure2(diagram=False):
     print("Running simulation #2")
 
     if diagram is True:
-        samples = 30
+        samples = 100
         freq = np.linspace(0.1, 15, samples)
         Iext = -1.0                                   
 
         name = "Fig2A"
         idx = 0
         for i, f in enumerate(freq):
-            period = np.round(1./f, 2) * 1000
-            # period = 1./f * 1000
-            dur = np.round(np.linspace(0, period, samples), 2)
+            period = 1./f * 1000
+            dur = np.linspace(0, period, samples)
             for j, d in enumerate(dur):
                 print('Frequency: {}, Duration: {}'.format(f, d))
                 nm = 'Fig2A_V'+str(idx)
@@ -94,7 +93,8 @@ def simulate_figure2(diagram=False):
                                        duration=d,
                                        inp_type='periodic',
                                        amplitude=Iext,
-                                       chunks=100)
+                                       chunks=100,
+                                       store_stim=False)
                 simulation.run(15*period, name+"_V"+str(idx),
                                'vode',
                                'Adams')
@@ -112,7 +112,7 @@ def simulate_figure2(diagram=False):
                                inp_type='periodic',
                                amplitude=Iext,
                                chunks=100)
-        simulation.run(2000, name+"_V"+str(i), 'vode', 'Adams')
+        simulation.run(5000, name+"_V"+str(i), 'vode', 'Adams')
 
 
 def simulate_figure3():
@@ -171,7 +171,7 @@ def simulate_figure5():
 def simulate_figure6():
     print("Running simulation #6")
 
-    Iext = np.round(np.genfromtxt('../data/data1.dat'), 3)
+    Iext = np.genfromtxt('../data/data1or.dat')[:, 0]
[…]

### 2016-08-25 article/detorakis-2016.tex
4th review - corrections

@@
 \usepackage{hyperref}
+\usepackage{bm,array}
 \hypersetup{breaklinks=true,
@@
 \newcommand{\Rm}[1]{\mathrm{#1}}
+
+\newcolumntype{C}{>{\centering\arraybackslash}p{3.5em}}
 
@@
 \date{\vspace{-5mm}
-    \sffamily \small \href{mailto:gdetorak@uci.edu}{gdetorak@uci.edu}}
+    \sffamily \small \href{mailto:gdetorak@uci.edu}{gdetorak@uci.edu} \\
+    \sffamily \small \href{mailto:gdetor@protonmail.com}{gdetor@protonmail.com}}
 
@@
 
-This work introduces a reference implementation of a neuron model for
-thalamocortical relay neurons, proposed by X-J Wang, \cite{wang:1994}.
-The model is conductance-based and takes advantage of an interplay between
-a T-type calcium current and a non-specific cation sag current and thus, it
-is able to generate spindle and delta rhythms. Another feature of this model
-is the presence of an intermittent phase-locking phenomenon where action 
-potentials of sodium take place in a non-periodic manner, despite the fact
-that they are phase-locked to the periodic input current. Finally, the model
-is capable of generating tonic spike patterns. The source code of
-reference implementation is written in Python (Numpy, Scipy, Matplotlib,
-and Scikit-image).
+The present paper replicates a neuron model for thalamocortical relay neurons,
+proposed by X-J Wang, \cite{wang:1994}. The model is conductance-based and 
+takes advantage of the interplay between a T-type calcium current and a 
+non-specific cation sag current and thus, it is able to generate spindle
+and delta rhythms. Another feature of this model is the presence of an
+intermittent phase-locking phenomenon where action potentials of sodium take
+place in a non-periodic manner, despite the fact that they are phase-locked to
+the periodic input current. Finally, the model is capable of generating tonic
+spike patterns. The source code for replicating Wang's model is written in
+Python (Numpy, Scipy, Matplotlib, and Scikit-image).
 
@@
 In this section, a detailed description of the model is given following 
-the paradigm of Nordlie et al, \cite{nordlie:2009}. To this end description of
-the model, equations, parameters, and inputs are given in the form of tables.
+the paradigm of Nordlie et al, \cite{nordlie:2009}. To this end, a description
+of the model's equations, parameters, and inputs are given in the form of
+tables.
 
-Table~\ref{Table:1} provides a description of the model, Table~\ref{Table:2}
-provides information about simulations duration and temporal integration time
-steps, Table~\ref{Table:3} gives a glimpse of the input signals used in this
-work. Table~\ref{Table:4} introduces the equations of the model and finally 
-Table~\ref{Table:5} summarizes all the parameters for each figure (simulation).
-The neuron model is conductance-based consisting of four differential equations
-describing the dynamics of membrane potential and the kinetics of a T-Type
-calcium current, a Sag current channel and a Potassium channel. The rest
-currents are described by algebraic equations.
+Table~\ref{Table:1} and~\ref{Table:2} provide a description of the model and 
+information about simulations duration and time steps, respectively. 
+Table~\ref{Table:3} gives a glimpse of the input signals used in this work and
+Table~\ref{Table:4} introduces the equations of the model. Table~\ref{Table:5}
+summarizes all the parameters used in this work.
+
+The neuron model is conductance-based, consisting of four differential
+equations describing the dynamics of membrane potential and the kinetics of
+a T-Type calcium current, a Sag current channel and a Potassium channel. The
+rest of the currents are described by algebraic equations.
 %%
@@
  %%
-The reference implementation has been done in a Python class (Python $3.5.1$)
-along with Numpy (version $1.10.4$), Scipy (version $0.17.0$), Matplotlib
-($1.5.1$) and Scikit-image (version $0.12.3$). The numerical integration
-has been done using the \emph{ode} method of Scipy \emph{integrate} package.
-Three different methods have been tested in this work (\emph{dopri5}, 
-\emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{Dopri5} is the closest
-numerical method to the one used by \cite{wang:1994} (fifth-order adaptive 
-size Runge-Kutta method). \emph{BDF} and \emph{Adams} provide similar numerical
-results as the first one, but they are faster. In the current implementation
-of \cite{wang:1994} the user has the option to choose one of the three methods
-at the stage of class instantiating. All three aforementioned methods have
-been tested by comparing, spike times, amplitudes and the coefficient
-of variation using as threshold value $0\, \Rm{mV}$. Spike events were
-overlapping, and $CV \simeq 3.5$ for all three methods. A difference found 
-in the amplitude of spikes generated by the three methods. Figure~\ref{Fig:1}
-shows two histograms of error distribution (error is defined to be the
-absolute difference of spike amplitude between two methods). Top histogram 
-depicts a distribution of differences in amplitude between \emph{dopri5} 
-and \emph{Adams} methods (gray color). The bottom histogram illustrates the 
-distribution of amplitude differences between \emph{dopri5} and \emph{BDF}
-methods (red color). In the rest of the article all the results have been 
-produced by using the \emph{Adams} method.
+
+Before proceeding to the simulations we introduce a few details regarding 
+the source code and the integration method which should be accurate and fast
+enough. The implementation is a Python class (Python $3.5.1$), which uses
+Numpy (version $1.10.4$), Scipy (version $0.17.0$), Matplotlib
+($1.5.1$) and Scikit-image (version $0.12.3$) packages. The numerical
+integration is carried out using the \emph{ode} method of Scipy 
+\emph{integrate} package. Three different methods were tested in this work 
+(\emph{dopri5}, \emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{Dopri5}
+is the closest numerical method to the one used by \cite{wang:1994}
+(fifth-order adaptive size Runge-Kutta method). The user of the present 
+implementation can choose one of the three methods at the stage of class
+instantiating. For all three methods we measured spike times, membrane
+potential amplitudes and the coefficient of variation using as threshold value
+$0\, \Rm{mV}$. We found that spike event times for \emph{dopri5}, \emph{Adams}
+and \emph{BDF} coincide and they all have the same $CV \simeq 2.3608$. 
+However, the amplitude of membrane potential was different. Figure~\ref{Fig:1}
+{\bfseries \sffamily A} shows the membrane potential (in $\Rm{mV}$) produced by
+\emph{dopri5} plotted against the membrane potential generated by \emph{Adams}
+method. Figure~\ref{Fig:1} {\bfseries \sffamily B} shows the membrane potential
+of a simulation using \emph{dopri5} against another using \emph{BDF}. The 
+\emph{Adams} method has less differences in amplitude from \emph{dopri5} 
+than \emph{BDF}. Therefore, we decided to use the \emph{Adams} method 
+since it runs faster than \emph{dopri5}. 
 %%
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{figs/Figure1.pdf}
-    \caption{{\bfseries \sffamily Amplitude difference distribution} 
-    {\bfseries \sffamily Top histogram} depicts the distribution of differences
-    in spikes amplitudes between \emph{dopri5} and \emph{Adams} integration
-        methods of Scipy \emph{integrate} package (black color). 
-    {\bfseries \sffamily Bottom histogram} illustrates the distribution of
-    differences in spikes amplitudes between \emph{dopri5} and \emph{BDF}
-    methods (red color). In both cases the number of bins is $30$. 
-    }
+    \includegraphics[width=0.9\textwidth]{figs/Figure1.pdf}
+    \caption{{\bfseries \sffamily Amplitude differences} The membrane potential
+    ($\Rm{mV}$) is generated using \emph{dopri5} integration method and is 
+    plotted against ({\bfseries \sffamily A}) the one generated by \emph{Adams}
+    method and ({\bfseries \sffamily B}) \emph{BDF} method.}
     \label{Fig:1}
@@
 generation i$7$ processor, $8\, \Rm{GB}$ of physical memory and running Arch
-Linux.
-The total execution time of all simulations was $526$ minutes and the peak
-consumed memory was $465\, \Rm{MB}$\footnote{Python memory profiler used 
+Linux. The total execution time of all simulations was $526$ minutes and the
+peak consumed memory was $465\, \Rm{MB}$\footnote{Python memory profiler used
     (\url{https://pypi.python.org/pypi/memory_profiler}).}\@. 
@@
 %%
-\begin{landscape}
 \begin{table}[!htbp]
     \centering
-    \begin{tabular}{cccccccc}
+    %\begin{tabular}{cccccccc}
+    \begin{tabular}{CCCCCCCC}
         \thickhline
         \multicolumn{8}{c}{Model Parameters} \\\thickhline
-        Figure & $V_0\, (\Rm{mV})$ & $g_T\, (\Rm{mS/cm^2})$ & $\theta_h\,
-        (\Rm{mV})$ & $k_h\, (\Rm{mV^{-1}})$ & $\sigma_{Na}\, (\Rm{mV})$ &
-        $g_L\, (\Rm{mS/cm^2})$ & $V_L\, (\Rm{mV})$ \\\rowcolor{LightGray}
+        Figure & \begin{tabular}[x]{@{}c@{}} $V_0$ \\ $(\Rm{mV})$ \end{tabular}
+            & \begin{tabular}[x]{@{}c@{}} $g_T$ \\ $(\Rm{mS/cm^2})$ \end{tabular}
+            & \begin{tabular}[x]{@{}c@{}} $\theta_h$ \\ $(\Rm{mV})$ \end{tabular}& 
+            \begin{tabular}[x]{@{}c@{}} $k_h$ \\ $(\Rm{mV^{-1}})$ \end{tabular}&
+            \begin{tabular}[x]{@{}c@{}} $\sigma_{Na}$ \\ $(\Rm{mV})$ \end{tabular}&
+            \begin{tabular}[x]{@{}c@{}} $g_L$ \\ $(\Rm{mS/cm^2})$ \end{tabular}&
+            \begin{tabular}[x]{@{}c@{}} $V_L$ \\ $(\Rm{mV})$ \end{tabular}\\\rowcolor{LightGray}
         \thickhline
-        $1$ & $-74$ & $1$ & $-79$ & $-5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
-        $2$ & $-74$ & $1$ & $-81$ & $6.25$ & $3$ & $0.1$ & $-72$ \\\rowcolor{LightGray}
-        $3$ & $-74$ & $0.3$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
+        $1$ & $-74$ & $1$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
+        $2$ & $-74$ & $0.3$ & $-81$ & $6.25$ & $3$ & $0.1$ & $-72$ \\\rowcolor{LightGray}
+        $3$ & $-74$ & $1$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{Gray}
         $4$ & $-72$/$-64$ & $1.0$ & $-79$ & $5$ & $6$ & $0.12$ & $-70$ \\\rowcolor{LightGray}
-        $5$ & $-74$ & $0.3$ & $-75$ & $5$ & $6$ & $0.08$ & $-70$ \\\rowcolor{Gray}
-        $6$ & $-74$ & $1$/$0.7$ & $-79$ & $5$ & $6$ & $0.1$/$0.04$ & $-70$ \\\rowcolor{LightGray}
+        $5$ & $-74$ & $1.0$ & $-75$ & $5$ & $6$ & $0.08$ & $-70$ \\\rowcolor{Gray}
+        $6$ & $-74$ & $1$/$0.7$ & $-79$ & $5$ & $6$ & $0.12$/$0.04$ & $-70$ \\\rowcolor{LightGray}
         $7$ & $-72$ & $0.3$/$0.25$ & $-81$ & $6.25$ & $3$ & $0.1$ & $-72$ \\
         \thickhline
-        \multicolumn{8}{l}{Common Parameters} \\\rowcolor{Gray}
-        \multicolumn{8}{c}{$C_m = 1\, \Rm{\mu F/cm^2}$, $\phi_h = 2$,
+        \multicolumn{8}{l}{Common Parameters} \\\rowcolor{LightGray}
+        \thickhline
+        \multicolumn{8}{p{36.5em}}{
+        $C_m = 1\, \Rm{\mu F/cm^2}$, $\phi_h = 2$,
         $V_{\Rm{Ca}} = 120\, \Rm{mV}$, $\phi_H = 1$, $g_h = 0.04\,
-        \Rm{mS/cm^2}$, $V_h = -40\, \Rm{mV}$, $g_{\Rm{K}} = 30\, \Rm{mS/cm^2}$,
-        $V_{\Rm{K}} = -80\, \Rm{mV}$, $g_{\Rm{Ca}} = 42\, \Rm{mS/cm^2}$} \\\rowcolor{Gray}
-        \multicolumn{8}{c}{$V_{\Rm{Ca}} = 55\, \Rm{mV}$, $\phi_{n} = 28.5$,
+        \Rm{mS/cm^2}$, $V_h = -40\, \Rm{mV}$,         
+        $g_{\Rm{K}} = 30\, \Rm{mS/cm^2}$,
+        $V_{\Rm{K}} = -80\, \Rm{mV}$, $g_{\Rm{Na}} = 42\,\Rm{mS/cm^2}$, 
+        $V_{\Rm{Na}} = 55\, \Rm{mV}$, $\phi_{n} = \frac{200}{7}$,}\\\rowcolor{LightGray}
+        \multicolumn{8}{p{36.5em}}{
         $\sigma_{\Rm{K}} = 10\, \Rm{mV}$, $V_{\Rm{Na(P)}} = 55\, \Rm{mV}$,
@@
 \end{table}
-\end{landscape}
 %%
@@
 
-We simulated the model described in Table~\ref{Table:4} using the parameters
-given in Table~\ref{Table:5} and the corresponding input (see 
-Table~\ref{Table:3}). First, we examined what is the response of the reference 
+We simulated the model described in Table~\ref{Table:4} using parameters
+given in Table~\ref{Table:5} and the corresponding inputs (see 
+Table~\ref{Table:3}). First, we examined the response of the present 
 implementation to rhythmic hyperpolarization. In \cite{wang:1994} this is 
-illustrated in Fig~$1$\footnote{From now and then original article's figures
-    will be referred as Fig}. 
-Thus, we applied a periodic current pulse of $-1\,\Rm{\mu A/cm^2}$ amplitude at
-several different frequencies ($\frac{1}{P_0}$ is the frequency in $\Rm{Hz}$
-and $P_0$ is the corresponding period in $\Rm{ms}$) ranging from 
-$0.1\,\Rm{Hz}$ to $15\,\Rm{Hz}$ with a resolution (discretization) of $100$ 
-points. The same number of samples used for discretizing the duration of the
-pulse for each frequency.
+illustrated in Wang's Fig~$1$. Thus, we applied a periodic current pulse of
+$-1\,\Rm{\mu A/cm^2}$ amplitude at several different frequencies
+($\frac{1}{P_0}$ is the frequency in $\Rm{Hz}$ and $P_0$ is the corresponding
+period in $\Rm{ms}$) ranging from $0.1\,\Rm{Hz}$ to $15\,\Rm{Hz}$ with a 
+resolution (discretization) of $100$ samples. The same number of samples used 
+for discretizing the duration (ON/OFF) of the pulse for each frequency. The ON
+duration of a pulse is defined by the value of $p$ and the OFF duration as
+$P_0 - p$.
 
-Upper panel shows two different simulations of the reference implementation at
-specific frequencies ($5,\, 0.5\, \Rm{Hz}$) and ratios $p/P_0$ ($0.6, 0.6$),
-respectively (following the upper panels of Fig~$1$ in \cite{wang:1994}).
-These panels are comparable to the ones presented in \cite{wang:1994} and the 
-timescales are exactly the same ($400\, \Rm{ms}$ for the left sub-panel and
-$4\, \Rm{s}$ for the second one). However, numbers regarding pulse ON duration
-($p$) in Fig~$1$ of \cite{wang:1994} are not correct. For instance, take the
-frequency to be $\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$ then
-the pulse ON duration must be $p = 120\, \Rm{ms}$ and not $160\, \Rm{ms}$.
-
-Bottom panel illustrates the total results of the $100 \times 100$ simulations
-we ran.  In order to create this figure, we counted the number of spikes
-generated per pulse, both supra-threshold and sub-threshold, and then we 
-computed the mean (this procedure is not described in a solid way in
-\cite{wang:1994}, however it is described in \cite{mccormick:1990}).  
-This figure is different from \cite{wang:1994}, where the shaded area that 
-contains an average number of spikes below $1.0$ (but not zero) is larger
-than in \cite{wang:1994}. Unfortunately, information regarding Fig~$1$
-is limited in \cite{wang:1994} in order to be able to follow a specific recipe
-and try to reproduce exactly the same figure.
+The results are depicted in Figure~\ref{Fig:2}. Upper panel shows two different
+simulations of the present implementation at specific frequencies 
+($5,\, 0.5\, \Rm{Hz}$) and ratios $p/P_0$ ($0.6, 0.6$), respectively (following
+the upper panels of Fig~$1$ in \cite{wang:1994}). The results in
+Figure~\ref{Fig:2} are comparable to the ones presented in \cite{wang:1994}
+and the timescales are exactly the same ($400\, \Rm{ms}$ for the left sub-panel
+and $4\, \Rm{s}$ for the second one). However, numbers regarding pulse ON 
+duration ($p$) in Wang's Fig~$1$ are possibly an error. As an example, assume
+that the frequency is $\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$
+then the pulse ON duration must be $p = 120\, \Rm{ms}$ and not $160\, \Rm{ms}$ 
+(that is reported in \cite{wang:1994}). Bottom panel illustrates the total 
+results of $100 \times 100$ simulations. In order to create this figure, we
+counted the number of spikes, both supra- and sub-threshold ones, generated per
+pulse and then we computed the mean (this procedure is not described 
+in full detail in \cite{wang:1994}, however more details can be found in
+\cite{mccormick:1990}). Figure~\ref{Fig:2} differs from the one in 
+\cite{wang:1994}, to the point that the shaded area that contains an average
+number of spikes below $1.0$ (but not 
+zero) is larger than in \cite{wang:1994}. Unfortunately, there are not enough
+details regarding Fig~$1$ in \cite{wang:1994} to make a faithful reproduction
+of that figure.
 
@@
     \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
-        We stimulate the model with a periodic pulse with varying frequency
-        $\frac{1}{P_0}$ in $[0.1, 15]\, \Rm{Hz}$ and ON pulse duration $p$. 
-        {\bfseries \sffamily Upper panels} Show the results of two specific
-        simulations out of the $100 \times 100$ we ran in total. The blue 
-        curve illustrates a case where bursts of four spike take place when
-        frequency is $\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$.
-        Black 
-        curve depicts a case of two spikes bursts. In this case the frequency
-        is $\frac{1}{P_0} = 0.5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$.
-        {\bfseries \sffamily Bottom panel} shows a range of frequencies
-        $\frac{1}{P_0}$ from $0.1\, \Rm{Hz}$ to $15\, \Rm{Hz}$ versus the ratio
-        $\frac{p}{P_0}$ in range $[0, 1]$. Numbers indicate the average number
-        of spikes per pulse. }
+    We stimulate the model given in Table~\ref{Table:4} using a periodic pulse
+    with frequency ($\frac{1}{P_0}$) varying from $0.1$ to $15\, \Rm{Hz}$.
+    {\bfseries \sffamily Upper panels} show the results of two simulations out
+    of $100 \times 100$. The blue curve indicates bursts of four spikes
+    ($\frac{1}{P_0} = 5\, \Rm{Hz}$ and $\frac{p}{P_0} = 0.6$). Black curve
+    depicts a case of two spikes bursts ($\frac{1}{P_0} = 0.5\, \Rm{Hz}$ and
+    $\frac{p}{P_0} = 0.6$).
+    {\bfseries \sffamily Bottom panel} shows the average number of spikes (sub-
+    or supra-threshold) measured per pulse. Frequency varies in the interval 
+    $[0.1, 15]\, \Rm{Hz}$ and the ratio $\frac{p}{P_0}$ is in interval $[0, 1]$.
+    Numbers signify the average number of spikes per pulse.}
     \label{Fig:2}
@@
 
-Then we tested the transition from subthreshold to bursting oscillations via 
-chaos. Therefore, we used a steady current varying only its amplitude keeping
-all the other parameters fixed. We used eight different values (exactly the
-same as in \cite{wang:1994}),
+The next simulation investigated the transition from subthreshold to bursting
+oscillations via chaos. This is illustrated in Fig~$3$ in \cite{wang:1994}. In
+this case a steady current is applied to the model described in
+Table~\ref{Table:4}. In \cite{wang:1994} Wang used eleven different
+amplitudes for the constant current in order to explore the model's behavior
+repertoire. Hence, 
 $$I_{\text{app}} = 
 3.0, 0.0, -0.45, -0.455, -0.47, -0.55, -0.6, -0.8, -1.3, -1.4, -2.0\, 
-\Rm{\mu A /cm^2}.$$ Results are depicted in Figure~\ref{Fig:3}. This figure
-corresponds to Fig~$3$ of \cite{wang:1994}. In this case, there was no
-difference between the reference implementation and the original one. 
+\Rm{\mu A /cm^2}.$$
+The results for these simulations are depicted in Figure~\ref{Fig:3}. In this
+case, there was no difference between the present implementation and the
+original one.
 %%
@@
     \includegraphics[width=1.\textwidth]{figs/Figure3.pdf}
-    \caption{{\bfseries \sffamily Dynamic behavior of neuron.} Here we
-    simulate the model with parameters given in Table~\ref{Table:5}. The
-    only parameter that varies from panel to panel in this figure is the
-    external current $I_{\text{ext}}$. These results correspond to Fig~$3$ 
-    of \cite{wang:1994}.}
+    \caption{{\bfseries \sffamily Dynamic behavior of neuron.} 
+    Eleven different external constant currents $I_{\text{app}} = 
+    3.0, 0.0, -0.45, -0.455, -0.47, -0.55, -0.6, -0.8, -1.3, -1.4, -2.0\, 
+    \Rm{\mu A /cm^2}$ have been applied as inputs to the model described by
+    equations given in Table~\ref{Table:4}. The behavior of the model spans
+    from tonic spiking to spikes bursts, subthreshold activity and no activity 
+    at all.}
     \label{Fig:3}
@@
 
-The next simulation we performed reproduces results related to hysteresis as
-in Fig~$4$ of \cite{wang:1994}. First, we run a simulation where the input 
-current $I_{ext}$ varied from $-0.433\, \Rm{\mu A / cm^2}$ to $-0.55 \mu A /cm^2$.
-At every iteration, we detected if there is subthreshold and/or suprathreshold 
-activity in the membrane voltage trace. Thus we created the hysteresis diagram
-given in Figure~\ref{Fig:4}, top panel. 
+\cite{wang:1994} has showed that the current model expresses a hysteresis
+(see Wang's Fig~$4$). In order to reproduce that case we simulated
+the model using a constant external current $I_{app}$. The simulation starts
+by applying a current with an amplitude of $-0.433\, \Rm{\mu A / cm^2}$ and 
+continues by increasing the amplitude up to $-0.55 \mu A /cm^2$. At every 
+iteration, sub- and/or supra-threshold activity of the membrane potential is 
+detected. The results are given in Figure~\ref{Fig:4} {\bfseries \sffamily A}. 
 %%
@@
     \caption{{\bfseries \sffamily Hysteresis near the transition from the
-    subthreshold to bursting oscillation.} {\bfseries \sffamily First panel:}
+    subthreshold to bursting oscillation.} ({\bfseries \sffamily A})
     Hysteresis diagram indicates a coexistence of bursting and subthreshold 
-    activity. {\bfseries \sffamily Second panel:} Bistability example
+    activity. ({\bfseries \sffamily B}) Bistability example
     shows how the same system can produce either subthreshold activity (black
-    curve) or burst activity (gray curve). {\bfseries \sffamily Third 
-    panel:} shows the activation of $H$ for the same simulation as in the upper
-    panel. {\bfseries  \sffamily Fourth panel:} Input to the model for
-    this simulation is a brief current pulse at $0.25\, \Rm{\mu A /cm^2}$
-    (black curve) and $-1.2\, \Rm{\mu A/cm^2}$ (gray curve) followed by a
-    steady current at $-0.47\, \Rm{\mu A/cm^2}$ for both simulations. }
+    curve) or burst activity (gray curve). ({\bfseries \sffamily C}) shows the
+    activation of $H$ for the same simulation as in the 
+    panel above. ({\bfseries  \sffamily D}) A brief current pulse at
+    $0.25\, \Rm{\mu A /cm^2}$ (black curve) or $-1.2\, \Rm{\mu A/cm^2}$ (gray
+    curve) followed by a steady current at $-0.47\, \Rm{\mu A/cm^2}$ is used 
+    as external input to the model. }
     \label{Fig:4}
@@
 %%
-In addition, we tested the bistability of the current model using the same 
-protocol as in \cite{wang:1994}. According to that protocol, we first apply 
-a brief pulse (for $100\, \Rm{ms}$) of $0.25\, \Rm{\mu A /cm^2}$ and
-$-1.2\, \Rm{\mu A/cm^2}$ followed by a steady current at $-0.47\, 
-\Rm{\mu A /cm^2}$. This causes the state of the model to switch from a purely 
-subthreshold activity pattern to a mixed sub- and suprathreshold activity
-pattern. The results are given in Figure~\ref{Fig:4}, right panel.
+In addition, the model expresses a bistability. This was tested by applying the
+same protocol as in \cite{wang:1994}. Thus first a brief pulse 
+(for $100\, \Rm{ms}$) of $0.25\, \Rm{\mu A /cm^2}$ (black line in
+Figure~\ref{Fig:4} {\bfseries \sffamily C}) or $-1.2\, \Rm{\mu A/cm^2}$ (gray
+line in Figure~\ref{Fig:4} {\bfseries \sffamily C}) is applied to the model
+followed by a steady current at $-0.47\, \Rm{\mu A /cm^2}$. Such a protocol
+causes the state of the model to switch from purely sub-threshold to mixed
+sub- and suprathreshold. These results are illustrated in
+Figure~\ref{Fig:4} {\bfseries \sffamily B}. In
+the same figure panels {\bfseries \sffamily C} and {\bfseries \sffamily D}
+depict the input to the model and the $H$ of the Sag channel, respectively.
 
-Another interesting behavior of the model is the development of a ``spiral''
-chaos (see Fig~$6$ in \cite{wang:1994}). The reference implementation is 
-capable of generating similar behavior as Figure~\ref{Fig:5} shows. In order
-to get these results we applied a constant external current with an amplitude
-of $-0.95\, \Rm{\mu A / cm^2}$. 
+Another interesting behavior of the model is the development of ``spiral''
+chaos (see Fig~$6$ in \cite{wang:1994}). In this case the external input 
+is constant and its amplitude is set to $-0.95\, \Rm{\mu A / cm^2}$. The
+results for this case are shown in Figure~\ref{Fig:5}, and the current model
+is capable of generating ``spiral'' chaos as in \cite{wang:1994}. 
 %%
@@
     \includegraphics[width=.6\textwidth]{figs/Figure5.pdf}
-    \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
-    we show that the model is capable of generating ``spiral chaos''
-    as it has been shown in Fig~$6$ in \cite{wang:1994}.
-    The external current in this simulation is constant ($-0.95\Rm{\mu A/cm^2}$). 
-    {\bfseries \sffamily Top panel} shows the phase portrait of the 
-    membrane potential and the Sag current ($V$ vs $h$).
+    \caption{{\bfseries \sffamily ``Spiral Chaos''.} It is generated when a
+    constant current with amplitude $-0.95\Rm{\mu A/cm^2}$ is injected to the 
+    model described in Table~\ref{Table:4} and using the parameters from
+    Table~\ref{Table:5}. {\bfseries \sffamily Top panel} shows the phase
+    portrait of the membrane potential and the Sag current ($V$ vs $h$).
     {\bfseries \sffamily Middle panel} illustrates the membrane potential ($V$)
@@
 
-The next simulation investigates how the bursting frequency is affected by 
-the injected external current (see Fig~$7$ in \cite{wang:1994}, two bursting
-modes). To address this issue we captured data\footnote{Data are available in
-    the accompanying github repository of the present article.} values for
-the external current from Fig~$7$ of the original article (dots in Fig $7$,
-pg. $27$ in \cite{wang:1994}) using a software called 
-PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
-The results from our simulations are illustrated in Figure~\ref{Fig:6}. Black
-curve represents the frequency in $\Rm{Hz}$ and the blue one the period in 
-$\Rm{sec}$. Black dots indicate frequency and open circles the corresponding
-period (reference implementation), respectively. Cyan crosses show the original
-data taken from \cite{wang:1994} and magenta pentagons the corresponding 
-period. The original points and the reference implementation share a similar 
-shape (and smoothness). However, there is a slight divergence for a few of 
-the points. 
+The bursting frequency of the model can be affected by the injected external
+current (see Fig~$7$ in \cite{wang:1994}, two bursting modes). To verify that
+the present implementation is capable of producing similar results we 
+captured data\footnote{Data are available in the accompanying github repository
+of the present article.} values for the external current from Fig~$7$ of the 
+original article (dots in Fig $7$, pg. $27$ in \cite{wang:1994}) using a 
+software called PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
+So for every data point (external current amplitude), we ran a simulation and
+the bursting frequency of the membrane potential was measured. The results are
+illustrated in Figure~\ref{Fig:6}. The black curve shows the frequency in 
+$\Rm{Hz}$ and the blue curve shows the period in $\Rm{sec}$. Black dots indicate
+the frequency and open circles the corresponding period (present implementation),
+respectively. Cyan crosses show the original data taken from \cite{wang:1994} 
+and magenta pentagons the corresponding period. It is apparent that the
+present implementation is able to reproduce Wang's Fig~$7$, despite the fact
+that some of the original points do not match the exactly the ones computed
+using the reference implementation code. 
 %%
@@
     \caption{{\bfseries \sffamily Frequency and period versus steady input 
-    current.} This figure illustrates how the external current (ranging
+    current.} This figure illustrates how an external current (ranging
     from $0$ to $-2\, \Rm{\mu A/cm^2}$ affects the bursts frequency (bursting
@@
     Cyan crosses (frequency) and magenta pentagons (period) indicate original
-    data taken from \cite{wang:1994}.}
+    data extracted from \cite{wang:1994}.}
     \label{Fig:6}
@@
 
-The last simulation is related to Fig~$2$ of \cite{wang:1994} and to an
-``intermittent'' phase-locking phenomenon. A periodic pulse with frequency 
-$10\,\Rm{Hz}$ and ratio $\frac{p}{P_0} = 0.8$ is applied to the model as
-external input current. The response of the model is registered and sub- and
-supra-threshold spikes are counted. We applied six different current amplitudes
-and two different values for the T-type calcium conductance. 
+Fig~$2$ of \cite{wang:1994} illustrates an ``intermittent'' phase-locking
+phenomenon, generated by integrating the model described in Table~\ref{Table:4}
+and using a periodic pulse with frequency $10\,\Rm{Hz}$ and ratio 
+$\frac{p}{P_0} = 0.8$ as external current. We applied the same protocol on 
+the current implementation by using six different amplitude values and two 
+different T-type calcium conductance values. 
 $I_{app} = -1.4, -1.5, -1.6\, \Rm{\mu A /cm^2}$ and $g_T = 0.3\, \Rm{mS/cm^2}$
 $I_{app} = -1.2, -1.5, -1.8\, \Rm{\mu A/cm^2}$ and  $g_T = 0.25\,
-\Rm{mS/cm^2}$. Figure~\ref{Fig:7} shows $2$ seconds of membrane potential 
-along with symbolic patterns of $0$ (dark circles) and $1$ (green line
-segments). Results given in Figure~\ref{Fig:7} are similar to the ones of
-Fig~$2$A and Fig~$2$B in \cite{wang:1994}.
+\Rm{mS/cm^2}$. After counting sub- and supra-threshold spikes, we computed the
+symbolic patterns for every simulation. Figure~\ref{Fig:7} shows $2$ seconds
+of membrane potential for every simulation along with symbolic patterns of $0$
+(dark circles) and $1$ (green line segments). Results given in 
+Figure~\ref{Fig:7} are similar to the ones of Fig~$2$A and Fig~$2$B.
 %%
@@
 
-\section{Conclusion}\label{conclusion}
+\section{Conclusion}
+\label{conclusion}
 
 A conductance-based model for relay thalamocortical neurons proposed 
-by \cite{wang:1994} was implemented in Python. The model tested 
-thoroughly in several examples taken from the original article. 
-In general, the original model was easy to be implemented since all the
-equations and the most of the parameters (except the initial time step of the
-integration method) are given. The reference implementation results are similar
-to the original ones except maybe of Figure~\ref{2} (Fig~$1$ of
-\cite{wang:1994}), where we found some divergence from the original results.
-Furthermore, not any other implementation of \cite{wang:1994} found in order
-to be compared with the current reference implementation.
+by \cite{wang:1994} was implemented in Python. The present model was tested 
+thoroughly by using several examples from the original article. In general, 
+the original model is easily implemented since all the equations and most
+of the parameters (except the initial time step of the integration method) are
+given. However, in some experimental protocols important information is
+missing (e.g.\@ Wang's Fig~$1$ in \cite{wang:1994}, where our results do not
+match exactly the original ones), leading to a less faithful reproduction of
+those results. Furthermore, to the best of our knowledge, no any other
+implementation of \cite{wang:1994} is available the thus we cannot compare
+the current implementation to any other.
 

### 2016-08-25 code/neuron_model.py

@@ -296,7 +296,7 @@ class WangModel():
             Iapp = np.ones((t_sim, )) * self.amplitude
         elif self.inp_type == "pulse":
             Iapp = np.zeros((t_sim, ))
-            dur = self.P / dt
+            dur = int(self.P / dt)
             Iapp[:dur] = self.pulse_amplitude
             Iapp[dur:] = self.amplitude
         elif self.inp_type == "periodic":

### 2016-08-25 code/plot_figures.py

@@ -29,90 +29,44 @@ from matplotlib import collections  as mc
 from skimage.feature import peak_local_max
 
 
-def plot_figure1_old():
-    dt = 0.05
+def plot_figure1():
     base = '../data/'
-    dopri = np.load(base+'test_dopri_V.npy')
-    adams = np.load(base+'test_adams_V.npy')
-    bdf = np.load(base+'test_bdf_V.npy')
 
-    def spk(x):
-        tmp = np.zeros((x.shape[0], ))
-        spikes = peak_local_max(x, min_distance=10)
-        spikes = spikes[x[spikes] > 0]
-        tmp[spikes] = 1
-        return tmp.astype('f')
+    data = []
+    data.append(np.load(base+'test_dopri_V.npy')[:, 1])   # dopri5
+    data.append(np.load(base+'test_adams_V.npy')[:, 1])   # adams
+    data.append(np.load(base+'test_bdf_V.npy')[:, 1])     # BDF
 
-    M, N = 4200, 4205
-    fig = plt.figure(figsize=(10, 10))
+    labels = ['Dopri5', 'Adams', 'BDF']
+    case = ['A', 'B']
+
+    fig = plt.figure(figsize=(15, 7))
+    fig.subplots_adjust(wspace=0.5, hspace=0.5)
     for i in range(2):
-        if i == 0:
-            M, N = 4270, 4310
-        else:
-            M, N = 6830, 6850
-        ax = fig.add_subplot(2, 1, i+1)
-        ax.plot(dopri[M:N, 0], dopri[M:N, 1], 'k', lw=2, alpha=1.0,
-                label='dopri5')
-        ax.plot(adams[M:N, 0], adams[M:N, 1], 'k--', lw=2, alpha=1.0,
-                label='Adams', zorder=0)
-        ax.plot(bdf[M:N, 0], bdf[M:N, 1], 'k-.', lw=2, alpha=1.0,
-                label='BDF', zorder=1)
-        if i == 1:
-            m = max(bdf[M:N, 1].max(), dopri[M:N, 1].max(), adams[M:N, 1].max())
-            mm = dopri[M:N, 1].max()
-            n = max(bdf[M:N, 1].argmax(), dopri[M:N, 1].argmax(),
-                    adams[M:N, 1].argmax())
-            print(m, n, mm)
-        ax.set_xlabel('Time (ms)', fontsize=16, weight='bold')
-        ax.set_ylabel('Membrane Potential (mV)', fontsize=16, weight='bold')
[…]

### 2016-08-25 code/simulations.py

@@ -40,9 +40,9 @@ def simulate_figure1():
     simulation = WangModel(base+"params_figure3.cfg",
                            inp_type='const',
                            amplitude=Iext)
-    simulation.run(6000, name[0]+"_V", 'dopri5')
-    simulation.run(6000, name[1]+"_V", 'vode', 'Adams')
-    simulation.run(6000, name[2]+"_V", 'vode', 'BDF')
+    simulation.run(3000, name[0]+"_V", 'dopri5')
+    simulation.run(3000, name[1]+"_V", 'vode', 'Adams')
+    simulation.run(3000, name[2]+"_V", 'vode', 'BDF')
 
     base = '../data/'
     def spk(x):

### 2016-08-26 article/detorakis-2016.tex
New comments addressed

@@
 potential amplitudes and the coefficient of variation using as threshold value
-$0\, \Rm{mV}$. We found that spike event times for \emph{dopri5}, \emph{Adams}
-and \emph{BDF} coincide and they all have the same $CV \simeq 2.3608$. 
-However, the amplitude of membrane potential was different. Figure~\ref{Fig:1}
-{\bfseries \sffamily A} shows the membrane potential (in $\Rm{mV}$) produced by
-\emph{dopri5} plotted against the membrane potential generated by \emph{Adams}
-method. Figure~\ref{Fig:1} {\bfseries \sffamily B} shows the membrane potential
+$0\, \Rm{mV}$. \emph{Dopri5} and \emph{Adams} methods differ in one spike 
+event time whilst \emph{dopri5} and \emph{BDF} have a difference of 37 spike
+events. However, all three methods (\emph{dopri5}, \emph{Adams} and \emph{BDF})
+have the same $CV \simeq 2.3608$. The most stroking difference found for the
+amplitude of membrane potential. Figure~\ref{Fig:1} {\bfseries \sffamily A}
+shows the membrane potential (in $\Rm{mV}$) produced by \emph{dopri5} plotted 
+against the membrane potential generated by \emph{Adams} method.
+Figure~\ref{Fig:1} {\bfseries \sffamily B} shows the membrane potential
 of a simulation using \emph{dopri5} against another using \emph{BDF}. The 
@@
         \thickhline
-        \multicolumn{3}{c}{Simulation Time} \\ \thickhline
+        \multicolumn{3}{c}{Simulated Time} \\ \thickhline
         Figure & Simulation Time ($\Rm{s}$) & Integration Step ($\Rm{ms}$) \\ 
@@
         $1$ & $6$ & $0.05$ \\ \rowcolor{Gray}
-        $2$ & $15\times period$, $5$ & $0.05$ \\ \rowcolor{LightGray}
+        $2$ & $15\times \text{period}$, $5$ & $0.05$ \\ \rowcolor{LightGray}
         $3$ & $6$ & $0.05$ \\ \rowcolor{Gray} 
@@
     ($\Rm{mV}$) is generated using \emph{dopri5} integration method and is 
-    plotted against ({\bfseries \sffamily A}) the one generated by \emph{Adams}
-    method and ({\bfseries \sffamily B}) \emph{BDF} method.}
+    plotted (black dots) against ({\bfseries \sffamily A}) the one generated
+    by \emph{Adams} method and ({\bfseries \sffamily B}) \emph{BDF} method.
+    The light shaded diagonal area indicates a $2 \, \Rm{mV}$ difference.}
     \label{Fig:1}
@@
 that some of the original points do not match the exactly the ones computed
-using the reference implementation code. 
+using the reference implementation code. Furthermore, it is worth mentioning 
+that for properly replicate Figure~\ref{Fig:6} we had to run quite long
+simulations (see Table~\ref{Table:2}). If the signal is too short the results 
+can be quite different from the ones given in \cite{wang:1994}.
 %%

### 2016-08-26 code/plot_figures.py

@@ -44,7 +44,12 @@ def plot_figure1():
     fig.subplots_adjust(wspace=0.5, hspace=0.5)
     for i in range(2):
         ax = fig.add_subplot(1, 2, i+1, aspect=1)
-        ax.plot(data[0], data[i+1], 'k')
+        ax.plot(data[0], data[i+1], 'k.')
+        plt.fill_between([-80, 20], [-82, 18], [-78, 22],
+                          edgecolor='none',
+                          facecolor='k', alpha=0.2)
+        ax.set_xlim([-80, 20])
+        ax.set_ylim([-80, 20])
         ax.spines['right'].set_visible(False)
         ax.spines['top'].set_visible(False)
         ax.yaxis.set_ticks_position('left')
@@ -57,8 +62,6 @@ def plot_figure1():
         ax.set_ylabel(labels[i+1]+' V (mV)', fontsize=17, weight='bold')
         ax.get_xaxis().set_tick_params(which='both', direction='out')
         ax.get_yaxis().set_tick_params(which='both', direction='out')
-        if i == 1:
-            ax.spines['left'].set_visible(False)
         if i == 0:
             ax.yaxis.set_ticks_position('left')
             ax.xaxis.set_ticks_position('bottom')

### 2016-08-26 code/simulations.py

@@ -58,14 +58,17 @@ def simulate_figure1():
     spks1 = spk(dopri[:, 1]).astype('i')
     spks2 = spk(adams[:, 1]).astype('i')
     spks3 = spk(bdf[:, 1]).astype('i')
+    s1 = np.where(spks1 > 0)[0]
+    s2 = np.where(spks2 > 0)[0]
+    s3 = np.where(spks3 > 0)[0]
     isi1 = np.diff(dopri[spks1==1, 0])
     isi2 = np.diff(adams[spks2==1, 0])
     isi3 = np.diff(bdf[spks3==1, 0])
     CV1 = np.mean(isi1) / np.std(isi1)
     CV2 = np.mean(isi2) / np.std(isi2)
     CV3 = np.mean(isi3) / np.std(isi3)
-    print('sum(Spikes dopri5 - spikes Adams):  ',(dopri[spks1,0]-adams[spks2,0]).sum())
-    print('sum(Spikes dopri5 - spikes BDF):  ',(dopri[spks1,0]-bdf[spks3,0]).sum())
+    print('sum(Spikes dopri5 - spikes Adams):  ', np.abs((s1 - s2).sum()))
+    print('sum(Spikes dopri5 - spikes BDF):  ', np.abs((s1 - s3).sum()))
     print('CV dopri5:', CV1)
     print('CV Adams:', CV2)
     print('CV BDF:', CV3)

### 2016-10-28 article/VRostami_JIto_MDenker_SGruen_2016.tex

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Vahid Rostami,
      Junji Ito,,
      Michael Denker,,
      Sonja Grün,
  },
            pdftitle=
{
[Re] Spike Synchronization and Rate Modulation Differentially Involved in
Motor Cortical Function
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{VRostami_JIto_MDenker_SGruen_2016.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Nov 2016 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Spike Synchronization and Rate Modulation Differentially Involved in
Motor Cortical Function}

    \usepackage{authblk}
                        \author[1]{Vahid Rostami}
                    \author[1]{Junji Ito,}
                    \author[1]{Michael Denker,}
                    \author[1,2]{Sonja Grün}
                            \affil[1]{Inst. of Neuroscience \& Medicine (INM-6) and Inst. for Advanced
Simulation (IAS-6), JARA Brain Institute I, Jülich Research Center,
Jülich, Germany}
                    \affil[2]{Theoretical Systems Neurobiology, RWTH Aachen University, Aachen,
Germany}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:v.rostami@fz-juelich.de}{v.rostami@fz-juelich.de}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Nov, 1, 2016\\
  {\bfseries Accepted}  Nov, 1, 2016\\
  {\bfseries Published} Nov, 1, 2016\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Spike synchronization and rate modulation differentially involved in
motor cortical function. Alexa Riehle, Sonja Grün, Markus Diesmann, and
Ad Aertsen (1997) Science 278:1950-1953. DOI:10.1126/science.278.5345.19
50
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Understanding how information is processed by networks of neurons in the
brain is a major goal in neuroscience. There has been a long-standing
debate in the community on the contribution of the firing activity of
individual neurons or populations of neurons to information processing:
one perspective is that the firing rate of the neurons, i.e.~the number
of spikes per time unit, represents information or is relevant for the
processing. Another perspective is that neurons communicate through
precisely timed coordination of spiking, a view that results from the
insight that a neuron fires most efficiently if it receives synchronous
spike input, i.e., in the range of a few ms \autocite{Abeles82}.

To test whether temporal coordination of spiking activity is indeed
relevant for neuronal information processing, advanced data analysis
methods are required that perform correlation analysis between
simultaneously recorded single unit spike trains. The Unitary Events
(UE) analysis method \autocites{GruenPhD}{Gruen99}{Gruen02a}{Gruen02b}
is able to extract significant spike synchrony between neuronal
activities that is beyond what is expected by chance (given the rates)
and to follow the dynamics of this coordination. The tricky issue with
such analyses is that one has to take into account that experimental
spike data are typically non-stationary in the sense that the firing
rates are not stationary in time and not homogeneous across trials, and
that the statistics of the individual spike trains deviate from those of
a Poisson process \autocites{Gruen09}{GruenRotter10_Chap10}. If such
features are ignored there is a considerable danger of occurrence of
false positives and thus wrong interpretation of the data. In
particular, changes in the firing rates are the most prominent
generators of false positives if ignored. The original UE method
considers this aspect by performing the analysis in a sliding time
window fashion. In later versions of the analysis method also other
features were corrected for by considering the respective features in
the null-hypothesis of the tests
\autocites{Gruen03b}{Maldonado08}{Louis10}{Pipa2013}, either by using
extended analytical descriptions of the null-hypotheses or by use of
surrogate methods \autocites{Gruen09}{GruenRotter10_Chap10}{Louis10}.

The UE method had been applied to experimental parallel spike data from,
e.g., the motor cortex of awake behaving non-human primates
\autocites{Riehle97}{Gruen99}{Grammont99}{Riehle2000}{Gruen03b}{Kilavik09}{Denker10}{Denker11}
and to data from visual cortices \autocites{Maldonado08}{Ito11}.
Generally it was found that UEs, i.e., synchronous spike events across
neurons that are in excess of the expectation, occur in relation to
behavior, e.g., when the animal expects a signal to occur, however the
signal does not occur \autocite{Riehle97}. This finding, in particular,
demonstrates the core feature of the method: by performing a time
resolved analysis, the method accounts for changes in the firing rates
and captures modulations of significant spike synchrony in time. It was
also shown that the time of occurrence of UEs may change to a new
requested timing in the behavior during a learning process
\autocite{Kilavik09}.

Several of these studies based on the UE method, in particular
\autocite{Riehle97}, have been widely cited and the method has been
recognized as one of the standard tools to analyze temporal coordination
of neuronal spiking activities \autocites{Brown04_456}{Nakahara02}. The
analysis method has been and is taught in international data analysis
courses. However, a publicly available and open source implementation of
the UE method had not been available. Only very recently, one of the
authors of this study (VR) reimplemented the UE analysis as part of the
Electrophysiology Analysis Toolbox\footnote{\url{http://neuralensemble.org/elephant/}}
(Elephant), a Python library that provides implementations for the
analysis of electrophysiological data. A reimplementation of the method,
rather than releasing the existing code, was justified by two factors.

First, the custom data object model used to represent the primary data
and metadata in the original analysis code was not documented.
Therefore, any data represented in a specific file format had to be
converted by implementing a custom data loading routine for this data
model. Our new implementation of the UE method is part of the Elephant
and Neo\footnote{\url{http://neuralensemble.org/neo/}} libraries and is
based on the internal data object model provided by the Neo library
\autocite{Garcia14}, a package for representing electrophysiology data
in Python. Neo provides support for reading a wide range of
neurophysiology proprietary file formats, and supports writing to a
subset of these formats, including non-proprietary formats such as HDF5.

Second, the implementation of the UE method in the original publication
has experienced several updates after its publication in order to
include improvements and extensions of the method to accommodate more
features of experimental data. Since no version control was employed in
this development process, the original code used in Riehle et al.
\autocite{Riehle97} was lost at some point. Considering that no
systematic testing was performed each time the code base of the Unitary
Event analysis was updated after the original publication to check
whether the code gives the same results as before, it was not clear
whether the latest version could exactly reproduce the results of
\autocite{Riehle97}.

In this paper we illustrate the successful reproduction of the results
shown in \autocite{Riehle97} using our new Python implementation of the
UE method. In particular, we reproduce Figure 2 and Figure 4A of the
original paper, which represent the central results of the original
study. The remainder of the original paper consisted of more example
analyses of individual data sets and a meta statistics across many data
sets, all of which were based on the same analysis method and thus do
not provide additional insight when reproduced. In the original
publication the authors used two different UE implementations: one
implemented by one of the coauthors of the present study (SG) in
IDL\footnote{\url{http://www.harrisgeospatial.com/ProductsandSolutions/GeospatialProducts/IDL.aspx}},
and the other in Matlab (Mathworks, Nattick, MA) implemented later by
Markus Diesmann (MD) and SG. The IDL implementation is not available
anymore. At a later point we were provided with a Matlab implementation
of the extended UE method, however it does not preserve the original
implementation used in \autocite{Riehle97} and was not considered in
this study.

For the reproduction of the original results we contacted and
communicated with Alexa Riehle (AR), CNRS-AMU, Marseille, and MD,
Research Centre Jülich. AR is the first author of the original
publication and performed the experiments and data analysis. MD is the
third author of the original publication and contributed with the Matlab
implementation and quality checks of the software implementations. The
reproduction of the results of \autocite{Riehle97} would not have been
possible without contacting these authors, since the information in the
original publication is not sufficient for reproducing the results. AR
provided us with the original data and information, including an old
written report by MD, which was crucial to reproduce Figure 4A.

\section{Methods}\label{methods}

For reproducing the results of \autocite{Riehle97} we use our
reimplementation of the UE analysis method in Python which is made
available in the \texttt{unitary\_event\_analysis} module of the
Elephant library (accepted pull-request: \autocite{Pullrequest_UE}). The
method is accompanied by unit tests for individual functions (test
coverage: \(88.54\%\)) and documented as part of the library
documentation. The structure and algorithm of our UE implementation is
explained in the pseudo-code shown below.

In the following, we will explain the algorithm in detail. The primary
data entering the UE method is a set of parallel, i.e., simultaneously
recorded, spike trains, recorded in one or multiple trials. In Elephant,
an individual spike train of a particular neuron in a particular trial
data is represented as a \texttt{Spiketrain} object in the data object
model provided by the Neo library, which stores the time points of spike
occurrences along with additional information describing the spike
train, such as the start and end times. The UE method consumes a nested
list of \texttt{Spiketrain} objects relating to the spike trains of
individual neurons in the individual trials. In order to perform the
necessary calculations, the spike data must be converted to an
alternative time-binned representation (line 1 in the pseudo code) where
the parallel spike trains are stored as binary sequences of ones
(marking time bins containing at least 1 spike) and zeros (bins with no
spike). The bin size is a parameter provided as input (parameter
\emph{\texttt{bin\_size}} in the pseudo code) to the analysis, and
defines the temporal precision in detecting spike synchrony. A spike
pattern, which is a representation of spike synchrony in the analysis,
is defined as a specific vector of zeros and ones in one time bin of a
given trial across all neurons. Thus, the number of possible spike
patterns is given by \(2^{N}-N-1\), where \(N\) is the number of
neurons. In order to limit the analysis to a set of patterns of
interest, the input parameter \emph{\texttt{pattern\_hash\_values}}
specifies the patterns to consider in the analysis in the form of a hash
value which uniquely represents each spike pattern (line 2 in the pseudo
code). The hash value is obtained by interpreting the binary spike
pattern as a binary number, where the \(n\)-th neuron is represented by
bit \(n-1\).

The UE analysis is performed in a sliding time window fashion, i.e.~the
data in each window are analyzed separately. This approach is chosen to
account for potential changes of the firing rates in time and to follow
the dynamics of the correlation. Here, a sliding time window is defined
based on trial time, meaning that a certain window position includes the
activity of all neurons in all trials in a certain time interval of the
trial. In the algorithm, at each position of the window, the data
contained in the window are extracted in order to compute the
significance of the specified patterns (line 3 in the pseudo code). For
doing that the UE analysis requires the empirical number of occurrences
of each pattern as well as its expected number given the firing rates.
While the empirical number is directly extracted from the data (line 5
in the pseudo code), the method to compute the expected number can be
chosen using the input parameter \emph{\texttt{method}} depending on the
assumptions regarding the data (line 6-19 in the pseudo code). For
cross-trial homogeneous data, selecting
\emph{``analytic\_TrialAverage''} as \emph{\texttt{method}} (line 7 in
the pseudo code) computes the expected number by estimating the rate of
each neuron from the average spike count across trials
\autocites{Gruen02a}{Gruen02b}. For data with cross-trial variability,
the \emph{``analytic\_TrialByTrial''} method (line 10 in the pseudo
code) should be used, which accounts for cross-trial changes in the
firing rates by computing the expected number of occurrences of the
spike pattern based on the product of single-trial estimates of firing
rates \autocite{Gruen03b}. Both methods perform a parametric test, where
the\\
\includegraphics{UE_algorithm.eps}\\
number of occurrences of the spike pattern is assumed to be a
Poisson-distributed. As an alternative non-parametric option, selecting
\emph{``surrogate\_TrialByTrial''} as \emph{\texttt{method}} (line 15 in
the pseudo code) will numerically compute the distribution of the
expected number by implementing the null-hypothesis based on surrogate
spike trains \autocite{Gruen09}. In the following, we will explain these
methods in greater detail.

In case of selecting \emph{``analytic\_TrialAverage''} - as used in the
remainder of this study for the reproduction of \autocite{Riehle97} -
the number of spikes per neuron within the sliding window is summed
across trials and divided by the number of trials and bins contained,
thus yielding the average probability \(p_{i}\) to have a spike of
neuron \(i\) in a bin of the time window. The probability to find a
particular pattern by chance in a bin is then computed by multiplication
of the relevant probabilities, e.g.~for a pattern {[}1,0,1{]} the
probability \(p\) of occurrence is given by
\(p_{[1,0,1]}=p_{1}*(1-p_{2})*p_{3}\). Note that (\(1-p_{2}\)) is the
probability for neuron 2 to contribute no spike to the pattern. The
expected number of pattern occurrences, computed as the product of the
occurrence probability \(p\) of the pattern, e.g. \(p_{[1,0,1]}\), and
the number of bins (across all trials) covered by the sliding window.
The distribution of pattern occurence numbers is given by a Poisson
distribution with the mean equal to the expected number.

Alternatively, for the \emph{``analytic\_TrialByTrial''} method, the
firing probability of each neuron is calculated in a trial-by-trial
manner based on the spike counts per trial. The probability \(p\) of
finding the pattern by chance is calculated by summing the products of
the firing probabilities obtained individually from each trial. As for
the trial-averaging method, the expected number of pattern occurrences
is given by multiplication with the number of bins of the trial in the
sliding window, used as the mean of a Poisson distribution to obtain the
distribution of expected pattern occurence numbers.

As a third alternative, a surrogate method for estimating the expected
number can be selected using \emph{``surrogate\_TrialByTrial''} for the
parameter \emph{\texttt{method}}. In this Monte-Carlo approach, a
surrogate version of the spike trains is generated repeatedly, and from
each surrogate the number of occurrences of the pattern of interest is
counted. The method by which surrogates are generated from the input
spike trains is spike time randomization of the spikes per trial and per
neuron within the sliding window. The pattern counts obtained from this
procedure form a distribution of the expected number of occurrences of
the pattern, thus implementing the null-hypothesis under the constraints
implied by the surrogate method.

The distribution obtained by either of the three methods above is then
used for the significance test of the pattern on the basis of the
empirical occurrence count. The p-value resulting from the test is then
transformed by a logarithmic transformation to the surprise value (line
20 in the pseudo code), which indicates by positive or negative values
more or less occurrences of the pattern than expected by chance,
respectively. If the p-value is below a fixed prescribed level
(e.g.~below \(5\%\), which corresponds to a surprise value exceeding
\(1.27\)), the occurrences of the spike pattern under investigation in
the sliding window are marked as UEs for that pattern. This procedure is
performed for each pattern of interest, and in each sliding window.

In the present study, in order to reproduce the original results we used
the \emph{``analytic\_TrialAverage''} method, which reflects the
analysis performed in the original publication \autocite{Riehle97}. The
\emph{``analytic\_TrialByTrial''} and \emph{``surrogate\_TrialByTrial''}
methods are extensions of the original UE method, which were developed
after the original publication and introduced in subsequent works
\autocites{Gruen03b}{Gruen09}.

Our reimplementation of the UE method is based on the data object model
provided by the Neo library, upon which the Elephant library is based.
The Neo library provides loading routines for a variety of data formats,
including proprietary and generic data formats. The data sets available
for reproducing Figures 2 and 4A of \autocite{Riehle97} were
tab-separated ASCII text files containing two columns of integers
(informally often referred to as ``GDF-format''): the first column
provides event codes (behavioral events or neuron IDs), and the second
column contains the time of the occurrence of these events (time
stamps). The units of the time stamps are not contained in the data
file. We partly extracted metadata information, in particular the time
units and the meaning of the event codes, from a Matlab routine
(provided by AR) operating on the GDF data file. However, only after
further communication with AR we were able to identify the exact meaning
of the content of the data files. Using this information, we wrote a new
loading routine that loads the GDF data as Neo data objects.

Our reimplementation uses the \texttt{conversion} module of Elephant for
converting the spike data (represented as a series of time stamps) into
the binary sequence to guarantee a unique, global binning mechanism for
all analysis methods provided in Elephant. The bin size to be set for
the analysis was extracted from the original publication. However,
defining the time point to start the binning of each single trial data
required to know the alignment event in each trial and how much time
before this event (pre-time) is considered. Since this information was
not documented in the original paper, we tried several possibilities
until we got an agreement with the original figures as will be shown in
the Results.

To check if our Python implementation produces the same results as the
implementation(s) used in the original publication, we compare each of
our figures visually in detail with the original figures. This is also
the test if we used the correct data files since the names of the data
sets are not mentioned in the original publication. A direct numerical
comparison to the original results plotted in these figures was not
possible as this data is not available.

\section{Results}\label{results}

For the reproduction of the original results in \autocite{Riehle97}, we
focus on reproducing Figure 2A-F and Figure 4A. Figure 2 is chosen
because it represents the main result of the study and includes the UE
analysis that underlies all subsequent analyses. Figure 4A is chosen
because this is an example of the application of the UE method to data
with more than 2 neurons. In terms of complexity of the code the
implementation of the UE analysis for three or more neurons is
considerably more demanding than for only two neurons. With this example
we show that our implementation is capable of performing the UE analysis
for the generic case of arbitrary numbers of neurons.

We apply our reimplementation of the UE method to preprocessed versions
of the spike train data available to us after communication with AR,
which in part were identical to those used in the original analysis.
Also, we learned from AR that Figure 2 was generated by the Matlab
implementation of the UE method while all remaining figures of the
original publication, including Figure 4A, were generated by the older
implementation in IDL (see Methods regarding versions of the original
code).

\begin{figure}[htbp]
\centering
\includegraphics{figure1.eps}
\caption{\label{fig:figure2alignedPS}Initial attempt to reproduce Figure
2 of the original publication with trial alignment to PS. \textbf{A)}
Raster plot of two neurons (neuron 2: top of panel; neuron 3: bottom of
panel) in 32 trials (sorted identically for both neurons). \textbf{B)}
Average firing rate of each neuron calculated across trials in a sliding
window of length 100~ms in steps of 5~ms. \textbf{C)} Same raster plot
as in panel A with spike coincidences (i.e., pattern {[}1,1{]}) between
the two neurons marked by cyan squares. \textbf{D)} Empirical (cyan) and
expected (magenta) number of coincidences calculated in a time-resolved
manner (parameters of sliding window identical to panel B). \textbf{E)}
Time course of the surprise measure, calculated in same sliding windows
as in panel B. Surprise values that correspond to positive and negative
significance levels \(\alpha+=0.05\) and \(\alpha-=0.95\) are shown with
by horizontal red and green lines, respectively. \textbf{F)} Same raster
plot as in panel A with significant coincidences, i.e.~UEs, marked by
red squares.}\label{fig:figure2alignedPS}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure2.eps}
\caption{\label{fig:figure2alignedRS}Reproduction of Figure 2 of the
original publication with trial alignment to RS. The same conventions as
in Figure ~\ref{fig:figure2alignedPS} apply to the respective
panels.}\label{fig:figure2alignedRS}
\end{figure}

Let us start with the reproduction of Figure 2 of the original
publication. We first give a brief description of the experiment (see
the original publication for details). After the monkey was presented
with the preparatory signal (PS) he had to sit still and wait for a
response signal (RS) to start his arm movement (i.e.~equivalent to a GO
signal). The duration of the waiting period was randomly selected on a
trial-by-trial basis to be either 600, 900, 1200 or 1500 ms. In Figure 2
of \autocite{Riehle97} only trials of the longest waiting period (1500
ms) were used for the analysis. In these trials, times marked as
expected signals ES1, ES2, and ES3 corresponded to the ends of the three
shorter waiting periods, at which the monkey could have gotten the RS
signal but did not. As the monkey was trained to recognize and
distinguish the four waiting periods, but was not informed of the
randomly selected period for a given trial, ES1-ES3 were time points at
which the monkey expected that a signal could occur.

Since the data file for Figure 2 contained the data as a continuous
recording of one recording session (``winny131.gdf''; 2 neurons, and
behavioral events), we extract the trials by cutting the data in a time
window around specific trigger events that belong to trials of the
longest waiting period, such that the complete trial is contained in the
cut-out. In a subsequent step, the spike times in the individual trials
are aligned to the trigger event, such that spike times in each trial
are given as relative to the trigger.

The original publication does not provide information which event was
used as the trigger. In this experiment, 2 events that occur in every
trial could serve as trigger events, the preparatory signal PS (event
code 114) and the response signal RS (event code 124). We noticed that
the time interval between PS and RS for the longest trials was not
identical across the respective trials and varied by \(\pm\) 1~ms. Given
the UE method is applied on a time scale of 5~ms, the analysis results
therefore are expected to depend on whether trials are aligned to PS or
RS. Thus, we decide to generate the results for both alignments.

Figures~\ref{fig:figure2alignedPS} and ~\ref{fig:figure2alignedRS} show
the results of performing the UE analysis for PS- and RS-aligned data,
respectively. Here, the analysis parameters are set to the identical
values as reported in the original publication (bin size: 5~ms, analysis
time window size: 100~ms, time step of the sliding window: 5~ms,
significance level \(\alpha= 0.05\)). The comparison of the two figures
to the original figure shows agreement in the raster displays (panel A)
and the time-resolved, trial-averaged firing rate estimates (B).
However, although the graphs of the number of coincidences per sliding
window (panel D) and the surprise measure (panel E) are similar in their
overall general behavior, they differ in the details. Thus, indeed the
choice of the alignment influences the analysis result. In order to test
if one of the two alignments is in agreement with the original
publication, we perform a detailed visual comparison of our two figures
and the original one on the basis of the spikes marked as coincident
(panels C) and as part of a UE (panel D). We notice that when aligning
to PS (Figure~\ref{fig:figure2alignedPS}), the marked spikes do not
agree in all details with the original figure. However,
Figure~\ref{fig:figure2alignedRS}, with trials aligned to RS, agrees
completely with the original figure. Although we cannot check the
identity to a further extent because the exact numerical values of the
original results are not available anymore in electronic form, we
believe that this visual inspection provides enough support for a
reproduction of the original result by the reimplementation of the UE
method.

As a next step we aim at reproducing Figure 4A of \autocite{Riehle97}.
This figure contains the result of the analysis of three neurons
recorded simultaneously, in contrast to Figure 2 where only two neurons
are considered. We analyze the original data for this figure provided by
AR with the parameter values given in \autocite{Riehle97} and compare
our result to the original figure. Figure 4A of the original publication
contains the raster displays of the data in the top panel, the raster
displays with the marked coincident spikes (blue marks) in the middle
panel, and the raster displays with the marked spikes that are part of a
UE (red marks) in the bottom panel. We find that the UE result is
different, as the UEs occur at different times and between different
neurons compared to the original publication. Thus we check whether the
spike times of the individual spikes are identical between the original
and our results. Figure~\ref{fig:comparison_raster} shows a segment of
the raster plot of the original figure and the corresponding segment of
our reproduced figure. We compare the positions of the single spikes and
find that there are small discrepancies between the two raster plots in
some of the spike times. Figure~\ref{fig:comparison_raster} shows
examples of clusters of spikes marked in red that should be identical in
both raster plots but contain a few individual spikes that are slightly
shifted in our figure compared to the original figure by a very small
amount.

\begin{figure}[htbp]
\centering
\includegraphics{PS_aligned_matlabdata_marked.eps}
\caption{\label{fig:comparison_raster}Close-ups of the original raster
display in Figure 4A of the original publication (left) and the first
data file available at hand for reproduction (right) reveal slight
differences in the positions of some spikes. Data on the right are
aligned (similar to the data on the left) to ES1 (event code 15 in the
GDF data file). The time before the alignment event is chosen as 700~ms,
and a bin size of 5~ms is used. The red marks indicate spike clusters
with identified differences between the left and the right panel, where
at least one spike is shifted in the right panel compared to the left
one.}\label{fig:comparisonux5fraster}
\end{figure}

This leads us to the suspicion that the data are binned in a fashion
that is not consistent with the data shown in the original publication.
Personal communication with AR revealed that while Figure 2 had been
generated by the Matlab implementation of the UE analysis, Figure 4 had
been generated by the IDL implementation (see Introduction). A report by
MD written before the time of the original publication summarized a
comparison of the IDL and the Matlab implementations, and concluded that
both were correct implementations of the method, but differed in their
results due to a slightly different implementation of the down-sampling
and binning of the raw data (recorded at 10 kHz). In the workflow for
the IDL implementation, as illustrated in
Figure~\ref{fig:workflow_report} (the leftmost branch of the diagram),
the raw data were first down-sampled to a temporal resolution of 0.5 ms
(by a program \texttt{2gdf}) and then further rounded to 1 ms resolution
integer values inside the IDL implementation. The data available to us
had a resolution of 1 ms, which must have been a result of another
down-sampling procedure than the one for the IDL implementation. This
explains the difference in the raster displays, and this difference is
likely also the cause that we were initially not able to reproduce the
original UE result.

\begin{figure}[htbp]
\centering
\includegraphics{CmpIDL_Matlab_3.eps}
\caption{\label{fig:workflow_report}Illustration of the data
preprocessing workflow, taken from the 1997 report of MD (in German) on
the comparison of the first UE implementation in IDL (left branch) and
the second implementation in Matlab \autocite{Diesmann16_personalcomm}
(right branch). Data entering both analysis branches have a resolution
of 0.1 ms (top box). In the IDL branch spike times \(t\) in the original
data are first transformed to a resolution of \(h=0.5\) ms (by the
\texttt{2gdf} program, middle left) using the method of binning
\(\lfloor t/h \rfloor\). Then the data are read into the ``IDL UE
Software'' (lower left box) and therein converted to \(h=1\) ms
resolution by the method of ``round half up''
\(\lfloor t/h+1./2 \rfloor\) prior to analysis. Alternatively, one can
load the 0.5 ms resolution data into the ``MATLAB UE Software'' (lower
right box). Here the bin width is a parameter of the analysis and thus
data can be converted to a 1 ms resolution but results are different
from the IDL branch. Results are only identical if a prior
transformation ``T'' (diagonal in center) performs a round half up and
no further binning is done in the MATLAB program. At a later point in
time the \texttt{alexa2gdf} converter function (written in MATLAB)
became available such that data in the original 0.1 ms resolution could
directly be converted to the 1 ms resolution by binning. The full report
(``Report\_by\_MarkusDiesmann.txt'') is included in the data folder of
the repository for this paper.}\label{fig:workflowux5freport}
\end{figure}

In our reproduction of Figure 2 of the original paper we use
preprocessed data available in 1 ms resolution, that likely experienced
the \texttt{alexa2gdf} program for conversion as shown in
Figure~\ref{fig:workflow_report} (the rightmost branch), before data are
loaded into our reimplementation of the UE analysis. However, according
to the aforementioned report by MD, we only have a chance to reproduce
Figure 4A of \autocite{Riehle97} if we have the original data or a
version of them with a time resolution lower than 1 ms available. The
original raw data with 0.1~ms resolution are presumably only available
on a storage medium and format that at present we are not able to read
and interpret. However, after we contacted AR she found the data
(``jenny201\_345\_preprocessed.gdf'') of Figure 4A of
\autocite{Riehle97} with a time resolution of 0.5~ms, which likely
experienced the \texttt{2gdf} program for conversion (middle left box in
Figure~\ref{fig:workflow_report}).

We loaded this data at 0.5~ms resolution into Python and converted the
data from the 0.5~ms to the required 1~ms resolution by the mathematical
operation \(\left\lfloor x+\frac{1}{2}\right\rfloor\) , called
``rounding half up''. In numerical software packages, including IDL,
this operation is typically implemented by a function named round().
However, the round() implementation of NumPy (version 1.11.0) performs
an even rounding, i.e., values exactly halfway between two integers are
rounded to the nearest even integer. Indeed, the latter implementation
of rounding did not reproduce the result of the original publication.
Thus, we used the expression \texttt{floor(x+0.5)} to perform rounding
as it is implemented by IDL. The procedure completely reproduces panel A
of Figure 4 in the original publication (see
Figure~\ref{fig:reproducedFig4A}).

\begin{figure}[htbp]
\centering
\includegraphics{figure3.eps}
\caption{\label{fig:reproducedFig4A}Reproduction of Figure 4A of the
original publication. The left part of the figure shows the UE analysis
result for the data aligned to ES1 (event code 15 in the GDF data file)
with a time before the event (pre-time) set to 699 ms. The right part of
the figure shows the analysis of the same data aligned to RS and with a
pre-time of 99 ms. The left and right parts of the figure show 96 and
128 trials, respectively.}\label{fig:reproducedFig4A}
\end{figure}

\section{Conclusion}\label{conclusion}

We are able to reproduce the original results of \autocite{Riehle97} by
applying a new reimplementation of the Unitary Events analysis method in
Python to the original data. The method involves a number of numerical
computations and is very sensitive, as we show here by the comparison of
Figures ~\ref{fig:figure2alignedPS} and ~\ref{fig:figure2alignedRS},
which differed in the events that the trials were aligned to. This
difference in the alignment would not make any difference in the results
if the time difference between the two events (PS and RS) were identical
across the trials. But since the latter was not the case due to hardware
features of the recording setup (as we learned from the first author of
the original publication), the binning of the data started at a slightly
different time points in different trials. This likely led to a loss or
an addition of a spike in a bin and thus to a small difference of the
number of spike synchrony events (see also the discussion on the issues
of exclusive binning in \autocite{Gruen99}). Nevertheless, this
sensitivity of the method is a strong indication that our reproductions
of Figure 2 and Figure 4A of the original publication are correct and
that our new implementation of the analysis faithfully implemented the
UE method.

The event to which the data were aligned and the cut time which then
defined the start of the (exclusive) binning was not documented in the
original publication. Also the original scripts for the analysis are not
available anymore, which could have revealed this information, even
without having the original UE software code at hand. Thus, due to the
lack of documentation we are only able to reproduce the results of
\autocite{Riehle97} by communicating with some of the authors of the
original publication.

The reproduction of Figure 4A of the original publication is a further
and important test whether our reimplementation is also correct for
N\textgreater{}2 neurons. This is relevant since the implementation
requires a more generic, complex algorithm for the analysis than the one
that can be used for only N=2. In the case of two neurons, there is only
one pattern type which has to be analyzed (i.e. {[}1,1{]}). However, in
the case of e.g.~3 neurons there are already 4 different spike patterns
to analyze ({[}1,1,0{]}, {[}0,1,1{]}, {[}1,0,1{]}, and {[}1,1,1{]}), and
even much more for more neurons (\(2^{N}-N-1\)). The statistics of each
of the patterns is performed separately and, therefore, the bookkeeping
needs to be carefully done.

The reproduction of Figure 4A is more complicated than reproducing
Figure 2 due to additional reasons. First of all, the data were not
available to us. After requesting them from the first author of the
original publication we received data and were not able to reproduce the
result - in terms of the UE results, but also the data seemed slightly
different. After further consultation with the original authors we
learned that the original Figure 4A was not generated by the Matlab
implementation used for Figure 2 but by another implementation in IDL.
Both are not available to us. However, we were told that the third
author of the original publication performed a thorough comparison of
the two implementations at the time and the final report on that
investigation was made available to us. This enables us to define the
correct workflow that reproduces the original result, given we have the
data in the correct resolution at hand.

The reimplemented UE analysis software contains extensions for improving
the statistics that were developed after the original publication. On
the one hand, it contains the option to adjust the statistics to take
into account cross-trial inhomogeneity by calculating the number of
expected spike synchrony events based on the firing rates in a
trial-by-trial fashion (option: \emph{``analytic\_TrialByTrial''}) as
suggested in \autocite{Gruen03b}, in contrast to using trial averages of
the firing rates. On the other hand, our reimplementation offers the
possibility to calculate the significance of the empirical number of
spike synchrony events based on a Monte Carlo approach (option:
\emph{``surrogate\_TrialByTrial''}). Instead of computing the
significance using a parametric distribution based on the estimate of
the firing rates, the null-hypothesis of independence is implemented by
surrogate data \autocites{Gruen09}{GruenRotter10_Chap10}{Louis10}. By
repeated intentional manipulation of the original data, potential spike
synchrony is deleted. Each of these surrogate data created by this
procedure are then searched - as the original data - for spike
synchrony, and these numbers create the distribution underlying the
significance test of the method. Obviously this version is considerably
more computationally expensive than the parametric approach used here
for the reproduction of \autocite{Riehle97} and we are currently working
on an HPC implementation to make use of parallelization. The Python
implementation of the UE method is publicly available in the open source
software package Elephant at http://neuralensemble.org/elephant/.

If the authors of the original paper would not have been accessible, we
would not have been able to reproduce the results. Nevertheless, here
our final validation of the reproduction is based only on a detailed
visual comparison of the original and reproduced figures. In an optimal
scenario, we would be able to exactly validate the results based on a
numerical comparison. To do so, all of the following pieces of
information would have had to be available at hand:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  the original primary data
\item
  metadata describing the primary data in detail
\item
  the original statistics software package (e.g.~Unitary Events)
\item
  the loading routine for the data
\item
  all specific code required to produce each figure of the original
  publication
\item
  detailed documentation of all code
\item
  the original software environment (with programs available in the
  original versions used), including, e.g.~the interpreter/compiler
  (here: Matlab) and operating system
\item
  unique identifiers of the data records that unambiguously identify
  data from within the analysis code
\end{enumerate}

In the analysis presented in this work, not even the original primary
data (1), recorded more than 20 years ago, but only a slightly
preprocessed version is available. However, even today many of the
pieces of information listed above are often not made available by
scientists. In part, this is due to the enormous complexity of the task
to record all information in fine detail leading from the experiment to
an analysis result. Moreover, there is still a lack of software tools to
support researchers in the process of acquiring, storing, and organizing
this information. Currently, there are emerging approaches suggested for
metadata annotation (2) of electrophysiological data, such as the
odML\footnote{\url{https://github.com/G-Node/python-odml}} framework
(see, e.g. \autocites{Grewe2011}{Zehl2016}) for storing hierarchical
collections of metadata or the NIX\footnote{\url{https://github.com/G-Node/nix/wiki}}
data format \autocite{Adrian2014} for linking data and metadata. In our
concrete example, the information about the hardware limitations in
storing the event times, would have been essential information contained
in the metadata. Using modern tools for version control, points (3)-(6)
can be easily addressed. There are emerging approaches to keep software
environment, i.e., the original Matlab version and the operating system
(7), e.g., by freezing the environment in a virtual machine. Point (8)
is still challenging, because it requires the data to be addressed in an
unambiguous manner from within the analysis scripts. Including data
within the code repositories is typically prohibitive due to the size of
the data. A solution would be to deposit data in public or private
databases that allow data to be identified using a unique identifier in
combination with a tool to generate a detailed provenance track of the
analysis process, but the implementation of tools and services for the
workflows used in data analysis of electrophysiological data is still an
ongoing endeavor \autocites{badia_incf_2015}{Denker2015_000}. In
summary, there are still components missing such that researchers are
put into a position to build complex data acquisition and analysis
workflows that enable optimal reproducibility in neuroscience.

\section{Acknowldegements}\label{acknowldegements}

This project received funding from the European Union's Horizon 2020
research and innovation programme under grant agreement No. 720270,
Deutsche Forschungsgemeinschaft Grant DE 2175/2-1 and GR 1753/4-2 of the
Priority Program (SPP 1665), the German-Japanese Computational
Neuroscience Project (German Federal Ministry for Education and
Research, BMBF Grant 01GQ1114), from the Helmholtz Portfolio Theme
``Supercomputing and Modeling for the Human Brain'', and from the Osaka
Univ for the project `Neural mechanism of active vision studied by
combining large-scale sampling of neural activity and advanced
computational analysis'.

We thank Alexa Riehle and Markus Diesmann for fruitful discussions.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2016-12-04 article/shifman-2016.tex
1st round revisions

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Aaron R. Shifman,
  },
            pdftitle=
{
[Re] Ionic Current Model of a Hypoglossal Motoneuron
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{shifman-2016.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2016 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Ionic Current Model of a Hypoglossal Motoneuron}

    \usepackage{authblk}
                        \author[1]{Aaron R. Shifman}
                            \affil[1]{Department of Biology, University of Ottawa, Ottawa, Ontario, Canada}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:ashifman@uottawa.ca}{ashifman@uottawa.ca}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{Ionic Current Model of a Hypoglossal Motoneuron}, Purvis LK.,
Butera RJ., (2005) Journal of Neurophysiology, 93, 723--733
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

This work serves as a reference implementation of the model proposed in
\autocite{purvis}, which is a biophysically detailed model of the ionic
currents involved in the rat hypoglossal motoneuron (HM). HMs innervate
the tongue and are involved in breathing, and issues with HMs have been
implicated in sleep apnea\autocite{purvis}. A biophysical model is
derived from known properties, and allows researchers to describe in
detail the role and interactions of all components of the phenomenon.
The model was originally solved (integrated) in XPP and analysed in
MATLAB. This version has been developed in Python both for the fact that
Python is relatively ubiquitous, and the availability of scientific
packages (NumPy, SciPy, \dots). A version of this model exists in cellML
format\autocite{cellml}, however the cellML model is not able to
replicate the original results (action potentials do not occur at 1nA
bias current), nor could it create action potentials with similar
appearance. Due to the nature of the model, debugging is nearly
impossible (multiple variables with the same name i.e. ``theta\_1'').

\section{Methods}\label{methods}

This work was prepared using the authors' description of the model
\autocite{purvis}, and personal communication with the authors. Some
slight inconsistencies were identified in the original paper which had a
negligible effect on the results of the paper. These have been addressed
in this implementation.

The mathematical description of the model was taken directly from page 2
of the original manuscript, and current dynamics were taken from the
appendix. The constants (parameters) were taken from table 2 and the
initial conditions were taken from table 1. It should be noted the the
given initial conditions, while near the equilibrium point (rest), were
not at equilibrium. In order to address this, the model was allowed to
evolve (in the absence of bias current) until numerical equilibrium was
reached, and this was taken to be the initial condition.

While the initial implementation used a Dormad-Prince adaptive solver,
this implementation used the LSODA algorithm (provided in the
scipy.integrate package), with absolute and relative tolerances of
\(10^{-10}\), and a maximum time step of \(\Delta T=0.05\) ms
(convergence of all results was tested with \(\Delta T=0.01\) ms and no
qualitative differences were found). The maximum time step is necessary
in order to account for the relatively short step function current (not
mentioned in the original paper). This implementation used python 3.5.2
(64 bit) on Linux with packages: SciPy v0.17.1, NumPy v1.11.1,
matplotlib v1.5.1.

\section{Results}\label{results}

All results from the original paper have been implemented, however some
slight inconsistencies are noted and will be discussed later.

The first figure in the original paper is the voltage response to a 1
ms, 1 nA current pulse. This is reproduced in figure \ref{fig:fig1}. As
seen, the results are very similar to the original figure, however the
amplitude of the ADP seems a bit below that of the original figure. This
discrepancy will be addressed in the discussion.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_1.pdf}
\caption{Model response (action potential) to a 1 nA pulse active for 1
ms. Various properties are outlined in the response. \label{fig:fig1}}
\end{figure}

The subsequent analysis performed in the original text is an analysis of
the timescales and amplitudes of the individual ionic currents. This
implementation is recreated in figure \ref{fig:fig2}. This
implementation is very close to the original figure, however it is
important to note that the I-SK current in this implementation has a
peak of \textasciitilde{}0.8 nA, whereas the peak amplitude in the
original paper is \textasciitilde{}1 nA, there are also slight
differences in the calcium currents.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_2.pdf}
\caption{Ionic currents underlying an action potential created by a 1 nA
pulse active for 1 ms. \label{fig:fig2}}
\end{figure}

Figure 3 in the text is an analysis of the dependency of the ADP on
calcium and pre-pulse hyperpolarization. This has been replicated in
figure \ref{fig:fig3}. Again there are slight discrepancies which will
be addressed later.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_3.pdf}
\caption{ADP dependancy on calcuim and voltage. Maximal T-type channel
conductance is varied, and resting potential is clamped at a
hyperpolarizing potential. \label{fig:fig3}}
\end{figure}

Validating the model against known experiments, in figure 4 in the text
the authors simulate the application of apamin by eliminating the
SK-current. This result is very well reproduced in figure
\ref{fig:fig4}. The authors also demonstrate spike frequency adaptation
which is known to occur with attenuated SK-current.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_4.pdf}
\caption{Comparison of model to known experimental results. Removal of
SK-current eliminates the ADP and attenuated SK-currents cause spike
frequency adaptation. \label{fig:fig4}}
\end{figure}

Next, the model response to elongated current application is illustrated
in figure \ref{fig:fig5}, this quite accurately reflects the results
presented in the corresponding plot in the original paper (figure 5).
The action potential plateau is accurately recreated and the length of
the plateau is given by the length of the stimulus (neuron returns to
rest after stimulus is removed). Discrepancies in the length of the
plateau are purely due to discrepancies in the stimulus duration (the
duration used in figure \ref{fig:fig5} is a guess that appears close).

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_5.pdf}
\caption{Model response to sustained 0.22 nA pulse which lasts for 2.3
seconds. \label{fig:fig5}}
\end{figure}

The next two figures are an analysis of frequency adaptation of the
neurons under varying T-type calcium currents. In the text figure 7 is
described as having a T-type current 10x that of figure 6. While true,
it would be more accurate to say that figure 6 has a T-type calcium
current 0.1x that of figure 7, as 0.1\(\mu S\) is the ``standard''
conductivity and figure 6 has a conductivity of 0.01\(\mu S\). These
results are quite well replicated in figures \ref{fig:fig6} and
\ref{fig:fig7}. Of note the ``double peak'' as the model stabilizes onto
its sustained spiking behavior is accurately replicated.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_6.pdf}
\caption{Model response to sustained current step with maximal T-type
conductance of 0.01\(\mu S\) \label{fig:fig6}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_7.pdf}
\caption{Model response to sustained current step with maximal T-type
conductance of 0.1\(\mu S\) \label{fig:fig7}}
\end{figure}

Following this, spike adaptation is quantified as a function of both
T-type and N-type calcium currents and varying the definition of total
calcium current. These results are very well replicated in figure
\ref{fig:fig8}.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_8.pdf}
\caption{Spike frequency adaptation to various calcium currents
conductances. \label{fig:fig8}}
\end{figure}

The last figure in the text, figure 9, could not implemented solely
based on the description given in the text. In the text the mAHP length
is defined as being ``measured from the repolarizing phase of the action
potential to the point in the AHP that reaches the resting potential''.
Through clarification with the authors, mAHP length is defined as the
length of time between the negative and positive crossing of the resting
potential value.

This is implemented in figure \ref{fig:fig9}, this very closely reflects
the original figure (digitized values provided in the code), however
there are some values that deviate slightly, again this result will be
discussed later.

\begin{figure}[htbp]
\centering
\includegraphics{../code/figures/figure_9.pdf}
\caption{mAHP length as a funtion of H-current amplitude.
\label{fig:fig9}}
\end{figure}

\section{Discussion}\label{discussion}

By and far, this implementation replicates the original paper. Slight
discrepancies in the details however are apparent. For example in figure
\ref{fig:fig1} the ADP amplitude is a few mV different in this
implementation compared to the original. Also for example in figure
\ref{fig:fig2} I-SK is \textasciitilde{}0.8nA in comparison to
\textasciitilde{}1nA in the original. These discrepancies continue
through several other figures. In order to track down these
discrepancies the original authors provided the XPP files they used to
create the figures. The XPP simulation had the same equations and
parameters as this implementation and closely matches this
implementation (code for figure 2 provides the option to compare the
I-SK that XPP calculates with this implementation). Given the fact that
the paper had the same parameters as this implementation and the XPP
simulation matches this implementation, its likely that some simulations
had a slightly different parameter set.

It is important to note however that this in no way detracts from the
results of the original manuscript - the only difference is in subtle
features. The major results: replication of frequency adaptations,
parametric variation, etc\ldots{} are all replicated despite these
discrepancies.

\section{Conclusion}\label{conclusion}

With confidence, this implementation has indicated that the the original
results were both true, and correctly implemented. This implementation
did identify some subtle discrepancies in the original manuscript, which
were addressed in this implementation.

\section{Acknowledgement}\label{acknowledgement}

I would like to Dr.~Butera for providing the XPP files for the original
paper. Funding for this work was provided in the form of an OGS grant to
ARS from the government of Ontario.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-01-10 article/shifman-2016.tex
Revisions round 2

@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 simulation matches this implementation, its likely that some simulations
-had a slightly different parameter set.
+had a slightly different parameter set. For example figures
+\ref{fig:fig2} and \ref{fig:fig6} have different parameter sets, and
+there are discrepancies in figure \ref{fig:fig2} but not in figure
+\ref{fig:fig6}. This may imply that the discrepancies may be in the
+differing parameters.
 
-It is important to note however that this in no way detracts from the
-results of the original manuscript - the only difference is in subtle
-features. The major results: replication of frequency adaptations,
-parametric variation, etc\ldots{} are all replicated despite these
-discrepancies.
+This implementation has identified subtle discrepancies in the original
+paper. However this replication proves that all major results:
+replication of frequency adaptations, parametric variation, etc\ldots{}
+are valid despite these discrepancies.
 

### 2017-03-29 article/VRostami_JIto_MDenker_SGruen_2016.tex
changes after reviewrs comments

@@
 insight that a neuron fires most efficiently if it receives synchronous
-spike input, i.e., in the range of a few ms \autocite{Abeles82}.
+spike input, i.e., in the range of a few milliseconds
+\autocite{Abeles82}.
 
@@
 considers this aspect by performing the analysis in a sliding time
-window fashion. In later versions of the analysis method also other
-features were corrected for by considering the respective features in
-the null-hypothesis of the tests
+window fashion. In later versions of the analysis method other features
+were corrected for by considering the respective features in the
+null-hypothesis of the tests
 \autocites{Gruen03b}{Maldonado08}{Louis10}{Pipa2013}, either by using
@@
 (Elephant), a Python library that provides implementations for the
-analysis of electrophysiological data. A reimplementation of the method,
-rather than releasing the existing code, was justified by two factors.
+analysis of electrophysiological data. This reimplementation of the
+method was particularly required for, in addition to the general
+motivation for a reimplementation to confirm the results not depending
+on small implementation details or mistakes, the following two specific
+reasons.
 
@@
 is defined as a specific vector of zeros and ones in one time bin of a
-given trial across all neurons. Thus, the number of possible spike
-patterns is given by \(2^{N}-N-1\), where \(N\) is the number of
-neurons. In order to limit the analysis to a set of patterns of
-interest, the input parameter \emph{\texttt{pattern\_hash\_values}}
-specifies the patterns to consider in the analysis in the form of a hash
-value which uniquely represents each spike pattern (line 2 in the pseudo
-code). The hash value is obtained by interpreting the binary spike
-pattern as a binary number, where the \(n\)-th neuron is represented by
-bit \(n-1\).
+given trial across all neurons. Since the aim of the UE method is to
+detect significant spike synchrony, the total number of spike patterns
+of concern is given by \(2^{N}-N-1\), where \(N\) is the number of
+neurons, the first term is the number of all possible spike patterns,
+the second term is the number of spike patterns composed of only one
+spike, and the third term is the number of the pattern without a spike.
+In order to limit the analysis to a set of patterns of interest, the
+input parameter \emph{\texttt{pattern\_hash\_values}} specifies the
+patterns to consider in the analysis in the form of a hash value which
+uniquely represents each spike pattern (line 2 in the pseudo code). The
+hash value is obtained by interpreting the binary spike pattern as a
+binary number, where the \(n\)-th neuron is represented by bit \(n-1\).
 
@@
 the number of bins (across all trials) covered by the sliding window.
-The distribution of pattern occurence numbers is given by a Poisson
+The distribution of pattern occurrence numbers is given by a Poisson
 distribution with the mean equal to the expected number.
@@
 sliding window, used as the mean of a Poisson distribution to obtain the
-distribution of expected pattern occurence numbers.
+distribution of expected pattern occurrence numbers.
 
@@
 Our reimplementation uses the \texttt{conversion} module of Elephant for
-converting the spike data (represented as a series of time stamps) into
+converting the spike data (represented as a series of timestamps) into
 the binary sequence to guarantee a unique, global binning mechanism for
@@
 implementation(s) used in the original publication, we compare each of
-our figures visually in detail with the original figures. This is also
-the test if we used the correct data files since the names of the data
-sets are not mentioned in the original publication. A direct numerical
-comparison to the original results plotted in these figures was not
-possible as this data is not available.
+our figures in detail with the original figures. For this comparison, as
+the original results used to generate these figures are not available to
+us, we first extract the times of the spikes and unitary events from the
+vector graphic image (PDF) in the original paper. Then, we directly
+compare these times to the times of spikes and unitary events in our
+reproduced results by plotting the former against the latter, as well as
+examining the distribution of the differences between them.
 
@@
 For the reproduction of the original results in \autocite{Riehle97}, we
-focus on reproducing Figure 2A-F and Figure 4A. Figure 2 is chosen
-because it represents the main result of the study and includes the UE
-analysis that underlies all subsequent analyses. Figure 4A is chosen
-because this is an example of the application of the UE method to data
-with more than 2 neurons. In terms of complexity of the code the
-implementation of the UE analysis for three or more neurons is
-considerably more demanding than for only two neurons. With this example
-we show that our implementation is capable of performing the UE analysis
-for the generic case of arbitrary numbers of neurons.
+have to focus on reproducing Figure 2A-F and Figure 4A, since for the
+other figures the data still available were either incomplete in respect
+to metadata (original Figure 3) or all the required data set were not
+available anymore (original Figure 4B,C) . Figure 2 represents the main
+result of the study and includes the UE analysis that underlies the
+subsequent analyses. Figure 4A is an example of the application of the
+UE method to data with more than 2 neurons. In terms of complexity of
+the code the implementation of the UE analysis for three or more neurons
+is considerably more demanding than for only two neurons. With this
+example we show that our implementation is capable of performing the UE
+analysis for the generic case of arbitrary numbers of neurons.
 
@@
 Average firing rate of each neuron calculated across trials in a sliding
-window of length 100~ms in steps of 5~ms. \textbf{C)} Same raster plot
+window of length 100 ms in steps of 5 ms. \textbf{C)} Same raster plot
 as in panel A with spike coincidences (i.e., pattern {[}1,1{]}) between
@@
 the time interval between PS and RS for the longest trials was not
-identical across the respective trials and varied by \(\pm\) 1~ms. Given
-the UE method is applied on a time scale of 5~ms, the analysis results
+identical across the respective trials and varied by \(\pm\) 1 ms. Given
+the UE method is applied on a time scale of 5 ms, the analysis results
 therefore are expected to depend on whether trials are aligned to PS or
@@
 respectively. Here, the analysis parameters are set to the identical
-values as reported in the original publication (bin size: 5~ms, analysis
-time window size: 100~ms, time step of the sliding window: 5~ms,
+values as reported in the original publication (bin size: 5 ms, analysis
+time window size: 100 ms, time step of the sliding window: 5 ms,
 significance level \(\alpha= 0.05\)). The comparison of the two figures
@@
 to PS (Figure~\ref{fig:figure2alignedPS}), the marked spikes do not
-agree in all details with the original figure. However,
-Figure~\ref{fig:figure2alignedRS}, with trials aligned to RS, agrees
-completely with the original figure. Although we cannot check the
-identity to a further extent because the exact numerical values of the
-original results are not available anymore in electronic form, we
-believe that this visual inspection provides enough support for a
-reproduction of the original result by the reimplementation of the UE
-method.
+agree in all details with the original figure. However, in
+Figure~\ref{fig:figure2alignedRS}, with trials aligned to RS, we find no
+disagreements with the original figure. After this visual comparison we
+check if our results are exactly identical to the original ones.
+Therefore, we extract the positions of the data points representing
+spikes and UEs in the original figure by reading the plotting commands
+in the PDF file of the original paper, and compare them to our
+reproduced results. In Figure~\ref{fig:validation_fig2} A the reproduced
+spike times in Figure~\ref{fig:figure2alignedRS} A are plotted against
+the extracted original spike times. The plotted data points lie on the
+diagonal line, indicating that the reproduced spike times correspond to
+to the original ones. Figure~\ref{fig:validation_fig2} B shows the same
+plot for the UEs, indicating the identity of the original and reproduced
+UE timings as well. To further confirm the identity, we plot in
+Figure~\ref{fig:validation_fig2} C the distributions of the differences
+between the original and the reproduced timings for the spikes (black)
+and the UEs (red). The differences are at most +/- 0.3 ms, which are
+considerably narrower than the +/- 1 ms differences caused by the
+misaligned data shown in Figure~\ref{fig:figure2alignedPS} A (see for
+comparison the gray plot in Figure~\ref{fig:validation_fig2} C, which
+shows the differences between the original spike times and the
+misaligned spike times). The remaining minor differences of the spike
+and UE times of the correctly aligned data and the original data are
+only due to slight errors in the extraction of the spike times from the
+original figure, which is inevitable because of a limited precision of
+the plotting commands in the PDF file. Thus, we confirm that spike
+timings in the correctly aligned data are identical to those in the
+original data, and our implementation of the UE analysis applied to
+these data reproduces exactly the same results as shown in the original
+paper.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{validation_results_fig2.eps}
+\caption{\label{fig:validation_fig2}\textbf{A)} The scatter plot of the
+reproduced spike times in Figure~\ref{fig:figure2alignedRS} A plotted
+against the extracted original spike times. \textbf{B)} Same plot as in
+A but for the time of occurrences of UEs. \textbf{C)} The distributions
+of the differences between the original and the reproduced timings for
+spikes (black) and UEs (red) in Figure~\ref{fig:figure2alignedRS}. The
+distribution of the differences between the original and the reproduced
+timings for spikes in Figure~\ref{fig:figure2alignedPS} is shown in
+gray.}\label{fig:validationux5ffig2}
+\end{figure}
 
@@
 aligned (similar to the data on the left) to ES1 (event code 15 in the
-GDF data file). The time before the alignment event is chosen as 700~ms,
-and a bin size of 5~ms is used. The red marks indicate spike clusters
+GDF data file). The time before the alignment event is chosen as 700 ms,
+and a bin size of 5 ms is used. The red marks indicate spike clusters
 with identified differences between the left and the right panel, where
@@
 version of them with a time resolution lower than 1 ms available. The
-original raw data with 0.1~ms resolution are presumably only available
+original raw data with 0.1 ms resolution are presumably only available
 on a storage medium and format that at present we are not able to read
@@
 (``jenny201\_345\_preprocessed.gdf'') of Figure 4A of
-\autocite{Riehle97} with a time resolution of 0.5~ms, which likely
+\autocite{Riehle97} with a time resolution of 0.5 ms, which likely
 experienced the \texttt{2gdf} program for conversion (middle left box in
@@
 
-We loaded this data at 0.5~ms resolution into Python and converted the
-data from the 0.5~ms to the required 1~ms resolution by the mathematical
+We loaded this data at 0.5 ms resolution into Python and converted the
+data from the 0.5 ms to the required 1 ms resolution by the mathematical
 operation \(\left\lfloor x+\frac{1}{2}\right\rfloor\) , called
@@
 number of spike synchrony events (see also the discussion on the issues
-of exclusive binning in \autocite{Gruen99}). Nevertheless, this
-sensitivity of the method is a strong indication that our reproductions
-of Figure 2 and Figure 4A of the original publication are correct and
-that our new implementation of the analysis faithfully implemented the
-UE method.
+of exclusive binning in \autocite{Gruen99}). In spite of this
+sensitivity of the method, we succeeded in generating results that are
+exactly identical to the original ones, as confirmed by the direct
+comparison of the positions of spikes and UEs in the original and the
+reproduced figures. This is a strong indication that our new
+implementation of the analysis faithfully implemented the UE method.
 
@@
 Both are not available to us. However, we were told that the third
-author of the original publication performed a thorough comparison of
-the two implementations at the time and the final report on that
+author, MD, of the original publication performed a thorough comparison
+of the two implementations at the time and the final report on that
 investigation was made available to us. This enables us to define the
@@
 
-\section{Acknowldegements}\label{acknowldegements}
+\section{Author Contribution}\label{author-contribution}
 
-This project received funding from the European Union's Horizon 2020
-research and innovation programme under grant agreement No. 720270,
-Deutsche Forschungsgemeinschaft Grant DE 2175/2-1 and GR 1753/4-2 of the
-Priority Program (SPP 1665), the German-Japanese Computational
-Neuroscience Project (German Federal Ministry for Education and
-Research, BMBF Grant 01GQ1114), from the Helmholtz Portfolio Theme
-``Supercomputing and Modeling for the Human Brain'', and from the Osaka
-Univ for the project `Neural mechanism of active vision studied by
-combining large-scale sampling of neural activity and advanced
-computational analysis'.
+\textbf{VR}: Implementing the UE analysis in Python, Reproduction of the
+figures, Discussion of the results, Drafting of the manuscript, Revision
+of the draft. \textbf{JI}: Reproduction of the figures (with VR),
+Discussion of the results (with all other authors), Drafting of the
+manuscript (with VR), Revision of the draft (with all other authors).
+\textbf{MDe}: Drafting of the manuscript (with VR), Revision of the
+draft (with all other authors), Assisted in integrating the UE analysis
+in the Elephant library. \textbf{SG}: Conceived the original idea of
+reproduction, mediated the communications with original authors,
+provided suggestions for correcting errors in intermediate reproduction
+results, discussed the results, revised the manuscript (with all other
+authors).
+
+\section{Acknowledgements}\label{acknowledgements}
+
+This project received funding from EU Grant 720270 (HBP), Deutsche
+Forschungsgemeinschaft Grant DE 2175/2-1 and GR 1753/4-2 of the Priority
+Program (SPP 1665), the German-Japanese Computational Neuroscience
+Project (German Federal Ministry for Education and Research, BMBF Grant
+01GQ1114), from the Helmholtz Portfolio Theme ``Supercomputing and
+Modeling for the Human Brain'', and from the Osaka Univ for the project
+`Neural mechanism of active vision studied by combining large-scale
+sampling of neural activity and advanced computational analysis'.
 

### 2017-03-29 code/Plotting_Figures.py

@@ -1,3 +1,19 @@
+'''
+Copyright (c) 2016-2017, Vahid Rostami, Junji ito, Michael Denker, Sonja Gruen
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without modification, are permitted provided that the following conditions are met:
+
+* Redistributions of source code must retain the above copyright notice, this list of conditions and the following disclaimer.
+* Redistributions in binary form must reproduce the above copyright notice, this list of conditions and the following disclaimer in the documentation and/or other materials provided with the distribution.
+Neither the names of the copyright holders nor the names of the contributors may be used to endorse or promote products derived from this software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+'''
+
+
+
+
 import numpy
 import quantities as pq
 import elephant.unitary_event_analysis as ue
@@ -43,7 +59,7 @@ plot_params = {
     'path_filename_format': 'figure1.eps',
 }
 
-print 'plotting Figure 1 with trigger: ', trigger, '...'
+print('plotting Figure 1 with trigger: ', trigger, '...')
 plot_figure1_2(spiketrain, UE, significance_level, binsize,
[…]

### 2017-03-29 code/utils.py

@@ -1,3 +1,16 @@
+'''
+Copyright (c) 2016-2017, Vahid Rostami, Junji ito, Michael Denker, Sonja Gruen
+All rights reserved.
+
+Redistribution and use in source and binary forms, with or without modification, are permitted provided that the following conditions are met:
+
+* Redistributions of source code must retain the above copyright notice, this list of conditions and the following disclaimer.
+* Redistributions in binary form must reproduce the above copyright notice, this list of conditions and the following disclaimer in the documentation and/or other materials provided with the distribution.
+Neither the names of the copyright holders nor the names of the contributors may be used to endorse or promote products derived from this software without specific prior written permission.
+
+THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+'''
+
 import numpy
 import quantities as pq
 import matplotlib.pyplot as plt
@@ -131,8 +144,7 @@ def plot_figure1_2(
     # figure format
     plot_params = plot_params_default
     plot_params.update(plot_params_user)
-    for key, val in plot_params.items():
-        exec(key + '=val')
+    globals().update(plot_params)
     if len(unit_real_ids) != N:
[…]

### 2017-03-29 article/VRostami_JIto_MDenker_SGruen_2016.tex
corrected authors format

@@
       Vahid Rostami,
-      Junji Ito,,
-      Michael Denker,,
+      Junji Ito,
+      Michael Denker,
       Sonja Grün,
@@
                         \author[1]{Vahid Rostami}
-                    \author[1]{Junji Ito,}
-                    \author[1]{Michael Denker,}
+                    \author[1]{Junji Ito}
+                    \author[1]{Michael Denker}
                     \author[1,2]{Sonja Grün}
@@
 would not have been able to reproduce the results. Nevertheless, here
-our final validation of the reproduction is based only on a detailed
-visual comparison of the original and reproduced figures. In an optimal
+our final validation of the reproduction is based the values extracted
+from the vector graphic image (PDF) in the original paper. In an optimal
 scenario, we would be able to exactly validate the results based on a

### 2017-04-05 article/images/pre_and_wide_resnets.tex

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Florian Muellerklein,
  },
            pdftitle=
{
[Re] Reproducing CIFAR-10 results from deep and wide preactivation residual
networks
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{pre_and_wide_resnets.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Reproducing CIFAR-10 results from deep and wide preactivation residual
networks}

    \usepackage{authblk}
                        \author[1]{Florian Muellerklein}
                            \affil[1]{}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:f.muellerklein@gmail.com}{f.muellerklein@gmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/FlorianMuellerklein/ReScience-submission/tree/master/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/FlorianMuellerklein/ReScience-submission/tree/master/code}{\faGithubAlt~Code repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/FlorianMuellerklein/ReScience-submission/tree/master/data}{\faGithubAlt~Data repository}
  \end{repobox}
      \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Original article (Identity Mappings in Deep Residual Networks, Kaiming
He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun, arXiv preprint
arXiv:1603.05027, 12, April 2016)
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

In 2015, Deep Residual Networks {[}1{]} were introduced as the winning
solutions to ImageNet detection, ImageNet localization, COCO detection,
and COCO segmentation, and they made it possible to train extremely deep
neural networks of up to 1000 or more layers. The main idea is that
Residual Networks ``reformulate the layers as learning residual
functions with reference to the layer inputs, instead of learning
unreferenced functions'' {[}1{]}. The basic idea is that the residual
functions (blocks) are a combination of convolution layers and skip
connections. The residual blocks are basically two branches that come
together with an element-wise addition. One branch of the residual block
is a stack of two convolution layers and the other is an identity
function.

\includegraphics{images/residual_block.png}

Diagram of the residual block taken from {[}1{]}.

Just like with normal convolution layers, these residual blocks can be
layered to create networks of increasing depth. Below is the basic
structure of the CIFAR-10 residual network {[}1{]}{[}2{]}, with the
depth being controlled by a multiplier \emph{n} which dictates how many
residual blocks to insert between each downsampling layer. Downsampling
is done by increasing the stride of the first convolution layer in a
residual block. Whenever the number of filters are increased the first
convolution layer within a residual block will do the downsampling.

\begin{longtable}[c]{@{}lcc@{}}
\toprule
Group & Size & Multiplier\tabularnewline
\midrule
\endhead
Conv1 & {[}3x3, 16{]} & -\tabularnewline
Conv2 & {[}3x3, 16{]}{[}3x3, 16{]} & \emph{n}\tabularnewline
Conv3 & {[}3x3, 32{]}{[}3x3, 32{]} & \emph{n}\tabularnewline
Conv4 & {[}3x3, 64{]}{[}3x3, 64{]} & \emph{n}\tabularnewline
Avg-Pool & 8x8 & -\tabularnewline
Softmax & 10 & -\tabularnewline
\bottomrule
\end{longtable}

Basic structure of the CIFAR-10 Residual Network. An initial convolution
layer is followed by residual blocks of two 3x3 convolutions which are
parallel to identity mappings, the output of the identity and
convolution stacks are added after each block. The depth is mostly
altered by the multiplier n which defines how many residual blocks to
use in each section.

In addition to the stacked 3x3 convolution architecture, ``Deep Residual
Learning for Image Recognition'' also introduces a bottleneck
architecture which is designed to increase the depth without
significantly increasing the amount of parameters.{[}1{]} The structure
of this architecture can be seen below, the depth is still controlled by
a multiplier \emph{n}. The main changes are in the residual blocks.

\begin{longtable}[c]{@{}lcc@{}}
\toprule
Group & Size & Multiplier\tabularnewline
\midrule
\endhead
Conv1 & {[}3x3, 16{]} & -\tabularnewline
Conv2 & {[}1x1, 16{]}{[}3x3, 16{]}{[}1x1, 64{]} &
\emph{n}\tabularnewline
Conv3 & {[}1x1, 32{]}{[}3x3, 32{]}{[}1x1, 128{]} &
\emph{n}\tabularnewline
Conv4 & {[}1x1, 64{]}{[}3x3, 64{]}{[}1x1, 256{]} &
\emph{n}\tabularnewline
Avg-Pool & 8x8 & -\tabularnewline
Softmax & 10 & -\tabularnewline
\bottomrule
\end{longtable}

This reproduction will focus on two recent improvements on the original
residual network design, Preactivation Residual Networks and Wide
Residual Networks.{[}2{]}{[}3{]} The preactivation architecture switches
up the order of the convolution, batch normalization and nonlinearities
within each residual block. The wide architecture simply increases the
number of convolution kernels within each preactivation residual block.

\textbf{Preactivation Residual Blocks:} ``Identity Mappings in Deep
Residual Networks'' {[}2{]} introduces the preactivation architecture
which changes the order of the convolution kernels, batch
normalizations, and nonlinearities. The preactivation residual block is
said to be easier to optimize and has implicit regularization.{[}2{]}
The changes from the original residual block to the preactivation
residual block are best viewed in the figure below.

\includegraphics{images/pre_block.png}

The alterations from the original ResNet to the Preactivation ResNet.

\textbf{Wide Residual Networks:} The preactivation residual networks are
very deep but also very thin, so ``Wide Residual Networks'' {[}3{]}
suggested that better improvements can be realized with a much more
shallow and wide architecture. With normal residual networks we get
diminishing returns in performance by increasing the depth. The wide
residual networks are an attempt to mitigate this observation. The
authors introduce ``wider deep residual networks that significantly
improve over {[}2{]}, having 50 times less layers and being more than 2
times faster.'' {[}3{]} Their 16-layer wide residual Network performs as
well as a thin 1001-layer preactivation residual network. The authors
also claim that the wider networks are much more efficient on a GPU than
the deeper thin networks.{[}3{]}

\includegraphics{images/wide_block.png}

An example of the wide ResNet, it's basically a Preactivation ResNet
with an increased filter count in the residual blocks and an optional
dropout between the two convolution layers.

The wide-ResNet simply adds another multiplier \emph{k} that increases
the number of filters used in each residual block. The idea of adding
dropout in between the two convolution layers is also introduced with
the wider residual blocks. The basic structure can be seen below.

\begin{longtable}[c]{@{}lcc@{}}
\toprule
Group & Size & Multiplier\tabularnewline
\midrule
\endhead
Conv1 & {[}3x3, 16{]} & -\tabularnewline
Conv2 & {[}3x3, 16 x \emph{k}{]}{[}3x3, 16 x \emph{k}{]} &
\emph{n}\tabularnewline
Conv3 & {[}3x3, 32 x \emph{k}{]}{[}3x3, 32 x \emph{k}{]} &
\emph{n}\tabularnewline
Conv4 & {[}3x3, 64 x \emph{k}{]}{[}3x3, 64 x \emph{k}{]} &
\emph{n}\tabularnewline
Avg-Pool & 8x8 & -\tabularnewline
Softmax & 10 & -\tabularnewline
\bottomrule
\end{longtable}

\section{Methods}\label{methods}

Both the original residual network and follow up preactivation residual
network papers use identical preprocessing, training and regularization
parameters. However, the wide residual paper uses different
preprocessing, training, and regularization while still comparing
results to the previous preactivation residual network. For wide
residual networks they used ``global contrast normalization and ZCA
whitening''{[}3{]} to preprocess the CIFAR-10 images. However, the
Microsoft Research Asia group only used ``32×32 images, with the
per-pixel mean subtracted''{[}1{]} as their network inputs. It is
possible that different network architectures would require different
parameters and input data to achieve their best performance. But there
should also be comparisons done where everything stays exactly the same
except for the networks. This reproduction will be done using the
preprocessing, training and regularization parameters from the original
and preactivation residual network papers.{[}1{]}{[}2{]}

\textbf{Preprocessing:} The only preprocessing done to the CIFAR-10
images are per-pixel mean subtraction as in {[}1{]} and {[}2{]}.

\textbf{Data Augmentation:} As the data were fed into the network they
were zero-padded with 4 pixels on every side and a random crop was taken
of the original size of 32x32. This effectively results in random
translations. Additionally, a horizontal flip was applied with
probability 0.5. Unfortunately batch sizes of 128 could not be used and
instead batch sizes of 64 had to be used due to hardware constraints.

\textbf{Training and regularization:} The networks were trained with 200
epochs (full passes through training dataset) with stochastic gradient
descent, nesterov momentum of 0.9, and cross-entropy loss. For the
bottleneck architecture, the initial learning rate was set to 0.01 to
warm up the network and was increased to 0.1 at epoch 10 then continued
on the same schedule as the other networks. For all other networks the
learning rate was adjusted by the following schedule \{0:0.1, 80: 0.01,
120: 0.001\}. L2 regularization of 0.0001 was used used as in
{[}1{]}{[}2{]} and not 0.0005 like in {[}3{]}.

\textbf{Dropout:} Dropout of 0.3 was used in the wide residual
network.{[}3{]}

\textbf{Hardware and software:} This reproduction was done using the
Theano and Lasagne software frameworks for mathematical computation and
neural networks.{[}4{]}{[}5{]} A Nvidia GTX 980 GPU was used to train
and evaluate the networks. Although the impact should be insignificant
it is worth noting that the original papers were both done with the
Torch software package and a GPU with the same underlying architecture
but more memory.{[}6{]}

\section{Results}\label{results}

The reproduction results are consistent with those of the original paper
within a reasonable margin. The training parameters and preprocessing
were kept the same to the extent possible. The most notable exceptions
are the batchsize (due to hardware constraints) and the training
parameters for the wide residual network.

\begin{longtable}[c]{@{}ccc@{}}
\toprule
\textbf{ResNet Type} & \textbf{Original Paper} & \textbf{Test
Results}\tabularnewline
\midrule
\endhead
ResNet-110 & 6.37 & 6.38\tabularnewline
ResNet-164 & 5.46 & 5.66\tabularnewline
WResNet-n2-k4 & 5.55 & 5.41\tabularnewline
\bottomrule
\end{longtable}

All results are presented from the first and only training run. I did
not run each network multiple times and choose the best score.

Additionally, the training plots seemed to have the same characteristics
as those presented in the original papers. Training plots for residual
networks are characterized by fairly noisy trends in the cross-entropy
during training and also much less of a quadratic trend than seen in
other neural networks. These trends are also seen here.

\subsubsection{ResNet-110}\label{resnet-110}

\begin{figure}[htbp]
\centering
\includegraphics{images/preresnet_110_plot.png}
\caption{ResNet-110}
\end{figure}

\subsubsection{ResNet-164}\label{resnet-164}

\begin{figure}[htbp]
\centering
\includegraphics{images/preresnet_164_plot.png}
\caption{ResNet-164}
\end{figure}

\subsubsection{Wide-ResNet n=2 k=4}\label{wide-resnet-n2-k4}

\begin{figure}[htbp]
\centering
\includegraphics{images/wideresnet_n2_k4_plot.png}
\caption{Wide-ResNet}
\end{figure}

Finally, the speed of the networks may be interesting to some users.
Wide residual networks do seem to allow for more parameters with a
minimal cost in training time.

\begin{longtable}[c]{@{}ccc@{}}
\toprule
\textbf{ResNet Type} & \textbf{Params} &
\textbf{Sec/Epoch}\tabularnewline
\midrule
\endhead
ResNet-110 & 1.7M & 90\tabularnewline
ResNet-164 & 1.7M & 183\tabularnewline
WResNet-n2-k4 & 2.7M & 113\tabularnewline
\bottomrule
\end{longtable}

\section{Conclusion}\label{conclusion}

The results from the original two papers were reproduced within a small
margin which is to be expected given the stochastic properties of neural
networks. Additionally the results from the wide residual networks were
able to be reproduced even with slightly different preprocessing and
training parameters. This underscores the robust qualities and high
performance of these new neural network architectures.

\section{References}\label{references}

\begin{itemize}
\tightlist
\item
  {[}1{]} He, Kaiming, et al. ``Deep residual learning for image
  recognition.'' arXiv preprint arXiv:1512.03385 (2015).
  \href{https://arxiv.org/abs/1512.03385}{link}
\item
  {[}2{]} He, Kaiming, et al. ``Identity mappings in deep residual
  networks.'' arXiv preprint arXiv:1603.05027 (2016).
  \href{https://arxiv.org/abs/1603.05027}{link}
\item
  {[}3{]} Sergey Zagoruyko, Nikos Komodakis, ``Wide Residual Neural
  Networks.'' arXiv preprint arXiv:1605.07146 (2016).
  \href{https://arxiv.org/abs/1605.07146}{link}
\item
  {[}4{]} The Theano Development Team, et al. ``Theano: A Python
  framework for fast computation of mathematical expressions.'' arXiv
  preprint arXiv:1605.02688 (2016).
  \href{https://arxiv.org/abs/1605.02688}{link}
\item
  {[}5{]} Dieleman, Sander, et al. ``Lasagne: First Release.'' Zenodo:
  Geneva, Switzerland (2015).
  \href{http://lasagne.readthedocs.io/en/latest/index.html}{link}
\item
  {[}6{]} Collobert, Ronan, Koray Kavukcuoglu, and Clément Farabet.
  ``Torch7: A matlab-like environment for machine learning.'' BigLearn,
  NIPS Workshop. No. EPFL-CONF-192376. (2011).
  \href{http://cs.nyu.edu/~koray/files/2011_torch7_nipsw.pdf}{link}
\end{itemize}

{\sffamily \small
  \printbibliography
}
\end{document}

### 2017-04-13 article/VRostami_JIto_MDenker_SGruen_2016.tex
fix typos and translation of Figure 5 to English

@@
 each neuron from the average spike count across trials
-\autocites{Gruen02a}{Gruen02b}. For data with cross-trial variability,
-the \emph{``analytic\_TrialByTrial''} method (line 10 in the pseudo
-code) should be used, which accounts for cross-trial changes in the
-firing rates by computing the expected number of occurrences of the
-spike pattern based on the product of single-trial estimates of firing
-rates \autocite{Gruen03b}. Both methods perform a parametric test, where
-the\\
+\autocites{Gruen02a}{Gruen02b}. For data with\\
 \includegraphics{UE_algorithm.eps}\\
-number of occurrences of the spike pattern is assumed to be a
-Poisson-distributed. As an alternative non-parametric option, selecting
-\emph{``surrogate\_TrialByTrial''} as \emph{\texttt{method}} (line 15 in
-the pseudo code) will numerically compute the distribution of the
-expected number by implementing the null-hypothesis based on surrogate
-spike trains \autocite{Gruen09}. In the following, we will explain these
-methods in greater detail.
+cross-trial variability, the \emph{``analytic\_TrialByTrial''} method
+(line 10 in the pseudo code) should be used, which accounts for
+cross-trial changes in the firing rates by computing the expected number
+of occurrences of the spike pattern based on the product of single-trial
+estimates of firing rates \autocite{Gruen03b}. Both methods perform a
+parametric test, where the number of occurrences of the spike pattern is
+assumed to be a Poisson-distributed. As an alternative non-parametric
+option, selecting \emph{``surrogate\_TrialByTrial''} as
+\emph{\texttt{method}} (line 15 in the pseudo code) will numerically
+compute the distribution of the expected number by implementing the
+null-hypothesis based on surrogate spike trains \autocite{Gruen09}. In
+the following, we will explain these methods in greater detail.
 
@@
 have to focus on reproducing Figure 2A-F and Figure 4A, since for the
-other figures the data still available were either incomplete in respect
-to metadata (original Figure 3) or all the required data set were not
-available anymore (original Figure 4B,C) . Figure 2 represents the main
-result of the study and includes the UE analysis that underlies the
-subsequent analyses. Figure 4A is an example of the application of the
-UE method to data with more than 2 neurons. In terms of complexity of
-the code the implementation of the UE analysis for three or more neurons
-is considerably more demanding than for only two neurons. With this
-example we show that our implementation is capable of performing the UE
-analysis for the generic case of arbitrary numbers of neurons.
+rest of the figures data were either incomplete in respect to metadata
+(original Figure 3) or not available (original Figure 4B,C). Figure 2
+represents the main result of the study and includes the UE analysis
+that underlies the subsequent analyses. Figure 4A is an example of the
+application of the UE method to data with more than 2 neurons. In terms
+of complexity of the code the implementation of the UE analysis for
+three or more neurons is considerably more demanding than for only two
+neurons. With this example we show that our implementation is capable of
+performing the UE analysis for the generic case of arbitrary number of
+neurons.
 
@@
 diagonal line, indicating that the reproduced spike times correspond to
-to the original ones. Figure~\ref{fig:validation_fig2} B shows the same
+the original ones. Figure~\ref{fig:validation_fig2} B shows the same
 plot for the UEs, indicating the identity of the original and reproduced
@@
 \caption{\label{fig:workflow_report}Illustration of the data
-preprocessing workflow, taken from the 1997 report of MD (in German) on
-the comparison of the first UE implementation in IDL (left branch) and
-the second implementation in Matlab \autocite{Diesmann16_personalcomm}
-(right branch). Data entering both analysis branches have a resolution
-of 0.1 ms (top box). In the IDL branch spike times \(t\) in the original
-data are first transformed to a resolution of \(h=0.5\) ms (by the
-\texttt{2gdf} program, middle left) using the method of binning
-\(\lfloor t/h \rfloor\). Then the data are read into the ``IDL UE
-Software'' (lower left box) and therein converted to \(h=1\) ms
-resolution by the method of ``round half up''
+preprocessing workflow, translated from the 1997 report of MD (in
+German) on the comparison of the first UE implementation in IDL (left
+branch) and the second implementation in Matlab
+\autocite{Diesmann16_personalcomm} (right branch). Data entering both
+analysis branches have a resolution of 0.1 ms (top box). In the IDL
+branch spike times \(t\) in the original data are first transformed to a
+resolution of \(h=0.5\) ms (by the \texttt{2gdf} program, middle left)
+using the method of binning \(\lfloor t/h \rfloor\). Then the data are
+read into the ``IDL UE Software'' (lower left box) and therein converted
+to \(h=1\) ms resolution by the method of ``round half up''
 \(\lfloor t/h+1./2 \rfloor\) prior to analysis. Alternatively, one can
@@
 which differed in the events that the trials were aligned to. This
-difference in the alignment would not make any difference in the results
-if the time difference between the two events (PS and RS) were identical
-across the trials. But since the latter was not the case due to hardware
-features of the recording setup (as we learned from the first author of
-the original publication), the binning of the data started at a slightly
+difference in the alignment would not affect the results if the time
+difference between the two events (PS and RS) were identical across the
+trials. But since the latter was not the case due to hardware features
+of the recording setup (as we learned from the first author of the
+original publication), the binning of the data started at a slightly
 different time points in different trials. This likely led to a loss or
@@
 would not have been able to reproduce the results. Nevertheless, here
-our final validation of the reproduction is based the values extracted
-from the vector graphic image (PDF) in the original paper. In an optimal
-scenario, we would be able to exactly validate the results based on a
-numerical comparison. To do so, all of the following pieces of
-information would have had to be available at hand:
+our final validation of the reproduction is based on the values
+extracted from the vector graphic image (PDF) in the original paper. In
+an optimal scenario, we would be able to exactly validate the results
+based on a numerical comparison. To do so, all of the following pieces
+of information would have had to be available at hand:
 

### 2017-04-23 article/VRostami_JIto_MDenker_SGruen_2016.tex
changes according to review

@@
 plot for the UEs, indicating the identity of the original and reproduced
-UE timings as well. To further confirm the identity, we plot in
-Figure~\ref{fig:validation_fig2} C the distributions of the differences
-between the original and the reproduced timings for the spikes (black)
-and the UEs (red). The differences are at most +/- 0.3 ms, which are
-considerably narrower than the +/- 1 ms differences caused by the
-misaligned data shown in Figure~\ref{fig:figure2alignedPS} A (see for
-comparison the gray plot in Figure~\ref{fig:validation_fig2} C, which
-shows the differences between the original spike times and the
-misaligned spike times). The remaining minor differences of the spike
-and UE times of the correctly aligned data and the original data are
-only due to slight errors in the extraction of the spike times from the
-original figure, which is inevitable because of a limited precision of
-the plotting commands in the PDF file. Thus, we confirm that spike
-timings in the correctly aligned data are identical to those in the
-original data, and our implementation of the UE analysis applied to
-these data reproduces exactly the same results as shown in the original
-paper.
+UE timings as well. To further confirm the identity, we plot the
+distributions of the differences between the original and the reproduced
+timings for the spikes (Figure~\ref{fig:validation_fig2} C; black) and
+the UEs (Figure~\ref{fig:validation_fig2} D). The differences are at
+most +/- 0.3 ms, which are considerably narrower than the +/- 1 ms
+differences caused by the misaligned data shown in
+Figure~\ref{fig:figure2alignedPS} A (see for comparison the gray plot in
+Figure~\ref{fig:validation_fig2} C, which shows the differences between
+the original spike times and the misaligned spike times). The remaining
+minor differences of the spike and UE times of the correctly aligned
+data and the original data are only due to slight errors in the
+extraction of the spike times from the original figure, which is
+inevitable because of a limited precision of the plotting commands in
+the PDF file. Thus, we confirm that spike timings in the correctly
+aligned data are identical to those in the original data, and our
+implementation of the UE analysis applied to these data reproduces
+exactly the same results as shown in the original paper.
 
@@
 of the differences between the original and the reproduced timings for
-spikes (black) and UEs (red) in Figure~\ref{fig:figure2alignedRS}. The
-distribution of the differences between the original and the reproduced
-timings for spikes in Figure~\ref{fig:figure2alignedPS} is shown in
-gray.}\label{fig:validationux5ffig2}
+spikes in Figure ~\ref{fig:figure2alignedRS} and
+Figure~\ref{fig:figure2alignedPS} are shown in black and gray,
+respectively. \textbf{D)} Same plot as in C but for UEs in Figure
+~\ref{fig:figure2alignedRS}.}\label{fig:validationux5ffig2}
 \end{figure}
@@
 \centering
-\includegraphics{figure3.eps}
+\includegraphics{figure6.eps}
 \caption{\label{fig:reproducedFig4A}Reproduction of Figure 4A of the

### 2017-04-23 code/Plotting_Figures.py

@@ -12,8 +12,6 @@ THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND 
 '''
 
 
-
-
 import numpy
 import quantities as pq
 import elephant.unitary_event_analysis as ue
@@ -56,7 +54,7 @@ plot_params = {
     'events': {'0\nPS': [300 * pq.ms], '600\nES1': [900 * pq.ms],
                'ES2': [1200 * pq.ms], 'ES3': [1500 * pq.ms],
                '1500\nRS': [1800 * pq.ms]},
-    'path_filename_format': 'figure1.eps',
+    'path_filename_format': '../article/figure1.eps',
 }
 
 print('plotting Figure 1 with trigger: ', trigger, '...')
@@ -82,7 +80,7 @@ UE = ue.jointJ_window_analysis(
     spiketrain, binsize, winsize, winstep, pattern_hash, method=method)
 
 # parameters for plotting
-plot_params = {'path_filename_format': 'figure2.eps'}
+plot_params = {'path_filename_format': '../article/figure2.eps'}
 
 print()
 print('plotting Figure 2 with trigger: ', trigger, '...')
@@ -105,7 +103,7 @@ t_post = 499 * pq.ms
 spiketrains2 = load_gdf2Neo(data_path + fname, trigger, t_pre, t_post)
 
 print()
-print('plotting Figure 3 ...')
+print('plotting Figure 6 ...')
 sts_lst = [spiketrains1, spiketrains2]
 Js_dict_lst_lst = []
 for spiketrains in sts_lst:
@@ -124,7 +122,7 @@ Js_dict = {
     'events1': {'PS\n0': [100 * pq.ms], 'ES1\n600': [700 * pq.ms]},
     'events2': {'RS\n': [100 * pq.ms]},
     'unit_real_ids': [3, 4, 5],
-    'path_filename_format': 'figure3.eps',
+    'path_filename_format': '../article/figure6.eps',
     'figsize': (8, 7),
     'fsize': 12,
     'ms': 2,
@@ -132,5 +130,5 @@ Js_dict = {
 }
 
 
-plot_figure3(sts_lst, Js_dict_lst_lst, significance_level, binsize,
+plot_figure6(sts_lst, Js_dict_lst_lst, significance_level, binsize,
              winsize, winstep, patterns_hash, N, N_comb, Js_dict)

### 2017-04-23 code/utils.py

@@ -23,7 +23,7 @@ plot_params_default = {
     # epochs to be marked on the time axis
     'events': [],
     # save figure
-    'save_fig': False,
+    'save_fig': True,
     # show figure
     'showfig': True,
     # figure size
@@ -325,7 +325,7 @@ def plot_figure1_2(
         plt.show()
 
 
-def plot_figure3(
+def plot_figure6(
         data_list, Js_dict_lst_lst, sig_level, binsize, winsize,
         winstep, patterns, N, N_comb, plot_params_user):
     """plots Figure 3 of the manuscript"""
@@ -365,7 +365,6 @@ def plot_figure3(
             ax0.set_yticklabels(['5', '4', '3'], fontsize=fsize)
             ax0.spines['right'].set_visible(False)
             ax0.yaxis.set_ticks_position('left')
-            # ax0.text(-110, 190, 'Neuron #', rotation=90, fontsize=fsize)
             ax0.set_ylabel('Neuron #', fontsize=fsize)
         else:
             ax0.spines['left'].set_visible(False)
@@ -389,7 +388,7 @@ def plot_figure3(
         ax1.set_xticks([])
         if data_cnt == 0:
             ax1.set_title('Coincident Events', fontsize=fsize)
-            ax1.text(-110, 190, 'Neuron #', rotation=90, fontsize=fsize)
+            ax1.set_ylabel('Neuron #', fontsize=fsize)
             ax1.set_yticks([num_tr / 2 + 1, num_tr * 3 / 2., num_tr * 5 / 2])
             ax1.set_yticklabels(['5', '4', '3'], fontsize=fsize)
             ax1.spines['right'].set_visible(False)
@@ -429,8 +428,7 @@ def plot_figure3(
         ax2.set_xticks([])
         if data_cnt == 0:
             ax2.set_title('Unitary Events', fontsize=fsize)
-            ax2.text(-110, 190, 'Neuron #', rotation=90,
-                     fontsize=fsize)
+            ax2.set_ylabel('Neuron #', fontsize=fsize)
             ax2.set_yticks([num_tr / 2 + 1, num_tr * 3 / 2.,
                             num_tr * 5 / 2])
             ax2.set_yticklabels(['5', '4', '3'], fontsize=fsize)
@@ -452,7 +450,8 @@ def plot_figure3(
                                  markersize=ms1, color='k')
                     if patt_tmp[n_cnt][0] == 1:
                         sig_idx_win = numpy.where(
[…]

### 2017-05-02 code/Plotting_Figures.py
add the code for generating Figure 3 and extracting values from eps

@@ -61,6 +61,8 @@ print('plotting Figure 1 with trigger: ', trigger, '...')
 plot_figure1_2(spiketrain, UE, significance_level, binsize,
                winsize, winstep, pattern_hash, N, plot_params)
 
+
+########################################################
 # Figure 2 (alignment of the trials on the RS)
 # This is the reproduction of Figure 2 of
 # the original article after cutting the data
@@ -88,7 +90,44 @@ plot_figure1_2(spiketrain, UE, significance_level, binsize,
                winsize, winstep, pattern_hash, N, plot_params)
 
 
+########################################################
 # Figure 3
+# load spike data of figure 2
+file_name = 'winny131_23.gdf'
+trigger = 'RS_4'
+t_pre = 1799 * pq.ms
+t_post = 300 * pq.ms
+spiketrain_fig2 = load_gdf2Neo(data_path + file_name, trigger, t_pre, t_post)
+
+# calculating UE ...
+UE = ue.jointJ_window_analysis(
+    spiketrain_fig2, binsize, winsize, winstep, pattern_hash, method=method)
+
+# load spike data of figure 1
+file_name = 'winny131_23.gdf'
+trigger = 'PS_4'
+t_pre = 300 * pq.ms
+t_post = 1800 * pq.ms
+spiketrain_fig1 = load_gdf2Neo(data_path + file_name, trigger, t_pre, t_post)
+
+# load extracted data from Riehle et al
+extracted_data = numpy.load(
+    data_path + 'extracted_data.npy', encoding='latin1').item()
+
+# parameters for plotting
+plot_params = {
+    'figsize': (7.5, 7.5),
+    'path_filename_format': '../article/figure3.eps'}
+
+print()
+print('plotting Figure 3 ...')
+plot_figure3([spiketrain_fig2, spiketrain_fig1], UE, extracted_data,
+             significance_level, binsize, winsize,
+             winstep, plot_params)
+
+
+########################################################
+# Figure 6
 # This is the reproduction of Figure 4A of the original article
 
 fname = 'jenny201_345_preprocessed.gdf'

### 2017-05-02 code/utils.py

@@ -325,10 +325,151 @@ def plot_figure1_2(
         plt.show()
 
 
+
+def plot_figure3(sts_lst, UE, extracted_data, significance_level,
+                 binsize, winsize, winstep, plot_params_user):
+    """plots Figure 3 of the manuscript"""
+
+    import matplotlib.pylab as plt
+    # figure format
+    plot_params = plot_params_default
+    plot_params.update(plot_params_user)
+    for key, val in plot_params.items():
+        exec(key + '=val')
+    t_start = sts_lst[0][0][0].t_start
+    t_stop = sts_lst[0][0][0].t_stop
+    t_winpos = ue._winpos(t_start, t_stop, winsize, winstep)
+    f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2)
+    f.set_size_inches(figsize)
+    plt.subplots_adjust(top=.9, right=.92, left=.12,
+                        bottom=.1, hspace=.4, wspace=.4)
+    fsize = 12
+    # scatter plot of panel A of figure 2 of the original publication
+    ax = ax1
+    ax.locator_params(nbins=4, axis='x')
+    ax.locator_params(nbins=4, axis='y')
+    ys_cnt = 0
+    diff_spikes_rep = []
+    spiketrain = sts_lst[0]
+    for cnt_i, i in enumerate(spiketrain.T):
+        for cnt_j, j in enumerate(i):
+            sp_trial = extracted_data['spikes'][ys_cnt]
+            if len(j) >= len(sp_trial):
+                for cnt, sp_t in enumerate(sp_trial):
+                    min_positive = numpy.argmin(numpy.abs(j.magnitude - sp_t))
+                    ax.plot(j.magnitude[min_positive], sp_t, ',', color='k')
+                    diff_spikes_rep.append(j.magnitude[min_positive] - sp_t)
+            else:
+                for cnt, sp_t in enumerate(j.magnitude):
+                    min_positive = numpy.argmin(numpy.abs(sp_trial - sp_t))
+                    ax.plot(sp_t, sp_trial[min_positive], ',', color='k')
+                    diff_spikes_rep.append(sp_t - sp_trial[min_positive])
+            ys_cnt += 1
+
+    ax.set_ylabel('spike time [ms] (reproduced)')
+    ax.set_xlabel('spike time [ms] (original article)')
+    ax.text(-0.1, 1.07, 'A', transform=ax.transAxes, size=fsize + 5,
+            weight='bold')
+
[…]

### 2017-05-23 article/Metzner-2017.tex
added Noise Exploration subsection and updated conclusion

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Metzner Christoph,
  },
            pdftitle=
{
[Re] Modeling GABA Alterations in Schizophrenia: A Link Between Impaired
Inhibition and Gamma and Beta Auditory Entrainment
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Modeling GABA Alterations in Schizophrenia: A Link Between Impaired
Inhibition and Gamma and Beta Auditory Entrainment}

    \usepackage{authblk}
                        \author[1]{Metzner Christoph}
                            \affil[1]{Centre of Computer Science and Informatics Research,University of
Hatfield, Hatfield, Hertfordshire, UK}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:c.metzner@herts.ac.uk}{c.metzner@herts.ac.uk}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Modeling GABA Alterations in Schizophrenia: A Link Between Impaired
Inhibition and Gamma and Beta Auditory Entrainment, D.
Vierling-Claassen, P. Siekmeier, S. Stufflebeam and N. Kopell, Journal
of Neurophysiology(99):2656-2671, 2008, doi:10.1152/jn.00870.2007
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

We provide an implementation of \autocite{Vierling2008}, which models
impaired auditory entrainment in the gamma range for schizophrenic
patients. Particularly, we only reimplement the simplified network model
and do not replicate the biophysically more detailed Genesis model which
is also developed in the article.

In the original article the authors conduct an MEG study of auditory
entrainment and find changes in gamma and beta range entrainment found
in schizophrenic patients. These changes in oscillatory dynamics in
patients are important for two reasons: First, gamma range oscillations
emerge in a variety of different tasks and are thought to underlie many
cognitive processes (see e.g. \autocite{Fries2005,Fries2015}). Furthermore, impaired gamma
range oscillations in schizophrenic patientas have been found in most of
these tasks and might offer an explanation for the cognitive deficits
found in patients. Therefore, the study of the mechanisms underlying
oscillatory entrainment might shed light on the basic principles that
are responsible for a wide range of deficits in schizophrenic patients
Second, oscillatory entrainment is a promising candidate for a
neurophysiological biomarker for schizophrenia
\autocite{Siekmeier2015}. In the original article they develop two
models that are able to account for the changes they find in their
experimental data. One is a biophysically detailed network model and the
other one a simplified network model. In both models a prolonged decay
at GABAergic inhibitory synapses causes the reduction in gamma
entrainment and the increase in beta entrainment. The simplified network
model, despite its many simplifications, offers insight into the
mechanisms underlying these entrainment deficits while focusing on a few
key parameters. This makes it easy to distill and understand underlying
dynamics and mechanisms and, at the same time, makes simulations
computationally very inexpensive allowing for an extensive exploration
of the parameter space. Additionally, again because of the simplicity,
the model can easily be extended to study the effect other parameters
(e.g.~sparsity of connections, different populations of inhibitory
neurons, different types of synaptic receptors,\ldots{}), without
loosing much of its simplicity.

We focus on the main results of the original article: an increase in
inhibitory decay time constants leads to a reduction of power in the
gamma range and an increase in power in the beta range, replicating
experimental findings for schizophrenic patients. Therefore, we
reproduce Figures 4,5,6,7,10 and 11 of the original paper. The original
model is implemented using Matlab but the source code is not publicly
available. The model and analysis scripts are implemented using Python
2.7.9. All simulations were run under Ubuntu 15.04.

\section{Methods}\label{methods}

The model was implemented solely from the paper description, since the
original code is not publicly available. The model is a simple model
consisting of two neural populations (excitatory and inhibitory cells).
Individual cells are modeled as theta neurons (for a detailed discussion
of this neuron model see \autocite{Boergers2003}). The neuron is
described by a single variable \(\theta\), which can be regarded as the
membrane voltage, subject to the following dynamics

\[\frac{d \theta}{dt}=1-\cos \theta + (b+S+N(t))\cdot(1+ \cos \theta),\]

where \(b\) is an externally applied current, \(S\) is the total
synaptic input to the cell and \(N(t)\) is a time-varying noise input.
Total synaptic input to a cell in a network is calculated as

\[S_k = \sum_{j=1}^n \alpha_j \cdot g_{jk} \cdot s_{jk},\]

where \(\alpha _j=\pm 1\) controls excitation and inhibition, \(g_{jk}\)
is the synaptic strength from cell \(j\) to cell \(k\) and \(s_{jk}\) is
the synaptic gating variable from cell \(j\) to cell \(k\). Synaptic
gating variables are subject to the following dynamics

\[\frac{ds_{jk}}{dt}= - \frac{s _{jk}}{\tau _j} + e ^{- \eta \cdot (1+ \cos \theta _j)} \cdot \frac{1-s _{jk}}{\tau _R},\]

where \(\tau _j\) is the synaptic decay time and \(\tau_R\) the synaptic
rise time. The network receives excitatory drive input at click train
frequency from a single pacemaker cell. Additionally, Poissonian noise
input is also given to all cells in the network. A noise spike at time
\(t_n\) elicits the following `EPSC'

\[N=H(t-t_n) \cdot \frac{A \cdot g _{gmax} \cdot (e^{-(t-t _n)/ \tau _{exc}} - e^{-(t-t _n)/ \tau _R} )}{\tau _{exc} - \tau _R},\]

where \(A \cdot g_{gmax}\) is the strength of the noise and again
\(\tau _{exc}\) is the synaptic decay time and \(\tau_R\) the synaptic
rise time.

Each population connects to itself and to the other with an all-to-all
connectivity. Both populations also have two sources of input, the
oscillatory drive input and a background noise input. The drive input
periodically sends spikes to both populations with a given frequency,
whereas the background noise input sends noise spikes at times drawn
from a Poisson distribution. Table \ref{tbl:summary} summarizes the
network model. Tables \ref{tbl:modparameters} and
\ref{tbl:simparameters} list the parameters of the model and the
simulations, their definitions and values, respectively.

The model was implemented using Python 2.7.9 using numpy 1.9.3.
Visualisation of results was also done in Python using the matplotlib
module (matplotlib 1.4.3). Furthermore, since the model is
computationally very inexpensive, we did not aim to provide the most
efficient implementation but rather an implementation that is clear, and
easy to understand and use.

All differential equations are solved using a simple forward Euler
scheme. As in the original article a single simulation simulatesa 500ms
trial and the time step is chosen such that this results in 2\^{}13 =
8192 data points. However, the main results are unaffected by a smaller
time step.

\begin{longtable}[]{@{}ll@{}}
\caption{Model summary }\label{tbl:summary}\tabularnewline
\toprule
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Populations\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
One excitatory and one inhibitory population\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Topology\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
None\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Connectivity\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
All-to-all\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Neuron model\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
Theta model\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Synapse model\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
(Quasi-)Instantaneous rise, exponential decay\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
External input\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
Poisson noise and periodic drive to both populations\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Recordings\strut
\end{minipage} & \begin{minipage}[t]{0.75\columnwidth}\raggedright\strut
Theta variables (both populations); `MEG' signal (summed EPSCs at exc.
neurons)\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}lll@{}}
\caption{Model parameters} \label{tbl:modparameters}\tabularnewline
\toprule
Parameter & Definition & Value\tabularnewline
\midrule
\endfirsthead
\toprule
Parameter & Definition & Value\tabularnewline
\midrule
\endhead
\(n_E\) & Exc. population size & 20\tabularnewline
\(n_I\) & Inh.population size & 10\tabularnewline
\(\tau_R\) & Synaptic rise time & 0.1\tabularnewline
\(\tau_{exc}\) & Excitatory decay time & 2.0\tabularnewline
\(\tau_{inh}\) & Inhibitory decay time (control) & 8.0\tabularnewline
\(\tau_{inh}\) & Inhibitory decay time (schizophrenia) &
28.0\tabularnewline
\(g_{ee}\) & E-E synaptic strength & 0.015\tabularnewline
\(g_{ei}\) & E-I synaptic strength & 0.025\tabularnewline
\(g_{ie}\) & I-E synaptic strength & 0.015\tabularnewline
\(g_{ii}\) & I-I synaptic strength & 0.02\tabularnewline
\(g_{de}\) & Synaptic strength of drive to E cells & 0.3\tabularnewline
\(g_{di}\) & Synaptic strength of drive to I cells & 0.08\tabularnewline
\(b\) & Applied current (regardless of cell type) & -0.1\tabularnewline
\(Ag_{max}\) & Scaling factor for noise EPSCs & 0.5\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}ll@{}}
\caption{Simulation parameters }\label{tbl:simparameters}\tabularnewline
\toprule
Parameter & Value\tabularnewline
\midrule
\endfirsthead
\toprule
Parameter & Value\tabularnewline
\midrule
\endhead
Time step (\(dt\)) & 0.061\tabularnewline
Number of data points & 8192 (\(=2^{13}\))\tabularnewline
Total time & 500ms\tabularnewline
\bottomrule
\end{longtable}

\section{Results}\label{results}

As explained in the introduction, we only replicated the simple model
from \autocite{Vierling2008}, and \textbf{not} the GENESIS model. We
aimed to reproduce Figures 4 (raw, simulated MEG signal) and 5 (power
spectra for MEG signals from Figure 4) from \autocite{Vierling2008},
which show the main results of the model (summarized in Table
\ref{tbl:mainresults}).

\begin{longtable}[]{@{}lll@{}}
\caption{Model parameters}\label{tbl:mainresults}\tabularnewline
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Drive\strut
\end{minipage} & \begin{minipage}[b]{0.41\columnwidth}\raggedright\strut
Control subjects\strut
\end{minipage} & \begin{minipage}[b]{0.42\columnwidth}\raggedright\strut
Schizophrenic patients\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Drive\strut
\end{minipage} & \begin{minipage}[b]{0.41\columnwidth}\raggedright\strut
Control subjects\strut
\end{minipage} & \begin{minipage}[b]{0.42\columnwidth}\raggedright\strut
Schizophrenic patients\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
40 Hz\strut
\end{minipage} & \begin{minipage}[t]{0.41\columnwidth}\raggedright\strut
Strong entrainment to the drive, no power in frequency bands apart from
40 Hz\strut
\end{minipage} & \begin{minipage}[t]{0.42\columnwidth}\raggedright\strut
Weaker entrainment to the drive, emergence of a subharmonic component
(at 20 Hz)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
30 Hz\strut
\end{minipage} & \begin{minipage}[t]{0.41\columnwidth}\raggedright\strut
Strong entrainment to the drive, no power in frequency bands apart from
30 Hz\strut
\end{minipage} & \begin{minipage}[t]{0.42\columnwidth}\raggedright\strut
Strong entrainment to the drive, no power in frequency bands apart from
30 Hz\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
20 Hz\strut
\end{minipage} & \begin{minipage}[t]{0.41\columnwidth}\raggedright\strut
Entrainment to the drive, however, more power in the harmonic 40 Hz
band\strut
\end{minipage} & \begin{minipage}[t]{0.42\columnwidth}\raggedright\strut
Stronger entrainment to the drive, less power in the harmonic band\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Figures \ref{fig:Vierling4} and \ref{fig:Vierling5} show the
output of the replicated model for the same simulations as for Figures 4
and 5 from the original article. The main charactersitics described
above can be clearly seen. However, in our model these main features are
a little bit less pronounced than in the original model. Since the
network model receives Poissonian noise (which is quite strong), this
difference may simply stem from a difference in noise. Furthermore, we
have to mention the differences in amplitude for the simulated MEG
signals (and accordingly the power spectra thereof) between the original
model and our replication, which we believe stems from a scaling in the
original model. However, since the original source code is not
available, we cannot verify that.

\begin{figure}
\centering
\includegraphics{Replication-Figure4.eps}
\caption{Replication of Figure 4: Raw simulated MEG signals (averaged
over 20 trials) for the control and the schizophrenic network at the
three different driving frequencies.}\label{fig:Vierling4}
\end{figure}

\begin{figure}
\centering
\includegraphics{Replication-Figure5.eps}
\caption{Replication of Figure 5: Power spectra of the averaged MEG
signals from \ref{fig:Vierling4}}\label{fig:Vierling5}
\end{figure}

After having looked at the model output averaged over 20 trials with
different noise, we also show single trial data which exemplify the main
model features , as was done in the original article. Figures
\ref{fig:Vierling6} and fig:Vierling7 show the model output in
response to 40 Hz drive input for the control and the schizophrenia
network, respectively. The strong entrainment in the control case, the
reduction of entrainment and the emergence of a subharmonic 20 Hz
component are again clearly visible. However, as before, in our model
implementation the emergent 20 Hz component is less pronounced than in
the original implementation (best seen in the excitatory population
activitydisplayed in the raster plot of \ref{fig:Vierling7}).

\begin{figure}
\centering
\includegraphics{Replication-Figure6.eps}
\caption{Replication of Figure 6: Single trial from the control network.
40 Hz drive. Network entrains to 40 Hz, as can be seen in the frequency
diagram, raster plot and MEG trace.}\label{fig:Vierling6}
\end{figure}

\begin{figure}
\centering
\includegraphics{Replication-Figure7.eps}
\caption{Replication of Figure 7: Single trial from the schizophrenia
network. 40 Hz drive. Network entrains to 40 Hz but also shows a strong
20 Hz component, as can be seen in the frequency diagram, raster plot
and MEG trace. Especially the inhibitory neurons only entrain to a 20 Hz
rhythm (see raster plot).}\label{fig:Vierling7}
\end{figure}

Figures \ref{fig:Vierling10} and \ref{fig:Vierling11} show the
model output in response to 20 Hz drive for the control and the
schizophrenia network, respectively. Again, main features of the
original model are faithfully reproduced.

\begin{figure}
\centering
\includegraphics{Replication-Figure10.eps}
\caption{Replication of Figure 10: Single trial from the control
network. 20 Hz drive. Network entrains to 20 Hz but also shows a 40 Hz
component, as can be seen in the frequency diagram, raster plot and MEG
trace.}\label{fig:Vierling10}
\end{figure}

\begin{figure}
\centering
\includegraphics{Replication-Figure11.eps}
\caption{Replication of Figure 11: Single trial from the schizophrenia
network. 20 Hz drive. Network entrains to 20 Hz without 40 Hz component,
as can be seen in the frequency diagram, raster plot and MEG trace. note
that the 40 Hz power in the power spectrum is a
harmonic.}\label{fig:Vierling11}
\end{figure}

\subsection{Exploration of the Discrepancies between Original and Reimplementation}
In this section we explore possible scenarios that might explain the less pronounced main features in the reimplementation.

We explore the influence of the background noise. The main mechanism behind the emergence of a 20 Hz component in response
to 40 Hz drive in the schizophrenia network is the prolonged inhibition which suppresses activity in every second 40 Hz cycle. In the absence
of noise the network responds with a pure 20 Hz oscillation (see Figures \ref{fig:NoiseExplorationMEG} (a) and \ref{fig:NoiseExplorationPSD} (a)). However, the background noise
gives sufficient input to some excitatory cells to overcome this suppression and also fire in between the 20 Hz cycles. This results in a mixed response, 
where power is split between 20 Hz and 40 Hz. The ratio of this split depends on the strength of the noise. Therefore, we asked whether the less pronounced
20 Hz component in our reimplementation might simply come from a too high background noise amplitude. Figures \ref{fig:NoiseExplorationMEG} (b) and (c) and \ref{fig:NoiseExplorationPSD} (b) and (c), show the 
response (MEG signal and power spectral density, respectivey) to a 40 Hz drive in the cases, where we scaled down the noise to 75\% and 50\%, respectively. It can clearly be seen that the 20 Hz component increases, when
the noise strength decreases (full noise strength can be seen in panel (d) of both Figures). However, as can be seen in panels (e) to (h), showing the response of the control network to 20 Hz drive,
is that downscaling of the background noise reduces the 40 Hz component in response to 20 Hz drive. This is not surprising since background noise drives excitatory cells to fire in between 20 Hz cycles in the control network, resulting
in the 40 Hz component. This component is suppressed in the schizophrenia network due to the prolonged inhibition.
Summarising, a difference in the strength of the background noise might be an explanation for the discrepancy between original and reimplementation. However, only a reduction of noise solely for the schizophrenia network 
fully replicates the original results.

\section{Conclusion}\label{conclusion}

Overall, we believe we have faithfully reproduced the main fetaures of
the simple model from \autocite{Vierling2008}. However, we note that
overall features a re a little bit less pronounced in our
reimplementation compared to the original model. We have explored possible explanations for this discrepancy
and conclude that it might come from minor inconsistencies with respect to parameters in the original model. Nevertheless,
the mechanism proposed in the original model could be reproduced and presents a possible explanation for deficits in 
auditory steady-state responses in patients suffering from schizophrenia.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-05-23 article/Metzner-2017.tex
update article; replace .md file with .tex file

@@
 \lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}
+
+\usepackage{placeins}
 
@@
 All differential equations are solved using a simple forward Euler
-scheme. As in the original article a single simulation simulatesa 500ms
+scheme. As in the original article a single simulation simulates a 500ms
 trial and the time step is chosen such that this results in 2\^{}13 =
@@
 \begin{longtable}[]{@{}lll@{}}
-\caption{Model parameters}\label{tbl:mainresults}\tabularnewline
+\caption{Main model result features}\label{tbl:mainresults}\tabularnewline
 \toprule
@@
 to 40 Hz drive in the schizophrenia network is the prolonged inhibition which suppresses activity in every second 40 Hz cycle. In the absence
-of noise the network responds with a pure 20 Hz oscillation (see Figures \ref{fig:NoiseExplorationMEG} (a) and \ref{fig:NoiseExplorationPSD} (a)). However, the background noise
+of noise the network responds with a pure 20 Hz oscillation (see Figures \ref{fig:NoiseExplorationMEG} left upper panel and \ref{fig:NoiseExplorationPSD} left upper panel).
+However, the background noise
 gives sufficient input to some excitatory cells to overcome this suppression and also fire in between the 20 Hz cycles. This results in a mixed response, 
@@
 20 Hz component in our reimplementation might simply come from a too high background noise amplitude.
-Figures \ref{fig:NoiseExplorationMEG} (b) and (c) and \ref{fig:NoiseExplorationPSD} (b) and (c), show the
+Figures \ref{fig:NoiseExplorationMEG} (left middle and lower panels  and \ref{fig:NoiseExplorationPSD} left middle and lower panels, show the
 response (MEG signal and power spectral density, respectivey) to a 40 Hz drive in the cases, where we scaled down the noise to 75\% and 50\%, respectively.
 It can clearly be seen that the 20 Hz component increases, when
-the noise strength decreases (full noise strength can be seen in panel (d) of both Figures). However, as can be seen in panels (e) to (h), showing the response of the control network to 20 Hz drive,
+the noise strength decreases. However, as can be seen in right panels, showing the response of the control network to 20 Hz drive,
 is that downscaling of the background noise reduces the 40 Hz component in response to 20 Hz drive.
@@
 fully replicates the original results.
+
+\begin{figure}
+\centering
+\includegraphics{Noise-Exploration-MEG.eps}
+\caption{Replication of Figure 11: Single trial from the schizophrenia
+network. 20 Hz drive. Network entrains to 20 Hz without 40 Hz component,
+as can be seen in the frequency diagram, raster plot and MEG trace. note
+that the 40 Hz power in the power spectrum is a
+harmonic.}\label{fig:NoiseExplorationMEG}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Noise-Exploration-PSD.eps}
+\caption{Replication of Figure 11: Single trial from the schizophrenia
+network. 20 Hz drive. Network entrains to 20 Hz without 40 Hz component,
+as can be seen in the frequency diagram, raster plot and MEG trace. note
+that the 40 Hz power in the power spectrum is a
+harmonic.}\label{fig:NoiseExplorationPSD}
+\end{figure}
+
+\FloatBarrier
 

### 2017-06-20 article/Metzner-2017.tex
fixed issues aaronshifman

@@
 in schizophrenic patients. These changes in oscillatory dynamics in
-patients are important for two reasons: First, gamma range oscillations
+patients are important for two reasons: first, gamma range oscillations
 emerge in a variety of different tasks and are thought to underlie many
@@
 oscillatory entrainment might shed light on the basic principles that
-are responsible for a wide range of deficits in schizophrenic patients
+are responsible for a wide range of deficits in schizophrenic patients.
 Second, oscillatory entrainment is a promising candidate for a
@@
 described by a single variable \(\theta\), which can be regarded as the
-membrane voltage, subject to the following dynamics
+neuron state, subject to the following dynamics
 
@@
 
-where \(\alpha _j=\pm 1\) controls excitation and inhibition, \(g_{jk}\)
+where \(\alpha _j\) controls excitation and inhibition, i.e. is $+1$ for excitatory synapses and $-1$ for inhibitory ones,
+\(g_{jk}\)
 is the synaptic strength from cell \(j\) to cell \(k\) and \(s_{jk}\) is
@@
 oscillatory drive input and a background noise input. The drive input
-periodically sends spikes to both populations with a given frequency,
-whereas the background noise input sends noise spikes at times drawn
+periodically sends spikes to both populations with a given frequency. In order to generate the spikes
+a drive cell is implemented (also a theta neuron), which receives an applied current so that its spike frequency
+matches the driving frequency. This drive cell is then connected to all cells in the network.
+The background noise input sends noise spikes at times drawn
 from a Poisson distribution. Table \ref{tbl:summary} summarizes the
@@
 
-All differential equations are solved using a simple forward Euler
-scheme. As in the original article a single simulation simulates a 500ms
-trial and the time step is chosen such that this results in 2\^{}13 =
+All differential equations were solved using a simple forward Euler
+scheme. As in the original article a single simulation simulated a 500ms
+trial and the time step was chosen such that this results in 2\^{}13 =
 8192 data points. However, the main results are unaffected by a smaller
@@
 model features , as was done in the original article. Figures
-\ref{fig:Vierling6} and fig:Vierling7 show the model output in
+\ref{fig:Vierling6} and \ref{fig:Vierling7} show the model output in
 response to 40 Hz drive input for the control and the schizophrenia

### 2017-06-20 article/Metzner-2017.tex
defined epsc and eta

@@
 Individual cells are modeled as theta neurons (for a detailed discussion
-of this neuron model see \autocite{Boergers2003}). The neuron is
-described by a single variable \(\theta\), which can be regarded as the
+of this neuron model see \autocite{Boergers2003}). The $k$th neuron in a network is
+described by a single variable \(\theta_k\), which can be regarded as the
 neuron state, subject to the following dynamics
 
-\[\frac{d \theta}{dt}=1-\cos \theta + (b+S+N(t))\cdot(1+ \cos \theta),\]
+\[\frac{d \theta_k}{dt}=1-\cos \theta_k + (b+S_k+N(t))\cdot(1+ \cos \theta_k),\]
 
@@
 
-where \(\tau _j\) is the synaptic decay time and \(\tau_R\) the synaptic
-rise time. The network receives excitatory drive input at click train
+where \(\tau _j\) is the synaptic decay time, \(\tau_R\) the synaptic
+rise time and \(\eta\) is a scaling parameter. The network receives excitatory drive input at click train
 frequency from a single pacemaker cell. Additionally, Poissonian noise
 input is also given to all cells in the network. A noise spike at time
-\(t_n\) elicits the following `EPSC'
+\(t_n\) elicits the following excitatory postsynaptic potential (`EPSC')
 

### 2017-08-02 article/detorakis-2017.tex
New submission

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Georgios Detorakis,
  },
            pdftitle=
{
[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
        Spiking Behavior
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

\usepackage{float,lscape}
\usepackage{xcolor,colortbl}
\usepackage{amssymb,amsmath,array,amsthm}
\usepackage{marvosym}
% \usepackage{makecell}
\usepackage{boldline}
\usepackage{bm,array}

\usepackage{adjustbox}

\newcommand{\Rm}[1]{\mathrm{#1}}

\newcolumntype{C}{>{\centering\arraybackslash}p{3.5em}}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\definecolor{Gray}{gray}{0.75}
\definecolor{LightGray}{gray}{0.95}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces
            Diverse Spiking Behaviors}

    \usepackage{authblk}
                        \author[1]{Georgios Detorakis}
                            \affil[1]{Department of Cognitive Sciences, UC Irvine,
                            Irvine, CA, USA}

\date{\vspace{-5mm}
      \sffamily \small
      \href{mailto:corresponding-author@mail.com}{gdetorak@uci.edu \\
      gdetor@protonmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The author has declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] A Generalized Linear Integrate-and-Fire Neural Model Produces
        Diverse Spiking Behaviors, Stefan Mihalas and Ernst Niebur, Neural
        Computation 21, 704--718, 2009.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Integrate-and-fire neurons are being used extensively in the field of 
neuroscience for modeling spiking behaviors~\cite{dayan:2001}. In this work we
provide a reference implementation of~\cite{mihalas:2009}, where the authors
have introduced a generalization of the leaky integrate-and-fire neuron model. 
The Mihalas-Niebur Neuron (MNN) model is a linear integrate-and-fire neuron
model capable of expressing a rich spiking behavior based on a set of 
parameters.

An MNN model expresses tonic and phasic spiking, class $1$ and $2$, spike
frequency adaptation, accommodation, threshold variability, rebound spike,
integrator, input bistability, hyperpolarizing spiking and bursting, tonic,
phasic and rebound bursting, mixed mode, afterpotentials, basal bistability, 
preferred frequency and spike latency. 
Due to its simplicity, the MNN model has been used in neuromorphic 
implementations such as~\cite{folowosele:2011}.

The model consists of linear differential equations, which describe the membrane
and threshold potentials and internal currents. All the results provided in 
\cite{mihalas:2009} have been obtained by using only two internal currents and
thus we use the exact same number of internal currents in this work. The
differential equations describe the subthreshold dynamics, while the 
instantaneous threshold potential controls when the neuron firing an action 
potential (spike) in a dynamic way. The ability of the MNN model to generate
such a diverse spiking behavior is due to the complex update rules. In this 
work the MNN model has been implemented in Python (version 3.6.1) using 
Numpy (version 1.13.1) and Matplotlib (version 2.0.2) packages.

\section{Methods}\label{methods}

In order to implement the model described in \cite{mihalas:2009}, we discretized
the dynamical system using the Euler's integration scheme. The time step is 
fixed to $0.1\, \Rm{ms}$ for all the simulations, and the simulation time
varies according to figure 1 of the original paper. We provide all equations
and parameters of the model in tables as it has been suggested
by~\cite{nordlie:2009}. 
Table~\ref{table:1} provides the summary of the model. Tables~\ref{table:2}
and~\ref{table:3} give the subthreshold dynamics (differential equations) 
describing the membrane and the threshold potentials as well as the two internal
currents and the update rules. The parameters for all the simulations are given
in table~\ref{table:4}. Table~\ref{table:5} provides the external current
intensities (and pulse durations), and initial conditions are described
in table~\ref{table:6}.

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ll}
        \thickhline
        \multicolumn{2}{c}{Model Summary} \\\thickhline
        \rowcolor{Gray}
        Populations  & No population -- single neuron model \\\rowcolor{LightGray}
        Topology     & -- \\ \rowcolor{Gray}
        Connectivity & -- \\ \rowcolor{LightGray}
        Neuron Model & Linear Integrate-and-Fire Neuron \\\rowcolor{Gray}
        Channel Models & Linear, first order ODEs  \\ \rowcolor{LightGray}
        Synapse Model & -- \\ \rowcolor{Gray}
        Plasticity & -- \\ \rowcolor{LightGray}
        Input & Constant current or rectangular pulses \\\rowcolor{Gray}
        Measurements & Membrane potential, phase plane \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Summary of the model}} 
    \label{table:1}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{3.5cm}ll}
        \thickhline
        \multicolumn{2}{c}{Neuron Model} \\\thickhline
        \rowcolor{Gray}
        Name  &  Mihalas-Niebur Neuron (MNN) \\ \rowcolor{LightGray}
        Type  &  Linear Leaky Integrate-and-Fire Neuron  \\ \rowcolor{Gray}
        Membrane Potential & $
            \begin{aligned}
                \frac{\Rm{d}V(t)}{\Rm{d}t} &= \frac{1}{C} \Big(I_e + I_1 + I_2 - G(V(t) -
                E_L)  \Big)
            \end{aligned}$ \\ \rowcolor{LightGray}
        Instantaneous Threshold Potential  & $
            \begin{aligned}
                \frac{\Rm{d}\Theta(t)}{\Rm{d}t} &= a(V(t) - E_L) - b(\Theta(t) -
                \Theta_{\infty}) 
            \end{aligned} $
            \\ \rowcolor{Gray}
        Internal Currents & $
            \begin{aligned}
                \frac{\Rm{d}I_{1}(t)}{\Rm{d}t} &= -k_1I_1(t)\\
                \frac{\Rm{d}I_{2}(t)}{\Rm{d}t} &= -k_2I_2(t) \\
            \end{aligned} $
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the subthreshold dynamics of
    Mihalas-Niebur neuron model.} $V(t)$ and $\Theta(t)$ are the membrane and 
    threshold potentials, respectively. $E_L$ and $\Theta_{\infty}$ are the
    reversal potentials for the membrane and the threshold variables,
    respectively. $a, b, k_1, k_2$ and $G$ are constant parameters. $I_e$ is
    the external current applied on the neuron model.} 
    \label{table:2}
\end{table}
%%

\begin{table}[!htbp]
    \centering
    \begin{tabular}{cc}
        \thickhline
        \multicolumn{2}{c}{Update rules} \\ \thickhline
        Variable & Rule  \\ \rowcolor{Gray}
        $V(t)$ & $V_r$  \\ \rowcolor{LightGray}
        $\Theta(t)$ & $\max\{\Theta_r, \Theta(t) \}$ \\ \rowcolor{Gray}
        $I_1(t)$ & $R_1 \times I_1(t) + A_1$ \\ \rowcolor{LightGray} 
        $I_2(t) $ & $R_2 \times I_2(t) + A_2$ \\\thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Update rules.} $V_r$ and $\Theta_r$
    are the reset values for the membrane and threshold potentials,
    respectively. $R_1, R_2, A_1$ and $A_2$ are constants.}
    \label{table:3}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{CCCCC}
        \thickhline
        \multicolumn{5}{c}{Model Parameters} \\\thickhline
           \begin{tabular}[x]{@{}c@{}} Figure \\  \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $a$ \\ $(\Rm{s^{-1}})$ \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $A_1/C$ \\ $(\Rm{V/s})$ \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $A_2/C$ \\ $(\Rm{V/s})$ \end{tabular}& 
           \begin{tabular}[x]{@{}c@{}} $t_f$ \\ $(\Rm{s})$ \end{tabular} \\
        \thickhline
        1A & $0$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
        1B & $0$ & $0$ & $0$ & $0.5$  \\\rowcolor{LightGray}
        1C & $5$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
        1D & $5$ & $0$ & $0$ & $0.5$ \\\rowcolor{LightGray}
        1E & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{Gray}
        1F & $5$ & $0$ & $0$ & $0.4$ \\\rowcolor{LightGray}
        1G & $5$ & $0$ & $0$ & $1.0$  \\\rowcolor{Gray}
        1H &  $5$ & $0$ & $0$ & $0.3$ \\\rowcolor{LightGray}
        1I & $5$  & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
        1J & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{LightGray}
        1K & $30$ & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
        1L & $30$  & $10$ & $-0.6$ & $0.4$ \\\rowcolor{LightGray}
        1M & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{Gray}
        1N & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{LightGray}
        1O & $5$ & $10$ & $-0.6$ & $1.0$ \\\rowcolor{Gray}
        1P & $5$  & $5$ & $-0.3$ & $0.5$ \\\rowcolor{LightGray}
        1Q & $5$  & $5$ & $-0.3$ & $0.2$ \\\rowcolor{Gray}
        1R & $0$ & $8$ & $-0.1$ & $0.2$ \\\rowcolor{LightGray}
        1S & $5$ & $-3$ & $0.5$ & $0.8$ \\\rowcolor{Gray}
        1T &  $-80$ & $0$ & $0$ & $0.05$ \\
        \thickhline
        \multicolumn{5}{l}{Common Parameters} \\\rowcolor{LightGray}
        \thickhline
        \multicolumn{5}{p{22.5em}}{
            $b = 10\, \Rm{s^{-1}}$, $G/C = 50\, \Rm{s^{-1}}$,
            $k_1 = 200\, \Rm{s^{-1}}$, $k_2 = 20\, \Rm{s^{-1}}$,
            $\Theta_{\infty} = -0.05\, \Rm{V}$,
            $R_1 = 0$, $R_2 = 1$, $E_l = -0.07\, \Rm{V}$,
            $V_r = -0.07\, \Rm{V}$, $\Theta_r = -0.06\, \Rm{V}$.
        }
        \\\thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulation Parameters}  }
    \label{table:4}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{CCp{28.2em}}
        \thickhline
        \multicolumn{3}{c}{Model Parameters} \\ \thickhline
        Figure & Type & $I_e/C (\Rm{V/s})$  \\
        \thickhline
        1A & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{Gray}
        1B & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1 + 10^{-6}$ \\\rowcolor{LightGray}
        1C & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
        1D &\includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{LightGray}
        1E &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
              $1.5 (0.1 \Rm{s}),\, 0 (0.5\Rm{s}),\, 0.5 (0.1\Rm{s}),\, 1
              (0.1\Rm{s}),\, 1.5 (0.1\Rm{s}),\, 0 (0.1 \Rm{s})$
              \\\rowcolor{Gray}
        1F &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.02
        \Rm{s}),\, 0 (0.18 \Rm{s}),\, -1.5 (0.025 \Rm{s}),\, 0 (0.025
        \Rm{s}),\, 1.5 (0.025 \Rm{s}),\, 0 (0.125 \Rm{s})$
        \\\rowcolor{LightGray} 
        1G & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.05
        \Rm{s}),\, -3.5 (0.756 \Rm{s}),\, 0 (0.194 \Rm{s})$ \\\rowcolor{Gray} 
        1H & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2(1 +
        10^{-6})$ \\\rowcolor{LightGray}
        1I & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
        \begin{tabular}[x]{@{}c@{}} 
        $1.5 (0.02 \Rm{s}),\, 0 (0.01 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\,
            0 (0.25 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\, 0 (0.02 \Rm{s}) $\\
            $ 1.5 (0.02 \Rm{s}),\, 0 (0.04 \Rm{s})$ 
        \end{tabular}
            \\\rowcolor{Gray}
        1J & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.1
        \Rm{s}),\, 1.7 (0.4 \Rm{s}),\, 1.5 (0.1 \Rm{s}),\, 1.7 (0.4 \Rm{s})$
        \\\rowcolor{LightGray}
        1K & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$  \\\rowcolor{Gray}
        1L & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$ \\\rowcolor{LightGray}
        1M & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
        1N & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$  \\\rowcolor{LightGray}
        1O &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.1 
        \Rm{s}),\, -3.5 (0.5 \Rm{s}),\, 0 (0.4 \Rm{s})$ \\\rowcolor{Gray}
        1P & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{LightGray}
        1Q & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $2 (0.015
        \Rm{s}),\, 0 (0.185 \Rm{s})$  \\\rowcolor{Gray}
        1R & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}& $5 (0.01
        \Rm{s}),\, 0 (0.09 \Rm{s}),\, 5 (0.01 \Rm{s}),\, 0 (0.09 \Rm{s})$
        \\\rowcolor{LightGray}
        1S & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}&
        \begin{tabular}[x]{@{}c@{}} 
            $5 (0.005 \Rm{s}),\, 0 (0.005 \Rm{s}),\, 4 (0.005 \Rm{s}),\, 0 (0.385 \Rm{s}),\,
            5 (0.005 \Rm{s}),\, 0 (0.045 \Rm{s}) $\\
            $4 (0.005\Rm{s}),\, 0 (0.345 \Rm{s}) $
        \end{tabular}
        \\\rowcolor{Gray}
        1T & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $8 (0.002 \Rm{s}),\, 0 (0.048 \Rm{s})$ \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily External current.} This table provides the
    external current for each panel in Figure~\ref{fig:1}. There are two types
    of external currents, constants and pulses. In the case of pulses the
    duration of each pulse is given in seconds along with its intensity.
    }
    \label{table:5}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{cc}
        \thickhline
        \multicolumn{2}{c}{Initial Conditions} \\ \thickhline
        Variable & Initial Value  \\ 
        \rowcolor{Gray}
        $V(t)$ & $-0.07\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{LightGray}
        $\Theta(t)$ & $-0.05\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{Gray}
        $I_1(t)$ & $0.01\, \Rm{V}$ \\ \rowcolor{LightGray} 
        $I_2(t) $ & $0.001\, \Rm{V}$ \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Initial conditions.} In all simulations
    have been used the same initial conditions, except from the one 
    illustrated in figure~\ref{fig:1}H.}
    \label{table:6}
\end{table}
%%

All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
generation i$7$ processor, $16\, \Rm{GB}$ of physical memory and running Arch
Linux. The total execution time of all simulations was $2.41$ seconds and the
peak consumed memory was $162\, \Rm{MB}$\footnote{Python memory profiler used
    (\url{https://pypi.python.org/pypi/memory_profiler}).}\@.

\clearpage

\section{Results}\label{results}

All three figures from the original article have been successfully replicated.
All the different spiking behaviors of the model are illustrated in 
Figure~\ref{fig:1}, where the black solid line indicates the membrane potential
($V(t)$), the red dashed line illustrates the instantaneous threshold potentials
($\Theta(t)$), and the gray line shows the input to the neuron ($I_e/C$).
Figures~\ref{fig:2} and~\ref{fig:3} depict the phase space of the phasic
spiking ($V(t)$ and $\Theta(t)$) and phasic bursting ($V(t)$,
$I_1(t)$, and $I_2(t)$). All the figures express the same qualitative
behavior as the original figures in~\cite{mihalas:2009}.

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=1.1\textwidth]{figs/Figure01.pdf}
    \caption{{\bfseries \sffamily Neural responses of MNN.} Black solid lines
    indicate the membrane potential ($V(t)$), the red dashed lines show the
    threshold potentials ($\Theta(t)$), and the gray lines the external
    currents applied on each case.
    {\bfseries \sffamily A} tonic spiking,
    {\bfseries \sffamily B} class $1$,
    {\bfseries \sffamily C} spike frequency adaptation,
    {\bfseries \sffamily D} phasic spiking,
    {\bfseries \sffamily E} accommodation,
    {\bfseries \sffamily F} threshold variability,
    {\bfseries \sffamily G} rebound spike,
    {\bfseries \sffamily H} class $2$,
    {\bfseries \sffamily I} integrator,
    {\bfseries \sffamily J} input bistability,
    {\bfseries \sffamily K} hyperpolarization induced spiking,
    {\bfseries \sffamily L} hyperpolarization induced bursting,
    {\bfseries \sffamily M} tonic bursting,
    {\bfseries \sffamily N} phasic bursting,
    {\bfseries \sffamily O} rebound burst,
    {\bfseries \sffamily P} mixed mode,
    {\bfseries \sffamily Q} afterpotentials,
    {\bfseries \sffamily R} basal bistability,
    {\bfseries \sffamily S} preferred frequency,
    {\bfseries \sffamily T} spike latency.} 
    \label{fig:1}
\end{figure}

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/Figure02.pdf}
    \caption{{\bfseries \sffamily Phase space of phasic spiking.} Blue
    solid lines indicate the trajectories of the model in the phase spiking
    behavior (Figure~\ref{fig:1}D). The dashed line corresponds to $V(t) =
    \Theta(t)$, and the black dots represent spike events. The parameters for 
    this simulation are the same as in Figure~\ref{fig:1}D.}
    \label{fig:2}
\end{figure}

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/Figure03.pdf}
    \caption{{\bfseries \sffamily Phase space of phasic bursting.} Blue solid
    lines represent the trajectories of the system and the black dots indicate
    spiking events. The parameters for this simulation are the same as in
    Figure~\ref{fig:1}N.}
    \label{fig:3}
\end{figure}

\section{Conclusion}\label{conclusion}

All figures in~\textcite{mihalas:2009} have been successfully replicated 
with high fidelity. Overall, the whole reproducing process was smooth and
without obscure points since most of the parameters are provided in the
original article. Only the external current time intervals and the initial
conditions are not provided explicitly and thus we had to extract that
information from figure $1$ of the original article. To conclude, the
article~\cite{mihalas:2009} has been successfully reproduced without any
discrepancy.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-08-02 code/neuron_model.py

@@ -1,0 +1,122 @@
+# -*- coding: utf-8 -*-
+# This script implements the Mihalas-Niebur Neuron model based on the paper
+# S. Mihalas and E. Niebur, "A Generalized Linear Integrate-and-Fire Neural
+# Model Produces Diverse Spiking Behaviors", Neural Computation 21, 2009.
+#
+# Copyright (C) 2017  Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software; you can redistribute it and/or modify it
+# under the terms of the GNU General Public License as published by the
+# Free Software Foundation; either version 2 of the License, or (at your
+# option) any later version.
+# This program is distributed in the hope that it will be useful, but WITHOUT
+# ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
+# FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for
+# more details. You should have received a copy of the GNU General Public
+# License along with this program; if not, write to
+# the Free Software Foundation, Inc.,
+# 51 Franklin Street,
+# Fifth Floor, Boston, MA  02110-1301, USA.
+import numpy as np
+# from memory_profiler import profile
+
+
+# Define simulation and model parameters dictionary
+params = {'k1': 0.2,             # I₁ decay term
+          'k2': 0.02,            # I₂ decay term
+          'b': 0.01,             # Θ decay term
+          'R1': 0.0,             # I₁ update constant
+          'R2': 1.0,             # I₂ update constant
+          'El': -70.0,           # Reverse potential
+          'Vr': -70.0,           # Resting potential
+          'Thetar': -60.0,       # Resting threshold
+          'a': 0.000,            # Specifier of spike response
+          'A1': 0.0,             # I₁ additive update constant
+          'A2': 0.0,             # I₂ additive update constant
+          'G': 0.05,             # Membrane potential decay term
+          'C': 1.0,              # Membrane capacitance
+          'Iext': np.zeros((200,)),    # External current
+          'ThetaInf': -50.0}     # Reverse threshold
+
+
+# @profile
+def mnn_model(pms=params, time=10, dt=1.0, IC=(0.01, 0.001, -70.0, -50.0)):
+    """ Main simulation function. Actual implementation of MNN model.
+
+        Args:
[…]

### 2017-08-02 code/run_all.py

@@ -1,0 +1,435 @@
+# This script implements the Mihalas-Niebur Neuron model based on the paper
+# S. Mihalas and E. Niebur, "A Generalized Linear Integrate-and-Fire Neural
+# Model Produces Diverse Spiking Behaviors", Neural Computation 21, 2009.
+#
+# Copyright (C) 2017  Georgios Is. Detorakis (gdetor@protonmail.com)
+#
+# This program is free software; you can redistribute it and/or modify it
+# under the terms of the GNU General Public License as published by the
+# Free Software Foundation; either version 2 of the License, or (at your
+# option) any later version.
+# This program is distributed in the hope that it will be useful, but WITHOUT
+# ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
+# FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for
+# more details. You should have received a copy of the GNU General Public
+# License along with this program; if not, write to
+# the Free Software Foundation, Inc.,
+# 51 Franklin Street,
+# Fifth Floor, Boston, MA  02110-1301, USA.
+import numpy as np
+from copy import copy
+from neuron_model import mnn_model
+import matplotlib.pylab as plt
+
+
+# Define simulation and model parameters dictionary
+params = {'k1': 0.2,             # I₁ decay term
+          'k2': 0.02,            # I₂ decay term
+          'b': 0.01,             # Θ decay term
+          'R1': 0.0,             # I₁ update constant
+          'R2': 1.0,             # I₂ update constant
+          'El': -70.0,           # Reverse potential
+          'Vr': -70.0,           # Resting potential
+          'Thetar': -60.0,       # Resting threshold
[…]

### 2017-08-10 article/Metzner-2017.tex
fixed typos, fiyed figure captions, rearranged figures and text

@@
 cognitive processes (see e.g. \autocite{Fries2005,Fries2015}). Furthermore, impaired gamma
-range oscillations in schizophrenic patientas have been found in most of
+range oscillations in schizophrenic patients have been found in most of
 these tasks and might offer an explanation for the cognitive deficits
@@
 neurons, different types of synaptic receptors,\ldots{}), without
-loosing much of its simplicity.
+losing much of its simplicity.
 
@@
 experimental findings for schizophrenic patients. Therefore, we
-reproduce Figures 4,5,6,7,10 and 11 of the original paper. The original
+reproduce Figures 4, 5, 6, 7, 10 and 11 of the original paper. The original
 model is implemented using Matlab but the source code is not publicly
@@
 
-where \(\alpha _j\) controls excitation and inhibition, i.e. is $+1$ for excitatory synapses and $-1$ for inhibitory ones,
+where \(n\) is the number of presynaptic neurons, \(\alpha _j\) controls excitation and inhibition, i.e. is $+1$ for excitatory synapses and $-1$ for inhibitory ones,
 \(g_{jk}\)
@@
 
-The model was implemented using Python 2.7.9 using numpy 1.9.3.
+The model was implemented using Python 2.7.9 and numpy 1.9.3.
 Visualisation of results was also done in Python using the matplotlib
@@
 \centering
-\includegraphics{Replication-Figure6.eps}
+\includegraphics[width=0.9\textwidth]{Replication-Figure6.eps}
 \caption{Replication of Figure 6: Single trial from the control network.
@@
 \centering
-\includegraphics{Replication-Figure7.eps}
+\includegraphics[width=0.9\textwidth]{Replication-Figure7.eps}
 \caption{Replication of Figure 7: Single trial from the schizophrenia
@@
 \centering
-\includegraphics{Replication-Figure10.eps}
+\includegraphics[width=0.9\textwidth]{Replication-Figure10.eps}
 \caption{Replication of Figure 10: Single trial from the control
@@
 
+\FloatBarrier
+
 \begin{figure}
 \centering
-\includegraphics{Replication-Figure11.eps}
+\includegraphics[width=0.9\textwidth]{Replication-Figure11.eps}
 \caption{Replication of Figure 11: Single trial from the schizophrenia
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.65\textwidth]{Noise-Exploration-MEG.eps}
+\caption{Raw simulated MEG signals (averaged
+over 20 trials) for the control and the schizophrenic network at 20 and 40 Hz driving frequencies 
+for three different levels of noise strength (0, 50 and 75 \% of the strength in the standard model).}\label{fig:NoiseExplorationMEG}
+\end{figure}
+
+\FloatBarrier
+
 \subsection{Exploration of the Discrepancies between Original and Reimplementation}
 In this section we explore possible scenarios that might explain the less pronounced main features in the reimplementation.
+We explore the influence of the background noise. The main mechanism behind the emergence of a 20 Hz component in response
+to 40 Hz drive in the schizophrenia network is the prolonged inhibition which suppresses activity in every second 40 Hz cycle.
 
-We explore the influence of the background noise. The main mechanism behind the emergence of a 20 Hz component in response
-to 40 Hz drive in the schizophrenia network is the prolonged inhibition which suppresses activity in every second 40 Hz cycle. In the absence
+In the absence
 of noise the network responds with a pure 20 Hz oscillation (see Figures \ref{fig:NoiseExplorationMEG} left upper panel and \ref{fig:NoiseExplorationPSD} left upper panel).
@@
 \centering
-\includegraphics{Noise-Exploration-MEG.eps}
-\caption{Replication of Figure 11: Single trial from the schizophrenia
-network. 20 Hz drive. Network entrains to 20 Hz without 40 Hz component,
-as can be seen in the frequency diagram, raster plot and MEG trace. note
-that the 40 Hz power in the power spectrum is a
-harmonic.}\label{fig:NoiseExplorationMEG}
+\includegraphics[width=\textwidth]{Noise-Exploration-PSD.eps}
+\caption{Power spectra of the raw MEG signals from \ref{fig:NoiseExplorationMEG}}\label{fig:NoiseExplorationPSD}
 \end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Noise-Exploration-PSD.eps}
-\caption{Replication of Figure 11: Single trial from the schizophrenia
-network. 20 Hz drive. Network entrains to 20 Hz without 40 Hz component,
-as can be seen in the frequency diagram, raster plot and MEG trace. note
-that the 40 Hz power in the power spectrum is a
-harmonic.}\label{fig:NoiseExplorationPSD}
-\end{figure}
-
-\FloatBarrier
 
@@
 the simple model from \autocite{Vierling2008}. However, we note that
-overall features a re a little bit less pronounced in our
+overall features are a little bit less pronounced in our
 reimplementation compared to the original model. We have explored possible explanations for this discrepancy

### 2017-08-10 article/detorakis-2017.tex
reviewers comments addressed

@@
 thus we use the exact same number of internal currents in this work. The
-differential equations describe the subthreshold dynamics, while the 
-instantaneous threshold potential controls when the neuron firing an action 
-potential (spike) in a dynamic way. The ability of the MNN model to generate
+subthreshold dynamics are defined by a set of linear ordinary differential 
+equations, while an instantaneous threshold potential controls when the neuron
+firing an action potential (spike) in a dynamic way. The ability of the MNN 
+model to generate 
 such a diverse spiking behavior is due to the complex update rules. In this 
@@
 In order to implement the model described in \cite{mihalas:2009}, we discretized
-the dynamical system using the Euler's integration scheme. The time step is 
-fixed to $0.1\, \Rm{ms}$ for all the simulations, and the simulation time
-varies according to figure 1 of the original paper. We provide all equations
-and parameters of the model in tables as it has been suggested
-by~\cite{nordlie:2009}. 
+the dynamical system using the forward Euler's integration scheme. The time step
+is fixed to $0.1\, \Rm{ms}$ for all the simulations, and the total simulation
+time $t_f$ varies according to figure 1 of the original paper. Our
+implementation differs from the one in the original paper, since in
+\cite{mihalas:2009}, authors numerically solve equation $3.5$ (algebraic 
+equation) under the constraint imposed by inequality $3.4$ and thus they
+compute the spike times. On the other hand, in this work we directly compute
+numerically the solution of the dynamical system defined by equations $2.1$
+and $2.2$ in \cite{mihalas:2009} (see tables~\ref{table:2} and~\ref{table:3}).
+
+We provide all equations and parameters of the model in tables as it has been 
+suggested by~\cite{nordlie:2009}. 
 Table~\ref{table:1} provides the summary of the model. Tables~\ref{table:2}
@@
 currents and the update rules. The parameters for all the simulations are given
-in table~\ref{table:4}. Table~\ref{table:5} provides the external current
-intensities (and pulse durations), and initial conditions are described
-in table~\ref{table:6}.
+in table~\ref{table:4}, while the external current intensities and pulse
+duration are provided in table~\ref{table:5}. The parameters in this work are 
+exactly the same used in the original paper (table 1, pg. $711$). We had to 
+infer the time intervals and the total simulation times for the pulses since
+they are not given explicitly in the original paper. Thus, we extracted the
+time intervals from figure $1$ of \cite{mihalas:2009} by visual inspection. 
+The initial conditions are given in table~\ref{table:6}.
+
+All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
+generation i$7$ processor, $16\, \Rm{GB}$ of physical memory and running Arch
+Linux (x$86\_64$). The total execution time of all simulations was $2.41$
+seconds and the peak consumed memory was $162\, \Rm{MB}$\footnote{Python memory
+profiler used (\url{https://pypi.python.org/pypi/memory_profiler}).}\@.
 
@@
 
+\vspace{-1cm}
+
 %%
@@
 %%
-
+\vspace{-2cm}
 %%
@@
     have been used the same initial conditions, except from the one 
-    illustrated in figure~\ref{fig:1}H.}
+    illustrated in Figure~\ref{fig:1}H.}
     \label{table:6}
@@
 %%
-
-All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
-generation i$7$ processor, $16\, \Rm{GB}$ of physical memory and running Arch
-Linux. The total execution time of all simulations was $2.41$ seconds and the
-peak consumed memory was $162\, \Rm{MB}$\footnote{Python memory profiler used
-    (\url{https://pypi.python.org/pypi/memory_profiler}).}\@.
 
@@
 ($\Theta(t)$), and the gray line shows the input to the neuron ($I_e/C$). 
+The $x$-axis scale in all the panels are exactly the same as in the original
+paper (indicating the total simulation time ($t_f$), while the $y$-axis scale 
+differs from the one in the original paper. In this work the $y$-axis scale
+is the same same for all the subplots ($[-95, -25]\, \Rm{mV}$), except from
+panels G and O ($[-145, -25]\, \Rm{mV}$).
+
 Figures~\ref{fig:2} and~\ref{fig:3} depict the phase space of the phasic
 spiking ($V(t)$ and $\Theta(t)$) and phasic bursting ($V(t)$,
-$I_1(t)$, and $I_2(t)$). All the figures express the same qualitative
-behavior as the original figures in~\cite{mihalas:2009}.
+$I_1(t)$, and $I_2(t)$). In both figures the blue curves and the black dots 
+indicate the trajectory of the system and spike events, respectively.
+In Figure~\ref{fig:2} the gray arrows show the evolution of the system
+(vector field of the system). Figure~\ref{fig:3} has a different orientation 
+from the original one but both illustrate the same trajectories and spike 
+events of the system. 
+All the figures express the same qualitative behavior as the original
+figures in~\cite{mihalas:2009}.
 
@@
     \centering
-    \includegraphics[width=1.1\textwidth]{figs/Figure01.pdf}
+    \includegraphics[width=0.9\textwidth]{figs/Figure01.pdf}
     \caption{{\bfseries \sffamily Neural responses of MNN.} Black solid lines
@@
 without obscure points since most of the parameters are provided in the
-original article. Only the external current time intervals and the initial
-conditions are not provided explicitly and thus we had to extract that
-information from figure $1$ of the original article. To conclude, the
+original article. Only the time intervals for which the external current is
+applied to the model and the initial conditions are not provided explicitly.
+Therefore, we had to extract that information from figure $1$ of the original
+article. To conclude, the
 article~\cite{mihalas:2009} has been successfully reproduced without any

### 2017-08-10 code/neuron_model.py

@@ -1,22 +1,33 @@
+#!/usr/bin/env python
 # -*- coding: utf-8 -*-
 # This script implements the Mihalas-Niebur Neuron model based on the paper
 # S. Mihalas and E. Niebur, "A Generalized Linear Integrate-and-Fire Neural
 # Model Produces Diverse Spiking Behaviors", Neural Computation 21, 2009.
 #
-# Copyright (C) 2017  Georgios Is. Detorakis (gdetor@protonmail.com)
+# Copyright (c) 2017 Georgios Is. Detorakis
+# All rights reserved.
 #
-# This program is free software; you can redistribute it and/or modify it
-# under the terms of the GNU General Public License as published by the
-# Free Software Foundation; either version 2 of the License, or (at your
-# option) any later version.
-# This program is distributed in the hope that it will be useful, but WITHOUT
-# ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-# FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for
-# more details. You should have received a copy of the GNU General Public
-# License along with this program; if not, write to
-# the Free Software Foundation, Inc.,
-# 51 Franklin Street,
-# Fifth Floor, Boston, MA  02110-1301, USA.
+# Redistribution and use in source and binary forms, with or without
+# modification, are permitted provided that the following conditions
+# are met:
+# 1. Redistributions of source code must retain the above copyright
+#    notice, this list of conditions and the following disclaimer.
+# 2. Redistributions in binary form must reproduce the above copyright
+#    notice, this list of conditions and the following disclaimer in the
+#    documentation and/or other materials provided with the distribution.
+# 3. The name of the author may not be used to endorse or promote products
+#    derived from this software without specific prior written permission.
+#
+# THIS SOFTWARE IS PROVIDED BY THE AUTHOR ``AS IS'' AND ANY EXPRESS OR
+# IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES
+# OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED.
+# IN NO EVENT SHALL THE AUTHOR BE LIABLE FOR ANY DIRECT, INDIRECT,
[…]

### 2017-08-10 code/run_all.py

@@ -1,25 +1,37 @@
+#!/usr/bin/env python
 # This script implements the Mihalas-Niebur Neuron model based on the paper
 # S. Mihalas and E. Niebur, "A Generalized Linear Integrate-and-Fire Neural
 # Model Produces Diverse Spiking Behaviors", Neural Computation 21, 2009.
 #
-# Copyright (C) 2017  Georgios Is. Detorakis (gdetor@protonmail.com)
+# Copyright (c) 2017 Georgios Is. Detorakis
+# All rights reserved.
 #
-# This program is free software; you can redistribute it and/or modify it
-# under the terms of the GNU General Public License as published by the
-# Free Software Foundation; either version 2 of the License, or (at your
-# option) any later version.
-# This program is distributed in the hope that it will be useful, but WITHOUT
-# ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-# FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for
-# more details. You should have received a copy of the GNU General Public
-# License along with this program; if not, write to
-# the Free Software Foundation, Inc.,
-# 51 Franklin Street,
-# Fifth Floor, Boston, MA  02110-1301, USA.
+# Redistribution and use in source and binary forms, with or without
+# modification, are permitted provided that the following conditions
+# are met:
+# 1. Redistributions of source code must retain the above copyright
+#    notice, this list of conditions and the following disclaimer.
+# 2. Redistributions in binary form must reproduce the above copyright
+#    notice, this list of conditions and the following disclaimer in the
+#    documentation and/or other materials provided with the distribution.
+# 3. The name of the author may not be used to endorse or promote products
+#    derived from this software without specific prior written permission.
+#
[…]

### 2017-08-16 article/detorakis-2017.tex
more corrections

@@
 equations, while an instantaneous threshold potential controls when the neuron
-firing an action potential (spike) in a dynamic way. The ability of the MNN 
+fires an action potential (spike) in a dynamic way. The ability of the MNN 
 model to generate 
@@
 In order to implement the model described in \cite{mihalas:2009}, we discretized
-the dynamical system using the forward Euler's integration scheme. The time step
+the dynamical system using the forward Euler integration scheme. The time step
 is fixed to $0.1\, \Rm{ms}$ for all the simulations, and the total simulation
@@
 ($\Theta(t)$), and the gray line shows the input to the neuron ($I_e/C$). 
-The $x$-axis scale in all the panels are exactly the same as in the original
+The $x$-axis scales in all panels are exactly the same as in the original
 paper (indicating the total simulation time ($t_f$), while the $y$-axis scale 
 differs from the one in the original paper. In this work the $y$-axis scale
-is the same same for all the subplots ($[-95, -25]\, \Rm{mV}$), except from
+is the same same for all the subplots ($[-95, -25]\, \Rm{mV}$), except for
 panels G and O ($[-145, -25]\, \Rm{mV}$).

### 2017-08-16 code/neuron_model.py

@@ -86,24 +86,24 @@ def mnn_model(pms=None, Iext=np.zeros((200,)), dt=1.0,
                                potential at the spike-event times.
             spikes (ndarray)  : The spike trains.
     """
-    sim_time = Iext.shape[0]                    # Simulation time
+    sim_steps = Iext.shape[0]                    # Total simulation steps
 
     if pms is None:
         pms = default_parameters()
 
     # ODEs Initial conditions
-    I1 = IC[0] * np.ones((sim_time, ))          # Internal current 1
-    I2 = IC[1] * np.ones((sim_time, ))          # Internal current 2
-    V = IC[2] * np.ones((sim_time, ))           # Membrane potential
-    Theta = IC[3] * np.ones((sim_time, ))       # Threshold potential
+    I1 = IC[0] * np.ones((sim_steps, ))          # Internal current 1
+    I2 = IC[1] * np.ones((sim_steps, ))          # Internal current 2
+    V = IC[2] * np.ones((sim_steps, ))           # Membrane potential
+    Theta = IC[3] * np.ones((sim_steps, ))       # Threshold potential
 
     # Auxiliary vectors
-    theta_ = np.zeros((sim_time, ))                 # Threshold track
-    time_ = np.zeros((sim_time, ))                  # Time vector
+    theta_ = np.zeros((sim_steps, ))                 # Threshold track
+    time_ = np.zeros((sim_steps, ))                  # Time vector
 
     spikes = []                                 # Spike-event times
     # Forward Euler integration of MNN
-    for t in range(1, sim_time):
+    for t in range(1, sim_steps):
         # ODEs
         I1[t] = I1[t-1] + dt * (-pms['k1'] * I1[t-1])
         I2[t] = I2[t-1] + dt * (-pms['k2'] * I2[t-1])

### 2017-08-17 article/stojic2017.tex
Stojic-2017: first draft of the Hinton and Nowlan 1987 replication

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Hrvoje Stojić,
  },
            pdftitle=
{
[Re] How learning can guide evolution
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

% typesetting algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{stojic2017.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] How learning can guide evolution}

    \usepackage{authblk}
                        \author[1]{Hrvoje Stojić}
                            \affil[1]{Department of Economics and Business, Universitat Pompeu Fabra,
Barcelona, Spain}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:hrvoje.stojic@protonmail.com}{hrvoje.stojic@protonmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Aug, 17, 2017\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/code}{\faGithubAlt~Code repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/data}{\faGithubAlt~Data repository}
  \end{repobox}
      \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{How learning can guide evolution}, G. E. Hinton, and S. J. Nowlan,
Complex Systems, 1 (3), 495-502, 1987.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Lamarckian hypothesis that experiences accumulated during the lifetime
of an individual organism are directly transmitted to the next
generation is considered to be incorrect. The reference paper
\autocite{hinton1987learning} argues that this does not necessarily mean
that individual learning cannot exert influence on evolution in an
indirect fashion, providing a simple but convincing computational
example of such interaction. The authors use an extreme scenario where
optimization landscape is completely flat except for a single peak. In
such environment curvature of the optimization surface does not provide
any guidance and evolutionary search alone cannot find the maximum.
Through a simulation the authors show that such hostile environment can
be tackled by a combination of evolutionary and individual learning. If
organisms can search during their lifetime in addition, those organisms
whose genomes are closer to the targeted peak are going to be able to
find the peak with individual search. These organisms will have higher
fitness and transmit their genes to the next generation. In effect,
capacity for individual learning ``creates a hill'' to the peak that the
evolution can climb, as illustrated in Figure 1 in
\textcite{hinton1987learning}\footnote{Note that captions of the figures
  are switched, caption of Figure 1 is incorrectly placed under Figure
  2, and vice versa}. This is one of the clearest illustrations of
benefits of individual learning for evolution and interaction between
the two learning processes.\footnote{An indirect influence of individual
  learning on evolutionary learning is often called Baldwin effect.}
This result has made a substantial impact in cognitive science
literature.

The main aim of this article is to replicate the simulations reported in
the reference paper \autocite{hinton1987learning}. I have inquired with
one of the authors and the original implementation of the simulations is
not available any more. I am also not aware of other implementations
elsewhere. Simulations are relatively simple and I propose a replication
based on description from the reference paper. Replication code relies
on R programming language \autocite{R} and several R packages
\autocites{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}.

\section{Methods}\label{methods}

Description of simulations can be found in the caption of Figure 1 in
the reference paper \autocite{hinton1987learning}. I consider the
description to be detailed enough for an implementation. I have followed
it closely, with two exceptions. First, I have repeated evolution many
times to average out potential noise in the process. In the reference
article such iterations are not explicitly mentioned, and moreover, when
interpreting their results from Figure 2, the authors state that Figure
2 presents results of ``\ldots{} a typical evolutionary search
\ldots{}'', suggesting it is a single run of the simulation. Second, I
let the population of agents evolve for 1000 generations instead of 50
generations reported in \textcite{hinton1987learning}. Reason for this
will become clear in the Results section.

I describe my implementation in detail in Algorithm 1. Parameters for
the simulation are summarized in Table ~\ref{tbl:parameters}.

\begin{algorithm}
\caption{Simulation description}
    \begin{algorithmic}[1]
    \For{$sim=1:noSim$}
        \State{Generate targeted genome, $g^*=\{a_i\}_{i=1}^{noAlleles}$: $a_i \sim Bernoulli(p)$, }
        \State {$p \sim Uniform(0,1)$ }
        \State{Generate initial genomes of the agents, $G=\{g_i\}_{i=1}^{noAgents}$: $\forall g_i$ generate} 
        \State {$\{a_i\}_{i=1}^{noAlleles}$ by sampling with replacement from $\{0,1,?\}$ according to $p_0,p_1,p_?$.}
        \For{$gen=1:noGenerations$}
            \For{$ag=1:noAgents$} \Comment {\textbf{Individual learning}}
                \For{$step=1:lifetime$}
                    \State {For all ? alleles in agent's genome: $a_i \sim Bernoulli(0.5)$}
                    \If{$g_i == g^*$ after individual learning} \State {\textbf{break}} \EndIf
                \EndFor
                \State {Evaluate fitness: $f_{ag} = 1 - 19(lifetime-step)/1000$}
                \State {Record frequency of Correct, Incorrect and Undecided alleles.}
            \EndFor
            \State Compute parenting probabilities: 
            \State $p_{ag}=f_{ag}/\sum_{i=1}^{noAgents}f_{i}$, $\forall ag$
            \For{$i=1:|G|$} \Comment {\textbf{Generate children genomes}}
                \State {Choose two parents by sampling with replacement according to $p_{ag}$}
                \State {Generate new genome, $g_i^{child}$: randomly choose a cross-over point, 
                \State copy all alleles from the first parent up to the cross-over point, 
                \State and from the second parent beyond the cross-over point.}
            \EndFor
            \State {Update genomes: $g_i \gets g_i^{child}$, $\forall i \in G$}
        \EndFor
    \EndFor
    \end{algorithmic}
\end{algorithm}

\hypertarget{tbl:parameters}{}
\begin{longtable}[]{@{}lll@{}}
\caption{\label{tbl:parameters}Simulation parameters. }\tabularnewline
\toprule
Table & Value & Description\tabularnewline
\midrule
\endfirsthead
\toprule
Table & Value & Description\tabularnewline
\midrule
\endhead
noSim & 100 & Number of times evolution is simulated\tabularnewline
noGenerations & 1000 & Number of generations for evolutionary
algorithm\tabularnewline
noAgents & 1000 & Number of agents in a population\tabularnewline
lifetime & 1000 & Number of cycles available for individual
learning\tabularnewline
noAlleles & 20 & Number of alleles in agent's genome\tabularnewline
\(p_0\) & 0.25 & Probability that allele is of type zero\tabularnewline
\(p_1\) & 0.25 & Probability that allele is of type one\tabularnewline
\(p_?\) & 0.50 & Probability that allele is of type ?\tabularnewline
\bottomrule
\end{longtable}

\section{Results}\label{results}

Simulation results are summarized in Figure ~\ref{fig:relFrequencies50},
which corresponds to Figure 2 from the reference article. Qualitatively
the results are very similar - there is an obvious increase in
proportion of alleles correctly set by the evolution (close to 75\% by
generation 50) and proportion of incorrect alleles decreases to zero by
generation 50. This is a clear evidence of individual learning guiding
evolutionary search even though there is no direct transmission of
knowledge acquired during individual learning from one generation to the
next.

There is an important difference with respect to the reference paper,
however. In original simulations proportion of Undecided alleles that
are left to individual learning is relatively constant, close to initial
50\%. The authors point out this as an interesting result - there is
little selective pressure to specify last few connections by evolution
because few switches are quickly set through individual learning. In
contrast, results in Figure ~\ref{fig:relFrequencies50} show that
proportion of Undecided alleles is continuously decreasing. In fact,
whereas proportion of Correct alleles in the reference article flattens
out in my case it continues increasing, with increase coming from
proportion of Undecided alleles decreasing further. To make sure this is
really a long run trend I have let populations evolve for larger number
of generations than in the reference article. Results of all 1000
generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
generation 1000 proportion of Undecided alleles drops to 9\%. Even
though decrease is smaller with each generation, it does not seem to
stop and flatten out by the end of the simulation.

What is the source of this particular difference in results? One
explanation is that results reported in Figure 2 in the reference
article stem from a non-typical draw. This is unlikely as upon examining
evolutionary paths of Undecided alleles in all simulation runs, they all
show a strong downward trend already in first 50 generations, differing
only in the exact generation when this decrease onsets. Another
explanation is that original implementation is flawed as there actually
is a selection pressure for eliminating individual learning after some
time. The fitness function described in the reference article is such
that the smaller the number of steps an agent takes in individual
lifetime to reach the peak, the higher the probability of mating and
transmitting the genes to the next generation. Number of steps is a
direct function of number of undecided alleles in the genome. Hence,
there is a selection pressure for decreasing the number of Undecided
alleles. In personal correspondence, one of the authors agreed with this
interpretation, stating that indeed there should be a selective pressure
for decrease in number of Undecided alleles. The pressure should
decrease as a power function of the mean number of steps agents need to
reach the peak, which explains rapidly decreasing improvements with each
generation.

\begin{figure}
\centering
\includegraphics{Figure2.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele. Correct alleles are those that are set by evolutionary
learning and correspond to the targeted pattern, Incorrect denote
proportion of incorrectly set alleles, while Undecided denote proportion
of alleles left for individual learning. Proportions are means of 100
simulation runs, and barely visible grey ribbons are standard errors.
This is the reproduction of Figure 2 in the original
article.}\label{fig:relFrequencies50}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figure2_1000.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele, with a larger number of generations. Proportions are
means of 100 simulation runs, and barely visible grey ribbons are
standard errors.}\label{fig:relFrequencies1000}
\end{figure}

\section{Conclusion}\label{conclusion}

Qualitatively the result is comparable to that reported in the reference
article, but there is relatively important quantitative difference.
Proportions of Undecided alleles in implementation proposed here
diminish over time, whereas in the original implementation they stay
close to initial values. I attribute the difference to a flaw in the
original implementation, providing arguments based on fitness function
proposed in the reference article. Although the main result of the
reference article - individual learning indirectly guiding evolutionary
search - has been reproduced, the difference I have found suggests that
individual learning would be eradicated over time.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-08-17 code/genFigure2.R

@@ -1,0 +1,133 @@
+# ----------------------------------------------------------------------
+# Description
+# ----------------------------------------------------------------------
+
+# Script for Hinton, Nowlan 1987 article
+# It produces the figure 2 from the article (note that the caption of
+# figure 2 is switched with caption of figure 1)
+
+
+# ----------------------------------------------------------------------
+# Loading the data
+# ----------------------------------------------------------------------
+
+# you will need to set the directory if using the script interactively, 
+# below you see example of my path to the folder
+# setwd("/home/hstojic/Research/replications/las_HintonNowlan1987_repl/code")
+
+# load libraries and auxiliary functions
+source("utils.R")
+
+# load the dataset
+evoRes <- read.csv(file = "../data/simulation.csv", stringsAsFactors = FALSE)
+
+
+# ----------------------------------------------------------------------
+# Reshaping the data
+# ----------------------------------------------------------------------
+
+# reshaping data for plotting
+colnames(evoRes) <- c("Simulation", "Correct", "Incorrect", "Undecided")
+
+plotFrame_means <- evoRes %>% 
+    group_by(Simulation) %>% 
+    mutate(Epoch = 1:n()) %>%
+    group_by(Epoch) %>% 
+    summarize(
+        Correct = mean(Correct),
+        Incorrect = mean(Incorrect),
+        Undecided = mean(Undecided))  %>%
+    melt( 
+        id.vars = "Epoch",
+        measure.vars = c("Correct", "Incorrect", "Undecided"),
+        value.name = "proportion_mean",
+        variable.name = "alleleType")
+
+plotFrame_ses <- evoRes %>% 
+    group_by(Simulation) %>% 
+    mutate(Epoch = 1:n()) %>%
+    group_by(Epoch) %>% 
+    summarize(
+        Correct = sd(Correct)/100,
+        Incorrect = sd(Incorrect)/100,
+        Undecided = sd(Undecided)/100) %>%
+    melt( 
+        id.vars = "Epoch",
+        measure.vars = c("Correct", "Incorrect", "Undecided"),
+        value.name = "proportion_se",
+        variable.name = "alleleType") 
+
+plotFrame <- left_join(plotFrame_means, plotFrame_ses, 
+    by = c("Epoch", "alleleType")) 
+
+
+# ----------------------------------------------------------------------
+# Plotting
+# ----------------------------------------------------------------------
+
+# plot of all 1000 epochs
[…]

### 2017-08-17 code/simulation.R

@@ -1,0 +1,202 @@
+# ----------------------------------------------------------------------
+# Description
+# ----------------------------------------------------------------------
+
+# Script for running the simulation from Hinton, Nowlan 1987 article
+
+
+# ----------------------------------------------------------------------
+# Setting the parameters
+# ----------------------------------------------------------------------
+
+# you will need to set the directory if using the script interactively, 
+# below you see example of my path to the folder
+# setwd("/home/hstojic/Research/replications/las_HintonNowlan1987_repl/code")
+
+# load libraries and auxiliary functions
+source("utils.R")
+
+# we set base probabilities  
+# for elements in the genome, according to the article
+baseProbs <- c(0.25, 0.25, 0.5)  # 0, 1, NA (? from the article)
+noGenerations <- 1000
+noAgents <- 1000
+lifetime <- 1000
+noSim <- 100
+seed <- 1234
+
+# to speed up simulations, we run it in parallel 
+noCores <-  ifelse(detectCores() == 1, 1, detectCores() - 1)
+registerDoParallel(noCores)
+registerDoRNG(seed)  
+
+
+# ----------------------------------------------------------------------
+# Defining functions
+# ----------------------------------------------------------------------
+
+# function for an agent going through all the rounds in its lifetime
+# trying to learn the target
+agent <- function(genotype, target, lifetime = 1000) {
+
+    # identify how much evolution left to an agent to learn
+    learningSpace <- is.na(genotype)
+    learningGoal <- target[learningSpace]
+
+    # if evolution did not leave anything for an agent to learn
+    # we return back the original genotype and lifetime 
+    if (!any(learningSpace)) {
+        return(list(trialSolved = lifetime, 
+                    genome = genotype))
+    
+    # if evolutionary learning is incorrect, no amount of individual 
+    # learning will yield any benefit, we end the lifetime
+    } else if (!all(genotype[!learningSpace] == target[!learningSpace])) {
+        learnSolution <- learn(learningGoal,1)$solution
+        genotype[learningSpace] <- learnSolution
+        return(list(trialSolved = lifetime, 
+                    genome = genotype))
+    
+    # else agent does individual learning
+    } else {
+        learningOutcome <- learn(learningGoal, lifetime)
+        genotype[learningSpace] <- learningOutcome$solution
+        return(list(trialSolved = learningOutcome$trialSolved, 
[…]

### 2017-08-17 code/utils.R

@@ -1,0 +1,47 @@
+# ----------------------------------------------------------------------
+# Information
+# ----------------------------------------------------------------------
+
+# this script contains ggplot theme and color specifications 
+# used in plotting figures, as well as some useful functions for plotting
+
+
+# ----------------------------------------------------------------------
+# Loading packages
+# ----------------------------------------------------------------------
+
+# packages I usually use for creating figures 
+packages <- c('ggplot2', 'dplyr', 'reshape2', 
+              'doParallel', 'foreach', 'doRNG')
+lapply(packages, library, character.only = TRUE)
+
+
+# ----------------------------------------------------------------------
+# Ggplot themes
+# ----------------------------------------------------------------------
+
+# theme with font sizes adjusted for plotting figures with tikz device
+pdftheme <- 
+    theme(
+        axis.line = element_line(colour = "black"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_blank(),
+        panel.background = element_blank(),
+        panel.margin = unit(1.3, "lines"),
+        axis.ticks.x = element_line(linetype = 0),
+        axis.ticks.y = element_line(lineend=4, linetype = 1, colour = "black"),
+        axis.ticks = element_line (colour = "black", size = 0.3), 
+        axis.text = element_text(size=12, colour = "black"),
+        axis.text.x = element_text(vjust = 0.5, size=12),
+        axis.title=element_text(size=13),
+        axis.title.y = element_text(vjust = 1.8),
+        axis.title.x = element_text(vjust = -.8),
+        legend.title = element_blank(),
+        legend.justification = c(1,0.5),
+        legend.position = c(1,0.5),
+        legend.text = element_text(size=12),
+        legend.key = element_rect(fill="#FFFFFF"),
+        strip.text = element_text(size=13),
+        strip.background = element_rect(fill="#FFFFFF")
+    )

### 2017-08-20 article/stojic2017.tex
corrected grammar in the article, labels in the figures closer to lines and removed axis

@@
 
-  {\bfseries Received}  Aug, 17, 2017\\
+  {\bfseries Received}  Aug, 20, 2017\\
   {\bfseries Accepted}  Sep, 1, 2015\\
@@
 
-Lamarckian hypothesis that experiences accumulated during the lifetime
-of an individual organism are directly transmitted to the next
-generation is considered to be incorrect. The reference paper
+The Lamarckian hypothesis that adaptations accumulated during the
+lifetime of an individual organism are directly transmitted to the next
+generation is considered to be incorrect. The reference article
 \autocite{hinton1987learning} argues that this does not necessarily mean
@@
 indirect fashion, providing a simple but convincing computational
-example of such interaction. The authors use an extreme scenario where
-optimization landscape is completely flat except for a single peak. In
-such environment curvature of the optimization surface does not provide
-any guidance and evolutionary search alone cannot find the maximum.
-Through a simulation the authors show that such hostile environment can
-be tackled by a combination of evolutionary and individual learning. If
-organisms can search during their lifetime in addition, those organisms
-whose genomes are closer to the targeted peak are going to be able to
-find the peak with individual search. These organisms will have higher
-fitness and transmit their genes to the next generation. In effect,
-capacity for individual learning ``creates a hill'' to the peak that the
-evolution can climb, as illustrated in Figure 1 in
-\textcite{hinton1987learning}\footnote{Note that captions of the figures
-  are switched, caption of Figure 1 is incorrectly placed under Figure
-  2, and vice versa}. This is one of the clearest illustrations of
-benefits of individual learning for evolution and interaction between
-the two learning processes.\footnote{An indirect influence of individual
-  learning on evolutionary learning is often called Baldwin effect.}
-This result has made a substantial impact in cognitive science
-literature.
+example of such an interaction. The authors use an extreme scenario
+where the optimization landscape is completely flat except for a single
+peak. In such an environment, curvature of the optimization surface does
+not provide any guidance and evolutionary search alone cannot find the
+maximum. Through a simulation the authors show that such a hostile
+environment can be tackled by a combination of evolutionary and
+individual learning. In addition, if organisms can search during their
+lifetime, those organisms whose genomes are closer to the targeted peak
+are going to be able to find the peak with individual search. Such
+organisms will have higher fitness and transmit their genes to the next
+generation. In effect, capacity for individual learning ``creates a
+hill'' to the peak that evolutionary search can climb, as illustrated in
+Figure 1 in \textcite{hinton1987learning}\footnote{Note that captions of
+  the figures are switched, caption of Figure 1 is incorrectly placed
+  under Figure 2, and vice versa}. This is one of the clearest
+illustrations of the benefits of individual learning for evolution and
+the interaction between the two learning processes.\footnote{An indirect
+  influence of individual learning on evolutionary learning is often
+  called Baldwin effect.} This result has made a substantial impact in
+the cognitive science literature.
 
 The main aim of this article is to replicate the simulations reported in
-the reference paper \autocite{hinton1987learning}. I have inquired with
-one of the authors and the original implementation of the simulations is
-not available any more. I am also not aware of other implementations
-elsewhere. Simulations are relatively simple and I propose a replication
-based on description from the reference paper. Replication code relies
-on R programming language \autocite{R} and several R packages
-\autocites{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}.
+the reference article \autocite{hinton1987learning}. I have corresponded
+with one of the authors and learned that the original implementation of
+the simulations is no longer available. I am also not aware of other
+implementations elsewhere. The simulations are relatively simple, and I
+thus propose a replication based on description from the reference
+article. Replication code relies on R programming language \autocite{R}
+and several R packages
+\autocites{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}{ggrepel}{directlabels}.
 
@@
 
-Description of simulations can be found in the caption of Figure 1 in
-the reference paper \autocite{hinton1987learning}. I consider the
-description to be detailed enough for an implementation. I have followed
-it closely, with two exceptions. First, I have repeated evolution many
-times to average out potential noise in the process. In the reference
-article such iterations are not explicitly mentioned, and moreover, when
-interpreting their results from Figure 2, the authors state that Figure
-2 presents results of ``\ldots{} a typical evolutionary search
-\ldots{}'', suggesting it is a single run of the simulation. Second, I
-let the population of agents evolve for 1000 generations instead of 50
-generations reported in \textcite{hinton1987learning}. Reason for this
-will become clear in the Results section.
+A description of the simulations can be found in the caption of Figure 1
+in the reference article \autocite{hinton1987learning}. I consider the
+description to be detailed enough for the implementation. I have
+followed it closely, with two exceptions. First, I have repeated the
+evolutionary search many times to average out potential noise. In the
+reference article, such iterations are not explicitly mentioned, and
+moreover, when interpreting their results from Figure 2, the authors
+state that Figure 2 presents results of ``\ldots{} a typical
+evolutionary search \ldots{}'', suggesting it is a single run of the
+simulation. Second, I let the population of agents evolve for 1000
+generations instead of 50 generations reported in
+\textcite{hinton1987learning}. Reason for this will become clear in the
+Results section.
 
@@
 \endhead
-noSim & 100 & Number of times evolution is simulated\tabularnewline
+noSim & 100 & Number of times evolutionary search is
+repeated\tabularnewline
 noGenerations & 1000 & Number of generations for evolutionary
@@
 
-Simulation results are summarized in Figure ~\ref{fig:relFrequencies50},
-which corresponds to Figure 2 from the reference article. Qualitatively
-the results are very similar - there is an obvious increase in
-proportion of alleles correctly set by the evolution (close to 75\% by
-generation 50) and proportion of incorrect alleles decreases to zero by
-generation 50. This is a clear evidence of individual learning guiding
-evolutionary search even though there is no direct transmission of
-knowledge acquired during individual learning from one generation to the
-next.
-
-There is an important difference with respect to the reference paper,
-however. In original simulations proportion of Undecided alleles that
-are left to individual learning is relatively constant, close to initial
-50\%. The authors point out this as an interesting result - there is
-little selective pressure to specify last few connections by evolution
-because few switches are quickly set through individual learning. In
-contrast, results in Figure ~\ref{fig:relFrequencies50} show that
-proportion of Undecided alleles is continuously decreasing. In fact,
-whereas proportion of Correct alleles in the reference article flattens
-out in my case it continues increasing, with increase coming from
-proportion of Undecided alleles decreasing further. To make sure this is
-really a long run trend I have let populations evolve for larger number
-of generations than in the reference article. Results of all 1000
-generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
-generation 1000 proportion of Undecided alleles drops to 9\%. Even
-though decrease is smaller with each generation, it does not seem to
-stop and flatten out by the end of the simulation.
-
-What is the source of this particular difference in results? One
-explanation is that results reported in Figure 2 in the reference
-article stem from a non-typical draw. This is unlikely as upon examining
-evolutionary paths of Undecided alleles in all simulation runs, they all
-show a strong downward trend already in first 50 generations, differing
-only in the exact generation when this decrease onsets. Another
-explanation is that original implementation is flawed as there actually
-is a selection pressure for eliminating individual learning after some
-time. The fitness function described in the reference article is such
-that the smaller the number of steps an agent takes in individual
-lifetime to reach the peak, the higher the probability of mating and
-transmitting the genes to the next generation. Number of steps is a
-direct function of number of undecided alleles in the genome. Hence,
-there is a selection pressure for decreasing the number of Undecided
-alleles. In personal correspondence, one of the authors agreed with this
-interpretation, stating that indeed there should be a selective pressure
-for decrease in number of Undecided alleles. The pressure should
-decrease as a power function of the mean number of steps agents need to
-reach the peak, which explains rapidly decreasing improvements with each
-generation.
+The simulation results are summarized in Figure
+~\ref{fig:relFrequencies50}, which corresponds to Figure 2 from the
+reference article. Qualitatively, the results are very similar - there
+is an obvious increase in the proportion of alleles correctly set by the
+evolutionary learning (close to 75\% by generation 50) and proportion of
+incorrect alleles decreases to zero by generation 50. This is clear
+evidence of individual learning guiding evolutionary search, even though
+there is no direct transmission of knowledge acquired during individual
+learning from one generation to the next.
 
@@
 types of allele. Correct alleles are those that are set by evolutionary
-learning and correspond to the targeted pattern, Incorrect denote
-proportion of incorrectly set alleles, while Undecided denote proportion
-of alleles left for individual learning. Proportions are means of 100
-simulation runs, and barely visible grey ribbons are standard errors.
-This is the reproduction of Figure 2 in the original
-article.}\label{fig:relFrequencies50}
+learning and correspond to the targeted pattern. Incorrect alleles
+represent the proportion of incorrectly set alleles, while Undecided
+represent the proportion of alleles left for individual learning.
+Proportions are means of 100 simulation runs, and barely visible grey
+ribbons are standard errors. This is the reproduction of Figure 2 in the
+reference article
+\autocite{hinton1987learning}.}\label{fig:relFrequencies50}
 \end{figure}
+
+There is an important difference with respect to the reference article,
+however. In the original simulations, the proportion of Undecided
+alleles that are left to individual learning is relatively constant,
+close to initial 50\%. The authors point this out as an interesting
+result, suggesting that there is little selective pressure to specify
+last few connections by evolutionary learning because those can be
+quickly set through individual learning. In contrast, the results in
+Figure ~\ref{fig:relFrequencies50} show that the proportion of Undecided
+alleles is continuously decreasing. In fact, whereas the proportion of
+Correct alleles in the reference article flattens out, in my case it
+continues increasing, with the increase coming from the proportion of
+Undecided alleles decreasing further. To make sure that this is really a
+long run trend I have let populations evolve for a larger number of
+generations than in the reference article. Results of all 1000
+generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
+generation 1000 the proportion of Undecided alleles drops to 9\%. Even
+though the decrease is smaller with each generation, it does not seem to
+stop and flatten out by the end of the simulation.
+
+What is the source of this particular difference in results? One
+explanation is that the results reported in Figure 2 in the reference
+article stem from a non-typical draw. This is unlikely, as upon
+examining the evolutionary paths of Undecided alleles in all simulation
+runs, they all show a strong downward trend already in the first 50
+generations, differing only in the exact generation when this decrease
+onsets. Another explanation is that the original implementation is
+flawed as there actually is a selection pressure for eliminating
+individual learning after some time. The fitness function described in
+the reference article is such that the smaller the number of steps an
+agent takes in an individual lifetime to reach the peak, the higher the
+probability of mating and transmitting genes to the next generation. The
+number of steps is a direct function of the number of undecided alleles
+in the genome. Hence, there is a selection pressure for decreasing the
+number of Undecided alleles. In personal correspondence, one of the
+authors agreed with this interpretation, stating that indeed there
+should be a selective pressure for a decrease in number of Undecided
+alleles. The pressure should decrease as a power function of the mean
+number of steps an agent needs to reach the peak, which explains rapidly
+decreasing improvements with each generation.
 
@@
 
-Qualitatively the result is comparable to that reported in the reference
-article, but there is relatively important quantitative difference.
-Proportions of Undecided alleles in implementation proposed here
+Qualitatively, the result is comparable to that reported in the
+reference article, but there are relatively important differences. The
+proportion of Undecided alleles in the implementation proposed here
 diminish over time, whereas in the original implementation they stay
-close to initial values. I attribute the difference to a flaw in the
-original implementation, providing arguments based on fitness function
+close to the initial values. I attribute the difference to a flaw in the
+original implementation, providing arguments based on a fitness function
 proposed in the reference article. Although the main result of the
-reference article - individual learning indirectly guiding evolutionary
-search - has been reproduced, the difference I have found suggests that
-individual learning would be eradicated over time.
+reference article - that individual learning can indirectly guide
+evolutionary search - has been reproduced, the difference in simulation
+results suggests that individual learning might be eradicated over time.
+This decrease in usefulness of individual learning occurs in static
+environments where targeted genome stays fixed, as in these simulations.
+In more realistic scenarios, where environments are non-stationary and
+the targeted genome would be changing over time, individual learning
+would have a more stable role.
 

### 2017-08-20 code/genFigure2.R

@@ -68,7 +68,9 @@ plotFrame <- left_join(plotFrame_means, plotFrame_ses,
 # plot of all 1000 epochs
 noEpochs <- 1000
 relFrequencies1000 <- 
-    ggplot(plotFrame, 
+    ggplot(
+        mutate(plotFrame, 
+            lineLabel = ifelse(Epoch == 700, as.character(alleleType), NA)), 
         aes(x = Epoch, y = proportion_mean)) +
     geom_line(aes(group = alleleType, linetype = alleleType)) + 
     geom_ribbon(
@@ -77,11 +79,17 @@ relFrequencies1000 <-
             ymax = ifelse((proportion_mean + proportion_se) > 1, 1, 
                           proportion_mean + proportion_se), 
         group = alleleType)) + 
-    scale_x_continuous("Generations",
+    geom_label_repel( 
+        aes(x = Epoch, y = proportion_mean, label = lineLabel),
+        colour = "black", 
+        segment.color = "black", segment.size = 0.2, force = 10,
+        size = fontSize,
+        inherit.aes = FALSE) + 
+    scale_x_continuous("\nGenerations",
                      limits = c(1, noEpochs),
                      breaks = c(1, seq(0, noEpochs, 100)[2:length(seq(0, noEpochs, 100))]),
                      labels = c(1, seq(0, noEpochs, 100)[2:length(seq(0, noEpochs, 100))])) +
-    scale_y_continuous("Relative frequency of allele",
+    scale_y_continuous("Relative frequency of allele\n",
                        limits = c(0, 1),
                        breaks = seq(0, 1, 0.1)) +
     scale_linetype_manual("",
@@ -89,14 +97,16 @@ relFrequencies1000 <-
                           labels = c("Correct alleles", 
                                      "Incorrect alleles",
                                      "Undecided alleles")) +
-    pdftheme
+    pdftheme + 
+    theme(legend.position = "none")
 
 
 # plot of first 50 epochs, same as in the article
 noEpochs <- 50
 relFrequencies50 <- 
     ggplot(
-        filter(plotFrame, Epoch <= noEpochs), 
+        filter(plotFrame, Epoch <= noEpochs) %>%
+        mutate(lineLabel = ifelse(Epoch == 45, as.character(alleleType), NA)), 
         aes(x = Epoch, y = proportion_mean)) +
     geom_line(aes(group = alleleType, linetype = alleleType)) + 
     geom_ribbon(
@@ -105,11 +115,17 @@ relFrequencies50 <-
[…]

### 2017-08-20 code/utils.R

@@ -11,37 +11,47 @@
 # ----------------------------------------------------------------------
 
 # packages I usually use for creating figures 
-packages <- c('ggplot2', 'dplyr', 'reshape2', 
+packages <- c('ggplot2', 'dplyr', 'reshape2', 'ggrepel',
               'doParallel', 'foreach', 'doRNG')
 lapply(packages, library, character.only = TRUE)
 
 
 # ----------------------------------------------------------------------
-# Ggplot themes
+# Fonts and ggplot themes
 # ----------------------------------------------------------------------
+
+# font setup
+# loadfonts()
+fontSetup <- "Helvetica"
+fontSize <- 3.15 
+pointSize <- 1.5
+themeFontSize <- 12
 
 # theme with font sizes adjusted for plotting figures with tikz device
 pdftheme <- 
     theme(
-        axis.line = element_line(colour = "black"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         panel.border = element_blank(),
         panel.background = element_blank(),
-        panel.margin = unit(1.3, "lines"),
-        axis.ticks.x = element_line(linetype = 0),
-        axis.ticks.y = element_line(lineend=4, linetype = 1, colour = "black"),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        axis.ticks.x = element_line(lineend = 4, linetype = 1),
+        axis.ticks.y = element_line(lineend = 4, linetype = 1),
         axis.ticks = element_line (colour = "black", size = 0.3), 
-        axis.text = element_text(size=12, colour = "black"),
-        axis.text.x = element_text(vjust = 0.5, size=12),
-        axis.title=element_text(size=13),
+        axis.text = element_text(size = themeFontSize, colour = "black"),
+        axis.text.x = element_text(vjust = 0.5),
+        axis.title = element_text(size = themeFontSize + 1),
         axis.title.y = element_text(vjust = 1.8),
         axis.title.x = element_text(vjust = -.8),
         legend.title = element_blank(),
-        legend.justification = c(1,0.5),
-        legend.position = c(1,0.5),
-        legend.text = element_text(size=12),
-        legend.key = element_rect(fill="#FFFFFF"),
-        strip.text = element_text(size=13),
-        strip.background = element_rect(fill="#FFFFFF")
+        legend.justification = c(1,0),
+        legend.position = c(1,0),
+        legend.text = element_text(size = themeFontSize),
[…]

### 2017-09-05 article/stojic2017.tex
updated article and the code based on reviews

@@
 environment can be tackled by a combination of evolutionary and
-individual learning. In addition, if organisms can search during their
+individual learning. For organisms that can search during their
 lifetime, those organisms whose genomes are closer to the targeted peak
@@
 Figure 1 in \textcite{hinton1987learning}\footnote{Note that captions of
-  the figures are switched, caption of Figure 1 is incorrectly placed
-  under Figure 2, and vice versa}. This is one of the clearest
+  the figures are switched in the reference article, caption of Figure 1
+  is incorrectly placed under Figure 2, and vice versa.}. In contrast to
+the Lamarckian hypothesis, this is an indirect transmission of
+individual learning on evolutionary learning, what is often called the
+Baldwin effect \autocite{Baldwin1896}. This is one of the clearest
 illustrations of the benefits of individual learning for evolution and
-the interaction between the two learning processes.\footnote{An indirect
-  influence of individual learning on evolutionary learning is often
-  called Baldwin effect.} This result has made a substantial impact in
-the cognitive science literature.
+the interaction between the two learning processes. This article is well
+regarded and widely cited in cognitive science literature.
 
-The main aim of this article is to replicate the simulations reported in
-the reference article \autocite{hinton1987learning}. I have corresponded
-with one of the authors and learned that the original implementation of
-the simulations is no longer available. I am also not aware of other
-implementations elsewhere. The simulations are relatively simple, and I
-thus propose a replication based on description from the reference
-article. Replication code relies on R programming language \autocite{R}
-and several R packages
-\autocites{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}{ggrepel}{directlabels}.
+I replicated the simulations reported in the reference article
+\autocite{hinton1987learning} in R \autocite{R}, using several packages
+\autocites{devtools}{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}{ggrepel}.
+In correspondence with Geoffrey Hinton I learned that the original
+implementation is not available.
 
@@
 
-I describe my implementation in detail in Algorithm 1. Parameters for
+Pseudocode of my implementation is shown in Algorithm 1. Parameters for
 the simulation are summarized in Table ~\ref{tbl:parameters}.
@@
         \State {$p \sim Uniform(0,1)$ }
-        \State{Generate initial genomes of the agents, $G=\{g_i\}_{i=1}^{noAgents}$: $\forall g_i$ generate} 
+        \State{Generate initial genomes of the agents, $G=\{g_i\}_{i=1}^{noAgents}$: for all $g_i$ generate} 
         \State {$\{a_i\}_{i=1}^{noAlleles}$ by sampling with replacement from $\{0,1,?\}$ according to $p_0,p_1,p_?$.}
@@
                 \EndFor
-                \State {Evaluate fitness: $f_{ag} = 1 - 19(lifetime-step)/1000$}
+                \State {Evaluate fitness: $f_{ag} = 1 + 19(lifetime-step)/1000$}
                 \State {Record frequency of Correct, Incorrect and Undecided alleles.}
@@
             \State Compute parenting probabilities: 
-            \State $p_{ag}=f_{ag}/\sum_{i=1}^{noAgents}f_{i}$, $\forall ag$
-            \For{$i=1:|G|$} \Comment {\textbf{Generate children genomes}}
+            \State $p_{ag}=f_{ag}/\sum_{i=1}^{noAgents}f_{i}$, for every agent $ag$
+            \For{$i=1:noAgents$} \Comment {\textbf{Generate children genomes}}
                 \State {Choose two parents by sampling with replacement according to $p_{ag}$}
@@
             \EndFor
-            \State {Update genomes: $g_i \gets g_i^{child}$, $\forall i \in G$}
+            \State {Update genomes: $g_i \gets g_i^{child}$, for every genome in $G$}
         \EndFor
@@
 noAlleles & 20 & Number of alleles in agent's genome\tabularnewline
-\(p_0\) & 0.25 & Probability that allele is of type zero\tabularnewline
-\(p_1\) & 0.25 & Probability that allele is of type one\tabularnewline
-\(p_?\) & 0.50 & Probability that allele is of type ?\tabularnewline
+\(p_0\) & 0.25 & Probability that allele's value is 0\tabularnewline
+\(p_1\) & 0.25 & Probability that allele's value is 1\tabularnewline
+\(p_?\) & 0.50 & Probability that allele's value is ?
+(learnable)\tabularnewline
 \bottomrule
@@
 represent the proportion of alleles left for individual learning.
-Proportions are means of 100 simulation runs, and barely visible grey
-ribbons are standard errors. This is the reproduction of Figure 2 in the
-reference article
+Proportions are means of 100 simulation runs, and grey shaded area
+around the mean line represent standard errors of the means. This is the
+reproduction of Figure 2 in the reference article
 \autocite{hinton1987learning}.}\label{fig:relFrequencies50}
@@
 generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
-generation 1000 the proportion of Undecided alleles drops to 9\%. Even
-though the decrease is smaller with each generation, it does not seem to
-stop and flatten out by the end of the simulation.
-
-What is the source of this particular difference in results? One
-explanation is that the results reported in Figure 2 in the reference
-article stem from a non-typical draw. This is unlikely, as upon
-examining the evolutionary paths of Undecided alleles in all simulation
-runs, they all show a strong downward trend already in the first 50
-generations, differing only in the exact generation when this decrease
-onsets. Another explanation is that the original implementation is
-flawed as there actually is a selection pressure for eliminating
-individual learning after some time. The fitness function described in
-the reference article is such that the smaller the number of steps an
-agent takes in an individual lifetime to reach the peak, the higher the
-probability of mating and transmitting genes to the next generation. The
-number of steps is a direct function of the number of undecided alleles
-in the genome. Hence, there is a selection pressure for decreasing the
-number of Undecided alleles. In personal correspondence, one of the
-authors agreed with this interpretation, stating that indeed there
-should be a selective pressure for a decrease in number of Undecided
-alleles. The pressure should decrease as a power function of the mean
-number of steps an agent needs to reach the peak, which explains rapidly
-decreasing improvements with each generation.
+generation 1000 the mean proportion of Undecided alleles drops to 9\%,
+with the decrease being smaller with each generation.
 
@@
 types of allele, with a larger number of generations. Proportions are
-means of 100 simulation runs, and barely visible grey ribbons are
-standard errors.}\label{fig:relFrequencies1000}
+means of 100 simulation runs, and grey shaded area around the mean line
+represent standard errors of the means.}\label{fig:relFrequencies1000}
 \end{figure}
+
+What is the source of this particular difference in results? The results
+reported in Figure 2 in the reference article might stem from a
+non-typical draw. This is unlikely, as illustrated in Figure
+~\ref{fig:Figure_relFrequencies_allSims}. Evolutionary paths of
+Undecided alleles in all simulation runs show a strong downward trend
+already in the first 50 generations, differing only in the exact
+generation when this decrease onsets. As shown in Figure
+~\ref{fig:Figure_endPointHistogram}, distribution of proportions of
+Undecided alleles at the end of the simulation (1000 generations) ranges
+from 0.05 to 0.20, much lower than 0.46 that was obtained in the
+reference article.
+
+\begin{figure}
+\centering
+\includegraphics{Figure_relFrequencies_allSims.pdf}
+\caption{The evolution of relative frequencies of Undecided alleles for
+all 100 simulations, red line marks the mean across all simulations,
+same as in Figure ~\ref{fig:relFrequencies1000}, while blue line marks
+approximately the end point of relative frequency from the reference
+article.}\label{fig:Figure_relFrequencies_allSims}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Figure_endPointHistogram.pdf}
+\caption{Histogram of relative frequencies of Undecided alleles at the
+end of simulations (1000
+generations).}\label{fig:Figure_endPointHistogram}
+\end{figure}
+
+The decrease in Undecided alleles is actually expected, once the fitness
+function described in the reference article is examined more closely.
+The function is such that the smaller the number of steps an agent takes
+in an individual lifetime to reach the peak, the higher the probability
+of mating and transmitting genes to the next generation. The number of
+steps is a direct function of the number of undecided alleles in the
+genome. In other words, there exists a selection pressure for decreasing
+the number of Undecided alleles. In personal correspondence, Geoffrey
+Hinton agreed with this argument, stating that the selective pressure
+should decrease as a power function of the mean number of steps an agent
+needs to reach the peak. Computational power at the time was orders of
+magnitude smaller than today, preventing them to explore the space more
+thoroughly.
 
@@
 diminish over time, whereas in the original implementation they stay
-close to the initial values. I attribute the difference to a flaw in the
-original implementation, providing arguments based on a fitness function
-proposed in the reference article. Although the main result of the
-reference article - that individual learning can indirectly guide
-evolutionary search - has been reproduced, the difference in simulation
-results suggests that individual learning might be eradicated over time.
-This decrease in usefulness of individual learning occurs in static
-environments where targeted genome stays fixed, as in these simulations.
-In more realistic scenarios, where environments are non-stationary and
-the targeted genome would be changing over time, individual learning
-would have a more stable role.
+close to the initial values. I have provided an argument based on a
+fitness function proposed in the reference article that clearly explains
+the decrease. Although the main result of the reference article - that
+individual learning can indirectly guide evolutionary search - has been
+reproduced, the difference in simulation results suggests that
+individual learning might be eradicated over time. This decrease in
+usefulness of individual learning occurs in static environments where
+targeted genome stays fixed, as in these simulations. In more realistic
+scenarios, where environments are non-stationary and the targeted genome
+would be changing over time, individual learning would have a more
+stable role.
 

### 2017-09-05 code/genFigure2.R

@@ -15,6 +15,9 @@
 # below you see example of my path to the folder
 # setwd("/home/hstojic/Research/replications/las_HintonNowlan1987_repl/code")
 
+# house keeping
+rm(list = ls())
+
 # load libraries and auxiliary functions
 source("utils.R")
 
@@ -26,9 +29,13 @@ evoRes <- read.csv(file = "../data/simulation.csv", stringsAsFactors = FALSE)
 # Reshaping the data
 # ----------------------------------------------------------------------
 
-# reshaping data for plotting
+# adding nice names for the data
 colnames(evoRes) <- c("Simulation", "Correct", "Incorrect", "Undecided")
 
+# detect the number of simulations in the data
+noSim <- max(evoRes$Simulation)
+
+# computing means
 plotFrame_means <- evoRes %>% 
     group_by(Simulation) %>% 
     mutate(Epoch = 1:n()) %>%
@@ -43,14 +50,15 @@ plotFrame_means <- evoRes %>%
         value.name = "proportion_mean",
         variable.name = "alleleType")
 
+# computing standard errors
 plotFrame_ses <- evoRes %>% 
     group_by(Simulation) %>% 
     mutate(Epoch = 1:n()) %>%
     group_by(Epoch) %>% 
     summarize(
-        Correct = sd(Correct)/100,
-        Incorrect = sd(Incorrect)/100,
-        Undecided = sd(Undecided)/100) %>%
+        Correct = sd(Correct)/sqrt(noSim),
+        Incorrect = sd(Incorrect)/sqrt(noSim),
+        Undecided = sd(Undecided)/sqrt(noSim)) %>%
     melt( 
         id.vars = "Epoch",
         measure.vars = c("Correct", "Incorrect", "Undecided"),
@@ -70,19 +78,15 @@ noEpochs <- 1000
 relFrequencies1000 <- 
     ggplot(
         mutate(plotFrame, 
-            lineLabel = ifelse(Epoch == 700, as.character(alleleType), NA)), 
+            lineLabel = ifelse(Epoch == 0.8*noEpochs, 
+                as.character(alleleType), NA)), 
         aes(x = Epoch, y = proportion_mean)) +
     geom_line(aes(group = alleleType, linetype = alleleType)) + 
-    geom_ribbon(
-        aes(ymin = ifelse((proportion_mean - proportion_se) < 0, 0,
-                          proportion_mean - proportion_se), 
[…]

### 2017-09-05 code/genFigureAux.R

@@ -1,0 +1,145 @@
+# ----------------------------------------------------------------------
+# Description
+# ----------------------------------------------------------------------
+
+# Script for Hinton, Nowlan 1987 article
+# It produces additional figure with path for undecided allele for each 
+# simulation run, and histogram of end points, with original results
+# from the reference paper overlaid 
+
+
+# ----------------------------------------------------------------------
+# Loading the data
+# ----------------------------------------------------------------------
+
+# you will need to set the directory if using the script interactively, 
+# below you see example of my path to the folder
+# setwd("/home/hstojic/Research/replications/las_HintonNowlan1987_repl/code")
+
+# house keeping
+rm(list = ls())
+
+# load libraries and auxiliary functions
+source("utils.R")
+
+# load the dataset
+evoRes <- read.csv(file = "../data/simulation.csv", stringsAsFactors = FALSE)
+
+
+# ----------------------------------------------------------------------
+# Reshaping the data
+# ----------------------------------------------------------------------
+
+# reshaping data for plotting
+colnames(evoRes) <- c("Simulation", "Correct", "Incorrect", "Undecided")
+
+# number of epochs we plot
+noEpochs <- 1000
+
+# data for evolution paths of all simulations for undecided alleles
+plotFrame <- evoRes %>% 
+    group_by(Simulation) %>% 
+    mutate(Epoch = 1:n()) %>%
+    mutate(lineLabel = ifelse(Epoch == 600 & Simulation == 2, "Single simulation", NA) )
+
+plotFrame_mean <- plotFrame %>% 
+    group_by(Epoch) %>% 
+    summarize(Undecided = mean(Undecided)) %>%
+    mutate(lineLabel = ifelse(Epoch == 400, "Mean", NA) )
+
+# data for the histogram of the end points
+plotHist <- filter(plotFrame, Epoch == max(Epoch))
+
+# the endpoint of the original Hinton & Nowlan allele frequencies 
+# aaproximate, from figure 2
+undecidedRefArticle <- 0.46
+refArticle <- 
+    data.frame(Epoch = 1:noEpochs, Undecided = undecidedRefArticle) %>%
+    mutate(lineLabel = ifelse(Epoch == 800, "Hinton & nowlan (1987)", NA) )
+
+
+# ----------------------------------------------------------------------
+# Plotting
[…]

### 2017-09-05 code/simulation.R

@@ -13,11 +13,15 @@
 # below you see example of my path to the folder
 # setwd("/home/hstojic/Research/replications/las_HintonNowlan1987_repl/code")
 
+# house keeping
+rm(list = ls())
+
 # load libraries and auxiliary functions
 source("utils.R")
 
 # we set base probabilities  
 # for elements in the genome, according to the article
+noAlleles <- 20
 baseProbs <- c(0.25, 0.25, 0.5)  # 0, 1, NA (? from the article)
 noGenerations <- 1000
 noAgents <- 1000
@@ -36,57 +40,61 @@ registerDoRNG(seed)
 # ----------------------------------------------------------------------
 
 # function for an agent going through all the rounds in its lifetime
-# trying to learn the target
-agent <- function(genotype, target, lifetime = 1000) {
+# trying to learn the targeted genome, gstar
+# it is used in individual learning part of the loop from Algorithm 1
+agentLearning <- function(g, gstar, lifetime = 1000) {
 
-    # identify how much evolution left to an agent to learn
-    learningSpace <- is.na(genotype)
-    learningGoal <- target[learningSpace]
+    # identify part of the genome, g, that evolution left to an agent to learn
+    glearnID <- is.na(g)
+    gstarL <- gstar[glearnID]
 
     # if evolution did not leave anything for an agent to learn
     # we return back the original genotype and lifetime 
-    if (!any(learningSpace)) {
-        return(list(trialSolved = lifetime, 
-                    genome = genotype))
+    if (!any(glearnID)) {
+        return(list(stepSolved = lifetime, genome = g))
     
     # if evolutionary learning is incorrect, no amount of individual 
     # learning will yield any benefit, we end the lifetime
-    } else if (!all(genotype[!learningSpace] == target[!learningSpace])) {
-        learnSolution <- learn(learningGoal,1)$solution
-        genotype[learningSpace] <- learnSolution
-        return(list(trialSolved = lifetime, 
-                    genome = genotype))
+    } else if (!all(g[!glearnID] == gstar[!glearnID])) {
+        learnSolution <- learning(gstarL,1)$solution
+        g[glearnID] <- learnSolution
+        return(list(stepSolved = lifetime, genome = g))
     
     # else agent does individual learning
     } else {
-        learningOutcome <- learn(learningGoal, lifetime)
-        genotype[learningSpace] <- learningOutcome$solution
[…]

### 2017-09-05 code/utils.R

@@ -10,10 +10,22 @@
 # Loading packages
 # ----------------------------------------------------------------------
 
-# packages I usually use for creating figures 
-packages <- c('ggplot2', 'dplyr', 'reshape2', 'ggrepel',
-              'doParallel', 'foreach', 'doRNG')
-lapply(packages, library, character.only = TRUE)
+# installing the packages, attempting to install the exact version 
+if ("devtools" %in% rownames(installed.packages()) == FALSE) {
+    install.packages("devtools")
+    library(devtools)
+}
+
+packages <- c('ggplot2' = "2.2.1", 'dplyr' = "0.5.0", 'reshape2' = "1.4.2", 
+              'ggrepel' = "0.6.5", 'doParallel' = "1.0.10", 
+              'foreach' = "1.4.3", 'doRNG' = "1.6.6")
+for (p in names(packages)) {
+    install_version(p, version = packages[[p]], 
+        repos = "http://cran.us.r-project.org")
+}
+
+# loading the packages 
+lapply(names(packages), library, character.only = TRUE)
 
 
 # ----------------------------------------------------------------------
@@ -21,7 +33,6 @@ lapply(packages, library, character.only = TRUE)
 # ----------------------------------------------------------------------
 
 # font setup
-# loadfonts()
 fontSetup <- "Helvetica"
 fontSize <- 3.15 
 pointSize <- 1.5

### 2017-10-04 article/detorakis-2017.tex
markdown file fixed

--- 2017-08-02 article/detorakis-2017.tex
@@
 \usepackage{hyperref}
+\usepackage{amssymb,amsmath,array,amsthm}
 \hypersetup{breaklinks=true,
@@
 [Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
-        Spiking Behavior
+Spiking Behaviors
 },
@@
 \lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}
-
-\usepackage{float,lscape}
-\usepackage{xcolor,colortbl}
-\usepackage{amssymb,amsmath,array,amsthm}
-\usepackage{marvosym}
-% \usepackage{makecell}
-\usepackage{boldline}
-\usepackage{bm,array}
-
-\usepackage{adjustbox}
-
-\newcommand{\Rm}[1]{\mathrm{#1}}
-
-\newcolumntype{C}{>{\centering\arraybackslash}p{3.5em}}
-
-\makeatletter
-\newcommand{\thickhline}{%
-    \noalign {\ifnum 0=`}\fi \hrule height 1pt
-    \futurelet \reserved@a \@xhline
-}
-\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
-
-\definecolor{Gray}{gray}{0.75}
-\definecolor{LightGray}{gray}{0.95}
 
@@
 % --- Document ----------------------------------------------------------------
-\title{[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces
-            Diverse Spiking Behaviors}
+\title{[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
+Spiking Behaviors}
 
@@
                         \author[1]{Georgios Detorakis}
-                            \affil[1]{Department of Cognitive Sciences, UC Irvine,
-                            Irvine, CA, USA}
+                            \affil[1]{Department of Cognitive Sciences, UC Irvine, Irvine CA, USA}
 
 \date{\vspace{-5mm}
-      \sffamily \small
-      \href{mailto:corresponding-author@mail.com}{gdetorak@uci.edu \\
-      gdetor@protonmail.com}}
+      \sffamily \small \href{mailto:gdetorak@uci.edu (gdetor@protonmail.com)}{gdetorak@uci.edu (gdetor@protonmail.com)}}
 
@@
   {\bfseries Competing Interests:}\\
-  The author has declared that no competing interests exist.
+  The authors have declared that no competing interests exist.
   \end{flushleft}
@@
 \begin{itemize}
-    \item[→] A Generalized Linear Integrate-and-Fire Neural Model Produces
-        Diverse Spiking Behaviors, Stefan Mihalas and Ernst Niebur, Neural
-        Computation 21, 704--718, 2009.
+    \item[→] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
+Spiking Behaviors, Stefan Mihalas and Ernst Niebur, Neural Computation
+21, 704--718, 2009.
   \end{itemize}\par
@@
 Integrate-and-fire neurons are being used extensively in the field of
-neuroscience for modeling spiking behaviors~\cite{dayan:2001}. In this work we
-provide a reference implementation of~\cite{mihalas:2009}, where the authors
-have introduced a generalization of the leaky integrate-and-fire neuron model. 
-The Mihalas-Niebur Neuron (MNN) model is a linear integrate-and-fire neuron
+neuroscience for modeling spiking behaviors~\autocite{dayan:2001}. In
+this work we provide a reference implementation
+of~\autocite{mihalas:2009}, where the authors have introduced a
+generalization of the leaky integrate-and-fire neuron model. The
+Mihalas-Niebur Neuron (MNN) model is a linear integrate-and-fire neuron
 model capable of expressing a rich spiking behavior based on a set of
@@
 
-An MNN model expresses tonic and phasic spiking, class $1$ and $2$, spike
-frequency adaptation, accommodation, threshold variability, rebound spike,
-integrator, input bistability, hyperpolarizing spiking and bursting, tonic,
-phasic and rebound bursting, mixed mode, afterpotentials, basal bistability, 
-preferred frequency and spike latency. 
-Due to its simplicity, the MNN model has been used in neuromorphic 
-implementations such as~\cite{folowosele:2011}.
+An MNN model expresses tonic and phasic spiking, class \(1\) and \(2\),
+spike frequency adaptation, accommodation, threshold variability,
+rebound spike, integrator, input bistability, hyperpolarizing spiking
+and bursting, tonic, phasic and rebound bursting, mixed mode,
+afterpotentials, basal bistability, preferred frequency and spike
+latency. Due to its simplicity, the MNN model has been used in
+neuromorphic implementations such as~\autocite{folowosele:2011}.
 
-The model consists of linear differential equations, which describe the membrane
-and threshold potentials and internal currents. All the results provided in 
-\cite{mihalas:2009} have been obtained by using only two internal currents and
-thus we use the exact same number of internal currents in this work. The
-differential equations describe the subthreshold dynamics, while the 
-instantaneous threshold potential controls when the neuron firing an action 
-potential (spike) in a dynamic way. The ability of the MNN model to generate
-such a diverse spiking behavior is due to the complex update rules. In this 
-work the MNN model has been implemented in Python (version 3.6.1) using 
-Numpy (version 1.13.1) and Matplotlib (version 2.0.2) packages.
+The model consists of linear differential equations, which describe the
+membrane and threshold potentials and internal currents. All the results
+provided in \autocite{mihalas:2009} have been obtained by using only two
+internal currents and thus we use the exact same number of internal
+currents in this work. The subthreshold dynamics are defined by a set of
+linear ordinary differential equations, while an instantaneous threshold
+potential controls when the neuron fires an action potential (spike) in
+a dynamic way. The ability of the MNN model to generate such a diverse
+spiking behavior is due to the complex update rules. In this work the
+MNN model has been implemented in Python (version 3.6.1) using Numpy
+(version 1.13.1) and Matplotlib (version 2.0.2) packages.
 
@@
 
-In order to implement the model described in \cite{mihalas:2009}, we discretized
-the dynamical system using the Euler's integration scheme. The time step is 
-fixed to $0.1\, \Rm{ms}$ for all the simulations, and the simulation time
-varies according to figure 1 of the original paper. We provide all equations
-and parameters of the model in tables as it has been suggested
-by~\cite{nordlie:2009}. 
-Table~\ref{table:1} provides the summary of the model. Tables~\ref{table:2}
-and~\ref{table:3} give the subthreshold dynamics (differential equations) 
-describing the membrane and the threshold potentials as well as the two internal
-currents and the update rules. The parameters for all the simulations are given
-in table~\ref{table:4}. Table~\ref{table:5} provides the external current
-intensities (and pulse durations), and initial conditions are described
-in table~\ref{table:6}.
+In order to implement the model described in \autocite{mihalas:2009}, we
+discretized the dynamical system using the forward Euler integration
+scheme. The time step is fixed to \(0.1\, \mathrm{ms}\) for all the
+simulations, and the total simulation time \(t_f\) varies according to
+figure 1 of the original paper. Our implementation differs from the one
+in the original paper, since in \autocite{mihalas:2009}, authors
+numerically solve equation \(3.5\) (algebraic equation) under the
+constraint imposed by inequality \(3.4\) and thus they compute the spike
+times. On the other hand, in this work we directly compute numerically
+the solution of the dynamical system defined by equations \(2.1\) and
+\(2.2\) in \autocite{mihalas:2009} (see tables~~\ref{tbl:2}
+and~~\ref{tbl:3}).
 
-%%
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{ll}
-        \thickhline
-        \multicolumn{2}{c}{Model Summary} \\\thickhline
-        \rowcolor{Gray}
-        Populations  & No population -- single neuron model \\\rowcolor{LightGray}
-        Topology     & -- \\ \rowcolor{Gray}
-        Connectivity & -- \\ \rowcolor{LightGray}
-        Neuron Model & Linear Integrate-and-Fire Neuron \\\rowcolor{Gray}
-        Channel Models & Linear, first order ODEs  \\ \rowcolor{LightGray}
-        Synapse Model & -- \\ \rowcolor{Gray}
-        Plasticity & -- \\ \rowcolor{LightGray}
-        Input & Constant current or rectangular pulses \\\rowcolor{Gray}
-        Measurements & Membrane potential, phase plane \\
-        \thickhline
-    \end{tabular}
-    \caption{{\bfseries \sffamily Summary of the model}} 
-    \label{table:1}
-\end{table}
-%%
+We provide all equations and parameters of the model in tables as it has
+been suggested by~\autocite{nordlie:2009}. Table~~\ref{tbl:1} provides
+the summary of the model. Tables~~\ref{tbl:2} and~~\ref{tbl:3} give the
+subthreshold dynamics (differential equations) describing the membrane
+and the threshold potentials as well as the two internal currents and
+the update rules. The parameters for all the simulations are given in
+table~~\ref{tbl:4}, while the external current intensities and pulse
+duration are provided in table~~\ref{tbl:5}. The parameters in this work
+are exactly the same used in the original paper (table 1, pg. \(711\)).
+We had to infer the time intervals and the total simulation times for
+the pulses since they are not given explicitly in the original paper.
+Thus, we extracted the time intervals from figure \(1\) of
+\autocite{mihalas:2009} by visual inspection. The initial conditions are
+given in table~~\ref{tbl:6}.
 
-%%
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{p{3.5cm}ll}
-        \thickhline
-        \multicolumn{2}{c}{Neuron Model} \\\thickhline
-        \rowcolor{Gray}
-        Name  &  Mihalas-Niebur Neuron (MNN) \\ \rowcolor{LightGray}
-        Type  &  Linear Leaky Integrate-and-Fire Neuron  \\ \rowcolor{Gray}
-        Membrane Potential & $
-            \begin{aligned}
-                \frac{\Rm{d}V(t)}{\Rm{d}t} &= \frac{1}{C} \Big(I_e + I_1 + I_2 - G(V(t) -
-                E_L)  \Big)
-            \end{aligned}$ \\ \rowcolor{LightGray}
-        Instantaneous Threshold Potential  & $
-            \begin{aligned}
-                \frac{\Rm{d}\Theta(t)}{\Rm{d}t} &= a(V(t) - E_L) - b(\Theta(t) -
-                \Theta_{\infty}) 
-            \end{aligned} $
-            \\ \rowcolor{Gray}
-        Internal Currents & $
-            \begin{aligned}
-                \frac{\Rm{d}I_{1}(t)}{\Rm{d}t} &= -k_1I_1(t)\\
-                \frac{\Rm{d}I_{2}(t)}{\Rm{d}t} &= -k_2I_2(t) \\
-            \end{aligned} $
-    \end{tabular}
-    \caption{{\bfseries \sffamily Description of the subthreshold dynamics of
-    Mihalas-Niebur neuron model.} $V(t)$ and $\Theta(t)$ are the membrane and 
-    threshold potentials, respectively. $E_L$ and $\Theta_{\infty}$ are the
-    reversal potentials for the membrane and the threshold variables,
-    respectively. $a, b, k_1, k_2$ and $G$ are constant parameters. $I_e$ is
-    the external current applied on the neuron model.} 
-    \label{table:2}
-\end{table}
-%%
+All simulations ran on a Dell OptiPlex \(7040\), equipped with a sixth
+generation i\(7\) processor, \(16\, \mathrm{GB}\) of physical memory and
+running Arch Linux (x\(86\_64\)). The total execution time of all
+simulations was \(2.41\) seconds and the peak consumed memory was
+\(162\, {\mathrm{MB}}\)\footnote{Python memory profiler used
+  (\url{https://pypi.python.org/pypi/memory_profiler}).}.
 
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{cc}
-        \thickhline
-        \multicolumn{2}{c}{Update rules} \\ \thickhline
-        Variable & Rule  \\ \rowcolor{Gray}
-        $V(t)$ & $V_r$  \\ \rowcolor{LightGray}
-        $\Theta(t)$ & $\max\{\Theta_r, \Theta(t) \}$ \\ \rowcolor{Gray}
-        $I_1(t)$ & $R_1 \times I_1(t) + A_1$ \\ \rowcolor{LightGray} 
-        $I_2(t) $ & $R_2 \times I_2(t) + A_2$ \\\thickhline
-    \end{tabular}
-    \caption{{\bfseries \sffamily Update rules.} $V_r$ and $\Theta_r$
+\hypertarget{tbl:1}{}
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tbl:1}\textbf{Summary of the model.} }\tabularnewline
+\toprule
+Model Summary &\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Model Summary &\tabularnewline
+\midrule
+\endhead
+Populations & No population -- single neuron model\tabularnewline
+Topology & --\tabularnewline
+Connectivity & --\tabularnewline
+Neuron Model & Linear Integrate-and-Fire Neuron\tabularnewline
+Channel Models & Linear, first order ODEs\tabularnewline
+Synapse Model & --\tabularnewline
+Plasticity & --\tabularnewline
+Input & Constant current/rectangular pulses\tabularnewline
+Measurements & Membrane potential, phase plane\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{tbl:2}{}
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tbl:2}Description of the subthreshold dynamics of
+Mihalas--Niebur neuron model. \(V(t)\) and \(\Theta(t)\) are the
+membrane and threshold potentials, respectively. \(E_L\) and
+\(\Theta_{\infty}\) are the reversal potentials for the membrane and
+threshold variables, respectively. \(a, b, k_1, k_2\) and \(G\) are
+constant parameters. \(I_e\) is the external current applied on the
+neuron model. }\tabularnewline
+\toprule
+Neuron Model &\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Neuron Model &\tabularnewline
+\midrule
+\endhead
+Name & Mihalas-Niebur Neuron (MNN)\tabularnewline
+Type & Linear Leaky Integrate-and-Fire Neuron\tabularnewline
+Membrane Potential &
+\( \frac{\mathrm{d}V(t)}{\mathrm{d}t} = \frac{1}{C} \Big(I_e + I_1 + I_2 - G(V(t) - E_L)  \Big) \)\tabularnewline
+Instantaneous Threshold Potential &
+\( \frac{\mathrm{d}\Theta(t)}{\mathrm{d}t} = a(V(t) - E_L) - b(\Theta(t) - \Theta_{\infty}) \)\tabularnewline
+Internal Currents &
+\( \frac{\mathrm{d}I_{1}(t)}{\mathrm{d}t} = -k_1I_1(t) \)\tabularnewline
+&
+\( \frac{\mathrm{d}I_{2}(t)}{\mathrm{d}t} = -k_2I_2(t) \)\tabularnewline
+\bottomrule
+\end{longtable}
+
+\pagebreak
+
+\hypertarget{tbl:3}{}
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tbl:3}\textbf{Update rules.} \(V_r\) and \(\Theta_r\)
 are the reset values for the membrane and threshold potentials,
-    respectively. $R_1, R_2, A_1$ and $A_2$ are constants.}
-    \label{table:3}
-\end{table}
-%%
+respectively. \(R_1, R_2, A_1\) and \(A_2\) are constants.
+}\tabularnewline
+\toprule
+Variable & Rule\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Variable & Rule\tabularnewline
+\midrule
+\endhead
+\(V(t)\) & \(V_r\)\tabularnewline
+\(\Theta(t)\) & \(\max\{\Theta_r, \Theta(t) \}\)\tabularnewline
+\(I_1(t)\) & \(R_1 \times I_1(t) + A_1\)\tabularnewline
+\(I_2(t)\) & \(R_2 \times I_2(t) + A_2\)\tabularnewline
+\bottomrule
+\end{longtable}
 
-%%
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{CCCCC}
-        \thickhline
-        \multicolumn{5}{c}{Model Parameters} \\\thickhline
-           \begin{tabular}[x]{@{}c@{}} Figure \\  \end{tabular}
-         & \begin{tabular}[x]{@{}c@{}} $a$ \\ $(\Rm{s^{-1}})$ \end{tabular}
-         & \begin{tabular}[x]{@{}c@{}} $A_1/C$ \\ $(\Rm{V/s})$ \end{tabular}
-         & \begin{tabular}[x]{@{}c@{}} $A_2/C$ \\ $(\Rm{V/s})$ \end{tabular}& 
-           \begin{tabular}[x]{@{}c@{}} $t_f$ \\ $(\Rm{s})$ \end{tabular} \\
-        \thickhline
-        1A & $0$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
-        1B & $0$ & $0$ & $0$ & $0.5$  \\\rowcolor{LightGray}
-        1C & $5$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
-        1D & $5$ & $0$ & $0$ & $0.5$ \\\rowcolor{LightGray}
-        1E & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{Gray}
-        1F & $5$ & $0$ & $0$ & $0.4$ \\\rowcolor{LightGray}
-        1G & $5$ & $0$ & $0$ & $1.0$  \\\rowcolor{Gray}
-        1H &  $5$ & $0$ & $0$ & $0.3$ \\\rowcolor{LightGray}
-        1I & $5$  & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
-        1J & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{LightGray}
-        1K & $30$ & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
-        1L & $30$  & $10$ & $-0.6$ & $0.4$ \\\rowcolor{LightGray}
-        1M & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{Gray}
-        1N & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{LightGray}
-        1O & $5$ & $10$ & $-0.6$ & $1.0$ \\\rowcolor{Gray}
-        1P & $5$  & $5$ & $-0.3$ & $0.5$ \\\rowcolor{LightGray}
-        1Q & $5$  & $5$ & $-0.3$ & $0.2$ \\\rowcolor{Gray}
-        1R & $0$ & $8$ & $-0.1$ & $0.2$ \\\rowcolor{LightGray}
-        1S & $5$ & $-3$ & $0.5$ & $0.8$ \\\rowcolor{Gray}
-        1T &  $-80$ & $0$ & $0$ & $0.05$ \\
-        \thickhline
-        \multicolumn{5}{l}{Common Parameters} \\\rowcolor{LightGray}
-        \thickhline
-        \multicolumn{5}{p{22.5em}}{
-            $b = 10\, \Rm{s^{-1}}$, $G/C = 50\, \Rm{s^{-1}}$,
-            $k_1 = 200\, \Rm{s^{-1}}$, $k_2 = 20\, \Rm{s^{-1}}$,
-            $\Theta_{\infty} = -0.05\, \Rm{V}$,
-            $R_1 = 0$, $R_2 = 1$, $E_l = -0.07\, \Rm{V}$,
-            $V_r = -0.07\, \Rm{V}$, $\Theta_r = -0.06\, \Rm{V}$.
-        }
-        \\\thickhline
-    \end{tabular}
-    \caption{{\bfseries \sffamily Simulation Parameters}  }
-    \label{table:4}
-\end{table}
-%%
+\hypertarget{tbl:4}{}
+\begin{longtable}[]{@{}lllll@{}}
+\caption{\label{tbl:4}\textbf{Simulation Parameters}. Common parameters
+for all simulations: \(b = 10\, \mathrm{s^{-1}}\),
+\(G/C = 50\, \mathrm{s^{-1}}\), \(k_1 = 200\, \mathrm{s^{-1}}\),
+\(k_2 = 20\, \mathrm{s^{-1}}\),
+\(\Theta_{\infty} = -0.05\, \mathrm{V}\), \(R_1 = 0\), \(R_2 = 1\),
+\(E_l = -0.07\, \mathrm{V}\), \(V_r = -0.07\, \mathrm{V}\),
+\(\Theta_r = -0.06\, \mathrm{V}\). }\tabularnewline
+\toprule
+Figure & \(a (\mathrm{s^{-1}})\) & \(\frac{A_1}{C} (\mathrm{V/s})\) &
+\(\frac{A_2}{C} (\mathrm{V/s})\) & \(t_f \mathrm{s}\)\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Figure & \(a (\mathrm{s^{-1}})\) & \(\frac{A_1}{C} (\mathrm{V/s})\) &
+\(\frac{A_2}{C} (\mathrm{V/s})\) & \(t_f \mathrm{s}\)\tabularnewline
+\midrule
+\endhead
+1A & \(0\) & \(0\) & \(0\) & \(0.2\)\tabularnewline
+1B & \(0\) & \(0\) & \(0\) & \(0.5\)\tabularnewline
+1C & \(5\) & \(0\) & \(0\) & \(0.2\)\tabularnewline
+1D & \(5\) & \(0\) & \(0\) & \(0.5\)\tabularnewline
+1E & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
+1F & \(5\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
+1G & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
+1H & \(5\) & \(0\) & \(0\) & \(0.3\)\tabularnewline
+1I & \(5\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
+1J & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
+1K & \(30\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
+1L & \(30\) & \(10\) & \(-0.6\) & \(0.4\)\tabularnewline
+1M & \(5\) & \(10\) & \(-0.6\) & \(0.5\)\tabularnewline
+1N & \(5\) & \(10\) & \(-0.6\) & \(0.5\)\tabularnewline
+1O & \(5\) & \(10\) & \(-0.6\) & \(1.0\)\tabularnewline
+1P & \(5\) & \(5\) & \(-0.3\) & \(0.5\)\tabularnewline
+1Q & \(5\) & \(5\) & \(-0.3\) & \(0.2\)\tabularnewline
+1R & \(0\) & \(8\) & \(-0.1\) & \(0.2\)\tabularnewline
+1S & \(5\) & \(-3\) & \(0.5\) & \(0.8\)\tabularnewline
+1T & \(-80\) & \(0\) & \(0\) & \(0.05\)\tabularnewline
+\bottomrule
+\end{longtable}
 
-%%
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{CCp{28.2em}}
-        \thickhline
-        \multicolumn{3}{c}{Model Parameters} \\ \thickhline
-        Figure & Type & $I_e/C (\Rm{V/s})$  \\
-        \thickhline
-        1A & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{Gray}
-        1B & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1 + 10^{-6}$ \\\rowcolor{LightGray}
-        1C & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
-        1D &\includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{LightGray}
-        1E &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
-              $1.5 (0.1 \Rm{s}),\, 0 (0.5\Rm{s}),\, 0.5 (0.1\Rm{s}),\, 1
-              (0.1\Rm{s}),\, 1.5 (0.1\Rm{s}),\, 0 (0.1 \Rm{s})$
-              \\\rowcolor{Gray}
-        1F &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.02
-        \Rm{s}),\, 0 (0.18 \Rm{s}),\, -1.5 (0.025 \Rm{s}),\, 0 (0.025
-        \Rm{s}),\, 1.5 (0.025 \Rm{s}),\, 0 (0.125 \Rm{s})$
-        \\\rowcolor{LightGray} 
-        1G & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.05
-        \Rm{s}),\, -3.5 (0.756 \Rm{s}),\, 0 (0.194 \Rm{s})$ \\\rowcolor{Gray} 
-        1H & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2(1 +
-        10^{-6})$ \\\rowcolor{LightGray}
-        1I & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
-        \begin{tabular}[x]{@{}c@{}} 
-        $1.5 (0.02 \Rm{s}),\, 0 (0.01 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\,
-            0 (0.25 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\, 0 (0.02 \Rm{s}) $\\
-            $ 1.5 (0.02 \Rm{s}),\, 0 (0.04 \Rm{s})$ 
-        \end{tabular}
-            \\\rowcolor{Gray}
-        1J & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.1
-        \Rm{s}),\, 1.7 (0.4 \Rm{s}),\, 1.5 (0.1 \Rm{s}),\, 1.7 (0.4 \Rm{s})$
-        \\\rowcolor{LightGray}
-        1K & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$  \\\rowcolor{Gray}
-        1L & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$ \\\rowcolor{LightGray}
-        1M & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
-        1N & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$  \\\rowcolor{LightGray}
-        1O &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.1 
-        \Rm{s}),\, -3.5 (0.5 \Rm{s}),\, 0 (0.4 \Rm{s})$ \\\rowcolor{Gray}
-        1P & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{LightGray}
-        1Q & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $2 (0.015
-        \Rm{s}),\, 0 (0.185 \Rm{s})$  \\\rowcolor{Gray}
-        1R & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}& $5 (0.01
-        \Rm{s}),\, 0 (0.09 \Rm{s}),\, 5 (0.01 \Rm{s}),\, 0 (0.09 \Rm{s})$
-        \\\rowcolor{LightGray}
-        1S & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}&
-        \begin{tabular}[x]{@{}c@{}} 
-            $5 (0.005 \Rm{s}),\, 0 (0.005 \Rm{s}),\, 4 (0.005 \Rm{s}),\, 0 (0.385 \Rm{s}),\,
-            5 (0.005 \Rm{s}),\, 0 (0.045 \Rm{s}) $\\
-            $4 (0.005\Rm{s}),\, 0 (0.345 \Rm{s}) $
-        \end{tabular}
-        \\\rowcolor{Gray}
-        1T & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $8 (0.002 \Rm{s}),\, 0 (0.048 \Rm{s})$ \\
-        \thickhline
-    \end{tabular}
-    \caption{{\bfseries \sffamily External current.} This table provides the
-    external current for each panel in Figure~\ref{fig:1}. There are two types
-    of external currents, constants and pulses. In the case of pulses the
-    duration of each pulse is given in seconds along with its intensity.
-    }
-    \label{table:5}
-\end{table}
-%%
+\pagebreak
 
-%%
-\begin{table}[!htbp]
-    \centering
-    \begin{tabular}{cc}
-        \thickhline
-        \multicolumn{2}{c}{Initial Conditions} \\ \thickhline
-        Variable & Initial Value  \\ 
-        \rowcolor{Gray}
-        $V(t)$ & $-0.07\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{LightGray}
-        $\Theta(t)$ & $-0.05\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{Gray}
-        $I_1(t)$ & $0.01\, \Rm{V}$ \\ \rowcolor{LightGray} 
-        $I_2(t) $ & $0.001\, \Rm{V}$ \\ \thickhline
-    \end{tabular}
-    \caption{{\bfseries \sffamily Initial conditions.} In all simulations
-    have been used the same initial conditions, except from the one 
-    illustrated in figure~\ref{fig:1}H.}
-    \label{table:6}
-\end{table}
-%%
+\hypertarget{tbl:5}{}
+\begin{longtable}[]{@{}lll@{}}
+\caption{\label{tbl:5}\textbf{External current}. This table provides the
+external current for each panel in Figure~~\ref{fig:1}. There are two
+types of external currents, constants and pulses. In the case of pulses
+the duration of each pulse is given in seconds along with its intensity.
+}\tabularnewline
+\toprule
+Figure & Type & \(I_e/C (V/s)\)\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Figure & Type & \(I_e/C (V/s)\)\tabularnewline
+\midrule
+\endhead
+1A & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(1.5\)\tabularnewline
+1B & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(1 + 10^{-6}\)\tabularnewline
+1C & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(2\)\tabularnewline
+1D & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(1.5\)\tabularnewline
+1E & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(1.5 (0.1 \mathrm{s}),\, 0 (0.5\mathrm{s}),\, 0.5 (0.1\mathrm{s}),\, 1 (0.1\mathrm{s}),\, 1.5 (0.1\mathrm{s}),\, 0 (0.1 \mathrm{s}) \)\tabularnewline
+1F & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(1.5 (0.02 \mathrm{s}),\, 0 (0.18 \mathrm{s}),\, -1.5 (0.025 \mathrm{s}),\, 0 (0.025 \mathrm{s}),\, 1.5 (0.025 \mathrm{s}),\, 0 (0.125 \mathrm{s})\)\tabularnewline
+1G & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(0 (0.05 \mathrm{s}),\, -3.5 (0.756 \mathrm{s}),\, 0 (0.194 \mathrm{s}) \)\tabularnewline
+1H & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\( 2(1 + 10^{-6}) \)\tabularnewline
+1I & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\( 1.5 (0.02 \mathrm{s}),\, 0 (0.01 \mathrm{s}),\, 1.5 (0.02 \mathrm{s}),\, 0 (0.25 \mathrm{s}),\, 1.5 (0.02 \mathrm{s}),\, 0 (0.02 \mathrm{s})\)\tabularnewline
+& & \( 1.5 (0.02 \mathrm{s}),\, 0 (0.04 \mathrm{s}) \)\tabularnewline
+1J & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\( 1.5 (0.1 \mathrm{s}),\, 1.7 (0.4 \mathrm{s}),\, 1.5 (0.1 \mathrm{s}),\, 1.7 (0.4 \mathrm{s}) \)\tabularnewline
+1K & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(-1\)\tabularnewline
+1L & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(-1\)\tabularnewline
+1M & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(2\)\tabularnewline
+1N & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(1.5\)\tabularnewline
+1O & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(0 (0.1 \mathrm{s}),\, -3.5 (0.5 \mathrm{s}),\, 0 (0.4 \mathrm{s})\)\tabularnewline
+1P & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
+\(2\)\tabularnewline
+1Q & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(2 (0.015 \mathrm{s}),\, 0 (0.185 \mathrm{s})\)\tabularnewline
+1R & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(5 (0.01 \mathrm{s}),\, 0 (0.09 \mathrm{s}),\, 5 (0.01 \mathrm{s}),\, 0 (0.09 \mathrm{s})\)\tabularnewline
+1S & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(5 (0.005 \mathrm{s}),\, 0 (0.005 \mathrm{s}),\, 4 (0.005 \mathrm{s}),\, 0 (0.385 \mathrm{s}),\, 5 (0.005 \mathrm{s}),\, 0 (0.045 \mathrm{s}) \)\tabularnewline
+& & \(4 (0.005\mathrm{s}),\, 0 (0.345 \mathrm{s}) \)\tabularnewline
+1T & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
+\(8 (0.002 {\mathrm{s}}),\, 0 (0.048 \mathrm{s})\)\tabularnewline
+\bottomrule
+\end{longtable}
 
-All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
-generation i$7$ processor, $16\, \Rm{GB}$ of physical memory and running Arch
-Linux. The total execution time of all simulations was $2.41$ seconds and the
-peak consumed memory was $162\, \Rm{MB}$\footnote{Python memory profiler used
-    (\url{https://pypi.python.org/pypi/memory_profiler}).}\@.
-
-\clearpage
+\hypertarget{tbl:6}{}
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tbl:6}\textbf{Initial Conditions} In all simulations the
+same initial conditions have been used, except from the one illustrated
+in Figure ~\ref{fig:1} H. }\tabularnewline
+\toprule
+Variable & Initial Value\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Variable & Initial Value\tabularnewline
+\midrule
+\endhead
+\(V(t)\) & \(-0.07\, \mathrm{V}\) / \(-0.03\, \mathrm{V}\) (Figure
+1H)\tabularnewline
+\(\Theta(t)\) & \(-0.05\, \mathrm{V}\) / \(-0.03\, \mathrm{V}\) (Figure
+1H)\tabularnewline
+\(I_1(t)\) & \(0.01\, \mathrm{V}\)\tabularnewline
+\(I_2(t)\) & \(0.001\, \mathrm{V}\)\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-All three figures from the original article have been successfully replicated.
-All the different spiking behaviors of the model are illustrated in 
-Figure~\ref{fig:1}, where the black solid line indicates the membrane potential
-($V(t)$), the red dashed line illustrates the instantaneous threshold potentials
-($\Theta(t)$), and the gray line shows the input to the neuron ($I_e/C$).
-Figures~\ref{fig:2} and~\ref{fig:3} depict the phase space of the phasic
-spiking ($V(t)$ and $\Theta(t)$) and phasic bursting ($V(t)$,
-$I_1(t)$, and $I_2(t)$). All the figures express the same qualitative
-behavior as the original figures in~\cite{mihalas:2009}.
+All three figures from the original article have been successfully
+replicated. All the different spiking behaviors of the model are
+illustrated in Figure~~\ref{fig:1}, where the black solid line indicates
+the membrane potential (\(V(t)\)), the red dashed line illustrates the
+instantaneous threshold potentials (\(\Theta(t)\)), and the gray line
+shows the input to the neuron (\(I_e/C\)). The \(x\)-axis scales in all
+panels are exactly the same as in the original paper (indicating the
+total simulation time (\(t_f\)), while the \(y\)-axis scale differs from
+the one in the original paper. In this work the \(y\)-axis scale is the
+same for all the subplots (\([-95, -25]\, {\mathrm{mV}}\)), except for
+panels G and O (\([-145, -25]\, {\mathrm{mV}}\)).
 
-\begin{figure}[htpb!]
+Figures~~\ref{fig:2} and~~\ref{fig:3} depict the phase space of the
+phasic spiking (\(V(t)\) and \(\Theta(t)\)) and phasic bursting
+(\(V(t)\), \(I_1(t)\), and \(I_2(t)\)). In both figures the blue curves
+and the black dots indicate the trajectory of the system and spike
+events, respectively. In Figure~~\ref{fig:2} the gray arrows show the
+evolution of the system (vector field of the system).
+Figure~~\ref{fig:3} has a different orientation from the original one
+but both the original figure and the replicated one illustrate the same
+trajectories and spike events of the system. All the figures express the
+same qualitative behavior as the original figures
+in~\autocite{mihalas:2009}.
+
+\begin{figure}
 \centering
-    \includegraphics[width=1.1\textwidth]{figs/Figure01.pdf}
-    \caption{{\bfseries \sffamily Neural responses of MNN.} Black solid lines
-    indicate the membrane potential ($V(t)$), the red dashed lines show the
-    threshold potentials ($\Theta(t)$), and the gray lines the external
-    currents applied on each case.
-    {\bfseries \sffamily A} tonic spiking,
-    {\bfseries \sffamily B} class $1$,
-    {\bfseries \sffamily C} spike frequency adaptation,
-    {\bfseries \sffamily D} phasic spiking,
-    {\bfseries \sffamily E} accommodation,
-    {\bfseries \sffamily F} threshold variability,
-    {\bfseries \sffamily G} rebound spike,
-    {\bfseries \sffamily H} class $2$,
-    {\bfseries \sffamily I} integrator,
-    {\bfseries \sffamily J} input bistability,
-    {\bfseries \sffamily K} hyperpolarization induced spiking,
-    {\bfseries \sffamily L} hyperpolarization induced bursting,
-    {\bfseries \sffamily M} tonic bursting,
-    {\bfseries \sffamily N} phasic bursting,
-    {\bfseries \sffamily O} rebound burst,
-    {\bfseries \sffamily P} mixed mode,
-    {\bfseries \sffamily Q} afterpotentials,
-    {\bfseries \sffamily R} basal bistability,
-    {\bfseries \sffamily S} preferred frequency,
-    {\bfseries \sffamily T} spike latency.} 
-    \label{fig:1}
+\includegraphics{figs/Figure01.pdf}
+\caption{\textbf{Neural responses of MNN.} Black solid lines indicate
+the membrane potential (\(V(t)\)), the red dashed lines show the
+threshold potentials (\(\Theta(t)\)), and the gray lines the external
+currents applied on each case. {\textbf{A}} tonic spiking, {\textbf{B}}
+class \(1\), {\textbf{C}} spike frequency adaptation, {\textbf{D}}
+phasic spiking, {\textbf{E}} accommodation, {\textbf{F}} threshold
+variability, {\textbf{G}} rebound spike, {\textbf{H}} class \(2\),
+{\textbf{I}} integrator, {\textbf{J}} input bistability, {\textbf{K}}
+hyperpolarization induced spiking, {\textbf{L}} hyperpolarization
+induced bursting, {\textbf{M}} tonic bursting, {\textbf{N}} phasic
+bursting, {\textbf{O}} rebound burst, {\textbf{P}} mixed mode,
+{\textbf{Q}} afterpotentials, {\textbf{R}} basal bistability,
+{\textbf{S}} preferred frequency, {\textbf{T}} spike
+latency.}\label{fig:1}
 \end{figure}
 
-\begin{figure}[htpb!]
+\begin{figure}
 \centering
-    \includegraphics[width=0.6\textwidth]{figs/Figure02.pdf}
-    \caption{{\bfseries \sffamily Phase space of phasic spiking.} Blue
-    solid lines indicate the trajectories of the model in the phase spiking
-    behavior (Figure~\ref{fig:1}D). The dashed line corresponds to $V(t) =
-    \Theta(t)$, and the black dots represent spike events. The parameters for 
-    this simulation are the same as in Figure~\ref{fig:1}D.}
-    \label{fig:2}
+\includegraphics{figs/Figure02.pdf}
+\caption{\textbf{Phase space of phasic spiking.} Blue solid lines
+indicate the trajectories of the model in the phase spiking behavior
+(Figure~~\ref{fig:1} D). The dashed line corresponds to
+\(V(t) =  \Theta(t)\), and the black dots represent spike events. The
+parameters for this simulation are the same as in Figure~~\ref{fig:1}
+D.}\label{fig:2}
 \end{figure}
 
-\begin{figure}[htpb!]
+\begin{figure}
 \centering
-    \includegraphics[width=0.8\textwidth]{figs/Figure03.pdf}
-    \caption{{\bfseries \sffamily Phase space of phasic bursting.} Blue solid
-    lines represent the trajectories of the system and the black dots indicate
+\includegraphics{figs/Figure03.pdf}
+\caption{\textbf{Phase space of phasic bursting.} Blue solid lines
+represent the trajectories of the system and the black dots indicate
 spiking events. The parameters for this simulation are the same as in
-    Figure~\ref{fig:1}N.}
-    \label{fig:3}
+Figure~~\ref{fig:1} N.}\label{fig:3}
 \end{figure}
@@
 All figures in~\textcite{mihalas:2009} have been successfully replicated
-with high fidelity. Overall, the whole reproducing process was smooth and
-without obscure points since most of the parameters are provided in the
-original article. Only the external current time intervals and the initial
-conditions are not provided explicitly and thus we had to extract that
-information from figure $1$ of the original article. To conclude, the
-article~\cite{mihalas:2009} has been successfully reproduced without any
-discrepancy.
+with high fidelity. Overall, the whole reproducing process was smooth
+and without obscure points since most of the parameters are provided in
+the original article. Only the time intervals for which the external
+current is applied to the model and the initial conditions are not
+provided explicitly. Therefore, we had to extract that information from
+figure \(1\) of the original article. To conclude, the
+article~\autocite{mihalas:2009} has been successfully reproduced without
+any discrepancy.
 

### 2017-10-04 article/detorakis-2017_bak.tex
markdown file fixed

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
% \hypersetup{breaklinks=true,
%             bookmarks=true,
%             pdfauthor=
% {
%       Georgios Detorakis,
%   },
%             pdftitle=
% {
% [Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
%         Spiking Behavior
% },
%             colorlinks=true,
%             citecolor=blue,
%             urlcolor=blue,
%             linkcolor=blue,
%             pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

\usepackage{float,lscape}
\usepackage{xcolor,colortbl}
\usepackage{amssymb,amsmath,array,amsthm}
\usepackage{marvosym}
% \usepackage{makecell}
\usepackage{boldline}
\usepackage{bm,array}

\usepackage{adjustbox}

\newcommand{\Rm}[1]{\mathrm{#1}}

\newcolumntype{C}{>{\centering\arraybackslash}p{3.5em}}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\definecolor{Gray}{gray}{0.75}
\definecolor{LightGray}{gray}{0.95}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces
            Diverse Spiking Behaviors}

    \usepackage{authblk}
                        \author[1]{Georgios Detorakis}
                            \affil[1]{Department of Cognitive Sciences, UC Irvine,
                            Irvine, CA, USA}

\date{\vspace{-5mm}
      \sffamily \small
      \href{mailto:corresponding-author@mail.com}{gdetorak@uci.edu \\
      gdetor@protonmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The author has declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] A Generalized Linear Integrate-and-Fire Neural Model Produces
        Diverse Spiking Behaviors, Stefan Mihalas and Ernst Niebur, Neural
        Computation 21, 704--718, 2009.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Integrate-and-fire neurons are being used extensively in the field of 
neuroscience for modeling spiking behaviors~\cite{dayan:2001}. In this work we
provide a reference implementation of~\cite{mihalas:2009}, where the authors
have introduced a generalization of the leaky integrate-and-fire neuron model. 
The Mihalas-Niebur Neuron (MNN) model is a linear integrate-and-fire neuron
model capable of expressing a rich spiking behavior based on a set of 
parameters.

An MNN model expresses tonic and phasic spiking, class $1$ and $2$, spike
frequency adaptation, accommodation, threshold variability, rebound spike,
integrator, input bistability, hyperpolarizing spiking and bursting, tonic,
phasic and rebound bursting, mixed mode, afterpotentials, basal bistability, 
preferred frequency and spike latency. 
Due to its simplicity, the MNN model has been used in neuromorphic 
implementations such as~\cite{folowosele:2011}.

The model consists of linear differential equations, which describe the membrane
and threshold potentials and internal currents. All the results provided in 
\cite{mihalas:2009} have been obtained by using only two internal currents and
thus we use the exact same number of internal currents in this work. The
subthreshold dynamics are defined by a set of linear ordinary differential 
equations, while an instantaneous threshold potential controls when the neuron
fires an action potential (spike) in a dynamic way. The ability of the MNN 
model to generate 
such a diverse spiking behavior is due to the complex update rules. In this 
work the MNN model has been implemented in Python (version 3.6.1) using 
Numpy (version 1.13.1) and Matplotlib (version 2.0.2) packages.

\section{Methods}\label{methods}

In order to implement the model described in \cite{mihalas:2009}, we discretized
the dynamical system using the forward Euler integration scheme. The time step
is fixed to $0.1\, \Rm{ms}$ for all the simulations, and the total simulation
time $t_f$ varies according to figure 1 of the original paper. Our
implementation differs from the one in the original paper, since in
\cite{mihalas:2009}, authors numerically solve equation $3.5$ (algebraic 
equation) under the constraint imposed by inequality $3.4$ and thus they
compute the spike times. On the other hand, in this work we directly compute
numerically the solution of the dynamical system defined by equations $2.1$
and $2.2$ in \cite{mihalas:2009} (see tables~\ref{table:2} and~\ref{table:3}).

We provide all equations and parameters of the model in tables as it has been 
suggested by~\cite{nordlie:2009}. 
Table~\ref{table:1} provides the summary of the model. Tables~\ref{table:2}
and~\ref{table:3} give the subthreshold dynamics (differential equations) 
describing the membrane and the threshold potentials as well as the two internal
currents and the update rules. The parameters for all the simulations are given
in table~\ref{table:4}, while the external current intensities and pulse
duration are provided in table~\ref{table:5}. The parameters in this work are 
exactly the same used in the original paper (table 1, pg. $711$). We had to 
infer the time intervals and the total simulation times for the pulses since
they are not given explicitly in the original paper. Thus, we extracted the
time intervals from figure $1$ of \cite{mihalas:2009} by visual inspection. 
The initial conditions are given in table~\ref{table:6}.

All simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
generation i$7$ processor, $16\, \Rm{GB}$ of physical memory and running Arch
Linux (x$86\_64$). The total execution time of all simulations was $2.41$
seconds and the peak consumed memory was $162\, \Rm{MB}$\footnote{Python memory
profiler used (\url{https://pypi.python.org/pypi/memory_profiler}).}\@.

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ll}
        \thickhline
        \multicolumn{2}{c}{Model Summary} \\\thickhline
        \rowcolor{Gray}
        Populations  & No population -- single neuron model \\\rowcolor{LightGray}
        Topology     & -- \\ \rowcolor{Gray}
        Connectivity & -- \\ \rowcolor{LightGray}
        Neuron Model & Linear Integrate-and-Fire Neuron \\\rowcolor{Gray}
        Channel Models & Linear, first order ODEs  \\ \rowcolor{LightGray}
        Synapse Model & -- \\ \rowcolor{Gray}
        Plasticity & -- \\ \rowcolor{LightGray}
        Input & Constant current or rectangular pulses \\\rowcolor{Gray}
        Measurements & Membrane potential, phase plane \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Summary of the model}} 
    \label{table:1}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{3.5cm}ll}
        \thickhline
        \multicolumn{2}{c}{Neuron Model} \\\thickhline
        \rowcolor{Gray}
        Name  &  Mihalas-Niebur Neuron (MNN) \\ \rowcolor{LightGray}
        Type  &  Linear Leaky Integrate-and-Fire Neuron  \\ \rowcolor{Gray}
        Membrane Potential & $
            \begin{aligned}
                \frac{\Rm{d}V(t)}{\Rm{d}t} &= \frac{1}{C} \Big(I_e + I_1 + I_2 - G(V(t) -
                E_L)  \Big)
            \end{aligned}$ \\ \rowcolor{LightGray}
        Instantaneous Threshold Potential  & $
            \begin{aligned}
                \frac{\Rm{d}\Theta(t)}{\Rm{d}t} &= a(V(t) - E_L) - b(\Theta(t) -
                \Theta_{\infty}) 
            \end{aligned} $
            \\ \rowcolor{Gray}
        Internal Currents & $
            \begin{aligned}
                \frac{\Rm{d}I_{1}(t)}{\Rm{d}t} &= -k_1I_1(t)\\
                \frac{\Rm{d}I_{2}(t)}{\Rm{d}t} &= -k_2I_2(t) \\
            \end{aligned} $
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the subthreshold dynamics of
    Mihalas-Niebur neuron model.} $V(t)$ and $\Theta(t)$ are the membrane and 
    threshold potentials, respectively. $E_L$ and $\Theta_{\infty}$ are the
    reversal potentials for the membrane and the threshold variables,
    respectively. $a, b, k_1, k_2$ and $G$ are constant parameters. $I_e$ is
    the external current applied on the neuron model.} 
    \label{table:2}
\end{table}
%%

\begin{table}[!htbp]
    \centering
    \begin{tabular}{cc}
        \thickhline
        \multicolumn{2}{c}{Update rules} \\ \thickhline
        Variable & Rule  \\ \rowcolor{Gray}
        $V(t)$ & $V_r$  \\ \rowcolor{LightGray}
        $\Theta(t)$ & $\max\{\Theta_r, \Theta(t) \}$ \\ \rowcolor{Gray}
        $I_1(t)$ & $R_1 \times I_1(t) + A_1$ \\ \rowcolor{LightGray} 
        $I_2(t) $ & $R_2 \times I_2(t) + A_2$ \\\thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Update rules.} $V_r$ and $\Theta_r$
    are the reset values for the membrane and threshold potentials,
    respectively. $R_1, R_2, A_1$ and $A_2$ are constants.}
    \label{table:3}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{CCCCC}
        \thickhline
        \multicolumn{5}{c}{Model Parameters} \\\thickhline
           \begin{tabular}[x]{@{}c@{}} Figure \\  \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $a$ \\ $(\Rm{s^{-1}})$ \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $A_1/C$ \\ $(\Rm{V/s})$ \end{tabular}
         & \begin{tabular}[x]{@{}c@{}} $A_2/C$ \\ $(\Rm{V/s})$ \end{tabular}& 
           \begin{tabular}[x]{@{}c@{}} $t_f$ \\ $(\Rm{s})$ \end{tabular} \\
        \thickhline
        1A & $0$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
        1B & $0$ & $0$ & $0$ & $0.5$  \\\rowcolor{LightGray}
        1C & $5$ & $0$ & $0$ & $0.2$ \\\rowcolor{Gray}
        1D & $5$ & $0$ & $0$ & $0.5$ \\\rowcolor{LightGray}
        1E & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{Gray}
        1F & $5$ & $0$ & $0$ & $0.4$ \\\rowcolor{LightGray}
        1G & $5$ & $0$ & $0$ & $1.0$  \\\rowcolor{Gray}
        1H &  $5$ & $0$ & $0$ & $0.3$ \\\rowcolor{LightGray}
        1I & $5$  & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
        1J & $5$ & $0$ & $0$ & $1.0$ \\\rowcolor{LightGray}
        1K & $30$ & $0$ & $0$ & $0.4$ \\\rowcolor{Gray}
        1L & $30$  & $10$ & $-0.6$ & $0.4$ \\\rowcolor{LightGray}
        1M & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{Gray}
        1N & $5$ & $10$ & $-0.6$ & $0.5$ \\\rowcolor{LightGray}
        1O & $5$ & $10$ & $-0.6$ & $1.0$ \\\rowcolor{Gray}
        1P & $5$  & $5$ & $-0.3$ & $0.5$ \\\rowcolor{LightGray}
        1Q & $5$  & $5$ & $-0.3$ & $0.2$ \\\rowcolor{Gray}
        1R & $0$ & $8$ & $-0.1$ & $0.2$ \\\rowcolor{LightGray}
        1S & $5$ & $-3$ & $0.5$ & $0.8$ \\\rowcolor{Gray}
        1T &  $-80$ & $0$ & $0$ & $0.05$ \\
        \thickhline
        \multicolumn{5}{l}{Common Parameters} \\\rowcolor{LightGray}
        \thickhline
        \multicolumn{5}{p{22.5em}}{
            $b = 10\, \Rm{s^{-1}}$, $G/C = 50\, \Rm{s^{-1}}$,
            $k_1 = 200\, \Rm{s^{-1}}$, $k_2 = 20\, \Rm{s^{-1}}$,
            $\Theta_{\infty} = -0.05\, \Rm{V}$,
            $R_1 = 0$, $R_2 = 1$, $E_l = -0.07\, \Rm{V}$,
            $V_r = -0.07\, \Rm{V}$, $\Theta_r = -0.06\, \Rm{V}$.
        }
        \\\thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulation Parameters}  }
    \label{table:4}
\end{table}
%%

\vspace{-1cm}

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{CCp{28.2em}}
        \thickhline
        \multicolumn{3}{c}{Model Parameters} \\ \thickhline
        Figure & Type & $I_e/C (\Rm{V/s})$  \\
        \thickhline
        1A & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{Gray}
        1B & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1 + 10^{-6}$ \\\rowcolor{LightGray}
        1C & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
        1D &\includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$ \\\rowcolor{LightGray}
        1E &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
              $1.5 (0.1 \Rm{s}),\, 0 (0.5\Rm{s}),\, 0.5 (0.1\Rm{s}),\, 1
              (0.1\Rm{s}),\, 1.5 (0.1\Rm{s}),\, 0 (0.1 \Rm{s})$
              \\\rowcolor{Gray}
        1F &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.02
        \Rm{s}),\, 0 (0.18 \Rm{s}),\, -1.5 (0.025 \Rm{s}),\, 0 (0.025
        \Rm{s}),\, 1.5 (0.025 \Rm{s}),\, 0 (0.125 \Rm{s})$
        \\\rowcolor{LightGray} 
        1G & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.05
        \Rm{s}),\, -3.5 (0.756 \Rm{s}),\, 0 (0.194 \Rm{s})$ \\\rowcolor{Gray} 
        1H & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2(1 +
        10^{-6})$ \\\rowcolor{LightGray}
        1I & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & 
        \begin{tabular}[x]{@{}c@{}} 
        $1.5 (0.02 \Rm{s}),\, 0 (0.01 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\,
            0 (0.25 \Rm{s}),\, 1.5 (0.02 \Rm{s}),\, 0 (0.02 \Rm{s}) $\\
            $ 1.5 (0.02 \Rm{s}),\, 0 (0.04 \Rm{s})$ 
        \end{tabular}
            \\\rowcolor{Gray}
        1J & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $1.5 (0.1
        \Rm{s}),\, 1.7 (0.4 \Rm{s}),\, 1.5 (0.1 \Rm{s}),\, 1.7 (0.4 \Rm{s})$
        \\\rowcolor{LightGray}
        1K & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$  \\\rowcolor{Gray}
        1L & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $-1$ \\\rowcolor{LightGray}
        1M & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{Gray}
        1N & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $1.5$  \\\rowcolor{LightGray}
        1O &\includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $0 (0.1 
        \Rm{s}),\, -3.5 (0.5 \Rm{s}),\, 0 (0.4 \Rm{s})$ \\\rowcolor{Gray}
        1P & \includegraphics[width=0.1\textwidth]{figs/const.pdf} & $2$ \\\rowcolor{LightGray}
        1Q & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $2 (0.015
        \Rm{s}),\, 0 (0.185 \Rm{s})$  \\\rowcolor{Gray}
        1R & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}& $5 (0.01
        \Rm{s}),\, 0 (0.09 \Rm{s}),\, 5 (0.01 \Rm{s}),\, 0 (0.09 \Rm{s})$
        \\\rowcolor{LightGray}
        1S & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf}&
        \begin{tabular}[x]{@{}c@{}} 
            $5 (0.005 \Rm{s}),\, 0 (0.005 \Rm{s}),\, 4 (0.005 \Rm{s}),\, 0 (0.385 \Rm{s}),\,
            5 (0.005 \Rm{s}),\, 0 (0.045 \Rm{s}) $\\
            $4 (0.005\Rm{s}),\, 0 (0.345 \Rm{s}) $
        \end{tabular}
        \\\rowcolor{Gray}
        1T & \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} & $8 (0.002 \Rm{s}),\, 0 (0.048 \Rm{s})$ \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily External current.} This table provides the
    external current for each panel in Figure~\ref{fig:1}. There are two types
    of external currents, constants and pulses. In the case of pulses the
    duration of each pulse is given in seconds along with its intensity.
    }
    \label{table:5}
\end{table}
%%
\vspace{-2cm}
%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{cc}
        \thickhline
        \multicolumn{2}{c}{Initial Conditions} \\ \thickhline
        Variable & Initial Value  \\ 
        \rowcolor{Gray}
        $V(t)$ & $-0.07\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{LightGray}
        $\Theta(t)$ & $-0.05\, \Rm{V}$ / $-0.03\, \Rm{V}$ (Figure 1H) \\ \rowcolor{Gray}
        $I_1(t)$ & $0.01\, \Rm{V}$ \\ \rowcolor{LightGray} 
        $I_2(t) $ & $0.001\, \Rm{V}$ \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Initial conditions.} In all simulations
    have been used the same initial conditions, except from the one 
    illustrated in Figure~\ref{fig:1}H.}
    \label{table:6}
\end{table}
%%

\clearpage

\section{Results}\label{results}

All three figures from the original article have been successfully replicated.
All the different spiking behaviors of the model are illustrated in 
Figure~\ref{fig:1}, where the black solid line indicates the membrane potential
($V(t)$), the red dashed line illustrates the instantaneous threshold potentials
($\Theta(t)$), and the gray line shows the input to the neuron ($I_e/C$). 
The $x$-axis scales in all panels are exactly the same as in the original
paper (indicating the total simulation time ($t_f$), while the $y$-axis scale 
differs from the one in the original paper. In this work the $y$-axis scale
is the same same for all the subplots ($[-95, -25]\, \Rm{mV}$), except for
panels G and O ($[-145, -25]\, \Rm{mV}$).

Figures~\ref{fig:2} and~\ref{fig:3} depict the phase space of the phasic
spiking ($V(t)$ and $\Theta(t)$) and phasic bursting ($V(t)$,
$I_1(t)$, and $I_2(t)$). In both figures the blue curves and the black dots 
indicate the trajectory of the system and spike events, respectively.
In Figure~\ref{fig:2} the gray arrows show the evolution of the system
(vector field of the system). Figure~\ref{fig:3} has a different orientation 
from the original one but both illustrate the same trajectories and spike 
events of the system. 
All the figures express the same qualitative behavior as the original
figures in~\cite{mihalas:2009}.

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=0.9\textwidth]{figs/Figure01.pdf}
    \caption{{\bfseries \sffamily Neural responses of MNN.} Black solid lines
    indicate the membrane potential ($V(t)$), the red dashed lines show the
    threshold potentials ($\Theta(t)$), and the gray lines the external
    currents applied on each case.
    {\bfseries \sffamily A} tonic spiking,
    {\bfseries \sffamily B} class $1$,
    {\bfseries \sffamily C} spike frequency adaptation,
    {\bfseries \sffamily D} phasic spiking,
    {\bfseries \sffamily E} accommodation,
    {\bfseries \sffamily F} threshold variability,
    {\bfseries \sffamily G} rebound spike,
    {\bfseries \sffamily H} class $2$,
    {\bfseries \sffamily I} integrator,
    {\bfseries \sffamily J} input bistability,
    {\bfseries \sffamily K} hyperpolarization induced spiking,
    {\bfseries \sffamily L} hyperpolarization induced bursting,
    {\bfseries \sffamily M} tonic bursting,
    {\bfseries \sffamily N} phasic bursting,
    {\bfseries \sffamily O} rebound burst,
    {\bfseries \sffamily P} mixed mode,
    {\bfseries \sffamily Q} afterpotentials,
    {\bfseries \sffamily R} basal bistability,
    {\bfseries \sffamily S} preferred frequency,
    {\bfseries \sffamily T} spike latency.} 
    \label{fig:1}
\end{figure}

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/Figure02.pdf}
    \caption{{\bfseries \sffamily Phase space of phasic spiking.} Blue
    solid lines indicate the trajectories of the model in the phase spiking
    behavior (Figure~\ref{fig:1}D). The dashed line corresponds to $V(t) =
    \Theta(t)$, and the black dots represent spike events. The parameters for 
    this simulation are the same as in Figure~\ref{fig:1}D.}
    \label{fig:2}
\end{figure}

\begin{figure}[htpb!]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/Figure03.pdf}
    \caption{{\bfseries \sffamily Phase space of phasic bursting.} Blue solid
    lines represent the trajectories of the system and the black dots indicate
    spiking events. The parameters for this simulation are the same as in
    Figure~\ref{fig:1}N.}
    \label{fig:3}
\end{figure}

\section{Conclusion}\label{conclusion}

All figures in~\textcite{mihalas:2009} have been successfully replicated 
with high fidelity. Overall, the whole reproducing process was smooth and
without obscure points since most of the parameters are provided in the
original article. Only the time intervals for which the external current is
applied to the model and the initial conditions are not provided explicitly.
Therefore, we had to extract that information from figure $1$ of the original
article. To conclude, the
article~\cite{mihalas:2009} has been successfully reproduced without any
discrepancy.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-10-16 article/Poisot-2017.tex
All figures done, some text in

--- 2017-10-04 article/detorakis-2017.tex
@@
 \usepackage{hyperref}
-\usepackage{amssymb,amsmath,array,amsthm}
 \hypersetup{breaklinks=true,
@@
 {
-      Georgios Detorakis,
+      Timothée Poisot,
   },
@@
 {
-[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
-Spiking Behaviors
+[Re] On the coexistence of specialists and generalists
 },
@@
 % --- Document ----------------------------------------------------------------
-\title{[Re] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
-Spiking Behaviors}
+\title{[Re] On the coexistence of specialists and generalists}
 
     \usepackage{authblk}
-                        \author[1]{Georgios Detorakis}
-                            \affil[1]{Department of Cognitive Sciences, UC Irvine, Irvine CA, USA}
+                        \author[1]{Timothée Poisot}
+                            \affil[1]{Département de Sciences Biologiques, Université de Montréal, Montréal,
+QC, Canada}
 
 \date{\vspace{-5mm}
-      \sffamily \small \href{mailto:gdetorak@uci.edu (gdetor@protonmail.com)}{gdetorak@uci.edu (gdetor@protonmail.com)}}
+      \sffamily \small \href{mailto:timothee.poisot@umontreal.ca}{timothee.poisot@umontreal.ca}}
 
@@
 \begin{itemize}
-    \item[→] A Generalized Linear Integrate-and-Fire Neural Model Produces Diverse
-Spiking Behaviors, Stefan Mihalas and Ernst Niebur, Neural Computation
-21, 704--718, 2009.
+    \item[→] On the coexistence of specialists and generalists, David Sloan Wilson
+and Jin Yoshimura, The American Naturalist 144:4, 692-707, 1994.
   \end{itemize}\par
@@
 
-Integrate-and-fire neurons are being used extensively in the field of
-neuroscience for modeling spiking behaviors~\autocite{dayan:2001}. In
-this work we provide a reference implementation
-of~\autocite{mihalas:2009}, where the authors have introduced a
-generalization of the leaky integrate-and-fire neuron model. The
-Mihalas-Niebur Neuron (MNN) model is a linear integrate-and-fire neuron
-model capable of expressing a rich spiking behavior based on a set of
-parameters.
-
-An MNN model expresses tonic and phasic spiking, class \(1\) and \(2\),
-spike frequency adaptation, accommodation, threshold variability,
-rebound spike, integrator, input bistability, hyperpolarizing spiking
-and bursting, tonic, phasic and rebound bursting, mixed mode,
-afterpotentials, basal bistability, preferred frequency and spike
-latency. Due to its simplicity, the MNN model has been used in
-neuromorphic implementations such as~\autocite{folowosele:2011}.
-
-The model consists of linear differential equations, which describe the
-membrane and threshold potentials and internal currents. All the results
-provided in \autocite{mihalas:2009} have been obtained by using only two
-internal currents and thus we use the exact same number of internal
-currents in this work. The subthreshold dynamics are defined by a set of
-linear ordinary differential equations, while an instantaneous threshold
-potential controls when the neuron fires an action potential (spike) in
-a dynamic way. The ability of the MNN model to generate such a diverse
-spiking behavior is due to the complex update rules. In this work the
-MNN model has been implemented in Python (version 3.6.1) using Numpy
-(version 1.13.1) and Matplotlib (version 2.0.2) packages.
+The coexistence of specialists and generalist within ecological
+communities is a long-standing question. \textcite{wils94csg} have
+suggested that this coexistence can be understood when examined in the
+light of (i) differential fitness loss associated to specialism, (ii)
+active habitat selection, (iii) negative density dependence due to
+competition, and (iv) stochastic changes in habitat quality, that allow
+combinations of species to persist even though coexistence would not be
+possible in a purely deterministic world. Here I propose an
+implementation of this model in \emph{Julia} \autocite{beza17jfa}, and
+show that it is able to reproduce most figures from the original
+manuscript.
 
@@
 
-In order to implement the model described in \autocite{mihalas:2009}, we
-discretized the dynamical system using the forward Euler integration
-scheme. The time step is fixed to \(0.1\, \mathrm{ms}\) for all the
-simulations, and the total simulation time \(t_f\) varies according to
-figure 1 of the original paper. Our implementation differs from the one
-in the original paper, since in \autocite{mihalas:2009}, authors
-numerically solve equation \(3.5\) (algebraic equation) under the
-constraint imposed by inequality \(3.4\) and thus they compute the spike
-times. On the other hand, in this work we directly compute numerically
-the solution of the dynamical system defined by equations \(2.1\) and
-\(2.2\) in \autocite{mihalas:2009} (see tables~~\ref{tbl:2}
-and~~\ref{tbl:3}).
+The \textcite{wils94csg} model describes three species across two
+patches of habitat. Species 1 is a specialist of habitat 1, species 2 is
+a specialist of habitat 2, and species 3 is a generalist. This results
+in the maximum density that these species can reach in both habitats:
 
-We provide all equations and parameters of the model in tables as it has
-been suggested by~\autocite{nordlie:2009}. Table~~\ref{tbl:1} provides
-the summary of the model. Tables~~\ref{tbl:2} and~~\ref{tbl:3} give the
-subthreshold dynamics (differential equations) describing the membrane
-and the threshold potentials as well as the two internal currents and
-the update rules. The parameters for all the simulations are given in
-table~~\ref{tbl:4}, while the external current intensities and pulse
-duration are provided in table~~\ref{tbl:5}. The parameters in this work
-are exactly the same used in the original paper (table 1, pg. \(711\)).
-We had to infer the time intervals and the total simulation times for
-the pulses since they are not given explicitly in the original paper.
-Thus, we extracted the time intervals from figure \(1\) of
-\autocite{mihalas:2009} by visual inspection. The initial conditions are
-given in table~~\ref{tbl:6}.
+\begin{equation}
+\mathbf{K} = \begin{bmatrix}
+  K_1 & aK_1 \\
+  a_K2 & K_2 \\
+  bK_1 & b_K2
+\end{bmatrix} \,.
+\end{equation}
 
-All simulations ran on a Dell OptiPlex \(7040\), equipped with a sixth
-generation i\(7\) processor, \(16\, \mathrm{GB}\) of physical memory and
-running Arch Linux (x\(86\_64\)). The total execution time of all
-simulations was \(2.41\) seconds and the peak consumed memory was
-\(162\, {\mathrm{MB}}\)\footnote{Python memory profiler used
-  (\url{https://pypi.python.org/pypi/memory_profiler}).}.
+In this matrix, \(K_1\) is the quality of habitat 1, \(K_2\) is the
+quality of habitat 2, \(a\) is the fitness cost of the specialist in its
+non-optimal environment, and \(b\) is the fitness cost of generalism.
+Note that \(1 > b > a > 0\).
 
-\hypertarget{tbl:1}{}
-\begin{longtable}[]{@{}ll@{}}
-\caption{\label{tbl:1}\textbf{Summary of the model.} }\tabularnewline
+Species distribute themselves across habitats in a way that minimizes
+the negative effect of other species on their fitness. This is modelled
+by each species having a value \(p_i\), which is the proportion of its
+species choosing habitat 1. Values of \(\mathbf{p}\) are found by
+measuring the negative density effect of each species in each habitat:
+
+\begin{equation}
+D_{l1} = \frac{\sum_{i\in l,m,m}p_iN_i}{K_{l1}}
+\end{equation}
+
+and
+
+\begin{equation}
+D_{l2} = \frac{\sum_{i\in l,m,m}(1-p_i)N_i}{K_{l1}} \,.
+\end{equation}
+
+The value of \(p_l\) for which \(D_{l1}=D_{l2}\) is
+
+\begin{equation}
+p_l = - \frac{(K_{l1}+K_{l2})(N_mp_m+N_np_n)-K_{l1}(N_l+N_m+N_n)}{N_l(K_{l1}+K_{l2})} \,.
+\end{equation}
+
+We fix \(p_m\) and \(p_n\), and find the value of \(p_l\), while
+enforcing the constraint of \(0 \leq p_l \leq 1\). Repeating this
+procedure a few times for the different species yields the optimal
+values of \(\mathbf{p}\); we can measure the density of individuals in
+both habitats. Before we do so, there is a proportion \(g\) of
+individuals that select habitats at random. Given a total population
+size of \(N_i\), there are \(N_i(g/2)\) individuals will go to either
+habitat, and \(N_i(1-g)p_i\) will pick habitat 1. With this information,
+we can write the matrix describing habitat selection:
+
+\begin{equation}
+\mathbf{N} = \begin{bmatrix}
+  N_1(\frac{g}{2}+(1-g)p_1) & N_1(\frac{g}{2}+(1-g)(1-p_1))\\
+  N_2(\frac{g}{2}+(1-g)p_2) & N_2(\frac{g}{2}+(1-g)(1-p_2))\\
+  N_3(\frac{g}{2}+(1-g)p_3) & N_3(\frac{g}{2}+(1-g)(1-p_3))
+\end{bmatrix} \,.
+\end{equation}
+
+Finally, the fitness of every species in each habitat is given by
+
+\begin{equation}
+W_{ij} = \text{exp}\left[r\left(1-\frac{N_{i1}+N_{i2}+N_{i3}}{K_{ij}}\right)\right] \,,
+\end{equation}
+
+where \(r\) is the growth rate (assumed equal). The population size at
+the next timestep is simply given by
+
+\begin{equation}
+\mathbf{N}_{t+1} = \mathbf{W}\odot \mathbf{N}_{t} \,,
+\end{equation}
+
+where \(\odot\) is the element-wise multiplication.
+
+\section{Results}\label{results}
+
+Original sources were not available, and no attempts were made to
+contact the authors. For some non-stochastic situations, it is possible
+to calculate expected values by hand. The original manuscript does
+provide some of these values, and they were used to test this
+implementation. Figure 2A and 2B in the original manuscript provide a
+good diagnostic value, as they are based on non-stochastic situations.
+In \textcite{fig:02},
+
+\begin{figure}
+\centering
+\includegraphics{figure02.pdf}
+\caption{Population dynamics and habitat preference of the generalist
+alone (left), and following invasion by the two specialists at the
+generalist equilibrium (right).}\label{fig:02}
+\end{figure}
+
+Results should be compared with original results and you have to explain
+why you think they are the same or why they may differ (qualitative
+result vs quantitative result). Note that it is not necessary to redo
+all the original analysis of the results.
+
+\section{Conclusion}\label{conclusion}
+
+Conclusion, at the very minimum, should indicate very clearly if you
+were able to replicate original results. If it was not possible but you
+found the reason why (error in the original results), you should exlain
+it.
+
+\begin{longtable}[]{@{}llllll@{}}
+\caption{Table caption \{\#tbl:table\}}\tabularnewline
 \toprule
-Model Summary &\tabularnewline
+\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Heading 1\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Heading 2\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage}\tabularnewline
 \midrule
@@
 \toprule
-Model Summary &\tabularnewline
+\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Heading 1\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+Heading 2\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
+\strut
+\end{minipage}\tabularnewline
 \midrule
 \endhead
-Populations & No population -- single neuron model\tabularnewline
-Topology & --\tabularnewline
-Connectivity & --\tabularnewline
-Neuron Model & Linear Integrate-and-Fire Neuron\tabularnewline
-Channel Models & Linear, first order ODEs\tabularnewline
-Synapse Model & --\tabularnewline
-Plasticity & --\tabularnewline
-Input & Constant current/rectangular pulses\tabularnewline
-Measurements & Membrane potential, phase plane\tabularnewline
+cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
+cell6 row 1\tabularnewline
+cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
+cell6 row 2\tabularnewline
+cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
+cell6 row 3\tabularnewline
 \bottomrule
@@
 
-\hypertarget{tbl:2}{}
-\begin{longtable}[]{@{}ll@{}}
-\caption{\label{tbl:2}Description of the subthreshold dynamics of
-Mihalas--Niebur neuron model. \(V(t)\) and \(\Theta(t)\) are the
-membrane and threshold potentials, respectively. \(E_L\) and
-\(\Theta_{\infty}\) are the reversal potentials for the membrane and
-threshold variables, respectively. \(a, b, k_1, k_2\) and \(G\) are
-constant parameters. \(I_e\) is the external current applied on the
-neuron model. }\tabularnewline
-\toprule
-Neuron Model &\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Neuron Model &\tabularnewline
-\midrule
-\endhead
-Name & Mihalas-Niebur Neuron (MNN)\tabularnewline
-Type & Linear Leaky Integrate-and-Fire Neuron\tabularnewline
-Membrane Potential &
-\( \frac{\mathrm{d}V(t)}{\mathrm{d}t} = \frac{1}{C} \Big(I_e + I_1 + I_2 - G(V(t) - E_L)  \Big) \)\tabularnewline
-Instantaneous Threshold Potential &
-\( \frac{\mathrm{d}\Theta(t)}{\mathrm{d}t} = a(V(t) - E_L) - b(\Theta(t) - \Theta_{\infty}) \)\tabularnewline
-Internal Currents &
-\( \frac{\mathrm{d}I_{1}(t)}{\mathrm{d}t} = -k_1I_1(t) \)\tabularnewline
-&
-\( \frac{\mathrm{d}I_{2}(t)}{\mathrm{d}t} = -k_2I_2(t) \)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\pagebreak
-
-\hypertarget{tbl:3}{}
-\begin{longtable}[]{@{}ll@{}}
-\caption{\label{tbl:3}\textbf{Update rules.} \(V_r\) and \(\Theta_r\)
-are the reset values for the membrane and threshold potentials,
-respectively. \(R_1, R_2, A_1\) and \(A_2\) are constants.
-}\tabularnewline
-\toprule
-Variable & Rule\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Variable & Rule\tabularnewline
-\midrule
-\endhead
-\(V(t)\) & \(V_r\)\tabularnewline
-\(\Theta(t)\) & \(\max\{\Theta_r, \Theta(t) \}\)\tabularnewline
-\(I_1(t)\) & \(R_1 \times I_1(t) + A_1\)\tabularnewline
-\(I_2(t)\) & \(R_2 \times I_2(t) + A_2\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{tbl:4}{}
-\begin{longtable}[]{@{}lllll@{}}
-\caption{\label{tbl:4}\textbf{Simulation Parameters}. Common parameters
-for all simulations: \(b = 10\, \mathrm{s^{-1}}\),
-\(G/C = 50\, \mathrm{s^{-1}}\), \(k_1 = 200\, \mathrm{s^{-1}}\),
-\(k_2 = 20\, \mathrm{s^{-1}}\),
-\(\Theta_{\infty} = -0.05\, \mathrm{V}\), \(R_1 = 0\), \(R_2 = 1\),
-\(E_l = -0.07\, \mathrm{V}\), \(V_r = -0.07\, \mathrm{V}\),
-\(\Theta_r = -0.06\, \mathrm{V}\). }\tabularnewline
-\toprule
-Figure & \(a (\mathrm{s^{-1}})\) & \(\frac{A_1}{C} (\mathrm{V/s})\) &
-\(\frac{A_2}{C} (\mathrm{V/s})\) & \(t_f \mathrm{s}\)\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Figure & \(a (\mathrm{s^{-1}})\) & \(\frac{A_1}{C} (\mathrm{V/s})\) &
-\(\frac{A_2}{C} (\mathrm{V/s})\) & \(t_f \mathrm{s}\)\tabularnewline
-\midrule
-\endhead
-1A & \(0\) & \(0\) & \(0\) & \(0.2\)\tabularnewline
-1B & \(0\) & \(0\) & \(0\) & \(0.5\)\tabularnewline
-1C & \(5\) & \(0\) & \(0\) & \(0.2\)\tabularnewline
-1D & \(5\) & \(0\) & \(0\) & \(0.5\)\tabularnewline
-1E & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
-1F & \(5\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
-1G & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
-1H & \(5\) & \(0\) & \(0\) & \(0.3\)\tabularnewline
-1I & \(5\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
-1J & \(5\) & \(0\) & \(0\) & \(1.0\)\tabularnewline
-1K & \(30\) & \(0\) & \(0\) & \(0.4\)\tabularnewline
-1L & \(30\) & \(10\) & \(-0.6\) & \(0.4\)\tabularnewline
-1M & \(5\) & \(10\) & \(-0.6\) & \(0.5\)\tabularnewline
-1N & \(5\) & \(10\) & \(-0.6\) & \(0.5\)\tabularnewline
-1O & \(5\) & \(10\) & \(-0.6\) & \(1.0\)\tabularnewline
-1P & \(5\) & \(5\) & \(-0.3\) & \(0.5\)\tabularnewline
-1Q & \(5\) & \(5\) & \(-0.3\) & \(0.2\)\tabularnewline
-1R & \(0\) & \(8\) & \(-0.1\) & \(0.2\)\tabularnewline
-1S & \(5\) & \(-3\) & \(0.5\) & \(0.8\)\tabularnewline
-1T & \(-80\) & \(0\) & \(0\) & \(0.05\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\pagebreak
-
-\hypertarget{tbl:5}{}
-\begin{longtable}[]{@{}lll@{}}
-\caption{\label{tbl:5}\textbf{External current}. This table provides the
-external current for each panel in Figure~~\ref{fig:1}. There are two
-types of external currents, constants and pulses. In the case of pulses
-the duration of each pulse is given in seconds along with its intensity.
-}\tabularnewline
-\toprule
-Figure & Type & \(I_e/C (V/s)\)\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Figure & Type & \(I_e/C (V/s)\)\tabularnewline
-\midrule
-\endhead
-1A & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(1.5\)\tabularnewline
-1B & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(1 + 10^{-6}\)\tabularnewline
-1C & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(2\)\tabularnewline
-1D & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(1.5\)\tabularnewline
-1E & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(1.5 (0.1 \mathrm{s}),\, 0 (0.5\mathrm{s}),\, 0.5 (0.1\mathrm{s}),\, 1 (0.1\mathrm{s}),\, 1.5 (0.1\mathrm{s}),\, 0 (0.1 \mathrm{s}) \)\tabularnewline
-1F & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(1.5 (0.02 \mathrm{s}),\, 0 (0.18 \mathrm{s}),\, -1.5 (0.025 \mathrm{s}),\, 0 (0.025 \mathrm{s}),\, 1.5 (0.025 \mathrm{s}),\, 0 (0.125 \mathrm{s})\)\tabularnewline
-1G & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(0 (0.05 \mathrm{s}),\, -3.5 (0.756 \mathrm{s}),\, 0 (0.194 \mathrm{s}) \)\tabularnewline
-1H & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\( 2(1 + 10^{-6}) \)\tabularnewline
-1I & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\( 1.5 (0.02 \mathrm{s}),\, 0 (0.01 \mathrm{s}),\, 1.5 (0.02 \mathrm{s}),\, 0 (0.25 \mathrm{s}),\, 1.5 (0.02 \mathrm{s}),\, 0 (0.02 \mathrm{s})\)\tabularnewline
-& & \( 1.5 (0.02 \mathrm{s}),\, 0 (0.04 \mathrm{s}) \)\tabularnewline
-1J & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\( 1.5 (0.1 \mathrm{s}),\, 1.7 (0.4 \mathrm{s}),\, 1.5 (0.1 \mathrm{s}),\, 1.7 (0.4 \mathrm{s}) \)\tabularnewline
-1K & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(-1\)\tabularnewline
-1L & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(-1\)\tabularnewline
-1M & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(2\)\tabularnewline
-1N & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(1.5\)\tabularnewline
-1O & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(0 (0.1 \mathrm{s}),\, -3.5 (0.5 \mathrm{s}),\, 0 (0.4 \mathrm{s})\)\tabularnewline
-1P & \includegraphics[width=0.10000\textwidth]{figs/const.pdf} &
-\(2\)\tabularnewline
-1Q & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(2 (0.015 \mathrm{s}),\, 0 (0.185 \mathrm{s})\)\tabularnewline
-1R & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(5 (0.01 \mathrm{s}),\, 0 (0.09 \mathrm{s}),\, 5 (0.01 \mathrm{s}),\, 0 (0.09 \mathrm{s})\)\tabularnewline
-1S & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(5 (0.005 \mathrm{s}),\, 0 (0.005 \mathrm{s}),\, 4 (0.005 \mathrm{s}),\, 0 (0.385 \mathrm{s}),\, 5 (0.005 \mathrm{s}),\, 0 (0.045 \mathrm{s}) \)\tabularnewline
-& & \(4 (0.005\mathrm{s}),\, 0 (0.345 \mathrm{s}) \)\tabularnewline
-1T & \includegraphics[width=0.10000\textwidth]{figs/pulse.pdf} &
-\(8 (0.002 {\mathrm{s}}),\, 0 (0.048 \mathrm{s})\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{tbl:6}{}
-\begin{longtable}[]{@{}ll@{}}
-\caption{\label{tbl:6}\textbf{Initial Conditions} In all simulations the
-same initial conditions have been used, except from the one illustrated
-in Figure ~\ref{fig:1} H. }\tabularnewline
-\toprule
-Variable & Initial Value\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Variable & Initial Value\tabularnewline
-\midrule
-\endhead
-\(V(t)\) & \(-0.07\, \mathrm{V}\) / \(-0.03\, \mathrm{V}\) (Figure
-1H)\tabularnewline
-\(\Theta(t)\) & \(-0.05\, \mathrm{V}\) / \(-0.03\, \mathrm{V}\) (Figure
-1H)\tabularnewline
-\(I_1(t)\) & \(0.01\, \mathrm{V}\)\tabularnewline
-\(I_2(t)\) & \(0.001\, \mathrm{V}\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\section{Results}\label{results}
-
-All three figures from the original article have been successfully
-replicated. All the different spiking behaviors of the model are
-illustrated in Figure~~\ref{fig:1}, where the black solid line indicates
-the membrane potential (\(V(t)\)), the red dashed line illustrates the
-instantaneous threshold potentials (\(\Theta(t)\)), and the gray line
-shows the input to the neuron (\(I_e/C\)). The \(x\)-axis scales in all
-panels are exactly the same as in the original paper (indicating the
-total simulation time (\(t_f\)), while the \(y\)-axis scale differs from
-the one in the original paper. In this work the \(y\)-axis scale is the
-same for all the subplots (\([-95, -25]\, {\mathrm{mV}}\)), except for
-panels G and O (\([-145, -25]\, {\mathrm{mV}}\)).
-
-Figures~~\ref{fig:2} and~~\ref{fig:3} depict the phase space of the
-phasic spiking (\(V(t)\) and \(\Theta(t)\)) and phasic bursting
-(\(V(t)\), \(I_1(t)\), and \(I_2(t)\)). In both figures the blue curves
-and the black dots indicate the trajectory of the system and spike
-events, respectively. In Figure~~\ref{fig:2} the gray arrows show the
-evolution of the system (vector field of the system).
-Figure~~\ref{fig:3} has a different orientation from the original one
-but both the original figure and the replicated one illustrate the same
-trajectories and spike events of the system. All the figures express the
-same qualitative behavior as the original figures
-in~\autocite{mihalas:2009}.
+A reference to table \textcite{tbl:table}. A reference to figure
+\textcite{fig:logo}. A reference to equation \textcite{eq:1}. A
+reference to citation \textcite{markdown}.
 
@@
 \centering
-\includegraphics{figs/Figure01.pdf}
-\caption{\textbf{Neural responses of MNN.} Black solid lines indicate
-the membrane potential (\(V(t)\)), the red dashed lines show the
-threshold potentials (\(\Theta(t)\)), and the gray lines the external
-currents applied on each case. {\textbf{A}} tonic spiking, {\textbf{B}}
-class \(1\), {\textbf{C}} spike frequency adaptation, {\textbf{D}}
-phasic spiking, {\textbf{E}} accommodation, {\textbf{F}} threshold
-variability, {\textbf{G}} rebound spike, {\textbf{H}} class \(2\),
-{\textbf{I}} integrator, {\textbf{J}} input bistability, {\textbf{K}}
-hyperpolarization induced spiking, {\textbf{L}} hyperpolarization
-induced bursting, {\textbf{M}} tonic bursting, {\textbf{N}} phasic
-bursting, {\textbf{O}} rebound burst, {\textbf{P}} mixed mode,
-{\textbf{Q}} afterpotentials, {\textbf{R}} basal bistability,
-{\textbf{S}} preferred frequency, {\textbf{T}} spike
-latency.}\label{fig:1}
+\includegraphics{rescience-logo.pdf}
+\caption{Figure caption}\label{fig:logo}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{figs/Figure02.pdf}
-\caption{\textbf{Phase space of phasic spiking.} Blue solid lines
-indicate the trajectories of the model in the phase spiking behavior
-(Figure~~\ref{fig:1} D). The dashed line corresponds to
-\(V(t) =  \Theta(t)\), and the black dots represent spike events. The
-parameters for this simulation are the same as in Figure~~\ref{fig:1}
-D.}\label{fig:2}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{figs/Figure03.pdf}
-\caption{\textbf{Phase space of phasic bursting.} Blue solid lines
-represent the trajectories of the system and the black dots indicate
-spiking events. The parameters for this simulation are the same as in
-Figure~~\ref{fig:1} N.}\label{fig:3}
-\end{figure}
-
-\section{Conclusion}\label{conclusion}
-
-All figures in~\textcite{mihalas:2009} have been successfully replicated
-with high fidelity. Overall, the whole reproducing process was smooth
-and without obscure points since most of the parameters are provided in
-the original article. Only the time intervals for which the external
-current is applied to the model and the initial conditions are not
-provided explicitly. Therefore, we had to extract that information from
-figure \(1\) of the original article. To conclude, the
-article~\autocite{mihalas:2009} has been successfully reproduced without
-any discrepancy.
+\[ A = \sqrt{\frac{B}{C}} \] \{\#eq:1\}
 

### 2017-10-16 code/WY94.jl

@@ -3,31 +3,19 @@ using StatsBase
 
 pyplot()
 
-function habitat_preference(N,K;trials=20)
-    p = rand(3)
-    for iteration in 1:trials
-        l, m, n = sample(1:3, 3, replace=false)
-        Kl1, Kl2 = K[l,1], K[l,2]
-        Nl, Nm, Nn = N[[l,m,n]]
-        pl, pm, pn = p[[l,m,n]]
-        initdiff = 5000.0
-        for test_p in linspace(0.0,1.0,2000)
-            pl = test_p
-            attempt_p = [test_p, pm, pn]
-            attempt_N = [Nl, Nm, Nn]
-            diff_1 = (attempt_p.*attempt_N)./Kl1
-            diff_2 = ((1.-attempt_p).*attempt_N)./Kl2
-            diff = abs(sum(diff_1)-sum(diff_2))
-            if diff < initdiff
-                initdiff = diff
-                p[l] = pl
-            end
-        end
-    end
-    return p
+function habitat_preference(N,K;trials=50)
+   p = [1.0, 0.0, rand()] # Because why not?
+   TN = sum(N)
+   for iteration in 1:trials
+      l, m, n = sample(1:3, 3, replace=false)
+      K1, K2 = K[l,1], K[l,2]
+      NUM = (K1+K2)*N[n]*p[n] + (K1+K2)*N[m]*p[m] - K1*N[l] - K1*N[m] - K1*N[n]
+      DEN = (K1+K2)*N[l]
+      RATIO = minimum([maximum([-(NUM/DEN), 0.0]), 1.0])
+      p[l] = DEN > 0.0 ? RATIO : 0.0
+   end
+   return p
 end
-
-
 
 function generate_K(k1, k2, a, b, sync)
    rnd1 = rand()
@@ -61,4 +49,10 @@ function WY94(N1, N2, N3; a=0.1, b=0.9, K1=[200,200], K2=[100,100], T=100, g=0.0
       output[(i+1),5:end] = p
    end
    return output
+end
+
+function shannon(n1, n2, n3)
+   t = n1 .+ n2 .+ n3
+   p1, p2, p3 = n1./t, n2./t, n3./t
+   return -(p1 .* log.(p1) .+ p2 .* log.(p2) .+ p3 .* log.(p3))./log(3)
 end

### 2017-10-16 code/figure02.jl

@@ -11,7 +11,7 @@ xaxis!(p1p, "Time")
 
 p2d = plot(o2[:,1], o2[:,4], m=:circle, c=:black, leg=false)
 plot!(p2d, o2[:,1], o2[:,2], m=:utriangle, c=:black, leg=false)
-plot!(p2d, o2[:,1], o2[:,3], m=:ltriangle, c=:black, leg=false)
+plot!(p2d, o2[:,1], o2[:,3], m=:dtriangle, c=:black, leg=false)
 
 p2p = plot(o2[:,1], o2[:,end], m=:circle, c=:black, leg=false)
 plot!(p2p, o2[:,1], o2[:,5], m=:utriangle, c=:black, leg=false)

### 2017-10-16 code/figureSurf.jl

@@ -1,0 +1,26 @@
+include("WY94.jl")
+
+gs = 25
+
+b = linspace(0.3, 0.9, gs)
+kvar = linspace(0.0, 100.0, gs)
+
+o = zeros(Float64, (gs, gs))
+
+for i in eachindex(b)
+    for j in eachindex(kvar)
+        hstore = zeros(10)
+        k = [100.0, 100.0].+[-kvar[j], kvar[j]]
+        for re in 1:length(hstore)
+            s = WY94(3,3,3; K1=k, K2=k, b=b[i])
+            hstore[re] = mean(shannon(s[:,2],s[:,3],s[:,4]))
+        end
+        o[i,j] = mean(hstore)
+    end
+end
+
+plot(2.0.*kvar, b, o, lt=:surface, c=:viridis, leg=false)
+xaxis!("Range of variation")
+yaxis!("Ability of generalist")
+
+savefig("figureSurf.pdf")

### 2017-10-16 article/Poisot-2017.tex
PDF compiled

@@
 \usepackage{fontawesome}
+\usepackage{amssymb,amsmath}
 \usepackage{caption}
@@
   K_1 & aK_1 \\
-  a_K2 & K_2 \\
+  aK_2 & K_2 \\
   bK_1 & b_K2
@@
 \begin{equation}
-D_{l1} = \frac{\sum_{i\in l,m,m}p_iN_i}{K_{l1}}
+D_{l1} = \frac{\sum_{i\in l,m,n}p_iN_i}{K_{l1}}
 \end{equation}
@@
 \begin{equation}
-D_{l2} = \frac{\sum_{i\in l,m,m}(1-p_i)N_i}{K_{l1}} \,.
+D_{l2} = \frac{\sum_{i\in l,m,n}(1-p_i)N_i}{K_{l1}} \,.
 \end{equation}
@@
 
+\hypertarget{tbl:prm}{}
+\begin{longtable}[]{@{}llr@{}}
+\caption{\label{tbl:prm}Default parameters. Unless otherwise specified,
+these parameters have been used for all figures. }\tabularnewline
+\toprule
+Parameter & Meaning & Default value\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Parameter & Meaning & Default value\tabularnewline
+\midrule
+\endhead
+\(r\) & \emph{per capita} growth rate & \(1.3\)\tabularnewline
+\(a\) & fitness cost of specialism & \(0.1\)\tabularnewline
+\(b\) & fitness cost of generalism & \(0.9\)\tabularnewline
+\(K_1\) & range of habitat quality (1) & \([200,200]\)\tabularnewline
+\(K_2\) & range of habitat quality (1) & \([100,100]\)\tabularnewline
+\(g\) & proportion picking a random habitat & \(0.0\)\tabularnewline
+\bottomrule
+\end{longtable}
+
 \section{Results}\label{results}
@@
 provide some of these values, and they were used to test this
-implementation. Figure 2A and 2B in the original manuscript provide a
-good diagnostic value, as they are based on non-stochastic situations.
-In \textcite{fig:02},
+implementation. All default parameter values are given in
+~\ref{tbl:prm}. In all figures, circles represent the generalist, upper
+and lower triangles represent the specialists of habitats 1 and 2
+(respectively), and a diamond is used when both specialists have the
+same behaviour.
+
+Figure 2A and 2B in the original manuscript provide a good diagnostic
+value, as they are based on non-stochastic situations. These are
+reproduced in ~\ref{fig:02}. As in the original simulations, the
+generalist alone overshoots its carrying capacity at \(t=5\), and the
+stabilizes at a value of \(N_3^\star = 270\). Note also that
+\(p_3 = 2/3\) , which represents the ratio of habitat qualities. When
+adding specialists at initial densities of \(N_1 = N_2 = 1\), we observe
+(i) that the preference of the generalist shifts towards habitat 1 to
+avoid competition in the smallest habitat, and (ii) the density of the
+specialist of habitat 2 stabilizes as soon as the generalist abandons
+habitat 2 (at \(t \approx 35\)).
 
@@
 
-Results should be compared with original results and you have to explain
-why you think they are the same or why they may differ (qualitative
-result vs quantitative result). Note that it is not necessary to redo
-all the original analysis of the results.
+In ~\ref{fig:03}, this implementation gives the same qualitative results
+as the original article. There are small quantitative differences in
+output, which are in my opinion explained by the initial population
+densities (which are not given in the original article). Absent
+variation in habitat quality, generalists cannot persist, and the
+specialists both reach \(N^\star = \bar K\). There is a value of
+variation for which the densities of specialists and generalists are
+equal (about 110 for tied variations, and 130 for independent
+variations).
+
+\begin{figure}
+\centering
+\includegraphics{figure03.pdf}
+\caption{Consequences of increasing the range of variation of both
+\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
+solid lines tied variations, as in the original article.}\label{fig:03}
+\end{figure}
+
+In ~\ref{fig:04}, the same qualitative result (generalists increase in
+abundance when one habitat becomes larger). The quantitative differences
+can, again, be explained by changes in the initial population densities.
+Finally, the results in ~\ref{fig:05} and ~\ref{fig:06} can also be
+replicated.
+
+\begin{figure}
+\centering
+\includegraphics{figure04.pdf}
+\caption{Consequences of increasing the range of variation of both
+\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
+solid lines tied variations, as in the original article.}\label{fig:04}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figure05.pdf}
+\caption{Consequences of increasing the range of variation of both
+\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
+solid lines tied variations, as in the original article.}\label{fig:05}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figure06.pdf}
+\caption{Consequences of increasing the range of variation of both
+\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
+solid lines tied variations, as in the original article.}\label{fig:06}
+\end{figure}
 
@@
 
-Conclusion, at the very minimum, should indicate very clearly if you
-were able to replicate original results. If it was not possible but you
-found the reason why (error in the original results), you should exlain
-it.
-
-\begin{longtable}[]{@{}llllll@{}}
-\caption{Table caption \{\#tbl:table\}}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Heading 1\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Heading 2\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Heading 1\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-Heading 2\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright\strut
-\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
-cell6 row 1\tabularnewline
-cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
-cell6 row 2\tabularnewline
-cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
-cell6 row 3\tabularnewline
-\bottomrule
-\end{longtable}
-
-A reference to table \textcite{tbl:table}. A reference to figure
-\textcite{fig:logo}. A reference to equation \textcite{eq:1}. A
-reference to citation \textcite{markdown}.
+I believe that this implementation faithfully reproduces the results of
+\textcite{wils94csg}. A single run of the model (using 100 generations
+as in the original article) completes in
+\(\approx 5\times 10^{-3} \text{s}\). This makes this implementation
+usable for teaching, as it is not time consuming to generate results. As
+an example, I merged the output of ~\ref{fig:03} and ~\ref{fig:05}, to
+generate the response of Shannon's evenness (which measures the extent
+to which all species have the same density) to changes in the rate of
+variation \emph{and} generalist ability. The results are presented in
+~\ref{fig:surf}.
 
@@
 \centering
-\includegraphics{rescience-logo.pdf}
-\caption{Figure caption}\label{fig:logo}
+\includegraphics{figureSurf.pdf}
+\caption{Response of the model to changing the ability of the generalist
+\emph{and} the range of variation. While the response to environmental
+change is bell-shaped when generalists are competitive (\(g\) close to
+unity), it becomes monotonously decreasing for generalists with lower
+competitive ability (\(g\) close to \(a\)).}\label{fig:surf}
 \end{figure}
-
-\[ A = \sqrt{\frac{B}{C}} \] \{\#eq:1\}
 

### 2017-10-26 article/Mondeel_Ogundipe_Westerhoff-2017.tex
initial article commit

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Thierry D.G.A. Mondeel,
      Vivian Ogundipe,
      Hans V. Westerhoff,
  },
            pdftitle=
{
[Re] Predicting metabolic biomarkers of human inborn errors of metabolism
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

\usepackage{amsmath,amssymb}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Predicting metabolic biomarkers of human inborn errors of metabolism}

    \usepackage{authblk}
                        \author[1,*]{Thierry D.G.A. Mondeel}
                    \author[1]{Vivian Ogundipe}
                    \author[1, 2, 3]{Hans V. Westerhoff}
                            \affil[1]{Synthetic Systems Biology and Nuclear Organization, Swammerdam Institute
for Life Sciences, University of Amsterdam, The Netherlands}
                    \affil[2]{Molecular Cell Physiology, Amsterdam Institute for Molecules, Medicines
and Systems, VU University Amsterdam, The Netherlands}
                    \affil[3]{Manchester Centre for Integrative Systems Biology, School of Chemical
Engineering and Analytical Science, The University of Manchester, UK}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}{d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] T. Shlomi, M.N. Cabili, E. Ruppin, Predicting metabolic biomarkers of
human inborn errors of metabolism, Mol. Syst. Biol. 5 (2009) 263.
doi:10.1038/msb.2009.22.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Shlomi et al. \autocite{Shlomi2009} have developed a way in which flux
variability analysis \autocite{Mahadevan2003} is used to predict
biomarkers for inborn errors of metabolism (IEMs). The approach focuses
on exchange reactions in a metabolic network that connect extracellular
metabolites that reside in a medium, which might correspond to the
blood, to a reservoir such as the urine. The method then determines the
ranges of exchange fluxes that are compatible with constraints set by
the network's topology, flux bounds on reactions and the requirement of
steady state for each metabolite. Shlomi et al. proposed that if any
such ranges shift convincingly when the network contains a mutation, the
concentration ranges of the corresponding metabolites in the reservoir
should shift in parallel and these metabolites would therefore be
biomarkers.

Shlomi et al. illustrated the method first on a virtual metabolic map
and further validated their technique on the reconstruction of the human
metabolic map Recon1 \autocite{Duarte2007} emperically with respect to
known biomarkers, but did not give any formal proof. For a set of 17
IEMs that affect amino-acid metabolism they found a 56\% fraction of the
biomarkers that are correctly predicted and a fraction of 76\% of
predicted biomarkers that were predicted correctly in terms of whether
they increased or decreased in concentration. Thiele et al.
\autocite{Thiele2013} later found an even higher precision using
essentially the same FVA methodology on an updated version of the
metabolic reconstruction.

The source code is not publicly available and the computation
environment used was not mentioned in the publication. The
\href{https://github.com/opencobra/cobratoolbox}{MATLAB COBRA toolbox}
contains code to reproduce the analysis by Thiele et al. but this was
also performed in MATLAB.

We here present an implementation of the biomarker prediction method
originally proposed by Shlomi et al., programmed in Python, and
reproduce the results presented in Figure 1 and 2 of the original
publication. These figures concern the method's application to a simple
metabolic map and then to a set of \(17\) amino acid metabolism
disorders. Our Python module provides the biomarker prediction algorithm
and is agnostic of the map used. It can therefore be applied to other
maps than the human metabolic reconstruction considered by Shlomi et al.
Jupyter notebooks are provided that generate the illustrative map and
reproduce all results referred to in the Figures 1 and 2 of Shlomi et
al. Our implementation is based on information given in the original
paper. However, some details critical to the reproduction were left
unexplained there and are highlighted here. We also include a more
detailed discussion of one of the amino acid IEMs to examine the
robustness of the predictions made using this approach.

\section{Methods}\label{methods}

\subsection{Overview}\label{overview}

We implemented the method of Shlomi et al. in Python. Our Python
function has several input parameters most of which are optional in
order to enable the user to perturb the network in various ways and to
try variations of the method that differ from the ones implemented here
or in Shlomi et al. The default settings are such that they mimic the
approach Shlomi et al. originally proposed. All flux-variability
analyses are performed using COBRApy \autocite{Ebrahim2013}.

All simulations discussed here were performed in Python 3.6 using
COBRApy \autocite{Ebrahim2013} (version 0.9) and Pandas (version
0.20.3). This repository also functions in conjunction with
\href{http://mybinder.org}{MyBinder} which allows for full
reproducibility in the `cloud' without any need for installation of
additional software.

The rest of this section introduces the approach proposed by Shlomi et
al. as we implemented it.

\subsection{Definitions}\label{definitions}

Exchange reactions are reactions in the metabolic network that allow
import and/or export of metabolites from the metabolic map. Exchange
reactions are represented by non-mass-balanced pseudoreactions, e.g.
\(X \leftrightarrow \varnothing\), where \(X\) is understood to
originate from or efflux to the biofluids. Positive flux indicates net
secretion of \(X\) by the cell and negative flux indicates the net
uptake of \(X\).

We define boundary metabolites to be metabolites that are involved in
exchange reactions with the environment, i.e.~they serve as inputs
and/or outputs for the metabolic network, e.g.~glucose and CO\(_2\).
These metabolites typically exist both inside and outside the cell.

Each boundary metabolite will be associated with the range of flux its
exchange reaction supports under the following constraints for the flux
distributions in the model: (i) network topology, (ii) mass-balance for
all metabolites, (iii) flux bounds of reactions in the network.
Optionally, there is an additional constraint of being optimal with
respect to attaining or maximizing a certain set of fluxes, the
``objective'', in the model.

Biomarkers are those boundary metabolites the range of exchange flux of
which differs between the wild-type (WT) and mutant simulations. Shlomi
et al. proposed a threshold of at least 10\% difference in the lower or
upper bounds of the wild-type vs.~the mutant flux intervals.

Each biomarker will be associated with a prediction for being either
elevated or decreased in the mutant case, as comparted to the wild-type.
A biomarker is considered to have an increased, or reduced,
extracellular concentration in the mutant case if, when plotting the
wild-type and mutant flux-variability (see below) intervals on a
horizontal axis, both the lower and upper boundary of the mutant
interval are shifted to the right, or left, respectively. If the two
borders of the mutant interval move in opposite directions with respect
to the wild-type interval the result is scored as `unchanged' and the
boundary metabolite is not considered to be a biomarker.

\subsection{Flux variability
calculations}\label{flux-variability-calculations}

Flux ranges for boundary metabolites are calculated by applying flux
variability analysis \autocite{Mahadevan2003}. Mathematically, flux
variability analysis for a specific exchange reaction \(v_i\) entails
the following linear programming problem:

\begin{align}
\text{Maximize } &\pm v_i, \text{ such that for all k} \nonumber \\
S v &= 0 \nonumber \\
Z  &\geq \phi Z_{\text{opt}} \nonumber \\
\alpha_k \leq & ~ v_k \leq \beta_k 
\end{align}

Here \(v\) is the column vector of fluxes representing all reactions in
the model, \(v_i\) is the flux of the exchange reaction of a given
boundary metabolite \(i\), \(\alpha_k\) is the, possibly negative, lower
bound for reaction \(k\) and similarly \(\beta_k\) is the, possibly
negative, upper bound for reaction \(k\). \(Z = c^T ~ v\) is the
objective function defined for the map, entailing a linear combination,
defined by column vector c, of one or more fluxes, and
\(Z_{\text{max}}\) its maximal value. We define \(0 \leq \phi \leq 1\)
to be an arbitrary minimal fraction of the maximal objective function
value that has to be achieved. When \(\phi=1\) we ask for the range of
flux allowed for reaction \(v_i\) while maintaining the maximal value of
the linear combination of objective fluxes. When setting \(\phi = 0\),
one is asking for the allowable flux range through reaction \(v_i\)
regardless of the value of the objective function (although the sign has
to be maintained). The latter, i.e. \(\phi=0\), is the case considered
in the method by Shlomi et al.

We here entertain the inborn errors of metabolism (IEM) due to
loss-of-function mutations in a single enzyme \(r\) that may catalyse
multiple chemical reactions. For IEMs that disrupt the activity of
several reactions Shlomi et al. proposed applying the approach to each
reaction individually and then combining the list of biomarkers.

Therefore, for a given IEM we loop the following sequence of steps for
each affected reaction \(r\) and for each boundary metabolite \(m\):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compute the exchange flux interval of \(m\), when \(r\) is forced to
  be active with a flux of at least \(\epsilon \geq 0\). By default
  \(\epsilon = 1\). We denote this the forward wild-type interval
  \(WT_{r,m}^+\).
\item
  For reversible reactions \(r\), if the forward interval yields a flux
  range of \([0,0]\) constrain the flux to be negative and lower than
  \(-\epsilon \leq 0\), and compute the backward wild-type interval
  \(WT_{r,m}^-\).
\item
  The wild-type interval \(WT_{r,m}\) is given by the union of the
  forward and backward intervals.\\
\item
  Compute the exchange flux interval of \(m\) when \(r\) is forced to be
  inactive by temporarily setting the lower and upper bound of \(r\)
  equal to zero. We denote this interval \(M_{r,m}\).
\item
  Compare the wild-type and mutant intervals and predict whether the
  external concentration of \(m\) is increased, decreased or unaltered
  when reaction \(r\) is inactivated. This is done by the following
  rule: if both boundaries are higher in the mutant than the
  corresponding boundaries in the wild-type, then \(m\) is predicted to
  be a biomarker and elevated. If both boundaries are lower in mutant as
  compared to wild type then the metabolite is predicted to be a
  biomarker and reduced. If the predicted changes are smaller than 10\%
  \(m\) is a low confidence biomarker. When the wild-type and mutant
  intervals are disjoint \(m\) is a highly confident (H.C) biomarker.
\end{enumerate}

After the loop is completed, contradictory predictions for \(m\) between
the affected reactions \(r\) are dealt with based on a majority rule.
When there is an equal number of elevated and reduced predictions the
boundary metabolite is considered to be unchanged, i.e.~not a biomarker.

\section{Results}\label{results}

\subsection{Robustness of the flux variability
method}\label{robustness-of-the-flux-variability-method}

Shlomi et al. did not discuss in detail the effects of changing various
parameters settings of, and assumptions inherent to, their FVA-based
method, nor did they present a proof of any generality. Because our
Python implementation allows the user to set various parameters it
becomes rather easy to illustrate how specific settings in the approach
proposed by Shlomi et al. affect the outcome of the simulations. An
example of the potential pitfalls and issues with this method is
considered below and looks at predictions for the inborn error of
metabolism known as Phenylketonuria. We here define this disease by one
of its potential causes, i.e.~dysfunctionality of the enzyme converting
phenylalanine into tyrosine. PKU patients are known for elevated levels
of phenylalanine and decreased levels of tyrosine in their serum. We
illustrate this briefly in an example also considered in Figure 2 of
Shlomi et al. In Figure ~\ref{fig:PKU_network} we sketched some
essential features of the metabolic network surrounding the
phenylalanine hydroxylase (PAH) enzyme that is malfunctioning in PKU.
Unless stated otherwise, the simulations are performed for a medium that
allows efflux through all exchange reactions with an upper bound of 1000
and influx through all exchange reactions with a lower bound of 1 and we
use the default settings, including a 10\% significance threshold). The
only considered biomarkers are the \(20\) proteinogenic amino acids. In
total we will consider six cases. The results are summarized in Table
~\ref{tbl:PKU_results}.

\begin{figure}
\centering
\includegraphics{./Figures/PKU_network.pdf}
\caption{Scheme of the network surrounding phenylalanine hydroxylase
(PAH) and its cofactors. The 4a-hydroxytetrahydrobiopterin produced by
PAH may spontaneously or when catalysed by PCBD1 (EC 4.2.1.96) dehydrate
to dihydrobiopterin which may be enzymatically reduced back by NADPH +
\(H^+\) to tetrahydrobiopterin through DHPR (EC 1.5.1.34). The solid
black lines indicate reactions that are contained in both Recon 1 and
Recon 2.2 \autocite{Swainston2016}. The solid green line refers to the
alternative PAH reaction that includes the spontaneous dehydradation
towards dihydrobiopterin that is only present in Recon 2.2. Dashed lines
indicate sources of the synthesis pathways or terminal products of the
indicated metabolites. The tyrosine transaminase (TYRTA) uses
oxoglutarate (AKG) and glutamate (Glu) as
cosubstrates.}\label{fig:PKU_network}
\end{figure}

Perhaps the most fundamental question is: can changes in network
topology result in different predictions? The answer is yes. One
illustration of this is the comparison of predictions between Recon 1
and Recon 2. As Table ~\ref{tbl:PKU_results} indicates, using the same
medium and algorithm settings for both Recon 1 and Recon 2.2 the two
networks led to different predictions of the effect of a defect of
phenylalanine hydroxylase on tyrosine. The application of the flux
variability method to Recon 1 correctly predicted both tyrosine and
phenylalanine as biomarkers and gets the directionality of the change in
their concentrations right. In contrast, the prediction based on Recon
2.2 lacks tyrosine as a biomarker.

With ever improving network annotations it becomes feasible to simulate
the gene knockout directly for most single gene defects leading to
inborn errors of metabolism. Shlomi et al. directly blocked or activated
reactions associated with a causal gene for a specific IEM. For enzymes
partaking in several reactions Shlomi et al. proposed activating and
blocking all reactions associated with the gene individually and then
taking the union of their predicted biomarkers. When a biomarker was
subject to qualitatively different predictions between the
enzyme-catalyzed reactions encoded by the gene, a majority rule was
applied. We will refer to this as the asynchronous setting. Blocking all
affected reactions simultaneously could yield different predictions. We
will refer to this as the synchronous setting. We added the option to
choose between the asynchronous and synchronous setting to our
implementation.

For example, consider a gene that encodes an enzyme that catalyzes two
chemically different reactions, \(R_1\) and \(R_2\). Using the
asynchronous setting we would run the algorithm once for \(R_1\) and
once for \(R_2\) and take the union of all predicted biomarkers.
Inconsistent predictions between \(R_1\) and \(R_2\) are dealt with
based on a majority rule (see Methods). In contrast, with the
synchronous setting, we would run the algorithm once. For the wild-type
we would simultaneously force flux through both reactions and in the
mutant we would block both reactions. The synchronous approach simulates
the complete catalytic inactivity or absence of the enzyme and therefore
the infeasibility of any reactions it catalyzes. This corresponds to a
deletion of the gene. The asynchronous approach might correspond to a
point mutation that changes the specificity of the enzyme.

\hypertarget{tbl:PKU_results}{}
\begin{longtable}[]{@{}lllllll@{}}
\caption{\label{tbl:PKU_results}Summary of predictions made using the
FVA-based biomarker prediction method for the inborn-error of metabolism
phenylketonuria for two networks (\(R_1\) = Recon 1 and \(R_2\) = Recon
2), with or without the synchronous setting (indicated by ``sync'' here)
and with high (maximum influx of 20) or low (no influx)
hydroxyphenylpyruvate (hpp +/-) in the medium. \(^*\) refers to a
situation without any tyrosine production in the mutant. \(^{**}\)
refers to the situation where tyrosine reduction is so small that it
does not reach the 10\% threshold. `-' indicates no prediction because
the boundaries did not change from wild-type to mutant. \(\epsilon = 0\)
refers to the case where we do not force any flux through the affected
reaction(s) in the wild-type. In all other columns, \(\epsilon = 1\).
The full output with interval bounds that the method generated can be
found in the relevant notebook in this repository. }\tabularnewline
\toprule
Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
\midrule
\endfirsthead
\toprule
Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
\midrule
\endhead
Phenylalanine & Elevated & Elevated & Elevated & Elevated & Elevated &
-\tabularnewline
Tyrosine & Reduced & - & Reduced & Reduced\(^*\) & Reduced\(^{**}\) &
Reduced\tabularnewline
\bottomrule
\end{longtable}

Returning to the example of phenylketonuria we simulated this for Recon
2.2, using the synchronous approach. Column 3 in Table 1, shows that the
result now agrees with the Recon1 asynchronous predictions. Using Recon
1 we came to the correct prediction because it contains only one
reaction (`PHETHPTOX2') linked to the PKU gene, i.e.~only phenylalanine
hydroxylase (Entrez ID: 5053, HGNC:8582):

\begin{align*}
&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
&\leftrightarrow \\
&\text{4a-hydroxy-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-tyrosine}
\end{align*}

In Recon1, the 4a-hydroxytetrahydrobiopterin produced may dehydrate to
6,7-dihydrobiopterin by the action of the `THBPT4ACAMDASE' reaction
catalyzed by the PCBD1 enzyme (EC 4.2.1.96). Recon 2.2 contains, in
addition to the PAH and PCBD1 catalyzed reactions indicated above, the
overall reaction where the spontaneous dehydration reaction is included,
again attributed to phenylalanine hydroxylase:

\begin{align*}
&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
&\leftrightarrow \\
&\text{6,7-dihydrobiopterin} + \text{L-tyrosine} + \text{H}_2\text{O}.
\end{align*}

This details are included in the pathway summary in our Figure
~\ref{fig:PKU_network}. When the reactions affected by the PKU gene are
knocked out simultaneously, Recon2 correctly predicts both tyrosine and
phenylalanine as biomarkers. When blocking only one reaction, there is a
way to produce tyrosine. Consequently, tyrosine is not predicted as
biomarker if Recon 2 is used as metabolic map.

Another significant modulation point for the method, and for
flux-balance-based models in general, is the choice of medium. More
specifically, it matters for the biomarker prediction which other
exchange reactions are allowed to have a non-zero flux and what the
bounds on these fluxes are. Changes herein leave the internal network
topology unchanged and instead alter how the model is allowed to
interface with biological compartments or environments that are not
explicitly modeled. Potentially this potentially has large consequences
as it may open up or shut down entire parts of the metabolic network
that rely on specific substrates. It additionally interacts with the
setting of the threshold for considering biomarkers significant since
the medium components together with the network pathways (and their
V\(_{\text{max}}\)'s) determine the amount of a given metabolite that
may be produced.

In order to examine these ramifications we continue with the synchronous
PKU example and note that tyrosine production flux goes down in the
mutant, but not to zero. This is due to an alternative tyrosine
synthesis pathway in the model originating from 4-Hydroxyphenylpyruvate
in the medium through the `TYRTA' reaction (tyrosine transaminase E.C.
2.6.1.5). This transaminase is considered to be reversible in Recon 2
and Equilibrator \autocite{Noor2012} annotates this reaction with a
\(K'_{\text{eq}} \approx 1\). It should be mentioned though that the
ability for transport of hydroxyphenylpyruvate is speculative. Column 4
and 5 in Table 1 show results when using the standard medium and a
medium without or with increased (i.e.~influx at -20)
4-Hydroxyphenylpyruvate influx respectively. In FBA and FVA changes in
medium concentrations of exchange metabolites are modelled by changing
the corresponding influx (inward exchange) V\(_{\text{max}}\) (bound).
In the former the ability to produce tyrosine indeed disappears in the
mutant case. In the latter there is so much left-over capacity that the
change in interval bounds slips below the 10\% cutoff. This sensitivity
to medium composition appropriately reflects the phenomenon that cells
behave differently in different culture media, and that organisms behave
differently depending on their nutrition. Our discussion here may
highlight that this feature should be expected to affect the pertinence
of biomarkers. Biomarkers may have a tendency to be non-robust,
dependent as they can be on network topology details and nutrition.

Finally, we consider the value of the \(\epsilon\) parameter, which
indicates the amount of flux forced through the reaction in the
wild-type condition. The last column in Table ~\ref{tbl:PKU_results}
highlights that phenylalanine does not to show up as a biomarker if
setting \(\epsilon = 0\) and performing the PKU simulation on Recon 1.
This may also be the reason why Sahoo et al. did not manage to find
phenylalanine as a biomarker for PKU in their study
\autocite{Sahoo2012}.

\subsection{\texorpdfstring{Reproduction of Figure 1 in
\autocite{Shlomi2009}}{Reproduction of Figure 1 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-1-in-shlomi2009}

\begin{figure}
\centering
\includegraphics{Figures/scheme_shlomi2009_figure1}
\caption{A metabolic network used to illustrate the predictions of
biomarkers of IEMs as originally considered in the work by Shlomi et al.
for their prediction of biomarkers of IEMs. All metabolites but M3 have
an exchange reaction allowing for uptake or secretion of the
metabolites. M2 synthesis is fully dependent on uptake of M1 and M5 or
M6. The IEM considered is one that leads to deletion of the enzyme
converting M1 plus M3 to M2 plus M4.}\label{fig:Shlomi2009_figure_1}
\end{figure}

The original publication's Figure 1 exemplified the biomarker prediction
method for an illustrative network. It is reproduced here as Figure
~\ref{fig:Shlomi2009_figure_1}. The simple nature of the example serves
to explain the reasoning behind the method and additionally helped us
validate our implementation of the method. This repository provides an
SBML file of the network, the Jupyter notebook that generates it and a
Jupyter notebook reproducing the biomarker prediction.

The network in Figure ~\ref{fig:Shlomi2009_figure_1} consists of seven
metabolites, \(6\) of which (i.e.~all except M3) have exchange
reactions. We consider a hypothetical deletion of the enzyme catalyzing
the conversion of M1 into M2 coupled to the conversion of M3 into M4.
Shlomi et al. provide the a biomarker prediction for this network
graphically in their Figure 1. We here do so numerically and a Jupyter
notebook reproducing the biomarker prediction accompanies this paper. In
our analysis all exchange reactions were given inward bounds of \(-10\)
and outward bounds of \(+1000\) so that the internal reaction bounds (of
\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
shows that all exchangeable metabolites except M5 and M7 are predicted
to be robust biomarkers.

The exchange interval we found for metabolite M6 shows that in the
wild-type case M6 can be either taken-up from biofluids or secreted. The
upper bound is lower in absolute value than the lower bound because in
the wild-type at least \(1\) unit of flux is needed to provide M3 as
substrate for the enzyme under investigation since the enzyme convert M1
and M3 to M2 and M4 is required to have a minimal flux of \(1\) unit. In
the disease case, M6 (synthesized through M5) can only be secreted to
biofluids. These results fully agree with the original publication.

\subsection{\texorpdfstring{Reproduction of Figure 2 in
\autocite{Shlomi2009}}{Reproduction of Figure 2 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-2-in-shlomi2009}

Shlomi et al. applied their method to the reconstructed human metabolic
map, examining a set of \(17\) IEMs affecting amino-acid metabolism.
Figure 2 in their publication compares predicted biomarkers with known
biomarkers from the OMIM database \autocite{McKusick2007}.

We reproduced the analysis for each IEM listed in Table
~\ref{tbl:amino_acid_iems} and ultimately found full agreement between
what was reported by Shlomi et al. and the results of our
implementation.

It is worthwhile to note, with an eye on the sensitivity of this
approach to the various settings, that we needed to set the medium
composition very specifically. Shlomi et al. did not report on their
chosen medium in their publication but we were able to reproduce their
results by setting all inward flux bounds to \(1\) and all outward flux
bounds to \(1000\).

We made use of the Excel file included as supplementary information in
\autocite{Shlomi2009} and included it in this repository as well. In
this Excel file the genes and reactions linked to a given IEM, are
detailed. We attempted to extract the genes for each IEM and through the
Recon 1 map deduce the coupled reactions but found out that this yielded
predictions different from the ones presented by Shlomi et al. In the
Excel file a list of reactions linked to the IEM is also provided.
However, these reactions are not necessarily the same reactions that are
linked in Recon 1 to the genes they provide. Assuming the reconstruction
of the metabolic network to be accurate this raises some doubts as to
the correctness of the results obtained by Shlomi et al. for those IEMs
where the listed reactions did not agree with the reconstructed map.

Additionally, we needed to manually add the information for
S-Adenosylhomocysteine hydrolase and Methionine adenosyltransferase
deficiency, since they were missing from the Excel file and to reduce
the tetrahydrobiopterin deficiency gene list to only consider QDPR
(quinoid dihydropteridine reductase). Lastly, Histidinemia was in the
Excel file associated to the reaction HISD whereas this should have
refered to HISDr to match the reaction ID in Recon 1. After these
changes, we recovered the results as originally presented and as
summarized here in Table ~\ref{tbl:amino_acid_iems}.

\hypertarget{tbl:toy_model_results}{}
\begin{longtable}[]{@{}lllc@{}}
\caption{\label{tbl:toy_model_results}Biomarker predictions using our
implementation with respect to the mutation affecting the conversion of
M1 plus M3 to M2 plus M4 in the network of Figure
~\ref{fig:Shlomi2009_figure_1}. H.C. stands for `highly confident'
indicating, as defined by Shlomi et al., i.e.~no overlap between the WT
and mutant intervals. Between brackets {[}{]} we report the lower and
upper bounds of the FVA intervals we obtained. }\tabularnewline
\toprule
ID & Prediction & WT & Mutant\tabularnewline
\midrule
\endfirsthead
\toprule
ID & Prediction & WT & Mutant\tabularnewline
\midrule
\endhead
M1 & H.C. Elevated & {[}-10.0, -1.0{]} & {[}0.0, 0.0{]}\tabularnewline
M2 & H.C. Reduced & {[}1.0, 10.0{]} & {[}0.0, 0.0{]}\tabularnewline
M4 & Reduced & {[}1.0, 20.0{]} & {[}0.0, 10.0{]}\tabularnewline
M6 & Elevated & {[}-10.0, 9.0{]} & {[}0.0, 10.0{]}\tabularnewline
M7 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
M5 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
\bottomrule
\end{longtable}

\section{Conclusion}\label{conclusion}

The results summarized in Figures 1 and 2 of \autocite{Shlomi2009} have
been successfully replicated with the software tools provided here.
Overall, the reproducing process was smooth since the method is
illustrated both on an illustrative example and on the human metabolic
reconstruction, resulting in a multitude of test cases for the
implementation.

The points of trouble we encountered were the lack of explicit
discussion on the set of fluxes allowed into and out of the model and a
set of discrepancies between the gene-reaction mappings listed in the
Supplementary Excel file and the actual gene-reaction mappings in the
metabolic network. However, trial-and-error and manual curation solved
these issues.

\hypertarget{tbl:amino_acid_iems}{}
\begin{longtable}[]{@{}llll@{}}
\caption{\label{tbl:amino_acid_iems}We used our implementation to
reproduce the biomarkers for the list of inborn-errors of metabolism
shown in Figure 2 in \autocite{Shlomi2009}. Using the Supplementary
Excel table of \autocite{Shlomi2009} linking IEMs to causative genes and
affected reactions we filtered out those that are listed here. The only
considered biomarkers are the \(20\) proteinogenic amino acids. We
report all biomarkers the change of which exceeds the 10\% threshold and
categorize them by their qualitative prediction: elevated or reduced
serum levels. }\tabularnewline
\toprule
IEM & Affected reactions & Elevated & Reduced\tabularnewline
\midrule
\endfirsthead
\toprule
IEM & Affected reactions & Elevated & Reduced\tabularnewline
\midrule
\endhead
S-Adenosylhomocysteine & SEAHCYSHYD, AHCi & L-Methionine &
L-Cysteine\tabularnewline
hydrolase & & &\tabularnewline
Alkaptonuria & HGNTOR & L-Tyrosine &\tabularnewline
Arginase deficiency & ARGN & &\tabularnewline
Cystinuria & CYSTSERex, & &\tabularnewline
& SERLYSNaex & &\tabularnewline
Lysinuric protein & SERLYSNaex & &\tabularnewline
intolerance & & &\tabularnewline
Glutamate formimino- & FTCD, GluForTx & L-Histidine &\tabularnewline
transferase deficiency & & &\tabularnewline
Histidinemia & HISDr & L-Histidine &\tabularnewline
Homocystinuria & CYSTS, MTHFR3, & L-Methionine &
L-Cysteine\tabularnewline
& METS & &\tabularnewline
Hyperprolinemia type I & PRO1xm, PROD2m & &\tabularnewline
maple syrup & OIVD1m, & L-Valine, &\tabularnewline
urine disease & OIVD2m, & L-Isoleucine &\tabularnewline
& OIVD3m & L-Leucine, &\tabularnewline
Methionine adenosyl- & SELMETAT, METAT & L-Methionine &
L-Cysteine\tabularnewline
transferase deficiency & & &\tabularnewline
methylmalonic acidemia & CBLATm, CBL2tm, & L-Isoleucine &\tabularnewline
& MMMm & &\tabularnewline
phenylketonuria & PHETHPTOX2 & L-Phenylalanine &
L-Tyrosine\tabularnewline
Tetrahydrobiopterin & DHPR & &\tabularnewline
deficiency & & &\tabularnewline
Tyrosinemia type 1 & FUMAC & L-Tyrosine &\tabularnewline
Tyrosinemia type III & PPOR, 34HPPOR & L-Tyrosine &\tabularnewline
Glycine encephalopathy & GCC2am, GCC2bim, & &\tabularnewline
& GCC2cm, GCCam, & &\tabularnewline
& GCCbim, GCCcm & &\tabularnewline
\bottomrule
\end{longtable}

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2017-10-26 notebook/findBiomarkers.py

@@ -1,0 +1,357 @@
+import operator  # for sorting dictionaries
+
+import cobra
+import pandas as pd
+from cobra import Metabolite, Reaction
+from tqdm import tqdm_notebook # progress bars
+
+pd.set_option('display.max_rows', 10000) # Show everything
+pd.set_option('display.max_colwidth', -1)
+pd.set_option('expand_frame_repr', False)
+
+def findBiomarkers(model,fvaRxns=[],mods=[],mode='',always_unite=False,synchronous=False,
+                eps=1.0,threshold=0.1,fracOpt=0,forceFlux=True,geneAssociationByKO=False):
+    ''' Implements the biomarker prediction algorithm proposed in (Shlomi et al., 2009). 
+    It returns a pandas dataframe listing predicted biomarkers with their WT and mutant FVA intervals. 
+    It considers as potential biomarkers those contained in the exchange reactions in fvaRxns, under
+    alteration of reactions/genes in mods.
+    We allow for various always_unite alterations through the parameters: eps,threshold,fracOpt,etc.
+
+    model:                  A COBRApy model object
+    fvaRxns:                A list with (a subset of) exchange reactions in M. FVA is performed on these reactions.
+    mods:                   A list of reactions or genes (IDs or objects) that will be altered. 
+                            E.g. the gene/reactions affected by the IEM.
+    mode:                   Two options: IEMrxn (closest to Schlomi et al. approach),
+                            IEMgene (same but with gene IDs to knockout).
+                            When left empty the algorithm deduces which.
+    always_unite            BOOLEAN. Indicates whether to take the union of the forward and
+                            backward healthy intervals (True) or only take it when the
+                            forward interval is [0,0] (False).
+    synchronous:            BOOLEAN. Whether, in the case of multiple modifiers, they should be knocked out simultaneously.
+    eps:                    The flux to force through the modifiers in the healthy case.
+    threshold:              The minimal change percentage in the interval bounds to be considered a biomarker
+    fracOpt:                Sets the minimal percentage of the optimum of the objective function to attain
+    forceFlux:              BOOLEAN. Whether or not to force flux through the reaction in the healthy case.
[…]

### 2017-10-26 article/Mondeel_Ogundipe_Westerhoff-2017.tex
License + manuscript edits + module versions

@@
 ranges of exchange fluxes that are compatible with constraints set by
-the network's topology, flux bounds on reactions and the requirement of
+the network's topology, flux bounds of reactions and the requirement of
 steady state for each metabolite. Shlomi et al. proposed that if any
@@
 Shlomi et al. illustrated the method first on a virtual metabolic map
-and further validated their technique on the reconstruction of the human
-metabolic map Recon1 \autocite{Duarte2007} emperically with respect to
-known biomarkers, but did not give any formal proof. For a set of 17
-IEMs that affect amino-acid metabolism they found a 56\% fraction of the
-biomarkers that are correctly predicted and a fraction of 76\% of
-predicted biomarkers that were predicted correctly in terms of whether
-they increased or decreased in concentration. Thiele et al.
+and further validated their technique using the reconstruction of the
+human metabolic map Recon1 \autocite{Duarte2007} empirically with
+respect to known biomarkers, but did not provide any formal proof. For a
+set of 17 IEMs that affect amino acid metabolism, they found a 56\%
+fraction of the biomarkers that are correctly predicted and a fraction
+of 76\% of predicted biomarkers that were predicted correctly in terms
+of whether they increased or decreased in concentration. Thiele et al.
 \autocite{Thiele2013} later found an even higher precision using
@@
 
-We here present an implementation of the biomarker prediction method
-originally proposed by Shlomi et al., programmed in Python, and
+Here we here present an implementation of the biomarker prediction
+method originally proposed by Shlomi et al., programmed in Python, and
 reproduce the results presented in Figure 1 and 2 of the original
@@
 Jupyter notebooks are provided that generate the illustrative map and
-reproduce all results referred to in the Figures 1 and 2 of Shlomi et
-al. Our implementation is based on information given in the original
-paper. However, some details critical to the reproduction were left
-unexplained there and are highlighted here. We also include a more
-detailed discussion of one of the amino acid IEMs to examine the
-robustness of the predictions made using this approach.
+reproduce all results referred to in Figures 1 and 2 of Shlomi et al.
+Our implementation is based on information given in the original paper.
+However, some details critical to the reproduction were left unexplained
+there and are highlighted here. We also include a more detailed
+discussion of one of the amino acid IEMs to examine the robustness of
+the predictions made using this approach.
 
@@
 We implemented the method of Shlomi et al. in Python. Our Python
-function has several input parameters most of which are optional in
+function has several input parameters, most of which are optional in
 order to enable the user to perturb the network in various ways and to
@@
 Each biomarker will be associated with a prediction for being either
-elevated or decreased in the mutant case, as comparted to the wild-type.
+elevated or decreased in the mutant case, as comparted to the wild type.
 A biomarker is considered to have an increased, or reduced,
 extracellular concentration in the mutant case if, when plotting the
-wild-type and mutant flux-variability (see below) intervals on a
-horizontal axis, both the lower and upper boundary of the mutant
-interval are shifted to the right, or left, respectively. If the two
-borders of the mutant interval move in opposite directions with respect
-to the wild-type interval the result is scored as `unchanged' and the
-boundary metabolite is not considered to be a biomarker.
+wild-type and mutant flux-variability intervals on a horizontal axis
+(see below), both the lower and upper boundary of the mutant interval
+are shifted to the right, or left, respectively. If the two borders of
+the mutant interval move in opposite directions with respect to the
+wild-type interval, the result is scored as `unchanged' and the boundary
+metabolite is not considered to be a biomarker.
 
@@
 
-We here entertain the inborn errors of metabolism (IEM) due to
+Here, we examine inborn errors of metabolism (IEMs) resulting from
 loss-of-function mutations in a single enzyme \(r\) that may catalyse
 multiple chemical reactions. For IEMs that disrupt the activity of
-several reactions Shlomi et al. proposed applying the approach to each
+several reactions, Shlomi et al. proposed applying the approach to each
 reaction individually and then combining the list of biomarkers.
@@
   be active with a flux of at least \(\epsilon \geq 0\). By default
-  \(\epsilon = 1\). We denote this the forward wild-type interval
+  \(\epsilon = 1\). We refer to this as the forward wild-type interval
   \(WT_{r,m}^+\).
@@
   For reversible reactions \(r\), if the forward interval yields a flux
-  range of \([0,0]\) constrain the flux to be negative and lower than
+  range of \([0,0]\), constrain the flux to be negative and lower than
   \(-\epsilon \leq 0\), and compute the backward wild-type interval
@@
 \item
-  The wild-type interval \(WT_{r,m}\) is given by the union of the
+  The wild-type interval \(WT_{r,m}\) is equal to the union of the
   forward and backward intervals.\\
@@
   Compute the exchange flux interval of \(m\) when \(r\) is forced to be
-  inactive by temporarily setting the lower and upper bound of \(r\)
-  equal to zero. We denote this interval \(M_{r,m}\).
+  inactive by temporarily setting the lower and upper bound of \(r\) to
+  zero. This interval is denoted by \(M_{r,m}\).
 \item
   Compare the wild-type and mutant intervals and predict whether the
-  external concentration of \(m\) is increased, decreased or unaltered
-  when reaction \(r\) is inactivated. This is done by the following
-  rule: if both boundaries are higher in the mutant than the
-  corresponding boundaries in the wild-type, then \(m\) is predicted to
-  be a biomarker and elevated. If both boundaries are lower in mutant as
-  compared to wild type then the metabolite is predicted to be a
-  biomarker and reduced. If the predicted changes are smaller than 10\%
+  external concentration of \(m\) increases, decreases or remains
+  unchanged when reaction \(r\) is forced to be inactive. This is done
+  by the following rule: if both boundaries are higher in the mutant
+  simulation than the corresponding boundaries in the wild-type
+  simulation, then \(m\) is predicted to be a biomarker and elevated. If
+  both boundaries are lower in the mutant simulation as compared to the
+  wild-type simulation, then the metabolite is predicted to be a
+  biomarker and reduced. If the predicted changes are smaller than 10\%,
   \(m\) is a low confidence biomarker. When the wild-type and mutant
-  intervals are disjoint \(m\) is a highly confident (H.C) biomarker.
+  intervals are disjoint intervals, \(m\) is a highly confident (H.C)
+  biomarker.
 \end{enumerate}
@@
 the affected reactions \(r\) are dealt with based on a majority rule.
-When there is an equal number of elevated and reduced predictions the
+When there is an equal number of elevated and reduced predictions, the
 boundary metabolite is considered to be unchanged, i.e.~not a biomarker.
@@
 Shlomi et al. did not discuss in detail the effects of changing various
-parameters settings of, and assumptions inherent to, their FVA-based
+parameter settings of, and assumptions inherent to, their FVA-based
 method, nor did they present a proof of any generality. Because our
-Python implementation allows the user to set various parameters it
-becomes rather easy to illustrate how specific settings in the approach
+Python implementation allows the user to set various parameters, it
+becomes possible to illustrate how specific settings in the approach
 proposed by Shlomi et al. affect the outcome of the simulations. An
 example of the potential pitfalls and issues with this method is
-considered below and looks at predictions for the inborn error of
-metabolism known as Phenylketonuria. We here define this disease by one
-of its potential causes, i.e.~dysfunctionality of the enzyme converting
-phenylalanine into tyrosine. PKU patients are known for elevated levels
-of phenylalanine and decreased levels of tyrosine in their serum. We
-illustrate this briefly in an example also considered in Figure 2 of
-Shlomi et al. In Figure ~\ref{fig:PKU_network} we sketched some
-essential features of the metabolic network surrounding the
-phenylalanine hydroxylase (PAH) enzyme that is malfunctioning in PKU.
-Unless stated otherwise, the simulations are performed for a medium that
-allows efflux through all exchange reactions with an upper bound of 1000
-and influx through all exchange reactions with a lower bound of 1 and we
-use the default settings, including a 10\% significance threshold). The
-only considered biomarkers are the \(20\) proteinogenic amino acids. In
-total we will consider six cases. The results are summarized in Table
+investigated below, in which we look at predictions for the inborn error
+of metabolism known as Phenylketonuria (PKU). Here, we simulate this
+disease by applying one of its potential causes, i.e.~dysfunctionality
+of the enzyme converting phenylalanine into tyrosine.
+
+PKU patients are known for elevated levels of phenylalanine and
+decreased levels of tyrosine in their serum. We illustrate this briefly
+in an example also considered in Figure 2 of Shlomi et al. In Figure
+~\ref{fig:PKU_network}, we sketched some essential features of the
+metabolic network surrounding the phenylalanine hydroxylase (PAH) enzyme
+that malfunctions in PKU. Unless stated otherwise, the simulations are
+performed with a medium that allows efflux through all exchange
+reactions with an upper bound of \(1000\) and influx through all
+exchange reactions with a lower bound of \(1\) and we use the default
+settings, including a 10\% significance threshold. The only considered
+biomarkers are the \(20\) proteinogenic amino acids. We will consider
+six cases in total. The results are summarized in Table
 ~\ref{tbl:PKU_results}.
@@
 and Recon 2. As Table ~\ref{tbl:PKU_results} indicates, using the same
-medium and algorithm settings for both Recon 1 and Recon 2.2 the two
+medium and algorithm settings for both Recon 1 and Recon 2.2, the two
 networks led to different predictions of the effect of a defect of
@@
 variability method to Recon 1 correctly predicted both tyrosine and
-phenylalanine as biomarkers and gets the directionality of the change in
-their concentrations right. In contrast, the prediction based on Recon
-2.2 lacks tyrosine as a biomarker.
+phenylalanine as biomarkers as well as the directionality of the change
+in their concentrations. In contrast, the prediction based on Recon 2.2
+fails to identify tyrosine as a biomarker.
 
-With ever improving network annotations it becomes feasible to simulate
+With ever-improving network annotations, it becomes feasible to simulate
 the gene knockout directly for most single gene defects leading to
@@
 reactions associated with a causal gene for a specific IEM. For enzymes
-partaking in several reactions Shlomi et al. proposed activating and
+partaking in several reactions, Shlomi et al. proposed activating and
 blocking all reactions associated with the gene individually and then
-taking the union of their predicted biomarkers. When a biomarker was
-subject to qualitatively different predictions between the
-enzyme-catalyzed reactions encoded by the gene, a majority rule was
-applied. We will refer to this as the asynchronous setting. Blocking all
-affected reactions simultaneously could yield different predictions. We
-will refer to this as the synchronous setting. We added the option to
-choose between the asynchronous and synchronous setting to our
-implementation.
+taking the union of their predicted biomarkers. A majority rule was
+applied when a biomarker was subject to qualitatively different
+predictions between the enzyme-catalyzed reactions encoded by the gene.
+We will refer to this as the asynchronous setting. Blocking all affected
+reactions simultaneously could yield different predictions. We will
+refer to this as the synchronous setting. In our implementation, we
+added the option to choose between the asynchronous and synchronous
+setting.
 
@@
 chemically different reactions, \(R_1\) and \(R_2\). Using the
-asynchronous setting we would run the algorithm once for \(R_1\) and
+asynchronous setting, we would run the algorithm once for \(R_1\) and
 once for \(R_2\) and take the union of all predicted biomarkers.
@@
 synchronous setting, we would run the algorithm once. For the wild-type
-we would simultaneously force flux through both reactions and in the
-mutant we would block both reactions. The synchronous approach simulates
-the complete catalytic inactivity or absence of the enzyme and therefore
-the infeasibility of any reactions it catalyzes. This corresponds to a
-deletion of the gene. The asynchronous approach might correspond to a
-point mutation that changes the specificity of the enzyme.
+simulation, we would simultaneously force flux through both reactions
+and in the mutant simulation, we would block both reactions. The
+synchronous approach simulates the complete catalytic inactivity or
+absence of the enzyme and therefore the infeasibility of any reactions
+it catalyzes. This corresponds to a deletion of the gene. The
+asynchronous approach might correspond to a point mutation that changes
+the specificity of the enzyme.
 
@@
 \caption{\label{tbl:PKU_results}Summary of predictions made using the
-FVA-based biomarker prediction method for the inborn-error of metabolism
+FVA-based biomarker prediction method for the inborn error of metabolism
 phenylketonuria for two networks (\(R_1\) = Recon 1 and \(R_2\) = Recon
@@
 hydroxyphenylpyruvate (hpp +/-) in the medium. \(^*\) refers to a
-situation without any tyrosine production in the mutant. \(^{**}\)
-refers to the situation where tyrosine reduction is so small that it
-does not reach the 10\% threshold. `-' indicates no prediction because
-the boundaries did not change from wild-type to mutant. \(\epsilon = 0\)
-refers to the case where we do not force any flux through the affected
-reaction(s) in the wild-type. In all other columns, \(\epsilon = 1\).
-The full output with interval bounds that the method generated can be
-found in the relevant notebook in this repository. }\tabularnewline
+situation without any tyrosine production in the mutant simulation.
+\(^{**}\) refers to the situation where tyrosine reduction is so small
+that it does not reach the 10\% threshold. `-' indicates no prediction
+because the boundaries did not change from wild-type to mutant.
+\(\epsilon = 0\) refers to the case where we do not force any flux
+through the affected reaction(s) in the wild-type simulation. In all
+other columns, \(\epsilon = 1\). The full output with interval bounds
+that the method generated can be found in the relevant notebook in this
+repository. }\tabularnewline
 \toprule
@@
 
-Returning to the example of phenylketonuria we simulated this for Recon
-2.2, using the synchronous approach. Column 3 in Table 1, shows that the
-result now agrees with the Recon1 asynchronous predictions. Using Recon
-1 we came to the correct prediction because it contains only one
-reaction (`PHETHPTOX2') linked to the PKU gene, i.e.~only phenylalanine
-hydroxylase (Entrez ID: 5053, HGNC:8582):
+Returning to the example of phenylketonuria, we simulated this for Recon
+2.2 using the synchronous approach. Column 3 in Table
+{[}tbl:PKU\_results{]} shows that the result now agrees with the Recon 1
+asynchronous predictions. Using Recon 1, we came to the correct
+prediction because it contains only one reaction (`PHETHPTOX2') linked
+to the PKU gene, i.e.~only phenylalanine hydroxylase (Entrez ID: 5053,
+HGNC:8582):
 
@@
 
-In Recon1, the 4a-hydroxytetrahydrobiopterin produced may dehydrate to
+In Recon 1, the 4a-hydroxytetrahydrobiopterin produced may dehydrate to
 6,7-dihydrobiopterin by the action of the `THBPT4ACAMDASE' reaction
@@
 
-This details are included in the pathway summary in our Figure
+These details are included in the pathway summary in our Figure
 ~\ref{fig:PKU_network}. When the reactions affected by the PKU gene are
-knocked out simultaneously, Recon2 correctly predicts both tyrosine and
+knocked out simultaneously, Recon 2 correctly predicts both tyrosine and
 phenylalanine as biomarkers. When blocking only one reaction, there is a
-way to produce tyrosine. Consequently, tyrosine is not predicted as
-biomarker if Recon 2 is used as metabolic map.
+way to produce tyrosine. Consequently, tyrosine is not predicted as a
+biomarker if Recon 2 is used as the metabolic map.
 
 Another significant modulation point for the method, and for
-flux-balance-based models in general, is the choice of medium. More
-specifically, it matters for the biomarker prediction which other
-exchange reactions are allowed to have a non-zero flux and what the
-bounds on these fluxes are. Changes herein leave the internal network
-topology unchanged and instead alter how the model is allowed to
-interface with biological compartments or environments that are not
-explicitly modeled. Potentially this potentially has large consequences
-as it may open up or shut down entire parts of the metabolic network
-that rely on specific substrates. It additionally interacts with the
-setting of the threshold for considering biomarkers significant since
-the medium components together with the network pathways (and their
+flux-balance-based models in general, is the medium. More specifically,
+it matters for the biomarker prediction which other exchange reactions
+are allowed to have a non-zero flux and what the bounds on these fluxes
+are. Changes in this leave the internal network topology unchanged and
+instead alter how the model is allowed to interface with biological
+compartments or environments that are not explicitly modeled. This
+potentially has major consequences, as it may open up or shut down
+entire parts of the metabolic network that rely on specific substrates.
+Additionally, it interacts with the significance threshold, since the
+medium components together with the network pathways (and their
 V\(_{\text{max}}\)'s) determine the amount of a given metabolite that
@@
 
-In order to examine these ramifications we continue with the synchronous
-PKU example and note that tyrosine production flux goes down in the
-mutant, but not to zero. This is due to an alternative tyrosine
-synthesis pathway in the model originating from 4-Hydroxyphenylpyruvate
-in the medium through the `TYRTA' reaction (tyrosine transaminase E.C.
-2.6.1.5). This transaminase is considered to be reversible in Recon 2
-and Equilibrator \autocite{Noor2012} annotates this reaction with a
-\(K'_{\text{eq}} \approx 1\). It should be mentioned though that the
-ability for transport of hydroxyphenylpyruvate is speculative. Column 4
-and 5 in Table 1 show results when using the standard medium and a
-medium without or with increased (i.e.~influx at -20)
-4-Hydroxyphenylpyruvate influx respectively. In FBA and FVA changes in
-medium concentrations of exchange metabolites are modelled by changing
-the corresponding influx (inward exchange) V\(_{\text{max}}\) (bound).
-In the former the ability to produce tyrosine indeed disappears in the
-mutant case. In the latter there is so much left-over capacity that the
-change in interval bounds slips below the 10\% cutoff. This sensitivity
-to medium composition appropriately reflects the phenomenon that cells
-behave differently in different culture media, and that organisms behave
-differently depending on their nutrition. Our discussion here may
-highlight that this feature should be expected to affect the pertinence
-of biomarkers. Biomarkers may have a tendency to be non-robust,
-dependent as they can be on network topology details and nutrition.
+In order to examine these ramifications, we continue with the
+synchronous PKU example and note that tyrosine production flux goes down
+in the mutant simulation, but not to zero. This is due to an alternative
+tyrosine synthesis pathway in the model originating from
+4-Hydroxyphenylpyruvate in the medium through the `TYRTA' reaction
+(tyrosine transaminase E.C. 2.6.1.5). This transaminase is considered to
+be reversible in Recon 2 and Equilibrator \autocite{Noor2012} annotates
+this reaction with a \(K'_{\text{eq}} \approx 1\). It should be
+mentioned, however, that the ability for transport of
+hydroxyphenylpyruvate is speculative. Column 4 and 5 in Table 1 show the
+results when using the standard medium and a medium without or with
+increased (i.e.~influx at -20) 4-Hydroxyphenylpyruvate influx
+respectively. In FBA and FVA, changes in medium concentrations of
+exchange metabolites are modelled by changing the corresponding influx
+(inward exchange) V\(_{\text{max}}\) (bound). In the former, the ability
+to produce tyrosine indeed disappears in the mutant case. In the latter,
+there is so much leftover capacity that the change in interval bounds
+slips below the 10\% cutoff. This sensitivity to medium composition
+appropriately reflects the phenomenon of cells behaving differently in
+different culture media, and organisms behaving differently depending on
+their nutrition. Our discussion here may highlight that this feature
+should be expected to affect the pertinence of biomarkers. Biomarkers
+may have a tendency to be non-robust, dependent as they can be on
+network topology details and nutrition.
 
@@
 wild-type condition. The last column in Table ~\ref{tbl:PKU_results}
-highlights that phenylalanine does not to show up as a biomarker if
-setting \(\epsilon = 0\) and performing the PKU simulation on Recon 1.
-This may also be the reason why Sahoo et al. did not manage to find
+indicates that phenylalanine does not show up as a biomarker if setting
+\(\epsilon = 0\) and the PKU simulation is performed on Recon 1. This
+may also be the reason why Sahoo et al. did not manage to find
 phenylalanine as a biomarker for PKU in their study
@@
 ~\ref{fig:Shlomi2009_figure_1}. The simple nature of the example serves
-to explain the reasoning behind the method and additionally helped us
-validate our implementation of the method. This repository provides an
-SBML file of the network, the Jupyter notebook that generates it and a
-Jupyter notebook reproducing the biomarker prediction.
+to explain the reasoning behind the method and also helped us validate
+our implementation of the method. This repository provides an SBML file
+of the network, the Jupyter notebook that generates it and a Jupyter
+notebook reproducing the biomarker prediction.
 
@@
 the conversion of M1 into M2 coupled to the conversion of M3 into M4.
-Shlomi et al. provide the a biomarker prediction for this network
-graphically in their Figure 1. We here do so numerically and a Jupyter
-notebook reproducing the biomarker prediction accompanies this paper. In
-our analysis all exchange reactions were given inward bounds of \(-10\)
-and outward bounds of \(+1000\) so that the internal reaction bounds (of
-\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
-shows that all exchangeable metabolites except M5 and M7 are predicted
-to be robust biomarkers.
+Shlomi et al. graphically provide the biomarker prediction for this
+network in their Figure 1. Here, we do so numerically and this paper is
+accompanied by a Jupyter notebook reproducing the biomarker prediction.
+In our analysis, all exchange reactions were given inward bounds of
+\(-10\) and outward bounds of \(+1000\) so that the internal reaction
+bounds (of \(\pm 1000\)) are never reached. Our Table
+~\ref{tbl:toy_model_results} shows that all exchangeable metabolites
+except M5 and M7 are predicted to be robust biomarkers.
 
 The exchange interval we found for metabolite M6 shows that in the
-wild-type case M6 can be either taken-up from biofluids or secreted. The
-upper bound is lower in absolute value than the lower bound because in
-the wild-type at least \(1\) unit of flux is needed to provide M3 as
+wild-type case, M6 can be either taken up from biofluids or secreted.
+The upper bound is lower in absolute value than the lower bound because
+in the wild type, at least \(1\) unit of flux is needed to provide M3 as
 substrate for the enzyme under investigation since the enzyme convert M1
@@
 the disease case, M6 (synthesized through M5) can only be secreted to
-biofluids. These results fully agree with the original publication.
+biofluids. These results are in full agreement with the original
+publication.
 
@@
 Figure 2 in their publication compares predicted biomarkers with known
-biomarkers from the OMIM database \autocite{McKusick2007}.
-
-We reproduced the analysis for each IEM listed in Table
-~\ref{tbl:amino_acid_iems} and ultimately found full agreement between
-what was reported by Shlomi et al. and the results of our
-implementation.
+biomarkers from the OMIM database \autocite{McKusick2007}. We reproduced
+the analysis for each IEM listed in Table ~\ref{tbl:amino_acid_iems} and
+ultimately found full agreement between what was reported by Shlomi et
+al. and the results of our implementation.
 
@@
 We made use of the Excel file included as supplementary information in
-\autocite{Shlomi2009} and included it in this repository as well. In
-this Excel file the genes and reactions linked to a given IEM, are
-detailed. We attempted to extract the genes for each IEM and through the
-Recon 1 map deduce the coupled reactions but found out that this yielded
-predictions different from the ones presented by Shlomi et al. In the
-Excel file a list of reactions linked to the IEM is also provided.
-However, these reactions are not necessarily the same reactions that are
-linked in Recon 1 to the genes they provide. Assuming the reconstruction
-of the metabolic network to be accurate this raises some doubts as to
-the correctness of the results obtained by Shlomi et al. for those IEMs
-where the listed reactions did not agree with the reconstructed map.
+\autocite{Shlomi2009} and included it in this repository as well. The
+genes and reactions linked to a given IEM are detailed in this Excel
+file. We attempted to extract the genes for each IEM and using the Recon
+1 map, deduced the coupled reactions, but found out that this yielded
+predictions different from the ones presented by Shlomi et al. However,
+the reactions listed in the Excel file are not necessarily the same
+reactions that are linked in Recon 1 to the genes linked to the IEMs.
+Assuming the reconstruction of the metabolic network to be accurate,
+this raises some doubts as to the correctness of the results obtained by
+Shlomi et al. for those IEMs where the listed reactions did not agree
+with the reconstructed map.
 
@@
 Excel file associated to the reaction HISD whereas this should have
-refered to HISDr to match the reaction ID in Recon 1. After these
-changes, we recovered the results as originally presented and as
+referred to HISDr to match the reaction identifier in Recon 1. After
+these changes, we recovered the results as originally presented and as
 summarized here in Table ~\ref{tbl:amino_acid_iems}.
@@
 ~\ref{fig:Shlomi2009_figure_1}. H.C. stands for `highly confident'
-indicating, as defined by Shlomi et al., i.e.~no overlap between the WT
-and mutant intervals. Between brackets {[}{]} we report the lower and
-upper bounds of the FVA intervals we obtained. }\tabularnewline
+indicating, as defined by Shlomi et al., no overlap between the WT and
+mutant intervals. Between brackets {[}{]} we report the lower and upper
+bounds of the FVA intervals we obtained. }\tabularnewline
 \toprule
@@
 been successfully replicated with the software tools provided here.
-Overall, the reproducing process was smooth since the method is
-illustrated both on an illustrative example and on the human metabolic
+Overall, the reproduction process went smoothly, since the method was
+applied to both an illustrative example and to the human metabolic
 reconstruction, resulting in a multitude of test cases for the
@@
 
-The points of trouble we encountered were the lack of explicit
-discussion on the set of fluxes allowed into and out of the model and a
-set of discrepancies between the gene-reaction mappings listed in the
-Supplementary Excel file and the actual gene-reaction mappings in the
-metabolic network. However, trial-and-error and manual curation solved
+The obstacles we encountered were the lack of explicit discussion on the
+set of fluxes allowed into and out of the model and a set of
+discrepancies between the gene-reaction mappings listed in the
+supplementary Excel file and the actual gene-reaction mappings in the
+metabolic network. However, trial and error and manual curation solved
 these issues.
@@
 \caption{\label{tbl:amino_acid_iems}We used our implementation to
-reproduce the biomarkers for the list of inborn-errors of metabolism
-shown in Figure 2 in \autocite{Shlomi2009}. Using the Supplementary
+reproduce the biomarkers for the list of inborn errors of metabolism
+shown in Figure 2 in \autocite{Shlomi2009}. Using the supplementary
 Excel table of \autocite{Shlomi2009} linking IEMs to causative genes and
-affected reactions we filtered out those that are listed here. The only
+affected reactions, we filtered out those that are listed here. The only
 considered biomarkers are the \(20\) proteinogenic amino acids. We
@@
 Hyperprolinemia type I & PRO1xm, PROD2m & &\tabularnewline
-maple syrup & OIVD1m, & L-Valine, &\tabularnewline
+Maple syrup & OIVD1m, & L-Valine, &\tabularnewline
 urine disease & OIVD2m, & L-Isoleucine &\tabularnewline
@@
 & MMMm & &\tabularnewline
-phenylketonuria & PHETHPTOX2 & L-Phenylalanine &
+Phenylketonuria & PHETHPTOX2 & L-Phenylalanine &
 L-Tyrosine\tabularnewline

### 2017-11-08 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-RomaroC-CeballosCC-RoqueAC-2017.tex
Add files via upload
Initial commit

\documentclass[10pt,a4paper,onecolumn]{article}
% * <aroquesilva@gmail.com> 2017-09-19T13:55:25.418Z:
%
% ^.
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\usepackage{cleveref} %added by Renan
\usepackage{multirow} %added by Rodrigo
\usepackage{amsmath,amssymb} %added by Roque
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Renan O. Shimoura,
      Nilton L. Kamiji,
      Rodrigo F.O. Pena,
  },
            pdftitle=
{
[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}

\newcommand{\blau}[1]{{\color{blue} #1}}
\newcommand{\rot}[1]{{\color{red} #1}}

\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-RomaroC-CeballosCC-RoqueAC-2017.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Nov 2017 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model}

    \usepackage{authblk}
                    \author[1]{Renan O. Shimoura}
                    \author[1]{Nilton L. Kamiji}
                    \author[1]{Rodrigo F.O. Pena}
                    \author[1]{Vinicius L. Cordeiro}
                    \author[1,2]{Cesar C. Ceballos}
                    \author[1]{Cecilia Romaro}
                    \author[1]{Antonio C. Roque}
                    \affil[1]{Faculty of Philosophy, Sciences and Letters at Ribeir\~{a}o Preto, Department of Physics, University of S\~{a}o Paulo, Ribeir\~{a}o Preto, S\~{a}o Paulo, Brazil}
                    \affil[2]{Department of Physiology, Ribeir\~ao Preto Medical School, School of Medicine, University of S\~ao Paulo, Ribeir\~ao Preto, Brazil}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:antonior@usp.br}{antonior@usp.br}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Month, day, year\\
  {\bfseries Accepted}  Month, day, year\\
  {\bfseries Published} Month, day, year\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C. Potjans,  M. Diesmann, Cerebral Cortex, 24 (3): 785-806, 2014. DOI: 10.1093/cercor/bhs358
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}
Most theoretical studies of cortical activity are based on networks of randomly connected units \cite{brunel2000,vogels2005,kriener2014,ostojic2014} 
or with architectures artificially
built from random networks \cite{tomov2014}. In spite of the usefulness of these models, in order to understand the interplay between network structure and cortical dynamics it is essential to have computational models which accurately represent the cortical network architecture. Recently, Potjans and Diesmann \cite{potjans2014} developed a network model of the local cortical microcircuit based on extensive experimental data on the intrinsic circuitry of striate cortex \cite{thomson2002,binzegger2004}.
The model contains two cell types (excitatory and inhibitory) distributed over four layers, L2/3, L4, L5, and L6, and represents the cortical network below a surface area of 1 $mm^{2}$ (a scheme is shown in Fig.~\ref{fig:diagram_net}).

 \begin{figure}[htbp]
 \centering
 \includegraphics[scale=0.4]{diagram.pdf}
 \caption{Schematic representation of the cortical network model (adapted from \cite{potjans2014}).
 The model consists of four layers (L2/3, L4, L5 and L6), each one populated with excitatory (triangles) and inhibitory (circles) neurons (Table \ref{table:population}). Arrows represent connections with probabilities $>$ 0.04: excitatory in red and inhibitory in blue (Table \ref{table:conectivity}). Black arrows represent background inputs.}\label{fig:diagram_net}
 \end{figure}

The original implementation was based on the NEST simulator \cite{gewaltig2007} and the source code is available at the Open Source Brain platform \cite{potjans2014pyNEST}.
Here, we reimplemented the full model in the Brian 2 simulator \cite{goodman2009} without direct reference to the original source code.

\section{Methods}\label{methods}

In this work, we replicated in Brian 2 every detail of the Potjans-Diesmann model as described in their original article \cite{potjans2014}. Hereafter, we will refer to the original NEST implementation of the Potjans-Diesmann model \cite{potjans2014} as reference (or original) article.  In this section we explain how this reimplementation was done. Further statistical analyses were performed using SciPy, NumPy, and Matplotlib libraries for the Python language.

\subsection{Neurons}

Network neurons are described by the leaky integrate-and-fire (LIF) neuron model. The subthreshold membrane voltage of neuron $i$ obeys the equation
\begin{equation}
\label{Eq:LIF}
\dot{V}_i(t) = -\left( V_i(t) - V_{\rm reset} \right) /\tau_m + I_i(t)/C_m,
\end{equation}
where $\tau_m$ is the membrane time constant, $C_m$ is the membrane capacitance, $V_{\rm reset}$ is the reset potential, and $I_i(t)$ is the total input current. When $V_i(t) \geq V_{\rm th}$ the neuron emits a spike and the voltage is reset to $V_{\rm reset}$, remaining fixed at $V_{\rm reset}$ for a refractory period $\tau_{\rm ref}$. The total input current $I_i(t)$ is divided into external $I_{i, \rm ext}(t)$ and synaptic $I_{i, \rm syn}(t)$. Whenever an excitatory (or inhibitory) neuron $j$ presynaptic to neuron $i$ fires at time $t_{j}^{f}$, the synaptic current to neuron $i$ changes at time $t_{j}^{f}+d_e$ (or $d_i$) by an amount $w$ (or $-g w$), where $w$ is the excitatory synaptic weight, $g$ is the relative weight of the inhibitory synapse, $d_e$ is the excitatory synapse transmission delay, and $d_i$ is the inhibitory synapse transmission delay. In the absence of synaptic inputs the synaptic current changes as
\begin{equation}
\label{Eq:Isyn}
\dot{I}_{i, \rm syn}(t) = -I_{i, \rm syn}(t)/\tau_{\rm syn} 
\end{equation}
\noindent where $\tau_{\rm syn}$ is the postsynaptic current time constant (parameter values are shown in Table \ref{table:parameters}).

\begin{table}[!ht]
  \caption{\label{table:parameters} Neuron and synaptic parameters. Neuron and synaptic parameters used in our simulations according to \cite{potjans2014}.}
 \begin{center}
  \begin{tabular}{lccccccccc}
   \toprule
   \midrule
 membrane time constant & $\tau_m$ & 10 ms \tabularnewline
 refractory period & $\tau_{ref}$ & 2 ms \tabularnewline
 postsynaptic current time constant & $\tau_{syn}$ & 0.5 ms \tabularnewline
 membrane capacitance & $C_m$ & 250 pF\tabularnewline
 reset voltage & $V_{\rm reset}$ & -65 mV\tabularnewline
 threshold voltage & $V_{\rm th}$ & -50 mV\tabularnewline
 excitatory synaptic weight & $w$ & $\mathcal{N}$($\mu = 87.8, \sigma = 8.8$) pA \tabularnewline
 relative inhibitory synaptic weight & $g$ & 4 \tabularnewline
 excitatory synaptic transmission delay & $d_e$ & $\mathcal{N}$($\mu=1.5 , \sigma = 0.75$) ms \tabularnewline
 inhibitory synaptic transmission delay & $d_i$ & $\mathcal{N}$($\mu = 0.8 , \sigma = 0.4$) ms \tabularnewline
 initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-58.0, \sigma=10.0$) mV \tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

\subsection{Network}
The procedure to set up the network connections is the following: 
\begin{itemize}
\item Start with a set of neurons $N=$ 77,169, with model parameter described in Table~\ref{table:parameters}.

\item The $N$ neurons are distributed over the eight different populations, L23e, L23i, etc, according to the numbers shown in Table~\ref{table:population}.

\item For each one of the sixty-four possible combinations of two from the eight cell populations, the total number $K$ of synapses is calculated using equation~(\ref{Eq:Ksyn}) (compare with equation~(1) of the original article), where $N_{pre/post}$ are the sizes of presynaptic/postsynaptic populations and $C_a$ is the corresponding connection probability given in Table~\ref{table:conectivity} (the subindex $a$ stands for `anatomical' in the terminology of the original article).

\begin{equation}
\label{Eq:Ksyn}
K = \frac{\log(1 - C_a)}{\log(1-1/(N_{pre}N_{post}))}
\end{equation}

\item For every one of the sixty-four two-cell populations, the $K$ synapses determined above are created by uniformly and randomly choosing $K$ pairs of neurons (one from each population) and placing a connection between them. This is done with repetition to allow the creation of multiple synaptic contacts between any pair of neurons.

\item The synaptic weight for connections originating from excitatory neurons is set to $w$ and the synaptic weight for connections from inhibitory neurons is set to $-gw$. In addition, the synaptic weight for connections from neurons of layer L4e to L23e is doubled \cite{potjans2014,wagatsuma2011}.
\end{itemize}

\begin{table}[!ht]
  \caption{\label{table:population} Layer population sizes (extracted from \cite{potjans2014}). Neurons were distributed over four different layers (L23, L4, L5 and L6), and for each layer they were divided into excitatory (L23e, L4e, L5e and L6e) and inhibitory (L23i, L4i, L5i and L6i) subpopulations.}
 \begin{center}
  \begin{tabular}{cccccccc}
   \toprule
   \midrule
 L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
 20683 & 5834 & 21915 & 5479 & 4850 & 1065 & 14395 & 2948\tabularnewline
   \bottomrule
  \end{tabular}
 \end{center}
\end{table}

\begin{table}[!ht]
\caption{\label{table:conectivity} Connectivity matrix between the different populations of the model (extracted from \cite{potjans2014}). The connectivity matrix describes the probabilities of the target-specific connections between populations of neurons.
}
\begin{center}
\begin{tabular}{lccccccccc}
 \toprule
 \midrule
 &  &  &  &  & from &  &  &  & \tabularnewline
  \midrule
 &  & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
 \midrule
 & L23e & 0.101 & 0.169 & 0.044 & 0.082 & 0.032 & 0 & 0.008 & 0\tabularnewline
 & L23i & 0.135 & 0.137 & 0.032 & 0.052 & 0.075 & 0 & 0.004 & 0\tabularnewline
 & L4e & 0.008 & 0.006 & 0.050 & 0.135 & 0.007 & 0.0003 & 0.045 & 0\tabularnewline
 to & L4i & 0.069 & 0.003 & 0.079 & 0.160 & 0.003 & 0 & 0.106 & 0\tabularnewline
 & L5e & 0.100 & 0.062 & 0.051 & 0.006 & 0.083 & 0.373 & 0.020 & 0\tabularnewline
 & L5i & 0.055 & 0.027 & 0.026 & 0.002 & 0.060 & 0.316 & 0.009 & 0\tabularnewline
 & L6e & 0.016 & 0.007 & 0.021 & 0.017 & 0.057 & 0.020 & 0.040 & 0.225\tabularnewline
 & L6i & 0.036 & 0.001 & 0.003 & 0.001 & 0.028 & 0.008 & 0.066 & 0.144\tabularnewline
\bottomrule
\end{tabular}
\end{center}
\end{table}

The ordinary differential equations were solved with the exact integration method for linear equations available in Brian 2 with a time step $\Delta t=0.1$ ms. All simulations were carried out in a cluster with 4 nodes each equipped with 2 Intel Xeon processors.

\subsection{External input}

We chose from the reference paper three different types of external (``background") inputs:

\begin{enumerate}
\item \textbf{Layer specific}: Neurons from each layer receive specific background spike-trains drawn from an $8$ Hz Poisson distribution. The number of inputs per neuron is given in the first row of Table \ref{table:externalinput}.
\item \textbf{Layer independent}: Spike-trains are drawn from an $8$ Hz Poisson distribution as above, but now the number of inputs per neuron is the same for all the excitatory layers and the same for all the inhibitory layers as shown in the second row of Table \ref{table:externalinput}.
\item \textbf{DC input}: The Poissonian background is replaced by constant DC currents to all neurons. The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}). 
%\item \textbf{Frequency sweep}: External spikes are drawn from Poisson distributions with frequency values in the range from 3 to 15 Hz. The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}).
\end{enumerate}

\begin{table}[!ht]
\caption{\label{table:externalinput} Estimated numbers of external inputs per neurons in all network layers. The total number of external inputs is rounded.
}
\begin{center}
\begin{tabular}{lcccccccccc}
 \toprule
 \midrule
 & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i %& Th
 \tabularnewline
 \midrule
 Layer specific & 1600 & 1500 & 2100 & 1900 & 2000 & 1900 & 2900 & 2100 \tabularnewline
 Layer independent & 2000 & 1850 & 2000 & 1850 & 2000 & 1850 & 2000 & 1850 \tabularnewline
 Background rate  & & & & 8 Hz \tabularnewline
\bottomrule
\end{tabular}
\end{center}
\end{table}

\subsection{Measures}

Here, we define the measures used to characterize the layer-specific activity of the network. They are the same ones used in the original article.

The spike train of a neuron $i$ is represented by a sequence of temporal events (sum of delta functions). The firing rate over an interval $T$ is obtained by summing the number of spikes during that interval and dividing by $T$ (spike-count firing rate). The average firing rate of a population of $N$ neurons is computed by calculating the firing rates of the neurons and dividing by $N$. With this procedure we calculated the average firing rates of the eight populations in the network.

To characterize irregularity in the network we use the coefficient of variation (CV) of the interspike interval (ISI) distribution. The $\rm CV_{\rm i}$ for each neuron $i$ is computed as the ratio of the standard deviation over the mean of its ISI distribution. Exponential distributions have $\rm CV \approx 1$ while more regular distributions have $\rm CV<1$.

Synchrony is characterized by the variability of the histogram of population spiking activity (bin size = 3 ms). The synchrony index is computed as the variance of the spike count histogram divided by its mean.

The degree of asynchronous and irregular activity in a population is quantified by a measure called AIness\%. This is the percentage of the population with mean firing rate $<$ 30 Hz, irregularity between 0.7 and 1.2, and synchrony $<$ 8.

To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test. To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.

\section{Results}\label{results}

In the following, we present results of the replicated studies (with the same data sampling sizes) done in the reference article for the network with parameters as defined in Methods. We did not replicate analyses which involve changes in the network structure.

\subsection{Spontaneous Activity}
The simulated spontaneous activity is asynchronous and irregular (Fig.~\ref{fig:fig2}\textbf{A}) and the cell-type specific firing rates are in agreement with the ones observed in the reference article, including the lowest rates for the excitatory cells of layers 2/3 and 6 and the highest rates for L5 cells (Fig.~\ref{fig:fig2}\textbf{B}). For all layers the inhibitory cell firing rates exceed the ones of excitatory cells. The firing rate variabilities (Fig.~\ref{fig:fig2}\textbf{B}) and the single-cell firing rate irregularities (Fig.~\ref{fig:fig2}\textbf{C}) are also similar to the ones reported in the original article. The irregularity measure is $>$ 0.80 for all cell populations (Fig.~\ref{fig:fig2}\textbf{C}). The profile of the synchrony of spiking activity across the cell populations is also consistent with the one reported in the reference article. The highest degree of synchrony is found in L5e and the lowest one in L6.

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.001]{fig2.pdf}
 \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article). In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz). (A) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells). Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer). (B–D) Statistics based on samples of 1000 neurons per layer recorded for 60 s. (B) Single-cell firing rate boxplot. Triangles indicate population mean firing rates, crosses show outliers. (C) Irregularity of single-cell firing rates. (D) Synchrony of spiking activity.}\label{fig:fig2}
\end{figure}

In order to compare the Brian 2 with the NEST implementations of the Potjans-Diesmann model, we used the available pyNEST code \cite{potjans2014pyNEST} to run NEST simulations of the model with the same parameters given in Methods. The comparisons were made in terms of the cumulative distributions of the CVs of ISIs and the firing rates of the eight cell populations in the network (plots not shown here but as in Figs.~\ref{fig:fig3} and~\ref{fig:fig4} below). The p-value (two-sample Kolmogorov-Smirnov test) was higher than 0.6 for all comparisons meaning that no significant difference was found between NEST and Brian 2 simulations.

Besides the creation of the network connections by the procedure described above, which is the one used in the original article, it is also possible to connect the neurons in the network using the alternative expression for the total number of synapses between two cell populations,

\begin{equation}
\label{Eq:Ksyn_alt}
K = C_a(N_{pre}N_{post}),
\end{equation}
\noindent which comes from the first-order Taylor series approximation to the connection probability $C_{a}$ as explained in the original article (see equation (2) and following text in the original article).

%inbuilt Brian 2 randomization option in the $connect(i,j,p)$ function. In other words, there are two ways of using the $connect$ function to create the network connections: in one way the user prescribes a list of pre- and postsynaptic pairs and feeds the function with this list. In the other way the user feeds $i$ and $j$ with the pre- and postsynaptic populations and $p$ acts as the connection probability among these populations. Here we used the first way to connect the neurons and perform the comparison between NEST and Brian 2 simulations, but we also tested the second way and found that it brings different results as shown in Figs.~\ref{fig:cv_hists} and~\ref{fig:f_hists}.
% * <aroquesilva@gmail.com> 2017-11-07T21:03:53.535Z:
%
% ^.
To test for possible effects of using this approximate equation, we constructed the network in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) instead of~(\ref{Eq:Ksyn}). We found that the use of the approximate equation brings discrepancies in comparison with the original model. The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations for simulations of the original model based on the pyNEST code \cite{potjans2014pyNEST} and the Brian 2 code with $K$ calculated from equation~(\ref{Eq:Ksyn_alt}).

Most of the layer-specific comparisons yielded p-values higher than 0.6 as before, indicating no significant differences. However, the p-value of the comparison between the CV of ISIs for the layer 5 excitatory neurons was $0.07$, and the p-values of the comparisons between the firing rates of excitatory cells from layers 2/3 and 6 were, respectively, $0.10$ and $0.08$. Therefore, in the cases of excitatory cells of layers 2/3, 5 and 6 there were statistically significant differences in the firing properties of neurons. The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have firing rate 12.2 $\pm$ 6.1 Hz while in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig3.pdf}
 \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue) and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
\end{figure}

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig4.pdf}
 \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue) and in NEST using equation~(\ref{Eq:Ksyn}) (orange). NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
\end{figure}

The differences found highlight the importance of using the exact expression for the connection probability $C_{a}$ given in equation (1) of the original article (which corresponds to equation~(\ref{Eq:Ksyn}) of this text)  in simulations of the Potjans-Diesmann model.

We now return to the reimplementation of the Potjans-Diesmann model in Brian 2 using equation~(\ref{Eq:Ksyn}), which will be kept for the rest of this replication work.

\subsection{Dependence of Spontaneous Activity on External Inputs}

In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs. These are shown in Fig.~\ref{fig:fig5}A, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}B, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs. In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig5.pdf}
 \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Figs. 7A and 7B of the original article). Color codes and sample cell sizes as in Fig. \rot{\ref{fig:fig2}}. (A1--A2) DC input. (B1--B2) Layer independent input.}\label{fig:fig5}
\end{figure}

\subsection{Stability of Network Activity}

The activity features obtained in the original article by changing the relative strength of inhibitory synapses and the background rate are reproduced in our reimplementation (Fig.~\ref{fig:fig6}).

The asynchronous and irregular activity of the reimplemented network model, as characterized by the AIness\%, is similar to the one found in the reference article for background rates >5 Hz and relative inhibitory synaptic strengths >4 Hz. In comparison with the reference article, the relative order of excitatory firing rates is maintained for every combination between background rate and relative synaptic strength, with highest values in L5 and smallest in L2/3 and L6. Similarly to the original article, the firing rate of L4e cells is the most sensitive to variations in the background rate whereas the firing rate of L5e cells is the most sensitive to variations in the relative inhibitory synaptic strength $g$.

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.01]{fig6.pdf}
 \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig. 8 of the original article). (A) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$. (B) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$. Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate. (C) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (A)).}\label{fig:fig6}
\end{figure}

\section{Conclusion}\label{conclusion}
Using the Brian 2 reimplementation of the Potjans-Diesmann model we were able to reproduce the main results of the original article \cite{potjans2014}. The spontaneous activity of the network reimplementation is asynchronous and irregular as evaluated by the different measures used to characterize spiking behavior.

We also have shown the importance of using the exact expression in equation (1) of the original article instead of the approximate one in equation (2) in implementations of the model. The use of the approximate expression leads to mean firing rates of L5e neurons significantly higher than in the original implementation of the model.

The successful replication of the results of the reference article confirms the correctness of the original implementation of the model.

\section{Acknowledgements}\label{acknowledgements}
The authors would like to thank Prof. Markus Diesmann for valuable comments on the manuscript. This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC. ACR is partially supported by the CNPq fellowship grant 306251/2014-0. RP and ACR are also part of FAPESP 2015/50122-0 and DFG-GRTK 1740/2. CCC, and CR are supported by a CAPES PhD scholarship.

{\sffamily \small
   \printbibliography[title=References]
}

\end{document}

### 2017-12-12 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Rename ShimouraR-KamijiNL-PenaRFO-CordeiroVL-RomaroC-CeballosCC-RoqueAC-2017.tex to ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Uploaded version of the MS

--- 2017-11-08 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-RomaroC-CeballosCC-RoqueAC-2017.tex
@@
 ]{biblatex}
-\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-RomaroC-CeballosCC-RoqueAC-2017.bib}
+\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.bib}
 
@@
  inhibitory synaptic transmission delay & $d_i$ & $\mathcal{N}$($\mu = 0.8 , \sigma = 0.4$) ms \tabularnewline
- initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-58.0, \sigma=10.0$) mV \tabularnewline
+ initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-65.0, \sigma=6.5$) mV \tabularnewline
    \bottomrule
@@
 The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}).
-%\item \textbf{Frequency sweep}: External spikes are drawn from Poisson distributions with frequency values in the range from 3 to 15 Hz.
-The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}).
+Observe that the number of inputs per neuron is multiplied by an effective factor which is given by $\nu \times \omega \times \tau_{syn}=0.3512$ pA.
 \end{enumerate}
+
+Additionally to these inputs, in Fig.~\ref{fig:fig5}C we simulate the network in several trials where the number of external inputs change.
+In every trial, the external inputs to each of the excitatory layers is a number randomly drawn between the layer specific and the layer independent inputs reported in Table \ref{table:externalinput}.
+The inhibitory layers have their number of external inputs randomly decided between the already chosen number of their respective excitatory layer and the number calculated by
+
+\begin{equation}
+\label{Eq:Ksyn_alt}
+C_{i} = \left( \frac{1-T}{1+T} \right)C_{e} ,
+\label{Eq:transitivity}
+\end{equation}
+
+\noindent where $T$ is the target specificity defined in \cite{potjans2014} and is assumed here to be $T=0.1$.
+However, there is an exception to this rule: L6i is allowed to have $T=0.2$ due to the high number of inputs to L6e.
 
@@
  \midrule
- & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i %& Th
+ & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i
  \tabularnewline
@@
  Layer independent & 2000 & 1850 & 2000 & 1850 & 2000 & 1850 & 2000 & 1850 \tabularnewline
- Background rate  & & & & 8 Hz \tabularnewline
+ Background rate  ($\nu$) & & & & 8 Hz \tabularnewline
 \bottomrule
@@
  \centering
- \includegraphics[scale=0.001]{fig2.pdf}
+ \includegraphics[scale=0.8]{fig2.pdf}
  \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article).
@@
 
-%inbuilt Brian 2 randomization option in the $connect(i,j,p)$ function.
-In other words, there are two ways of using the $connect$ function to create the network connections: in one way the user prescribes a list of pre- and postsynaptic pairs and feeds the function with this list.
-In the other way the user feeds $i$ and $j$ with the pre- and postsynaptic populations and $p$ acts as the connection probability among these populations.
-Here we used the first way to connect the neurons and perform the comparison between NEST and Brian 2 simulations, but we also tested the second way and found that it brings different results as shown in Figs.~\ref{fig:cv_hists} and~\ref{fig:f_hists}.
-% * <aroquesilva@gmail.com> 2017-11-07T21:03:53.535Z:
-%
-% ^.
-To test for possible effects of using this approximate equation, we constructed the network in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) instead of~(\ref{Eq:Ksyn}).
+To test for possible effects of using this approximate equation, we constructed the network in Brian 2 using both equations (\ref{Eq:Ksyn_alt}) and (\ref{Eq:Ksyn}).
 We found that the use of the approximate equation brings discrepancies in comparison with the original model.
-The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations for simulations of the original model based on the pyNEST code \cite{potjans2014pyNEST} and the Brian 2 code with $K$ calculated from equation~(\ref{Eq:Ksyn_alt}).
-
-Most of the layer-specific comparisons yielded p-values higher than 0.6 as before, indicating no significant differences.
-However, the p-value of the comparison between the CV of ISIs for the layer 5 excitatory neurons was $0.07$, and the p-values of the comparisons between the firing rates of excitatory cells from layers 2/3 and 6 were, respectively, $0.10$ and $0.08$.
-Therefore, in the cases of excitatory cells of layers 2/3, 5 and 6 there were statistically significant differences in the firing properties of neurons.
-The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have firing rate 12.2 $\pm$ 6.1 Hz while in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
+The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations for simulations of the original model based on the pyNEST code \cite{potjans2014pyNEST} and the Brian 2 code with $K$ calculated from equations (\ref{Eq:Ksyn_alt}) and (\ref{Eq:Ksyn}).
 
@@
  \centering
- \includegraphics[scale=0.01]{fig3.pdf}
- \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue) and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
+ \includegraphics[scale=0.02]{fig3.pdf}
+ \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
 NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
@@
  \includegraphics[scale=0.01]{fig4.pdf}
- \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue) and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
+ \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
 NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
@@
 
+When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, most of the layer-specific comparisons yielded p-values higher than 0.6 as before, indicating no significant differences.
+However, the p-value of the comparison between the CV of ISIs for the layer 5 excitatory neurons was $0.07$, and the p-values of the comparisons between the firing rates of excitatory cells from layers 2/3 and 6 were, respectively, $0.10$ and $0.08$.
+Therefore, in the cases of excitatory cells of layers 2/3, 5 and 6 there were statistically significant differences in the firing properties of neurons.
+The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have firing rate 12.2 $\pm$ 6.1 Hz while in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
+
 The differences found highlight the importance of using the exact expression for the connection probability $C_{a}$ given in equation (1) of the original article (which corresponds to equation~(\ref{Eq:Ksyn}) of this text)  in simulations of the Potjans-Diesmann model.
+Observe the close agreement of the cumulative histograms when Brian 2 (K from equation~(\ref{Eq:Ksyn})) and NEST (K from equation~(\ref{Eq:Ksyn})) are compared.
 
@@
 In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
+In Fig.~\ref{fig:fig5}C we present the population firing rates for 100 trials with the rule explained in the methods section.
+This latter experiment presented an excellent agreement with the histograms observed in Fig.
+7 of the original article.
 
@@
  \includegraphics[scale=0.01]{fig5.pdf}
- \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Figs.
-7A and 7B of the original article).
+ \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig.
+7 of the original article).
 Color codes and sample cell sizes as in Fig.
@@
 (A1--A2) DC input.
-(B1--B2) Layer independent input.}\label{fig:fig5}
+(B1--B2) Layer independent input.
+(C1--C4) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods).
+}\label{fig:fig5}
 \end{figure}
@@
 ACR is partially supported by the CNPq fellowship grant 306251/2014-0.
-RP and ACR are also part of FAPESP 2015/50122-0 and DFG-GRTK 1740/2.
+RP and ACR are also part of the IRTG 1740 / TRP 2015/50122-0, funded by DFG / FAPESP.
 CCC, and CR are supported by a CAPES PhD scholarship.

### 2018-01-02 article/Mondeel_Ogundipe_Westerhoff-2017.tex
[Text] further updates following Aaron's comments
- Move definitions to the introduction
- Refer to the subpanels A and B of Figure 1
- update table IEM names to match original
- up-to-date pdf

@@
 
-The source code is not publicly available and the computation
-environment used was not mentioned in the publication. The
-\href{https://github.com/opencobra/cobratoolbox}{MATLAB COBRA toolbox}
-contains code to reproduce the analysis by Thiele et al. but this was
-also performed in MATLAB.
-
-Here we here present an implementation of the biomarker prediction
-method originally proposed by Shlomi et al., programmed in Python, and
-reproduce the results presented in Figure 1 and 2 of the original
-publication. These figures concern the method's application to a simple
-metabolic map and then to a set of \(17\) amino acid metabolism
-disorders. Our Python module provides the biomarker prediction algorithm
-and is agnostic of the map used. It can therefore be applied to other
-maps than the human metabolic reconstruction considered by Shlomi et al.
-Jupyter notebooks are provided that generate the illustrative map and
-reproduce all results referred to in Figures 1 and 2 of Shlomi et al.
-Our implementation is based on information given in the original paper.
-However, some details critical to the reproduction were left unexplained
-there and are highlighted here. We also include a more detailed
-discussion of one of the amino acid IEMs to examine the robustness of
-the predictions made using this approach.
-
-\section{Methods}\label{methods}
-
-\subsection{Overview}\label{overview}
-
-We implemented the method of Shlomi et al. in Python. Our Python
-function has several input parameters, most of which are optional in
-order to enable the user to perturb the network in various ways and to
-try variations of the method that differ from the ones implemented here
-or in Shlomi et al. The default settings are such that they mimic the
-approach Shlomi et al. originally proposed. All flux-variability
-analyses are performed using COBRApy \autocite{Ebrahim2013}.
-
-All simulations discussed here were performed in Python 3.6 using
-COBRApy \autocite{Ebrahim2013} (version 0.9) and Pandas (version
-0.20.3). This repository also functions in conjunction with
-\href{http://mybinder.org}{MyBinder} which allows for full
-reproducibility in the `cloud' without any need for installation of
-additional software.
-
-The rest of this section introduces the approach proposed by Shlomi et
-al. as we implemented it.
+Below, we first introduce the essential vocabulary and concepts
+necessary for understanding the method proposed by Shlomi et al.
 
@@
 
-Flux ranges for boundary metabolites are calculated by applying flux
-variability analysis \autocite{Mahadevan2003}. Mathematically, flux
-variability analysis for a specific exchange reaction \(v_i\) entails
-the following linear programming problem:
+Flux ranges for boundary metabolites, consistent with the requirements
+of a steady state and constraints on reaction reversibilities, may be
+calculated by applying flux variability analysis
+\autocite{Mahadevan2003}. Mathematically, flux variability analysis for
+a specific exchange reaction \(v_i\) entails the following linear
+programming problem:
 
@@
 S v &= 0 \nonumber \\
-Z  &\geq \phi Z_{\text{opt}} \nonumber \\
+Z  &\geq \phi Z_{\text{max}} \nonumber \\
 \alpha_k \leq & ~ v_k \leq \beta_k 
@@
 bound for reaction \(k\) and similarly \(\beta_k\) is the, possibly
-negative, upper bound for reaction \(k\). \(Z = c^T ~ v\) is the
-objective function defined for the map, entailing a linear combination,
-defined by column vector c, of one or more fluxes, and
-\(Z_{\text{max}}\) its maximal value. We define \(0 \leq \phi \leq 1\)
-to be an arbitrary minimal fraction of the maximal objective function
-value that has to be achieved. When \(\phi=1\) we ask for the range of
-flux allowed for reaction \(v_i\) while maintaining the maximal value of
-the linear combination of objective fluxes. When setting \(\phi = 0\),
-one is asking for the allowable flux range through reaction \(v_i\)
-regardless of the value of the objective function (although the sign has
-to be maintained). The latter, i.e. \(\phi=0\), is the case considered
-in the method by Shlomi et al.
+negative, upper bound for reaction \(k\). The reaction bounds are the
+\(V_{max}\)'s of the reactions and are part of the metabolic network
+definition, typically they are set to either \(0\) or \(\pm 1000\) if
+the true, biological, \(V_{max}\)'s are not known. These values allow
+specification of irreversible reactions by setting the lower bound (or
+upper bound) to zero. The index \(i\) specifies a single reaction that
+is to be maximized, whereas the index \(k\) is used to index the bound
+constraints that exist for each reaction, including \(v_i\). \(v_i\) may
+be maximized in the forward (positive flux) or the reverse (negative
+flux) direction if allowed by the bounds on \(v_i\). This is indicated
+by the \(\pm\) symbol. \(S\) is the stoichiometry matrix for the
+metabolic network of size \(mxr\) where \(m\) is the number of
+metabolites and \(r\) the number of reactions. The numeric, typically
+integer, elements of \(S\) represent the stoichiometry coefficients of
+each metabolite \(m\) in each reaction \(r\). Consequently, for large
+networks, this matrix is sparse (has many zero entries). \(Z = c^T ~ v\)
+is the objective function defined for the map, entailing a linear
+combination, defined by column vector \(c\), of one or more reactions.
+The vector \(c\) is an indicator of objective reactions, i.e.~it
+contains a value of \(1\) in rows corresponding to fluxes that are to be
+included in the objective, and zeros in all other rows. Due to the
+vector multiplication \(c^T ~ v\), the objective \(Z\) sums the fluxes
+of the reactions that correspond to rows containing a value of \(1\) in
+\(c\). \(Z_{\text{max}}\) denotes the maximal value of the objective
+function. We define \(0 \leq \phi \leq 1\) to be an arbitrary minimal
+fraction of the maximal objective function value that has to be
+achieved. When \(\phi=1\) we ask for the range of flux allowed for
+reaction \(v_i\) while maintaining the maximal value of the linear
+combination of objective fluxes. When setting \(\phi = 0\), one is
+asking for the allowable flux range through reaction \(v_i\) regardless
+of the value of the objective function (although the sign has to be
+maintained). The latter, i.e. \(\phi=0\), is the case considered in the
+method by Shlomi et al.
+
+\subsection{Reference implementation}\label{reference-implementation}
+
+The source code for the original publication is not publicly available
+and the computation environment used was not mentioned in the
+publication. The \href{https://github.com/opencobra/cobratoolbox}{MATLAB
+COBRA toolbox} contains code to reproduce the analysis by Thiele et al.
+but this was also performed in MATLAB.
+
+We here present an implementation of the biomarker prediction method
+originally proposed by Shlomi et al., programmed in Python, and
+reproduce the results presented in Figure 1A-B and 2 of the original
+publication. These figures concern the method's application to a simple
+metabolic map and then to a set of \(17\) amino acid metabolism
+disorders. Our Python module provides the biomarker prediction algorithm
+and is agnostic of the map used. It can therefore be applied to other
+maps than the human metabolic reconstruction considered by Shlomi et al.
+Jupyter notebooks are provided that generate the illustrative map and
+reproduce all results referred to in Figures 1A-B and 2 of Shlomi et al.
+Our implementation is based on information given in the original paper.
+However, some details critical to the reproduction were left unexplained
+there and are highlighted here. We also include a more detailed
+discussion of one of the amino acid IEMs to examine the robustness of
+the predictions made using this approach. We feel this adds to the
+utility of this reference implementation and the understanding of the
+approach introduced by Shlomi et al.
+
+\section{Methods}\label{methods}
+
+\subsection{Overview}\label{overview}
+
+We implemented the method of Shlomi et al. in Python. Our Python
+function has several input parameters, most of which are optional in
+order to enable the user to perturb the network in various ways and to
+try variations of the method that differ from the ones implemented here
+or in Shlomi et al. The default settings are such that they mimic the
+approach Shlomi et al. originally proposed. All flux-variability
+analyses are performed using COBRApy \autocite{Ebrahim2013}.
+
+All simulations discussed here were performed in Python 3.6 using
+COBRApy \autocite{Ebrahim2013} (version 0.9) and Pandas (version
+0.20.3). This repository also functions in conjunction with
+\href{http://mybinder.org}{MyBinder} which allows for full
+reproducibility in the `cloud' without any need for installation of
+additional software.
+
+The rest of this section introduces the approach proposed by Shlomi et
+al. as we implemented it.
+
+\subsection{Implementation of the
+algorithm}\label{implementation-of-the-algorithm}
 
@@
   The wild-type interval \(WT_{r,m}\) is equal to the union of the
-  forward and backward intervals.\\
+  forward and backward intervals.
 \item
@@
 \section{Results}\label{results}
+
+In this section we start with an analysis of the sensitivity of the
+method which was not clearly discussed in the original publication but
+is easily facilitated with our reference implementation. Subsequently,
+we discuss the reproduction of Figures 1 and 2 from the original
+publication.
 
@@
 
-\subsection{\texorpdfstring{Reproduction of Figure 1 in
-\autocite{Shlomi2009}}{Reproduction of Figure 1 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-1-in-shlomi2009}
+\subsection{\texorpdfstring{Reproduction of Figure 1 panel A and B in
+\autocite{Shlomi2009}}{Reproduction of Figure 1 panel A and B in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-1-panel-a-and-b-in-shlomi2009}
 
@@
 
-The original publication's Figure 1 exemplified the biomarker prediction
-method for an illustrative network. It is reproduced here as Figure
-~\ref{fig:Shlomi2009_figure_1}. The simple nature of the example serves
-to explain the reasoning behind the method and also helped us validate
-our implementation of the method. This repository provides an SBML file
-of the network, the Jupyter notebook that generates it and a Jupyter
-notebook reproducing the biomarker prediction.
+The original publication's Figure 1, panel A and B, exemplified the
+biomarker prediction method for an illustrative network. It is
+reproduced here as Figure ~\ref{fig:Shlomi2009_figure_1}. The simple
+nature of the example serves to explain the reasoning behind the method
+and also helped us validate our implementation of the method. This
+repository provides an SBML file of the network, the Jupyter notebook
+that generates it and a Jupyter notebook reproducing the biomarker
+prediction.
 
@@
 Shlomi et al. graphically provide the biomarker prediction for this
-network in their Figure 1. Here, we do so numerically and this paper is
+network in their Figure 1A. Here, we do so numerically and this paper is
 accompanied by a Jupyter notebook reproducing the biomarker prediction.
@@
 Alkaptonuria & HGNTOR & L-Tyrosine &\tabularnewline
-Arginase deficiency & ARGN & &\tabularnewline
+Argininemia & ARGN & &\tabularnewline
 Cystinuria & CYSTSERex, & &\tabularnewline
@@
 L-Tyrosine\tabularnewline
-Tetrahydrobiopterin & DHPR & &\tabularnewline
-deficiency & & &\tabularnewline
-Tyrosinemia type 1 & FUMAC & L-Tyrosine &\tabularnewline
+Phenylketonuria II & DHPR & &\tabularnewline
+Tyrosinemia type I & FUMAC & L-Tyrosine &\tabularnewline
 Tyrosinemia type III & PPOR, 34HPPOR & L-Tyrosine &\tabularnewline

### 2018-01-25 article/Caze_2018.tex
rename .tex file and correct the README

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
		Romain Cazé,
		Marcel Stimberg,
		Benoît Girard
  },
            pdftitle=
{
[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\usepackage{siunitx}
\newcommand{\blau}[1]{{\color{blue} #1}}
\newcommand{\rot}[1]{{\color{red} #1}}

\lstnewenvironment{code}{\lstset{language=Python,basicstyle=\small\ttfamily,showstringspaces=false,stringstyle=\color{blue},commentstyle=\color{blue},columns=spaceflexible}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{biblio.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Jan 2018 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks}

\usepackage{authblk}
                \author[1]{Romain Cazé}
                \author[2]{Marcel Stimberg}
                \author[1]{Benoît Girard}
                \affil[1]{Sorbonne Université, ISIR \'equipe AMAC, Paris, France}
                \affil[2]{Sorbonne Université, INSERM, CNRS, Institut de la Vision, Paris, France}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:romain.caze@upmc.fr}{romain.caze@upmc.fr}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

% --- Additional packages ----
% Package for subfigures
\usepackage[position=top,labelformat=simple,labelfont=bf]{subcaption}
\renewcommand*{\thesubfigure}{\Alph{subfigure}} % Upper case letters for subfigures

%% Useful packages
\usepackage{amsmath}
\usepackage[colorinlistoftodos]{todonotes}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Month, day, year\\
  {\bfseries Accepted}  Month, day, year\\
  {\bfseries Published} Month, day, year\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks, R.M. Memmesheimer, M. Timme, PLoS Computational Biology, 8(4): e1002384, 2012. DOI: 10.1371/journal.pcbi.1002384
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section*{Introduction}
Dendritic non-linearities increase neurons' computation capacity, turning them into complex computing units \cite{Poirazi2003a}. However network studies are usually based on point neuron models that do not incorporate dendrites and their non-linearities. In contrast, the study replicated here \cite{Memmesheimer2012} uses a simple point-neuron model that contains an effective description of dendrites by a non-linear summation of its excitatory synaptic input. Due to the simplicity of the model, both large-scale parameter exploration of a medium-sized network, as well as an analytical investigation of its properties are feasible.

The original study was based on simulation and analysis code in C and Mathematica, but this code is not publicly available. Here, we replicate the study using the neural simulator Brian~2 \cite{Goodman2009,Stimberg2014}, a simulator based on the Python language that has become a common choice in computational neuroscience \cite{Muller2015}. This simulator offers a good trade-off between flexibility and performance and is therefore a suitable choice for this study of a non-standard neuron model.

\section*{Methods}
\subsection*{Neural network model}
The simulations in the original paper were done in phase representation, but our implementation uses the more standard representation in terms of the neurons' membrane potential. The membrane potential dynamic $V_i$ (measured relative to the resting potential) of neuron $i$ is described by (a slightly rewritten version equivalent to equation (1) of \cite{Memmesheimer2012}):

\begin{equation}
\tau_m \frac{dV_i(t)}{dt} = - V_i(t) + V_0 + \sigma \big( \sum_{j\in E_i} w_\text{ex} \epsilon_{j}(t-\tau) \big) - \sum_{j \in I_i} w_\text{in}\epsilon_{j}(t-\tau),
\end{equation}

where $V(t)$ is the membrane voltage at time $t$, $\tau_m$ the membrane time constant, $V_0$ the displacement of the membrane potential due to a constant external input, $w_\text{ex}$ and $w_\text{in}$ respectively the excitatory and inhibitory weights, $\tau$ the transmission delay, and $E_i$ and $I_i$ the indices of neurons that connect with respectively excitatory and inhibitory weights to neuron $i$. If the membrane potential crosses a threshold $\Theta$, it emits the spike and its membrane potential is reset to its resting potential, i.e. $\SI{0}{\milli\volt}$. The function $\epsilon_{j}$ is 1 whenever neuron $j$ spikes, and 0 otherwise

The originality of this model lays in a filtering of the excitatory inputs by the function $\sigma$ to model dendritic non-linearities. The study considers two functions $\sigma$ (see Fig.~1 of the original paper and insets in Fig.~\ref{fig:markov}). Linear coupling, defined by $\sigma(x)=x$, and non-linear coupling with:
\[
    \sigma(x)= 
\begin{cases}
    x,              & \text{if } x\leq V_a\\
    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{if } V_a<x \leq V_b\\
    V_c,            & \text{otherwise}
\end{cases}
\]

For this non-linear function, the biological interpretation is that below a first threshold $V_a$, excitatory inputs are propagated passively by the dendrite and sum linearly. Above this threshold, they trigger a dendritic spike, resulting in an amplified somatic voltage. If they are even higher and cross the threshold $V_b$, they stop increasing the somatic voltage and saturate. Note that this non-linearity only affects spikes received synchronously as has been observed experimentally. With the introduction of this non-linearity, the modelled point neuron becomes an effective model of a neuron with one non-linear dendrite.

The network consists of $N$ neurons, where a directed synaptic connection between a pair of neurons is established with probability $p_0$. Each of these connections is taken to be excitatory with probability $p_\text{ex}$, and inhibitory otherwise. Note that this connection scheme implies that neurons cannot be separated into groups of excitatory and inhibitory neurons, a neuron typically projects with both excitatory and inhibitory connections to other neurons.

The parameter values used in the original study as well as in our replication are summarized in Table~\ref{tab:parameters}.

\begin{table}[ht]
\begin{tabular}{ l  c  S p{2.85cm}}
  \toprule
  {Parameter} & {Symbol} & {Value} &\\
  \midrule
  Number of neurons & $N$ & 1000 &\\
  Connection probability & $p_0$ & 0.3 &\\
  Probability connection being excitatory & $p_\text{ex}$ & 0.5 &\\
  Excitatory connection weight & $w_\text{ex}$ & \SI{0.2}{\milli\volt} & (varied in Fig.~\ref{fig:grid})\\
  Inhibitory connection weight & $w_\text{in}$ & \SI{0.2}{\milli\volt} & (varied in Fig.~\ref{fig:grid})\\
  Membrane time constant & $ \tau_m $ & \SI{8}{\milli\second} &\\
  Firing threshold & $\Theta$ & \SI{16}{\milli\volt} &\\
  Synaptic delay & $\tau$ & \SI{5}{\milli\second} & (but see Implementation section)\\
  Threshold for supra-linear summation & $V_a$ & \SI{2}{\milli\volt} \\
  Threshold for saturation & $V_b$ & \SI{4}{\milli\volt} \\
  Saturation value & $V_c$ & \SI{6}{\milli\volt} \\
  \bottomrule
\end{tabular}
\caption{Parameter values used in the simulations}\label{tab:parameters}
\end{table}

\subsection*{Implementation and numerical methods}
The authors of the original study provided us with the principal C and Mathematica code used in their study. However, this code does not compile and run any more with recent Mathematica versions. Our implementation is not based on this code, but only on the descriptions in the original paper (with minor clarifications provided by the authors where necessary).

The authors of the original study used an event-based simulation scheme, exploiting the linearity of the sub-threshold dynamics. Due to this linearity, the membrane potential can be advanced for arbitrarily large time spans between spikes, and spike times do not have to be bound to any temporal grid. In contrast, the Brian simulator is built on a clock-driven paradigm where time is advanced in discrete steps. This is because the simulator supports a wide range of neural models of and event-driven methods can only be applied to a small number of simple neuron models, and also becomes computationally demanding in large networks with many spikes.

All simulations presented here have been performed with a step size of \SI{0.1}{\milli\second} and the forward Euler integration algorithm (changing to an exact integration based on the analytical solution of the linear equations did not change the results).

Note that the event-based method used in the original study has important consequences for the specific implementation of non-linear synaptic summation: only spikes that arrive exactly synchronously can sum up non-linearly. This, together with the synaptic delay that is constant over all connections, implies that the non-linearity will only apply to spikes that originate from neurons that spiked perfectly synchronously one synaptic delay earlier; randomly occurring spikes have only an infinitesimally small chance to arrive at the receiving neuron synchronously. When using a clock-driven algorithm, as we do in this replication, this is no longer the case. All neurons that spike within the same time step will deliver their spikes to their target neurons at the same time step, potentially triggering non-linear summation. While the clock-driven method is a less accurate method in general, we would argue that in the context of non-linear summation it is actually closer to the biological phenomenon it models. While dendrites detect synchronous events with sub-millisecond precision \cite{Softky1994}, this precision is of course finite. With that said, the qualitative results of the original study can be replicated with a clock-driven method, as we will show in the Results section.

For linearly coupled networks, the neuron model is a standard leaky-integrate-and-fire neuron with ``delta synapses'', i.e. incoming synaptic events instantaneously affect the membrane potential as soon as they are received. In Brian, this model is described as follows (omitting the code to connect the synapses and set their weights):
\begin{code}
eq = "dV/dt = -gamma*V + V0 : volt"
G = br2.NeuronGroup(N, model=eq, threshold="V>Theta",
                    reset="V=0*mV", method='euler')
exc_syn = br2.Synapses(G, G, 'w : volt (constant, shared)',
                       on_pre='V += w', delay=5 * ms - dt)
inh_syn = br2.Synapses(G, G, 'w : volt (constant, shared)',
                       on_pre='V -= w', delay=5 * ms - dt)
\end{code}

Note that the delay has been set as one time step (\verb|dt|) less than \SI{5}{\milli\second}. This is because in Brian, when a neuron receives synaptic input, this input can only trigger a threshold crossing of the receiving neuron in the following time step. To get chains of propagated synchronous spikes every \SI{5}{\milli\second} as in the original study, we therefore have to reduce the delay by one time step.

For non-linearly coupled networks, we have to use a slightly more complicated approach where we sum up all the received excitatory input in a variable \verb|ve|, apply the non-linear function $\sigma$ to this input, and then reset the variable \verb|ve| back to 0 to prepare for the next time step. In Brian syntax (the definition of the inhibitory synapses is identical to the linearly-coupled network):
\begin{code}
eq = """dV/dt = -gamma*V + V0 : volt
        ve : volt
     """
G = br2.NeuronGroup(N, model=eq, threshold="V>Theta",
                    reset="V=0*mV", method='euler')
exc_syn = br2.Synapses(G, G, 'w : volt (constant, shared)',
                       on_pre='ve += w', delay=5 * ms - dt)
G.run_regularly('''
                V += clip(ve, 0*mV, 2*mV) + clip(2*(ve-2*mV), 0*mV, 4*mV)
                ve = 0*mV
                ''', when='after_synapses')
\end{code}
The \verb|run_regularly| operation that is run every time step to implement the non-linear $\sigma$ function refers to the predefined \verb|clip| function which clips the value to a given range, following the syntax
\begin{code}
clip(value, lower_bound, upper_bound)
\end{code}

To facilitate simulation, we use the \emph{joblib} library to automatically store results and to distribute computation over processor cores. Several simulation and analysis functions are annotated with a \verb|@mem.cache| decorator, meaning that their input arguments and outputs will be automatically stored to disk and calling the function will return the stored values if available. This feature does not only ease development (e.g. plotting functions can be changed without re-running the underlying simulations), but also allows to interrupt long-running parameter explorations without losing the results obtained so far. The same result could be achieved by manually storing results to disk, but joblib's caching mechanism helps to avoid common errors (e.g. to re-use an old result obtained with an outdated version of the code) while keeping the code readable. To reduce the total simulation time, the grid exploration in Fig.~\ref{fig:grid}, and the analysis of group size evolution in Fig.~\ref{fig:markov} have been distributed over processor cores using joblib's \verb|Parallel| construct. For the simulations presented in Fig.~\ref{fig:grid}, multiple repetitions for the same combination of synaptic weights are distributed over processor cores, for the simulations presented in Fig.~\ref{fig:markov}, simulations are evaluated in parallel over group sizes. In both cases, a number of $C-1$ parallel processes is used, where $C$ is the number of processor cores in the user's system.

We wanted reproducible simulations despite the use of random connectivity and initial conditions. Setting a single global seed was not sufficient because we used parallel simulations executed in non-deterministic order. Instead, we used a set of pre-determined seeds, one for each figure, and used those ``meta-seeds'' to generate a set of random seeds for each individual simulation.

For the grid search presented in Fig.~\ref{fig:grid}, we varied inhibitory and excitatory synaptic weights between \SI{0.16}{\milli\volt} and \SI{0.4}{\milli\volt} (given that each neuron receives on average 150 excitatory and 150 inhibitory inputs, this corresponds to a ``total weight'' between \SI{24}{\milli\volt} and \SI{60}{\milli\volt}, as used on the axes of Fig.~\ref{fig:grid}), using 100 steps for each variable (comparable to the 96 steps used in the original publication). Each network was simulated 10 times with different initial conditions (synaptic connections and initial membrane potential values), whereas the original publication used 20 simulations per weight combination and additionally added ``1--50 random spikes initially in transit''.

To numerically derive the evolution of synchronous pulse sizes presented in Fig.~\ref{fig:markov} (for the semi-analytical approach, see the next section), we performed simulations of \SI{56}{\milli\second} each. During each simulation, we stimulated the network at 50ms with a synchronous pulse of activity, consisting of between 1 and 176 neurons (varied in steps of 5 -- the original study used group sizes between 1 and 181 varied in steps of 6). We then observed the number of active neurons at \SI{55}{\milli\second}, i.e. at one $\tau$ after the stimulation. For each stimulation group size, we ran the simulations 50 times with different initial conditions, consistent with the original study.

\subsection*{Semi-analytical and analytical methods}
We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}).
To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows. First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.  We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential. From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
\begin{equation}
F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
\end{equation}
Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e. $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$. Let us assume that in the last iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e. that $N-g_i$ neurons are available to spike in this iteration. To calculate the probability for a single (non-refractory) neuron $j$ to spike in response to these $g_i$ neurons, we will first calculate the probability that out of the $g_i$ neurons that spiked, exactly $n_\text{ex}$ neurons are connected with excitatory connections and $n_\text{in}$ neurons are connected with inhibitory connections to neuron $j$. There are $\binom{g_i}{n_\text{ex} + n_\text{in}}$ possibilities to chose $n_\text{ex} + n_\text{in}$ out of $g_i$ neurons, and $\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}}$ possibilities to divide these neurons into groups of size $n_\text{ex}$ and $n_\text{in}$. This yields the total number of combinations $c(g_i, n_\text{ex}, n_\text{in})$:
\begin{equation}
c(g_i, n_\text{ex}, n_\text{in}) = \binom{g_i}{n_\text{ex} + n_\text{in}}\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}} = \frac{g_i!}{n_\text{ex}!n_\text{in}!(g_i - n_\text{ex} - n_\text{in})!}
\end{equation}
The probability that these selected $n_\text{ex}$ neurons are actually connected via excitatory connections to neuron $j$, that the $n_\text{in}$ neurons are actually connected via inhibitory connections to neuron $j$, and that the remaining $g_i - n_\text{ex} - n_\text{in}$ neurons are \emph{not} connected to neuron $j$ is:
\begin{equation}
p(g_i, n_\text{ex}, n_\text{in}) = (p_0 p_\text{ex})^{n_\text{ex}} (p_0 p_\text{in})^{n_\text{in}} (1 - p_0)^{g_i - n_\text{ex} - n_\text{in}}
\end{equation}
Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike. This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
\begin{equation}
P^s(g_i) = \sum_{n_\text{ex}=1}^{g_i}\sum_{n_\text{in}=0}^{g_i-n_\text{ex}} F\big(\epsilon(n_\text{ex}, n_\text{in})\big) c(g_i, n_\text{ex}, n_\text{in}) p(g_i, n_\text{ex}, n_\text{in})
\end{equation}
In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, i.e.:
\begin{equation}
E(g_{i+1}|g_i = g_i') = (N - g_i') P^s(g_i')
\end{equation}

\section*{Results}
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:raster}
\textbf{Propagation of synchronous spiking in linearly (A) and non-linearly (B) coupled networks.} We initiate a chain of synchronous pulses by applying external supra-threshold inputs to the first 100 neurons at time $t=\SI{150}{\milli\second}$ (red coloured spikes, grey vertical dotted lines indicate times where spikes occur as part of the chain). Top row: total size $g'$ of synchronized groups within the chain. Middle row: network rate over all neurons (rate in \si{\kilo\hertz}, bin size \SI{1}{\milli\second}). Bottom row: spiking activity of the first 200 neurons in a network of a 1000 neurons versus time.\\
Replication of Fig.~2 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g002}).}
\end{figure}

We first replicate a simulation showing the propagation of synchronous spiking activity in networks using linear (Fig.~\ref{fig:raster}A) and non-linear (Fig.~\ref{fig:raster}B) coupling. This figure corresponds to Fig.~2 of the original article. Given that the simulations are based on random synaptic connections and random initial conditions, we do not expect exact replication of spike times, but the qualitative features of the simulation are faithfully replicated: synchronous activity quickly dies out and disappears in the background activity for linearly coupled networks (Fig.~\ref{fig:raster}A) but is persistently propagated in non-linearly coupled networks (Fig.~\ref{fig:raster}B).

To confirm the equivalence of our network's behaviour with the original study over a wide range of synaptic connection strength we ran a grid search exploration for linearly (Fig.~\ref{fig:grid}A) and non-linearly (Fig.~\ref{fig:grid}B) coupled networks. Network behaviour is colour-coded as being part of one out of four classes: stable activity, but no persistent propagation (green); stable activity with persistent propagation (blue); unstable activity after synchronous stimulation (yellow); unstable activity before synchronous stimulation (red). For quantitative definitions, see Table~\ref{tab:color}. We run each combination of connection strengths 10 times, and combine the results to yield a color in RGB notation (values between 0 and 1 for the red, green, and blue component), according to:
\begin{equation}
R = U_1 + U_2;\quad G = E + U_2;\quad B = S,
\end{equation}
where $U_1$, and $U_2$ stand for the fraction of trials displaying unstable network activity before respectively after onset of propagation, and $S$ and $E$ for the fraction of trials displaying stable background activity with respectively without persistent propagation. We obtained this scheme directly from the authors of the original study, it was not mentioned in the publication. Combining individual results in this way does not necessarily lead to easily interpretable colours, but using a different scheme, e.g. colouring according to the class that occurred the most often, leads to visually similar results (not shown).

Our replication does not reproduce all details of the original plots (in particular at the top of the plots), but looks very similar on a coarse scale. Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not. The minor differences in our replication could be due to either our use of a clock-based simulation method, or to the slightly different initial conditions (no spikes ``in transit'').
\begin{table}
\begin{tabular}{ c  l  p{0.32\textwidth}}
  \toprule
   & Class & Definition \\
  \midrule
  $U_1$ (red) & Unstable before stimulation & $\exists t<t_\text{stim}: a_b(t) \geq 100$ \\
  $U_2$ (yellow) & Unstable after stimulation & $\exists t>t_\text{stim}: a_b(t) \geq 100$ \\
  $E$ (green) &  Stable activity, no propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} < 10$\\
  $S$ (blue) & Stable activity, persistent propagation & $\forall t: a_b(t) < 100\, \wedge\, g_\text{p} \geq 10$\\
  \bottomrule
\end{tabular}
\caption{Classification of network behaviour for Fig.~\ref{fig:grid}. Here, $a_b$ refers to the background activity, the number of neurons active during a \SI{1}{\milli\second} time bin, ignoring the time steps at $t_\text{p}^k = t_\text{stim} + k\cdot\tau$, i.e. time steps that follow the synchronous stimulation by multiples of the synaptic delay. The variable $g_\text{p}$ refers to the number of successful propagations of the synchronous activity, i.e. the number of time steps $t_\text{p}^k$ where number of active neurons exceeds the maximal background activity $a_b$.}\label{tab:color}
\end{table}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_mean.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_mean.pdf}
\end{subfigure}
\caption{\label{fig:grid}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B) coupled networks.} For each combination, we initiated synchronous activity with 100 neurons and assessed the stability of the temporal evolution. Blue colouring indicates stable propagation of synchrony, red indicates unstable background activity before onset of propagation, yellow indicates unstable background activity after onset of propagation, and green colouring indicates the absence of propagation (see Methods for details).\\
Replication of Fig.~3a and b in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g003})}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks.} To numerically estimate the probability distribution, we show occurrences of pulse-sizes $g'_1$ in response to a pulse of size $g'_0$ by grey lines; green squares represent the corresponding mean response group sizes.\\
Replication of Fig.~4 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g004})}
\end{figure}

Finally, we looked at the evolution of synchronous group sizes for one combination of synaptic weights ($w_\text{ex}=w_\text{in}=\SI{0.2}{\milli\volt}$) in detail (Fig.~\ref{fig:markov}). The results confirms our previous conclusion that we can faithfully replicate the network behaviour. They match both qualitatively and quantitatively, the group size consistently declines with each iteration in linearly coupled networks (Fig.~\ref{fig:markov}A), whereas it increases for intermediate group sizes and declines otherwise in non-linearly coupled networks (Fig.~\ref{fig:markov}B). While the semi-analytic solution (red dotted line) shows a good match to the numerical results (grey lines, means as green dots), it shows some quantitative differences to the semi-analytic solution presented in Fig.~4 of the original article. In particular, our semi-analytical solution seems to mostly slightly underestimate the size of the group in the next iteration while the semi-analytical solution presented in the original article seems to slightly overestimate it.

\section*{Discussion}

We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences. These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision. However, we tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained similar results (not shown). Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.

The reusable code we provide with this work paves the way to future studies of networks of non-linearly coupled integrate-and-fire neurons. It gives researchers the means to verify that the original study's conclusion hold under variations of its model assumptions, such as the details of the connectivity scheme or the synaptic model. This article has already provided a first such verification, demonstrating that the original studies main findings do not critically depend on the event-based method and the entailing ``perfect time resolution''. Finally, we hope that the provided code can also serve as a basis for future studies of functional paradigms going beyond the mere propagation of synchronous activity.

\section*{Acknowledgements}
We would like to thank Raoul Martin Memmesheimer for providing implementation details about the original study. This work has received funding from the European Union’s Horizon 2020 research and innovation programme under grant agreement No 640891 (DREAM Project).

{\sffamily \small
   \printbibliography[title=References]
%}

\end{document}

### 2018-01-31 article/Name-YEAR.tex
add Literature and first changes

--- 2017-10-26 article/Mondeel_Ogundipe_Westerhoff-2017.tex
@@
 {
-      Thierry D.G.A. Mondeel,
-      Vivian Ogundipe,
-      Hans V. Westerhoff,
+      René Larisch,
+      Name Surname,,
   },
@@
 {
-[Re] Predicting metabolic biomarkers of human inborn errors of metabolism
+[Re] This is the title
 },
@@
 \bibliography{bibliography.bib}
-
-\usepackage{amsmath,amssymb}
 
@@
 % --- Document ----------------------------------------------------------------
-\title{[Re] Predicting metabolic biomarkers of human inborn errors of metabolism}
+\title{[Re] This is the title}
 
     \usepackage{authblk}
-                        \author[1,*]{Thierry D.G.A. Mondeel}
-                    \author[1]{Vivian Ogundipe}
-                    \author[1, 2, 3]{Hans V. Westerhoff}
-                            \affil[1]{Synthetic Systems Biology and Nuclear Organization, Swammerdam Institute
-for Life Sciences, University of Amsterdam, The Netherlands}
-                    \affil[2]{Molecular Cell Physiology, Amsterdam Institute for Molecules, Medicines
-and Systems, VU University Amsterdam, The Netherlands}
-                    \affil[3]{Manchester Centre for Integrative Systems Biology, School of Chemical
-Engineering and Analytical Science, The University of Manchester, UK}
+                        \author[1]{René Larisch}
+                    \author[2, 3]{Name Surname,}
+                            \affil[1]{Affiliation Dept/Program/Center, Institution Name, City, State, Country}
+                    \affil[2]{Affiliation Dept/Program/Center, Institution Name, City, State, Country}
+                    \affil[3]{Affiliation Dept/Program/Center, Institution Name, City, State, Country}
 
 \date{\vspace{-5mm}
-      \sffamily \small \href{mailto:d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}{d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}}
+      \sffamily \small \href{mailto:corresponding-author@mail.com}{corresponding-author@mail.com}}
 
@@
 \begin{itemize}
-    \item[→] T. Shlomi, M.N. Cabili, E. Ruppin, Predicting metabolic biomarkers of
-human inborn errors of metabolism, Mol. Syst. Biol. 5 (2009) 263.
-doi:10.1038/msb.2009.22.
+    \item[→] Original article (title, authors, journal, doi)
   \end{itemize}\par
@@
 
-Shlomi et al. \autocite{Shlomi2009} have developed a way in which flux
-variability analysis \autocite{Mahadevan2003} is used to predict
-biomarkers for inborn errors of metabolism (IEMs). The approach focuses
-on exchange reactions in a metabolic network that connect extracellular
-metabolites that reside in a medium, which might correspond to the
-blood, to a reservoir such as the urine. The method then determines the
-ranges of exchange fluxes that are compatible with constraints set by
-the network's topology, flux bounds on reactions and the requirement of
-steady state for each metabolite. Shlomi et al. proposed that if any
-such ranges shift convincingly when the network contains a mutation, the
-concentration ranges of the corresponding metabolites in the reservoir
-should shift in parallel and these metabolites would therefore be
-biomarkers.
-
-Shlomi et al. illustrated the method first on a virtual metabolic map
-and further validated their technique on the reconstruction of the human
-metabolic map Recon1 \autocite{Duarte2007} emperically with respect to
-known biomarkers, but did not give any formal proof. For a set of 17
-IEMs that affect amino-acid metabolism they found a 56\% fraction of the
-biomarkers that are correctly predicted and a fraction of 76\% of
-predicted biomarkers that were predicted correctly in terms of whether
-they increased or decreased in concentration. Thiele et al.
-\autocite{Thiele2013} later found an even higher precision using
-essentially the same FVA methodology on an updated version of the
-metabolic reconstruction.
-
-The source code is not publicly available and the computation
-environment used was not mentioned in the publication. The
-\href{https://github.com/opencobra/cobratoolbox}{MATLAB COBRA toolbox}
-contains code to reproduce the analysis by Thiele et al. but this was
-also performed in MATLAB.
-
-We here present an implementation of the biomarker prediction method
-originally proposed by Shlomi et al., programmed in Python, and
-reproduce the results presented in Figure 1 and 2 of the original
-publication. These figures concern the method's application to a simple
-metabolic map and then to a set of \(17\) amino acid metabolism
-disorders. Our Python module provides the biomarker prediction algorithm
-and is agnostic of the map used. It can therefore be applied to other
-maps than the human metabolic reconstruction considered by Shlomi et al.
-Jupyter notebooks are provided that generate the illustrative map and
-reproduce all results referred to in the Figures 1 and 2 of Shlomi et
-al. Our implementation is based on information given in the original
-paper. However, some details critical to the reproduction were left
-unexplained there and are highlighted here. We also include a more
-detailed discussion of one of the amino acid IEMs to examine the
-robustness of the predictions made using this approach.
+The introduction should introduce the original paper and put it in
+context (e.g.~is it an important paper in the domain ?). You must also
+specify if there was an implementation available somewhere and provide a
+link to it if relevant (and in such a case, you have to specify if the
+proposed replication is based on this original implementation). You
+should also introduce your implementation by listing language, tools,
+libraries, etc. and motivate choices if relevant.
 
@@
 
-\subsection{Overview}\label{overview}
+The methods section should explain how you replicated the original
+results:
 
-We implemented the method of Shlomi et al. in Python. Our Python
-function has several input parameters most of which are optional in
-order to enable the user to perturb the network in various ways and to
-try variations of the method that differ from the ones implemented here
-or in Shlomi et al. The default settings are such that they mimic the
-approach Shlomi et al. originally proposed. All flux-variability
-analyses are performed using COBRApy \autocite{Ebrahim2013}.
-
-All simulations discussed here were performed in Python 3.6 using
-COBRApy \autocite{Ebrahim2013} (version 0.9) and Pandas (version
-0.20.3). This repository also functions in conjunction with
-\href{http://mybinder.org}{MyBinder} which allows for full
-reproducibility in the `cloud' without any need for installation of
-additional software.
-
-The rest of this section introduces the approach proposed by Shlomi et
-al. as we implemented it.
-
-\subsection{Definitions}\label{definitions}
-
-Exchange reactions are reactions in the metabolic network that allow
-import and/or export of metabolites from the metabolic map. Exchange
-reactions are represented by non-mass-balanced pseudoreactions, e.g.
-\(X \leftrightarrow \varnothing\), where \(X\) is understood to
-originate from or efflux to the biofluids. Positive flux indicates net
-secretion of \(X\) by the cell and negative flux indicates the net
-uptake of \(X\).
-
-We define boundary metabolites to be metabolites that are involved in
-exchange reactions with the environment, i.e.~they serve as inputs
-and/or outputs for the metabolic network, e.g.~glucose and CO\(_2\).
-These metabolites typically exist both inside and outside the cell.
-
-Each boundary metabolite will be associated with the range of flux its
-exchange reaction supports under the following constraints for the flux
-distributions in the model: (i) network topology, (ii) mass-balance for
-all metabolites, (iii) flux bounds of reactions in the network.
-Optionally, there is an additional constraint of being optimal with
-respect to attaining or maximizing a certain set of fluxes, the
-``objective'', in the model.
-
-Biomarkers are those boundary metabolites the range of exchange flux of
-which differs between the wild-type (WT) and mutant simulations. Shlomi
-et al. proposed a threshold of at least 10\% difference in the lower or
-upper bounds of the wild-type vs.~the mutant flux intervals.
-
-Each biomarker will be associated with a prediction for being either
-elevated or decreased in the mutant case, as comparted to the wild-type.
-A biomarker is considered to have an increased, or reduced,
-extracellular concentration in the mutant case if, when plotting the
-wild-type and mutant flux-variability (see below) intervals on a
-horizontal axis, both the lower and upper boundary of the mutant
-interval are shifted to the right, or left, respectively. If the two
-borders of the mutant interval move in opposite directions with respect
-to the wild-type interval the result is scored as `unchanged' and the
-boundary metabolite is not considered to be a biomarker.
-
-\subsection{Flux variability
-calculations}\label{flux-variability-calculations}
-
-Flux ranges for boundary metabolites are calculated by applying flux
-variability analysis \autocite{Mahadevan2003}. Mathematically, flux
-variability analysis for a specific exchange reaction \(v_i\) entails
-the following linear programming problem:
-
-\begin{align}
-\text{Maximize } &\pm v_i, \text{ such that for all k} \nonumber \\
-S v &= 0 \nonumber \\
-Z  &\geq \phi Z_{\text{opt}} \nonumber \\
-\alpha_k \leq & ~ v_k \leq \beta_k 
-\end{align}
-
-Here \(v\) is the column vector of fluxes representing all reactions in
-the model, \(v_i\) is the flux of the exchange reaction of a given
-boundary metabolite \(i\), \(\alpha_k\) is the, possibly negative, lower
-bound for reaction \(k\) and similarly \(\beta_k\) is the, possibly
-negative, upper bound for reaction \(k\). \(Z = c^T ~ v\) is the
-objective function defined for the map, entailing a linear combination,
-defined by column vector c, of one or more fluxes, and
-\(Z_{\text{max}}\) its maximal value. We define \(0 \leq \phi \leq 1\)
-to be an arbitrary minimal fraction of the maximal objective function
-value that has to be achieved. When \(\phi=1\) we ask for the range of
-flux allowed for reaction \(v_i\) while maintaining the maximal value of
-the linear combination of objective fluxes. When setting \(\phi = 0\),
-one is asking for the allowable flux range through reaction \(v_i\)
-regardless of the value of the objective function (although the sign has
-to be maintained). The latter, i.e. \(\phi=0\), is the case considered
-in the method by Shlomi et al.
-
-We here entertain the inborn errors of metabolism (IEM) due to
-loss-of-function mutations in a single enzyme \(r\) that may catalyse
-multiple chemical reactions. For IEMs that disrupt the activity of
-several reactions Shlomi et al. proposed applying the approach to each
-reaction individually and then combining the list of biomarkers.
-
-Therefore, for a given IEM we loop the following sequence of steps for
-each affected reaction \(r\) and for each boundary metabolite \(m\):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
+\begin{itemize}
 \tightlist
 \item
-  Compute the exchange flux interval of \(m\), when \(r\) is forced to
-  be active with a flux of at least \(\epsilon \geq 0\). By default
-  \(\epsilon = 1\). We denote this the forward wild-type interval
-  \(WT_{r,m}^+\).
+  did you use paper description
 \item
-  For reversible reactions \(r\), if the forward interval yields a flux
-  range of \([0,0]\) constrain the flux to be negative and lower than
-  \(-\epsilon \leq 0\), and compute the backward wild-type interval
-  \(WT_{r,m}^-\).
+  did you contact authors ?
 \item
-  The wild-type interval \(WT_{r,m}\) is given by the union of the
-  forward and backward intervals.\\
+  did you use original sources ?
 \item
-  Compute the exchange flux interval of \(m\) when \(r\) is forced to be
-  inactive by temporarily setting the lower and upper bound of \(r\)
-  equal to zero. We denote this interval \(M_{r,m}\).
+  did you modify some parts ?
 \item
-  Compare the wild-type and mutant intervals and predict whether the
-  external concentration of \(m\) is increased, decreased or unaltered
-  when reaction \(r\) is inactivated. This is done by the following
-  rule: if both boundaries are higher in the mutant than the
-  corresponding boundaries in the wild-type, then \(m\) is predicted to
-  be a biomarker and elevated. If both boundaries are lower in mutant as
-  compared to wild type then the metabolite is predicted to be a
-  biomarker and reduced. If the predicted changes are smaller than 10\%
-  \(m\) is a low confidence biomarker. When the wild-type and mutant
-  intervals are disjoint \(m\) is a highly confident (H.C) biomarker.
-\end{enumerate}
+  etc.
+\end{itemize}
 
-After the loop is completed, contradictory predictions for \(m\) between
-the affected reactions \(r\) are dealt with based on a majority rule.
-When there is an equal number of elevated and reduced predictions the
-boundary metabolite is considered to be unchanged, i.e.~not a biomarker.
+If relevevant in your domain, you should also provide a new standardized
+description of the work.
 
@@
 
-\subsection{Robustness of the flux variability
-method}\label{robustness-of-the-flux-variability-method}
+Results should be compared with original results and you have to explain
+why you think they are the same or why they may differ (qualitative
+result vs quantitative result). Note that it is not necessary to redo
+all the original analysis of the results.
 
-Shlomi et al. did not discuss in detail the effects of changing various
-parameters settings of, and assumptions inherent to, their FVA-based
-method, nor did they present a proof of any generality. Because our
-Python implementation allows the user to set various parameters it
-becomes rather easy to illustrate how specific settings in the approach
-proposed by Shlomi et al. affect the outcome of the simulations. An
-example of the potential pitfalls and issues with this method is
-considered below and looks at predictions for the inborn error of
-metabolism known as Phenylketonuria. We here define this disease by one
-of its potential causes, i.e.~dysfunctionality of the enzyme converting
-phenylalanine into tyrosine. PKU patients are known for elevated levels
-of phenylalanine and decreased levels of tyrosine in their serum. We
-illustrate this briefly in an example also considered in Figure 2 of
-Shlomi et al. In Figure ~\ref{fig:PKU_network} we sketched some
-essential features of the metabolic network surrounding the
-phenylalanine hydroxylase (PAH) enzyme that is malfunctioning in PKU.
-Unless stated otherwise, the simulations are performed for a medium that
-allows efflux through all exchange reactions with an upper bound of 1000
-and influx through all exchange reactions with a lower bound of 1 and we
-use the default settings, including a 10\% significance threshold). The
-only considered biomarkers are the \(20\) proteinogenic amino acids. In
-total we will consider six cases. The results are summarized in Table
-~\ref{tbl:PKU_results}.
+\section{Conclusion}\label{conclusion}
+
+Conclusion, at the very minimum, should indicate very clearly if you
+were able to replicate original results. If it was not possible but you
+found the reason why (error in the original results), you should exlain
+it.
+
+\hypertarget{tbl:table}{}
+\begin{longtable}[]{@{}llllll@{}}
+\caption{\label{tbl:table}Table caption }\tabularnewline
+\toprule
+Heading 1 & & & Heading 2 & &\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Heading 1 & & & Heading 2 & &\tabularnewline
+\midrule
+\endhead
+cell1 row1 & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 &
+cell6 row 1\tabularnewline
+cell1 row2 & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 &
+cell6 row 2\tabularnewline
+cell1 row3 & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 &
+cell6 row 3\tabularnewline
+\bottomrule
+\end{longtable}
+
+A reference to table ~\ref{tbl:table}. A reference to figure
+~\ref{fig:logo}. A reference to equation ~\ref{eq:1}. A reference to
+citation \textcite{markdown}.
 
@@
 \centering
-\includegraphics{./Figures/PKU_network.pdf}
-\caption{Scheme of the network surrounding phenylalanine hydroxylase
-(PAH) and its cofactors. The 4a-hydroxytetrahydrobiopterin produced by
-PAH may spontaneously or when catalysed by PCBD1 (EC 4.2.1.96) dehydrate
-to dihydrobiopterin which may be enzymatically reduced back by NADPH +
-\(H^+\) to tetrahydrobiopterin through DHPR (EC 1.5.1.34). The solid
-black lines indicate reactions that are contained in both Recon 1 and
-Recon 2.2 \autocite{Swainston2016}. The solid green line refers to the
-alternative PAH reaction that includes the spontaneous dehydradation
-towards dihydrobiopterin that is only present in Recon 2.2. Dashed lines
-indicate sources of the synthesis pathways or terminal products of the
-indicated metabolites. The tyrosine transaminase (TYRTA) uses
-oxoglutarate (AKG) and glutamate (Glu) as
-cosubstrates.}\label{fig:PKU_network}
+\includegraphics{rescience-logo.pdf}
+\caption{Figure caption}\label{fig:logo}
 \end{figure}
 
-Perhaps the most fundamental question is: can changes in network
-topology result in different predictions? The answer is yes. One
-illustration of this is the comparison of predictions between Recon 1
-and Recon 2. As Table ~\ref{tbl:PKU_results} indicates, using the same
-medium and algorithm settings for both Recon 1 and Recon 2.2 the two
-networks led to different predictions of the effect of a defect of
-phenylalanine hydroxylase on tyrosine. The application of the flux
-variability method to Recon 1 correctly predicted both tyrosine and
-phenylalanine as biomarkers and gets the directionality of the change in
-their concentrations right. In contrast, the prediction based on Recon
-2.2 lacks tyrosine as a biomarker.
-
-With ever improving network annotations it becomes feasible to simulate
-the gene knockout directly for most single gene defects leading to
-inborn errors of metabolism. Shlomi et al. directly blocked or activated
-reactions associated with a causal gene for a specific IEM. For enzymes
-partaking in several reactions Shlomi et al. proposed activating and
-blocking all reactions associated with the gene individually and then
-taking the union of their predicted biomarkers. When a biomarker was
-subject to qualitatively different predictions between the
-enzyme-catalyzed reactions encoded by the gene, a majority rule was
-applied. We will refer to this as the asynchronous setting. Blocking all
-affected reactions simultaneously could yield different predictions. We
-will refer to this as the synchronous setting. We added the option to
-choose between the asynchronous and synchronous setting to our
-implementation.
-
-For example, consider a gene that encodes an enzyme that catalyzes two
-chemically different reactions, \(R_1\) and \(R_2\). Using the
-asynchronous setting we would run the algorithm once for \(R_1\) and
-once for \(R_2\) and take the union of all predicted biomarkers.
-Inconsistent predictions between \(R_1\) and \(R_2\) are dealt with
-based on a majority rule (see Methods). In contrast, with the
-synchronous setting, we would run the algorithm once. For the wild-type
-we would simultaneously force flux through both reactions and in the
-mutant we would block both reactions. The synchronous approach simulates
-the complete catalytic inactivity or absence of the enzyme and therefore
-the infeasibility of any reactions it catalyzes. This corresponds to a
-deletion of the gene. The asynchronous approach might correspond to a
-point mutation that changes the specificity of the enzyme.
-
-\hypertarget{tbl:PKU_results}{}
-\begin{longtable}[]{@{}lllllll@{}}
-\caption{\label{tbl:PKU_results}Summary of predictions made using the
-FVA-based biomarker prediction method for the inborn-error of metabolism
-phenylketonuria for two networks (\(R_1\) = Recon 1 and \(R_2\) = Recon
-2), with or without the synchronous setting (indicated by ``sync'' here)
-and with high (maximum influx of 20) or low (no influx)
-hydroxyphenylpyruvate (hpp +/-) in the medium. \(^*\) refers to a
-situation without any tyrosine production in the mutant. \(^{**}\)
-refers to the situation where tyrosine reduction is so small that it
-does not reach the 10\% threshold. `-' indicates no prediction because
-the boundaries did not change from wild-type to mutant. \(\epsilon = 0\)
-refers to the case where we do not force any flux through the affected
-reaction(s) in the wild-type. In all other columns, \(\epsilon = 1\).
-The full output with interval bounds that the method generated can be
-found in the relevant notebook in this repository. }\tabularnewline
-\toprule
-Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
-\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
-\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
-\midrule
-\endhead
-Phenylalanine & Elevated & Elevated & Elevated & Elevated & Elevated &
--\tabularnewline
-Tyrosine & Reduced & - & Reduced & Reduced\(^*\) & Reduced\(^{**}\) &
-Reduced\tabularnewline
-\bottomrule
-\end{longtable}
-
-Returning to the example of phenylketonuria we simulated this for Recon
-2.2, using the synchronous approach. Column 3 in Table 1, shows that the
-result now agrees with the Recon1 asynchronous predictions. Using Recon
-1 we came to the correct prediction because it contains only one
-reaction (`PHETHPTOX2') linked to the PKU gene, i.e.~only phenylalanine
-hydroxylase (Entrez ID: 5053, HGNC:8582):
-
-\begin{align*}
-&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
-&\leftrightarrow \\
-&\text{4a-hydroxy-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-tyrosine}
-\end{align*}
-
-In Recon1, the 4a-hydroxytetrahydrobiopterin produced may dehydrate to
-6,7-dihydrobiopterin by the action of the `THBPT4ACAMDASE' reaction
-catalyzed by the PCBD1 enzyme (EC 4.2.1.96). Recon 2.2 contains, in
-addition to the PAH and PCBD1 catalyzed reactions indicated above, the
-overall reaction where the spontaneous dehydration reaction is included,
-again attributed to phenylalanine hydroxylase:
-
-\begin{align*}
-&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
-&\leftrightarrow \\
-&\text{6,7-dihydrobiopterin} + \text{L-tyrosine} + \text{H}_2\text{O}.
-\end{align*}
-
-This details are included in the pathway summary in our Figure
-~\ref{fig:PKU_network}. When the reactions affected by the PKU gene are
-knocked out simultaneously, Recon2 correctly predicts both tyrosine and
-phenylalanine as biomarkers. When blocking only one reaction, there is a
-way to produce tyrosine. Consequently, tyrosine is not predicted as
-biomarker if Recon 2 is used as metabolic map.
-
-Another significant modulation point for the method, and for
-flux-balance-based models in general, is the choice of medium. More
-specifically, it matters for the biomarker prediction which other
-exchange reactions are allowed to have a non-zero flux and what the
-bounds on these fluxes are. Changes herein leave the internal network
-topology unchanged and instead alter how the model is allowed to
-interface with biological compartments or environments that are not
-explicitly modeled. Potentially this potentially has large consequences
-as it may open up or shut down entire parts of the metabolic network
-that rely on specific substrates. It additionally interacts with the
-setting of the threshold for considering biomarkers significant since
-the medium components together with the network pathways (and their
-V\(_{\text{max}}\)'s) determine the amount of a given metabolite that
-may be produced.
-
-In order to examine these ramifications we continue with the synchronous
-PKU example and note that tyrosine production flux goes down in the
-mutant, but not to zero. This is due to an alternative tyrosine
-synthesis pathway in the model originating from 4-Hydroxyphenylpyruvate
-in the medium through the `TYRTA' reaction (tyrosine transaminase E.C.
-2.6.1.5). This transaminase is considered to be reversible in Recon 2
-and Equilibrator \autocite{Noor2012} annotates this reaction with a
-\(K'_{\text{eq}} \approx 1\). It should be mentioned though that the
-ability for transport of hydroxyphenylpyruvate is speculative. Column 4
-and 5 in Table 1 show results when using the standard medium and a
-medium without or with increased (i.e.~influx at -20)
-4-Hydroxyphenylpyruvate influx respectively. In FBA and FVA changes in
-medium concentrations of exchange metabolites are modelled by changing
-the corresponding influx (inward exchange) V\(_{\text{max}}\) (bound).
-In the former the ability to produce tyrosine indeed disappears in the
-mutant case. In the latter there is so much left-over capacity that the
-change in interval bounds slips below the 10\% cutoff. This sensitivity
-to medium composition appropriately reflects the phenomenon that cells
-behave differently in different culture media, and that organisms behave
-differently depending on their nutrition. Our discussion here may
-highlight that this feature should be expected to affect the pertinence
-of biomarkers. Biomarkers may have a tendency to be non-robust,
-dependent as they can be on network topology details and nutrition.
-
-Finally, we consider the value of the \(\epsilon\) parameter, which
-indicates the amount of flux forced through the reaction in the
-wild-type condition. The last column in Table ~\ref{tbl:PKU_results}
-highlights that phenylalanine does not to show up as a biomarker if
-setting \(\epsilon = 0\) and performing the PKU simulation on Recon 1.
-This may also be the reason why Sahoo et al. did not manage to find
-phenylalanine as a biomarker for PKU in their study
-\autocite{Sahoo2012}.
-
-\subsection{\texorpdfstring{Reproduction of Figure 1 in
-\autocite{Shlomi2009}}{Reproduction of Figure 1 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-1-in-shlomi2009}
-
-\begin{figure}
-\centering
-\includegraphics{Figures/scheme_shlomi2009_figure1}
-\caption{A metabolic network used to illustrate the predictions of
-biomarkers of IEMs as originally considered in the work by Shlomi et al.
-for their prediction of biomarkers of IEMs. All metabolites but M3 have
-an exchange reaction allowing for uptake or secretion of the
-metabolites. M2 synthesis is fully dependent on uptake of M1 and M5 or
-M6. The IEM considered is one that leads to deletion of the enzyme
-converting M1 plus M3 to M2 plus M4.}\label{fig:Shlomi2009_figure_1}
-\end{figure}
-
-The original publication's Figure 1 exemplified the biomarker prediction
-method for an illustrative network. It is reproduced here as Figure
-~\ref{fig:Shlomi2009_figure_1}. The simple nature of the example serves
-to explain the reasoning behind the method and additionally helped us
-validate our implementation of the method. This repository provides an
-SBML file of the network, the Jupyter notebook that generates it and a
-Jupyter notebook reproducing the biomarker prediction.
-
-The network in Figure ~\ref{fig:Shlomi2009_figure_1} consists of seven
-metabolites, \(6\) of which (i.e.~all except M3) have exchange
-reactions. We consider a hypothetical deletion of the enzyme catalyzing
-the conversion of M1 into M2 coupled to the conversion of M3 into M4.
-Shlomi et al. provide the a biomarker prediction for this network
-graphically in their Figure 1. We here do so numerically and a Jupyter
-notebook reproducing the biomarker prediction accompanies this paper. In
-our analysis all exchange reactions were given inward bounds of \(-10\)
-and outward bounds of \(+1000\) so that the internal reaction bounds (of
-\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
-shows that all exchangeable metabolites except M5 and M7 are predicted
-to be robust biomarkers.
-
-The exchange interval we found for metabolite M6 shows that in the
-wild-type case M6 can be either taken-up from biofluids or secreted. The
-upper bound is lower in absolute value than the lower bound because in
-the wild-type at least \(1\) unit of flux is needed to provide M3 as
-substrate for the enzyme under investigation since the enzyme convert M1
-and M3 to M2 and M4 is required to have a minimal flux of \(1\) unit. In
-the disease case, M6 (synthesized through M5) can only be secreted to
-biofluids. These results fully agree with the original publication.
-
-\subsection{\texorpdfstring{Reproduction of Figure 2 in
-\autocite{Shlomi2009}}{Reproduction of Figure 2 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-2-in-shlomi2009}
-
-Shlomi et al. applied their method to the reconstructed human metabolic
-map, examining a set of \(17\) IEMs affecting amino-acid metabolism.
-Figure 2 in their publication compares predicted biomarkers with known
-biomarkers from the OMIM database \autocite{McKusick2007}.
-
-We reproduced the analysis for each IEM listed in Table
-~\ref{tbl:amino_acid_iems} and ultimately found full agreement between
-what was reported by Shlomi et al. and the results of our
-implementation.
-
-It is worthwhile to note, with an eye on the sensitivity of this
-approach to the various settings, that we needed to set the medium
-composition very specifically. Shlomi et al. did not report on their
-chosen medium in their publication but we were able to reproduce their
-results by setting all inward flux bounds to \(1\) and all outward flux
-bounds to \(1000\).
-
-We made use of the Excel file included as supplementary information in
-\autocite{Shlomi2009} and included it in this repository as well. In
-this Excel file the genes and reactions linked to a given IEM, are
-detailed. We attempted to extract the genes for each IEM and through the
-Recon 1 map deduce the coupled reactions but found out that this yielded
-predictions different from the ones presented by Shlomi et al. In the
-Excel file a list of reactions linked to the IEM is also provided.
-However, these reactions are not necessarily the same reactions that are
-linked in Recon 1 to the genes they provide. Assuming the reconstruction
-of the metabolic network to be accurate this raises some doubts as to
-the correctness of the results obtained by Shlomi et al. for those IEMs
-where the listed reactions did not agree with the reconstructed map.
-
-Additionally, we needed to manually add the information for
-S-Adenosylhomocysteine hydrolase and Methionine adenosyltransferase
-deficiency, since they were missing from the Excel file and to reduce
-the tetrahydrobiopterin deficiency gene list to only consider QDPR
-(quinoid dihydropteridine reductase). Lastly, Histidinemia was in the
-Excel file associated to the reaction HISD whereas this should have
-refered to HISDr to match the reaction ID in Recon 1. After these
-changes, we recovered the results as originally presented and as
-summarized here in Table ~\ref{tbl:amino_acid_iems}.
-
-\hypertarget{tbl:toy_model_results}{}
-\begin{longtable}[]{@{}lllc@{}}
-\caption{\label{tbl:toy_model_results}Biomarker predictions using our
-implementation with respect to the mutation affecting the conversion of
-M1 plus M3 to M2 plus M4 in the network of Figure
-~\ref{fig:Shlomi2009_figure_1}. H.C. stands for `highly confident'
-indicating, as defined by Shlomi et al., i.e.~no overlap between the WT
-and mutant intervals. Between brackets {[}{]} we report the lower and
-upper bounds of the FVA intervals we obtained. }\tabularnewline
-\toprule
-ID & Prediction & WT & Mutant\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-ID & Prediction & WT & Mutant\tabularnewline
-\midrule
-\endhead
-M1 & H.C. Elevated & {[}-10.0, -1.0{]} & {[}0.0, 0.0{]}\tabularnewline
-M2 & H.C. Reduced & {[}1.0, 10.0{]} & {[}0.0, 0.0{]}\tabularnewline
-M4 & Reduced & {[}1.0, 20.0{]} & {[}0.0, 10.0{]}\tabularnewline
-M6 & Elevated & {[}-10.0, 9.0{]} & {[}0.0, 10.0{]}\tabularnewline
-M7 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
-M5 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\section{Conclusion}\label{conclusion}
-
-The results summarized in Figures 1 and 2 of \autocite{Shlomi2009} have
-been successfully replicated with the software tools provided here.
-Overall, the reproducing process was smooth since the method is
-illustrated both on an illustrative example and on the human metabolic
-reconstruction, resulting in a multitude of test cases for the
-implementation.
-
-The points of trouble we encountered were the lack of explicit
-discussion on the set of fluxes allowed into and out of the model and a
-set of discrepancies between the gene-reaction mappings listed in the
-Supplementary Excel file and the actual gene-reaction mappings in the
-metabolic network. However, trial-and-error and manual curation solved
-these issues.
-
-\hypertarget{tbl:amino_acid_iems}{}
-\begin{longtable}[]{@{}llll@{}}
-\caption{\label{tbl:amino_acid_iems}We used our implementation to
-reproduce the biomarkers for the list of inborn-errors of metabolism
-shown in Figure 2 in \autocite{Shlomi2009}. Using the Supplementary
-Excel table of \autocite{Shlomi2009} linking IEMs to causative genes and
-affected reactions we filtered out those that are listed here. The only
-considered biomarkers are the \(20\) proteinogenic amino acids. We
-report all biomarkers the change of which exceeds the 10\% threshold and
-categorize them by their qualitative prediction: elevated or reduced
-serum levels. }\tabularnewline
-\toprule
-IEM & Affected reactions & Elevated & Reduced\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-IEM & Affected reactions & Elevated & Reduced\tabularnewline
-\midrule
-\endhead
-S-Adenosylhomocysteine & SEAHCYSHYD, AHCi & L-Methionine &
-L-Cysteine\tabularnewline
-hydrolase & & &\tabularnewline
-Alkaptonuria & HGNTOR & L-Tyrosine &\tabularnewline
-Arginase deficiency & ARGN & &\tabularnewline
-Cystinuria & CYSTSERex, & &\tabularnewline
-& SERLYSNaex & &\tabularnewline
-Lysinuric protein & SERLYSNaex & &\tabularnewline
-intolerance & & &\tabularnewline
-Glutamate formimino- & FTCD, GluForTx & L-Histidine &\tabularnewline
-transferase deficiency & & &\tabularnewline
-Histidinemia & HISDr & L-Histidine &\tabularnewline
-Homocystinuria & CYSTS, MTHFR3, & L-Methionine &
-L-Cysteine\tabularnewline
-& METS & &\tabularnewline
-Hyperprolinemia type I & PRO1xm, PROD2m & &\tabularnewline
-maple syrup & OIVD1m, & L-Valine, &\tabularnewline
-urine disease & OIVD2m, & L-Isoleucine &\tabularnewline
-& OIVD3m & L-Leucine, &\tabularnewline
-Methionine adenosyl- & SELMETAT, METAT & L-Methionine &
-L-Cysteine\tabularnewline
-transferase deficiency & & &\tabularnewline
-methylmalonic acidemia & CBLATm, CBL2tm, & L-Isoleucine &\tabularnewline
-& MMMm & &\tabularnewline
-phenylketonuria & PHETHPTOX2 & L-Phenylalanine &
-L-Tyrosine\tabularnewline
-Tetrahydrobiopterin & DHPR & &\tabularnewline
-deficiency & & &\tabularnewline
-Tyrosinemia type 1 & FUMAC & L-Tyrosine &\tabularnewline
-Tyrosinemia type III & PPOR, 34HPPOR & L-Tyrosine &\tabularnewline
-Glycine encephalopathy & GCC2am, GCC2bim, & &\tabularnewline
-& GCC2cm, GCCam, & &\tabularnewline
-& GCCbim, GCCcm & &\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{equation} A = \sqrt{\frac{B}{C}} \label{eq:1}\end{equation}
 

### 2018-02-06 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Update version of figures and manuscript

@@
 \usepackage{amsmath,amssymb} %added by Roque
+\usepackage[normalem]{ulem} % added by Nilton to use \sout
 \hypersetup{breaklinks=true,
@@
 ]{biblatex}
-\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.bib}
+\bibliography{ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017}
 
@@
 \label{Eq:LIF}
-\dot{V}_i(t) = -\left( V_i(t) - V_{\rm reset} \right) /\tau_m + I_i(t)/C_m,
+\dot{V}_i(t) = - \left( V_i(t) - V_{\rm reset} \right)/\tau_m + I_i(t)/C_m,
 \end{equation}
@@
  inhibitory synaptic transmission delay & $d_i$ & $\mathcal{N}$($\mu = 0.8 , \sigma = 0.4$) ms \tabularnewline
- initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-65.0, \sigma=6.5$) mV \tabularnewline
+ initial membrane potential & $V_0$ & $\mathcal{N}$($\mu=-58.0, \sigma=10.0$) mV \tabularnewline
    \bottomrule
@@
   \caption{\label{table:population} Layer population sizes (extracted from \cite{potjans2014}).
-Neurons were distributed over four different layers (L23, L4, L5 and L6), and for each layer they were divided into excitatory (L23e, L4e, L5e and L6e) and inhibitory (L23i, L4i, L5i and L6i) subpopulations.}
+Neurons were distributed over four different layers (L23, L4, L5 and L6), and for each layer they were divided into excitatory (L23e, L4e, L5e and L6e) and inhibitory (L23i, L4i, L5i and L6i) subpopulations.
+The thalamic population is represented by Th.}
  \begin{center}
-  \begin{tabular}{cccccccc}
+  \begin{tabular}{ccccccccc}
    \toprule
    \midrule
- L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
- 20683 & 5834 & 21915 & 5479 & 4850 & 1065 & 14395 & 2948\tabularnewline
+ L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i & Th\tabularnewline
+ 20683 & 5834 & 21915 & 5479 & 4850 & 1065 & 14395 & 2948 & 902\tabularnewline
    \bottomrule
@@
 \begin{center}
-\begin{tabular}{lccccccccc}
+\begin{tabular}{lcccccccccc}
  \toprule
@@
   \midrule
- &  & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i\tabularnewline
+ &  & L23e & L23i & L4e & L4i & L5e & L5i & L6e & L6i & Th\tabularnewline
  \midrule
- & L23e & 0.101 & 0.169 & 0.044 & 0.082 & 0.032 & 0 & 0.008 & 0\tabularnewline
- & L23i & 0.135 & 0.137 & 0.032 & 0.052 & 0.075 & 0 & 0.004 & 0\tabularnewline
- & L4e & 0.008 & 0.006 & 0.050 & 0.135 & 0.007 & 0.0003 & 0.045 & 0\tabularnewline
- to & L4i & 0.069 & 0.003 & 0.079 & 0.160 & 0.003 & 0 & 0.106 & 0\tabularnewline
- & L5e & 0.100 & 0.062 & 0.051 & 0.006 & 0.083 & 0.373 & 0.020 & 0\tabularnewline
- & L5i & 0.055 & 0.027 & 0.026 & 0.002 & 0.060 & 0.316 & 0.009 & 0\tabularnewline
- & L6e & 0.016 & 0.007 & 0.021 & 0.017 & 0.057 & 0.020 & 0.040 & 0.225\tabularnewline
- & L6i & 0.036 & 0.001 & 0.003 & 0.001 & 0.028 & 0.008 & 0.066 & 0.144\tabularnewline
+ & L23e & 0.101 & 0.169 & 0.044 & 0.082 & 0.032 & 0.0 & 0.008 & 0.0&0.0\tabularnewline
+ & L23i & 0.135 & 0.137 & 0.032 & 0.052 & 0.075 & 0.0 & 0.004 & 0.0&0.0\tabularnewline
+ & L4e & 0.008 & 0.006 & 0.050 & 0.135 & 0.007 & 0.0003 & 0.045 & 0.0&0.0983\tabularnewline
+ to & L4i & 0.069 & 0.003 & 0.079 & 0.160 & 0.003 & 0.0 & 0.106 & 0.0&0.0619\tabularnewline
+ & L5e & 0.100 & 0.062 & 0.051 & 0.006 & 0.083 & 0.373 & 0.020 & 0.0&0.0\tabularnewline
+ & L5i & 0.055 & 0.027 & 0.026 & 0.002 & 0.060 & 0.316 & 0.009 & 0.0&0.0\tabularnewline
+ & L6e & 0.016 & 0.007 & 0.021 & 0.017 & 0.057 & 0.020 & 0.040 & 0.225&0.0512\tabularnewline
+ & L6i & 0.036 & 0.001 & 0.003 & 0.001 & 0.028 & 0.008 & 0.066 & 0.144&0.0196\tabularnewline
 \bottomrule
@@
 The ordinary differential equations were solved with the exact integration method for linear equations available in Brian 2 with a time step $\Delta t=0.1$ ms.
-All simulations were carried out in a cluster with 4 nodes each equipped with 2 Intel Xeon processors.
+All simulations were carried out in a computer equipped with 2 Intel Xeon processors (E-2650v3, 2.3GHz) with 128GB of RAM.
 
@@
 \begin{enumerate}
-\item \textbf{Layer specific}: Neurons from each layer receive specific background spike-trains drawn from an $8$ Hz Poisson distribution.
+\item \textbf{Layer specific}: Neurons from each layer receive specific background spike-trains drawn from a Poisson distribution with firing rate $\nu = 8$Hz.
 The number of inputs per neuron is given in the first row of Table \ref{table:externalinput}.
@@
 The number of inputs per neuron follows the layer specific configuration (first row in Table \ref{table:externalinput}).
-Observe that the number of inputs per neuron is multiplied by an effective factor which is given by $\nu \times \omega \times \tau_{syn}=0.3512$ pA.
+Observe that the number of inputs per neuron is multiplied by an effective factor which is given by $\nu \times w \times \tau_{syn}=0.3512$ pA.
+This factor is determined by calculating the time average of the synaptic current delivered by $n$ spikes of the Poisson background to a given neuron.
+\item \textbf{Transient thalamic input}: A thalamic population (Th) of $902$ neurons, represented by Poissonian spike trains with firing rate of $120$ Hz, is connected with the network following the connection probabilities in Table \ref{table:conectivity}.
+In a simulated time period of $1$ s, the thalamic input is applied at time $t = 700$ ms and lasts for $10$ ms.
+This protocol is repeated hundred times to compute the average activity of $2.5$\% randomly sampled neurons from each layer of the network.
 \end{enumerate}
@@
 \begin{equation}
-\label{Eq:Ksyn_alt}
+  \label{Eq:transitivity}
   C_{i} = \left( \frac{1-T}{1+T} \right)C_{e} ,
-\label{Eq:transitivity}
 \end{equation}
@@
 
-To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test.
+To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test (KS-statistics).
 To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.
-
+The null hypothesis that two sample sets were drawn from the same distribution is accepted, for an error $\alpha = 0.01$, if the KS-statistic $\leq 1.63$.
 \section{Results}\label{results}
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!h]
  \centering
- \includegraphics[scale=0.8]{fig2.pdf}
+ \includegraphics[scale=0.001]{fig2.pdf}
  \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article).
@@
 
+\begin{figure}[!htp]
+ \centering
+ \includegraphics[scale=0.01]{fig3.pdf}
+ \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
+NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
+\end{figure}
+
 Besides the creation of the network connections by the procedure described above, which is the one used in the original article, it is also possible to connect the neurons in the network using the alternative expression for the total number of synapses between two cell populations,
+
+\begin{figure}[!ht]
+ \centering
+ \includegraphics[scale=0.005]{fig4.pdf}
+ \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
+NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
+\end{figure}
 
@@
 
-\begin{figure}[!ht]
- \centering
- \includegraphics[scale=0.02]{fig3.pdf}
- \caption{Cumulative distribution of CVs of ISIs for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
-NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
-\end{figure}
-
-\begin{figure}[!ht]
- \centering
- \includegraphics[scale=0.01]{fig4.pdf}
- \caption{Cumulative distribution of firing rate for the eight cell populations (indicated atop each panel) for simulations implemented in Brian 2 using equation~(\ref{Eq:Ksyn_alt}) (blue), in Brian 2 using equation~(\ref{Eq:Ksyn}) (green), and in NEST using equation~(\ref{Eq:Ksyn}) (orange).
-NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig4}
-\end{figure}
-
-When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, most of the layer-specific comparisons yielded p-values higher than 0.6 as before, indicating no significant differences.
-However, the p-value of the comparison between the CV of ISIs for the layer 5 excitatory neurons was $0.07$, and the p-values of the comparisons between the firing rates of excitatory cells from layers 2/3 and 6 were, respectively, $0.10$ and $0.08$.
-Therefore, in the cases of excitatory cells of layers 2/3, 5 and 6 there were statistically significant differences in the firing properties of neurons.
-The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have firing rate 12.2 $\pm$ 6.1 Hz while in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
+When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences for an error $\alpha = 0.01$.
+The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
 
@@
 
-In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs.
-These are shown in Fig.~\ref{fig:fig5}A, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}B, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs.
-In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
-In Fig.~\ref{fig:fig5}C we present the population firing rates for 100 trials with the rule explained in the methods section.
-This latter experiment presented an excellent agreement with the histograms observed in Fig.
-7 of the original article.
-
-\begin{figure}[!ht]
+\begin{figure}[!htb]
  \centering
- \includegraphics[scale=0.01]{fig5.pdf}
+ \includegraphics[scale=0.005]{fig5.pdf}
  \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig.
@@
 
+In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs.
+These are shown in Fig.~\ref{fig:fig5}A, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}B, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs.
+In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
+In Fig.~\ref{fig:fig5}C we present the population firing rates for 100 trials with the rule explained in the methods section.
+This latter experiment presented an excellent agreement with the histograms observed in Fig.
+7 of the original article.
+
 \subsection{Stability of Network Activity}
@@
 
-The asynchronous and irregular activity of the reimplemented network model, as characterized by the AIness\%, is similar to the one found in the reference article for background rates >5 Hz and relative inhibitory synaptic strengths >4 Hz.
-In comparison with the reference article, the relative order of excitatory firing rates is maintained for every combination between background rate and relative synaptic strength, with highest values in L5 and smallest in L2/3 and L6.
-Similarly to the original article, the firing rate of L4e cells is the most sensitive to variations in the background rate whereas the firing rate of L5e cells is the most sensitive to variations in the relative inhibitory synaptic strength $g$.
-
 \begin{figure}[!ht]
  \centering
- \includegraphics[scale=0.01]{fig6.pdf}
+ \includegraphics[scale=0.1]{fig6.pdf}
  \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig.
@@
 
+The asynchronous and irregular activity of the reimplemented network model, as characterized by the AIness\%, is similar to the one found in the reference article for background rates >5 Hz and relative inhibitory synaptic strengths >4 Hz.
+In comparison with the reference article, the relative order of excitatory firing rates is maintained for every combination between background rate and relative synaptic strength, with highest values in L5 and smallest in L2/3 and L6.
+Similarly to the original article, the firing rate of L4e cells is the most sensitive to variations in the background rate whereas the firing rate of L5e cells is the most sensitive to variations in the relative inhibitory synaptic strength $g$.
+
+\subsection{Propagation of Transient Thalamic Inputs}
+
+In this section we make use of the thalamic inputs as described in the last column of Table \ref{table:population}.
+For further information on how thalamic inputs are applied we refer to the Methods section.
+
+As described in the results section of the original article, when thalamic input is delivered to the network the output is rather stereotypical.
+This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian2 to transient thalamic input.
+By comparing histograms in Fig.~\ref{fig:fig7}\textbf{B} with the ones in Fig.
+10 \textbf{B} of the original article one can see the similarities.
+
+\begin{figure}[!ht]
+ \centering
+% * <aroquesilva@gmail.com> 2018-02-06T01:02:54.595Z:
+%
+% ^.
+ \includegraphics[scale=0.01]{fig7.pdf}
+ \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.10 from the original article).
+Note that the time axis is centralized at the beginning of the thalamic input, which lasts for 10 ms.
+(A) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section.
+(B) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
+\end{figure}
+
+\subsection{Important information needed during replication}
+
+Most of the information necessary for the replication done here was available in the original article.
+However, a few details were found only in the NEST implementation of the code available in the Open Source Brain (OSB) platform \cite{potjans2014pyNEST}.
+In this section we list these details.
+
+\begin{itemize}
+
+\item When replacing the Poissonian background by a DC current the number of inputs per neuron was multiplied by the factor $0.3512$ pA as explained in the Methods section of this article.
+This factor is not explicitly explained in the original article.
+
+\item There is no mention to the initial voltage values in the original article.
+This information was found in the OSB NEST code where one can observe that the initial voltage values are randomly drawn from the normal distribution $\mathcal{N}$($\mu=-58 , \sigma = 10$) mV.
+
+\item In the original article, the equation for the integrate-and-fire model does not explicitly state the  resting potential.
+This could suggest to a na\"ive reader that the resting potential is $0$ mV, whereas the correct resting potential is the reset voltage $V_{\rm reset}$ (cf.
+equation~(\ref{Eq:LIF})).
+
+\item The thalamic input in the original article is described as having $15$ Hz, however in the OSB NEST code it is actually $120$ Hz, which is the correct value to be used.
+
+\end{itemize}
+
 \section{Conclusion}\label{conclusion}
@@
 
-The successful replication of the results of the reference article confirms the correctness of the original implementation of the model.
+The successful replication of the results of the reference article confirms the robustness of the original implementation of the model.
 
@@
 The authors would like to thank Prof.
-Markus Diesmann for valuable comments on the manuscript.
+Markus Diesmann, Dr.
+Maximilian Schmidt and Dr.
+Sacha van Albada for valuable comments on the manuscript.
 This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC.

### 2018-02-07 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
added article files

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Mario Senden,
      Jannis Schuecker,,
      Jan Hahne,,
      Markus Diesmann,,
      Rainer Goebel,,
  },
            pdftitle=
{
[Re] A neural model of the saccade generator in the reticular formation
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{senden-schuecker-hahne-diesmann-goebel-2018.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Month 2018 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] A neural model of the saccade generator in the reticular formation}

    \usepackage{authblk}
                        \author[1, 2]{Mario Senden}
                    \author[3]{Jannis Schuecker,}
                    \author[4]{Jan Hahne,}
                    \author[3, 5, 6]{Markus Diesmann,}
                    \author[1, 2, 7]{Rainer Goebel,}
                            \affil[1]{Department of Cognitive Neuroscience, Faculty of Psychology and
Neuroscience, Maastricht University, 6201BC Maastricht, The Netherlands}
                    \affil[2]{Maastricht Brain Imaging Centre, Faculty of Psychology and Neuroscience,
Maastricht University, P.O. Box 616, 6200 MD Maastricht, The Netherlands}
                    \affil[3]{nstitute of Neuroscience and Medicine (INM-6) and Institute for Advanced
Simulation (IAS-6) and JARA Institute Brain Structure-Function
Relationships (INM-10), Jülich Research Centre, Jülich, Germany}
                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universit"at
Wuppertal, Wuppertal, Germany}
                    \affil[5]{Department of Psychiatry, Psychotherapy and Psychosomatics, Medical
Faculty, RWTH Aachen University, 52062 Aachen, Germany}
                    \affil[6]{Department of Physics, Faculty 1, RWTH Aachen University, 52062 Aachen,
Germany}
                    \affil[7]{Department of Neuroimaging and Neuromodeling, Netherlands Institute for
Neuroscience, an Institute of the Royal Netherlands Academy of Arts and
Sciences (KNAW), 1105BA Amsterdam, The Netherlands}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:mario.senden@maastrichtuniversity.nl}{mario.senden@maastrichtuniversity.nl}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Month, Day, 2018\\
  {\bfseries Accepted}  Month, Day, 2018\\
  {\bfseries Published} Month, Day, 2018\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{A neural model of the saccade generator in the reticular
formation}, G. Gancarz, S. Grossberg, Neural Networks, 1159-1174, 1998
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

We provide an implementation of the saccade generator (SG); a rate
neuron model of neural circuitry in the reticular formation proposed by
Gancarz \& Grossberg \autocite{Gancarz1998}. The SG is an integral part
of the eye movement system \autocite{Grossberg2012} and as such vital
for developing large-scale architectures of visuo-motor integration. For
this reason it is of interest to implement the model in a publicly
available, widely used, and actively developed neural simulation
framework such as NEST \autocite{Gewaltig2007}. We show that the model
translates well to the NEST framework as our implementation faithfully
reproduces all simulation results reported in the original publication.
Our code uses the Python interface \autocite{Eppler2008} for legibility
with both model and analysis scripts being implemented using Python
2.7.12.

\section{Methods}\label{methods}

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. In what follows we briefly describe the dynamics of neurons
within the horizontal component (equivalent descriptions apply to the
vertical component).

\subsubsection{Long-lead burst neurons}\label{long-lead-burst-neurons}

Each long-lead burst neuron (\(L\)) receives excitatory external input
\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\)):

\[
\begin{array}{ll}
\tau\frac{dL_l}{dt} &= -1.3L_l+I_l-2B_l \\\\
\tau\frac{dL_r}{dt} &= -1.3L_r+I_r-2B_r \\
\end{array}
\]

\subsubsection{Excitatory burst neurons}\label{excitatory-burst-neurons}

Each excitatory burst neuron (\(E\)) receives excitatory input from the
ipsilateral LLBN, a constant arousal signal equal to \(1\), and
inhibitory input from the omnipause neuron (\(P\)):

\[
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5E_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5E_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\] with a nonlinear gain function \(g(x)\) given by

\[
g(x) = \frac{x^4}{0.1^4+x^4} \textrm{.}
\]

\subsubsection{Inhibitory burst neurons}\label{inhibitory-burst-neurons}

Each inhibitory burst neuron (\(B\)) receives excitatory input from the
ipsilateral EBN:

\[
\begin{array}{ll}
\tau\frac{dB_l}{dt} &= -2.4B_l+3E_l \\\\
\tau\frac{dB_r}{dt} &= -2.4B_r+3E_r \\
\end{array}
\]

\subsubsection{Omnipause neuron}\label{omnipause-neuron}

The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs:

\[
\tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
\] where \(J\) represents external electrical stimulation.

\subsubsection{Tonic neurons}\label{tonic-neurons}

Each tonic neuron (\(T\)) receives excitatory input from the ipsilateral
EBN and inhibitory input from the contralateral EBN:

\[
\begin{array}{ll}
\tau\frac{dT_l}{dt} &= 0.1(E_l-E_r) \\\\
\tau\frac{dT_r}{dt} &= 0.1(E_r-E_l) \textrm{.} \\
\end{array}
\] Horizontal eye position (\(\theta\)) depends on activity of the right
TN; i.e. \(\theta = 260(T_r-0.5)\).

In implementing this model, we largely followed the descriptions
provided in the original publication with a number of well-motivated
exceptions. Specifically, the original model description has two
features which cannot be straightforwardly translated to NEST. First, a
nonlinear gain function is applied to a subset of inputs to EBNs and the
OPN while a linear gain function is applied to their remaining inputs.
Since NEST only applies a single gain function per neuron to each of its
inputs, we opted for using a linear gain function for EBNs and the OPN.
Furthermore, we passed those inputs requiring an additional nonlinear
gain function through an auxiliary unit instantaneously applying the
desired nonlinearity before passing the result on to EBNs and the OPN.
Second, constant input to a neuron was not hard-coded but rather
provided by an appropriately weighted bias node. Neither of these
changes lead to discrepancies with original results.

In all simulations we used the Exponential Euler method for numerical
integration of rate neurons \autocite{Hahne2016} at a time step of
\(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
should be noted that rates of all neurons were initialized to zero and
were thus not at resting equilibrium. In order to address this, the
model was allowed to evolve for \(100\,\mathrm{ms}\), to relax towards
equilibrium before applying any input. Furthermore, we always simulated
the full model; i.e.~both its horizontal and vertical components even if
input was applied only to one of the two.

\section{Results}\label{results}

All results from the original publication have been implemented. Our
results accord very well with those reported by Gancarz \& Grossberg
\autocite{Gancarz1998}. Only a single discrepancy was observed which we
examined in depth.

The first simulation showcases the evolution of activity for each neuron
type in the horizontal SG to a constant input applied to the left LLBN.
The original publication does not report exact activation values
observed for each neuron rendering a quantitative analysis of the
accuracy of our replication impossible. However, qualitatively
activation profiles shown in figure 3 in the original publication and
those shown in figure \ref{fig:fig_1} show good correspondence. In both
implementations, the left LLBN showed a prelude of activity with left
EBN bursts beginning after the onset of LLBN activity. The right EBN
produced a small burst at the end of a saccade. Furthermore, increases
in activity of the left TN were mirrored by decreases in the right.
Finally, OPN activity dropped to zero during production of a saccade.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.45000\textwidth]{figures/fig1.eps}
\caption{\textbf{Activity profiles in the left (A) and right (B) SG.}
All activities are in response to constant input (\(\mathrm{I=1}\))
applied to the left LLBN for \(265\,\mathrm{ms}\).\label{fig:fig_1}}
\end{figure}

The second simulation shows the relation between input strength and
burst amplitude for LLBNs and EBNs. With increasing input strength, both
amplitude and duration of activation increased in long-lead and
excitatory burst neurons. As before, these results accord well with
those shown in figure 5 of the original publication.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.36000\textwidth]{figures/fig2.eps}
\caption{\textbf{Activity profiles in LLBN (A) and EBN (B).} Increased
input strength resulted in larger LLBN and EBN burst size with inputs
equal to 1 (blue), 1.75 (green), and 2.5 (red) each applied to the left
LLBN for \(85\,\mathrm{ms}\).\label{fig:fig_2}}
\end{figure}

The third simulation generates saccades in response to different input
strengths applied to the horizontal and vertical SG (figure
\ref{fig:fig_3}). As in the original publication, saccades were
generally straight with a slight tendency to curve. We extracted
quantitative estimates of saccade amplitude for the original publication
from their figure 6 and calculated the root-mean-squared error (RMSE)
between these estimates and saccade amplitudes produced by our model.
The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
0.22\textdegree, 0.14\textdegree~for the blue, green, red, turqoiuse,
and purple curves, respectively. Given that saccade amplitudes are
between 10\textdegree~and 15\textdegree, these RMSE values indicate a
close match between the two implementations.

\begin{figure}
\centering
\includegraphics[width=0.48750\textwidth,height=0.48750\textwidth]{figures/fig3.eps}
\caption{\textbf{Oblique saccades.} Inputs to the right and upward LLBNs
were \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.22}\) (green);
\(\mathrm{{I}_{r}=0.74}\) \& \(\mathrm{{I}_{u}=0.4}\) (red);
\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turqoiuse); and
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and were
applied for \(75\,\mathrm{ms}\). As expected, saccades were fairly
straight with a slight tendency to curve.\label{fig:fig_3}}
\end{figure}

The fourth simulation implements a staircase of three saccades in
response to continuous input (figure \ref{fig:fig_4}). Our results agree
with those reported in the original publication (their figure 7) with
saccades being of equal length and the smaller component (horizontal)
being stretched to produce straight oblique saccades.

\begin{figure}
\centering
\includegraphics[width=0.33000\textwidth,height=0.45000\textwidth]{figures/fig4.eps}
\caption{\textbf{Saccadic staircase.} Three saccades in a staircase
continued in the same direction as the initial saccade. Inputs to the
right and upward LLBNs were equal to \(\mathrm{{I}_{r}=0.2}\) and
\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). Eye position was
sampled every \(2\,\mathrm{ms}\).\label{fig:fig_4}}
\end{figure}

In the fifth simulation the average activity of the left EBN was
obtained for a series of saccades with different directions. Figure
\ref{fig:fig_5} shows a polar plot of average activity corresponding to
each saccade reflecting the neuron's tuning curve. The tuning curve we
observed for the left EBN exhibits a cardioid-like shape as was the case
in the original publication (figure 8).

\begin{longtable}[]{@{}lllllllllllllll@{}}
\caption{Direction specific inputs to SG to produce EBN tuning curve.
\label{tbl:table_1}}\tabularnewline
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endfirsthead
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endhead
\(\mathrm{{I}_{l}}\) & .00 & .00 & .00 & .00 & .20 & .45 & .63 & .70 &
.63 & .45 & .20 & .00 & .00 & .00\tabularnewline
\(\mathrm{{I}_{r}}\) & .70 & .45 & .20 & .00 & .00 & .00 & .00 & .00 &
.00 & .00 & .00 & .00 & .20 & .45\tabularnewline
\(\mathrm{{I}_{d}}\) & .00 & .00 & .00 & .00 & .00 & .00 & .00 & .00 &
.20 & .45 & .63 & .70 & .63 & .45\tabularnewline
\(\mathrm{{I}_{u}}\) & .00 & .45 & .63 & .70 & .63 & .45 & .20 & .00 &
.00 & .00 & .00 & .00 & .00 & .00\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=0.37500\textwidth,height=0.37500\textwidth]{figures/fig5.eps}
\caption{\textbf{EBN tuning curve.} Tuning curve of the left EBN
exhibiting a cardioid-like shape. Inputs to the SG producing the desired
saccades can be found in table \ref{tbl:table_1}. Each of these inputs
was applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
\end{figure}

\pagebreak
The sixth simulation reported in the original publication was designed
to replicate results of Stanford \textit{et al.}
\autocite{Stanford1996}. These authors stimulated the superior
colliculus (SC) at various frequencies and measured the resulting
saccade amplitude, duration, and velocity; showing that ampltidude
saturates before velocity. Our implementation of the SG model was
capable of replicating these results. However, reproducing simulation
results reported by Gancarz \& Grossberg \autocite{Gancarz1998} with our
implementation was complicated by the fact that the stimulation protocol
given by the authors lead to the production of two rather than a single
saccade for larger stimulation intensities. Furthermore, the authors did
not report their criteria for identifying saccade on- and offsets.
Stanford \textit{et al.} \autocite{Stanford1996} used velocity criteria
to determine onset (\(v\textgreater30\,\mathrm{deg/s}\)) and offset
(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provided us
with explicit criteria, velocity did not always drop below
\(30\,\mathrm{deg/s}\) after the first saccade before rising again with
the second. To determine the offset of a saccade in those cases, we
found the local minimum between the end of the first and the beginning
of the second saccade. With these criteria in place, we stimulated the
SC. Results of our simulation are shown in figure \ref{fig:fig_6}. We
extracted quantitative results of the original publication from their
figure 9 and calculated the RMSE between these estimates and our results
for each curve. The RMSE was 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
\(4.45\,\mathrm{deg/s}\) for saccade amplitude, duration, and peak
velocity, respectively. Our results accord thus very well with those
reported in the original publication. This includes the observation that
saccade amplitude and especially saccade duration were largest for a
stimulation intensity of 1.2. It is conceivable that this intensity
marks the point after which the SG generator starts producing two rather
than a single saccade with the eye movement at this intensity merging
two saccades and thus being stretched out.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.37500\textwidth]{figures/fig6.eps}
\caption{\textbf{Effect of stimulation frequency on saccade (A)
amplitude, (B) duration, and (C) velocity.} Saccade amplitude (A)
saturated before saccade velocity (C). A range of unitless values
varying between 1 and 2.4 at increments of 0.2 were applied to the SC
(reflecting stimulation at different frequencies). The connection weight
from SC to LLBN was \(\mathrm{W=2}\) and stimulation duration was
\(125\,\mathrm{ms}\).\label{fig:fig_6}}
\end{figure}

The seventh simulation shows that saccade velocity and duration can be
traded while keeping amplitude constant. To produce a high-velocity
saccade, the SC was stimulated at a high frequency. Similarly, to
produce a low-velocity saccade, the SC was stimulated at a low
frequency. Figure \ref{fig:fig_7} shows the results of our simulation.
In line with the original publication, the saccade amplitude reflected
by TN activity was identical after high- and low-frequency saccades thus
confirming the reported effect. However, we observe differences in the
specific details of our implementation with respect to the original,
starting with the shape of the input curves. While the rise of the input
curves in the original publication and our implementation coincide well,
the decay in our curves started later and proceeded with a shorter time
constant (\(\approx25\,\mathrm{ms}\), i.e.~half the simulation time
constant). This resulted in more total input to the model in our
implementation and hence lead to the generation of a second saccade not
observed in the original publication. When repeating our simulation
using an ad-hoc solution to create matching input curves (by halving the
time constant after external stimulation), we observed results identical
to those reported in the original publication. We proceeded to
investigate the origin of the discrepancy with regard to the shapes of
the input curves. First, we tested whether discrepancies were specific
to our NEST implementation by solving the expressions describing the
input analytically. Specifically, according to equations A13-A15 of
\autocite{Gancarz1998} the input \(I\) is described by \[
\begin{array}{lll}
    \tau \frac{dA}{dt} &=-A+F(t)\\
    f(x) &=
    \left\{
    \begin{array}{lll}
        0 \quad &\textrm{if} \quad x<0 \\
        x \quad &\textrm{if} \quad 0<x<1 \\
        1 \quad &\textrm{if} \quad x \geq 1 \\
    \end{array}
    \right. \\
    I(t) &=Wf(A)
\end{array}
\]\\
with the stimulus \[
F(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t<{t}_{on} \\
    F \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.} \\
    0 \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\]\\
Given this, the analytic expression for \(I(t)\) is \begin{equation}
I(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t<{t}_{on} \\
    W \cdot f(F \cdot (1-{e}^{-(t-{t}_{on})/ \tau})) \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.}\\
    W \cdot f(F \cdot (1-{e}^{-({t}_{off}-{t}_{on})/ \tau}) \cdot {e}^{-(t-{t}_{off})/ \tau}) \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\label{eq:1}\end{equation} Evaluating the analytic expression given
parameter values reported in the original manuscript exactly reproduced
our numerical results and thus produced curves equally deviatng from
those shown in the original publication. Next, we contacted one of the
original authors to rule out the possibility that we used erroneous
parameter settings but no mistake was pointed out to us. The origin of
the discrepancy thus remains unclear. However, it is likely it stems
from the description of the input rather than from problems with the SG
model itself.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.75000\textwidth]{figures/fig7.eps}
\caption{\textbf{Trading saccade velocity for duration.} Duration and
velocity of a saccade can be traded while keeping amplitude constant. To
produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
publication). To produce a low velicity saccade an input of
\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
input curve produced by our implementation of the original equations
(A), a second saccade can be observed for both high and low velocity
saccades. If the input curves reported in the original manuscript were
recreated in an ad-hoc fashion (B), the responses of all model neurons
matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
\end{figure}

The eighth simulation reported by Gancarz \& Grossberg
\autocite{Gancarz1998} shows how strong sustained input to the SG
produces smooth eye movements. Our results, shown in figure
\ref{fig:fig_8}, reproduced these findings as they strongly resembled
those shown in figure 11 of the original publication. Specifically, an
initial burst exhibited by the EBN was followed by sustained lower
activity. This was due to inhibitory feedback being insufficient to
silence the LLBN when input remains continuously strong and resulted, in
turn, in the observed smooth eye movement.

\begin{figure}
\centering
\includegraphics[width=0.60000\textwidth,height=0.46200\textwidth]{figures/fig8.eps}
\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
SG neurons accompanying smooth eyes movement as a result of strong
(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
input.\label{fig:fig_8}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.46200\textwidth,height=0.70000\textwidth]{figures/fig9.eps}
\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
OPN stimulation interrupted the saccade, which remained accurate
nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
uninterrupted saccade.\label{fig:fig_9}}
\end{figure}

\clearpage

The final simulation showcases the evolution of activity exhibited by SG
neurons when the OPN is briefly electrically stimulated while a constant
input is applied to the LLBN. External stimulation temporarily restored
activation in the OPN and hence also inhibition of the EBN, leading to
an interruption of the saccade. As is shown in figure \ref{fig:fig_9},
the saccade remained accurate despite this disruption. This is in
agreement with results shown in figure 12 of the original publication.

\section{Conclusion}\label{conclusion}

The reproduced results show very good qualitative correspondence with
those reported by Gancarz and Grossberg \autocite{Gancarz1998} and
accorded well quantitatively whenever such information was available.
The only discrepancy between our and the original implementation was
that the input to the saccade generator took more time to decay in our
implementation. While the exact cause of this discrepancy is unclear, we
could rule out that it was specific to the NEST implementation.
Furthermore, it does not affect reproduction of the main finding of
simulation seven; namely that saccade velocity and duration can be
traded while keeping its amplitude constant.

In conclusion, our reproduction confirms the results of the original
publication and shows that an implementation in the NEST framework is
feasible. This allows for the straightforward integration of the saccade
generator into models of the visuo-motor system forming the interface
between sensory processing and motor control.

\section{Acknowledgments}\label{acknowledgments}

All network simulations carried out with NEST
(http://www.nest-simulator.org). This research was funded by the
European Union's Horizon 2020 Research and Innovation Programme under
Grant Agreement No. 7202070 (HBP SGA1).

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2018-02-07 article/compile_pdf.sh

@@ -1,0 +1,6 @@
+pandoc --standalone --filter pandoc-fignos --filter pandoc-tablenos --filter pandoc-eqnos --template=rescience-template.tex --latex-engine=xelatex --biblatex --bibliography=senden-schuecker-hahne-diesmann-goebel-2018.bib -M "crossrefYaml=crossref.yaml" --output senden-schuecker-hahne-diesmann-goebel-2018.tex senden-schuecker-hahne-diesmann-goebel-2018.md
+
+xelatex senden-schuecker-hahne-diesmann-goebel-2018
+biber senden-schuecker-hahne-diesmann-goebel-2018
+xelatex senden-schuecker-hahne-diesmann-goebel-2018
+xelatex senden-schuecker-hahne-diesmann-goebel-2018

### 2018-02-07 code/Exp1.py
added scripts

@@ -1,0 +1,185 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 1: Activity profiles in left and right SG (figure 3 in original publication). 
+# Input I to the left side of the SG was set equal to 1 for 265 ms
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model		 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary	 ##
+###########################################
+
+# additional variables
+cm2inch		= .394 # inch/cm
+t_relax 	= 100  # ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig1.eps'
+fig_size 	= np.multiply([17.6,11.6],cm2inch)
+fig_rows 	= 6
+fig_cols 	= 2
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 			= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
+	col 		= np.mod(i,fig_cols)
+	ax[i] 		= fig.add_subplot(fig_rows,fig_cols,i+1)
+	ax[i].get_xaxis().set_visible(False)
+	ax[i].get_yaxis().set_ticks([])
+	ax[i].set_ylim([-1.,1.5])
+	ax[i].tick_params(right='off')
+	ax[i].tick_params(top='off')
+	if (col):
[…]

### 2018-02-07 code/Exp2.py

@@ -1,0 +1,147 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 2: Activity profiles in LLBN and EBN (figure 5 in original publication).  
+# LLBN and EBN discharge rates for 5, 10, and 20 degree saccades. 
+# Input to left SG was set equal to 1, 1.75, and 2.5; in each case for 85 ms
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394 # inch/cm
+t_relax 	= 100  # ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig2.eps'
+fig_size 	= np.multiply([17.6,8.5],cm2inch)
+fig_rows 	= 1
+fig_cols 	= 2
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
+	col 	= np.mod(i,2)	
+	ax[i] 	= fig.add_subplot(fig_rows,fig_cols,i+1)
+	ax[i].axhline(color='k',linestyle='--')
+	ax[i].set_xlabel('time (ms)')	
+	ax[i].set_ylim([-1.,1.5])
+	ax[i].locator_params(axis='y',nbins=3)
+
+ax[0].text(-0.075, 1.1, 'A', transform=ax[0].transAxes, size=16, weight='bold')
[…]

### 2018-02-07 code/Exp3.py

@@ -1,0 +1,150 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 3: Oblique saccades (figure 6 in original publication). 
+# Input to the horizontal (right) and vertical (up) circuits were: (0.67, 0.08), (0.7, 0.22), (0.74,0.4), (0.75, 0.6), (0.7, 0.9).
+# These inputs were left on for 75 ms.
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394 # inch/cm
+t_relax 	= 100  # ms
+g_pos		= 260. # gain eye position
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig3.eps'
+fig_size 	= np.multiply([11.6,11.6],cm2inch)
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+ax 			= fig.add_subplot(1,1,1)
+
+ax.axhline(color='k',linestyle='--')
+ax.axvline(color='k',linestyle='--')
+ax.set_xlim([-20.,20.])
+ax.set_ylim([-20.,20.])
+ax.locator_params(axis='x',nbins=5)
+ax.locator_params(axis='y',nbins=5)
+ax.set_xlabel('horizontal eye position (deg)')
+ax.set_ylabel('vertical eye position (deg)')
+
+
[…]

### 2018-02-07 code/Exp4.py

@@ -1,0 +1,140 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 4: Saccadic staircase (figure 7 in original publication). 
+# Inputs to the horizontal and vertical circuits were held at (0.2, 0.33) for 250 ms
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+from scipy.signal import argrelextrema
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394 # inch/cm
+t_relax 	= 100  # ms
+g_pos		= 260. # gain eye position
+sr 			= 40   # sampling rate
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig4.eps'
+fig_size 	= np.multiply([8.5,11.6],cm2inch)
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+ax 			= fig.add_subplot(1,1,1)
+
+ax.set_xlim([0.,10.])
+ax.set_ylim([0.,15.5])
+ax.set_xlabel('horizontal eye position (deg)')
+ax.set_ylabel('vertical eye position (deg)')
+
+
+###########################################
+#### 		set up experiment			 ##
+###########################################
+
+# input & external electric stimulation
+I_horizontal = .20
+I_vertical   = .33 
+J   		 = .0
+
+# timing protocol (in ms)
+preStim  	 =   0
+Stim     	 = 250
+postStim 	 =  50
+
[…]

### 2018-02-07 code/Exp5.py

@@ -1,0 +1,149 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 5: EBN tuning curve (figure 8 in original publication). 
+# Inputs to the horizontal and vertical circuits: 
+# (.2, 0., .63, .0),(0., 0., .7, 0.),(0., .2, .63, 0.),(0., .45, .45, 0.),
+# (0., .7, 0., 0.),(0., .45, 0., .45),(0., .2, 0., .63),(0., 0., 0., .7),
+# (.2, 0., 0., .63),(.45, 0., 0., .45),(0.63, 0., 0., .2),(0.7, 0., 0., 0.),
+# (0.63, 0., .2, 0.),(0.45, 0., .45, 0.).
+# The input was applied for 50 ms.
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394 			# inch/cm
+t_relax 	= 100  			# ms
+r			= np.zeros(15)	# EBN activity
+a			= np.zeros(15)	# saccade angle
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig5.eps'
+fig_size 	= np.multiply([8.5,8.5],cm2inch)
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
[…]

### 2018-02-07 code/Exp6.py

@@ -1,0 +1,182 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 6: Effect of stimulation frequency on saccade amplitude, duration, and velocity (figure 9 in original publication).  
+# SC stimulation frequency F was varied between 1 and 2.4 at increments of 0.2. The weight W was set equal to 2, and stimulation duration was 125 ms.
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+from scipy.signal import argrelextrema
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394			# inch/cm
+t_relax 	= 100  			# ms
+g_pos		= 260.			# gain eye position	
+threshold	= 30. 			# eye velocity marking onset of saccade		
+Amplitude	= np.zeros(8)	# saccade amplitude
+Duration	= np.zeros(8)	# saccade duration
+Velocity	= np.zeros(8)	# saccade velocity
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig6.eps'
+fig_size 	= np.multiply([17.6,8.5],cm2inch)
+fig_rows 	= 1
+fig_cols 	= 3
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
[…]

### 2018-02-07 code/Exp7a.py

@@ -1,0 +1,175 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation."
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 7: Trading saccade velocity for duration (figure 10 in original publication).
+# Only the shape of the input signal to the model was varied.
+# The weight W was set equal to 2. For the high velocity trial (dashed line), stimulation frequency F was 3,
+# and stimulation duration was 62 ms. 
+# For the low velocity trial (solid line), F was 1.3, and stimulation duration was 117 ms.
+#
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+from itertools import cycle
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394	# inch/cm
+t_relax 	= 100   # ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig7a.eps'
+fig_size 	= np.multiply([11.6,17.6],cm2inch)
+fig_rows 	= 5
+fig_cols 	= 1
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
[…]

### 2018-02-07 code/Exp7b.py

@@ -1,0 +1,170 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation."
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# As Exp7a with the difference that generation of the input is adjusted to exactly reproduce
+# the curve shown in the original publication (by reducing the time constant after stimulus offset). 
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+from itertools import cycle
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394	# inch/cm
+t_relax 	= 100   # ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig7b.eps'
+fig_size 	= np.multiply([11.6,17.6],cm2inch)
+fig_rows 	= 5
+fig_cols 	= 1
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
+	ax[i] 	= fig.add_subplot(fig_rows,fig_cols,i+1)
+	ax[i].get_xaxis().set_visible(False)
+	ax[i].get_yaxis().set_ticks([])
+	ax[i].set_ylim([-.5,2.])
+	ax[i].tick_params(right='off')
+	ax[i].tick_params(top='off')
+
+lines 	   	= ['k--','k-']
+linecycler 	= cycle(lines)
+
+
[…]

### 2018-02-07 code/Exp8.py

@@ -1,0 +1,159 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 8: Smooth staircase eye movement (figure 11 in original publication).  
+# Input I to the left side of the SG was set equal to 3 for 300 ms
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394	# inch/cm
+t_relax 	= 100  	# ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig8.eps'
+fig_size 	= np.multiply([17.6,11.6],cm2inch)
+fig_rows 	= 5
+fig_cols 	= 1
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
+	ax[i] 	= fig.add_subplot(fig_rows,fig_cols,i+1)
+	ax[i].get_xaxis().set_visible(False)
+	ax[i].get_yaxis().set_ticks([])
+	ax[i].set_ylim([-1.,1.5])
+	ax[i].tick_params(right='off')
+	ax[i].tick_params(top='off')
+
+
+###########################################
+#### 		set up experiment			 ##
+###########################################
+
[…]

### 2018-02-07 code/Exp9.py

@@ -1,0 +1,186 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Simulates reimplemented saccade generation (SG) model of Gancarz & Grossberg (1998).
+# Generates Fig. 9: Interrupted saccade resulting from OPN stimulation (figure 12 in original publication). 
+# Input to left side of saccade generator was set to 0.7 for 100 ms. 
+# OPN stimulation J was set to 1.8 for 5 ms.
+# -----------------------------------------------------------------------------
+
+import pylab as pl
+from matplotlib import rcParams
+
+
+###########################################
+#### 		set up model				 ##
+###########################################
+
+execfile('setup_model.py')
+
+
+###########################################
+#### 			auxiliary				 ##
+###########################################
+
+# additional variables
+cm2inch		= .394	# inch/cm
+t_relax 	= 100   # ms
+
+# figure setup
+rcParams.update({'figure.autolayout': True})
+
+fig_name	= '../article/figures/fig9.eps'
+fig_size 	= np.multiply([11.6,17.6],cm2inch)
+fig_rows 	= 6
+fig_cols 	= 1
+fig_plots	= fig_rows*fig_cols
+ppi			= 1200
+face	 	= 'white'
+edge	 	= 'white'
+
+ax 		 	= [None]*fig_plots
+fig 		= pl.figure(facecolor = face, edgecolor = edge, figsize = fig_size)
+for i in range(0,fig_plots):
+	ax[i] 	= fig.add_subplot(fig_rows,fig_cols,i+1)
+	ax[i].get_xaxis().set_visible(False)
+	ax[i].get_yaxis().set_ticks([])
+	ax[i].set_xlim([0,200])	
[…]

### 2018-02-07 code/setup_model.py

@@ -1,0 +1,164 @@
+# -----------------------------------------------------------------------------
+# Distributed under the GNU General Public License.
+#
+# Contributors: Mario Senden mario.senden@maastrichtuniversity.nl
+# -----------------------------------------------------------------------------
+# References:
+#
+# Gancarz, G., Grossberg, S. "A Neural Model of the Saccade Generator in the Reticular Formation." 
+# Neural Networks 11, no. 7-8 (October 1998): 1159-74. doi:10.1016/S0893-6080(98)00096-3.
+# -----------------------------------------------------------------------------
+# File description:
+# 
+# Sets up the SG model developed by Gancarz & Grossberg (1998) in PyNEST for subsequent simulation.
+# -----------------------------------------------------------------------------
+
+import nest
+import numpy as np
+
+
+###########################################
+#### general setup & NEST initialization ##
+###########################################
+
+# simulation parameters 
+dt     		=  0.05		# time step for numerical integration (in ms)
+delay  		=  0.05		# conductance delay (in ms) - setting this equal to time step implies no delay
+tau    		= 50.00		# time constant (in ms) 
+sigma  		=  0.00		# scaling factor of additive noise
+
+
+# NEST kernel initialization
+nest.ResetKernel()	 
+nest.SetKernelStatus({'resolution': dt, 'use_wfr': False})
+
+
+###########################################
+#### creation of neuron & device objects ##
+###########################################
+
+# neuron parameters
+# alpha				: upper limit for piecewise linear unit (2nd threshold for threshold_lin_rate)
+# g_in				: gain in multiplicative coupling for inhibitory input
+# lambda			: passive decay rate 
+# linear_summation	: if false, gain function is applied to input before summation; if true, it is applied afterwards
+# mult_coupling 	: if false, neuron does not exhibit multiplicative coupling; if true, it is does
+# rectify_output	: if false, activations updated according to numerical integration; if true, activations are rectified after each update
[…]

### 2018-02-07 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
updated equations

@@
 neuron (OPN) which tonically inhibits each EBN to suppress unwanted
-saccades. In what follows we briefly describe the dynamics of neurons
-within the horizontal component (equivalent descriptions apply to the
-vertical component).
+saccades. While this constitutes the core SG model, the superior
+collicuus (SC) is relevant for some simulations. In what follows we
+briefly describe the dynamics of neurons within the horizontal component
+(equivalent descriptions apply to the vertical component).
 
@@
 
-\[
+\begin{equation}
 \begin{array}{ll}
@@
 \end{array}
-\]
+\label{eq:llbn}\end{equation}
 
@@
 \end{array}
-\] with a nonlinear gain function \(g(x)\) given by
+\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
+given by
 
-\[
+\begin{equation}
 g(x) = \frac{x^4}{0.1^4+x^4} \textrm{.}
-\]
+\label{eq:gain}\end{equation}
 
@@
 \end{array}
-\]
+\label{eq:ibn}\end{equation}
 
@@
 
-\[
+\begin{equation}
 \tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
-\] where \(J\) represents external electrical stimulation.
+\label{eq:opn}\end{equation} where \(J\) represents external electrical
+stimulation and \(g(\cdot)\) again given by equation \ref{eq:gain}.
 
@@
 \end{array}
-\] Horizontal eye position (\(\theta\)) depends on activity of the right
-TN; i.e. \(\theta = 260(T_r-0.5)\).
+\label{eq:tn}\end{equation} Horizontal eye position (\(\theta\)) depends
+on activity of the right TN; i.e. \(\theta = 260(T_r-0.5)\).
+
+\subsubsection{Superior colliculus}\label{superior-colliculus}
+
+The superior colliculus (\(A\)) receives electrical stimulation F
+
+\begin{equation}
+\tau \frac{dA}{dt} =-A+F(t) \textrm{.}
+\label{eq:sc}\end{equation}
+
+\subsubsection{SC input to the saccade
+generator}\label{sc-input-to-the-saccade-generator}
+
+For a range of simulations, external input to the saccade generator is
+obtained by applying a nonlinearity \(f(\cdot)\) given by
+\begin{equation}
+f(x) = 
+\left\{
+ \begin{array}{lll}
+    0 \quad &\textrm{if} \quad x<0 \\
+    x \quad &\textrm{if} \quad 0<x<1 \\
+    1 \quad &\textrm{if} \quad x \geq 1 \\
+  \end{array}
+\right.\
+\label{eq:pw}\end{equation} to the activity \(A\) of the SC and scaling
+the result by a weight \(W\) such that
+
+\begin{equation}
+I = Wf(A)\textrm{.}
+\label{eq:stim}\end{equation}
 
@@
 In all simulations we used the Exponential Euler method for numerical
-integration of rate neurons \autocite{Hahne2016} at a time step of
+integration of rate neurons \autocite{Hahne2017} at a time step of
 \(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.90000\textwidth,height=0.90000\textwidth]{figures/fig7.eps}
+\caption{\textbf{Trading saccade velocity for duration.} Duration and
+velocity of a saccade can be traded while keeping amplitude constant. To
+produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
+to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
+publication). To produce a low velicity saccade an input of
+\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
+input curve produced by our implementation of the original equations
+(A), a second saccade can be observed for both high and low velocity
+saccades. If the input curves reported in the original manuscript were
+recreated in an ad-hoc fashion (B), the responses of all model neurons
+matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.70000\textwidth,height=0.56200\textwidth]{figures/fig8.eps}
+\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
+SG neurons accompanying smooth eyes movement as a result of strong
+(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
+input.\label{fig:fig_8}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.46200\textwidth,height=0.60000\textwidth]{figures/fig9.eps}
+\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
+OPN stimulation interrupted the saccade, which remained accurate
+nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
+for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
+input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
+\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
+uninterrupted saccade.\label{fig:fig_9}}
+\end{figure}
+
 The seventh simulation shows that saccade velocity and duration can be
@@
 to our NEST implementation by solving the expressions describing the
-input analytically. Specifically, according to equations A13-A15 of
-\autocite{Gancarz1998} the input \(I\) is described by \[
-\begin{array}{lll}
-    \tau \frac{dA}{dt} &=-A+F(t)\\
-    f(x) &=
-    \left\{
-    \begin{array}{lll}
-        0 \quad &\textrm{if} \quad x<0 \\
-        x \quad &\textrm{if} \quad 0<x<1 \\
-        1 \quad &\textrm{if} \quad x \geq 1 \\
-    \end{array}
-    \right. \\
-    I(t) &=Wf(A)
-\end{array}
-\]\\
-with the stimulus \[
-F(t)=
-\left\{
-\begin{array}{lll}
-    0 \quad &\textrm{if} \quad t<{t}_{on} \\
-    F \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.} \\
-    0 \quad &\textrm{if} \quad t \geq {t}_{off} \\
-\end{array}
-\right.
-\]\\
-Given this, the analytic expression for \(I(t)\) is \begin{equation}
+input analytically. Specifically, the analytic expression of the input
+\(I(t)\) described by equations \ref{eq:sc} - \ref{eq:stim} is
+\begin{equation}
 I(t)=
@@
 \right.
-\label{eq:1}\end{equation} Evaluating the analytic expression given
-parameter values reported in the original manuscript exactly reproduced
-our numerical results and thus produced curves equally deviatng from
-those shown in the original publication. Next, we contacted one of the
-original authors to rule out the possibility that we used erroneous
-parameter settings but no mistake was pointed out to us. The origin of
-the discrepancy thus remains unclear. However, it is likely it stems
-from the description of the input rather than from problems with the SG
-model itself.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.75000\textwidth,height=0.75000\textwidth]{figures/fig7.eps}
-\caption{\textbf{Trading saccade velocity for duration.} Duration and
-velocity of a saccade can be traded while keeping amplitude constant. To
-produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
-to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
-publication). To produce a low velicity saccade an input of
-\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
-input curve produced by our implementation of the original equations
-(A), a second saccade can be observed for both high and low velocity
-saccades. If the input curves reported in the original manuscript were
-recreated in an ad-hoc fashion (B), the responses of all model neurons
-matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
-\end{figure}
+\label{eq:analytic}\end{equation} Evaluating the analytic expression
+given parameter values reported in the original manuscript exactly
+reproduced our numerical results and thus produced curves equally
+deviating from those shown in the original publication. Next, we
+contacted one of the original authors to rule out the possibility that
+we used erroneous parameter settings but no mistake was pointed out to
+us. The origin of the discrepancy thus remains unclear. However, it is
+likely it stems from the description of the input rather than from
+problems with the SG model itself.
 
@@
 turn, in the observed smooth eye movement.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.60000\textwidth,height=0.46200\textwidth]{figures/fig8.eps}
-\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
-SG neurons accompanying smooth eyes movement as a result of strong
-(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
-input.\label{fig:fig_8}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.46200\textwidth,height=0.70000\textwidth]{figures/fig9.eps}
-\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
-OPN stimulation interrupted the saccade, which remained accurate
-nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
-for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
-input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
-\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
-uninterrupted saccade.\label{fig:fig_9}}
-\end{figure}
-
-\clearpage
 

### 2018-02-09 article/Cazé_2018.tex
update the tex (super minor changes) and simplify README

--- 2018-01-25 article/Caze_2018.tex
@@
 \end{equation}
-
 where $V(t)$ is the membrane voltage at time $t$, $\tau_m$ the membrane time constant, $V_0$ the displacement of the membrane potential due to a constant external input, $w_\text{ex}$ and $w_\text{in}$ respectively the excitatory and inhibitory weights, $\tau$ the transmission delay, and $E_i$ and $I_i$ the indices of neurons that connect with respectively excitatory and inhibitory weights to neuron $i$.
@@
 \]
-
 For this non-linear function, the biological interpretation is that below a first threshold $V_a$, excitatory inputs are propagated passively by the dendrite and sum linearly.
@@
 \end{code}
-
 Note that the delay has been set as one time step (\verb|dt|) less than \SI{5}{\milli\second}.
@@
 \end{code}
-
 To facilitate simulation, we use the \emph{joblib} library to automatically store results and to distribute computation over processor cores.
@@
 We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}).
-To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows.
-First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.
-We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential.
-From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
-\begin{equation}
-F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
-\end{equation}
-Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e.
+ $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e.
 $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$.
@@
 where $U_1$, and $U_2$ stand for the fraction of trials displaying unstable network activity before respectively after onset of propagation, and $S$ and $E$ for the fraction of trials displaying stable background activity with respectively without persistent propagation.
-We obtained this scheme directly from the authors of the original study, it was not mentioned in the publication.
+We obtained this scheme directly from the authors of the original study, it was only coarsely described in the original publication.
 Combining individual results in this way does not necessarily lead to easily interpretable colours, but using a different scheme, e.g.
@@
 The variable $g_\text{p}$ refers to the number of successful propagations of the synchronous activity, i.e.
-the number of time steps $t_\text{p}^k$ where number of active neurons exceeds the maximal background activity $a_b$.}\label{tab:color}
+the number of time steps $t_\text{p}^k$ where the number of active neurons exceeds the maximal background activity $a_b$.}\label{tab:color}
 \end{table}
@@
 These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision.
-However, we tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained similar results (not shown).
+We tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained qualitatively the same results (not shown).
 Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.

### 2018-02-09 article/Caze_2018.tex
Readd accidentally deleted text

@@
 For non-linearly coupled networks, we have to use a slightly more complicated approach where we sum up all the received excitatory input in a variable \verb|ve|, apply the non-linear function $\sigma$ to this input, and then reset the variable \verb|ve| back to 0 to prepare for the next time step.
-In Brian syntax (the definition of the inhibitory synapses is identical to the linearly-coupled network):
+In Brian syntax (the definition of the inhibitory synapses is identical to the linearly coupled network):
 \begin{code}

### 2018-02-14 article/Mondeel_Ogundipe_Westerhoff-2017.tex
RE: Federico
- Discussion of the flux variability cobrapy function.
- bold math notation in the definition of the FVA problem (changed the ReScience template to include the bm package)
- consistent definition of exchange reactions in the notebooks
- Some references to the manuscript in the notebook comments.
- extra discussion of phi = 0 or phi > 0
- Redraw the Fig. 1 network and include the fluxes

@@
 \usepackage{hyperref}
+\usepackage{bm}
 \hypersetup{breaklinks=true,
@@
 extracellular concentration in the mutant case if, when plotting the
-wild-type and mutant flux-variability intervals on a horizontal axis
-(see below), both the lower and upper boundary of the mutant interval
-are shifted to the right, or left, respectively. If the two borders of
-the mutant interval move in opposite directions with respect to the
-wild-type interval, the result is scored as `unchanged' and the boundary
-metabolite is not considered to be a biomarker.
+wild-type and mutant flux-variability intervals on a horizontal axis,
+both the lower and upper boundary of the mutant interval are shifted to
+the right, or left, respectively (see Figures
+~\ref{fig:Shlomi2009_figure_1}B and ~\ref{fig:Shlomi2009_figure_3}B). If
+the two borders of the mutant interval move in opposite directions with
+respect to the wild-type interval, the result is scored as `unchanged'
+and the boundary metabolite is not considered to be a biomarker.
 
@@
 \begin{align}
-\text{Maximize } &\pm v_i, \text{ such that for all k} \nonumber \\
-S v &= 0 \nonumber \\
+\text{Maximize } &\pm v_i, \text{ such that for all } k \nonumber \\
+\bm{S} \bm{v} &= \bm{0} \nonumber \\
 Z  &\geq \phi Z_{\text{max}} \nonumber \\
@@
 
-Here \(v\) is the column vector of fluxes representing all reactions in
-the model, \(v_i\) is the flux of the exchange reaction of a given
-boundary metabolite \(i\), \(\alpha_k\) is the, possibly negative, lower
-bound for reaction \(k\) and similarly \(\beta_k\) is the, possibly
-negative, upper bound for reaction \(k\). The reaction bounds are the
-\(V_{max}\)'s of the reactions and are part of the metabolic network
-definition, typically they are set to either \(0\) or \(\pm 1000\) if
-the true, biological, \(V_{max}\)'s are not known. These values allow
-specification of irreversible reactions by setting the lower bound (or
-upper bound) to zero. The index \(i\) specifies a single reaction that
-is to be maximized, whereas the index \(k\) is used to index the bound
-constraints that exist for each reaction, including \(v_i\). \(v_i\) may
-be maximized in the forward (positive flux) or the reverse (negative
-flux) direction if allowed by the bounds on \(v_i\). This is indicated
-by the \(\pm\) symbol. \(S\) is the stoichiometry matrix for the
-metabolic network of size \(mxr\) where \(m\) is the number of
-metabolites and \(r\) the number of reactions. The numeric, typically
-integer, elements of \(S\) represent the stoichiometry coefficients of
-each metabolite \(m\) in each reaction \(r\). Consequently, for large
-networks, this matrix is sparse (has many zero entries). \(Z = c^T ~ v\)
-is the objective function defined for the map, entailing a linear
-combination, defined by column vector \(c\), of one or more reactions.
-The vector \(c\) is an indicator of objective reactions, i.e.~it
-contains a value of \(1\) in rows corresponding to fluxes that are to be
-included in the objective, and zeros in all other rows. Due to the
-vector multiplication \(c^T ~ v\), the objective \(Z\) sums the fluxes
-of the reactions that correspond to rows containing a value of \(1\) in
-\(c\). \(Z_{\text{max}}\) denotes the maximal value of the objective
-function. We define \(0 \leq \phi \leq 1\) to be an arbitrary minimal
-fraction of the maximal objective function value that has to be
+Here \(\bm{v}\) is the column vector of fluxes representing all
+reactions in the model, \(v_i\) is the flux of the exchange reaction of
+a given boundary metabolite \(i\), \(\alpha_k\) is the, possibly
+negative, lower bound for reaction \(k\) and similarly \(\beta_k\) is
+the, possibly negative, upper bound for reaction \(k\). The reaction
+bounds are the \(V_{max}\)'s of the reactions and are part of the
+metabolic network definition, typically they are set to either \(0\) or
+\(\pm 1000\) if the true, biological, \(V_{max}\)'s are not known. These
+values allow specification of irreversible reactions by setting the
+lower bound (or upper bound) to zero. The index \(i\) specifies a single
+reaction that is to be maximized, whereas the index \(k\) is used to
+index the bound constraints that exist for each reaction, including
+\(v_i\). \(v_i\) may be maximized in the forward (positive flux) or the
+reverse (negative flux) direction if allowed by the bounds on \(v_i\).
+This is indicated by the \(\pm\) symbol. \(\bm{S}\) is the stoichiometry
+matrix for the metabolic network of size \(m \times r\) where \(m\) is
+the number of metabolites and \(r\) the number of reactions. The
+numeric, typically integer, elements of \(\bm{S}\) represent the
+stoichiometry coefficients of each metabolite \(m\) in each reaction
+\(r\). Consequently, for large networks, this matrix is sparse (has many
+zero entries). \(Z = \bm{c}^T \bm{v}\) is the objective function defined
+for the map, entailing a linear combination, defined by column vector
+\(\bm{c}\), of one or more reactions. The vector \(\bm{c}\) is an
+indicator of objective reactions, i.e.~it contains a value of \(1\) in
+rows corresponding to fluxes that are to be included in the objective,
+and zeros in all other rows. Due to the vector multiplication
+\(\bm{c}^T \bm{v}\), the objective \(Z\) sums the fluxes of the
+reactions that correspond to rows containing a value of \(1\) in
+\(\bm{c}\). \(Z_{\text{max}}\) denotes the maximal value of the
+objective function. We define \(0 \leq \phi \leq 1\) to be an arbitrary
+minimal fraction of the maximal objective function value that has to be
 achieved. When \(\phi=1\) we ask for the range of flux allowed for
@@
 maintained). The latter, i.e. \(\phi=0\), is the case considered in the
-method by Shlomi et al.
+method by Shlomi et al. The choice of \(\phi\) matters because when
+\(phi \neq 0\) we are requiring flux to flow through the set of
+objective reactions. This may entail a forced directionality through
+reactions that are required to ultimately produce flux in the objective
+reactions. These additional limitation in freedom of the flux pattern
+may subsequently impact the biomarker predictions. The approach
+discussed here entails predictions purely based on the network topology
+and therefore sets \(\phi = 0\).
 
@@
 
-After the loop is completed, contradictory predictions for \(m\) between
-the affected reactions \(r\) are dealt with based on a majority rule.
-When there is an equal number of elevated and reduced predictions, the
-boundary metabolite is considered to be unchanged, i.e.~not a biomarker.
+After the loop (steps 1-5) is completed, contradictory predictions for
+\(m\) between the affected reactions \(r\) are dealt with based on a
+majority rule. When there is an equal number of elevated and reduced
+predictions, the boundary metabolite is considered to be unchanged,
+i.e.~not a biomarker.
+
+The flux variability analyses that yield the exchange flux intervals
+computed in step 1, 2 and 4 are handled by the
+flux\_analysis.variability.flux\_variability\_analysis function in
+Cobrapy. The input for this function comprises the model, along with its
+associated flux bounds which change during step 1, 2 and 4, the list of
+reactions for which to calculate the flux intervals and a fraction of
+the optimal objective flux to minimally obtain (fraction\_of\_optimum).
+In our case we pass the model with flux bounds as described above, lists
+of exchange reactions for the metabolites of interest and a fraction of
+optimum of zero.
 
@@
 \centering
-\includegraphics{Figures/scheme_shlomi2009_figure1}
-\caption{A metabolic network used to illustrate the predictions of
+\includegraphics{Figures/Reproduction_Shlomi_fig_1}
+\caption{(A) A metabolic network used to illustrate the predictions of
 biomarkers of IEMs as originally considered in the work by Shlomi et al.
@@
 M6. The IEM considered is one that leads to deletion of the enzyme
-converting M1 plus M3 to M2 plus M4.}\label{fig:Shlomi2009_figure_1}
+converting M1 plus M3 to M2 plus M4. (B) Visual illustration of the
+predicted exchange intervals for each of the metabolites with an
+exchange reaction. The wild-type (WT) interval is indicated in black,
+the mutant interval in red. Each interval is associated with a numeric
+interval predicted by the biomarker prediction algorithm and with a
+qualitative prediction for the level of the metabolite in the biofluids
+(see methods). H.C. is an abbreviation for a highly confident biomarker,
+see main text.}\label{fig:Shlomi2009_figure_1}
 \end{figure}
@@
 
-The network in Figure ~\ref{fig:Shlomi2009_figure_1} consists of seven
+The network in Figure ~\ref{fig:Shlomi2009_figure_1}A consists of seven
 metabolites, \(6\) of which (i.e.~all except M3) have exchange
@@
 Shlomi et al. graphically provide the biomarker prediction for this
-network in their Figure 1A. Here, we do so numerically and this paper is
-accompanied by a Jupyter notebook reproducing the biomarker prediction.
-In our analysis, all exchange reactions were given inward bounds of
-\(-10\) and outward bounds of \(+1000\) so that the internal reaction
-bounds (of \(\pm 1000\)) are never reached. Our Table
-~\ref{tbl:toy_model_results} shows that all exchangeable metabolites
-except M5 and M7 are predicted to be robust biomarkers.
+network in their Figure 1A. Here, we do so graphically and numerically
+in Figure ~\ref{fig:Shlomi2009_figure_1}B and this paper is accompanied
+by a Jupyter notebook reproducing the biomarker prediction. In our
+analysis, all exchange reactions were given inward bounds of \(-10\) and
+outward bounds of \(+1000\) so that the internal reaction bounds (of
+\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
+shows that all exchangeable metabolites except M5 and M7 are predicted
+to be robust biomarkers.
 
@@
 summarized here in Table ~\ref{tbl:amino_acid_iems}.
-
-\hypertarget{tbl:toy_model_results}{}
-\begin{longtable}[]{@{}lllc@{}}
-\caption{\label{tbl:toy_model_results}Biomarker predictions using our
-implementation with respect to the mutation affecting the conversion of
-M1 plus M3 to M2 plus M4 in the network of Figure
-~\ref{fig:Shlomi2009_figure_1}. H.C. stands for `highly confident'
-indicating, as defined by Shlomi et al., no overlap between the WT and
-mutant intervals. Between brackets {[}{]} we report the lower and upper
-bounds of the FVA intervals we obtained. }\tabularnewline
-\toprule
-ID & Prediction & WT & Mutant\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-ID & Prediction & WT & Mutant\tabularnewline
-\midrule
-\endhead
-M1 & H.C. Elevated & {[}-10.0, -1.0{]} & {[}0.0, 0.0{]}\tabularnewline
-M2 & H.C. Reduced & {[}1.0, 10.0{]} & {[}0.0, 0.0{]}\tabularnewline
-M4 & Reduced & {[}1.0, 20.0{]} & {[}0.0, 10.0{]}\tabularnewline
-M6 & Elevated & {[}-10.0, 9.0{]} & {[}0.0, 10.0{]}\tabularnewline
-M7 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
-M5 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
-\bottomrule
-\end{longtable}
 

### 2018-02-15 article/Mondeel_Ogundipe_Westerhoff-2017.tex
RE: federico
Inclusion of figure 3

@@
 
-\section{Conclusion}\label{conclusion}
-
-The results summarized in Figures 1 and 2 of \autocite{Shlomi2009} have
-been successfully replicated with the software tools provided here.
-Overall, the reproduction process went smoothly, since the method was
-applied to both an illustrative example and to the human metabolic
-reconstruction, resulting in a multitude of test cases for the
-implementation.
-
-The obstacles we encountered were the lack of explicit discussion on the
-set of fluxes allowed into and out of the model and a set of
-discrepancies between the gene-reaction mappings listed in the
-supplementary Excel file and the actual gene-reaction mappings in the
-metabolic network. However, trial and error and manual curation solved
-these issues.
-
 \hypertarget{tbl:amino_acid_iems}{}
@@
 
+\subsection{\texorpdfstring{Reproduction of Figure 3
+\autocite{Shlomi2009}}{Reproduction of Figure 3 {[}@Shlomi2009{]}}}\label{reproduction-of-figure-3-shlomi2009}
+
+\begin{figure}
+\centering
+\includegraphics{Figures/Reproduction_Shlomi_fig_3}
+\caption{Reproduction of Figure 3 in \autocite{Shlomi2009}. (A)
+Illustration of the subnetwork of Recon 1 relevant to the effect of
+homocystinuria on the metabolism and transport of methionine. Nodes
+represent metabolites and edges represent reactions. For simplicity,
+only abbreviations of metabolite names are given in the Figure. The full
+names are: L-Methionine (met), S-Adenosyl-L-methionine (amet),
+S-Adenosyl-L-homocysteine (ahcys), L-Homocysteine (hcys), L-Serine
+(ser), L-Cystathionine (cyst), L-Cysteine (cys),
+S-Adenosylmethioninamine (ametam), 5-Methylthioadenosine (5mta),
+5-Methylthio-5-deoxy-D-ribose 1-phosphate (5mdr1p),
+5-Methylthio-5-deoxy-D-ribulose 1-phosphate (dmdru1p),
+2,3-diketo-5-methylthio-1-phosphopentane (dkmpp),
+2-keto-4-methylthiobutyrate (2kmb). Metabolites with a green background
+color participate in additional reactions not shown here. Reactions are
+indicated by their identifiers in Recon 1 and with a corresponding
+enzyme commision number if available. We highlighted the mutations for
+Homocystinuria (dysfunctional CBS) and hypermethioninemia (dysfunctional
+AHCY). (B) Prediction of concentration changes of methionine in
+homocystinuria and S-adenosylhomocysteine hydrolase deficiency based on
+the biomarker prediction algorithm.}\label{fig:Shlomi2009_figure_3}
+\end{figure}
+
+Figure 3 in the original publication entailed an in-depth look at the
+results highlighted in Figure 2 for methionine metabolism. We reproduced
+this Figure as Figure ~\ref{fig:Shlomi2009_figure_3}. The numerical
+results illustrated in ~\ref{fig:Shlomi2009_figure_3}B may be reproduced
+using the notebook `Reproduce\_figure2\_Shlomi2009', the detailed
+results may be inspected using the last cell of that notebook. We note
+that in the original publication the WT interval for the CBS IEM was
+drawn as an interval whereas our results indicate an interval of
+\([-1,-1]\).
+
+\section{Conclusion}\label{conclusion}
+
+The results summarized in Figures 1 (panel A and B), 2 and 3 of
+\autocite{Shlomi2009} have been successfully replicated with the
+software tools provided here. Overall, the reproduction process went
+smoothly, since the method was applied to both an illustrative example
+and to the human metabolic reconstruction, resulting in a multitude of
+test cases for the implementation.
+
+The obstacles we encountered were the lack of explicit discussion on the
+set of fluxes allowed into and out of the model and a set of
+discrepancies between the gene-reaction mappings listed in the
+supplementary Excel file and the actual gene-reaction mappings in the
+metabolic network. However, trial and error and manual curation solved
+these issues.
+
 {\sffamily \small

### 2018-02-15 article/Mondeel_Ogundipe_Westerhoff-2017.tex
small changes
- Vivian has moved to Groningen
- flux bounds of -1 not 1
- fix missing table reference
- consistently report we reproduce fig. 1 through 3.

@@
                         \author[1,*]{Thierry D.G.A. Mondeel}
-                    \author[1]{Vivian Ogundipe}
-                    \author[1, 2, 3]{Hans V. Westerhoff}
+                    \author[2]{Vivian Ogundipe}
+                    \author[1, 3, 4]{Hans V. Westerhoff}
                             \affil[1]{Synthetic Systems Biology and Nuclear Organization, Swammerdam Institute
 for Life Sciences, University of Amsterdam, The Netherlands}
-                    \affil[2]{Molecular Cell Physiology, Amsterdam Institute for Molecules, Medicines
+                    \affil[2]{Department of Cell Biology, University Medical Center Groningen,
+University of Groningen, The Netherlands}
+                    \affil[3]{Molecular Cell Physiology, Amsterdam Institute for Molecules, Medicines
 and Systems, VU University Amsterdam, The Netherlands}
-                    \affil[3]{Manchester Centre for Integrative Systems Biology, School of Chemical
+                    \affil[4]{Manchester Centre for Integrative Systems Biology, School of Chemical
 Engineering and Analytical Science, The University of Manchester, UK}
@@
 Each biomarker will be associated with a prediction for being either
-elevated or decreased in the mutant case, as comparted to the wild type.
+elevated or decreased in the mutant case, as compared to the wild type.
 A biomarker is considered to have an increased, or reduced,
@@
 method by Shlomi et al. The choice of \(\phi\) matters because when
-\(phi \neq 0\) we are requiring flux to flow through the set of
-objective reactions. This may entail a forced directionality through
-reactions that are required to ultimately produce flux in the objective
-reactions. These additional limitation in freedom of the flux pattern
-may subsequently impact the biomarker predictions. The approach
-discussed here entails predictions purely based on the network topology
-and therefore sets \(\phi = 0\).
+\(\phi \neq 0\) we are requiring flux to flow through the set of
+objective reactions (unless \(Z_{\text{max}} = 0\)). This may entail a
+forced directionality through reactions that are required to ultimately
+produce flux in the objective reactions. These additional limitations in
+the freedom of the flux pattern may subsequently impact the biomarker
+predictions. The approach discussed here entails predictions purely
+based on the network topology and therefore sets \(\phi = 0\).
 
@@
 originally proposed by Shlomi et al., programmed in Python, and
-reproduce the results presented in Figure 1A-B and 2 of the original
+reproduce the results presented in Figure 1A-B, 2 and 3 of the original
 publication. These figures concern the method's application to a simple
@@
 Jupyter notebooks are provided that generate the illustrative map and
-reproduce all results referred to in Figures 1A-B and 2 of Shlomi et al.
-Our implementation is based on information given in the original paper.
-However, some details critical to the reproduction were left unexplained
-there and are highlighted here. We also include a more detailed
-discussion of one of the amino acid IEMs to examine the robustness of
-the predictions made using this approach. We feel this adds to the
-utility of this reference implementation and the understanding of the
-approach introduced by Shlomi et al.
+reproduce all results referred to in Figures 1A-B, 2 and 3 of Shlomi et
+al. Our implementation is based on information given in the original
+paper. However, some details critical to the reproduction were left
+unexplained there and are highlighted here. We also include a more
+detailed discussion of one of the amino acid IEMs to examine the
+robustness of the predictions made using this approach. We feel this
+adds to the utility of this reference implementation and the
+understanding of the approach introduced by Shlomi et al.
 
@@
 is easily facilitated with our reference implementation. Subsequently,
-we discuss the reproduction of Figures 1 and 2 from the original
+we discuss the reproduction of Figures 1, 2 and 3 from the original
 publication.
@@
 reactions with an upper bound of \(1000\) and influx through all
-exchange reactions with a lower bound of \(1\) and we use the default
+exchange reactions with a lower bound of \(-1\) and we use the default
 settings, including a 10\% significance threshold. The only considered
@@
 outward bounds of \(+1000\) so that the internal reaction bounds (of
-\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
+\(\pm 1000\)) are never reached. Figure ~\ref{fig:Shlomi2009_figure_1}B
 shows that all exchangeable metabolites except M5 and M7 are predicted
@@
 chosen medium in their publication but we were able to reproduce their
-results by setting all inward flux bounds to \(1\) and all outward flux
+results by setting all inward flux bounds to \(-1\) and all outward flux
 bounds to \(1000\).
@@
 represent metabolites and edges represent reactions. For simplicity,
-only abbreviations of metabolite names are given in the Figure. The full
+only abbreviations of metabolite names are given in the figure. The full
 names are: L-Methionine (met), S-Adenosyl-L-methionine (amet),

### 2018-02-20 article/Mondeel_Ogundipe_Westerhoff-2017.tex
RE: Aaron 2.0
- Small text changes
- medium is not -10 inward not -1. This results in correct predictions for Figure 3.
- Include text and code to show result for -1 related to Figure 3
- Fix ordering Figure 1

@@
 respect to known biomarkers, but did not provide any formal proof. For a
-set of 17 IEMs that affect amino acid metabolism, they found a 56\%
-fraction of the biomarkers that are correctly predicted and a fraction
-of 76\% of predicted biomarkers that were predicted correctly in terms
-of whether they increased or decreased in concentration. Thiele et al.
-\autocite{Thiele2013} later found an even higher precision using
-essentially the same FVA methodology on an updated version of the
-metabolic reconstruction.
+set of 17 IEMs that affect amino acid metabolism, they found 56\% of the
+biomarkers that are correctly predicted and 76\% of predicted biomarkers
+that were predicted correctly in terms of whether they increased or
+decreased in concentration. Thiele et al. \autocite{Thiele2013} later
+found an even higher precision using essentially the same FVA
+methodology on an updated version of the metabolic reconstruction.
 
@@
 reactions are represented by non-mass-balanced pseudoreactions, e.g.
-\(X \leftrightarrow \varnothing\), where \(X\) is understood to
-originate from or efflux to the biofluids. Positive flux indicates net
-secretion of \(X\) by the cell and negative flux indicates the net
-uptake of \(X\).
+\(X \leftrightarrow \varnothing\). Positive flux indicates net secretion
+of \(X\) by the cell and negative flux indicates the net uptake of
+\(X\).
 
@@
 \begin{align}
-\text{Maximize } &\pm v_i, \text{ such that for all } k \nonumber \\
+\text{Min } v_i \text{ or Max } &v_i, \text{ such that for all } k \nonumber \\
 \bm{S} \bm{v} &= \bm{0} \nonumber \\
@@
 reverse (negative flux) direction if allowed by the bounds on \(v_i\).
-This is indicated by the \(\pm\) symbol. \(\bm{S}\) is the stoichiometry
-matrix for the metabolic network of size \(m \times r\) where \(m\) is
-the number of metabolites and \(r\) the number of reactions. The
-numeric, typically integer, elements of \(\bm{S}\) represent the
-stoichiometry coefficients of each metabolite \(m\) in each reaction
-\(r\). Consequently, for large networks, this matrix is sparse (has many
-zero entries). \(Z = \bm{c}^T \bm{v}\) is the objective function defined
-for the map, entailing a linear combination, defined by column vector
-\(\bm{c}\), of one or more reactions. The vector \(\bm{c}\) is an
+For a network with \(m\) metabolites and \(r\) reactions, i.e.~fluxes,
+\(\bm{S}\) is the stoichiometry matrix for the network of size
+\(m \times r\). The numeric, typically integer, elements of \(\bm{S}\)
+represent the stoichiometry coefficients of each metabolite \(m\) in
+each reaction \(r\). \(Z = \bm{c}^T \bm{v}\) is the objective function
+defined for the map, entailing a linear combination, defined by column
+vector \(\bm{c}\), of one or more reactions. The vector \(\bm{c}\) is an
 indicator of objective reactions, i.e.~it contains a value of \(1\) in
@@
 2.2 using the synchronous approach. Column 3 in Table
-{[}tbl:PKU\_results{]} shows that the result now agrees with the Recon 1
+~\ref{tbl:PKU_results} shows that the result now agrees with the Recon 1
 asynchronous predictions. Using Recon 1, we came to the correct
@@
 
-\begin{align*}
-&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
-&\leftrightarrow \\
-&\text{4a-hydroxy-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-tyrosine}
-\end{align*}
+\begin{gather*}
+\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
+\updownarrow \\
+\text{4a-hydroxy-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-tyrosine}
+\end{gather*}
 
@@
 
-\begin{align*}
-&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
-&\leftrightarrow \\
-&\text{6,7-dihydrobiopterin} + \text{L-tyrosine} + \text{H}_2\text{O}.
-\end{align*}
+\begin{gather*}
+\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
+ \updownarrow \\
+\text{6,7-dihydrobiopterin} + \text{L-tyrosine} + \text{H}_2\text{O}.
+\end{gather*}
 
@@
 that generates it and a Jupyter notebook reproducing the biomarker
-prediction.
+prediction. The image itself was produced using Adobe Illustrator.
 
@@
 chosen medium in their publication but we were able to reproduce their
-results by setting all inward flux bounds to \(-1\) and all outward flux
-bounds to \(1000\).
+results by setting all inward flux bounds to \(-10\) and all outward
+flux bounds to \(1000\).
 
@@
 homocystinuria and S-adenosylhomocysteine hydrolase deficiency based on
-the biomarker prediction algorithm.}\label{fig:Shlomi2009_figure_3}
+the biomarker prediction algorithm. The wild-type (WT) interval is
+indicated in black, the mutant interval in red. Each interval is
+associated with a numeric interval predicted by the biomarker prediction
+algorithm and with a qualitative prediction for the level of the
+metabolite in the biofluids (see methods). H.C. is an abbreviation for a
+highly confident biomarker, see main
+text.}\label{fig:Shlomi2009_figure_3}
 \end{figure}
@@
 Figure 3 in the original publication entailed an in-depth look at the
-results highlighted in Figure 2 for methionine metabolism. We reproduced
-this Figure as Figure ~\ref{fig:Shlomi2009_figure_3}. The numerical
-results illustrated in ~\ref{fig:Shlomi2009_figure_3}B may be reproduced
-using the notebook `Reproduce\_figure2\_Shlomi2009', the detailed
-results may be inspected using the last cell of that notebook. We note
-that in the original publication the WT interval for the CBS IEM was
-drawn as an interval whereas our results indicate an interval of
-\([-1,-1]\).
+effect of inborn errors in AHCY and CBS on methionine metabolism. We
+reproduced this Figure as Figure ~\ref{fig:Shlomi2009_figure_3}. We used
+Adobe Illustrator to redraw the network from the original publication
+and based the intervals in panel B on the numerical results the
+algorithm predicts. The qualitative results are already contained in
+~\ref{tbl:amino_acid_iems}. The numerical results illustrated in
+~\ref{fig:Shlomi2009_figure_3}B may be reproduced using the notebook
+\emph{Reproduce\_figure\_2\_and\_3\_Shlomi2009.ipynb}. For clarity we
+included a section in the notebook that details the results relevant to
+Figure ~\ref{fig:Shlomi2009_figure_3}B.
+
+We note that in early versions of this reproduction we were utilising a
+medium that allowed influx of \(-1\) for all medium components, as
+opposed to \(-10\) now. It turns out that a setting of \(-1\) does
+accurately reproduce the results for Figure 2, but not for Figure 3. For
+the CBS IEM the WT interval would correspond to \([-1,-1]\) and would
+therefore be a point instead of a wide interval as Shlomi et al. draw
+it. We included this simulation in the last cell of the notebook
+\emph{Reproduce\_figure\_2\_and\_3\_Shlomi2009.ipynb} This is another
+example of the inherent sensitivity of the approach to various settings
+in the algorithm.
 
@@
 
+We hope this reference implementation is useful to the community and can
+serve as a starting point for further implementations of this approach
+and serves as a warning of the accompanying sensitivities to various
+algorithm settings we have lain out in this manuscript.
+
 {\sffamily \small

### 2018-02-20 article/Mondeel_Ogundipe_Westerhoff-2017.tex
Additional explanation of the Figure 3 interval prediction

@@
 therefore be a point instead of a wide interval as Shlomi et al. draw
-it. We included this simulation in the last cell of the notebook
-\emph{Reproduce\_figure\_2\_and\_3\_Shlomi2009.ipynb} This is another
-example of the inherent sensitivity of the approach to various settings
-in the algorithm.
+it. The reason for this stems from the forced flux \(\epsilon = 1\) in
+WT. When the maximum influx of methionine is \(-1\) (as a medium
+component) it is also required at a flux of \(-1\) due to the forced
+active flux in the WT. When the maximum influx is increased to \(-10\)
+the lower bound of the predicted interval increases to \(-10\) but the
+upper bound does not because \(\epsilon\) is still equal to \(1\).
+
+We included this simulation in the last cell of the notebook. This is
+another example of the inherent sensitivity of the approach to various
+settings in the algorithm.
 

### 2018-02-21 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
updates according to Tiziano Zito's suggestions

@@
 
-We provide an implementation of the saccade generator (SG); a rate
-neuron model of neural circuitry in the reticular formation proposed by
-Gancarz \& Grossberg \autocite{Gancarz1998}. The SG is an integral part
-of the eye movement system \autocite{Grossberg2012} and as such vital
-for developing large-scale architectures of visuo-motor integration. For
-this reason it is of interest to implement the model in a publicly
-available, widely used, and actively developed neural simulation
-framework such as NEST \autocite{Gewaltig2007}. We show that the model
-translates well to the NEST framework as our implementation faithfully
-reproduces all simulation results reported in the original publication.
-Our code uses the Python interface \autocite{Eppler2008} for legibility
-with both model and analysis scripts being implemented using Python
-2.7.12.
+In 2013 the European Comission launched the Human Brain Project
+(\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
+research infrastructure spurring pan-european collaboration.
+Collaboration at such a scale, involving both empirically-oriented and
+theoretical neuroscientists, offers the unique opportunity to develop
+large-scale models of the brain. Specifically, individual research
+groups might develop models of distinct brain regions which can
+subsequently be combined into a unified whole. To facilitate this
+colaboration, the HBP encourages the utilization of publicly available,
+widely used, and actively developed neural simulation frameworks such as
+NEST \autocite{Gewaltig2007}. In light of this, the NEST framework has
+recently been extended to support simulation of functionally inspired
+rate neuron models in addition to biologically grounded spiking neuron
+models \autocite{Hahne2017}. Here we make use of this new functionality
+of the NEST framework (v2.16.0) to provide an implementation of the
+saccade generator (SG); a rate neuron model of neural circuitry in the
+reticular formation proposed by Gancarz \& Grossberg
+\autocite{Gancarz1998}. The SG is an integral part of the eye movement
+system \autocite{Grossberg2012} and as such vital for developing
+large-scale architectures of visuo-motor integration. We show that the
+model translates well to the NEST framework as our implementation
+faithfully reproduces all simulation results reported in the original
+publication. Our code uses the Python interface \autocite{Eppler2008}
+for legibility with both model and analysis scripts being implemented
+using Python 2.7.12.
 
@@
 \begin{array}{ll}
-\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5E_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
-\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5E_r+1)-(E_r+1)(10L_l+20g(P))) \\
+\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5L_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
+\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5L_r+1)-(E_r+1)(10L_l+20g(P))) \\
 \end{array}
@@
 
-In all simulations we used the Exponential Euler method for numerical
-integration of rate neurons \autocite{Hahne2017} at a time step of
-\(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
-should be noted that rates of all neurons were initialized to zero and
-were thus not at resting equilibrium. In order to address this, the
-model was allowed to evolve for \(100\,\mathrm{ms}\), to relax towards
-equilibrium before applying any input. Furthermore, we always simulated
-the full model; i.e.~both its horizontal and vertical components even if
-input was applied only to one of the two.
+In all simulations we used the Exponential Euler (EE) method for
+numerical integration of all except of tonic neurons
+\autocite{Hahne2017} at a time step of \(0.05\,\mathrm{ms}\) and a time
+constant of \(50\,\mathrm{ms}\). Tonic neurons were integrated using the
+Euler-Maruyama (EM) method since they do not exhibit passive decay
+leading to division-by-zero when using the EE method. It should be noted
+that initial firing rates of neurons were not at resting equilibrium. In
+order to address this, the model was allowed to evolve for
+\(100\,\mathrm{ms}\), to relax towards equilibrium before applying any
+input. Furthermore, as in the original publication, all activations were
+bounded from below at zero. Finally, we always simulated the full model;
+i.e.~both its horizontal and vertical components even if input was
+applied only to one of the two.
 
@@
 feasible. This allows for the straightforward integration of the saccade
-generator into models of the visuo-motor system forming the interface
-between sensory processing and motor control.
+generator with computational models of other components of the
+visuo-motor system (e.g.~salience computation) within a shared
+framework.
 
@@
 
+\section{Appendix}\label{appendix}
+
+\subsection{A1 - Rate neurons in NEST}\label{a1---rate-neurons-in-nest}
+
+For a detailed, comprehensive, treatment of rate-based neuron models in
+NEST see Hahne et al.\autocite{Hahne2017}. Here we provide only a brief
+overview to facilitate readability/interpretability of the code provided
+with this publication. In general, rate-based neurons in NEST consist of
+two components; a neuron and a gain function.
+
+\subsubsection{Neuron}\label{neuron}
+
+Two types of neuron base classes exist in NEST v2.16.0:
+\texttt{rate\_neuron} and\\
+\texttt{rate\_transformer\_node}. The \texttt{rate\_neuron} implements
+continuous rate dynamics with and without multiplicative coupling. It
+can apply a nonlinear gain function to its inputs either before or after
+their weighted summation. It is numerically integrated using the EE
+method unless a neuron does not exhibit exponential decay in which case
+the EM method is used. Furthermore, rates can be bounded (from below at
+zero) or unbounded. Finally, noise (in the form of a Wiener process) can
+be added to its input or to its output. The \texttt{rate\_neuron} has a
+total of seven parameters (see table \ref{tbl:table_2}) and two further
+properties \texttt{rate} (current firing rate) and \texttt{noise}
+(current noise level).
+
+\begin{longtable}[]{@{}lll@{}}
+\caption{\texttt{rate\_neuron} parameters.
+\label{tbl:table_2}}\tabularnewline
+\toprule
+parameter & default value & description\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+parameter & default value & description\tabularnewline
+\midrule
+\endhead
+\texttt{tau} & \(10\,\mathrm{ms}\) & time constant\tabularnewline
+\texttt{lambda} & \(1\) & passive decay rate\tabularnewline
+\texttt{std} & \(1\) & noise standard deviation\tabularnewline
+\texttt{mean} & \(0\) & mean firing rate\tabularnewline
+\texttt{linear\_summation} & \texttt{true} & if true, apply gain
+function after summation\tabularnewline
+\texttt{rectify\_output} & \texttt{false} & if true, multiplicative
+coupling\tabularnewline
+\texttt{mult\_coupling} & \texttt{false} & if true, rate is
+bounded\tabularnewline
+\bottomrule
+\end{longtable}
+
+For the implementation of this neuron see
+\texttt{rate\_neuron\_ipn\_impl.h} and\\
+\texttt{rate\_neuron\_opn\_impl.h} in
+\texttt{../nest/nest-simulator-master/models} for input and output
+noise, respectively.
+
+The \texttt{rate\_transformer\_node} instantaneously applies a
+nonlinearity to its inputs but does not exhibit temporal dynamics. It
+can be used to apply different nonlinearities to different inputs to a
+single neuron. This class has only the \texttt{rate} property and no
+parameters. For its implementation see
+\texttt{rate\_transformer\_node\_impl.h} in
+\texttt{../nest/nest-simulator-master/models}.
+
+\subsubsection{Gain function}\label{gain-function}
+
+Five gain functions exist in NEST v2.16.0: \texttt{lin\_rate},
+\texttt{sigmoid\_rate},\\
+\texttt{sigmoid\_rate\_gg\_1998}, \texttt{tanh\_rate}, and
+\texttt{threshold\_lin\_rate}. Documentation of each gain function is
+given in the header file with the corresponding name within
+\texttt{../nest/nest-simulator-master/models}. We added
+\texttt{sigmoid\_rate\_gg\_1998} specifically for the reimplementation
+of the saccade generator presented here.
+
+Gain functions are combined with a neuron class to obtain a specific
+neuron model. For instance, the combination of the \texttt{lin\_rate}
+gain function with the \texttt{rate\_neuron\_ipn} class gives the
+\texttt{lin\_rate\_ipn} neuron model. The parameters of a model are a
+combination of all parameters of the neuron base class and all
+parameters of the gain function. Within Python the command
+\texttt{nest.Models()} lists all existing model types including all
+implemented combinations of gain function and neuron class.
+
 {\sffamily \small

### 2018-02-21 nest/nest-simulator-master/models/hh_psc_alpha.h
added nest v2.16.0-beta to submission

@@ -1,0 +1,414 @@
+/*
+ *  hh_psc_alpha.h
+ *
+ *  This file is part of NEST.
+ *
+ *  Copyright (C) 2004 The NEST Initiative
+ *
+ *  NEST is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License as published by
+ *  the Free Software Foundation, either version 2 of the License, or
+ *  (at your option) any later version.
+ *
+ *  NEST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with NEST.  If not, see <http://www.gnu.org/licenses/>.
+ *
+ */
+
+#ifndef HH_PSC_ALPHA_H
+#define HH_PSC_ALPHA_H
+
+// Generated includes:
+#include "config.h"
+
+#ifdef HAVE_GSL
+
+// External includes:
+#include <gsl/gsl_errno.h>
+#include <gsl/gsl_matrix.h>
+#include <gsl/gsl_odeiv.h>
+#include <gsl/gsl_sf_exp.h>
+
+// Includes from nestkernel:
+#include "archiving_node.h"
+#include "connection.h"
+#include "event.h"
+#include "nest_types.h"
+#include "recordables_map.h"
+#include "ring_buffer.h"
+#include "universal_data_logger.h"
+
+namespace nest
+{
+/**
+ * Function computing right-hand side of ODE for GSL solver.
+ * @note Must be declared here so we can befriend it in class.
+ * @note Must have C-linkage for passing to GSL. Internally, it is
+ *       a first-class C++ function, but cannot be a member function
+ *       because of the C-linkage.
+ * @note No point in declaring it inline, since it is called
+ *       through a function pointer.
+ * @param void* Pointer to model neuron instance.
+ */
+extern "C" int hh_psc_alpha_dynamics( double, const double*, double*, void* );
+
+/* BeginDocumentation
+Name: hh_psc_alpha - Hodgkin Huxley neuron model.
+
+Description:
+
+ hh_psc_alpha is an implementation of a spiking neuron using the Hodkin-Huxley
+ formalism.
+
+ (1) Post-synaptic currents
+ Incoming spike events induce a post-synaptic change of current modelled
+ by an alpha function. The alpha function is normalised such that an event of
[…]

### 2018-02-21 nest/nest-simulator-master/models/iaf_tum_2000.h

@@ -1,0 +1,403 @@
+/*
+ *  iaf_tum_2000.h
+ *
+ *  This file is part of NEST.
+ *
+ *  Copyright (C) 2004 The NEST Initiative
+ *
+ *  NEST is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License as published by
+ *  the Free Software Foundation, either version 2 of the License, or
+ *  (at your option) any later version.
+ *
+ *  NEST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with NEST.  If not, see <http://www.gnu.org/licenses/>.
+ *
+ */
+
+
+#ifndef IAF_TUM_2000_H
+#define IAF_TUM_2000_H
+
+// Includes from nestkernel:
+#include "archiving_node.h"
+#include "connection.h"
+#include "event.h"
+#include "nest_types.h"
+#include "ring_buffer.h"
+#include "universal_data_logger.h"
+
+namespace nest
+{
+/* BeginDocumentation
+   Name: iaf_tum_2000 - Leaky integrate-and-fire neuron model with exponential
+                        PSCs.
+
+   Description:
+
+   iaf_tum_2000 is an implementation of a leaky integrate-and-fire model
+   with exponential shaped postsynaptic currents (PSCs) according to [1].
+   The postsynaptic currents have an infinitely short rise time.
+   In particular, this model allows setting an absolute and relative
+   refractory time separately, as required by [1].
+
+   The threshold crossing is followed by an absolute refractory period
+   (t_ref_abs) during which the membrane potential is clamped to the resting
+   potential. During the total refractory period (t_ref_tot), the membrane
+   potential evolves, but the neuron will not emit a spike, even if the
+   membrane potential reaches threshold. The total refractory time must be
+   larger or equal to the absolute refractory time. If equal, the
+   refractoriness of the model if equivalent to the other models of NEST.
+
+   The linear subthreshold dynamics is integrated by the Exact
+   Integration scheme [2]. The neuron dynamics is solved on the time
+   grid given by the computation step size. Incoming as well as emitted
[…]

### 2018-02-21 nest/nest-simulator-master/pynest/nest/tests/test_events.py

@@ -1,0 +1,96 @@
+# -*- coding: utf-8 -*-
+#
+# test_events.py
+#
+# This file is part of NEST.
+#
+# Copyright (C) 2004 The NEST Initiative
+#
+# NEST is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 2 of the License, or
+# (at your option) any later version.
+#
+# NEST is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with NEST.  If not, see <http://www.gnu.org/licenses/>.
+
+"""
+Test of events
+"""
+
+import unittest
+import nest
+
+
+@nest.check_stack
+class EventsTestCase(unittest.TestCase):
+    """Tests of the Connect API"""
+
+    def test_Events_1(self):
+        """Recorder Events"""
+
+        nest.ResetKernel()
+
+        sd = nest.Create('spike_detector', 1, {'withtime': True})
+        d = nest.GetStatus(sd, 'events')[0]
+
+        senders = d['senders']
+        times = d['times']
+
+        vm = nest.Create('voltmeter', 1, {'withtime': True})
+        d = nest.GetStatus(vm, 'events')[0]
+
+        senders = d['V_m']
+        times = d['times']
+
+    def test_EventsVoltage(self):
+        """Voltage Events"""
+
+        nest.ResetKernel()
+
+        nest.sr('20 setverbosity')
+        n = nest.Create('iaf_psc_alpha')
+        vm = nest.Create('voltmeter', 1, {'withtime': True, 'interval': 1.})
+
+        nest.Connect(vm, n)
+        nest.SetKernelStatus({'print_time': False})
+        nest.Simulate(10)
+
+        d = nest.GetStatus(vm, 'events')[0]
+
+        self.assertEqual(len(d['V_m']), 9)
+
+    def test_EventsSpikes(self):
+        """Spike Events"""
+
+        nest.ResetKernel()
+
+        nest.sr('20 setverbosity')
+
+        n = nest.Create('iaf_psc_alpha', 1, {'I_e': 1000.})
+        sd = nest.Create('spike_detector', 1, {'withtime': True})
+
[…]

### 2018-02-21 nest/nest-simulator-master/pynest/nest/tests/test_threads.py

@@ -1,0 +1,129 @@
+# -*- coding: utf-8 -*-
+#
+# test_threads.py
+#
+# This file is part of NEST.
+#
+# Copyright (C) 2004 The NEST Initiative
+#
+# NEST is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 2 of the License, or
+# (at your option) any later version.
+#
+# NEST is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with NEST.  If not, see <http://www.gnu.org/licenses/>.
+
+"""
+UnitTests for multithreaded pynest
+"""
+
+import unittest
+import nest
+
+
+@nest.check_stack
+class ThreadTestCase(unittest.TestCase):
+    """Tests for multi-threading"""
+
+    def nest_multithreaded(self):
+        """Return True, if we have a thread-enabled NEST, False otherwise"""
+
+        nest.sr("statusdict/threading :: (no) eq not")
+        return nest.spp()
+
+    def test_Threads(self):
+        """Multiple threads"""
+
+        if not self.nest_multithreaded():
+            self.skipTest("NEST was compiled without multi-threading")
+
+        nest.ResetKernel()
+        self.assertEqual(nest.GetKernelStatus()['local_num_threads'], 1)
+
+        nest.SetKernelStatus({'local_num_threads': 8})
+        n = nest.Create('iaf_psc_alpha', 8)
+        st = list(nest.GetStatus(n, 'vp'))
+        st.sort()
+        self.assertEqual(st, [0, 1, 2, 3, 4, 5, 6, 7])
+
+    def test_ThreadsGetConnections(self):
+        """GetConnections with threads"""
+
+        if not self.nest_multithreaded():
+            self.skipTest("NEST was compiled without multi-threading")
+
+        nest.ResetKernel()
+        nest.SetKernelStatus({'local_num_threads': 8})
+        pre = nest.Create("iaf_psc_alpha")
+        post = nest.Create("iaf_psc_alpha", 6)
+
+        nest.Connect(pre, post)
+
+        conn = nest.GetConnections(pre)
+        # Because of threading, targets may be in a different order than
[…]

### 2018-02-21 nest/nest-simulator-master/sli/slidict.cc

@@ -1,0 +1,1091 @@
+/*
+ *  slidict.cc
+ *
+ *  This file is part of NEST.
+ *
+ *  Copyright (C) 2004 The NEST Initiative
+ *
+ *  NEST is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU General Public License as published by
+ *  the Free Software Foundation, either version 2 of the License, or
+ *  (at your option) any later version.
+ *
+ *  NEST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU General Public License for more details.
+ *
+ *  You should have received a copy of the GNU General Public License
+ *  along with NEST.  If not, see <http://www.gnu.org/licenses/>.
+ *
+ */
+
+/*
+    SLI Dictionary access
+*/
+
+#include "slidict.h"
+
+// C++ includes:
+#include <typeinfo>
+
+// Includes from sli:
+#include "arraydatum.h"
+#include "booldatum.h"
+#include "dictdatum.h"
+#include "dictstack.h"
+#include "integerdatum.h"
+#include "iostreamdatum.h"
+#include "namedatum.h"
+#include "tokenutils.h"
+
+/*
+BeginDocumentation
+
+   Name: dict - Create new, empty dictionary
+
+   Synopsis: dict -> <<>>
+
+   Description:
+    dict creates a new dictionary with no entries. This command is
+    equivalent to
+    <<>>
+
+   Diagnostics: No errors are raised.
+
+   Author: Marc-Oliver Gewaltig
+
+   SeeAlso: <<>>, clonedict, cleardict
+
+*/
+
+void
+DictFunction::execute( SLIInterpreter* i ) const
+{
+  // create a new dictionary
+
+  i->EStack.pop(); // never forget me
+  i->OStack.push( DictionaryDatum( new Dictionary ) );
+}
+
+/*
+BeginDocumentation
+
+   Name: put_d - Add an entry to a dictionary
+
+   Synopsis: <<dict>> /key val -> <<dict>>
+
+   Description:
+     put_d adds the pair /key value to the supplied dictionary.
+     If /key was already defined in that dictionary, its old
+     value is lost.
+
+     In contrast to PostScript dictionaries, only literal names can
+     be used as keys in a dictionary.
+
[…]

### 2018-03-02 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Uploaded version of the manuscript

@@
 \begin{itemize}
-    \item[→] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C.
+    \item[â†’] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C.
 Potjans,  M.
@@
 
-Additionally to these inputs, in Fig.~\ref{fig:fig5}C we simulate the network in several trials where the number of external inputs change.
+Additionally to these inputs, in Fig.~\ref{fig:fig5}\textbf{C} we simulate the network in several trials where the number of external inputs change.
 In every trial, the external inputs to each of the excitatory layers is a number randomly drawn between the layer specific and the layer independent inputs reported in Table \ref{table:externalinput}.
@@
 To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.
-The null hypothesis that two sample sets were drawn from the same distribution is accepted, for an error $\alpha = 0.01$, if the KS-statistic $\leq 1.63$.
+
 \section{Results}\label{results}
@@
 In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz).
-(A) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells).
+(\textbf{A}) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells).
 Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer).
-(B–D) Statistics based on samples of 1000 neurons per layer recorded for 60 s.
-(B) Single-cell firing rate boxplot.
+(\textbf{B--€“D}) Statistics based on samples of 1000 neurons per layer recorded for 60 s.
+(\textbf{B}) Single-cell firing rate boxplot.
 Triangles indicate population mean firing rates, crosses show outliers.
-(C) Irregularity of single-cell firing rates.
-(D) Synchrony of spiking activity.}\label{fig:fig2}
+(\textbf{C}) Irregularity of single-cell firing rates.
+(\textbf{D}) Synchrony of spiking activity.}\label{fig:fig2}
 \end{figure}
@@
 The comparisons were made in terms of the cumulative distributions of the CVs of ISIs and the firing rates of the eight cell populations in the network (plots not shown here but as in Figs.~\ref{fig:fig3} and~\ref{fig:fig4} below).
-The p-value (two-sample Kolmogorov-Smirnov test) was higher than 0.6 for all comparisons meaning that no significant difference was found between NEST and Brian 2 simulations.
+The two-sample Kolmogorov-Smirnov test pointed that no difference was found between NEST and Brian 2 simulations.
 
@@
 
-When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences for an error $\alpha = 0.01$.
-The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
+When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences (for an error $\alpha = 0.01$) between the firing rate and CV distributions.
+However, when comparing the mean firing rates and their standard deviations, in general, slightly differences were found being major for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
 
@@
 \rot{\ref{fig:fig2}}.
-(A1--A2) DC input.
-(B1--B2) Layer independent input.
-(C1--C4) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods).
+(\textbf{A1--A2}) DC input.
+(\textbf{B1--B2}) Layer independent input.
+(\textbf{C1--C4}) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods).
 }\label{fig:fig5}
@@
 In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs.
-These are shown in Fig.~\ref{fig:fig5}A, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}B, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs.
+These are shown in Fig.~\ref{fig:fig5}\textbf{A}, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}\textbf{B}, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs.
 In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
-In Fig.~\ref{fig:fig5}C we present the population firing rates for 100 trials with the rule explained in the methods section.
+In Fig.~\ref{fig:fig5}\textbf{C} we present the population firing rates for 100 trials with the rule explained in the methods section.
 This latter experiment presented an excellent agreement with the histograms observed in Fig.
@@
 8 of the original article).
-(A) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$.
-(B) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$.
+(\textbf{A}) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$.
+(\textbf{B}) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$.
 Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate.
-(C) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (A)).}\label{fig:fig6}
+(\textbf{C}) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (\textbf{A})).}\label{fig:fig6}
 \end{figure}
@@
 As described in the results section of the original article, when thalamic input is delivered to the network the output is rather stereotypical.
-This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian2 to transient thalamic input.
+This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian 2 to transient thalamic input.
 By comparing histograms in Fig.~\ref{fig:fig7}\textbf{B} with the ones in Fig.
-10 \textbf{B} of the original article one can see the similarities.
+10\textbf{B} we think that our model replicates the original article.
 
@@
  \includegraphics[scale=0.01]{fig7.pdf}
- \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.10 from the original article).
+ \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.~10 from the original article).
 Note that the time axis is centralized at the beginning of the thalamic input, which lasts for 10 ms.
-(A) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section.
-(B) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
+(\textbf{A}) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section.
+(\textbf{B}) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
 \end{figure}
@@
 Sacha van Albada for valuable comments on the manuscript.
-This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC.
+This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to ROS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC.
 ACR is partially supported by the CNPq fellowship grant 306251/2014-0.

### 2018-03-13 article/Caze_2018.tex
Update manuscript and code for the new revision
Co-authored-by: Romain Caze <romain.caze@gmail.com>

@@
 Dendritic non-linearities increase neurons' computation capacity, turning them into complex computing units \cite{Poirazi2003a}.
-However network studies are usually based on point neuron models that do not incorporate dendrites and their non-linearities.
+However, network studies are usually based on point neuron models that do not incorporate dendrites and their non-linearities.
 In contrast, the study replicated here \cite{Memmesheimer2012} uses a simple point-neuron model that contains an effective description of dendrites by a non-linear summation of its excitatory synaptic input.
@@
 The simulations in the original paper were done in phase representation, but our implementation uses the more standard representation in terms of the neurons' membrane potential.
-The membrane potential dynamic $V_i$ (measured relative to the resting potential) of neuron $i$ is described by (a slightly rewritten version equivalent to equation (1) of \cite{Memmesheimer2012}):
+The membrane potential $V_i$ of neuron $i$, measured relative to the resting potential, is described by:
 
 \begin{equation}
-\tau_m \frac{dV_i(t)}{dt} = - V_i(t) + V_0 + \sigma \big( \sum_{j\in E_i} w_\text{ex} \epsilon_{j}(t-\tau) \big) - \sum_{j \in I_i} w_\text{in}\epsilon_{j}(t-\tau),
+\label{eq:neuron_model}
+\tau_m \frac{dV_i(t)}{dt} = - V_i(t) + V_0 + \sigma \left( \sum_{j\in E_i} w_\text{ex} \epsilon_{j}(t-\tau) \right) - \sum_{j \in I_i} w_\text{in}\epsilon_{j}(t-\tau),
 \end{equation}
@@
 $\SI{0}{\milli\volt}$.
-The function $\epsilon_{j}$ is 1 whenever neuron $j$ spikes, and 0 otherwise
+The function $\epsilon_{j}$ is 1 whenever neuron $j$ spikes, and 0 otherwise.
+Equation~\ref{eq:neuron_model} is an alternative, mathematically equivalent formulation of equation (1) of \cite{Memmesheimer2012}.
 
 The originality of this model lays in a filtering of the excitatory inputs by the function $\sigma$ to model dendritic non-linearities.
-The study considers two functions $\sigma$ (see Fig.~1 of the original paper and insets in Fig.~\ref{fig:markov}).
+The study considers two functions $\sigma$ (see Fig.~1 of the original paper and insets in Fig.~\ref{fig:markov_corrected}).
 Linear coupling, defined by $\sigma(x)=x$, and non-linear coupling with:
-\[
+\begin{equation}
     \sigma(x)= 
 \begin{cases}
-    x,              & \text{if } x\leq V_a\\
-    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{if } V_a<x \leq V_b\\
-    V_c,            & \text{otherwise}
+    x              & \text{, if } x\leq V_a\\
+    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{, if } V_a<x \leq V_b\\
+    V_c            & \text{, otherwise.}
 \end{cases}
-\]
+\label{eq:non-linear}
+\end{equation}
 For this non-linear function, the biological interpretation is that below a first threshold $V_a$, excitatory inputs are propagated passively by the dendrite and sum linearly.
@@
 If they are even higher and cross the threshold $V_b$, they stop increasing the somatic voltage and saturate.
-Note that this non-linearity only affects spikes received synchronously as has been observed experimentally.
+Note that this non-linearity only affects spikes received synchronously as has been observed experimentally \cite{Softky1994}.
 With the introduction of this non-linearity, the modelled point neuron becomes an effective model of a neuron with one non-linear dendrite.
@@
   Firing threshold & $\Theta$ & \SI{16}{\milli\volt} &\\
+  External input & $V_0$ & \SI{17.6}{\milli\volt}&(effect on $V$)\\
   Synaptic delay & $\tau$ & \SI{5}{\milli\second} & (but see Implementation section)\\
@@
 
-For non-linearly coupled networks, we have to use a slightly more complicated approach where we sum up all the received excitatory input in a variable \verb|ve|, apply the non-linear function $\sigma$ to this input, and then reset the variable \verb|ve| back to 0 to prepare for the next time step.
+We used a the predefined \verb|clip| function to implemented non-linear coupling into a network. The function clips a value to a given range, following the syntax
+\begin{code}
+clip(value, lower_bound, upper_bound)
+\end{code}
+
+We employed this functions on the received excitatory inputs stored in a placed holder variable \verb|ve|.
+The variable corresponds to x in eq.~\ref{eq:non-linear}.
+We then reset the variable \verb|ve| back to 0 to prepare for the next time step.
 In Brian syntax (the definition of the inhibitory synapses is identical to the linearly coupled network):
+
 \begin{code}
@@
 \end{code}
-The \verb|run_regularly| operation that is run every time step to implement the non-linear $\sigma$ function refers to the predefined \verb|clip| function which clips the value to a given range, following the syntax
-\begin{code}
-clip(value, lower_bound, upper_bound)
-\end{code}
+
+The \verb|run_regularly| operation applies the non-linearity to the summed input at every time step.
+
+In addition to random initial conditions, i.e.
+random initial values of the membrane potential, the simulation includes 1--50 random spikes ``initially in transit''.
+The original article does not go into further details about their implementation, so we decided to use the following: at the beginning of the simulation, a uniformly distributed number between 1 and 50 is drawn to determine the total number of initial spikes.
+These spikes are ``in transit'', meaning that they are modelled as arising from events before the start of the simulation.
+Since the synaptic delay in the network is \SI{5}{\milli\second}, we draw the arrival times for these spikes between \SI{0}{\milli\second} and \SI{5}{\milli\second}.
+Consistent with the probability for excitatory connections in the network ($p_\text{ex}=0.5$), we take half of these spikes as arising from excitatory and half from inhibitory sources, modelling their effect as described above.
+
 To facilitate simulation, we use the \emph{joblib} library to automatically store results and to distribute computation over processor cores.
@@
 to re-use an old result obtained with an outdated version of the code) while keeping the code readable.
-To reduce the total simulation time, the grid exploration in Fig.~\ref{fig:grid}, and the analysis of group size evolution in Fig.~\ref{fig:markov} have been distributed over processor cores using joblib's \verb|Parallel| construct.
-For the simulations presented in Fig.~\ref{fig:grid}, multiple repetitions for the same combination of synaptic weights are distributed over processor cores, for the simulations presented in Fig.~\ref{fig:markov}, simulations are evaluated in parallel over group sizes.
+To reduce the total simulation time, the grid exploration in Fig.~\ref{fig:grid}, and the analysis of group size evolution in Fig.~\ref{fig:markov_corrected} have been distributed over processor cores using joblib's \verb|Parallel| construct.
+For the simulations presented in Fig.~\ref{fig:grid}, multiple repetitions for the same combination of synaptic weights are distributed over processor cores, for the simulations presented in Fig.~\ref{fig:markov_corrected}, simulations are evaluated in parallel over group sizes.
 In both cases, a number of $C-1$ parallel processes is used, where $C$ is the number of processor cores in the user's system.
@@
 
-For the grid search presented in Fig.~\ref{fig:grid}, we varied inhibitory and excitatory synaptic weights between \SI{0.16}{\milli\volt} and \SI{0.4}{\milli\volt} (given that each neuron receives on average 150 excitatory and 150 inhibitory inputs, this corresponds to a ``total weight'' between \SI{24}{\milli\volt} and \SI{60}{\milli\volt}, as used on the axes of Fig.~\ref{fig:grid}), using 100 steps for each variable (comparable to the 96 steps used in the original publication).
-Each network was simulated 10 times with different initial conditions (synaptic connections and initial membrane potential values), whereas the original publication used 20 simulations per weight combination and additionally added ``1--50 random spikes initially in transit''.
+For the grid search presented in Fig.~\ref{fig:grid}, we varied inhibitory and excitatory synaptic weights between \SI{0.16}{\milli\volt} and \SI{0.4}{\milli\volt}.
+Given that each neuron receives on average 150 excitatory and 150 inhibitory inputs, this corresponds to a ``total weight'' between \SI{24}{\milli\volt} and \SI{60}{\milli\volt}, as used on the axes of Fig.~\ref{fig:grid}.
+We used a resolution of 100 steps for each variable, comparable to the 96 steps used in the original publication.
+Each network was simulated 10 times with different initial conditions (synaptic connections and initial membrane potential values), whereas the original publication used 20 simulations per weight combination.
 
-To numerically derive the evolution of synchronous pulse sizes presented in Fig.~\ref{fig:markov} (for the semi-analytical approach, see the next section), we performed simulations of \SI{56}{\milli\second} each.
-During each simulation, we stimulated the network at 50ms with a synchronous pulse of activity, consisting of between 1 and 176 neurons (varied in steps of 5 -- the original study used group sizes between 1 and 181 varied in steps of 6).
+To numerically derive the evolution of synchronous pulse sizes presented in Fig.~\ref{fig:markov_corrected} (for the semi-analytical approach, see the next section), we performed simulations of \SI{56}{\milli\second} each.
+During each simulation, we stimulated the network at \SI{50}{\milli\second} with a synchronous pulse of activity, consisting of between 1 and 181 neurons (varied in steps of 5 -- the original study used group sizes in the same range varied in steps of 6).
 We then observed the number of active neurons at \SI{55}{\milli\second}, i.e.
 at one $\tau$ after the stimulation.
-For each stimulation group size, we ran the simulations 50 times with different initial conditions, consistent with the original study.
+For each stimulation group size, we ran the simulations 50~times with different initial conditions, consistent with the original study.
+To achieve a better quantitative match with the original study, we applied a correction for the stimulation group size $g'_0$ and the observed group size $g'_1$ in the next iteration (for the uncorrected results, see Fig.~\ref{fig:markov}).
+We refer to the corrected values as $\hat{g}_0$ and $\hat{g}_1$, respectively.
+These corrections are necessary because of the background activity and our clock-based simulation method and are motivated as follows.
+The activity of the full network in the absence of stimulation is about \SI{55}{\kilo\hertz} (see Fig.~\ref{fig:raster}).
+Given that our simulation uses a time step of \SI{0.1}{\milli\second}, this means that at every time step about 5.5 neurons on average are synchronously active.
+When we stimulate a network with a pulse of $g'_0$ synchronous neurons, the actual number of synchronously active neurons is therefore higher than this stimulation.
+Similarly, we have to correct the size of the observed group in the next iteration.
+In contrast to the simulation strategy of the original publication that could make use of an ``infinite precision'' to detect synchronous spikes triggered by the stimulation, our simulation will also detect all the synchronous spikes resulting from the background activity.
+In the corrected values shown in Fig.~\ref{fig:markov_corrected} we therefore used the actually observed synchronous spikes $\hat{g}'_0 \geq g'_0$ at the time of the stimulation, and the number of added spikes due to the stimulation $\hat{g}'_1 = g'_{1 \text{stim}} - g'_{1 \text{no stim}}$, where $g'_{1 \text{stim}}$ is the number of synchronous spikes we observe in a network with stimulation, and $g'_{1 \text{no stim}}$ the number of observed spikes in a network without stimulation.
+Both networks are simulated with the same random seed and therefore have the same network connectivity and the same initial conditions; the only difference lies in the presence or absence of external synchronous stimulation.
+Note that the correction for $\hat{g}'_1$ slightly ``overcorrects'', as it discards all spikes that are resulting from background activity, even those that also would have been triggered by the stimulation.
 
 \subsection*{Semi-analytical and analytical methods}
-We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}).
-To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows.
-First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.
-We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential.
-From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
+We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}), since the original publication did not give a detailed description of this approach.
+However, we replicate the semi-analytical work that was described in more detail.
+
+To calculate the semi-analytical solutions for Fig.~\ref{fig:markov_corrected}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows.
+First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second}.
+Each of these simulations was randomly connected and had random initial conditions, but received no additional synchronous stimulation.
+The original study additionally simulated each of the 100 networks 10 times, with the same synaptic connectivity but differing in the initial conditions.
+We then binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$ to obtain an estimate of $P(V)$, the distribution of the membrane potential.
+Note that the original study did not further specify how $P(V)$ was estimated from the simulation results.
+
+We then calculate the expected synchronous group size after one iteration as in the original publication.
+To make this calculation clearer, we now describe its derivation in detail.
+From the distribution $P(V)$ we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
 \begin{equation}
@@
 $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$.
-Let us assume that in the last iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e.
+Let us assume that in the previous iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e.
 that $N-g_i$ neurons are available to spike in this iteration.
@@
 \end{equation}
-In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, i.e.:
+In Fig.~\ref{fig:markov_corrected}, we plot the expected value of spiking neurons during the next iteration, given a certain initial synchronous pulse $g_i'$:
 \begin{equation}
@@
 Combining individual results in this way does not necessarily lead to easily interpretable colours, but using a different scheme, e.g.
-colouring according to the class that occurred the most often, leads to visually similar results (not shown).
+colouring according to the class that occurred the most often, leads to visually similar results (Fig.~\ref{fig:grid_max}).
 
-Our replication does not reproduce all details of the original plots (in particular at the top of the plots), but looks very similar on a coarse scale.
-Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not.
-The minor differences in our replication could be due to either our use of a clock-based simulation method, or to the slightly different initial conditions (no spikes ``in transit'').
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100_mean.pdf}
+\end{subfigure}%
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75_mean.pdf}
+\end{subfigure}
+\caption{\label{fig:grid}
+\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B, C) coupled networks.} For each combination, we initiated synchronous activity with 100 (A, B) or 75 (C) neurons and assessed the stability of the temporal evolution.
+Blue colouring indicates stable propagation of synchrony, red indicates unstable background activity before onset of propagation, yellow indicates unstable background activity after onset of propagation, and green colouring indicates the absence of propagation (see Methods for details).\\
+Replication of Fig.~3 in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g003})}
+\end{figure}
+
 \begin{table}
@@
 
-\begin{figure}
-\centering
-\begin{subfigure}[t]{0.5\textwidth}
-\caption{}\vspace{-1ex}
-\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_mean.pdf}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\caption{}\vspace{-1ex}
-\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_mean.pdf}
-\end{subfigure}
-\caption{\label{fig:grid}
-\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B) coupled networks.} For each combination, we initiated synchronous activity with 100 neurons and assessed the stability of the temporal evolution.
-Blue colouring indicates stable propagation of synchrony, red indicates unstable background activity before onset of propagation, yellow indicates unstable background activity after onset of propagation, and green colouring indicates the absence of propagation (see Methods for details).\\
-Replication of Fig.~3a and b in \cite{Memmesheimer2012} (\url{https://doi.org/10.1371/journal.pcbi.1002384.g003})}
-\end{figure}
+Our replication does not reproduce all the quantitative details of the original plots.
+The parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article.
+For linear networks, we do observe almost no simulations classified in this category, while the original article has a thin line of such classifications between the stable regime without propagation (green) and the regime where simulations are unstable even without stimulation (red).
+Similarly, for non-linearly coupled networks the parameter exploration in the original article shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} where the network is unstable only after the synchronous stimulation (yellow).
+In contrast, our simulations show a much smaller such zone when the inhibitory connection strength is small (Fig.~\ref{fig:grid}B).
+It seems that these differences mostly stem from the fact that the transition to the regime where simulations are unstable without stimulation (red) happens for lower excitatory and higher inhibitory connection strength in our simulations than in the original article, i.e.
+that the red zone is shifted towards the bottom left.
+This is particularly evident for strong excitatory and inhibitory connections in non-linearly coupled networks (bottom right in Fig.~\ref{fig:grid}B and C) which are unstable in our simulations, but stable or only unstable after stimulation in the original publication.
+Simulating the parameter combination corresponding to the point in the bottom right corner of the plot at different time resolutions (Fig.~\ref{fig:raster_strong}), we find that this difference does indeed seem to be triggered by limited time resolution in the model.
+The forced synchronisation of all spikes to a grid of \SI{0.1}{\milli\second}, together with the homogeneity of the synaptic delays, leads to strong oscillations in our simulation (Fig.~\ref{fig:raster_strong}A), which is consequently classified as unstable.
+When we use a simulation time step of \SI{0.01}{\milli\second}, these osciallations disappear and the network activity therefore gets classified as ``stable without persistent propagation''.
+Finally, non-linearly coupled networks show a discontinuity for weak inhibitory connection strengths in the original publication.
+For increasing excitatory connection strengths, network behaviour goes from stable (blue) to unstable after stimulation (yellow) to unstable before stimulation (red), but then get classified again as stable (blue) in a small range of parameters before finally transitioning to unstable activity (yellow and red).
+We do not observe this in our simulations, but simulations with a finer time resolution (Fig.~\ref{fig:grid_detail}C) indicate that this difference could be caused by the limited time resolution of our simulations as well.
+
+Despite these differences, the qualitative network behaviour looks very similar on a coarse scale.
+Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not.
+The differences in our replication are most likely due to our use of a clock-based simulation method with a limited time resolution (see Fig.~\ref{fig:grid_detail} and Fig.~\ref{fig:raster_strong}).
+In principle, another reason could be implementation details of the initial ``spikes in transit'' that were not fully specified in the original article, but this seems unlikely since completely removing them did not change the results in a noticeable way (not shown).
 
@@
 \caption{}\vspace{-1ex}
-\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
+\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us_corrected.pdf}
 \end{subfigure}%
@@
 \caption{}\vspace{-1ex}
-\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
+\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us_corrected.pdf}
 \end{subfigure}
-\caption{\label{fig:markov}
+\caption{\label{fig:markov_corrected}
 \textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks.} To numerically estimate the probability distribution, we show occurrences of pulse-sizes $g'_1$ in response to a pulse of size $g'_0$ by grey lines; green squares represent the corresponding mean response group sizes.\\
@@
 
-Finally, we looked at the evolution of synchronous group sizes for one combination of synaptic weights ($w_\text{ex}=w_\text{in}=\SI{0.2}{\milli\volt}$) in detail (Fig.~\ref{fig:markov}).
-The results confirms our previous conclusion that we can faithfully replicate the network behaviour.
-They match both qualitatively and quantitatively, the group size consistently declines with each iteration in linearly coupled networks (Fig.~\ref{fig:markov}A), whereas it increases for intermediate group sizes and declines otherwise in non-linearly coupled networks (Fig.~\ref{fig:markov}B).
-While the semi-analytic solution (red dotted line) shows a good match to the numerical results (grey lines, means as green dots), it shows some quantitative differences to the semi-analytic solution presented in Fig.~4 of the original article.
-In particular, our semi-analytical solution seems to mostly slightly underestimate the size of the group in the next iteration while the semi-analytical solution presented in the original article seems to slightly overestimate it.
+Finally, we looked at the evolution of synchronous group sizes for one combination of synaptic weights ($w_\text{ex}=w_\text{in}=\SI{0.2}{\milli\volt}$) in detail (Fig.~\ref{fig:markov_corrected}).
+The results confirms our previous conclusion that we can faithfully replicate the network behaviour which shows identical qualitative behaviour; the group size consistently declines with each iteration in linearly coupled networks (Fig.~\ref{fig:markov_corrected}A), whereas it increases for intermediate group sizes and declines otherwise in non-linearly coupled networks (Fig.~\ref{fig:markov_corrected}B).
+Quantitatively, the semi-analytical results also match well.
+Based on the plotted values in the original publication (Fig.~4 in \cite{Memmesheimer2012}), the curve based on the semi-analytical solution crosses the identity line at about 85 ($G_1$) and 135 ($G_2$) for non-linearly coupled networks, peaking in between for $g'_0\approx125$ and $g'_1\approx139$.
+In our simulations, it crosses the line at about 86 and 134, and peaks at $g'_0\approx126$ and $g'_1\approx137$.
+As in the original article, our numerical results show slightly lower values for the group size in the next iteration, with the biggest difference for values around the peak of the function for non-linearly coupled networks (Fig.~\ref{fig:markov_corrected}B).
+However, we see a quantitative difference for low initial group sizes.
+In these simulations, the original publication shows a close match between the numerical results and the semi-analytical solution.
+In our simulation, the semi-analytical solution overestimates the numerical one.
+Be reminded that to obtain this good quantitative match, we had to apply a correction to the numerical results, as explained in the implementation section (for the uncorrected values, see Fig.~\ref{fig:markov}).
+
+Despite the small quantitative differences, our work strengthens the original result by demonstrating a qualitative difference between linearly coupled and non-linearly coupled networks.
 
@@
 
-We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences.
-These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision.
-We tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained qualitatively the same results (not shown).
-Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.
+We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some quantitative differences.
+The grid search over parameters showed a higher prevalence of networks with unstable activity before the onset of stimulation compared to the original study.
+In the analysis of the evolution of synchronous pulses of different sizes, our results need a correction for the use of a clock-based simulation method to show a good quantative match with the original results.
+However, a marked difference remains for small initial group sizes.
+The differences we see are most likely due to our use of a clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision.
+However, the main qualitative finding---non-linearly coupled networks have a large parameter range supporting persistent propagation of synchronous activity whereas linearly coupled networks do not---remains if we use finer (\SI{0.05}{\milli\second}) or coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions (data not shown).
 
@@
 %}
+\clearpage
+\appendix
+\section*{Supplementary Figures}
+\begin{figure}[h]
+\centering
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100.pdf}
+\end{subfigure}%
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75.pdf}
+\end{subfigure}
+\caption{\label{fig:grid_max}
+\textbf{Most-often observed qualitative network behaviour as a function of excitatory and inhibitory connection
+strength} All conventions as in Fig.~\ref{fig:grid}, but points are coloured according the observed qualitative behaviour that occurred the most often among the 10 independent simulations for given connection strengths.}
+\end{figure}
 
+\begin{figure}[h]
+\centering
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean_detail.pdf}
+\end{subfigure}%
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_50us_gs100_mean_detail.pdf}
+\end{subfigure}
+\begin{subfigure}[t]{0.33\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/grid_nl_dt_10us_gs100_mean_detail.pdf}
+\end{subfigure}
+\caption{\label{fig:grid_detail}
+\textbf{Zoom on qualitative network behaviour with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}.
+Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}, and only for non-linearly coupled networks with an synchronous stimulation of 100 neurons.
+The total inhibitory connection strength was varied between \SI{24}{\milli\volt} and \SI{34.55}{\milli\volt}, the total excitatory connection strength was varied between \SI{34.91}{\milli\volt} and \SI{45.45}{\milli\volt}, i.e.
+the parameter range corresponds to a square region at the top of Fig.~\ref{fig:grid}B.
+Simulations where performed with simulation time steps of \SI{0.1}{\milli\second} (A; also the value used for all other simulations shown), \SI{0.05}{\milli\second} (B), and \SI{0.01}{\milli\second} (C).}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.5\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_100us.pdf}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_10us.pdf}
+\end{subfigure}
+\caption{\label{fig:raster_strong}
+\textbf{Propagation of synchronous spiking in strongly connected networks when simulated with a time step of \SI{0.1}{\milli\second} (A), or \SI{0.01}{\milli\second} (B)}.
+All conventions as in Fig.~\ref{fig:raster}, but for strong recurrent connections ($w_\text{ex} = w_\text{in} = \SI{0.4}{\milli\volt}$).
+Both simulated networks use non-linearly coupling, the network shown in panel (B) is simulated with a 10-times finer temporal resolution compared to the other results shown.}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}[t]{0.5\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\caption{}\vspace{-1ex}
+\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
+\end{subfigure}
+\caption{\label{fig:markov}
+\textbf{Evolution of synchronous pulses without correction in linearly (A) and non-linearly (B) coupled networks.} All conventions as in Fig.~\ref{fig:markov_corrected}.
+The numerical results shown here (gray lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
+\end{figure}
 \end{document}

### 2018-03-13 code/lib.py

@@ -1,7 +1,6 @@
 from __future__ import print_function, division
 
 from collections import Counter
-from math import factorial
 
 import numpy as np
 from scipy.special import comb
@@ -13,35 +12,68 @@ from brian2 import mV, ms
 mem = Memory(cachedir='./__joblib_cache__', verbose=0, compress=True)
 
 
-def group_size_evolutions(group_size, n_rep=1, linear=True, dt=0.1*ms, seeds=None):
-    """Return the evolution of the group size after one cycle"""
+def group_size_evolutions(group_size, n_rep=1, linear=True, dt=0.1*ms,
+                          bias_correction=False, seeds=None):
+    """Return the evolution of the group size after one cycle. Returns a
+    tuple (g0, g1), containing the size of the group size at stimulation time
+    (g0), and the group size in the next iteration (g1). If ``bias_correction``
+    is set to True, both of these values are corrected for the bias introduced
+    by the finite time step size."""
     group_ev = []
     for rep in range(n_rep):
         if seeds is None:
             seed = None
         else:
             seed = seeds[rep]
-        results = run_with_cache(56 * ms, group_size=group_size, stim_time=50 * ms,
-                                 linear=linear, repetition=rep, dt=dt, seed=seed)
-        group_ev.append(np.sum(np.abs(results['t'] - 55 * ms) < dt / 2))
+        if bias_correction:
+            # Run without stimulation, but otherwise identical (same seed)
+            results_no_stim = run_with_cache(56 * ms, group_size=0,
+                                             stim_time=None, linear=linear,
+                                             repetition=rep, dt=dt, seed=seed)
+        results = run_with_cache(56 * ms, group_size=group_size,
+                                 stim_time=50 * ms, linear=linear,
+                                 repetition=rep, dt=dt, seed=seed)
+        next_iteration = np.sum(np.abs(results['t'] - 55 * ms) < dt / 2)
+        if bias_correction:
+            # Use the actual number of synchronous spike at the time of the
+            # stimulation -- due to the background activity, this is usually
+            # higher than the size of the stimulation group
+            g0 = np.sum(np.abs(results['t'] - 50 * ms) < dt / 2)
[…]

### 2018-03-13 code/main.py

@@ -19,7 +19,8 @@ br2.prefs.codegen.target = 'cython'
 
 FIG_FOLDER = "../article/Figs/"
 
-main_seeds = [3982293089, 1861487226, 2570878100, 3472600360, 2780156760]
+main_seeds = [29732904, 94614906, 1861487226, 2570878100, 3472600360, 2780156760,
+              1966509933, 1687607157, 238607587]
 
 if not os.path.exists(FIG_FOLDER):
     os.makedirs(FIG_FOLDER)
@@ -27,33 +28,55 @@ if not os.path.exists(FIG_FOLDER):
 # Set figure options from configuration file
 plt.style.use('figures.conf')
 
+
 def generate_seeds(main_seed, n_seeds):
     np.random.seed(main_seed)
     return np.random.randint(0, 2**32-1, size=n_seeds, dtype=np.uint32)
 
 
-def fig2(quiet=False, dt=0.1*ms, seed=None):
+def fig2(linear=True, quiet=False, dt=0.1*ms, seed=None):
     """Generate the raster plot and population statistics for Figure 2
     of the original paper (https://doi.org/10.1371/journal.pcbi.1002384.g002)"""
     if not quiet:
-        print('Running simulations for Figure 2')
-    seeds = generate_seeds(seed, 2)
-    results = run_with_cache(w_ex=0.2*mV, w_in=0.2*mV, linear=True, dt=dt, seed=seeds[0])
-    results_nl = run_with_cache(w_ex=0.2*mV, w_in=0.2*mV, linear=False, dt=dt, seed=seeds[1])
+        coupling = 'linear' if linear else 'non-linear'
+        print('Running simulations for Figure 2 ({} coupling)'.format(coupling))
+    results = run_with_cache(w_ex=0.2*mV, w_in=0.2*mV, linear=linear, dt=dt,
+                             seed=seed)
 
-    figname = "spikes_l_dt_%.0fus.pdf" % (dt/us)
+    if linear:
+        figname = "spikes_l_dt_%.0fus.pdf" % (dt/us)
+    else:
+        figname = "spikes_nl_dt_%.0fus.pdf" % (dt / us)
     plot_population_activity(results, save=FIG_FOLDER + figname, dt=dt)
-    figname = "spikes_nl_dt_%.0fus.pdf" % (dt/us)
-    plot_population_activity(results_nl, save=FIG_FOLDER + figname, dt=dt)
 
 
-def fig3(linear=True, short=True, quiet=False, dt=0.1*ms, seed=None):
[…]

### 2018-03-13 code/plot.py

@@ -29,16 +29,21 @@ def plot_population_activity(results, ymax1=250, save=None, dt=0.1*ms):
     plt.figure()
     # Draw the raster plot
     ax_raster.vlines(np.arange(150, 250, 5), 0, 200, color="gray", linestyle='dashed')
-    ax_raster.plot(spikes_t/ms, spikes_i, '|', color="black")
+    ax_raster.plot(spikes_t/ms, spikes_i + 1, '|', color="black")  # Plot a 1-based neuron index
     red_spikes = ((spikes_t >= 150*ms) &
                   ((spikes_t % (5*ms) < dt / 2) |
                    (spikes_t % (5*ms) > (5*ms - dt / 2))))
 
-    ax_raster.plot(spikes_t[red_spikes]/ms, spikes_i[red_spikes], 'r|')
+    ax_raster.plot(spikes_t[red_spikes]/ms, spikes_i[red_spikes] + 1, 'r|')
 
     ax_raster.set(xlim=(110, 220), ylim=(0, 200), xlabel='Time (ms)',
-                  ylabel='Neuron')
+                  ylabel='Neuron', xticks=[120, 150, 180, 210],
+                  yticks=[1, 100, 200])
 
+    # Note that we use the simulation dt as the bin size here, so that we don't
+    # lose any time resolution when checking for spikes triggered by the
+    # synchronous stimulation. When plotting the population rate, we however
+    # bin into bins of 1ms
     bins = np.arange(0, 250 + dt/ms, dt/ms)
     binned_activity, _ = np.histogram(spikes_t / ms, bins)
     # Select the spikes that (most likely) arise from the synchronous
@@ -48,33 +53,45 @@ def plot_population_activity(results, ymax1=250, save=None, dt=0.1*ms):
     delay_steps = int(round(5*ms/dt))
     group_size = binned_activity[int(round(150*ms/dt))::delay_steps]
 
-    # Draw the population rate
-    ax_binned.bar(bins[:-1], binned_activity, color="black")
-    ax_binned.set(ylim=(0, ymax1), ylabel='Rate')
+    # Draw the population rate (with 1ms bins)
+    steps_per_bin = int(round(1*ms/dt))
+    binned_activity_1ms = binned_activity.reshape(-1, steps_per_bin).sum(axis=1)
+    ax_binned.bar(bins[:-1:steps_per_bin], binned_activity_1ms,
+                  color="black")
[…]

### 2018-03-20 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Updated version of manuscript

@@
 \documentclass[10pt,a4paper,onecolumn]{article}
+% * <aroquesilva@gmail.com> 2018-03-18T20:10:02.103Z:
+%
+% ^.
 % * <aroquesilva@gmail.com> 2017-09-19T13:55:25.418Z:
@@
 \maketitle
-
 \marginpar{
@@
 \begin{itemize}
-    \item[â†’] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C.
+    \item[→] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C.
 Potjans,  M.
@@
 To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test (KS-statistics).
-To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.
+To apply this non-parametric test, cumulated histograms were constructed with bin size chosen as the square root of data size.
 
@@
 Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer).
-(\textbf{B--€“D}) Statistics based on samples of 1000 neurons per layer recorded for 60 s.
+(\textbf{B--D}) Statistics based on samples of 1000 neurons per layer recorded for 60 s.
 (\textbf{B}) Single-cell firing rate boxplot.
@@
 In order to compare the Brian 2 with the NEST implementations of the Potjans-Diesmann model, we used the available pyNEST code \cite{potjans2014pyNEST} to run NEST simulations of the model with the same parameters given in Methods.
-The comparisons were made in terms of the cumulative distributions of the CVs of ISIs and the firing rates of the eight cell populations in the network (plots not shown here but as in Figs.~\ref{fig:fig3} and~\ref{fig:fig4} below).
-The two-sample Kolmogorov-Smirnov test pointed that no difference was found between NEST and Brian 2 simulations.
+The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations in the network.
+The two-sample Kolmogorov-Smirnov test pointed that no difference was found between NEST and Brian 2 simulations, as observed by the close agreement of the cumulative histograms (orange lines with filled circles and green lines, respectively) in Figs.~\ref{fig:fig3} and \ref{fig:fig4}.
 
-\begin{figure}[!htp]
+Besides the creation of the network connections by the procedure described above, which is the one used in the original article, it is also possible to connect the neurons in the network using the alternative expression for the total number of synapses between two cell populations,
+
+\begin{figure}[!b]
  \centering
@@
 NEST code taken from \cite{potjans2014pyNEST}.}\label{fig:fig3}
-\end{figure}
+% \end{figure}
 
-Besides the creation of the network connections by the procedure described above, which is the one used in the original article, it is also possible to connect the neurons in the network using the alternative expression for the total number of synapses between two cell populations,
-
-\begin{figure}[!ht]
- \centering
+% \begin{figure}[!ht]
+%  \centering
  \includegraphics[scale=0.005]{fig4.pdf}
@@
 We found that the use of the approximate equation brings discrepancies in comparison with the original model.
-The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations for simulations of the original model based on the pyNEST code \cite{potjans2014pyNEST} and the Brian 2 code with $K$ calculated from equations (\ref{Eq:Ksyn_alt}) and (\ref{Eq:Ksyn}).
+The comparisons were made in terms of the cumulative distributions of the CVs of ISIs (Fig.~\ref{fig:fig3}) and the firing rates (Fig.~\ref{fig:fig4}) of the eight cell populations for simulations of the Brian 2 code with $K$ calculated from equations (\ref{Eq:Ksyn_alt}) and (\ref{Eq:Ksyn}).
 
-When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences (for an error $\alpha = 0.01$) between the firing rate and CV distributions.
-However, when comparing the mean firing rates and their standard deviations, in general, slightly differences were found being major for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.
+When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed with equation~(\ref{Eq:Ksyn}) (the same used in the original NEST code), the  KS-statistics for layer L5e in the CVs of ISIs (2.43), and layers L23e, L5e and L6e in the firing rate (2.96, 1.79 and 1.87, respectively), revealed significant differences between the two implemented networks ($\geq$ 1.63 for $\alpha=0.01$).
+For the remaining layers, all comparisons indicated no significant differences for the firing rate and CV distributions.
+These results are summarized in Figs.~\ref{fig:fig3} and \ref{fig:fig4}, where the cumulative histogram for simulations using equation~(\ref{Eq:Ksyn_alt}) are shown in blue dashed lines.
 
-The differences found highlight the importance of using the exact expression for the connection probability $C_{a}$ given in equation (1) of the original article (which corresponds to equation~(\ref{Eq:Ksyn}) of this text)  in simulations of the Potjans-Diesmann model.
-Observe the close agreement of the cumulative histograms when Brian 2 (K from equation~(\ref{Eq:Ksyn})) and NEST (K from equation~(\ref{Eq:Ksyn})) are compared.
+The differences found highlight the importance of using the exact expression for the total number of synapses $K$ given in equation~(\ref{Eq:Ksyn}) (which corresponds to equation (1) of the original article)  in simulations of the Potjans-Diesmann model.
 

### 2018-04-16 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
R1-GC-1,2,3-TEXT-1,2: made adjustments based on general comments, added explanation subscripts, corrected typos

@@
 Maastricht University, P.O. Box 616, 6200 MD Maastricht, The Netherlands}
-                    \affil[3]{nstitute of Neuroscience and Medicine (INM-6) and Institute for Advanced
-Simulation (IAS-6) and JARA Institute Brain Structure-Function
+                    \affil[3]{Institute of Neuroscience and Medicine (INM-6) and Institute for
+Advanced Simulation (IAS-6) and JARA Institute Brain Structure-Function
 Relationships (INM-10), Jülich Research Centre, Jülich, Germany}
-                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universit"at
+                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universität
 Wuppertal, Wuppertal, Germany}
@@
 
-In 2013 the European Comission launched the Human Brain Project
+In 2013 the European Commission launched the Human Brain Project
 (\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
-research infrastructure spurring pan-european collaboration.
+research infrastructure spurring pan-European collaboration.
 Collaboration at such a scale, involving both empirically-oriented and
@@
 subsequently be combined into a unified whole. To facilitate this
-colaboration, the HBP encourages the utilization of publicly available,
+collaboration, the HBP encourages the utilization of publicly available,
 widely used, and actively developed neural simulation frameworks such as
@@
 models \autocite{Hahne2017}. Here we make use of this new functionality
-of the NEST framework (v2.16.0) to provide an implementation of the
+of the NEST framework (2.16.0) to provide an implementation of the
 saccade generator (SG); a rate neuron model of neural circuitry in the
@@
 for legibility with both model and analysis scripts being implemented
-using Python 2.7.12.
+using Python 2.7.12. However, the code is compatible with Python 3 as
+well (tested with version 3.5.2).
 
@@
 saccades. While this constitutes the core SG model, the superior
-collicuus (SC) is relevant for some simulations. In what follows we
+colliculus (SC) is relevant for some simulations. In what follows we
 briefly describe the dynamics of neurons within the horizontal component
-(equivalent descriptions apply to the vertical component).
+(equivalent descriptions apply to the vertical component). Please note
+that subscripts \(\mathrm{l}\), \(\mathrm{r}\), \(\mathrm{d}\) and
+\(\mathrm{u}\) indicate whether a neuron controls upward, downward,
+leftward or rightward saccade direction, respectively.
 
@@
 Each long-lead burst neuron (\(L\)) receives excitatory external input
-\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\)):
+\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\))
 
@@
 \begin{array}{ll}
-\tau\frac{dL_l}{dt} &= -1.3L_l+I_l-2B_l \\\\
-\tau\frac{dL_r}{dt} &= -1.3L_r+I_r-2B_r \\
+\tau\dot L_\mathrm{l} &= -1.3L_\mathrm{l}+I_\mathrm{l}-2B_\mathrm{l} \\\\
+\tau\dot L_\mathrm{r} &= -1.3L_\mathrm{r}+I_\mathrm{r}-2B_\mathrm{r} \textrm{.} \\
 \end{array}
@@
 ipsilateral LLBN, a constant arousal signal equal to \(1\), and
-inhibitory input from the omnipause neuron (\(P\)):
+inhibitory input from the contralateral LLBN and the omnipause neuron
+(\(P\))
 
@@
 \begin{array}{ll}
-\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5L_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
-\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5L_r+1)-(E_r+1)(10L_l+20g(P))) \\
+\tau\dot E_\mathrm{l} &= -3.5E_\mathrm{l}+(2-E_\mathrm{l})(5L_\mathrm{l}+1)-(E_\mathrm{l}+1)(10L_\mathrm{r}+20g(P)) \\\\
+\tau\dot E_\mathrm{r} &= -3.5E_\mathrm{r}+(2-E_\mathrm{r})(5L_\mathrm{r}+1)-(E_\mathrm{r}+1)(10L_\mathrm{l}+20g(P)) \textrm{,} \\
 \end{array}
-\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
-given by
+\label{eq:ebn}\end{equation} where the nonlinear gain function
+\(g(\cdot)\) is given by
 
@@
 Each inhibitory burst neuron (\(B\)) receives excitatory input from the
-ipsilateral EBN:
+ipsilateral EBN
 
@@
 \begin{array}{ll}
-\tau\frac{dB_l}{dt} &= -2.4B_l+3E_l \\\\
-\tau\frac{dB_r}{dt} &= -2.4B_r+3E_r \\
+\tau\dot B_\mathrm{l} &= -2.4B_\mathrm{l}+3E_\mathrm{l} \\\\
+\tau\dot B_\mathrm{r} &= -2.4B_\mathrm{r}+3E_\mathrm{r} \textrm{.} \\
 \end{array}
@@
 
-The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs:
+The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs
 
 \begin{equation}
-\tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
+\tau\dot P = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_\mathrm{l})+g(L_\mathrm{r})+g(L_\mathrm{u})+g(L_\mathrm{d})) \textrm{,}
 \label{eq:opn}\end{equation} where \(J\) represents external electrical
-stimulation and \(g(\cdot)\) again given by equation \ref{eq:gain}.
+stimulation and \(g(\cdot)\) is again given by equation \ref{eq:gain}.
 
@@
 Each tonic neuron (\(T\)) receives excitatory input from the ipsilateral
-EBN and inhibitory input from the contralateral EBN:
+EBN and inhibitory input from the contralateral EBN
 
@@
 \begin{array}{ll}
-\tau\frac{dT_l}{dt} &= 0.1(E_l-E_r) \\\\
-\tau\frac{dT_r}{dt} &= 0.1(E_r-E_l) \textrm{.} \\
+\tau\dot T_\mathrm{l} &= 0.1(E_\mathrm{l}-E_\mathrm{r}) \\\\
+\tau\dot T_\mathrm{r} &= 0.1(E_\mathrm{r}-E_\mathrm{l}) \textrm{.} \\
 \end{array}
-\label{eq:tn}\end{equation} Horizontal eye position (\(\theta\)) depends
-on activity of the right TN; i.e. \(\theta = 260(T_r-0.5)\).
+\label{eq:tn}\end{equation} The horizontal eye position (\(\theta\))
+depends on the activity of the right TN; i.e.
+\(\theta = 260(T_\mathrm{r}-0.5)\).
 
@@
 
-The superior colliculus (\(A\)) receives electrical stimulation F
+The superior colliculus (\(A\)) receives electrical stimulation \(F\)
+and evolves according to
 
 \begin{equation}
-\tau \frac{dA}{dt} =-A+F(t) \textrm{.}
+\tau \dot A =-A+F \textrm{.}
 \label{eq:sc}\end{equation}
@@
  \begin{array}{lll}
-    0 \quad &\textrm{if} \quad x<0 \\
+    0 \quad &\textrm{if} \quad x \leq 0 \\
     x \quad &\textrm{if} \quad 0<x<1 \\
@@
 \begin{equation}
-I = Wf(A)\textrm{.}
+I = Wf(A) \textrm{.}
 \label{eq:stim}\end{equation}
 
-In implementing this model, we largely followed the descriptions
-provided in the original publication with a number of well-motivated
-exceptions. Specifically, the original model description has two
-features which cannot be straightforwardly translated to NEST. First, a
-nonlinear gain function is applied to a subset of inputs to EBNs and the
-OPN while a linear gain function is applied to their remaining inputs.
-Since NEST only applies a single gain function per neuron to each of its
-inputs, we opted for using a linear gain function for EBNs and the OPN.
-Furthermore, we passed those inputs requiring an additional nonlinear
-gain function through an auxiliary unit instantaneously applying the
-desired nonlinearity before passing the result on to EBNs and the OPN.
-Second, constant input to a neuron was not hard-coded but rather
-provided by an appropriately weighted bias node. Neither of these
-changes lead to discrepancies with original results.
+In implementing this model, we largely follow the descriptions provided
+in the original publication with a number of motivated exceptions.
+Specifically, the original model description has two features which
+cannot be straightforwardly translated to NEST. First, a nonlinear gain
+function is applied to a subset of inputs to EBNs and the OPN while a
+linear gain function is applied to their remaining inputs. Since NEST
+applies the same gain function to all inputs of a neuron, we opt for
+using a linear gain function for EBNs and the OPN, but pass those inputs
+requiring an additional nonlinear gain function through an auxiliary
+unit instantaneously applying the desired nonlinearity before passing
+the result on to EBNs and the OPN. Second, constant input to a neuron is
+not hard-coded but rather provided by an appropriately weighted bias
+node. Neither of these changes leads to discrepancies with original
+results.
 
-In all simulations we used the Exponential Euler (EE) method for
-numerical integration of all except of tonic neurons
-\autocite{Hahne2017} at a time step of \(0.05\,\mathrm{ms}\) and a time
-constant of \(50\,\mathrm{ms}\). Tonic neurons were integrated using the
-Euler-Maruyama (EM) method since they do not exhibit passive decay
-leading to division-by-zero when using the EE method. It should be noted
-that initial firing rates of neurons were not at resting equilibrium. In
-order to address this, the model was allowed to evolve for
-\(100\,\mathrm{ms}\), to relax towards equilibrium before applying any
-input. Furthermore, as in the original publication, all activations were
-bounded from below at zero. Finally, we always simulated the full model;
-i.e.~both its horizontal and vertical components even if input was
-applied only to one of the two.
+In all simulations we use the Exponential Euler (EE) method for
+numerical integration of all except tonic neurons \autocite{Hahne2017}
+at a time step of \(0.05\,\mathrm{ms}\) and a time constant of
+\(50\,\mathrm{ms}\). Tonic neurons are integrated using the forward
+Euler method since they do not exhibit passive decay leading to
+division-by-zero when using the EE method. It should be noted that the
+initial firing rates of neurons are not in equilibrium. In order to
+address this, the model evolves for \(100\,\mathrm{ms}\) to relax
+towards equilibrium before any input is applied. Furthermore, as in the
+original publication, any activity is bounded from below at zero.
+Finally, we always simulate the full model; i.e.~both its horizontal and
+vertical components even if input is applied only to one of the two.
 
@@
 
-All results from the original publication have been implemented. Our
-results accord very well with those reported by Gancarz \& Grossberg
-\autocite{Gancarz1998}. Only a single discrepancy was observed which we
-examined in depth.
+All simulations of the original publication are repeated. Our results
+accord very well with those reported by Gancarz \& Grossberg
+\autocite{Gancarz1998}. Only a single discrepancy is observed which we
+examine in depth.
 
 The first simulation showcases the evolution of activity for each neuron
-type in the horizontal SG to a constant input applied to the left LLBN.
-The original publication does not report exact activation values
+type in the horizontal SG for a constant input applied to the left LLBN.
+The original publication does not report the exact activation values
 observed for each neuron rendering a quantitative analysis of the
@@
 those shown in figure \ref{fig:fig_1} show good correspondence. In both
-implementations, the left LLBN showed a prelude of activity with left
+implementations, the left LLBN exhibits a prelude of activity with left
 EBN bursts beginning after the onset of LLBN activity. The right EBN
-produced a small burst at the end of a saccade. Furthermore, increases
-in activity of the left TN were mirrored by decreases in the right.
-Finally, OPN activity dropped to zero during production of a saccade.
+produces a small burst at the end of a saccade. Furthermore, increases
+in activity of the left TN are mirrored by decreases in the right.
+Finally, OPN activity drops to zero during production of a saccade.
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
@@
 burst amplitude for LLBNs and EBNs. With increasing input strength, both
-amplitude and duration of activation increased in long-lead and
+amplitude and duration of activation increase in long-lead and
 excitatory burst neurons. As before, these results accord well with
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
@@
 strengths applied to the horizontal and vertical SG (figure
-\ref{fig:fig_3}). As in the original publication, saccades were
-generally straight with a slight tendency to curve. We extracted
-quantitative estimates of saccade amplitude for the original publication
-from their figure 6 and calculated the root-mean-squared error (RMSE)
-between these estimates and saccade amplitudes produced by our model.
-The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
-0.22\textdegree, 0.14\textdegree~for the blue, green, red, turqoiuse,
+\ref{fig:fig_3}). As in the original publication, saccades are generally
+straight with a slight tendency to curve. Using
+\href{http://plotdigitizer.sourceforge.net/}{PlotDigitizer}, we extract
+saccade endpoints displayed in figure 6 of the original publication.
+This provides us with estimates of saccade amplitude allowing us to
+quantitatively asses our results in terms of the root-mean-squared error
+(RMSE). The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
+0.22\textdegree, 0.14\textdegree~for the blue, green, red, turquoise,
 and purple curves, respectively. Given that saccade amplitudes are
@@
 \caption{\textbf{Oblique saccades.} Inputs to the right and upward LLBNs
-were \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
+are \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
 \(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.22}\) (green);
 \(\mathrm{{I}_{r}=0.74}\) \& \(\mathrm{{I}_{u}=0.4}\) (red);
-\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turqoiuse); and
-\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and were
-applied for \(75\,\mathrm{ms}\). As expected, saccades were fairly
-straight with a slight tendency to curve.\label{fig:fig_3}}
+\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turquoise); and
+\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and are
+applied for \(75\,\mathrm{ms}\).\label{fig:fig_3}}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.33000\textwidth,height=0.45000\textwidth]{figures/fig4.eps}
-\caption{\textbf{Saccadic staircase.} Three saccades in a staircase
-continued in the same direction as the initial saccade. Inputs to the
-right and upward LLBNs were equal to \(\mathrm{{I}_{r}=0.2}\) and
-\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). Eye position was
+\caption{\textbf{Saccades in a staircase.} Eye positions exhibited in a
+saccade staircase simulation. Dense clusters reflect saccade endpoints.
+A total of three saccades are exhibited with subsequent saccades
+continuing in the same direction as the initial saccade. Inputs to the
+right and upward LLBNs are equal to \(\mathrm{{I}_{r}=0.2}\) and
+\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). The eye position is
 sampled every \(2\,\mathrm{ms}\).\label{fig:fig_4}}
@@
 
-In the fifth simulation the average activity of the left EBN was
-obtained for a series of saccades with different directions. Figure
+In the fifth simulation the average activity of the left EBN is obtained
+for a series of saccades with different directions. Figure
 \ref{fig:fig_5} shows a polar plot of average activity corresponding to
-each saccade reflecting the neuron's tuning curve. The tuning curve we
-observed for the left EBN exhibits a cardioid-like shape as was the case
-in the original publication (figure 8).
+each saccade. This is the neuron's tuning curve. The tuning curve we
+observe for the left EBN exhibits a cardioid-like shape as in the
+original publication (figure 8).
 
-\begin{longtable}[]{@{}lllllllllllllll@{}}
+\pagebreak
+
+\begin{longtable}[c]{@{}lllllllllllllll@{}}
 \caption{Direction specific inputs to SG to produce EBN tuning curve.
@@
 \caption{\textbf{EBN tuning curve.} Tuning curve of the left EBN
-exhibiting a cardioid-like shape. Inputs to the SG producing the desired
-saccades can be found in table \ref{tbl:table_1}. Each of these inputs
-was applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
+exhibiting a cardioid-like shape. Inputs to the SG producing the
+saccades are given in table \ref{tbl:table_1}. Each of these inputs is
+applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
 \end{figure}
 
-\pagebreak
-The sixth simulation reported in the original publication was designed
-to replicate results of Stanford \textit{et al.}
-\autocite{Stanford1996}. These authors stimulated the superior
-colliculus (SC) at various frequencies and measured the resulting
-saccade amplitude, duration, and velocity; showing that ampltidude
-saturates before velocity. Our implementation of the SG model was
-capable of replicating these results. However, reproducing simulation
-results reported by Gancarz \& Grossberg \autocite{Gancarz1998} with our
-implementation was complicated by the fact that the stimulation protocol
-given by the authors lead to the production of two rather than a single
-saccade for larger stimulation intensities. Furthermore, the authors did
-not report their criteria for identifying saccade on- and offsets.
-Stanford \textit{et al.} \autocite{Stanford1996} used velocity criteria
-to determine onset (\(v\textgreater30\,\mathrm{deg/s}\)) and offset
-(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provided us
-with explicit criteria, velocity did not always drop below
+The sixth simulation reported in the original publication is designed to
+replicate results of Stanford \textit{et al.} \autocite{Stanford1996}.
+These authors stimulated the superior colliculus (SC) at various
+frequencies and measured the resulting saccade amplitude, duration, and
+velocity; showing that amplitude saturates before velocity. Our
+implementation of the SG model is capable of replicating these results.
+However, reproducing simulation results reported by Gancarz \& Grossberg
+\autocite{Gancarz1998} with our implementation is complicated by the
+fact that the stimulation protocol given by the authors leads to the
+production of two rather than a single saccade for larger stimulation
+intensities. Furthermore, the authors do not report their criteria for
+identifying saccade on- and offsets. Stanford \textit{et al.}
+\autocite{Stanford1996} use velocity criteria to determine onset
+(\(v\textgreater30\,\mathrm{deg/s}\)) and offset
+(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provides us
+with explicit criteria, velocity does not always drop below
 \(30\,\mathrm{deg/s}\) after the first saccade before rising again with
-the second. To determine the offset of a saccade in those cases, we
-found the local minimum between the end of the first and the beginning
-of the second saccade. With these criteria in place, we stimulated the
-SC. Results of our simulation are shown in figure \ref{fig:fig_6}. We
-extracted quantitative results of the original publication from their
-figure 9 and calculated the RMSE between these estimates and our results
-for each curve. The RMSE was 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
+the second. To determine the offset of a saccade in those cases, we find
+the local minimum between the end of the first and the beginning of the
+second saccade. With these criteria in place, we stimulate the SC.
+Results of our simulation are shown in figure \ref{fig:fig_6}. Again we
+use PlotDigitizer to extract a vector of saccade amplitude, duration and
+peak velocity observed at each stimulation frequency in the original
+publication from their figure 9 and calculate the RMSE between these and
+our curves. The RMSE is 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
 \(4.45\,\mathrm{deg/s}\) for saccade amplitude, duration, and peak
@@
 reported in the original publication. This includes the observation that
-saccade amplitude and especially saccade duration were largest for a
+saccade amplitude and especially saccade duration are largest for a
 stimulation intensity of 1.2. It is conceivable that this intensity
-marks the point after which the SG generator starts producing two rather
-than a single saccade with the eye movement at this intensity merging
-two saccades and thus being stretched out.
+marks the point after which the SG starts producing two rather than a
+single saccade with the eye movement at this intensity merging two
+saccades and thus being stretched out.
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.75000\textwidth,height=0.37500\textwidth]{figures/fig6.eps}
-\caption{\textbf{Effect of stimulation frequency on saccade (A)
-amplitude, (B) duration, and (C) velocity.} Saccade amplitude (A)
-saturated before saccade velocity (C). A range of unitless values
-varying between 1 and 2.4 at increments of 0.2 were applied to the SC
-(reflecting stimulation at different frequencies). The connection weight
-from SC to LLBN was \(\mathrm{W=2}\) and stimulation duration was
+\caption{\textbf{Effect of stimulation frequency on saccade amplitude
+(A), duration (B), and velocity (C).} A range of unitless values varying
+between 1 and 2.4 at increments of 0.2, reflecting stimulation at
+different frequencies ,is applied to the SC. The connection weight from
+SC to LLBN is \(\mathrm{W=2}\) and stimulation duration is
 \(125\,\mathrm{ms}\).\label{fig:fig_6}}
@@
 to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
-publication). To produce a low velicity saccade an input of
-\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
-input curve produced by our implementation of the original equations
-(A), a second saccade can be observed for both high and low velocity
-saccades. If the input curves reported in the original manuscript were
-recreated in an ad-hoc fashion (B), the responses of all model neurons
-matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
+publication; solid curve). To produce a low velocity saccade an input of
+\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\) (dashed curve). Given the
+shape of the input curve produced by our implementation of the original
+equations (A), a second saccade can be observed for both high and low
+velocity saccades. If the input curves reported in the original
+manuscript are recreated in an ad-hoc fashion (B), the responses of all
+model neurons matches those in \autocite{Gancarz1998}\label{fig:fig_7}}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
@@
 \caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
-OPN stimulation interrupted the saccade, which remained accurate
-nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
-for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
-input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
-\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
-uninterrupted saccade.\label{fig:fig_9}}
+OPN stimulation interrupts the saccade, which remains accurate
+nonetheless. An input of (\(\mathrm{I=0.7}\)) is applied to the LLBN for
+\(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the input,
+the OPN is stimulated (\(\mathrm{J=1.8}\)) for \(5\,\mathrm{ms}\). The
+dashed curve shows TN activity for an uninterrupted
+saccade.\label{fig:fig_9}}
 \end{figure}
@@
 traded while keeping amplitude constant. To produce a high-velocity
-saccade, the SC was stimulated at a high frequency. Similarly, to
-produce a low-velocity saccade, the SC was stimulated at a low
-frequency. Figure \ref{fig:fig_7} shows the results of our simulation.
-In line with the original publication, the saccade amplitude reflected
-by TN activity was identical after high- and low-frequency saccades thus
-confirming the reported effect. However, we observe differences in the
-specific details of our implementation with respect to the original,
-starting with the shape of the input curves. While the rise of the input
-curves in the original publication and our implementation coincide well,
-the decay in our curves started later and proceeded with a shorter time
-constant (\(\approx25\,\mathrm{ms}\), i.e.~half the simulation time
-constant). This resulted in more total input to the model in our
-implementation and hence lead to the generation of a second saccade not
-observed in the original publication. When repeating our simulation
-using an ad-hoc solution to create matching input curves (by halving the
-time constant after external stimulation), we observed results identical
-to those reported in the original publication. We proceeded to
-investigate the origin of the discrepancy with regard to the shapes of
-the input curves. First, we tested whether discrepancies were specific
-to our NEST implementation by solving the expressions describing the
-input analytically. Specifically, the analytic expression of the input
-\(I(t)\) described by equations \ref{eq:sc} - \ref{eq:stim} is
-\begin{equation}
+saccade, the SC is stimulated at a high frequency. Conversely, to
+produce a low-velocity saccade, the SC is stimulated at a low frequency.
+Figure \ref{fig:fig_7} shows the results of our simulation. In line with
+the original publication, the saccade amplitude reflected by TN activity
+is identical after high- and low-frequency saccades thus confirming the
+reported effect. However, we observe differences in the specific details
+of our implementation with respect to the original, starting with the
+shape of the input curves. We evaluate rise and decay of input curves
+reported in the original publication by estimating their onset as well
+as time constants from points obtained from figure 10 using
+PlotDigitizer. While the rise of the input curves in the original
+publication and our implementation coincide well, the decay in the
+original curves starts later and proceeds with a time constant of about
+half the magnitude in our simulation (\(\approx25\,\mathrm{ms}\)). This
+results in more total input to the model in our implementation and hence
+leads to the generation of a second saccade not observed in the original
+publication. When repeating our simulation using an ad-hoc solution to
+create matching input curves by halving the time constant after external
+stimulation, we observe results identical to those reported in the
+original publication. Next we investigate the origin of the discrepancy
+with regard to the shapes of the input curves. First, we test whether
+discrepancies are specific to our NEST implementation by solving the
+expressions describing the input analytically. Specifically, the
+analytic expression of the input \(I(t)\) described by equations
+\ref{eq:sc} - \ref{eq:stim} is \begin{equation}
 I(t)=
@@
 \begin{array}{lll}
-    0 \quad &\textrm{if} \quad t<{t}_{on} \\
+    0 \quad &\textrm{if} \quad t \leq {t}_{on} \\
     W \cdot f(F \cdot (1-{e}^{-(t-{t}_{on})/ \tau})) \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.}\\
@@
 given parameter values reported in the original manuscript exactly
-reproduced our numerical results and thus produced curves equally
-deviating from those shown in the original publication. Next, we
+reproduces our numerical results and thus produces curves equally
+deviating from those shown in the original publication. At this point we
 contacted one of the original authors to rule out the possibility that
-we used erroneous parameter settings but no mistake was pointed out to
-us. The origin of the discrepancy thus remains unclear. However, it is
-likely it stems from the description of the input rather than from
-problems with the SG model itself.
+we use erroneous parameter settings but no mistake was found. While the
+precise origin of the discrepancy thus remains unclear, most likely the
+original code uses different time constants for rise and decay.
 
@@
 produces smooth eye movements. Our results, shown in figure
-\ref{fig:fig_8}, reproduced these findings as they strongly resembled
+\ref{fig:fig_8}, reproduce these findings as they strongly resemble
 those shown in figure 11 of the original publication. Specifically, an
-initial burst exhibited by the EBN was followed by sustained lower
-activity. This was due to inhibitory feedback being insufficient to
-silence the LLBN when input remains continuously strong and resulted, in
+initial burst exhibited by the EBN is followed by sustained lower
+activity. This is due to inhibitory feedback being insufficient to
+silence the LLBN when input remains continuously strong and results, in
 turn, in the observed smooth eye movement.
@@
 neurons when the OPN is briefly electrically stimulated while a constant
-input is applied to the LLBN. External stimulation temporarily restored
+input is applied to the LLBN. External stimulation temporarily restores
 activation in the OPN and hence also inhibition of the EBN, leading to
 an interruption of the saccade. As is shown in figure \ref{fig:fig_9},
-the saccade remained accurate despite this disruption. This is in
+the saccade remains accurate despite this disruption. This is in
 agreement with results shown in figure 12 of the original publication.
@@
 those reported by Gancarz and Grossberg \autocite{Gancarz1998} and
-accorded well quantitatively whenever such information was available.
-The only discrepancy between our and the original implementation was
-that the input to the saccade generator took more time to decay in our
-implementation. While the exact cause of this discrepancy is unclear, we
-could rule out that it was specific to the NEST implementation.
-Furthermore, it does not affect reproduction of the main finding of
-simulation seven; namely that saccade velocity and duration can be
-traded while keeping its amplitude constant.
+accord well quantitatively wherever such information is available. The
+only discrepancy between our and the original implementation is that our
+implementation produces a second saccade in simulation 7 not observed in
+the original publication. While the exact cause of this discrepancy
+remains unclear, we could trace it back to the shape of the input
+curves. An analytical treatment reveals that the discrepant curves
+result from the description of the input allowing us to rule out that
+they originate from the NEST implementation. Furthermore, when the SG
+model is provided with corrected input curves, it faithfully reproduces
+the results reported in the original publication for simulation 7;
+namely that saccade velocity and duration can be traded while keeping
+the amplitude constant. There thus appear to be no issues with either
+the NEST implementation nor the SG model.
 
@@
 European Union's Horizon 2020 Research and Innovation Programme under
-Grant Agreement No. 7202070 (HBP SGA1).
+Grant Agreements No. 7202070 (HBP SGA1) and 737691 (HBP SGA2).
 
@@
 For a detailed, comprehensive, treatment of rate-based neuron models in
-NEST see Hahne et al.\autocite{Hahne2017}. Here we provide only a brief
-overview to facilitate readability/interpretability of the code provided
-with this publication. In general, rate-based neurons in NEST consist of
-two components; a neuron and a gain function.
+NEST see Hahne et al. \autocite{Hahne2017}. Here we provide only a brief
+overview to facilitate readability and interpretability of the code
+provided with this publication. In general, rate-based model neurons in
+NEST consist of two components; an abstract neuron class and a gain
+function.
 
-\subsubsection{Neuron}\label{neuron}
+\subsubsection{Neuron class}\label{neuron-class}
 
-Two types of neuron base classes exist in NEST v2.16.0:
+Two types of neuron base classes exist in NEST 2.16.0:
 \texttt{rate\_neuron} and\\
@@
 can apply a nonlinear gain function to its inputs either before or after
-their weighted summation. It is numerically integrated using the EE
-method unless a neuron does not exhibit exponential decay in which case
-the EM method is used. Furthermore, rates can be bounded (from below at
-zero) or unbounded. Finally, noise (in the form of a Wiener process) can
-be added to its input or to its output. The \texttt{rate\_neuron} has a
-total of seven parameters (see table \ref{tbl:table_2}) and two further
-properties \texttt{rate} (current firing rate) and \texttt{noise}
-(current noise level).
+their weighted summation. The model is numerically integrated using the
+EE method unless a neuron does not exhibit exponential decay in which
+case the Euler-Maruyama method is used. Furthermore, rates can be
+bounded (from below at zero) or unbounded. Finally, noise (in the form
+of a Wiener process) can be added to its input or to its output. The
+\texttt{rate\_neuron} has a total of seven parameters (see table
+\ref{tbl:table_2}) and two further properties \texttt{rate} (current
+firing rate) and \texttt{noise} (current noise level).
 
-\begin{longtable}[]{@{}lll@{}}
+\begin{longtable}[c]{@{}lll@{}}
 \caption{\texttt{rate\_neuron} parameters.
@@
 function after summation\tabularnewline
-\texttt{rectify\_output} & \texttt{false} & if true, multiplicative
+\texttt{rectify\_output} & \texttt{false} & if true, rate is
+bounded\tabularnewline
+\texttt{mult\_coupling} & \texttt{false} & if true, multiplicative
 coupling\tabularnewline
-\texttt{mult\_coupling} & \texttt{false} & if true, rate is
-bounded\tabularnewline
 \bottomrule
@@
 
-Five gain functions exist in NEST v2.16.0: \texttt{lin\_rate},
+Five gain functions exist in NEST 2.16.0: \texttt{lin\_rate},
 \texttt{sigmoid\_rate},\\

### 2018-04-17 article/Caze_2018.tex

@@
 \begin{cases}
-    x              & \text{, if } x\leq V_a\\
-    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{, if } V_a<x \leq V_b\\
-    V_c            & \text{, otherwise.}
+    x,              & \text{ if } x\leq V_a,\\
+    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{ if } V_a<x \leq V_b,\\
+    V_c,            & \text{ otherwise.}
 \end{cases}
@@
 
-The parameter values used in the original study as well as in our replication are summarized in Table~\ref{tab:parameters}.
+The parameter values used in the original study as well as in our replication are summarised in Table~\ref{tab:parameters}.
 
@@
 
-We used a the predefined \verb|clip| function to implemented non-linear coupling into a network. The function clips a value to a given range, following the syntax
+We used the predefined \verb|clip| function to implement non-linear coupling in the network. The function clips a value to a given range, following the syntax
 \begin{code}
@@
 
-We employed this functions on the received excitatory inputs stored in a placed holder variable \verb|ve|.
-The variable corresponds to x in eq.~\ref{eq:non-linear}.
+We employed this function on the received excitatory inputs stored in a placed holder variable \verb|ve|.
+The variable corresponds to $x$ in eq.~\ref{eq:non-linear}.
 We then reset the variable \verb|ve| back to 0 to prepare for the next time step.
@@
 Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike.
-This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
+This yields the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
 \begin{equation}
@@
 \textbf{Propagation of synchronous spiking in linearly (A) and non-linearly (B) coupled networks.} We initiate a chain of synchronous pulses by applying external supra-threshold inputs to the first 100 neurons at time $t=\SI{150}{\milli\second}$ (red coloured spikes, grey vertical dotted lines indicate times where spikes occur as part of the chain).
-Top row: total size $g'$ of synchronized groups within the chain.
+Top row: total size $g'$ of synchronised groups within the chain.
 Middle row: network rate over all neurons (rate in \si{\kilo\hertz}, bin size \SI{1}{\milli\second}).
@@
 For quantitative definitions, see Table~\ref{tab:color}.
-We run each combination of connection strengths 10 times, and combine the results to yield a color in RGB notation (values between 0 and 1 for the red, green, and blue component), according to:
+We run each combination of connection strengths 10 times, and combine the results to yield a colour in RGB notation (values between 0 and 1 for the red, green, and blue component), according to:
 \begin{equation}
@@
 
-Our replication does not reproduce all the quantitative details of the original plots.
-The parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article.
-For linear networks, we do observe almost no simulations classified in this category, while the original article has a thin line of such classifications between the stable regime without propagation (green) and the regime where simulations are unstable even without stimulation (red).
-Similarly, for non-linearly coupled networks the parameter exploration in the original article shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} where the network is unstable only after the synchronous stimulation (yellow).
+Our replication does not reproduce all quantitative details of the original plots.
+For non-linearly coupled networks, the parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article.
+The original article shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} with such a behaviour.
 In contrast, our simulations show a much smaller such zone when the inhibitory connection strength is small (Fig.~\ref{fig:grid}B).
@@
 The forced synchronisation of all spikes to a grid of \SI{0.1}{\milli\second}, together with the homogeneity of the synaptic delays, leads to strong oscillations in our simulation (Fig.~\ref{fig:raster_strong}A), which is consequently classified as unstable.
-When we use a simulation time step of \SI{0.01}{\milli\second}, these osciallations disappear and the network activity therefore gets classified as ``stable without persistent propagation''.
+When we use a simulation time step of \SI{0.01}{\milli\second}, these oscillations disappear and the network activity therefore gets classified as ``stable without persistent propagation''.
 Finally, non-linearly coupled networks show a discontinuity for weak inhibitory connection strengths in the original publication.
@@
 The grid search over parameters showed a higher prevalence of networks with unstable activity before the onset of stimulation compared to the original study.
-In the analysis of the evolution of synchronous pulses of different sizes, our results need a correction for the use of a clock-based simulation method to show a good quantative match with the original results.
+In the analysis of the evolution of synchronous pulses of different sizes, our results need a correction for the use of a clock-based simulation method to show a good quantitative match with the original results.
 However, a marked difference remains for small initial group sizes.
-The differences we see are most likely due to our use of a clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision.
+The differences we see are most likely due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision.
 However, the main qualitative finding---non-linearly coupled networks have a large parameter range supporting persistent propagation of synchronous activity whereas linearly coupled networks do not---remains if we use finer (\SI{0.05}{\milli\second}) or coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions (data not shown).
@@
 \caption{\label{fig:grid_detail}
-\textbf{Zoom on qualitative network behaviour with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}.
+\textbf{Zoom on qualitative network behaviour with different simulation time steps.} Colour code as in Fig.~\ref{fig:grid}.
 Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}, and only for non-linearly coupled networks with an synchronous stimulation of 100 neurons.
@@
 \textbf{Evolution of synchronous pulses without correction in linearly (A) and non-linearly (B) coupled networks.} All conventions as in Fig.~\ref{fig:markov_corrected}.
-The numerical results shown here (gray lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
+The numerical results shown here (grey lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
 \end{figure}

### 2018-04-24 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
corrected index sequence & forward Euler -> Forward Euler

@@
 (equivalent descriptions apply to the vertical component). Please note
-that subscripts \(\mathrm{l}\), \(\mathrm{r}\), \(\mathrm{d}\) and
-\(\mathrm{u}\) indicate whether a neuron controls upward, downward,
+that subscripts \(\mathrm{u}\), \(\mathrm{d}\), \(\mathrm{l}\) and
+\(\mathrm{r}\) indicate whether a neuron controls upward, downward,
 leftward or rightward saccade direction, respectively.
@@
 at a time step of \(0.05\,\mathrm{ms}\) and a time constant of
-\(50\,\mathrm{ms}\). Tonic neurons are integrated using the forward
+\(50\,\mathrm{ms}\). Tonic neurons are integrated using the Forward
 Euler method since they do not exhibit passive decay leading to

### 2018-05-08 article/Poisot-2017.tex
Update gitignore
Compile the paper

@@
 \usepackage{marginnote}
+\usepackage{amssymb,amsmath}
 \usepackage{graphicx}
@@
 \usepackage{fontawesome}
-\usepackage{amssymb,amsmath}
 \usepackage{caption}
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 combinations of species to persist even though coexistence would not be
-possible in a purely deterministic world. Here I propose an
-implementation of this model in \emph{Julia} \autocite{beza17jfa}, and
-show that it is able to reproduce most figures from the original
-manuscript.
+possible in a purely deterministic world. This was an influential paper
+in guiding subsequent research on the co-existence between specialists
+and generalists, as it showed a strong effect on the structure of
+environmental change. Here I propose an implementation of this model in
+\emph{Julia} \autocite{beza17jfa}, and show that it is able to reproduce
+most figures from the original manuscript.
 
-\section{Methods}\label{methods}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 non-optimal environment, and \(b\) is the fitness cost of generalism.
-Note that \(1 > b > a > 0\).
+Note that \(1 > b > a > 0\). The initial value of \(K_1\) is chosen
+randomly, using
+
+\[K_1 = K_{1,\text{min}} + \text{rnd}_1(K_{1,\text{max}}-K_{1,\text{min}})\,.\]
+
+If the variation of the two environments is linked, then
+
+\[K_2 = K_{2,\text{min}} + (1-\text{rnd}_1)(K_{2,\text{max}}-K_{2,\text{min}})\,.\]
+
+If the variation is independent, then
+
+\[K_2 = K_{2,\text{min}} + \text{rnd}_2(K_{2,\text{max}}-K_{2,\text{min}})\,.\]
+
+Both \(\text{rnd}_1\) and \(\text{rnd}_2\) are random numbers uniformly
+distributed in \([0,1]\).
 
@@
 
-The value of \(p_l\) for which \(D_{l1}=D_{l2}\) is
+The value of \(p_l\) for which \(D_{l1}=D_{l2}\) (checked using
+\textcite{maxi14mca}) is
 
@@
 We fix \(p_m\) and \(p_n\), and find the value of \(p_l\), while
-enforcing the constraint of \(0 \leq p_l \leq 1\). Repeating this
-procedure a few times for the different species yields the optimal
-values of \(\mathbf{p}\); we can measure the density of individuals in
-both habitats. Before we do so, there is a proportion \(g\) of
-individuals that select habitats at random. Given a total population
-size of \(N_i\), there are \(N_i(g/2)\) individuals will go to either
-habitat, and \(N_i(1-g)p_i\) will pick habitat 1. With this information,
-we can write the matrix describing habitat selection:
+enforcing the constraint of \(0 \leq p_l \leq 1\). Iterating this
+procedure a few times (10 was sufficient in all cases examined) for the
+different species yields the optimal values of \(\mathbf{p}\); we can
+measure the density of individuals in both habitats. Before we do so,
+however, we allow a proportion \(g\) of individuals that select habitats
+at random. Given a total population size of \(N_i\), there are
+\(N_i(g/2)\) individuals will go to either habitat, and \(N_i(1-g)p_i\)
+will pick habitat 1. With this information, we can write the matrix
+describing habitat selection:
 
@@
 
+The entire sequence within a timestep is: generate \(K\) for each patch;
+find preferences of species for both patches; distribute the species
+across patches; update population sizes based on the local fitness. This
+is iterated over as many timesteps as required to reach equilibrium.
+
 \hypertarget{tbl:prm}{}
@@
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 provide some of these values, and they were used to test this
-implementation. All default parameter values are given in
+implementation. All default parameter values are given in Table
 ~\ref{tbl:prm}. In all figures, circles represent the generalist, upper
@@
 value, as they are based on non-stochastic situations. These are
-reproduced in ~\ref{fig:02}. As in the original simulations, the
+reproduced in Fig. ~\ref{fig:02}. As in the original simulations, the
 generalist alone overshoots its carrying capacity at \(t=5\), and the
@@
 
+\hypertarget{fig:02}{%
 \begin{figure}
@@
 alone (left), and following invasion by the two specialists at the
-generalist equilibrium (right).}\label{fig:02}
+generalist equilibrium (right). Circles represent the generalist, upper
+and lower triangles represent the specialists of habitats 1 and 2
+(respectively), and a diamond is used when both specialists have the
+same behaviour.}\label{fig:02}
 \end{figure}
+}
 
-In ~\ref{fig:03}, this implementation gives the same qualitative results
-as the original article. There are small quantitative differences in
-output, which are in my opinion explained by the initial population
-densities (which are not given in the original article). Absent
-variation in habitat quality, generalists cannot persist, and the
+In Fig. ~\ref{fig:03}, this implementation gives the same qualitative
+results as the original article. There are small quantitative
+differences in output, which are in my opinion explained by the initial
+population densities (which are not given in the original article).
+Absent variation in habitat quality, generalists cannot persist, and the
 specialists both reach \(N^\star = \bar K\). There is a value of
@@
 
+\hypertarget{fig:03}{%
 \begin{figure}
@@
 \(K_1\) and \(K_2\). Dashed lines represent independent variation, and
-solid lines tied variations, as in the original article.}\label{fig:03}
+solid lines tied variations, as in the original article. Circles
+represent the generalist, upper and lower triangles represent the
+specialists of habitats 1 and 2 (respectively), and a diamond is used
+when both specialists have the same behaviour.}\label{fig:03}
 \end{figure}
+}
 
-In ~\ref{fig:04}, the same qualitative result (generalists increase in
-abundance when one habitat becomes larger). The quantitative differences
-can, again, be explained by changes in the initial population densities.
-Finally, the results in ~\ref{fig:05} and ~\ref{fig:06} can also be
-replicated.
+In Fig. ~\ref{fig:04}, the same qualitative result (generalists increase
+in abundance when one habitat becomes larger). The quantitative
+differences can, again, be explained by changes in the initial
+population densities. Finally, the results in Fig. ~\ref{fig:05} and
+Fig. ~\ref{fig:06} can also be replicated; it should be noted that for
+Fig. ~\ref{fig:06}, the lines cross at a different point, and diverge at
+different rates, than in the original paper. Additional analyses suggest
+that this depends on the number of timesteps used, as well as on the
+starting populations. The original manuscript does not provides these
+information, and so we will have to accept a \emph{qualitative}
+replication of this result.
 
+\hypertarget{fig:04}{%
 \begin{figure}
@@
 \includegraphics{figure04.pdf}
-\caption{Consequences of increasing the range of variation of both
-\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
-solid lines tied variations, as in the original article.}\label{fig:04}
+\caption{Consequences of changing the relative quality of habitats 1 and
+2 on coexistence. Circles represent the generalist, upper and lower
+triangles represent the specialists of habitats 1 and 2 (respectively),
+and a diamond is used when both specialists have the same
+behaviour.}\label{fig:04}
 \end{figure}
+}
 
+\hypertarget{fig:05}{%
 \begin{figure}
@@
 \includegraphics{figure05.pdf}
-\caption{Consequences of increasing the range of variation of both
-\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
-solid lines tied variations, as in the original article.}\label{fig:05}
+\caption{Consequences of changing \(b\), the ability of the generalist,
+on coexistence. Dashed lines represent independent variation, and solid
+lines tied variations, as in the original article. Circles represent the
+generalist, upper and lower triangles represent the specialists of
+habitats 1 and 2 (respectively), and a diamond is used when both
+specialists have the same behaviour. Note that the direction of the axis
+is \emph{reversed} with regard to the original figure.}\label{fig:05}
 \end{figure}
+}
 
+\hypertarget{fig:06}{%
 \begin{figure}
@@
 \includegraphics{figure06.pdf}
-\caption{Consequences of increasing the range of variation of both
-\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
-solid lines tied variations, as in the original article.}\label{fig:06}
+\caption{Consequences of changing the proportion of individuals picking
+their habitat at random. Circles represent the generalist, upper and
+lower triangles represent the specialists of habitats 1 and 2
+(respectively), and a diamond is used when both specialists have the
+same behaviour.}\label{fig:06}
 \end{figure}
+}
 
-\section{Conclusion}\label{conclusion}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
@@
 usable for teaching, as it is not time consuming to generate results. As
-an example, I merged the output of ~\ref{fig:03} and ~\ref{fig:05}, to
-generate the response of Shannon's evenness (which measures the extent
-to which all species have the same density) to changes in the rate of
-variation \emph{and} generalist ability. The results are presented in
-~\ref{fig:surf}.
+an example, I merged the output of Fig. ~\ref{fig:03} and Fig.
+~\ref{fig:05}, to generate the response of Pielou's evenness
+\autocite{piel66sps},
 
+\[
+J' = -\left(\sum_{i=1}^3\frac{N_i}{N}\text{ln}\frac{N_i}{N}\right)\frac{1}{\text{ln}\,3} \,,
+\]
+
+which measures the extent to which all species have the same density. It
+appears that \(J'\), a measure of coexistence, changes in response to
+both the rate of variation \emph{and} generalist ability. The results
+are presented in Fig. ~\ref{fig:surf}. This figure is primarily intended
+to show that this model can still generate new insights when we vary
+multiple parameters: namely, the response of equitability to changes in
+multiple variables is non-linear, and maximum equitability is reach for
+high generalist ability but moderate range of variation (this was not
+predictable from the outputs of Figs. ~\ref{fig:04} and ~\ref{fig:05}
+alone).
+
+\hypertarget{fig:surf}{%
 \begin{figure}
@@
 \emph{and} the range of variation. While the response to environmental
-change is bell-shaped when generalists are competitive (\(g\) close to
+change is hump-shaped when generalists are competitive (\(b\) close to
 unity), it becomes monotonously decreasing for generalists with lower
-competitive ability (\(g\) close to \(a\)).}\label{fig:surf}
+competitive ability (\(b\) close to \(a\)).}\label{fig:surf}
 \end{figure}
+}
 

### 2018-05-17 article/Bavard-Thero-2018.tex
Add files via upload

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Sophie Bavard,
      Héloïse Théro,
  },
            pdftitle=
{
[Re] Adaptive properties of differential learning rates for positive and
negative outcomes
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Adaptive properties of differential learning rates for positive and
negative outcomes}

    \usepackage{authblk}
                        \author[1]{Sophie Bavard}
                    \author[1]{Héloïse Théro}
                            \affil[1]{Laboratoire de Neurosciences Cognitives et Computationnelles (ENS -
INSERM), Département d'Études Cognitives, École normale supérieure, PSL
Research University, 29 rue d'Ulm, 75005 Paris, France}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:sophie.bavard@gmail.com, thero.heloise@gmail.com}{sophie.bavard@gmail.com, thero.heloise@gmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Olivia Guest\\

  {\bfseries Reviewers}\\
        Benoît Girard\\
        Xavier Hinault\\

  {\bfseries Received}  Feb, 9, 2018\\
  {\bfseries Accepted}  ?, ?, ?\\
  {\bfseries Published} ?, ?, ?\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Cazé, R. D., \& van der Meer, M. A. (2013). Adaptive properties of
differential learning rates for positive and negative outcomes.
Biological cybernetics, 107(6), 711-719.
https://doi.org/10.1007/s00422-013-0571-5
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Reinforcement learning represents a fundamental cognitive process:
learning by trial and error to maximize rewards and minimize
punishments. Current and most influential theoretical models of
reinforcement learning assume a unique learning rate parameter,
independently of the outcome valence \cite{sutton1998introduction, o2004dissociable, behrens2007learning}. However
human participants were shown to integrate differently positive and
negative outcomes \cite{frank2004carrot, frank2007genetic, sharot2011unrealistic}. This
motivated Cazé and van der Meer to implement a modified version of the
reinforcement learning model, with two distinct learning rates for
positive and negative outcomes \cite{caze2013adaptive}.

They have shown that although differential learning rates shifted reward
predictions and could thus be seen as a maladaptive bias, this model can
outperform the classical reinforcement learning model on tasks with
specific outcome probabilities. Following Cazé and van der Meer's
predictions \cite{caze2013adaptive}, a subsequent empirical article have modeled human behavior
on these specific tasks \cite{gershman2015learning}. The question
is still an active research area, as various articles have further
investigated the difference learning rates bias
\cite{garrett2014robust, moutsiana2015human, shah2016pessimistic, 
garrett2017optimistic, lefebvre2017behavioural, palminteri2017confirmation}.

At the time we first submitted this manuscript, we were not aware that the
 code of the reference article was freely available \href{https://github.com/rcaze/13_02CaVa}{online}. 
Our model and analysis scripts 
were therefore implemented from scratch using Python (version 3.6.3), with 
the NumPy (version 1.14.0) and Matplotlib (version 2.1.1) libraries.

% online (\url{https://github.com/rcaze/13_02CaVa}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We first implemented our scripts on Matlab, as we were more familiar
with this language, and then adapted them on Python.

We used the modeling description of the reference article to implement
our replication. They used standard Q-learners with a softmax action
selection rule \cite{sutton1998introduction}, and their precise
description enabled us to implement them with low difficulty. But we
found four ambiguities in the simulation procedure.

First, the authors described their analytical results to be valid for
``\(Q_0 \neq \{-1, 1\}\)'' in section 2, but did not specify what value
of \(Q_0\) they used in all the following simulations. We chose to use
\(Q_0 = 0\), as this initial value is the middle point between the two
possible outcomes (i.e., -1 and 1). As we replicated all the original
figures, even the dynamics in the beginning of the learning curves (see
Figures ~\ref{fig:figure2}A, ~\ref{fig:figure3} and ~\ref{fig:figure4}B),
we believe the reference article must have used similar initial
Q-values.

Second, regarding the parameter setting for Figure ~\ref{fig:figure1}'s
simulations, the ratio of \(\alpha^+\) over \(\alpha^-\) was said to be
either 0.25, 1 or 4, but they did not specify what were the exact values
of \(\alpha^+\) and \(\alpha^-\) used. We thus set them according to the
following description of the pessimistic, rational and optimistic agents
in section 3, i.e.:

\begin{itemize}
\tightlist
\item
  \(\alpha^+ = 0.1\) and \(\alpha^- = 0.4\) for the ratio of 0.25
\item
  \(\alpha^+ = 0.1\) and \(\alpha^- = 0.1\) for the ratio of 1
\item
  \(\alpha^+ = 0.4\) and \(\alpha^- = 0.1\) for the ratio of 4
\end{itemize}

Third, the number of iterations made to generate Figures
 \ref{fig:figure3} and \ref{fig:figure4} were not indicated, and we
assumed the authors used the same number as in Figures
 \ref{fig:figure1} and \ref{fig:figure2} (i.e., 5,000 runs).

Finally, it was unclear in the reference article whether the performances 
were averaged based on the probabilities to select the correct action, 
or based on the proportion of correct choices. We define the proportion 
of correct choices as the number of trials in which the correct action 
was chosen divided by the total number of trials. In contrast, the 
probability to select the correct action is computed by the model 
on each trial, and then used to choose an action. For example
Figure \ref{fig:figure2}'s legend indicated: ``Mean probability of choosing the best
arm'', suggesting that the probabilities themselves were used. However,
when commenting the figure in section 3, the authors appeared to say
that the actual choices were rather used: ``the optimistic agent learns
to take the best action significantly more than the rational agent''.
For our analyses, we started by using the probabilities of choice, as
this would lead to more clear, less noise-corrupted results. However we
then obtained very smooth learning curves, and were unable to reproduce
the spikiness of the original Figures \ref{fig:figure2}, 
\ref{fig:figure3} and \ref{fig:figure4}. We thus computed the
proportions of implemented choices for all our figures.

\hypertarget{results}{%
\section{Results}\label{results}}

All our figures reproduced the patterns of the original results. We were
even able to replicate the fine-grained details of the learning curves,
like the early bumps in performance in the high-reward task (Figures
 \ref{fig:figure2}A, \ref{fig:figure3} and \ref{fig:figure4}B, right
panels, around 50-100 trials). In Figure \ref{fig:figure1}, the mean
and the variance of the Q-values were also very similar as the ones in
the original figure. We numbered our figures in the same way as the reference article.

The only discrepancy we found was in Figure \ref{fig:figure4}A.
Although the general pattern was replicated, our learning curves
appeared smoother than in the reference article. As the number of
simulations were not explicitly specified for this figure, we cannot
know if this is due to us running a higher number of simulations than
the reference article, or from another difference in model
implementation.

\hypertarget{fig:figure1}{%
\begin{figure}
\centering
\includegraphics{Figure1.pdf}
\caption{Average estimated Q-values after 800 trials averaged for
different ratios of \(\alpha^+\) and \(\alpha^-\). The dotted lines
represent the underlying average reward: 0.8, 0.6, -0.6, -0.8. The error
bars represent the variance of the estimated
Q-values.}\label{fig:figure1}
\end{figure}
}

\hypertarget{fig:figure2}{%
\begin{figure}
\centering
\includegraphics{Figure2.png}
\caption{\textbf{A.} Performance, i.e., proportion of choices for the
best action, for the three agents: Rational (R, \(\alpha^+=\alpha^-\),
blue line), Optimistic (O, \(\alpha^+>\alpha^-\), green line) and
Pessimistic (P, \(\alpha^+<\alpha^-\), red line). In this figure and the
following ones, the left (resp. right) panel corresponds to the
low-reward (resp. high-reward) task. \textbf{B.} Proportion of action
switch after 800 trials for each agent, in the two different
tasks.}\label{fig:figure2}
\end{figure}
}

\hypertarget{fig:figure3}{%
\begin{figure}
\centering
\includegraphics{Figure3.png}
\caption{The performances of the Meta-learner (N) are shown in
\emph{purple} and those of the Rational agents (R) in different colors
of blue (in \emph{teal} for \(\alpha = 0.01\), in \emph{royal blue} for
\(\alpha = 0.1\) and in \emph{navy blue} for
\(\alpha = 0.4\)).}\label{fig:figure3}
\end{figure}
}

\hypertarget{fig:figure4}{%
\begin{figure}
\centering
\includegraphics{Figure4.png}
\caption{The performances of the Meta-learner, Optimistic, Rational and
Pessimistic agents \textbf{A.} in a task where the probabilities of
reward are 0.75 and 0.25 for the two choices. \textbf{B.} in a
``three-armed bandit'' task.}\label{fig:figure4}
\end{figure}
}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

All the figures in Cazé and van der Meer \cite{caze2013adaptive} have been successfully
reproduced with high fidelity, and we confirm the validity of their
simulations. Overall the whole replication procedure was smooth: the
models were implemented with low difficulty, and the simulations were
quite straightforward apart from a few obscure details. We hope this
replication can foster future research in the domain.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2018-05-21 article/Hathway-Goodman-2018.tex
add article as md, tex, pdf and add figures

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Pamela Hathway,
      Dan Goodman,,
  },
            pdftitle=
{
[Re] Spike Timing Dependent Plasticity Finds the Start of Repeating Patterns
in Continuous Spike Trains
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\defaultfontfeatures{Path = /usr/local/texlive/2018/texmf-dist/fonts/opentype/public/fontawesome/}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Spike Timing Dependent Plasticity Finds the Start of Repeating Patterns
in Continuous Spike Trains}

    \usepackage{authblk}
                        \author[1]{Pamela Hathway}
                    \author[1]{Dan Goodman,}
                            \affil[1]{Department of Electrical and Electronic Engineering, Imperial College,
London, UK}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:p.hathway16@imperial.ac.uk}{p.hathway16@imperial.ac.uk}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Spike Timing Dependent Plasticity Finds the Start of Repeating Patterns
in Continuous Spike Trains, Masquelier T, Guyonneau R, Thorpe SJ, PLoS
ONE 3(1): e1377, 2008. https://doi.org/10.1371/journal.pone.0001377
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Neurons communicate through repeated, specifically timed action
potential sequences (spike patterns) to convey information
\cite{Fellous2004, Prut1998}. Since neuronal activity is noisy and
neurons are likely involved in a multitude of spike patterns of various
lengths and extent, it can be hard to find spike patterns at first
glance. The more neurons are recorded, the more difficult the task
becomes due to the exponential increase of possible combinations of
spikes that could make up a pattern \cite{Buonomano2009}. It is unclear
how neurons in the brain may extract relevant information from such
input. In a 2008 paper, Masquelier and colleagues demonstrated that a
single neuron with spike timing dependent plasticity (STDP) synapses is
able to find the start of a repeating pattern in noisy (artificial) data
\cite{Masq2008}.

\textcite{Masq2008} use a simple feedforward network in which 2000 input
neurons connect to one output neuron via excitatory STDP synapses. Half
of the input neurons spike according to a spike pattern of 50 ms length
for about 25\% of the time. Finding the pattern is additionally made
difficult by jittering the pattern spikes, adding random noise spikes to
all neurons, and a constant population rate as well as no differences in
overall firing rate between neurons.

We replicated their findings using the spiking neural network simulator
Brian \cite{Goodman2009, Stimberg2014}, whereas the the original study
implemented the simulations in Matlab, with the main functions being
computed in C/C++ through mex files. In addition, we examined some of
the implementation parameters and investigated whether they were
essential to the success of the algorithm.

\section{Methods}\label{methods}

All simulations were performed using the spiking neural network
simulator Brian (Brian2, version 2.0, http://briansimulator.org/). We
attempted to stay as true as possible to the original study. Simulation
parameters were taken from the text and we additionally obtained the
source code for the standard parameter configuration from the authors to
ensure equivalency of the implementations. The source code for deleting
spikes within the pattern (Figure ~\ref{fig:para} E) was not provided.

\subsection{Running simulations}\label{running-simulations}

Brian calculates neuron properties such as membrane voltage at discrete
time steps. The time step for these simulations was 10\(^{-4}\) s unless
otherwise indicated. Each parameter combination was run 100 times and
the success of a run determined as in the original study: a hit rate of
over 98\%, no false alarms and an average latency of under 10 ms (as
calculated over the last 150 s of the simulation).

\hypertarget{tbl:param}{}
\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tbl:param}Parameter variations}\tabularnewline
\toprule
parameters & standard & variations\tabularnewline
\midrule
\endfirsthead
\toprule
parameters & standard & variations\tabularnewline
\midrule
\endhead
w\(_{initial}\) & 0.475 & 0.275, 0.325, 0.375, 0.425\tabularnewline
jitter (sd) {[}ms{]} & 1 & 0, 2, 3, 4, 5, 6\tabularnewline
pattern frequency & 0.25 & 0.05, 0.1, 0.15, 0.5\tabularnewline
prop. neurons in pattern & 0.5 & 0.3, 0.4, 0.6\tabularnewline
spike deletion & 0 & 0.1, 0.2, 0.3\tabularnewline
\bottomrule
\end{longtable}

\subsection{Spike trains}\label{spike-trains}

The spike trains of the 2000 input neurons were created in the same way
as in the original study: a Poisson process with a variable
instantaneous firing rate (30-90 Hz, 64 Hz on average) with no
refractory period.

The absence of a refractory period in the input neurons has the effect
inter spike intervals can be as small as 10\(^{-9}\) s in some cases.
The original implementation calculates the simulation variables (output
neuron voltage etc.) at every input neuron or output neuron spike, so
such small inter spike intervals are not a problem.

Brian on the other hand uses discrete time steps for its simulations and
updates its simulation variables once per time step. In order to convert
the original spike trains into Brian, we had to modify the spike trains
slightly: whenever two spikes from the same neuron happened in the same
time step, we deleted the second spike. At a resolution of 10\(^{-4}\) s
(standard time step) this affected 0.25\% of all spikes and 0.0025\% at
a resolution of 10\(^{-6}\) s. Deleting spikes does not affect the
amount of incoming voltage significantly since the initial firing rate
of the output neuron is the same with or without the close spikes.

\subsection{Leaky Integrate and Fire
neuron}\label{leaky-integrate-and-fire-neuron}

The original study models the potential of the output neuron with the
Gerstner's Spike Response model (SRM) \cite{Gerstner2002}, which uses
kernels to calculate the effect of incoming spikes on the postsynaptic
voltage. Brian on the other hand uses differential equations to model
the system parameters and evaluates those equations for each time step.
We converted the kernels of the posystnaptic potential and the spike
afterpotential into the following differential equations:

\begin{equation}\frac{du}{dt} = \frac{Xx - u}{\tau_{m}} + \frac{Aa}{\tau_{s}} \label{eq:u} \label{eq:1}\end{equation}

\begin{equation}\frac{dx}{dt} = -\frac{x}{\tau_{syn}}\label{eq:2}\end{equation}

\begin{equation}\frac{da}{dt} = -\frac{A}{\tau_{s}}\label{eq:3}\end{equation}

The values for the parameters can be found in Table ~\ref{tbl:param}.
Equation ~\ref{eq:1} describes the postsynaptic membrane potential,
which is influenced by presynaptic excitatory postsynaptic potentials
(EPSPs, first term) and a negative afterpotential after a spike (second
term). In the event of a presynaptic spike, x is set to 1, which
initiates the voltage increase in the postsynaptic neuron. When the
postsynaptic voltage reaches the threshold, a spike occurs: the voltage
u is set to twice the threshold, all EPSPs are flushed (all x set to 0)
and the negative afterpotential is set into motion (a is set to 1).
Presynaptic spikes had the same effect on the postsynaptic neuron
potential as in the original publication as shown in Figure
~\ref{fig:pot}.

\hypertarget{tbl:simt}{}
\begin{longtable}[]{@{}lrllr@{}}
\caption{\label{tbl:simt}Simulation parameters}\tabularnewline
\toprule
LIF neuron & value & \ldots{}\ldots{}\ldots{}\ldots{}\ldots{} & STDP &
value\tabularnewline
\midrule
\endfirsthead
\toprule
LIF neuron & value & \ldots{}\ldots{}\ldots{}\ldots{}\ldots{} & STDP &
value\tabularnewline
\midrule
\endhead
\(\tau_{m}\) {[}ms{]} & 10 & & \(\tau_{+}\) {[}ms{]} &
16.8\tabularnewline
\(\tau_{s}\) {[}ms{]} & 2.5 & & \(\tau_{-}\) {[}ms{]} &
33.7\tabularnewline
\(\tau_{syn}\) {[}ms{]} & 2.5 & & \(a_{+}\) & \(2^{-5}\)\tabularnewline
T {[}a.u.{]} & 500 & & \(a_{-}\) & 0.85 \(a_{+}\)\tabularnewline
X & \(\frac{\tau_{s}}{\tau_{m}}^{\frac{\tau_{m}}{\tau_{s} -\tau_{m}}}\)
& & \(w_{min}\) & 0\tabularnewline
A & -3 T & & \(w_{max}\) & 1\tabularnewline
\(\Delta x\) & 1 & & - & -\tabularnewline
\(\Delta a\) & 1 & & - & -\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics{figures/fig1_potentials.pdf}
\caption{Postsynaptic potential. The postsynaptic potential of the
replication implementation behaves the same as in the original study as.
Both show the same EPSP shapes and the negative afterpotential after a
postsynaptic spike. \textbf{A}) The postsynaptic potential of the
original study was calculated using the equations given in the original
paper. \textbf{B}) The postsynpatic potential in the replication
implementation was calculated from the differential equations specified
in section \textbf{Leaky Integrate and Fire neuron}}\label{fig:pot}
\end{figure}

\subsection{Spike Timing Dependent
Plasticity}\label{spike-timing-dependent-plasticity}

For synaptic plasticity, the original study uses the reduced nearest
neighbor rule (RNN) cite\{Morrison2008\} - a more restrictive version of
the nearest neighbor (NN) STDP rule. In the standard NN rule, every
spike causes a weight change (with the amount depending on the timing of
the nearest spike), while for RNN a weight change only happens for the
first postsynaptic spike immediately following a presynaptic spike (or
vice versa). This means that for NN, there can be a series of
potentiations or depressions, whereas for rNN potentiations and
depressions strictly alternate (due to neurons being forced to spike at
least once every 50 ms, see \cite{Masq2008}).

A visualisation of three STDP rules is shown in Figure ~\ref{fig:stdp}:
the RNN, the standard NN and the all-to-all (ATA) rule, in which all
pairs of spikes are considered to calculate the weight change. A case is
considered in which the presynaptic neuron (the input neurons in this
study) spikes more often than the postsynaptic neuron (the output neuron
in this study), leading to synaptic weight decreases for ATA and NN, but
no visible net change under RNN (Figure ~\ref{fig:stdp}\textbf{C}).

\begin{figure}
\centering
\includegraphics{figures/fig2_learningrules.pdf}
\caption{Effect of different learning rules on synaptic weight. We
compare the all-to-all (ATA, \textbf{A}), nearest neighbor (NN,
\textbf{B}), and reduced nearest neighbor (RNN, \textbf{C}) learning
rules for a case in which the presynaptic neuron (green dots) spikes
more often than the postsynaptic neuron (blue dots). Grey lines indicate
which spike pairs are taken into account for the calculation of the
synaptic weight change; the sign denotes a corresponding increase (+) or
decrease (-) in weight at that spike. The synaptic weight in ATA
(\textbf{A}) and NN (\textbf{B}) experience more weight changes and also
a net decrease in weight after only a few spikes. Under the RNN rule
(\textbf{C}), there are fewer changes and - in this example - no visible
net weight change.}\label{fig:stdp}
\end{figure}

\section{Results}\label{results}

The ability to find repeating patterns was reproduced in the replication
implementation. We could qualitatively reproduce all results of the
original paper and show the reproduced figures relevant to the model
behaviour: latency development (Figure ~\ref{fig:lat}), pattern
specificity after convergence (Figure ~\ref{fig:conv}), and robustness
(Figure ~\ref{fig:para}). In addition, we comment on some implementation
details that turned out to be relevant to replicating the original study
successfully: simulation time step, learning rule, EPSP shape.

\subsection{Pattern finding}\label{pattern-finding}

The latency measures the time of the output neuron spike relative to the
start of the pattern. If the spike occurs outside of the pattern
(\textgreater{}50 ms after the start of the pattern), the latency for
that spike is set to 0 as in the original paper.

In order to assess the time of pattern finding, we used the same spike
train (standard conditions) as input for both the original code and in
the replication implementation. The latency development looks similar in
both implementations (see Figure ~\ref{fig:lat}). The time until a
stable state arises is longer in the replication implementation (1400
discharges instead of 700 or 20 s instead of 13.5 s). This discrepancy
is due to the introduction of discrete time steps.

\begin{figure}
\centering
\includegraphics{figures/fig3_latency.pdf}
\caption{Latency. Comparing the latency development of the original
study (\textbf{A}) and the replication implementation (\textbf{B}). In
both implementations, the timing of the output spike in relation to the
start of the pattern is random at first, but becomes selective after a
few hundred discharges. From then on the postsynaptic neuron always
spikes only within the pattern (no spikes at latency = 0). The number of
discharges until this happens is smaller in the original paper than in
the replication implementation.}\label{fig:lat}
\end{figure}

At the end of the simulation, the synaptic weights that are maximally
potentiated (close to w\(_{max}\) = 1) belong exclusively to neurons
involved in the pattern, whereas weights from neurons not involved in
the pattern are depressed nearly completely (see Figure
~\ref{fig:conv}). Neurons that spike at the beginning of the pattern are
potentiated, causing the postsynaptic neuron to spike at a low latency.

\begin{figure}
\centering
\includegraphics{figures/fig4_convergence.pdf}
\caption{Weights at converged state. The weights in the converged state
look very similar in the the original study (\textbf{A)}) and the
replication implementation (\textbf{B)}). Weights from neurons not
involved in the pattern are close to 0 (neurons 1000-1999), whereas
weights from some neurons involved in the pattern (neurons 0-999) are
close to 1 (maximum value). The weights of neurons that spike at the
beginning of the pattern are nearly all close to 1 and their added
postsynaptic potentials cause a spike at the beginning of the pattern
(blue rectangle).}\label{fig:conv}
\end{figure}

Both latency development and the potentiation of only synaptic weights
of pattern neurons in the converged state were successfully reproduced
by the replication implementation.

\subsection{Robustness}\label{robustness}

Our simulations showed largely the same resilience to degradations as in
the original paper, but despite the very similar implementations, there
were some differences.

The replication implementation behaves the same as the original study
when subjected to different amounts of jitter (Figure ~\ref{fig:para}
\textbf{B}), various proportions of pattern neuron (Figure
~\ref{fig:para} \textbf{C}), different initial weights (Figure
~\ref{fig:para} \textbf{D}) and when a percentage of spikes within the
pattern are deleted (Figure ~\ref{fig:para} \textbf{E}).

\begin{figure}
\centering
\includegraphics{figures/fig5_robustness.pdf}
\caption{Robustness. Pattern finding abilities of the system were tested
against different pattern presentation frequencies (\textbf{A}), jitter
(\textbf{B}), number of neurons spiking according to the pattern
(\textbf{C}), initial weights (\textbf{D}) and deletion of spikes in the
pattern (\textbf{E}). For each parameter combination 100 simulations
were run and the number of successful runs are reported as percentage of
success. The replication implementation was run at two different time
steps: 10\(^{-6}\) s (solid black lines) and 10\(^{-4}\) s (dashed black
lines). The results from the original study were calculated using the
source code provided (solid green lines in \textbf{A}, \textbf{B},
\textbf{C}, \textbf{D}) or estimated from the original publication
(dashed green line in \textbf{E}).}\label{fig:para}
\end{figure}

At a high pattern repetition frequency of 0.5 (Figure ~\ref{fig:para}
\textbf{A}), when the pattern is presented every 100 ms for 50 ms, the
performance of the replication differs from the original: at larger time
steps of 10\(^{-4}\) s the replication version does not perform as well
as the original, but shows good results at smaller time steps
(10\(^{-6}\) s).

\begin{figure}
\centering
\includegraphics{figures/fig6_findtime.pdf}
\caption{Time until finding pattern. Using larger simulation time steps
leads to the pattern to be found later. The horizontal striped green
line is the reported time when the pattern is found by
\textcite{Masq2008}. Errorbars represent standard deviation from 100
successful runs.}\label{fig:find}
\end{figure}

\subsection{Implementation details}\label{implementation-details}

We noticed that small implementation details can affect the behaviour of
the network significantly. We summarised the relevant details in Table
~\ref{tbl:impl}\}.

\subsubsection{Simulation time step}\label{simulation-time-step}

In the standard parameter configuration, the time step chosen does not
have an effect on overall pattern finding success, as long as the time
step is shorter or equal to 10\(^{-4}\) s. At larger time steps
(10\(^{-3}\) s), no specificity emerges, instead there is a systematic
potentiation of all synaptic weights leading to very high firing rates
in the output neuron. In contrast, success rates are above 95\% for
10\(^{-4}\) s, 10\(^{-5}\) s, 10\(^{-6}\) s, and 10\(^{-7}\) s.

Although the success rate stays roughly the same, the time until the
pattern is found (no spikes outside of the pattern happen after this
time point) increases with larger time step size. An example can be seen
in Figure ~\ref{fig:lat}: the pattern is found after about 700 output
neuron spikes in the original publication (left) and after about 1400 in
the replication implementation (right) at a time step of 10\(^{-4}\) s
for the same input. At smaller times steps (10\(^{-6}\) s), the pattern
is found after about 700 discharges - the same as in the original
publication (Figure ~\ref{fig:find}).

\hypertarget{tbl:impl}{}
\begin{longtable}[]{@{}lrr@{}}
\caption{\label{tbl:impl}Implementation details}\tabularnewline
\toprule
& works & does not work\tabularnewline
\midrule
\endfirsthead
\toprule
& works & does not work\tabularnewline
\midrule
\endhead
simulation time step & continuous, discrete &\tabularnewline
time step size & \(\leq\) 10\(^{-4}\) s & \textgreater{} 10\(^{-4}\)
s\tabularnewline
learning rule & RNN & ATA, NN\tabularnewline
EPSP shape & kernel & immediate voltage increase\tabularnewline
\bottomrule
\end{longtable}

\subsubsection{Version of STDP learning
rule}\label{version-of-stdp-learning-rule}

The choice of learning rule is crucial for the pattern finding abilities
of the network. The RNN rule results in stable posystnaptic firing and
reliable finding of the pattern. The use of other learning rules does
not result in stable learning.

The spike trains used here involve neurons with a continuously high
firing rate in the input neurons (on average 64 Hz, min 30 Hz, max 90
Hz) and a significantly lower firing rate in the output neuron (63 Hz
initially, 5Hz after reaching specificity). This means that input
neurons fire often in the time between output neuron spikes. Therefore -
with any other than the reduced NN learning rule - the input neurons
will experience a decrease in weight a lot more often (every time an
input neuron spikes) than an increase in weight (every time the
postsynaptic neuron spikes). This leads to a strong overall depression
of the synaptic weights in the first few seconds of the simulation. With
the parameters specified in \textcite{Masq2008} and an ATA or a
conventional NN learning rule, the output neuron stops firing after a
few seconds because the output neuron voltage does not reach the voltage
threshold necessary to evoke an output neuron spike anymore (see Figure
~\ref{fig:meanw}\}). In the case of the standard parameters, the output
neuron stops firing after less than one second. When no output neuron
spikes occur, learning cannot take place and no specificity can emerge.

\begin{figure}
\centering
\includegraphics{figures/fig7_ata.pdf}
\caption{Effect of learning rules on synaptic weights. \textbf{A)} Both
the ATA and the NN rules lead to a very rapid depression of all synapses
leading to silence in the output neuron. \textbf{B)} In contrast, the
average weight of all neurons declines at a much slower rate for the RNN
rule. \textbf{C)} and \textbf{D)} When tweaking the ATA rule (increasing
a\(_{+}\) substantially) it is possible to achieve a behaviour that
resembles pattern finding. For a short amount of time the output neuron
becomes specific to the pattern, but loses this ability again and does
not regain it.}\label{fig:meanw}
\end{figure}

In the case of an ATA learning rule, a behaviour resembling pattern
finding can be evoked under the right circumstances: reducing the
learning rate by a factor of 5 and increasing the value of a\(_{+}\)
(the maximum weight increase) to double the value of a\(_{-}\) (the
maximum weight decrease). This setup works because the large amount of
depression (on every input spike) is counteracted by large potentiation
(due to the increased a\(_{+}\)). In such runs, the output neuron will
reach specificity and trace back through the latency and then starts
firing outside of the pattern again. This system is not stable since the
ATA rule leads to ``too much learning'' as the synaptic weights change
after every single spike. A further reduction in the learning rule will
not results in pattern specificity at all.

The standard NN learning rule, although exhibiting slower learning when
compared to the ATA rule, is unable to accomplish any pattern finding
because the limitation of potentiation to a certain maximum value means
that depression will always outweigh potentiation as it occurs far more
often.

\subsubsection{EPSP shape}\label{epsp-shape}

The kernels used in the original study simulate a gradual increase of
the postsynaptic neuron voltage as can be seen in Figure ~\ref{fig:pot}.
Other studies sometimes also model the effect of the presynaptic spike
as an immediate jump in postsynaptic voltage instead of that gradual
increase. In this system, using an immediate increase in postsynaptic
voltage does not lead to stable finding. This might have to to with the
fact that the kernel shape and the immediate increase exhibit slightly
different spike times as seen in Figure ~\ref{fig:epsps}.

\begin{figure}
\centering
\includegraphics{figures/fig8_kernel_vs_imm.pdf}
\caption{Choice of EPSP. For the same presynaptic spikes, the
postsynaptic voltage behaves similarly, but the postsynaptic spike time
is slightly different (green and black dots).}\label{fig:epsps}
\end{figure}

When modelling the immediate voltage increase, one needs to set the
height of the voltage increase (for Figure ~\ref{fig:epsps} \(\Delta u\)
was set to 1.2). It is very difficult to find the \(\Delta u\) that
corresponds to the same amount of voltage increase from the kernel. If
the value of \(\Delta u\) is too small, then the output neuron stops
firing after a short period of time without having gained specificity
for the pattern. If the value for \(\Delta u\) is too high all input
neurons are potentiated and the output firing rate rockets. For the
scope of this paper, no value for \(\Delta u\) was found to induce a
stable pattern finding behaviour.

\section{Conclusion}\label{conclusion}

We could successfully replicate the results from \textcite{Masq2008} - a
neuron with STDP synapses could reliably find a repeating spike pattern
and afterwards spike only when the pattern is presented. The time when
the pattern is found depends on the time step size chosen for the
simulation, whereas the success of reliably finding the pattern relies
on the precise learning rule and the shape of the EPSPs. Reproducing the
paper was made easy by the fact that simulation parameters were stated
in the text and the source code was shared by the authors of the paper
on request.

\textbf{Introducing discrete time steps.} To run the simulations in
Brian, we introduced discrete time steps. This did not affect pattern
finding abilities (as long as the time steps are smaller than
\(10^{-3}\) s), but it did increase the time until a pattern was found
for large time steps (\(10^{-4}\) s), but not for small time steps
(\(10^{-6}\) s).

We verified that the delay in finding did not stem from neither the
deletion of some input spikes (to avoid input neurons spiking twice in
one time step) nor the spike timing being at the start of each 0.1 ms
time bin. When we fed those modified input spikes in the source code,
timing of finding the pattern was not affected. The delay therefore
stems from forcing the output neuron to spike at the beginning of a time
step and the associated consequences for STDP learning.

The timing of the output neuron spikes is slightly different in the
original study and the replication. In the original study spike times of
the input and output neurons are unique, so no two neurons will spike at
the same time. In Brian, the output neuron reaches spikes at the
beginning of a time step and will therefore happen in the same time bin
as some input neuron spikes. Brian treats these input neuron spikes as
if happening just before the output neuron spike (due to scheduling of
events in Brian) and will increase all those weights instead of
increasing some and decreasing others. This higher number of
potentiations makes it more difficult for the system to systematically
depress unimportant weights in order to become selective to the pattern.
If the learning rule is modified so that the synaptic weight is
increased as well as depressed when input neuron and output neuron
spikes happen during the same time step (effectively causing a smaller
weight increase), finding the pattern occurs earlier, at around 17 s or
850 spikes which is close to the performance for smaller time steps and
the original paper (both 14 s or 700 spikes). This modification of the
learning rule is less successful at continuously spiking during the
pattern in longer simulations, it was therefore only used as a proof of
concept.

This difficulty to depress non-relevant input neurons is the reason for
the lower success rate at a high pattern presentation frequency at large
time steps (\(10^{-4}\) s) as seen in Figure ~\ref{fig:para}
\textbf{A}). The time until the pattern is found during this condition
is notably longer (\textgreater{}30 s instead of 20 s) and points
towards towards the difficulties of the system to properly depress the
synaptic weights of the non-pattern neurons.

Running the simulations using discrete time steps does not negatively
impact pattern finding abilities, but delays pattern finding for large
time steps due to a larger number of potentiations.

\textbf{Choice of learning rule.} The learning rule used is one specific
version of a NN STDP rule, which was not clearly stated in the original
article. The comparison with other learning rules shows that the use of
this particular learning rule enables the pattern finding behaviour.

The usage of the RNN rule has two interesting consequences. Firstly, it
slows down the rate of synaptic weight change, since it considers fewer
spike-spike interactions than other learning rules. This slows down
overall weight changes significantly and gives the system more time to
learn the spike sequences of the pattern. Neurons that spike together
during the pattern will experience similar weight changes more often
than non-pattern neurons. Over the course of hundreds of output neuron
spikes - until pattern specificity arises - these synchronised weight
changes lead to higher synaptic weights for the pattern neurons than for
the non-pattern neurons.

Secondly, for this particular input the restrictions of the RNN STDP
rule lead to the weights alternating between increase and decrease. The
stabilising effect of this is most clear during the converged phase. The
output neuron only spikes when a pattern is presented more or less at
the same latency. The input neurons (which are forced to spike at least
once every 50 ms) reliably spike at least once before or after or both.
In the latter case the input neurons will always experience the same
increase or decrease in synaptic weight per output neuron spike
disregarding small changes due to jitter and noise. If the input neuron
only spikes once per pattern, then either the increase or decrease of
weight will be nearly constant whereas the respective decrease or
increase in weight will be random. On average the weights of all input
neurons will slowly increase to 1 or decrease to 0 over time, resulting
in a very stable system.

\textbf{Choice of EPSP shape.} It seems to be essential to use a kernel
EPSP shape since using immediate voltage increases as the effect of an
input neuron spike does not lead to learning of the pattern. This might
be due to the difficulty in finding the correct parameters that create
an equivalent voltage increase. It seems likely that the EPSP shape is
responsible for the decrease in success, since small changes in output
neuron spike times also occur when using different time step sizes and
do not affect pattern finding abilities under standard conditions.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2018-05-22 article/build-with-docker.sh
Added working Dockerfile workflow

@@ -1,0 +1,9 @@
+#!/bin/sh
+docker run \
+  -v `pwd`/:/root/ \
+  pandoc \
+  --standalone --latex-engine=xelatex \
+  --template=rescience-template.tex \
+  --biblatex --bibliography=bibliography.bib \
+  -M "crossrefYaml=/crossref.yaml" \
+  --output Hathway-Goodman-2018.pdf Hathway-Goodman-2018.md

### 2018-05-27 article/Hathway-Goodman-2018.tex
Added final (?) versions of article, readme files, code

@@
       Pamela Hathway,
-      Dan Goodman,,
+      Dan F. M. Goodman,,
   },
@@
 \usepackage{fontspec}
-\defaultfontfeatures{Path = /usr/local/texlive/2018/texmf-dist/fonts/opentype/public/fontawesome/}
+%\defaultfontfeatures{Path = /usr/local/texlive/2018/texmf-dist/fonts/opentype/public/fontawesome/}
 \usepackage{fontawesome}
@@
                         \author[1]{Pamela Hathway}
-                    \author[1]{Dan Goodman,}
+                    \author[1]{Dan F. M. Goodman,}
                             \affil[1]{Department of Electrical and Electronic Engineering, Imperial College,
@@
 potential sequences (spike patterns) to convey information
-\cite{Fellous2004, Prut1998}. Since neuronal activity is noisy and
+\autocites{Fellous2004}{Prut1998}. Since neuronal activity is noisy and
 neurons are likely involved in a multitude of spike patterns of various
@@
 becomes due to the exponential increase of possible combinations of
-spikes that could make up a pattern \cite{Buonomano2009}. It is unclear
-how neurons in the brain may extract relevant information from such
-input. In a 2008 paper, Masquelier and colleagues demonstrated that a
-single neuron with spike timing dependent plasticity (STDP) synapses is
-able to find the start of a repeating pattern in noisy (artificial) data
-\cite{Masq2008}.
+spikes that could make up a pattern \autocite{Buonomano2009}. It is
+unclear how neurons in the brain may extract relevant information from
+such input. In a 2008 paper, Masquelier and colleagues demonstrated that
+a single neuron with afferent synapses exhibiting spike timing dependent
+plasticity (STDP) is able to find the start of a repeating pattern in
+noisy (artificial) data \autocite{Masq2008}.
 
@@
 of the input neurons spike according to a spike pattern of 50 ms length
-for about 25\% of the time. Finding the pattern is additionally made
-difficult by jittering the pattern spikes, adding random noise spikes to
-all neurons, and a constant population rate as well as no differences in
-overall firing rate between neurons.
+for about 25\% of the time. Finding the pattern is made more difficult
+by jittering the pattern spikes, adding random noise spikes to all
+neurons, and ensuring a constant population rate as well as no
+differences in overall firing rate between neurons.
 
 We replicated their findings using the spiking neural network simulator
-Brian \cite{Goodman2009, Stimberg2014}, whereas the the original study
-implemented the simulations in Matlab, with the main functions being
-computed in C/C++ through mex files. In addition, we examined some of
-the implementation parameters and investigated whether they were
-essential to the success of the algorithm.
+Brian \autocites{Goodman2009}{Stimberg2014}, whereas the the original
+study implemented the simulations in Matlab, with the main functions
+being computed in C/C++ through mex files. In addition, we examined some
+of the implementation details and parameters and investigated whether
+they were essential to the success of the algorithm.
 
@@
 ensure equivalency of the implementations. The source code for deleting
-spikes within the pattern (Figure ~\ref{fig:para} E) was not provided.
+spikes within the pattern (Fig.~\ref{fig:para} E) was not provided.
 
@@
 \begin{longtable}[]{@{}lrr@{}}
-\caption{\label{tbl:param}Parameter variations}\tabularnewline
+\caption{\label{tbl:param}Parameter variations }\tabularnewline
 \toprule
@@
 The absence of a refractory period in the input neurons has the effect
-inter spike intervals can be as small as 10\(^{-9}\) s in some cases.
-The original implementation calculates the simulation variables (output
-neuron voltage etc.) at every input neuron or output neuron spike, so
-such small inter spike intervals are not a problem.
+that inter spike intervals can be as small as 10\(^{-9}\) s in some
+cases. The original implementation uses an event-based simulation method
+in which the simulation variables (output neuron voltage etc.) are
+calculated at every input neuron or output neuron spike, and so such
+small inter spike intervals are not a problem.
 
@@
 time step, we deleted the second spike. At a resolution of 10\(^{-4}\) s
-(standard time step) this affected 0.25\% of all spikes and 0.0025\% at
-a resolution of 10\(^{-6}\) s. Deleting spikes does not affect the
-amount of incoming voltage significantly since the initial firing rate
-of the output neuron is the same with or without the close spikes.
+(standard time step) this affected 0.25\% of all spikes, and at a
+resolution of 10\(^{-6}\) s it affected 0.0025\%. Deleting spikes does
+not affect the input drive to the neuron significantly since the initial
+firing rate of the output neuron is the same with or without the close
+spikes.
 
@@
 The original study models the potential of the output neuron with the
-Gerstner's Spike Response model (SRM) \cite{Gerstner2002}, which uses
-kernels to calculate the effect of incoming spikes on the postsynaptic
-voltage. Brian on the other hand uses differential equations to model
-the system parameters and evaluates those equations for each time step.
-We converted the kernels of the posystnaptic potential and the spike
-afterpotential into the following differential equations:
+Gerstner's Spike Response model (SRM) \autocite{Gerstner2002}, which
+uses kernels to calculate the effect of incoming spikes on the
+postsynaptic voltage. Brian on the other hand uses differential
+equations to model the system parameters and evaluates those equations
+for each time step. We converted the kernels of the posystnaptic
+potential and the spike afterpotential into the following differential
+equations:
 
@@
 
-The values for the parameters can be found in Table ~\ref{tbl:param}.
-Equation ~\ref{eq:1} describes the postsynaptic membrane potential,
-which is influenced by presynaptic excitatory postsynaptic potentials
-(EPSPs, first term) and a negative afterpotential after a spike (second
-term). In the event of a presynaptic spike, x is set to 1, which
-initiates the voltage increase in the postsynaptic neuron. When the
-postsynaptic voltage reaches the threshold, a spike occurs: the voltage
-u is set to twice the threshold, all EPSPs are flushed (all x set to 0)
-and the negative afterpotential is set into motion (a is set to 1).
-Presynaptic spikes had the same effect on the postsynaptic neuron
-potential as in the original publication as shown in Figure
-~\ref{fig:pot}.
+The values for the parameters can be found in Tbl.~\ref{tbl:param}.
+Eq.~\ref{eq:1} describes the postsynaptic membrane potential, which is
+influenced by presynaptic excitatory postsynaptic potentials (EPSPs,
+first term) and a negative afterpotential after a spike (second term).
+In the event of a presynaptic spike, \(x\) is increased by 1
+(\(x\leftarrow x+1\)), which initiates the voltage increase in the
+postsynaptic neuron. When the postsynaptic voltage reaches the
+threshold, a spike occurs: the voltage \(u\) is set to twice the
+threshold (\(u\leftarrow 2T\)), all EPSPs are flushed (all \(x\) set to
+0, \(x\leftarrow 0\)) and the negative afterpotential is set into motion
+(\(a\) is set to 1, \(a\leftarrow 1\)). Presynaptic spikes had the same
+effect on the postsynaptic neuron potential as in the original
+publication as shown in Fig.~\ref{fig:pot}.
 
@@
 \begin{longtable}[]{@{}lrllr@{}}
-\caption{\label{tbl:simt}Simulation parameters}\tabularnewline
+\caption{\label{tbl:simt}Simulation parameters }\tabularnewline
 \toprule
@@
 \caption{Postsynaptic potential. The postsynaptic potential of the
-replication implementation behaves the same as in the original study as.
+replication implementation behaves the same as in the original study.
 Both show the same EPSP shapes and the negative afterpotential after a
@@
 original study was calculated using the equations given in the original
-paper. \textbf{B}) The postsynpatic potential in the replication
+paper. \textbf{B}) The postsynaptic potential in the replication
 implementation was calculated from the differential equations specified
@@
 For synaptic plasticity, the original study uses the reduced nearest
-neighbor rule (RNN) cite\{Morrison2008\} - a more restrictive version of
-the nearest neighbor (NN) STDP rule. In the standard NN rule, every
+neighbor rule (RNN) \autocite{Morrison2008} - a more restrictive version
+of the nearest neighbor (NN) STDP rule. In the standard NN rule, every
 spike causes a weight change (with the amount depending on the timing of
@@
 vice versa). This means that for NN, there can be a series of
-potentiations or depressions, whereas for rNN potentiations and
-depressions strictly alternate (due to neurons being forced to spike at
-least once every 50 ms, see \cite{Masq2008}).
+potentiations or depressions, whereas for RNN potentiations and
+depressions strictly alternate.
 
-A visualisation of three STDP rules is shown in Figure ~\ref{fig:stdp}:
-the RNN, the standard NN and the all-to-all (ATA) rule, in which all
-pairs of spikes are considered to calculate the weight change. A case is
+A visualisation of three STDP rules is shown in Fig.~\ref{fig:stdp}: the
+RNN, the standard NN and the all-to-all (ATA) rule, in which all pairs
+of spikes are considered to calculate the weight change. A case is
 considered in which the presynaptic neuron (the input neurons in this
@@
 in this study), leading to synaptic weight decreases for ATA and NN, but
-no visible net change under RNN (Figure ~\ref{fig:stdp}\textbf{C}).
+no visible net change under RNN (Fig.~\ref{fig:stdp} \textbf{C}).
 
@@
 original paper and show the reproduced figures relevant to the model
-behaviour: latency development (Figure ~\ref{fig:lat}), pattern
-specificity after convergence (Figure ~\ref{fig:conv}), and robustness
-(Figure ~\ref{fig:para}). In addition, we comment on some implementation
+behaviour: latency development (Fig.~\ref{fig:lat}), pattern specificity
+after convergence (Fig.~\ref{fig:conv}), and robustness
+(Fig.~\ref{fig:para}). In addition, we comment on some implementation
 details that turned out to be relevant to replicating the original study
@@
 the replication implementation. The latency development looks similar in
-both implementations (see Figure ~\ref{fig:lat}). The time until a
-stable state arises is longer in the replication implementation (1400
+both implementations (see Fig.~\ref{fig:lat}). The time until a stable
+state arises is longer in the replication implementation (1400
 discharges instead of 700 or 20 s instead of 13.5 s). This discrepancy
-is due to the introduction of discrete time steps.
+is due to the size of the discrete time steps used (see section
+\textbf{Implementation details}).
 
@@
 both implementations, the timing of the output spike in relation to the
-start of the pattern is random at first, but becomes selective after a
-few hundred discharges. From then on the postsynaptic neuron always
-spikes only within the pattern (no spikes at latency = 0). The number of
-discharges until this happens is smaller in the original paper than in
-the replication implementation.}\label{fig:lat}
+start of the pattern (vertical axis, set to latency=0 if output spike
+happens outside the pattern) is random at first, but becomes selective
+after a few hundred discharges. From then on the postsynaptic neuron
+always spikes only within the pattern (no spikes at latency = 0). The
+number of discharges until this happens is smaller in the original paper
+than in the replication implementation.}\label{fig:lat}
 \end{figure}
@@
 involved in the pattern, whereas weights from neurons not involved in
-the pattern are depressed nearly completely (see Figure
-~\ref{fig:conv}). Neurons that spike at the beginning of the pattern are
-potentiated, causing the postsynaptic neuron to spike at a low latency.
+the pattern are depressed nearly completely (see Fig.~\ref{fig:conv}).
+Neurons that spike at the beginning of the pattern are potentiated,
+causing the postsynaptic neuron to spike at a low latency.
 
@@
 replication implementation (\textbf{B)}). Weights from neurons not
-involved in the pattern are close to 0 (neurons 1000-1999), whereas
-weights from some neurons involved in the pattern (neurons 0-999) are
-close to 1 (maximum value). The weights of neurons that spike at the
-beginning of the pattern are nearly all close to 1 and their added
-postsynaptic potentials cause a spike at the beginning of the pattern
-(blue rectangle).}\label{fig:conv}
+involved in the pattern are close to 0 (black, neurons 1000-1999),
+whereas weights from some neurons involved in the pattern (neurons
+0-999) are close to 1 (white, maximum value). The weights of neurons
+that spike at the beginning of the pattern are nearly all close to 1 and
+their added postsynaptic potentials cause a spike at the beginning of
+the pattern (blue rectangle).}\label{fig:conv}
 \end{figure}
@@
 The replication implementation behaves the same as the original study
-when subjected to different amounts of jitter (Figure ~\ref{fig:para}
-\textbf{B}), various proportions of pattern neuron (Figure
-~\ref{fig:para} \textbf{C}), different initial weights (Figure
-~\ref{fig:para} \textbf{D}) and when a percentage of spikes within the
-pattern are deleted (Figure ~\ref{fig:para} \textbf{E}).
+when subjected to different amounts of jitter (Fig.~\ref{fig:para}
+\textbf{B}), various proportions of neurons involved in the pattern
+(Fig.~\ref{fig:para} \textbf{C}), different initial weights
+(Fig.~\ref{fig:para} \textbf{D}) and when a percentage of spikes within
+the pattern are deleted (Fig.~\ref{fig:para} \textbf{E}).
 
@@
 success. The replication implementation was run at two different time
-steps: 10\(^{-6}\) s (solid black lines) and 10\(^{-4}\) s (dashed black
+steps: 10\(^{-4}\) s (solid black lines) and 10\(^{-6}\) s (dashed black
 lines). The results from the original study were calculated using the
@@
 
-At a high pattern repetition frequency of 0.5 (Figure ~\ref{fig:para}
+At a high pattern repetition frequency of 0.5 (Fig.~\ref{fig:para}
 \textbf{A}), when the pattern is presented every 100 ms for 50 ms, the
@@
 the network significantly. We summarised the relevant details in Table
-~\ref{tbl:impl}\}.
+tbl.~\ref{tbl:impl}.
 
@@
 have an effect on overall pattern finding success, as long as the time
-step is shorter or equal to 10\(^{-4}\) s. At larger time steps
-(10\(^{-3}\) s), no specificity emerges, instead there is a systematic
+step is less than or equal to 10\(^{-4}\) s. At larger time steps
+(10\(^{-3}\) s), no specificity emerges. Instead there is a systematic
 potentiation of all synaptic weights leading to very high firing rates
@@
 Although the success rate stays roughly the same, the time until the
-pattern is found (no spikes outside of the pattern happen after this
-time point) increases with larger time step size. An example can be seen
-in Figure ~\ref{fig:lat}: the pattern is found after about 700 output
-neuron spikes in the original publication (left) and after about 1400 in
-the replication implementation (right) at a time step of 10\(^{-4}\) s
-for the same input. At smaller times steps (10\(^{-6}\) s), the pattern
-is found after about 700 discharges - the same as in the original
-publication (Figure ~\ref{fig:find}).
+pattern is found (defined as the time after which no output spikes
+happen outside of the pattern) increases with larger time step size. An
+example can be seen in Fig.~\ref{fig:lat}: the pattern is found after
+about 700 output neuron spikes in the original publication (left) and
+after about 1400 in the replication implementation (right) at a time
+step of 10\(^{-4}\) s for the same input. At smaller times steps
+(10\(^{-6}\) s), the pattern is found after about 700 discharges - the
+same as in the original publication (Fig.~\ref{fig:find}).
 
@@
 \begin{longtable}[]{@{}lrr@{}}
-\caption{\label{tbl:impl}Implementation details}\tabularnewline
+\caption{\label{tbl:impl}Implementation details }\tabularnewline
 \toprule
@@
 initially, 5Hz after reaching specificity). This means that input
-neurons fire often in the time between output neuron spikes. Therefore -
-with any other than the reduced NN learning rule - the input neurons
-will experience a decrease in weight a lot more often (every time an
-input neuron spikes) than an increase in weight (every time the
-postsynaptic neuron spikes). This leads to a strong overall depression
-of the synaptic weights in the first few seconds of the simulation. With
-the parameters specified in \textcite{Masq2008} and an ATA or a
-conventional NN learning rule, the output neuron stops firing after a
-few seconds because the output neuron voltage does not reach the voltage
-threshold necessary to evoke an output neuron spike anymore (see Figure
-~\ref{fig:meanw}\}). In the case of the standard parameters, the output
-neuron stops firing after less than one second. When no output neuron
-spikes occur, learning cannot take place and no specificity can emerge.
+neurons fire often in the time between output neuron spikes. Therefore,
+with any other than the reduced NN learning rule, the input neurons will
+experience a decrease in weight a lot more often (every time an input
+neuron spikes) than an increase in weight (every time the postsynaptic
+neuron spikes). This leads to a strong overall depression of the
+synaptic weights in the first few seconds of the simulation. With the
+parameters specified in \textcite{Masq2008} and an ATA or a conventional
+NN learning rule, the output neuron stops firing after a few seconds
+because the output neuron voltage does not reach the voltage threshold
+necessary to evoke an output neuron spike anymore (see
+Fig.~\ref{fig:meanw}). In the case of the standard parameters, the
+output neuron stops firing after less than one second. When no output
+neuron spikes occur, learning cannot take place and no specificity can
+emerge.
 
@@
 rule. \textbf{C)} and \textbf{D)} When tweaking the ATA rule (increasing
-a\(_{+}\) substantially) it is possible to achieve a behaviour that
+\(a_{+}\) substantially) it is possible to achieve a behaviour that
 resembles pattern finding. For a short amount of time the output neuron
@@
 finding can be evoked under the right circumstances: reducing the
-learning rate by a factor of 5 and increasing the value of a\(_{+}\)
-(the maximum weight increase) to double the value of a\(_{-}\) (the
+learning rate by a factor of 5 and increasing the value of \(a_{+}\)
+(the maximum weight increase) to double the value of \(a_{-}\) (the
 maximum weight decrease). This setup works because the large amount of
 depression (on every input spike) is counteracted by large potentiation
-(due to the increased a\(_{+}\)). In such runs, the output neuron will
-reach specificity and trace back through the latency and then starts
-firing outside of the pattern again. This system is not stable since the
-ATA rule leads to ``too much learning'' as the synaptic weights change
-after every single spike. A further reduction in the learning rule will
-not results in pattern specificity at all.
+(due to the increased \(a_{+}\)). In such runs, the output neuron will
+correctly reach specificity and trace back through the latency, but then
+starts firing outside of the pattern again. This system is not stable
+since the ATA rule leads to ``too much learning'' as the synaptic
+weights change after every single spike. A further reduction in the
+learning rule will not result in pattern specificity at all.
 
-The standard NN learning rule, although exhibiting slower learning when
-compared to the ATA rule, is unable to accomplish any pattern finding
-because the limitation of potentiation to a certain maximum value means
-that depression will always outweigh potentiation as it occurs far more
-often.
+We were unable to find parameters for the standard NN learning rule that
+allowed for pattern finding, despite this rule exhibiting slower
+learning when compared to the ATA rule. In the case of both the ATA and
+NN learning rules, it seems likely that stable pattern finding relies on
+a precise balance of \(a_{+}\) and \(a_{-}\), with runaway potentiation
+or depression likely if the balance is wrong. By contrast, the RNN rule
+is automatically balanced and is not subject to this issue.
+
+\subsubsection{\texorpdfstring{Effect of learning rule at
+\(\Delta t=0\)}{Effect of learning rule at \textbackslash{}Delta t=0}}\label{effect-of-learning-rule-at-delta-t0}
+
+The timing of the output neuron spikes is slightly different in the
+original study and the replication. In the original study spike times of
+the input and output neurons are not restricted to fixed multiples of
+the timestep, so it is extremely unlikely that two neurons will spike at
+the same time. In Brian, the output neuron spikes at the beginning of a
+time step and will therefore happen in the same time bin as some input
+neuron spikes leading to a time difference between the spikes of
+\(\Delta t=0\) where the STDP rule is undefined. Brian treats these
+input neuron spikes as if they happened just before the output neuron
+spike (\(\Delta t<0\), due to the scheduling of events in Brian) and
+will therefore increase all those weights instead of increasing some and
+decreasing others. This higher number of potentiations makes it more
+difficult for the system to systematically depress unimportant weights
+in order to become selective to the pattern. If the learning rule is
+modified so that the synaptic weight is increased as well as depressed
+when input neuron and output neuron spikes happen during the same time
+step (effectively causing a smaller weight increase), the pattern is
+found earlier, at around 17 s or 850 spikes (for a time step of
+\(10^{-4}s\)) which is close to the performance for smaller time steps
+(\(10^{-6}s\)) and the original paper (both 14 s or 700 spikes). We
+didn't use this modified learning rule in the rest of the paper because
+it differs from the implementation used in the majority of modelling
+papers.
+
+This difficulty to depress non-relevant input neurons is the reason for
+the lower success rate at a high pattern presentation frequency at large
+time steps (\(10^{-4}\) s) as seen in Fig.~\ref{fig:para} \textbf{A}).
+The time until the pattern is found during this condition is notably
+longer (\textgreater{}30 s instead of 20 s) and points towards towards
+the difficulties of the system to properly depress the synaptic weights
+of the non-pattern neurons.
 
@@
 The kernels used in the original study simulate a gradual increase of
-the postsynaptic neuron voltage as can be seen in Figure ~\ref{fig:pot}.
+the postsynaptic neuron voltage as can be seen in Fig.~\ref{fig:pot}.
 Other studies sometimes also model the effect of the presynaptic spike
@@
 increase. In this system, using an immediate increase in postsynaptic
-voltage does not lead to stable finding. This might have to to with the
-fact that the kernel shape and the immediate increase exhibit slightly
-different spike times as seen in Figure ~\ref{fig:epsps}.
+voltage does not lead to stable pattern finding. This might have to to
+with the fact that the kernel shape and the immediate increase exhibit
+slightly different spike times as seen in Fig.~\ref{fig:epsps}.
 
@@
 When modelling the immediate voltage increase, one needs to set the
-height of the voltage increase (for Figure ~\ref{fig:epsps} \(\Delta u\)
+magnitude of the voltage increase (for Fig.~\ref{fig:epsps} \(\Delta u\)
 was set to 1.2). It is very difficult to find the \(\Delta u\) that
@@
 the pattern is found depends on the time step size chosen for the
-simulation, whereas the success of reliably finding the pattern relies
-on the precise learning rule and the shape of the EPSPs. Reproducing the
+simulation, whereas the success of reliably finding the pattern requires
+the precise learning rule and the shape of the EPSPs. Reproducing the
 paper was made easy by the fact that simulation parameters were stated
@@
 Brian, we introduced discrete time steps. This did not affect pattern
-finding abilities (as long as the time steps are smaller than
-\(10^{-3}\) s), but it did increase the time until a pattern was found
-for large time steps (\(10^{-4}\) s), but not for small time steps
+finding abilities (as long as the time steps are \(10^{-4}\) s or
+smaller), but it did increase the time until a pattern was found for
+large time steps (\(10^{-4}\) s), but not for small time steps
 (\(10^{-6}\) s).
 
-We verified that the delay in finding did not stem from neither the
-deletion of some input spikes (to avoid input neurons spiking twice in
-one time step) nor the spike timing being at the start of each 0.1 ms
-time bin. When we fed those modified input spikes in the source code,
-timing of finding the pattern was not affected. The delay therefore
-stems from forcing the output neuron to spike at the beginning of a time
-step and the associated consequences for STDP learning.
-
-The timing of the output neuron spikes is slightly different in the
-original study and the replication. In the original study spike times of
-the input and output neurons are unique, so no two neurons will spike at
-the same time. In Brian, the output neuron reaches spikes at the
-beginning of a time step and will therefore happen in the same time bin
-as some input neuron spikes. Brian treats these input neuron spikes as
-if happening just before the output neuron spike (due to scheduling of
-events in Brian) and will increase all those weights instead of
-increasing some and decreasing others. This higher number of
-potentiations makes it more difficult for the system to systematically
-depress unimportant weights in order to become selective to the pattern.
-If the learning rule is modified so that the synaptic weight is
-increased as well as depressed when input neuron and output neuron
-spikes happen during the same time step (effectively causing a smaller
-weight increase), finding the pattern occurs earlier, at around 17 s or
-850 spikes which is close to the performance for smaller time steps and
-the original paper (both 14 s or 700 spikes). This modification of the
-learning rule is less successful at continuously spiking during the
-pattern in longer simulations, it was therefore only used as a proof of
-concept.
-
-This difficulty to depress non-relevant input neurons is the reason for
-the lower success rate at a high pattern presentation frequency at large
-time steps (\(10^{-4}\) s) as seen in Figure ~\ref{fig:para}
-\textbf{A}). The time until the pattern is found during this condition
-is notably longer (\textgreater{}30 s instead of 20 s) and points
-towards towards the difficulties of the system to properly depress the
-synaptic weights of the non-pattern neurons.
+We verified that the delay in pattern finding did not stem from either
+the deletion of some input spikes (to avoid input neurons spiking twice
+in one time step) or the spike timing being at the start of each 0.1 ms
+time bin. When we fed those modified input spikes to the original
+implementation, the timing of finding the pattern was not affected. The
+delay therefore stems from forcing the output neuron to spike at the
+beginning of a time step and the associated consequences for STDP
+learning.
 

### 2018-05-27 code/cluster/main_cluster.py

@@ -1,0 +1,406 @@
+from brian2 import *
+import time as time
+import argparse
+import pickle
+from numba import jit
+import numpy
+
+set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone')
+
+argparser = argparse.ArgumentParser()
+argparser.add_argument("runid")
+args = argparser.parse_args()
+runid = args.runid
+rest, numbr = runid.split('=')
+seedyy = int(numbr)
+savedir = '/home/ph416/pf/res'
+para = load('/home/ph416/pf/para.npy')
+
+seedy = mod(seedyy, 100)  # get random seed for simulation, keep it between 1 and 100 as for other simulations
+pararow = seedyy // 100
+
+win = para[pararow, 1]  # 1.9*T/2000
+jitter_sd = para[pararow, 2]
+number_pat = int(para[pararow, 3])
+pattern_freq = para[pararow, 4]
+spike_del = para[pararow, 5]
+T = para[pararow, 6]  # 0.5*(1-spike_del)*number_pat
+K2 = 3  # 4
+A = - K2 * T  # 4 * T  # 1500
+refract = 1
+
+name_run = 'newrule_f50'
+defaultclock.dt = 1e-4 * second  # length of time step for simulation
+runduration = 450  # length of simulation [second], full length = 450
+tripling = True
+now = datetime.datetime.now()
+
+result_onerun = {}
+result_onerun['name'] = name_run
+result_onerun['runid'] = runid
+result_onerun['seedyy'] = seedyy
+result_onerun['win'] = win
+result_onerun['jitter'] = jitter_sd
+result_onerun['n_pat'] = number_pat
+result_onerun['freq_pat'] = pattern_freq
+result_onerun['del'] = spike_del
+result_onerun['threshold'] = T
+result_onerun['runduration'] = runduration
+result_onerun['date'] = now.strftime('%y%m%d')
+result_onerun['t_res'] = defaultclock.dt / second
+result_onerun['K2'] = K2
+result_onerun['A'] = A
+result_onerun['refract_N2'] = refract
+
+with open('%s/result_seed%s_%s.pickle' % (savedir, runid, name_run), 'wb') as handle:
+    pickle.dump(result_onerun, handle, protocol=pickle.HIGHEST_PROTOCOL)
+
+
+@jit(nopython=True) # numba decorator compiles this function into low level code to run faster
+def make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern):
[…]

### 2018-05-27 code/cluster/run_main_cluster.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+#PBS -l walltime=05:00:00
+#PBS -l select=1:ncpus=1:mem=10gb
+#PBS -N rescience
+
+module load anaconda3/personal
+
+python /THIS/PATH/NEEDS/TO/BE/CHANGED/run_main_cluster.py runid=$PBS_ARRAY_INDEX

### 2018-05-27 code/create_input.py

@@ -2,26 +2,6 @@ from brian2 import *
 from numba import jit
 import numpy
 
-if __name__=='__main__':
-	import time
-	def timeit(method):
-		def timed(*args, **kw):
-			ts = time.time()
-			result = method(*args, **kw)
-			te = time.time()
-
-			if 'log_time' in kw:
-				name = kw.get('log_name', method.__name__.upper())
-				kw['log_time'][name] = int((te - ts) * 1000)
-			else:
-				print('%r  %2.2f ms' %
-					  (method.__name__, (te - ts) * 1000))
-			return result
-
-		return timed
-else:
-	def timeit(method):
-		return method
 
 @jit(nopython=True) # numba decorator compiles this function into low level code to run faster
 def make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern):
@@ -47,7 +27,6 @@ def make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, r
 	return array(st)
 
 
-@timeit
 def make_input(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern):
 	# make input spikes
 	spiketimes = []
@@ -65,7 +44,6 @@ def make_input(min_rate, max_rate, max_time_wo_spike, max_change_speed, rundurat
 	return afferents, spiketimes
 
 
-@timeit
 def make_pattern_presentation_array(runduration, patternlength, pattern_freq):
 	# make array with a 1 (pattern) or 0 (not pattern) for each time window (window size = length of pattern (0.05s))
 	runduration1 = min(runduration, 150)
@@ -87,7 +65,6 @@ def make_pattern_presentation_array(runduration, patternlength, pattern_freq):
 	return position_copypaste
 
 
-@timeit
 def copy_and_paste_jittered_pattern(times, indices, position_copypaste, patternlength, jitter_sd, spike_del, number_pat):
 	# choose first segment as pattern to be copy-pasted
 	startCPindex = where(position_copypaste == 1)[0][0]
@@ -152,7 +129,6 @@ def copy_and_paste_jittered_pattern(times, indices, position_copypaste, patternl
 	return times1, indices1
 
 
-@timeit
 def add_noise(times, indices, times_add, indices_add):
 	# combine the basic activity and the 10Hz additional noise to one input
 	times = concatenate((times, times_add))
[…]

### 2018-05-27 code/figure_1.py

@@ -1,70 +1,86 @@
 from brian2 import *
-
-plt.rcParams['savefig.dpi'] = 75
-plt.rcParams['figure.autolayout'] = False
-plt.rcParams['figure.figsize'] = 8, 5
-plt.rcParams['axes.labelsize'] = 17
-plt.rcParams['axes.titlesize'] = 17
-plt.rcParams['font.size'] = 17
-plt.rcParams['lines.linewidth'] = 2.0
-plt.rcParams['lines.markersize'] = 3
-plt.rcParams['legend.fontsize'] = 16
+import time
 
 
-runduration = 0.08
-defaultclock.dt = 1e-4 * second
+def fig_1():
+    print('%s Preparing Figure 1: Potentials' % time.strftime('%H:%M'))
+    plt.rcParams['savefig.dpi'] = 75
+    plt.rcParams['figure.autolayout'] = False
+    plt.rcParams['figure.figsize'] = 6, 5
+    plt.rcParams['axes.labelsize'] = 17
+    plt.rcParams['axes.titlesize'] = 17
+    plt.rcParams['font.size'] = 17
+    plt.rcParams['lines.linewidth'] = 2.0
+    plt.rcParams['lines.markersize'] = 3
+    plt.rcParams['legend.fontsize'] = 16
 
-# output neuron
-win = 1  # initial weight
-taum = 10 * ms
-taus = 2.5 * ms
-tausyn = 2.5 * ms
-T = 2.8
-u_rest = 0
-X = (taus/taum)**(taum/(taus - taum))
-A = -8.5
-deltax = 1
-deltaa = 1
+    device.reinit()
+    device.activate()
 
-# equations
-eqs = '''du/dt = (A*a)/taus + (X*x-u)/taum : 1
-            dx/dt = -x/tausyn : 1 
-            da/dt = -a/taus : 1'''
-eqs_reset = '''x = 0
-            u = 2*T
-            a = deltaa'''
-stdp_model = ''' wi : 1'''
-stdp_on_pre = ''' x_post += deltax*wi '''
+    runduration = 0.08
+    defaultclock.dt = 1e-4 * second
 
-# input
-times = array([5, 20, 40, 43, 46, 60]) * ms
-indices = zeros(len(times))
+    # output neuron
+    win = 1  # initial weight
+    taum = 10 * ms
+    taus = 2.5 * ms
+    tausyn = 2.5 * ms
+    T = 2.8
+    u_rest = 0
[…]

### 2018-05-27 code/figure_2.py

@@ -1,112 +1,127 @@
 from brian2 import *
-
-plt.rcParams['savefig.dpi'] = 75
-plt.rcParams['figure.autolayout'] = False
-plt.rcParams['figure.figsize'] = 12, 5
-plt.rcParams['axes.labelsize'] = 17
-plt.rcParams['axes.titlesize'] = 17
-plt.rcParams['font.size'] = 17
-plt.rcParams['lines.linewidth'] = 2.0
-plt.rcParams['lines.markersize'] = 3
-plt.rcParams['legend.fontsize'] = 16
-
-# spike times
-times_pre = array([2, 4, 6, 14, 16, 18, 22, 24, 26])*ms
-indices_pre = zeros(len(times_pre))
-times_post = array([8, 12, 20])*ms
-indices_post = zeros(len(times_post))
-factor = array([3])  # how long the interval between spikes should be
-
-# parameters STDP synapses
-tauplus = 16.8 * ms
-tauminus = 33.7 * ms
-aplus = 2.0 ** -5  # 0.03125# 0.0008#0.03125
-aminus = 0.85 * aplus
-defaultclock.dt = 1e-5 * second  # length of time step for simulation
-
-# equations synapses
-syn_eqs = ''' wi : 1
-dLTPtrace/dt = -LTPtrace / tauplus : 1 (clock-driven)
-dLTDtrace/dt = -LTDtrace / tauminus : 1 (clock-driven)'''
-
-# equations ATA rule
-synATA_pre = '''LTPtrace += aplus
-wi = clip(wi + LTDtrace, 0, 1)'''
-
-synATA_post = ''' LTDtrace -= aminus
-wi = clip(wi + LTPtrace, 0, 1)'''
-
-# equations NN rule
-synNSme_pre = '''LTPtrace = aplus
-wi = clip(wi + LTDtrace, 0, 1)'''
-
-synNSme_post = ''' LTDtrace = -aminus
-wi = clip(wi + LTPtrace, 0, 1)'''
-
-# equations RNN
-synRNN_pre = '''LTPtrace = aplus
-wi = clip(wi + LTDtrace, 0, 1)
-LTDtrace = 0'''
-
-synRNN_post = ''' LTDtrace = -aminus
-wi = clip(wi + LTPtrace, 0, 1)
-LTPtrace = 0'''
+import time
 
 
-Nin = SpikeGeneratorGroup(1, indices_pre, times_pre * factor)
[…]

### 2018-05-27 code/figure_3_4_7AB.py

@@ -1,0 +1,109 @@
+from brian2 import *
+
+
+def fig_3(N2spm, latency):
+	plt.rcParams['savefig.dpi'] = 75
+	plt.rcParams['figure.autolayout'] = False
+	plt.rcParams['figure.figsize'] = 6, 5
+	plt.rcParams['axes.labelsize'] = 17
+	plt.rcParams['axes.titlesize'] = 17
+	plt.rcParams['font.size'] = 17
+	plt.rcParams['lines.linewidth'] = 2.0
+	plt.rcParams['lines.markersize'] = 8
+	plt.rcParams['legend.fontsize'] = 16
+	fig3 = figure()
+	suptitle('Figure 3')
+	title('latency vs output neuron spikes')
+	plot(range(len(latency)), latency * 1000, 'g.', markersize=3)
+	ylabel('Postsynaptic spike latency [ms]')
+	xlabel('# discharge')
+	yticks([0, 10, 20, 30, 40, 50])
+	ylim([-2, 52])
+
+	tight_layout()
+
+	savefig('../article/figures/figure_3_created_%s.pdf' % datetime.datetime.now().strftime('%y%m%d'))
+
+	return fig3
+
+
+def fig_4(N1spm, syn12, timing_pattern):
+	plt.rcParams['savefig.dpi'] = 75
+	plt.rcParams['figure.autolayout'] = False
+	plt.rcParams['figure.figsize'] = 12, 5
+	plt.rcParams['axes.labelsize'] = 17
+	plt.rcParams['axes.titlesize'] = 17
+	plt.rcParams['font.size'] = 17
+	plt.rcParams['lines.linewidth'] = 2.0
+	plt.rcParams['lines.markersize'] = 8
+	plt.rcParams['legend.fontsize'] = 16
+	# get spikes from end of simulation
+	figsta = searchsorted(N1spm.t /second, timing_pattern[-2] - 0.025)
+	figsto = searchsorted(N1spm.t /second, timing_pattern[-2] + 0.075)
+	greyness = [str((255 - syn12.wi[int(i)]) / 255.) for i in N1spm.i[figsta:figsto]]  # high w == white
+
+	ax = figure(figsize=(8, 6)).add_subplot(1, 1, 1)
+	ax.set_facecolor("k")
[…]

### 2018-05-27 code/figure_5.py

@@ -1,0 +1,218 @@
+from brian2 import *
+import matplotlib.pyplot as plt
+import pickle as pickle
+import time
+
+
+
+def fig_5_saved():
+
+    import pickle as pickle
+    print('%s Preparing Figure 5: Robustness' % time.strftime('%H:%M'))
+    plt.rcParams['savefig.dpi'] = 75
+    plt.rcParams['figure.autolayout'] = False
+    plt.rcParams['figure.figsize'] = 8, 5
+    plt.rcParams['axes.labelsize'] = 17
+    plt.rcParams['axes.titlesize'] = 17
+    plt.rcParams['font.size'] = 17
+    plt.rcParams['lines.linewidth'] = 2.0
+    plt.rcParams['lines.markersize'] = 3
+    plt.rcParams['legend.fontsize'] = 13
+
+    # ### load parameter arrays and original paper results
+    with open('../data/results_masquelier2008.pickle', 'rb') as handle:
+            results_masq = pickle.load(handle)
+
+    # ### load results from different parameter conditions
+    runs_to_plot = ['e4', 'e6']
+    names_to_plot = ['rep 10$^{-4}$', 'rep 10$^{-6}$']
+    styles = ['-', '--']
+    colors = ['k', 'k']
+
+    fig = figure(figsize=(12, 4))
+    suptitle('Figure 5 saved results')
+
+    for idx, run_name in enumerate(runs_to_plot):
+
+        run_ids = arange(0, 22, 1)
+        success_rates = empty(22)
+
+        # save all success results in one array
+        for run_idx in run_ids:
+
+            with open('../data/results_saved_%s_%s.pickle' % (run_idx, run_name), 'rb') as handle:
+                run = pickle.load(handle)
+
+            success_rates[run_idx] = run['success_overall']
+
+        # make separate arrays for each condition and add result from standard condition (0)
+        win_res = concatenate((success_rates[1:5], array([success_rates[0]])))
+        jit_res = concatenate((array([success_rates[5]]), array([success_rates[0]]), success_rates[6:11]))
+        npat_res = concatenate((success_rates[11:14], array([success_rates[0]]), array([success_rates[14]])))
+        freq_res = concatenate((success_rates[15:18], array([success_rates[0]]), array([success_rates[18]])))
[…]

### 2018-05-27 code/figure_6.py

@@ -1,0 +1,260 @@
+from brian2 import *
+import pickle as pickle
+
+
+def fig_6_saved():
+    print('%s Preparing Figure 6: Time of finding pattern' % time.strftime('%H:%M'))
+    plt.rcParams['savefig.dpi'] = 75
+    plt.rcParams['figure.autolayout'] = False
+    plt.rcParams['figure.figsize'] = 6, 5
+    plt.rcParams['axes.labelsize'] = 17
+    plt.rcParams['axes.titlesize'] = 17
+    plt.rcParams['font.size'] = 17
+    plt.rcParams['lines.linewidth'] = 2.0
+    plt.rcParams['lines.markersize'] = 3
+    plt.rcParams['legend.fontsize'] = 16
+
+    resolutions = [10**(-4), 10**(-5), 10**(-6)]
+
+    # ### load arrays from saved runs
+    with open('../data/results_saved_0_e4.pickle', 'rb') as handle:
+            run = pickle.load(handle)
+    # only load runs that were successful in finding
+    find_time_e4 = run['find_t'][run['success'] > 0]
+
+    with open('../data/results_saved_0_e5.pickle', 'rb') as handle:
+        run = pickle.load(handle)
+    # only load runs that were successful in finding
+    find_time_e5 = run['find_t'][run['success'] > 0]
+
+    with open('../data/results_saved_0_e6.pickle', 'rb') as handle:
+            run = pickle.load(handle)
+    # only load runs that were successful in finding
+    find_time_e6 = run['find_t'][run['success'] > 0]
+
+    # ### plot time finding
+    fig = figure(figsize=(6, 5))
+    title('Figure 6 saved results')
+    errorbar(resolutions, [mean(find_time_e4), mean(find_time_e5), mean(find_time_e6)], [std(find_time_e4), std(find_time_e5), std(find_time_e6)], color='k')
+    ylabel('pattern finding time [s]')
+    xlabel('simulation time step [s]')
+    semilogx(10**(-4), 10**(-6), 'k-')
+    xlim([0.0000006, 0.00015])
+    ylim([0, 30])
+    # yticks(array([12, 14, 16, 18, 20, 22]))
+    axhline(y=13.5, linestyle='--', color='g')
+    tight_layout()
+
+    savefig('../article/figures/figure_6_from_saved_created_%s.pdf' % datetime.datetime.now().strftime('%y%m%d'))
+
+    return fig
+
+
[…]

### 2018-05-27 code/figure_7CD.py

@@ -1,0 +1,173 @@
+from brian2 import *
+import time as time
+
+
+def fig_7CD(run_idx):
+	from create_input import make_input, make_pattern_presentation_array, \
+		copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
+
+	device.reinit()
+	device.activate()
+	set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone_ata3')
+	print('%s Preparing Figure 7 C and D: success with ATA rule' % time.strftime('%H:%M'))
+
+	# parameters that were varied for the publication
+	random_seed = run_idx
+	defaultclock.dt = 1e-4 * second  # length of time step for simulation
+
+	jitter_sd = 0  # amount of jitter
+	number_pat = 1000  # number of neurons that take part in the pattern presentation
+	number_neurons = 2000
+	pattern_freq = 0.25  # frequency at which the pattern is presented
+	spike_del = 0  # percentage of spikes within the pattern to be deleted
+	T = 500  # threshold
+	win = 0.2375  # initial weight for 2000 input neurons - this number was not changed
+
+
+	# parameters for making input - same as in Masquelier et al., 2008
+	runduration = 150  # length of simulation [s], full length = 450
+	tripling = False  # instead of creating 450 s of unique input, make 150 s and concatenate those spikes
+	dt_createpattern = 0.001  # [s]
+	number_npat = number_neurons - number_pat  # number of neurons not in pattern
+	patternlength = 0.05  # [s]
+	# parameters for basic spikes (2000 neurons, firing rate changes over time 0-90 Hz, avg 54Hz)
+	max_rate_pat = 90
+	min_rate_pat = 0
+	max_time_wo_spike_pat = 0.05
+	max_change_speed_pat = max_rate_pat / max_time_wo_spike_pat
+	# parameters for additional noise spikes (2000 neurons, firing rate does not change, avg 10 Hz)
+	max_rate_add = 10
+	min_rate_add = 10
+	max_time_wo_spike_add = 1000
+	max_change_speed_add = 0
+
+	# LIF output neuron model parameters
+	taum = 10 * ms
+	taus = 2.5 * ms
+	tausyn = 2.5 * ms
+	u_rest = 0
[…]

### 2018-05-27 code/figure_8.py

@@ -1,85 +1,96 @@
 from brian2 import *
 
-plt.rcParams['savefig.dpi'] = 75
-plt.rcParams['figure.autolayout'] = False
-plt.rcParams['figure.figsize'] = 12, 5
-plt.rcParams['axes.labelsize'] = 17
-plt.rcParams['axes.titlesize'] = 17
-plt.rcParams['font.size'] = 17
-plt.rcParams['lines.linewidth'] = 2.0
-plt.rcParams['lines.markersize'] = 3
-plt.rcParams['legend.fontsize'] = 16
+
+def fig_8():
+	print('%s Preparing Figure 8: EPSP shape' % time.strftime('%H:%M'))
+	device.reinit()
+	device.activate()
+
+	plt.rcParams['savefig.dpi'] = 75
+	plt.rcParams['figure.autolayout'] = False
+	plt.rcParams['figure.figsize'] = 12, 5
+	plt.rcParams['axes.labelsize'] = 17
+	plt.rcParams['axes.titlesize'] = 17
+	plt.rcParams['font.size'] = 17
+	plt.rcParams['lines.linewidth'] = 2.0
+	plt.rcParams['lines.markersize'] = 3
+	plt.rcParams['legend.fontsize'] = 16
+
+	runduration = 0.08
+	defaultclock.dt = 1e-4 * second
+
+	# parameters neuron and STDP
+	taum = 10 * ms
+	taus = 2.5 * ms
+	tausyn = 2.5 * ms
+	T = 2.8
+	u_rest = 0
+	X = (taus/taum)**(taum/(taus - taum))
+	A = -8.5
+	deltax = 1
+	deltaa = 1
+	deltau = 1.2
+	win = 1
+
+	# equations
+	eqs = '''du/dt = (A*a)/taus + (X*x-u)/taum : 1
+				dx/dt = -x/tausyn : 1 
+				da/dt = -a/taus : 1'''
+
+	eqs_reset = '''x = 0
+				u = 2*T
+				a = deltaa'''
+
+	stdp_model = ''' wi : 1'''
+
+	stdp_on_pre = ''' x_post += deltax*wi '''
+
+	stdp_on_pre_imm = ''' u_post += deltau*wi '''
+
+	# input
+	times = array([5, 20, 40, 43, 46, 60]) * ms
+	indices = zeros(len(times))
+
+	N0 = SpikeGeneratorGroup(1, indices, times)
[…]

### 2018-05-27 code/main.py

@@ -1,0 +1,57 @@
+from figure_1 import fig_1
+from figure_2 import fig_2
+from figure_5 import fig_5_saved, fig_5_new
+from figure_6 import fig_6_saved, fig_6_new
+from run_simulation import run_sim
+from figure_7CD import fig_7CD
+from figure_8 import fig_8
+from write_param_inputfile import write_para_file
+import matplotlib.pyplot as plt
+import argparse
+import time
+
+
+# ### get command line options
+ap = argparse.ArgumentParser()
+ap.add_argument("-n", "--new", required=False, default='False',
+	help="whether to run calculations for figures 5 and 6 (True) or use saved values (False, default)")
+args = vars(ap.parse_args())
+
+
+# ### plot figures from paper
+fig1 = fig_1()
+
+
+fig2 = fig_2()
+
+
+print('%s Preparing Figures 3, 4, 7AB' % time.strftime('%H:%M'))
+fig3, fig4, fig7AB = run_sim(0)
+
+
+if args['new'] == 'False':
+    fig5 = fig_5_saved()
+    fig6 = fig_6_saved()
+
+elif args['new'] == 'True':
+    write_para_file()
+
+    reps_per_combination = 1
+    fig5 = fig_5_new(reps_per_combination)
+
+    reps_per_resolution = 2
+    fig6 = fig_6_new(reps_per_resolution)
+
+else:
+    'input for --new was not recognised. Please try again'
+
+
+fig7CD = fig_7CD(3)
+
+
+fig8 = fig_8()
+
+
+plt.show()
+
+

### 2018-05-27 code/run_simulation.py

@@ -1,225 +1,284 @@
 from brian2 import *
 import time as time
 
-start_time = time.time()
 
-from plot_functions import make_figure_3, make_figure_4
-from create_input import make_input, make_pattern_presentation_array, copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
+def run_sim(run_idx, pararow=0, only_success=False):
+	from figure_3_4_7AB import fig_3, fig_4, fig_7AB
+	from create_input import make_input, make_pattern_presentation_array, \
+		copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
 
-set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone')
+	device.reinit()
+	device.activate()
+	set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone')
 
+	if only_success == True:
+		run_idx = pararow * 100 + run_idx
 
-# parameters that were varied for the publication
-random_seed = 1
-defaultclock.dt = 1e-4 * second  # length of time step for simulation
+	# parameters that were varied for the publication
+	random_seed = run_idx
+	defaultclock.dt = 1e-4 * second  # length of time step for simulation
 
-jitter_sd = 1  # amount of jitter
-number_pat = 1000  # number of neurons that take part in the pattern presentation
-pattern_freq = 0.25  # frequency at which the pattern is presented
-spike_del = 0  # percentage of spikes within the pattern to be deleted
-T = 0.5*(1-spike_del)*number_pat  # threshold
-win = 1.9*T/2000  # initial weight for 2000 input neurons - this number was not changed
-
-print('####################################')
-print('Simulation parameters')
-print('####################################')
-print('random seed =            ', random_seed)
-print('dt =                     ', defaultclock.dt)
-print('initial weight =         ', win)
-print('jitter (SD) =            ', jitter_sd)
-print('% of neurons in pattern =', 100*number_pat/2000)
-print('pattern freq =           ', pattern_freq)
-print('% spikes deleted =       ', spike_del*100)
-
-
-# parameters for making input - same as in Masquelier et al., 2008
[…]

### 2018-05-27 code/write_param_inputfile.py

@@ -1,0 +1,76 @@
+from brian2 import *
+
+
+'''
+Parameter combinations as saved in param.npy
+
+row     win     jit     # pat neur      pat freq        % deleted
+0       0.475   1       1000            0.25            0.0
+
+1       0.275   1       1000            0.25            0.0
+2       0.325   1       1000            0.25            0.0
+3       0.375   1       1000            0.25            0.0
+4       0.425   1       1000            0.25            0.0
+
+5       0.475   0       1000            0.25            0.0
+6       0.475   2       1000            0.25            0.0
+7       0.475   3       1000            0.25            0.0
+8       0.475   4       1000            0.25            0.0
+9       0.475   5       1000            0.25            0.0
+10      0.475   6       1000            0.25            0.0
+
+11      0.190   1       400             0.25            0.0
+12      0.285   1       600             0.25            0.0
+13      0.380   1       800             0.25            0.0
+14      0.570   1       1200            0.25            0.0
+
+15      0.475   1       1000            0.05            0.0
+16      0.475   1       1000            0.10            0.0
+17      0.475   1       1000            0.20            0.0
+18      0.475   1       1000            0.50            0.0
+
+19      0.4275  1       1000            0.25            0.1
+20      0.380   1       1000            0.25            0.2
+21      0.3325  1       1000            0.25            0.3
+
+'''
+
+
+def write_para_file():
+	r_ = array([4])
+	w_ = array([0.275, 0.325, 0.375, 0.425])
+	j_ = array([0, 2, 3, 4, 5, 6])
+	n_ = array([400, 600, 800, 1200])
+	f_ = array([0.05, 0.10, 0.15, 0.50])
+	d_ = array([0.10, 0.20, 0.30])
+	combinations = len(r_) + len(w_) + len(j_) + len(n_) + len(f_) + len(d_)
+	stand = array([4, 0.475, 1, 1000, 0.25, 0, 500])
+
+	para = empty((combinations, 7))
+	for i in range(combinations):
+		para[i] = stand
+	count = 0
+	for x in r_:
+		para[count][0] = x
+		count += 1
+	for x in w_:
+		para[count][1] = x
+		count += 1
+	for x in j_:
+		para[count][2] = x
+		count += 1
+	for x in n_:
+		para[count][3] = x
[…]

### 2018-06-13 Bavard-Thero-2018.tex
Add files via upload
We added one sentence on the results section according to one of the reviewers' advice.

--- 2018-05-17 article/Bavard-Thero-2018.tex
@@
                     \author[1]{Héloïse Théro}
-                            \affil[1]{Laboratoire de Neurosciences Cognitives et Computationnelles (ENS -
+                            \affil[1]{Laboratoire de Neurosciences Cognitives Computationnelles (ENS -
 INSERM), Département d'Études Cognitives, École normale supérieure, PSL
@@
 the reference article, or from another difference in model
-implementation.
+implementation. However, we obtained similar noise when dividing 
+the number of simulations by ten, suggesting the original authors 
+used around 500 iterations for this part of the figure.
 

### 2018-06-19 article/Hathway-Goodman-2018.tex
Fixed random seeds: code is now deterministic. Added Fig 9 and Fig10

@@
 a net decrease in weight after only a few spikes. Under the RNN rule
-(\textbf{C}), there are fewer changes and - in this example - no visible
-net weight change.}\label{fig:stdp}
+(\textbf{C}), there are fewer changes and --- in this example --- no
+visible net weight change.}\label{fig:stdp}
 \end{figure}
@@
 \includegraphics{figures/fig4_convergence.pdf}
-\caption{Weights at converged state. The weights in the converged state
-look very similar in the the original study (\textbf{A)}) and the
-replication implementation (\textbf{B)}). Weights from neurons not
-involved in the pattern are close to 0 (black, neurons 1000-1999),
-whereas weights from some neurons involved in the pattern (neurons
-0-999) are close to 1 (white, maximum value). The weights of neurons
-that spike at the beginning of the pattern are nearly all close to 1 and
-their added postsynaptic potentials cause a spike at the beginning of
-the pattern (blue rectangle).}\label{fig:conv}
+\caption{Weights at converged state. This figure shows a raster plot of
+the input spikes at the end of the simulation (converged state) with the
+color of the dots indicating the synaptic weight. The final weights in
+the converged state look very similar in the the original study
+(\textbf{A)}) and the replication implementation (\textbf{B)}). Weights
+from neurons not involved in the pattern are close to 0 (black, neurons
+1000-1999), whereas weights from some neurons involved in the pattern
+(neurons 0-999) are close to 1 (white, maximum value). The weights of
+neurons that spike at the beginning of the pattern are nearly all close
+to 1 and their added postsynaptic potentials cause a spike at the
+beginning of the pattern (blue rectangle).}\label{fig:conv}
 \end{figure}
@@
 
-The timing of the output neuron spikes is slightly different in the
+The spike times of the output neuron are slightly different in the
 original study and the replication. In the original study spike times of
@@
 neuron spikes leading to a time difference between the spikes of
-\(\Delta t=0\) where the STDP rule is undefined. Brian treats these
-input neuron spikes as if they happened just before the output neuron
-spike (\(\Delta t<0\), due to the scheduling of events in Brian) and
-will therefore increase all those weights instead of increasing some and
-decreasing others. This higher number of potentiations makes it more
-difficult for the system to systematically depress unimportant weights
-in order to become selective to the pattern. If the learning rule is
-modified so that the synaptic weight is increased as well as depressed
-when input neuron and output neuron spikes happen during the same time
-step (effectively causing a smaller weight increase), the pattern is
-found earlier, at around 17 s or 850 spikes (for a time step of
-\(10^{-4}s\)) which is close to the performance for smaller time steps
-(\(10^{-6}s\)) and the original paper (both 14 s or 700 spikes). We
-didn't use this modified learning rule in the rest of the paper because
-it differs from the implementation used in the majority of modelling
-papers.
+\(\Delta t=0\) where the STDP rule is undefined. Brian treats all of the
+input neuron spikes in this time bin as if they happened just before the
+output neuron spike (\(\Delta t<0\), due to the scheduling of events in
+Brian) and will therefore increase all those weights instead of
+increasing some and decreasing others. This higher number of
+potentiations makes it more difficult for the system to systematically
+depress unimportant weights in order to become selective to the pattern.
+
+If the learning rule is modified so that the change in synaptic weight
+reflects that on average half the input neurons spike before the output
+neuron and the other half afterwards (by adding the mean of LTP and LTD
+traces), the pattern is found earlier, at around 17 s or 850 spikes (for
+a time step of \(10^{-4}s\)) which is close to the performance for
+smaller time steps (\(10^{-6}s\)) and the original paper (both 14 s or
+700 spikes). This modified learning rule was only used to determine the
+time until finding (Fig.~\ref{fig:find}) because it differs from the
+implementation used in the majority of modelling papers.
 

### 2018-06-19 code/cluster/main_cluster.py

@@ -64,15 +64,45 @@ with open('%s/result_seed%s_%s.pickle' % (savedir, runid, name_run), 'wb') as ha
     pickle.dump(result_onerun, handle, protocol=pickle.HIGHEST_PROTOCOL)
 
 
+def make_input(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern, random_seed):
+    # this code was adapted from the original Matlab code
+    # this function creates input spikes one neuron at a time
+    spiketimes = []
+    afferents = []
+    runduration1 = min(runduration, 150)  # input will be tripled later, only need 150s
+    for n in range(number_neurons):
+        # call function to make a single spike train with a random seed for numba function
+        # (seed for this number is defined in run_simulation.py, so this is deterministic)
+        st = numpy.array(
+            make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons,
+                              dt_createpattern, numpy.random.randint(2**30)))
+        spiketimes.append(st)
+        afferents.append(n * ones(len(st)))
+    # convert spike trains into array and sort by time
+    spiketimes = hstack(spiketimes)
+    afferents = hstack(afferents)
+    sortarray = argsort(spiketimes)
+    spiketimes = spiketimes[sortarray]
+    afferents = afferents[sortarray]
+    return afferents, spiketimes
+
+
 @jit(nopython=True)  # numba decorator compiles this function into low level code to run faster
 def make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons,
-                      dt_createpattern):
+                      dt_createpattern, random_seed):
+    # this code was adapted from the original Matlab code
+    # get random seed from make_input() (changes for every neuron)
+    numpy.random.seed(int(random_seed))
     runduration1 = min(runduration, 150)  # input will be tripled later, only need 150s
     st = []
+    # add a virtual spike at beginning to avoid all neurons spiking at 0
     virtual_pre_sim_spike = - numpy.random.rand() * max_time_wo_spike
+    # firing rates change at each time step
     firing_rate = min_rate + numpy.random.rand() * (max_rate - min_rate)
     rate_change = 2 * (numpy.random.rand() - .5) * max_change_speed
     mtws = max_time_wo_spike
+    # create spike trains from changing firing rates: go through 150s 1ms at a time and
+    # add spikes according to current firing rate (changes every 1ms)
[…]

### 2018-06-19 code/cluster/run_main_cluster.sh

@@ -5,6 +5,5 @@
 #PBS -N rescience
 
 module load anaconda3/personal
-module load gcc-5.4
 
 python /THIS/PATH/NEEDS/TO/BE/CHANGED/main_cluster.py runid=$PBS_ARRAY_INDEX

### 2018-06-19 code/create_input.py

@@ -3,7 +3,11 @@
 #
 # Contributors: Pamela Hathway p.hathway16@imperial.ac.uk
 # -----------------------------------------------------------------------------
-# These functions are called by run_simulation.py to create the spike trains. 
+# These functions are called by run_simulation.py to create the spike trains.
+# They were adapted to Python from the Matlab code from
+# Masquelier T, Guyonneau R, Thorpe SJ (2008). Spike Timing Dependent Plasticity Finds the Start of
+# Repeating Patterns in Continuous Spike Trains. PLoS ONE 3(1): e1377 doi:10.1371/journal.pone.0001377
+
 # This script can be run independently.
 # -----------------------------------------------------------------------------
 
@@ -13,212 +17,238 @@ import numpy
 from numpy import array
 
 
-@jit(nopython=True) # numba decorator compiles this function into low level code to run faster
-def make_single_train(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern):
-	runduration1 = min(runduration, 150)  # input will be tripled later, only need 150s
-	st = []
-	virtual_pre_sim_spike = - numpy.random.rand() * max_time_wo_spike
-	firing_rate = min_rate + numpy.random.rand() * (max_rate - min_rate)
-	rate_change = 2 * (numpy.random.rand() - .5) * max_change_speed
-	mtws = max_time_wo_spike
-	for t in numpy.arange(dt_createpattern, runduration1, dt_createpattern):
-		if numpy.random.rand() < dt_createpattern * firing_rate or \
-			(len(st) < 1 and t - virtual_pre_sim_spike > mtws) or \
-			(len(st) > 0 and t - st[-1] > mtws):
-			tmp = t - dt_createpattern * numpy.random.rand()
-			tmp = max(0, tmp)
-			tmp = min(runduration1, tmp)
-			st.append(tmp)
-			mtws = max_time_wo_spike
-		firing_rate = firing_rate + rate_change * dt_createpattern
-		rate_change = rate_change + 1 / 5 * 2 * (numpy.random.rand() - .5) * max_change_speed
-		rate_change = max(min(rate_change, max_change_speed), -max_change_speed)
-		firing_rate = max(min(firing_rate, max_rate), min_rate)
-	return st
+def make_input(min_rate, max_rate, max_time_wo_spike, max_change_speed, runduration, number_neurons, dt_createpattern, random_seed):
+    # this code was adapted from the original Matlab code
[…]

### 2018-06-19 code/figure_3_4_7AB.py

@@ -10,7 +10,7 @@
 from brian2 import *
 
 
-def fig_3(N2spm, latency):
+def fig_3(N2spm, latency, random_seed):
     plt.rcParams['savefig.dpi'] = 75
     plt.rcParams['figure.autolayout'] = False
     plt.rcParams['figure.figsize'] = 6, 5
@@ -31,12 +31,12 @@ def fig_3(N2spm, latency):
 
     tight_layout(rect=[0, 0.03, 1, 0.95])
 
-    savefig('../article/figures/figure_3_created_%s.pdf' % datetime.datetime.now().strftime('%y%m%d'))
+    savefig('../article/figures/figure_3_seed%s_created_%s.pdf' % (random_seed, datetime.datetime.now().strftime('%y%m%d')))
 
     return fig3
 
 
-def fig_4(N1spm, syn12, timing_pattern):
+def fig_4(N1spm, syn12, timing_pattern, random_seed):
     plt.rcParams['savefig.dpi'] = 75
     plt.rcParams['figure.autolayout'] = False
     plt.rcParams['figure.figsize'] = 12, 5
@@ -49,11 +49,16 @@ def fig_4(N1spm, syn12, timing_pattern):
     # get spikes from end of simulation
     figsta = searchsorted(N1spm.t / second, timing_pattern[-2] - 0.025)
     figsto = searchsorted(N1spm.t / second, timing_pattern[-2] + 0.075)
-    greyness = [str((255 - syn12.wi[int(i)]) / 255.) for i in N1spm.i[figsta:figsto]]  # high w == white
+    greyness = [str((255 - syn12.wi[int(i)]*255) / 255.) for i in N1spm.i[figsta:figsto]]  # high w == white
 
     ax = figure(figsize=(8, 6)).add_subplot(1, 1, 1)
     ax.set_facecolor("k")
-    ax.scatter(N1spm.t[figsta:figsto] / second, N1spm.i[figsta:figsto], c=greyness, s=3, lw=0, cmap='Greys')
+    sc = scatter(N1spm.t[figsta:figsto] / second, N1spm.i[figsta:figsto], c=greyness, s=3, lw=0, cmap='Greys')
+    cbar = colorbar(sc)
+    cbar.set_ticks([round(1 - t, 1) for t in cbar.ax.get_yticks()])
+    cbar.set_ticklabels([round(1 - t, 1) for t in cbar.ax.get_yticks()])
+    cbar.set_label('synaptic weight')
+
     title('Figure 4')
     xlabel('time [s]')
[…]

### 2018-06-19 code/figure_6.py

@@ -189,14 +189,14 @@ def run_find(res, rep):
     					wi = clip(wi + LTPtrace, wmin, wmax)
     					LTPtrace = 0''')
 
-    seed(int(random_seed))
+    numpy.random.seed(int(random_seed))
     indices, times = make_input(min_rate_pat, max_rate_pat, max_time_wo_spike_pat,
-                                max_change_speed_pat, runduration, number_neurons, dt_createpattern)
+                                max_change_speed_pat, runduration, number_neurons, dt_createpattern, random_seed)
     indices_add, times_add = make_input(min_rate_add, max_rate_add, max_time_wo_spike_add,
-                                        max_change_speed_add, runduration, number_neurons, dt_createpattern)
-    position_copypaste = make_pattern_presentation_array(runduration, patternlength, pattern_freq)
+                                        max_change_speed_add, runduration, number_neurons, dt_createpattern, random_seed)
+    position_copypaste = make_pattern_presentation_array(runduration, patternlength, pattern_freq, random_seed)
     times, indices = copy_and_paste_jittered_pattern(times, indices, position_copypaste,
-                                                     patternlength, jitter_sd, spike_del, number_pat)
+                                                     patternlength, jitter_sd, spike_del, number_pat, random_seed)
     times, indices = add_noise(times, indices, times_add, indices_add)
     if tripling and runduration > 300:
         times, indices = triple_input_runtime(times, indices)

### 2018-06-19 code/figure_7CD.py

@@ -12,171 +12,174 @@ import time as time
 
 
 def fig_7CD(run_idx):
-	from create_input import make_input, make_pattern_presentation_array, \
-		copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
+    from create_input import make_input, make_pattern_presentation_array, \
+        copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
 
-	device.reinit()
-	device.activate()
-	set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone_ata3')
-	print('%s Preparing Figure 7 C and D: success with ATA rule' % time.strftime('%H:%M'))
+    device.reinit()
+    device.activate()
+    set_device('cpp_standalone', build_on_run=True, directory='STDP_standalone_ata3')
+    print('%s Preparing Figure 7 C and D: success with ATA rule' % time.strftime('%H:%M'))
 
-	# parameters that were varied for the publication
-	random_seed = run_idx
-	defaultclock.dt = 1e-4 * second  # length of time step for simulation
+    # parameters that were varied for the publication
+    random_seed = run_idx
+    defaultclock.dt = 1e-4 * second  # length of time step for simulation
 
-	jitter_sd = 0  # amount of jitter
-	number_pat = 1000  # number of neurons that take part in the pattern presentation
-	number_neurons = 2000
-	pattern_freq = 0.25  # frequency at which the pattern is presented
-	spike_del = 0  # percentage of spikes within the pattern to be deleted
-	T = 500  # threshold
-	win = 0.2375  # initial weight for 2000 input neurons - this number was not changed
+    jitter_sd = 0  # amount of jitter
+    number_pat = 1000  # number of neurons that take part in the pattern presentation
+    number_neurons = 2000
+    pattern_freq = 0.25  # frequency at which the pattern is presented
+    spike_del = 0  # percentage of spikes within the pattern to be deleted
+    T = 500  # threshold
+    win = 0.2375  # initial weight for 2000 input neurons - this number was not changed
 
+    # parameters for making input - same as in Masquelier et al., 2008
+    runduration = 150  # length of simulation [s], full length = 450
[…]

### 2018-06-19 code/main.py

@@ -35,7 +35,7 @@ fig1 = fig_1()
 fig2 = fig_2()
 
 print('%s Preparing Figures 3, 4, 7AB' % time.strftime('%H:%M'))
-fig3, fig4, fig7AB = run_sim(0)
+fig3, fig4, fig7AB, fig9, fig10 = run_sim(1)
 
 if args['new'] == 'False':
     fig5 = fig_5_saved()
@@ -55,7 +55,7 @@ elif args['new'] == 'True':
 else:
     'input for --new was not recognised. Please try again'
 
-fig7CD = fig_7CD(3)
+fig7CD = fig_7CD(28)
 
 fig8 = fig_8()
 

### 2018-06-19 code/run_simulation.py

@@ -12,7 +12,7 @@ import time as time
 
 
 def run_sim(run_idx, pararow=0, only_success=False):
-    from figure_3_4_7AB import fig_3, fig_4, fig_7AB
+    from figure_3_4_7AB import fig_3, fig_4, fig_7AB, fig_9, fig_10
     from create_input import make_input, make_pattern_presentation_array, \
         copy_and_paste_jittered_pattern, add_noise, triple_input_runtime, remove_close_spikes
 
@@ -96,6 +96,7 @@ def run_sim(run_idx, pararow=0, only_success=False):
 
     # plotting parameters
     record_wi_dt = 0.05
+    record_u_dt = 0.0001
 
     # equations
     eqs = '''du/dt = (A*a)/taus + (X*x-u)/taum : 1
@@ -112,38 +113,38 @@ def run_sim(run_idx, pararow=0, only_success=False):
 
     stdp_on_pre = (''' x_post += deltax*wi
                         LTPtrace = aplus
-						wi = clip(wi + LTDtrace, wmin, wmax)
-						LTDtrace = 0''')
+                        wi = clip(wi + LTDtrace, wmin, wmax)
+                        LTDtrace = 0''')
 
     stdp_on_post = ('''LTDtrace = -aminus
-						wi = clip(wi + LTPtrace, wmin, wmax)
-						LTPtrace = 0''')
+                        wi = clip(wi + LTPtrace, wmin, wmax)
+                        LTPtrace = 0''')
 
     stdp_on_pre_nn = (''' x_post += deltax*wi
-						LTPtrace = aplus
-						wi = clip(wi + LTDtrace, wmin, wmax)''')
+                        LTPtrace = aplus
+                        wi = clip(wi + LTDtrace, wmin, wmax)''')
 
     stdp_on_post_nn = ('''LTDtrace = -aminus
-						wi = clip(wi + LTPtrace, wmin, wmax)''')
+                        wi = clip(wi + LTPtrace, wmin, wmax)''')
 
     stdp_on_pre_ata = (''' x_post += deltax*wi
-						LTPtrace += aplus
-						wi = clip(wi + LTDtrace, wmin, wmax)''')
+                        LTPtrace += aplus
+                        wi = clip(wi + LTDtrace, wmin, wmax)''')
 
     stdp_on_post_ata = ('''LTDtrace -= aminus
-						wi = clip(wi + LTPtrace, wmin, wmax)''')
+                        wi = clip(wi + LTPtrace, wmin, wmax)''')
 
[…]

### 2018-06-19 code/write_param_inputfile.py

@@ -10,7 +10,6 @@
 
 from brian2 import *
 
-
 '''
 Parameter combinations as saved in param.npy
 
@@ -36,7 +35,7 @@ row     win     jit     # pat neur      pat freq        % deleted
 
 15      0.475   1       1000            0.05            0.0
 16      0.475   1       1000            0.10            0.0
-17      0.475   1       1000            0.20            0.0
+17      0.475   1       1000            0.15            0.0
 18      0.475   1       1000            0.50            0.0
 
 19      0.4275  1       1000            0.25            0.1
@@ -47,40 +46,40 @@ row     win     jit     # pat neur      pat freq        % deleted
 
 
 def write_para_file():
-	r_ = array([4])
-	w_ = array([0.275, 0.325, 0.375, 0.425])
-	j_ = array([0, 2, 3, 4, 5, 6])
-	n_ = array([400, 600, 800, 1200])
-	f_ = array([0.05, 0.10, 0.15, 0.50])
-	d_ = array([0.10, 0.20, 0.30])
-	combinations = len(r_) + len(w_) + len(j_) + len(n_) + len(f_) + len(d_)
-	stand = array([4, 0.475, 1, 1000, 0.25, 0, 500])
+    r_ = array([4])
+    w_ = array([0.275, 0.325, 0.375, 0.425])
+    j_ = array([0, 2, 3, 4, 5, 6])
+    n_ = array([400, 600, 800, 1200])
+    f_ = array([0.05, 0.10, 0.15, 0.50])
+    d_ = array([0.10, 0.20, 0.30])
+    combinations = len(r_) + len(w_) + len(j_) + len(n_) + len(f_) + len(d_)
+    stand = array([4, 0.475, 1, 1000, 0.25, 0, 500])
 
-	para = empty((combinations, 7))
-	for i in range(combinations):
-		para[i] = stand
-	count = 0
-	for x in r_:
-		para[count][0] = x
-		count += 1
-	for x in w_:
-		para[count][1] = x
-		count += 1
-	for x in j_:
-		para[count][2] = x
-		count += 1
-	for x in n_:
-		para[count][3] = x
-		para[count][6] = 0.5 * (1 - stand[5]) * x
-		para[count][1] = 1.9 * 0.5 * (1 - stand[5]) * x / 2000
-		count += 1
-	for x in f_:
-		para[count][4] = x
-		count += 1
-	for x in d_:
-		para[count][5] = x
-		para[count][6] = 0.5 * (1 - x) * stand[3]
[…]

### 2018-06-27 article/Hathway-Goodman-2018.tex
changes in response to review2

@@
 \usepackage{fontspec}
-%\defaultfontfeatures{Path = /usr/local/texlive/2018/texmf-dist/fonts/opentype/public/fontawesome/}
 \usepackage{fontawesome}
@@
 equations to model the system parameters and evaluates those equations
-for each time step. We converted the kernels of the posystnaptic
-potential and the spike afterpotential into the following differential
-equations:
+for each time step. The SRM kernels can be modelled using alpha
+functions \autocite{Brette2007}, so we converted the kernels of the
+postsynaptic potential and the spike afterpotential into the following
+differential equations:
 
@@
 
-The values for the parameters can be found in Tbl.~\ref{tbl:param}.
-Eq.~\ref{eq:1} describes the postsynaptic membrane potential, which is
-influenced by presynaptic excitatory postsynaptic potentials (EPSPs,
-first term) and a negative afterpotential after a spike (second term).
-In the event of a presynaptic spike, \(x\) is increased by 1
-(\(x\leftarrow x+1\)), which initiates the voltage increase in the
-postsynaptic neuron. When the postsynaptic voltage reaches the
-threshold, a spike occurs: the voltage \(u\) is set to twice the
-threshold (\(u\leftarrow 2T\)), all EPSPs are flushed (all \(x\) set to
-0, \(x\leftarrow 0\)) and the negative afterpotential is set into motion
-(\(a\) is set to 1, \(a\leftarrow 1\)). Presynaptic spikes had the same
-effect on the postsynaptic neuron potential as in the original
-publication as shown in Fig.~\ref{fig:pot}.
+The values for the parameters can be found in Tbl.~\ref{tbl:param}. We
+chose the exact method for integrating the differential equations
+(called `linear' in Brian). Eq.~\ref{eq:1} describes the postsynaptic
+membrane potential, which is influenced by presynaptic excitatory
+postsynaptic potentials (EPSPs, first term) and a negative
+afterpotential after a spike (second term). In the event of a
+presynaptic spike at neuron \(i\), \(x\) is increased by the respecitve
+synaptic weight \(w_{i}\) (\(x\leftarrow x+w_{i}\)), which initiates the
+voltage increase in the postsynaptic neuron. When the postsynaptic
+voltage reaches the threshold, a spike occurs: the voltage \(u\) is set
+to twice the threshold (\(u\leftarrow 2T\)), all EPSPs are flushed (all
+\(x\) set to 0, \(x\leftarrow 0\)) and the negative afterpotential is
+set into motion (\(a\) is set to 1, \(a\leftarrow 1\)). Presynaptic
+spikes had the same effect on the postsynaptic neuron potential as in
+the original publication as shown in Fig.~\ref{fig:pot}.
 
@@
 the input spikes at the end of the simulation (converged state) with the
-color of the dots indicating the synaptic weight. The final weights in
-the converged state look very similar in the the original study
-(\textbf{A)}) and the replication implementation (\textbf{B)}). Weights
-from neurons not involved in the pattern are close to 0 (black, neurons
-1000-1999), whereas weights from some neurons involved in the pattern
-(neurons 0-999) are close to 1 (white, maximum value). The weights of
-neurons that spike at the beginning of the pattern are nearly all close
-to 1 and their added postsynaptic potentials cause a spike at the
-beginning of the pattern (blue rectangle).}\label{fig:conv}
+color of the dots indicating the synaptic weight (see colorbar). The
+final weights in the converged state look very similar in the the
+original study (\textbf{A)}) and the replication implementation
+(\textbf{B)}). Weights from neurons not involved in the pattern are
+close to 0 (black, neurons 1000-1999), whereas weights from some neurons
+involved in the pattern (neurons 0-999) are close to 1 (white, maximum
+value). The weights of neurons that spike at the beginning of the
+pattern are nearly all close to 1 and their added postsynaptic
+potentials cause a spike at the beginning of the pattern (blue
+rectangle).}\label{fig:conv}
 \end{figure}
@@
 700 spikes). This modified learning rule was only used to determine the
-time until finding (Fig.~\ref{fig:find}) because it differs from the
-implementation used in the majority of modelling papers.
+time until finding (Fig.~\ref{fig:find}).
 
@@
 postsynaptic voltage behaves similarly, but the postsynaptic spike time
-is slightly different (green and black dots).}\label{fig:epsps}
+is slightly different
+(/home/ph416/Documents/ReScience-submission/article/figures/fig8\_kernel\_vs\_imm.pdf).}\label{fig:epsps}
 \end{figure}

### 2018-06-27 code/figure_3_4_7AB.py

@@ -55,8 +55,9 @@ def fig_4(N1spm, syn12, timing_pattern, random_seed):
     ax.set_facecolor("k")
     sc = scatter(N1spm.t[figsta:figsto] / second, N1spm.i[figsta:figsto], c=greyness, s=3, lw=0, cmap='Greys')
     cbar = colorbar(sc)
-    cbar.set_ticks([round(1 - t, 1) for t in cbar.ax.get_yticks()])
-    cbar.set_ticklabels([round(1 - t, 1) for t in cbar.ax.get_yticks()])
+    cbar.ax.invert_yaxis()
+    cbar.set_ticks([0, 0.2, 0.4, 0.6, 0.8, 1])
+    cbar.set_ticklabels([1, 0.8, 0.6, 0.4, 0.2, 0])
     cbar.set_label('synaptic weight')
 
     title('Figure 4')

### 2018-06-27 code/run_simulation.py

@@ -98,28 +98,33 @@ def run_sim(run_idx, pararow=0, only_success=False):
     record_wi_dt = 0.05
     record_u_dt = 0.0001
 
-    # equations
+    # equations for the LIF enuron
     eqs = '''du/dt = (A*a)/taus + (X*x-u)/taum : 1
             dx/dt = -x/tausyn : 1
             da/dt = -a/taus : 1'''
 
+    # variable changes when a postsynaptic spike happens
     eqs_reset = '''x = 0
                     u = 2*T
                     a = deltaa'''
 
+    # equations for the STDP implementations: traces of presynaptic (LTDtrace) and postsynaptic (LTPtrace) activity
     stdp_model = ''' wi : 1
                     dLTPtrace/dt = -LTPtrace / tauplus  : 1 (event-driven)
                     dLTDtrace/dt = -LTDtrace / tauminus : 1 (event-driven)'''
 
+    # variable changes when a presynaptic spike happens
     stdp_on_pre = (''' x_post += deltax*wi
                         LTPtrace = aplus
                         wi = clip(wi + LTDtrace, wmin, wmax)
                         LTDtrace = 0''')
 
+    # variable changes when a postsynaptic spike happens
     stdp_on_post = ('''LTDtrace = -aminus
                         wi = clip(wi + LTPtrace, wmin, wmax)
                         LTPtrace = 0''')
-
+    
+    # different STDP equations for NN learning rule
     stdp_on_pre_nn = (''' x_post += deltax*wi
                         LTPtrace = aplus
                         wi = clip(wi + LTDtrace, wmin, wmax)''')
@@ -127,6 +132,7 @@ def run_sim(run_idx, pararow=0, only_success=False):
     stdp_on_post_nn = ('''LTDtrace = -aminus
                         wi = clip(wi + LTPtrace, wmin, wmax)''')
 
+    # different STDP equations for ATA learning rule
     stdp_on_pre_ata = (''' x_post += deltax*wi
                         LTPtrace += aplus
                         wi = clip(wi + LTDtrace, wmin, wmax)''')
@@ -134,6 +140,7 @@ def run_sim(run_idx, pararow=0, only_success=False):
     stdp_on_post_ata = ('''LTDtrace -= aminus
                         wi = clip(wi + LTPtrace, wmin, wmax)''')
 
+    # creating input spikes for presynaptic neurons
     if only_success == False:
         print('    #### Creating input')
     numpy.random.seed(int(random_seed))
[…]

### 2019-01-11 article/CRESSOT-CONINX-MERCKLING-PERRIN-DONCIEUX-2019.tex
ready for pull request, modified Makefile to compile .tex only, no .md

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{fontspec}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\captionsetup[figure]{name=Fig.,labelfont={sf,bf},textfont=sl,labelsep=quad}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            pdfauthor=
{
        Loic Cressot
            Alexandre Coninx
            Astrid Merckling
            Nicolas Perrin
            Stéphane Doncieux
  },
            pdftitle=
{
[Re] Learning state representations with robotic priors
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\usepackage{siunitx}
\newcommand{\blau}[1]{{\color{blue} #1}}
\newcommand{\rot}[1]{{\color{red} #1}}

\usepackage{ulem}
\definecolor{myred}{rgb}{0.8,0,0}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{myblue}{rgb}{0,0,0.7}
\newcommand{\np}[1]{\textcolor{myred}{\fbox{\bf NP}{\em #1}}}

\lstnewenvironment{code}{\lstset{language=Python,basicstyle=\small\ttfamily,showstringspaces=false,stringstyle=\color{blue},commentstyle=\color{blue},columns=spaceflexible}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newtcolorbox{repobox2}
   {colback=black, colframe=black!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Jan 2018 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Learning state representations with robotic priors}

\usepackage{authblk}
                \author[1]{Loic Cressot}
                \author[1]{Alexandre Coninx}
                \author[1]{Astrid Merckling}
                \author[1]{Nicolas Perrin}
                \author[1]{Stéphane Doncieux}
                \affil[1]{Sorbonne Université, CNRS, Institut des Systèmes Intelligents et de Robotique, ISIR, F-75005 Paris, France}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:Lcressot@gmail.com}{Lcressot@gmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

% --- Additional packages ----
% Package for subfigures
%\usepackage[position=top,labelformat=simple,labelfont=bf]{subcaption}
%\renewcommand*{\thesubfigure}{\Alph{subfigure}} % Upper case letters for subfigures

%% Useful packages
\usepackage{amsmath}
\usepackage[colorinlistoftodos]{todonotes}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Month, day, year\\
  {\bfseries Accepted}  Month, day, year\\
  {\bfseries Published} Month, day, year\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}

    \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Learning state representations with robotic priors, Jonschkowski, R. and Brock, O., 2015.  Autonomous Robots, 39(3), pp.407-428
\end{itemize}\par
\end{flushleft}
\end{rebox}

%------------------------------------------------------------------------
\section*{Introduction}
%------------------------------------------------------------------------
Under Markovian assumptions, an agent interacting with an environment only needs its current observation in this environment to be able to optimally choose its next action. In this scenario, which we consider in this paper, state representation learning aims at learning to compress observations of an agent  into compact and task oriented representations.
 By means of a dimensionality reduction and the removal of irrelevant information from observations, these compact representations should simplify the task of algorithms exploiting them, such as reinforcement learning algorithms. Jonschkowski and Brock \cite{jonschkowski2015learning} proposed a method for learning such a transformation by minimizing a loss with gradient descent. The loss implements priors that enforce the representations to be coherent with a physically plausible description of the robot state. 
The utility of the learned representations was evaluated using the fitted Q iteration reinforcement learning algorithm \cite{ernst2005tree}.

In this paper, we propose novel implementations of the state representation learning introduced in \cite{jonschkowski2015learning} and the fitted Q iteration algorithm, along with open-source code for the generation of new learning datasets \cite{RoundBotRepo}\cite{RoundBotArchive}. We used these implementations to reproduce the main results of the original manuscript on a mobile robot navigation task in simulation. We did not reproduce the sections of \cite{jonschkowski2015learning} corresponding to a slot car racing task and a navigation task with a real robot.\\

%------------------------------------------------------------------------
\section*{Methods}
%------------------------------------------------------------------------

We didn't contact the authors of \cite{jonschkowski2015learning}, and as they did not provide code for the data generation and the fitted Q iteration algorithm, we reimplemented those two. This enabled us to generate our own training data and make our own testing of the trained policies using the learned representations as states for these policies.
In addition, in order to maximize the independence of our results from those of the original paper, we also recoded the core algorithm with a different library: we used Keras with a Tensorflow backend instead of Sonnet and Tensforflow in the original code.\\

\subsubsection*{Data generation}
\label{sec:DataGeneration}

Our simulator is an OpenAI gym environment \cite{openAIgymref} that can be easily installed with pip, the python package installer, and can be found on a public git repository \cite{RoundBotRepo} and a public archive \cite{RoundBotArchive}. 
Figures \ref{subfig:obsSimple} and \ref{subfig:obsdistractors} display observations received in the simulator with an egocentric view, and Fig.~\ref{subfig:obsTopDown} displays observations with an exocentric view (from above). In addition, Fig.~\ref{fig:simpleVSdistrators} presents the environment with and without visual distractors.

\subsubsection*{Fitted Q iteration}
\label{sec:FittedQIteration}

We reimplemented the fitted Q iteration algorithm in Python, using exactly the same method and hyper-parameters as the one described in the original paper.

\subsubsection*{Representation learning}
\label{sec:RepresentationLearning}

We implemented our own new version of the state representation algorithm in Python with the Keras library using a Tensorflow back end, and used the same hyper-parameters as in the original paper, except for the regularization parameter $\lambda$ in the experiment with a view from above and distractors (see sections \textsf{\textbf{Invariance to Perspective}} and \textsf{\textbf{Ignoring distractors}}).

In three of the losses, conditional expectation are computed considering states, actions and rewards at two different time steps. In our opinion, the original paper has a minor ambiguity concerning the choice of the set of these time pairs $t_1$ and $t_2$. The original code shared by the authors of \cite{jonschkowski2015learning} allowed us to understand that for a given batch containing observations at different times $t_i$, all the possible pairs inside the batch are taken, provided that they satisfy some constraints present in the loss formulation (e.g. all the pairs $(t_1,t_2)$ for which the actions are identical: $a_{t_1}=a_{t_2}$,
or all the pairs with identical actions and different subsequent rewards: $a_{t_1}=a_{t_2} \wedge r_{t_1+1} \neq r_{t_2+1}$).

%------------------------------------------------------------------------
\section*{Results}
%------------------------------------------------------------------------

\subsection*{Simple navigation task}
%----------------------------------

In this section we reproduce the first experiment presented in \cite{jonschkowski2015learning}: \textit{the learning process on the simple navigation task}. As shown in figure \ref{subfig:obsSimple}, we choose the parameters of our data generator that produce observations resembling to those chosen by the authors in this first experiment : one-colored walls with 300° field of view. Then we run our implementation of the state representation learning algorithm with the same hyper-parameters as in the article and found qualitatively similar learned 2D states, as shown in Fig.~\ref{subfig:repSimple}.\\

We also run the same learning with 5 dimensional states and performed a principal component analysis (PCA) to show that the learned representation predominantly lies on only two principal dimensions (see Fig.~\ref{fig:pca}), as observed in \cite{jonschkowski2015learning}.\\

Afterwards, keeping the same training set, we run a 25 steps gradient descent training, and evaluate the learned representations at each step, as presented on figure \ref{fig:qleval}. As in the article, we perform 10 fitted Q iteration trainings for each step of the representation learning, and for each of the policies learned with fitted Q iteration, we perform 20 evaluations in the same environment as the one used for generating the training data. We sum the rewards over each evaluation and average them for each fitted Q learning, and gather the points on the plot shown on figure \ref{fig:qleval}. This experience shows that our results are not only qualitatively similar to those in the paper but also quantitatively. We observe that the losses, stacked on figure \ref{fig:qleval}, are less smooth than those in the article, particularly the causality loss. We also use a smaller gradient descent step than in the article because it seems that the representations are converging too fast for a proper step by step analysis.

\begin{figure}[ht]
\centering
  \begin{tabular}{cc}
  \centering
  \subfloat[]{
  \includegraphics[width=.39\linewidth,]{pictures/observations_simple_crop.png}
  \label{subfig:obsSimple}
  }
  & 
  \subfloat[]{
    \includegraphics[width=0.56\linewidth]{pictures/representation_simple.png}
  \label{subfig:repSimple}
  }
  \end{tabular}
\caption[Simple navigation task]{Simple navigation task experiment : 
\protect\subref{subfig:obsSimple} \textnormal{Observations}
\protect\subref{subfig:repSimple} \textnormal{2D state representation learned.}
\label{fig:simpleNavTask}
} 
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.6\linewidth]{pictures/PCA.png}
\caption[]{Eigenvalues on a 5D representation learned with the same data and parameters as in Fig.~\ref{fig:simpleNavTask}.}
\label{fig:pca}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pictures/ql_eval.png}
\caption[]{Fitted Q iteration evaluation }
\label{fig:qleval}
\end{figure}

\subsection*{Invariance to perspective}
\label{sec:inv2pers}
%--------------------------------------
In this section we reproduce the second experiment of the article: \textit{The invariance to perspective on the navigation task}. Again we use our own gym environment as a data generator to produce top-down perspective observations, as shown on figure \ref{subfig:obsTopDown}, where we artificially set a bigger radius to the robot, as in the article, so as to make it more visible. Then we run the state representation learning algorithm with the same hyper-parameters as the authors, except for the $L_1$ regularization parameter $\lambda$, see figure \ref{subfig:repTopDown}.\\

Indeed, when setting the regularization parameter $\lambda$ to a strong value such as $1e-1$ like in the paper, our implementation cannot learn any relevant representation, but converges instead to a blurry and mixed point cloud. However, for a low value of regularization, the learned representation is qualitatively similar to the one in the paper: a blurry version of the egocentric version. We should mention that we tried the initial code of the authors with the strong regularization value and it worked as presented in their paper.

\begin{figure}[ht]
\centering
  \begin{tabular}{cc}
  \centering
  \subfloat[]{
    \includegraphics[width=.39\linewidth]{pictures/observations_topdown_crop.png}
    \label{subfig:obsTopDown}
  }
  & 
  \subfloat[]{
    \includegraphics[width=.56\linewidth]{pictures/representation_topdown.png}
    \label{subfig:repTopDown}
  }
  \end{tabular}
\caption[Invariance to perspective]{Invariance to perspective experiment : \protect\subref{subfig:obsTopDown} \textnormal{Observations from a global point of view}
\protect\subref{subfig:repTopDown} \textnormal{2D state representation with $\lambda=1e-4$} \label{fig:invarianceToPerspective}} 
\end{figure}

\subsection*{Ignoring distractors}
%---------------------------------
In this section we reproduce the experiment on the navigation task with visual distractors. As shown on figure \ref{fig:simpleVSdistrators}, we implemented an option in our simulator to produce visual distractors with colored rectangles moving randomly on the surface of the walls and ground at a lower speed than the speed of the robot.\\

The obtained training observations, presented on figure \ref{subfig:obsdistractors} are used to train a state representation with the same hyper-parameters as in the paper, except again for the $\lambda$ parameter which is set to the slightly lower value $0.1$ instead of $0.3$, simply because it seems to give a qualitatively more similar result. The learned representation, presented on figure \ref{subfig:repdistractors}, is again coherent with the results presented in the paper.

\begin{figure}[ht]
\begin{minipage}[c]{.5\textwidth}
\centering
\subfloat[]{
  \includegraphics[width=0.90\linewidth]{pictures/simpleNav.png}
  \label{subfig:simpleNav}
}
\end{minipage}
\begin{minipage}[c]{.5\textwidth}
\subfloat[]{
  \includegraphics[width=0.80\linewidth]{pictures/distractors.png}
  \label{subfig:distractors}
}
\end{minipage}
\caption[]{Adding visual distractors in the setup : 
\protect\subref{subfig:simpleNav} \textnormal{Original setup}
\protect\subref{subfig:distractors} \textnormal{Setup with visual distractors}
\label{fig:simpleVSdistrators}
} 
\end{figure}

\begin{figure}
\centering
  \begin{tabular}{cc}
  \centering
  \subfloat[]{
    \includegraphics[width=0.39\linewidth]{pictures/observations_distractors_crop.png}
  \label{subfig:obsdistractors}
  }
  & 
  \subfloat[]{
    \includegraphics[width=0.56\linewidth]{pictures/representation_distractors.png}
  \label{subfig:repdistractors}
  }
  \end{tabular}
\caption[Ignoring distractors]{Ignoring distractors experiment : 
\protect\subref{subfig:obsdistractors} \textnormal{Observations with visual distractors}
\protect\subref{subfig:repdistractors} \textnormal{2D state representation with $\lambda=0.1$}
\label{fig:ignoringDistrators}
} 
\end{figure}

%------------------------------------------------------------------------
\section*{Conclusion}
 Our implementation produces results that seems faithful to those of the original paper. However, its computational speed is lower than the code provided by the authors, probably because we had to use some tricks to force the Keras library to deal with unsupervised learning and conditional expectations on pairwise operations.
%------------------------------------------------------------------------

\sffamily \small
\printbibliography[title=References]

\end{document}

### 2019-02-20 article/Caron-Lessard-Wu-Poisot-2018.tex
figures updated

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{amssymb,amsmath}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Dominique Caron,
      Vincent Lessard,
      Qile Wu,
      Timothée Poisot,
  },
            pdftitle=
{
[Re] Insect natural enemies as regulating factors
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Insect natural enemies as regulating factors}

    \usepackage{authblk}
                        \author[1]{Dominique Caron}
                    \author[1]{Vincent Lessard}
                    \author[1]{Qile Wu}
                    \author[1]{Timothée Poisot}
                            \affil[1]{Département de sciences biologiques, Université de Montréal, Montréal,
Québec, Canada}
                    \affil[2]{Québec Centre for Biodiversity Sciences, Montréal, Québec, Canada}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:timothee.poisot@umontreal.ca}{timothee.poisot@umontreal.ca}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/BIO6032/2018\_replication\_hassell\_1985/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/BIO6032/2018\_replication\_hassell\_1985/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] Hassell, M. P. Insect Natural Enemies as Regulating Factors. Journal of
Animal Ecology, vol.~54, no. 1, 1985, pp.~323--334.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Parasitism is a special case of predation. In both interactions, a
species (parasitoid or predator) feeds on the other species (host or
prey), acting as a regulating factor (\textcite{Anderson78}). However,
the population dynamics of both systems are very different.
\textcite{Thompson24} was the first to propose a model to describe this
host-parasitoid system. In his model, parasites are limited by the
number of eggs they lay. Depending on the relative increase rate of
hosts and parasites, either both population increase indefinitely or
decrease to extinction. Later, \textcite{Nicholson35} proposed other
models for which the rate of increase of the parasites is limited by
their capacity to find hosts. These were the basis for many other models
where parasites act as regulating factors (\textcite{Hassell78};
\textcite{Rockwood15}).

In 1983, \textcite{Dempster83} proposed that natural enemies may not be
an important regulating factor in insect dynamics. In fact, he failed to
detect density-dependence due to natural enemies in most of the studies
on Lepidoptera he reviewed. His proposition really contrasted with what
was thought at the time. In response to this article,
\textcite{Hassell85} analyzed an insect dynamic model in which the only
regulating factor was natural enemies. He showed that the difficulties
to detect the density-dependent effect of natural enemies was due to
time delays and stochasticity. This paper is still considered a classic
in fields of insect and host-parasitoid population dynamics. It
introduced an important argument on the role of natural enemies on
insect populations, a controversial topic that aroused ecologists to
debate for almost a decade (\textcite{Turchin90}).

We used information from \textcite{Hassell85} to replicate the model. We
were able to replicate the results central to the article. In addition,
we made new analyses for the stochastic models that bring additional
support to \textcite{Hassell85} arguments. To our best knowledge, the
original implementation was not available. The code for the simulations
and the figures were written in \emph{Julia} v0.6.2.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The mathematical formulation used in this paper to show the difficulties
of detecting natural enemies as regulating factors are the same that
were used in the original paper by \textcite{Hassell85} . First of all,
the host population dynamics are given as

\begin{equation}N_{t+1} = F \times N_t \times f(N_t,P_t) \times D\label{eq:1}\end{equation}

where \(N_t\) and \(N_{t+1}\) represent the host population at
generation \(t\) and at the next generation, \(F\) is the rate of
increase of the population and \(D\) is the density independent
probability of survival of the hosts (mortality). The specialist
parasitoids population dynamics are represented by

\begin{equation}P_{t+1} = c \times N_t \times [1-f(N_t, P_t)]\label{eq:2}\end{equation}

where \(P_t\) and \(P_{t+1}\) are the number of parasitoids at
generation \(t\) and at the next one, while \(c\) is the number of
female parasitoids emerging from each host parasitized. In both
eq.~\ref{eq:1} and eq.~\ref{eq:2}, \(f(N_t,P_t)\) represents the
probability of escaping mortality from natural enemies (parasitoids) and
is given by eq.~\ref{eq:3}.

\begin{equation}f(N_t,P_t) = [1 + (a \times P_t) / (m \times (1 + a \times T_h \times N_t))]^{-m} \label{eq:3}\end{equation}

where \(a\) is the per capita searching efficiency of the parasitoids,
\(m\) is the extent of clumping of the parasitoids attacks and \(T_h\)
is the handling time as a proportion of the total time. This paper also
explores the relationship between the hosts and a generalist parasitoid
population. Generalist parasite dynamics follows the equation

\begin{equation}P_t = h \times \left(1 - \text{exp}\left(-\frac{N_t}{b}\right)\right)\label{eq:4}\end{equation}

where \(h\) is the saturation number of parasitoids and \(b\) is the
rate of approaching this saturation number.

To determine if the natural enemies can be declared as density-dependent
factors, the host population mortality due to parasitism (\(k-value\))
is plotted against population density for each simulated generation. The
correlation coefficient \(r\) of the resulting scatter plot indicates
the strength of the density-dependence of natural enemies. The higher
\(r\) is, the strongest the relation between hosts and parasites is. The
host mortality is given by

\begin{equation}k_\text{value} = \text{log}_{10}\frac{N_t}{S}\label{eq:5}\end{equation}

where \(S\) is the number of hosts that survived parasitism. This number
is given by the host population density multiplied by the probability of
escaping mortality from natural enemies (eq.~\ref{eq:6}).

\begin{equation}S = N_t \times f(N_t,P_t)\label{eq:6}\end{equation}

The objective in this paper is to reproduce the main results of the
original publication. Therefore, we did not reproduce figures 1, 2 and
7. Figure 1 was a schematic representation of an insect life cycle.
Figure 2 and Figure 7 represent relationship between parameters and
population size/proportion of parasited host. Therefore, they are not
necessary in order to show how difficult it is to detect the regulating
effect of natural enemies on a host population. For every figure
reproduced, we used the exact same values that were used in the original
paper for the different parameters.

The software used to code and run the models and to generate the figures
is \emph{Julia} version 0.6.2 (\textcite{Bezanson17}) All the code used
to replicate the original paper is available alongside the article.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{deterministic-model}{%
\subsubsection{Deterministic model}\label{deterministic-model}}

The reproduced population dynamics for the host population and for the
specialist parasitoids (fig.~\ref{fig:figure3} (a) and (b)) are very
similar to the ones in Hassell's paper. When the level of clumping of
parasitoid attacks is high, both populations decrease during the first
10 generations before they stabilize, with the host population twice as
big as the specialist parasitoid population. When the extent of clumping
is lower, both populations show decreasing oscillations during the 50
simulated generations. The difference between this article and the
original publication resides in the fig.~\ref{fig:figure3} (c) and (d),
where we standardized the axes of the graphs. At first sight, in
Hassell's paper, the \(k-values\) in fig.~\ref{fig:figure3} (c) seem
almost as high as in fig.~\ref{fig:figure3} (d). However, when using the
same scale for the two graphs, we can see that the \(k-values\) in the
case of high level of clumping show a lot less variation than with a
lower level of clumping.

The simulations performed to study the relationship between the
population dynamics of the hosts and the generalist parasitoids are also
very similar to the ones done by Hassell (fig.~\ref{fig:figure4}). This
shows a density dependant relationship between the two populations,
where the natural enemy regulates the host population until they both
reach a stable equilibrium. This is also the conclusion when we observe
the relation between the two populations in fig.~\ref{fig:figure3} (a)
and (c).

\hypertarget{stochastic-model}{%
\subsubsection{Stochastic model}\label{stochastic-model}}

When the models include stochasticity, the resulting population dynamics
between hosts and parasites are very different compared to the
deterministic models. Overall, the results obtained in the replications
are very similar to the original paper. Whether we look at the model for
the specialist parasitoids or the one for the generalist parasitoids,
both show irregular oscillations in the host-parasitoid population
dynamics (fig.~\ref{fig:figure5} and fig.~\ref{fig:figure6}, (a), (b)
and (c)). The addition of stochastic parameters prevents the
stabilization of host populations and makes it more difficult to
identify parasitoids as a density-dependent control factor, except in
fig.~\ref{fig:figure6} (a). The relationships between the \(k-values\)
and the host density are similar in Hassell's publication and in ours.
However, the regression for these relationships in our replication tend
to have a determination coefficient (\(R^2\)) higher than the one found
by Hassell, but they are generally really close. Discrepancies in the
coefficient of determination can be explained by different routines for
pseudo random number generation.

Moreover, the oscillations we obtained in fig.~\ref{fig:figure6} (b) are
not the same range compared to the original paper. More precisely, the
oscillations we obtained with \(h\) stochastic are smaller compared to
the original paper (fig.~\ref{fig:figure6} (b)).

Because the inclusion of stochastic parameters in the population
dynamics causes variability in the outputs, the results from two
successive simulations can be very different. In order to account for
this variability and to show how it can affect the population dynamics
of the hosts and parasites, we added fig.~\ref{fig:figure7} and
fig.~\ref{fig:figure8}. These figures show the extent of the variation
of the correlation coefficient (\(r\)) obtained in 5000 different
simulations (as opposed to a single simulation in the original article).
A dotted line was added to represent the value of the correlation
coefficient (\(r\)) that came out of the deterministic models. The
values of \(r\) vary greatly for every stochastic parameter. In every
case, the mean value of \(r\) is lower in the stochastic models than in
the deterministic models. This is in agreement with Hassell's results,
and shows that stochasticity makes it harder to detect the density
dependence effect of the parasites, whether they are generalist or
specialist.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Overall, we were able to replicate most results from
\textcite{Hassell85}. We found the exact same results for the
deterministic model. We standardized the limits for the axes, which was
not the case in the original paper. This allows a more convenient
comparison of the different results.

As expected, we did not find the exact same dynamic for the stochastic
model. The figures we added (fig.~\ref{fig:figure7} and
fig.~\ref{fig:figure8}) showed how adding stochasticity into the model
can cause great variability in the output. For example, in the
specialist parasitoid model with a stochastic density-dependent host
survival (\(D\)), the correlation we found (fig.~\ref{fig:figure7} (a))
was sometimes very weak (\(r \approx 0.2\)) and some other times almost
as strong as the deterministic model (\(r \approx 0.7\)). Also, the
correlation between the mortality from parasitism (\(k\)-value) and host
density (\(N\)) found in the stochastic model was almost always weaker
than in the deterministic model (fig.~\ref{fig:figure7} and
fig.~\ref{fig:figure8}). Therefore, the results we added strongly
support the main argument from the original paper: adding stochasticity
almost always obscures the density-dependent effect of natural enemies.

The discrepancies we noted in the dynamics of the stochastic model with
the generalist parasite and a stochastic saturation number of parasitoid
(\(h\); in fig.~\ref{fig:figure6} (b)) are difficult to explain. It
seems unlikely that it is caused by either our implementation of the
model, or by errors during runtime, since we successfully replicated
results from all the other numerical experiments. Without the original
implementation, we can only speculate on the difference between the
original implementation and ours. This combination of parameters is the
only one with a stochastic parameter normally distributed with a mean
not equal to its associated standard deviation. We tried with an \(h\)
normally distributed with a mean and a standard deviation of 5
(fig.~\ref{fig:figure9}), and the results were a lot more similar to the
one from \textcite{Hassell85} than what we originally had
(fig.~\ref{fig:figure6} (b)). Again, this is only an hypothesis on the
kind of error that could explain the differences between the original
paper and ours. Other possible sources of error include the
pseudo-random numbers generator used, errors in the original
implementation, or errors in the parameter values reported in the
original publication; sadly, these are virtually impossible to rule out.

The mathematical model from the original paper was well detailed, which
allowed us to create our own implementation. The equation for the number
of survivors from parasitism (\(S\)) was the only one we needed to
deduce from our own interpretation. This variable is used in the
computation of the mortality (\(k_\text{value}\)) which is a well
documented index. Therefore, this has not limited us in the replication
of the article, and the fact that the deterministic simulations match
these of the original paper suggests that we used the same formulation
for \(S\).

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We were able to replicate the original results. Even if we did not find
the exact same dynamics for the stochastic models, we draw the same
conclusions : the density-dependent effect from natural enemies is
obscured by time delays and/or stochasticity. This makes it very
difficult to detect natural enemies as regulating factors from life
table data. In addition, we added density plots for the correlation
coefficient from 5000 iterations of each stochastic model. This allowed
us to determine that the differences between our results and Hassell's
were explained by the stochasticity of the models. Also, these new
results add a strong support to the arguments of the original paper. To
conclude, the reproduction of the reference article \textcite{Hassell85}
was successful and we hope it adds to the legacy left by this
significant paper in the history of population dynamics.

\begin{figure}
\hypertarget{fig:figure3}{%
\centering
\includegraphics{figures/figure3.pdf}
\caption{(a) and (b) Deterministic simulations of the host and
specialist parasite population dynamics (Eq.1, Eq.2, Eq.3) using two
different level of clumping in the parasitoid attacks: (a) m = 0.2; (b)
m = 0.8. The other parameters used are the same in both (a) and (b): F =
4, D = 0.5, c = 1, a = 0.5 and T\emph{h} = 0. (c) and (d) The
relationship between the mortality caused each generation by parasitism
(k-values) and the log10 host density for the fifty first generations,
linked to (a) and (b) respectively.}\label{fig:figure3}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure4}{%
\centering
\includegraphics{figures/figure4.pdf}
\caption{(a) Deterministic simulations of the host and generalist
parasite population dynamics (Eq.1, Eq.3, Eq.4) with the following
parameters: F = 4, D = 0.5, h = 10, b = 25, a = 0.5 and T\emph{h} = 0
and m = 0.5. (b) The relationship between the mortality caused each
generation by parasitism (k-values) and the log10 host density for the
fifty first generations of the population dynamics in
(a).}\label{fig:figure4}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure5}{%
\centering
\includegraphics{figures/figure5.pdf}
\caption{(a) - (c) Deterministic simulations of the host and specialist
parasite population dynamics (same as in Figure 3a) except for one
parameter that is treated as a normally distributed stochastic variable:
(a) D = 0.5 ± 0.5, (b) c = 0.5 ± 0.5 and (c) a = 0.5 ± 0.5. The other
parameters are the same as in Figure 3a. (d) - (f) The relationship
between the mortality caused each generation by parasitism (k-values)
and the log10 host density for the fifty first generations, linked to
(a), (b) and (c) respectively. The regression statistics for each
relationship go as follows: (d) \emph{y} = 0.049 + 0.148\emph{x};
\emph{r\^{}2} = 0.564. (e) \emph{y} = 0.082 + 0.113\emph{x};
\emph{r\^{}2} = 0.034. (f) \emph{y} = 0.083 + 0.158\emph{x};
\emph{r\^{}2} = 0.082.}\label{fig:figure5}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure6}{%
\centering
\includegraphics{figures/figure6.pdf}
\caption{(a) - (c) Deterministic simulations of the host and generalist
parasite population dynamics (same as in Figure 4a) except for one
parameter that is treated as a normally distributed stochastic variable:
(a) D = 0.5 ± 0.5, (b) h = 10 ± 5 and (c) a = 0.5 ± 0.5. The other
parameters are the same as in Figure 3a. (d) - (f) The relationship
between the mortality caused each generation by parasitism (k-values)
and the log10 host density for the fifty first generations, linked to
(a), (b) and (c) respectively. The regression statistics for each
relationship go as follows: (d) \emph{y} = 0.111 + 0.174\emph{x};
\emph{r\^{}2} = 0.843. (e) \emph{y} = -0.010 + 0.307\emph{x};
\emph{r\^{}2} = 0.187. (f) \emph{y} = -0.106 + 0.444\emph{x};
\emph{r\^{}2} = 0.262.}\label{fig:figure6}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure7}{%
\centering
\includegraphics{figures/figure7.pdf}
\caption{(a) - (c) Distribution of the values of correlation coefficient
(\(r\)) obtained for the 5000 simulations done in Figure 5(d) to Figure
5(f) respectively. The dotted line represents the value of \(r\)
obtained in the deterministic model.}\label{fig:figure7}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure8}{%
\centering
\includegraphics{figures/figure8.pdf}
\caption{(a) - (c) Distribution of the values of correlation coefficient
(\(r\)) obtained for the 5000 simulations done in Figure 6(d) to Figure
6(f), respectively. The dotted line represents the value of \(r\)
obtained in the deterministic model.}\label{fig:figure8}
}
\end{figure}

\begin{figure}
\hypertarget{fig:figure9}{%
\centering
\includegraphics{figures/figure9.pdf}
\caption{(a) Deterministic simulations of the host and generalist
parasite population dynamics (same as in Figure 6b) except that we used
\emph{h} = 5 ± 5. (b) The relationship between the mortality caused each
generation by parasitism (k-values) and the log10 host density for the
fifty first generations, linked to (a). The regression statistics go as
follows: \emph{y} = 0.030 + 0.192\emph{x}; \emph{r\^{}2} =
0.095.}\label{fig:figure9}
}
\end{figure}

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}

### 2019-05-24 article-latex/content.tex
Created new latex based article

\section{Introduction}

Ecological resources are often managed on the basis of outputs derived
from models of species-environment relationships, variously referred to
as: habitat suitability models, species distribution models, resource
selection functions, or ecological niche models. Therefore, evaluating
the predictive ability of such models is a necessary prerequisite for
robust decision-making. While such evaluations are ideally based on
independent data collected purely for the purpose of model testing, in
ecological studies fully independent data is often unavailable due to
logistical or financial constraints. Therefore, statistical resampling
methods are the most important tool we have for evaluating predictive
ability.

\bigskip

To briefly summarise, resampling methods repeatedly resample the full
dataset to create training and testing data subsets that are independent
of one another. The model is then iteratively fit using each training
data subset and prediction error estimated using the independent testing
data subset. An overall estimate of prediction error is then calculated
as the average prediction error across all resampled data subsets. While
resampling methods are not a replacement for independent data, they can
be used to conduct an internal evaluation that penalises for optimism
from overfitting\supercite{verbyla-1989}.

\bigskip

In their seminal paper on model evaluation Fielding and Bell\supercite{fielding-1997}
state with reference to resampling methods ``The
ecological literature seems to have paid little attention to how the
partitioning method can influence the error rates. Verbyla and Litvaitis
briefly reviewed a range of partitioning methods in their
assessment of resampling methods for evaluating classification
accuracy.'' The work of Verbyla and Litvaitis\supercite{verbyla-1989} remains the
only comparison of resampling methods for evaluating species-environment
relationship models. Therefore, given the importance of this work, we
endeavoured to replicate the study using an open computational approach.

\section{Prediction error}

When developing a species-environment relationship model, we will
usually have a dataset $d$ that contains observations or measurements
that form a species response variable $y$ and a set of one or more
environmental explanatory variables $x$. Using a species-environment
relationship model function $f^{}$ trained on the dataset $f^{(d)}$ a
prediction of the response variable $\hat{y}$ can be created from the
environmental explanatory variables for each sample $i$:

\begin{equation}
\hat{y_i} = f^{(d)}(x_i)
\end{equation}

We can then define the prediction error $Err$ for each sample $i$ as the
absolute difference between the observed response variable $y_i$ and
predicted response variable $\hat{y_i}$:

\begin{equation}
Err_i = | y_i - \hat{y_i} |
\end{equation}

This definition of $Err$ is equivalent to the binary misclassification
error rate used by Verbyla and Litvaitis\supercite{verbyla-1989}, as when
$y_i = \hat{y}$ then $Err_i = 0$, and when $y_i \neq \hat{y}$ then
$Err_i = 1$. But by expressing $Err$ in these terms generalises the
approach to situations in which the species-environment relationships
are measured or modelled on a continuous rather than binary scale, which
has become more common practice since the original computational
experiment was conducted.

\section{Resampling methods}

During the replication process it became apparent that the terminology
for resampling methods has developed over time, and has been used
somewhat inconsistently. Therefore, we begin by naming and formally
defining each of the resampling methods we have used based on
descriptions within the references.

\subsection{Resubstitution}

Given a dataset $d$ of size $n$ the resubstitution method calculates the
mean prediction error across all samples $i$ from a modelling function
trained on the entire dataset $f^{(d)}$.

\begin{equation}
Err^{R} = \frac{1}{n} \sum_{i=1}^{n} | y_i - f^{(d)}(x_i) |
\label{eq:resubstitution}
\end{equation}

The value $Err^{R}$ is called the resubstitution (or apparent) error
rate and is likely to provide an optimistic estimate of prediction
error, as the same data is used to train and test the model.

\subsection{Hold-out cross-validation}

Hold-out (or: split-sample, randomised, Monte Carlo) cross-validation
randomly partitions the dataset into training and testing subsets. 
Verbyla and Litvaitis\supercite{verbyla-1989} referred to this approach as 
simply ``cross-validation'' but we have chosen to use the more specific 
term of hold-out cross-validation to clarify which of the many types of 
cross-validation we are referring to. The proportion $p$ of data `held-out' 
from the dataset $d$ forms a testing
dataset $t$ of sample size $\sharp t$, with the remaining data forming
training dataset $\{d-t\}$. The model is fitted using the training
dataset $f^{(\{d-t\})}$, and the prediction error is estimated as the
mean prediction error for all $i$ in $t$ across a number of $H$
repetitions.

\begin{equation}
Err^{H}_{p} = \frac{1}{H } \sum_{t=1}^{H} \frac{1}{\sharp t} \sum_{i \in t} | y_i - f^{(\{d-t\})}(x_i) |
\end{equation}

In general this method can be considered an improvement over
resubstitution as the data used to train the model is separated from the
data used to test the model.

\subsection{$K$-fold cross-validation}

Verbyla and Litvaitis\supercite{verbyla-1989} describe a resampling method called ten-fold
cross-validation and another method called $n$-fold cross-validation or the jackknife.
Both these methods are variations of $K$-fold cross-validation. The $K$-fold
cross-validation method begins by randomly partitioning the
dataset into $k$ equally sized sets. Then the prediction for each sample
$i$ is calculated from a model fitted to the set $\{d-k : i \in k\}$,
which is the dataset $d$ excluding the set $k$ where $k$ includes $i$.

\begin{equation}
Err^{K}  = \frac{1}{n} \sum_{i=1}^{n} | y_i - f^{(\{d-k : i \in k\})}(x_i) |
\end{equation}

When $k = n$ then we produce a special form of $K$-fold cross-validation
called the jackknife (or: leave-one-out, $n$-fold) cross-validation.

\subsection{Bootstrap cross-validation}

Bootstrapping\supercite{diaconis-1983} is based upon a set $B$ of bootstrap sample datasets $b$,
for which $b$ is of size $n$ and is generated by randomly sampling with
replacement from the full dataset $d$. The model is
iteratively fitted to each $b$ and prediction errors calculated for all
samples $i$ in the test set $\{d - b\}$ that consists of the dataset $d$ with
all the $i$ in the bootstrap sample $b$ removed. This results in around
0.632 of the dataset $d$ occurring at least once in the training set $b$,
and the remaining 0.368 of $d$ occurring in testing set\supercite{hastie-2009}. The estimated prediction error rate is then the mean
prediction error across all bootstrap samples\supercite{efron-1983, jain-1987}.

\begin{equation}
Err^{B} = \dfrac{\sum\limits_{b=1}^{B}\sum\limits_{{i \in \{ d-b \} }}| y_i - f^{(b)}(x_i) |}{\sum\limits_{b=1}^{B} \sharp \{ d-b \}}
\label{eq:original-bootstrap}
\end{equation}

It is worth noting in the context of a replication study that the
equation used to calculate $Err^{B}$ was later changed, as this caused
some confusion during our replication. The second equation\supercite{efron-1993} works through
each sample $i$, and calculates the mean prediction error for a set
$C_i$ of size $\sharp C_i$ that is equal to the all bootstrap samples
that do not contain $i$, $C_i = \{ b \in B : i \notin b \}$.

\begin{equation}
Err^{B} = \frac{1}{n} \sum_{i=1}^{n} \frac{1}{\sharp C_i} \sum_{b \in C_i} | y_i - f^{(b)}(x_i) |
\end{equation}

This second definition was later termed the `leave-one-out bootstrap'
where it was also noted by Efron\supercite{efron-1997} that these ``two definitions agree as
$B \rightarrow \infty$ and produced nearly the same results in our
simulations'' which were based on $B=50$. Although the later
defintion of $Err^{B}$ has become more commonly used\supercite{hastie-2009}, 
as the two methods produce nearly identical results, we
have used the original definition (Equation \ref{eq:original-bootstrap})
in our replication as it is simpler to compute, and was the version used
by Verbyla and Litvaitis\supercite{verbyla-1989} that we are trying to replicate.

\section{Computational experiment replication}

Verbyla and Litvaitis\supercite{verbyla-1989} based their computational experiment
on applying linear discriminant analysis models to a random dataset.
Their premise was that by creating a random dataset the predictive
ability of a model should be no better than chance, and hence as the
true prediction error was known exactly each resampling method could be
assessed for bias and precision.

\subsection{Random datasets}

Each of 1000 computational experiments began by creating a random
dataset. The response variable consisted of 30 observations that were
randomly assigned a presence $=1$ or absence $=0$ value. These 30
observations were then matched with 10 explanatory variables. Verbyla and Litvaitis\supercite{verbyla-1989} 
state that the ``ten predictor variables were
generated with univariate normal distributions and equal variances'' but
neither the mean nor variance used was reported. Therefore, we assumed a
standard normal distribution of $\mu$ = 0 and $\sigma$ = 1. For each random
dataset a linear discriminant analysis model was fitted and prediction error
calculated using each resampling method.

\subsection{Resubstitution}

The resubstitution method is the simplest approach and the most
consistently reported in the literature, therefore we applied the
methodology exactly as described.

\subsection{$K$-fold cross-validation)}

We conducted $K$-fold cross-validation with $K=10$ and $K=n$ as in
Verbyla and Litvaitis\supercite{verbyla-1989} -- remembering that $K=n$ is
equivalent to the jackknife method. We also included $K=3$ as this
represents a situation with a similar proportion of the dataset forming
training and testing sets as with the bootstrap. This was done to assess
if the proportion of data within training and testing sets was affecting
comparisons of the resampling methods.

\subsection{Bootstrap cross-validation}

The bootstrap method was applied with bootstrap samples $B=200$ as
specified in the example code provided by Verbyla and Litvaitis\supercite{verbyla-1989}.

\subsection{Hold-out cross-validation}

We included three cases of hold-out cross-validation as while Verbyla and Litvaitis\supercite{verbyla-1989} 
stated that for this method the ``the estimate
of model classification accuracy will not be very precise'' no
experiments or citations were provided to support this claim. In
addition we felt there was some inconsistency in the two citations given with
reference to this method.  While Verbyla and Litvaitis\supercite{verbyla-1989} state 
``only one estimate of accuracy 
is made'' which matches the citation\supercite{lachenbruch-1968} using hold-out cross-validation
with $H=1$, a second citation\supercite{capen-1986} referred to hold-out
cross-validation with $H=10$.  Given this uncertainty we
explored three different versions of hold-out cross-validation. We used
$p=0.368$ twice to match the proportion of test data in the bootstrap
approach, and with $H=1$ for one method, and $H=200$ in the second
method for consistency with the bootstrap. We also used $H=200$ with
$p=0.200$ to examine sensitivity to the test data proportion.

\section{Results}

The results for each resampling method were presented by Verbyla and Litvaitis\supercite{verbyla-1989} 
as a ``smoothed frequency distribution'' but the
smoothing method was not reported. As their results appeared to be
normally distributed, to mimic the original results to aid comparisons
we produced our smoothed distributions using a one-dimensional Gaussian
kernel density estimator with a bandwidth of one (Figure \ref{fig:smoothed-distributions}).

\begin{figure}[tbh]
\centering
\includegraphics[width=\textwidth]{../code/resampling-results.png}
\caption{Smoothed distributions of estimates of prediction error for various resampling methods from 1000 computational experiments with the known prediction error of 0.5 also marked.}
\label{fig:smoothed-distributions}
\end{figure}

\bigskip

Looking at the distribution of $Err$ across the 1000 computational
experiments, the resubstitution method produced clearly biased estimates
of prediction error. All the other methods produced unbiased estimates,
but there was notable variation in the precision of those estimates,
with $Err^{B=200}$ ($\mu=0.497$, $\sigma=0.069$) producing the most
precise estimates.

\section{Conclusion}

While a lack of method description means our implementation will be
different to that of Verbyla and Litvaitis\supercite{verbyla-1989}, we would
conclude that our results are sufficiently similar to have replicated
their computational experiments.

\bigskip

Our findings confirm that resubstitution is a biased estimate of
prediction error, and that bootstrap cross-validation produces the most
precise unbiased estimate. We also found that hold-out cross-validation
produced unbiased but highly variable estimates of $Err$ as the method
is clearly sensitive to the choice of parameters. We found little
difference between any of the $K$-fold cross-validation methods.

\bigskip

Given the findings from our replication, we would support Verbyla and Litvaitis\supercite{verbyla-1989} 
in advocating the use of the bootstrap, as it
produced the most precise estimate, and unlike other resampling methods
it does not require a arbitrary choice of dataset partitions or splits
that could confound inter-study comparisons of model evaluations.

\bigskip

We conclude that while not a substitute for truly independent data,
resampling methods should be considered an important part of
species-environment relationship model evaluation, and would encourage
the use of the bootstrap cross-validation method in particular.

\section{Acknowledgements}

This research was funded by internal investment by Manaaki Whenua –
Landcare Research.

### 2019-05-24 article-latex/metadata.tex
Created new latex based article

% DO NOT EDIT - automatically generated from metadata.yaml

\def \codeURL{https://github.com/rescience-c/template}
\def \codeDOI{}
\def \dataURL{}
\def \dataDOI{}
\def \editorNAME{}
\def \editorORCID{}
\def \reviewerINAME{}
\def \reviewerIORCID{}
\def \reviewerIINAME{}
\def \reviewerIIORCID{}
\def \dateRECEIVED{}
\def \dateACCEPTED{}
\def \datePUBLISHED{}
\def \articleTITLE{[Re] Resampling methods for evaluating classification accuracy of wildlife habitat models}
\def \articleTYPE{Replication}
\def \articleDOMAIN{Ecology}
\def \articleBIBLIOGRAPHY{bibliography.bib}
\def \articleYEAR{2019}
\def \reviewURL{}
\def \articleABSTRACT{Three years have passed since ReScience published its first article and since September 2015, things have been going steadily. We're still alive, independent and without a budget. In the meantime, we have published around 20 articles (mostly in computational neuroscience \& computational ecology) and the initial has grown from around 10 to roughly 100 members (editors and reviewers), we have advertised ReScience at several conferences worldwide, gave some interviews and we published an article introducing ReScience in PeerJ. Based on our experience at managing the journal during these three years, we think the time is ripe for proposing some changes.}
\def \replicationCITE{}
\def \replicationBIB{}
\def \replicationURL{}
\def \replicationDOI{}
\def \contactNAME{Thomas R. Etherington}
\def \contactEMAIL{etheringtont@landcareresearch.co.nz}
\def \articleKEYWORDS{ecology, wildlife, habitat, model}
\def \journalNAME{ReScience C}
\def \journalVOLUME{4}
\def \journalISSUE{1}
\def \articleNUMBER{}
\def \articleDOI{}
\def \authorsFULL{Thomas R. Etherington and David J. Lieske}
\def \authorsABBRV{T.R. Etherington and D.J. Lieske}
\def \authorsSHORT{Etherington and Lieske}
\title{\articleTITLE}
\date{}
\author[1,\orcid{0000-0002-3187-075X}]{Thomas R. Etherington}
\author[2]{David J. Lieske}
\affil[1]{Manaaki Whenua -- Landcare Research, PO Box 69040, Lincoln 7608, New Zealand}
\affil[2]{Department of Geography and Environment, Mount Allison University, 144 Main Street, Sackville, New Brunswick, E4L 1A7, Canada}

### 2019-05-24 article-latex/article.py

@@ -1,0 +1,296 @@
+# ReScience yaml parser
+# Released under the BSD two-clauses licence
+
+import yaml
+
+class Contributor:
+    def __init__(self, role, name, orcid="", email="", affiliations=[]):
+        self.role = role
+        self.name = name
+        self.fullname = name
+        self.lastname = self.get_lastname(name)
+        self.abbrvname = self.get_abbrvname(name)
+        self.orcid = orcid
+        self.email = email
+        self.affiliations = affiliations
+
+    def get_abbrvname(self, name):
+        if not name: return ""
+        
+        if ',' in name:
+            lastname = name.split(",")[0]
+            firstnames = name.split(",")[1].strip().split(" ")
+        else:
+            lastname = name.split(" ")[-1]
+            firstnames = name.split(" ")[:-1]
+        abbrvname = ""
+        for firstname in firstnames:
+            if "-" in firstname:
+                for name in firstname.split("-"):
+                    abbrvname += name[0].strip().upper() + '.-'
+                abbrvname = abbrvname[:-1]
+            else:
+                abbrvname += firstname[0].strip().upper() + '.'
+        return abbrvname + " " + lastname
+
+
+    def get_lastname(self, name):
+        if not name: return ""
+        # Rougier, Nicolas P.
+        if ',' in name:
+            lastname = name.split(",")[0].strip()
+            firstname = name.split(",")[1].strip()
+        # Nicolas P. Rougier
+        else:
+            lastname = name.split(" ")[-1]
+            firstname = name.split(" ")[:-1]
+        return lastname
+    
+
+class Affiliation:
+    def __init__(self, code, name, address=""):
+        self.code = code
+        self.name = name
+        self.address = address
+
+class Repository:
+    def __init__(self, name, url, doi):
+        self.name = name
+        self.url = url
+        self.doi = doi
+
+class Replication:
+    def __init__(self, cite, bib, url, doi):
+        self.cite = cite
+        self.bib = bib
+        self.url = url
+        self.doi = doi
+
+class Review:
+    def __init__(self, url, doi):
+        self.url = url
+        self.doi = doi
+
+class Date:
+    def __init__(self, date):
+        try:
+            import dateutil.parser
+
+            date = dateutil.parser.parse(date)
+            self.date = date
+            self.year = date.year
+            self.month = date.month
+            self.day = date.day
+            self.textual = self.date.strftime("%d %B %Y")
+        except:
+            import datetime
+            now = datetime.datetime.now()
[…]

### 2019-05-24 article-latex/yaml-to-latex.py

@@ -1,0 +1,95 @@
+#! /usr/bin/env python3
+# ReScience yaml to latex converter
+# Released under the BSD two-clauses licence
+
+def generate_latex_metadata(filename, article):
+
+    abstract = article.abstract.replace("&", "\&")
+    
+    content = (
+        "% DO NOT EDIT - automatically generated from {filename}\n\n"
+        "\\def \\codeURL{{{_.code.url}}}\n"
+        "\\def \\codeDOI{{{_.code.doi}}}\n"
+        "\\def \\dataURL{{{_.data.url}}}\n"
+        "\\def \\dataDOI{{{_.data.doi}}}\n"
+        "\\def \\editorNAME{{{_.editors[0].name}}}\n"
+        "\\def \\editorORCID{{{_.editors[0].orcid}}}\n"
+        "\\def \\reviewerINAME{{{_.reviewers[0].name}}}\n"
+        "\\def \\reviewerIORCID{{{_.reviewers[0].orcid}}}\n"
+        "\\def \\reviewerIINAME{{{_.reviewers[1].name}}}\n"
+        "\\def \\reviewerIIORCID{{{_.reviewers[1].orcid}}}\n"
+        "\\def \\dateRECEIVED{{{_.date_received}}}\n"
+        "\\def \\dateACCEPTED{{{_.date_accepted}}}\n"
+        "\\def \\datePUBLISHED{{{_.date_published}}}\n"
+        "\\def \\articleTITLE{{{_.title}}}\n"
+        "\\def \\articleTYPE{{{_.type}}}\n"
+        "\\def \\articleDOMAIN{{{_.domain}}}\n"
+        "\\def \\articleBIBLIOGRAPHY{{{_.bibliography}}}\n"
+        "\\def \\articleYEAR{{{_.date_published.year}}}\n"
+        "\\def \\reviewURL{{{_.review.url}}}\n"
+        # "\\def \\articleABSTRACT{{{_.abstract}}}\n"
+        "\\def \\articleABSTRACT{{{abstract}}}\n"
+        "\\def \\replicationCITE{{{_.replication.cite}}}\n"
+        "\\def \\replicationBIB{{{_.replication.bib}}}\n"
+        "\\def \\replicationURL{{{_.replication.url}}}\n"
+        "\\def \\replicationDOI{{{_.replication.doi}}}\n"
+        "\\def \\contactNAME{{{_.contact.name}}}\n"
[…]
